### 2012-10-12 paper.tex
Added data and code

%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
\documentclass[journal]{IEEEtran}

\usepackage{mathbf-abbrevs}

%\newcommand{\dealias}{\operatorname{dealias}}

\input{defs}

%\newcommand{\term}{\emph}
%\renewcommand{\index}[1]{}

%\title{Polynomial phase estimation by phase unwrapping 1: Identifiability and strong consistency}
\title{Polynomial phase estimation by phase unwrapping}

\author{Robby~G.~McKilliam, Barry~G.~Quinn, I.~Vaughan~L.~Clarkson, Bill~Moran and Badri~N.~Vellambi%
    \thanks{%A preliminary version of some of this material is contained in Part 3 of Robby McKilliam's PhD thesis \cite{McKilliam2010thesis}. 
Robby~McKilliam and Badri Vellambi are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.  Barry~Quinn is with the Department of Statistics, Macquarie University, Sydney, NSW, 2109, Australia.   Vaughan~Clarkson is with the School of Information Technology \& Electrical Engineering, The University of Queensland, QLD., 4072, Australia.  B. Moran is with the Department of Electrical Engineering and Computer
Science, Melbourne Systems Lab, Dept of Elec \& Electronic Eng, Uni of Melbourne, Vic. 3010, Australia.}}
% The paper headers 
\markboth{Polynomial phase estimation by phase unwrapping}{DRAFT \today}

% make the title area

\begin{document} 
\maketitle

\begin{abstract}
Estimating the coefficients of a noisy polynomial phase signal is important in many fields including radar, biology and radio communications. One approach attempts to perform polynomial regression on the phase of the signal.  This is complicated by the fact that the phase is \emph{wrapped} modulo $2\pi$ and must be \emph{unwrapped} before regression can be performed. %A recent approach suggested by the authors is to perform the unwrapping in a least squares manner.  %It was shown how the resulting estimator could be posed as a \emph{nearest lattice point problem} in a specific lattice and could be solved using existing algorithms. 
%It was shown by Monte Carlo simulation that this produces a remarkably accurate estimator.  
In this paper we consider an estimator that performs phase unwrapping in a least squares manner.  We describe the asymptotic properties of this estimator, showing that it is strongly consistent and asymptotically normally distributed. %We hypothesise that the estimator produces very near maximum likelihood performance.
\end{abstract}

\begin{keywords}
Polynomial phase signals, phase unwrapping, asymptotic properties, nearest lattice point problem
\end{keywords}

%\bibliographystyle{unsrt}

\section{Introduction} \label{intro}

Polynomial phase signals arise in numerous fields including radar, sonar, geophysics, biology, and radio communication \cite{Angeby_estimating_2000}. In radar and sonar applications polynomial phase signals arise when acquiring radial velocity and acceleration (and higher order motion descriptors) of a target from a reflected signal, and also in continuous wave radar and low probability of intercept radar~\cite{Levanon_Radar_signals_2004}.  In biology, polynomial phase signals are used to describe the sounds emitted by bats and dolphins for echo location \citep{Suga_1975_bats_echolocation, Moss_2005echolocation}.

A polynomial phase signal of order $m$ is a function of the form
\[
s(t) = e^{2\pi j y(t)},
\]
where $j = \sqrt{-1}$, and $t$ is a real number, often representing time, and 
\[
y(t) = \tilde{\mu}_0 +\tilde{\mu}_1 t + \tilde{\mu}_2 t^2 + \dots \tilde{\mu}_m t^m
\]
is a polynomial of order $m$.  In practice the signal is typically sampled at discrete points in `time', $t$. In this paper we only consider uniform sampling, where the gap between consecutive samples is constant. In this case we can always consider the samples to be taken at some set of consecutive integers and our sampled polynomial phase signal looks like
\[
s_n = s(n) = e^{2\pi j y(n)},
\] 
where $n$ is an integer.  Of practical importance is the estimation of the coefficients $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ from a number, say $N$, of observations of the noisy sampled signal
\begin{equation}\label{eq:Y_nsamplednoisey}
Y_n = \rho s_n + X_n,
\end{equation}
where $\rho$ is a real number greater than zero representing the (usually unknown) signal amplitude and $\{X_n, n \in \ints\}$ is a sequence complex noise variables. In order to ensure identifiability it is necessary to restrict the $m+1$ coefficients to a region of $m+1$ dimensional Euclidean space $\reals^{m+1}$ called an \emph{identifiable region}.  It was shown in \cite{McKilliam2009IndentifiabliltyAliasingPolyphase} that an identifiable region tessellates a particular $m+1$ dimensional lattice.  We discuss this in Section \ref{sec:ident_aliasing}.

%A popular estimator is based on the \emph{discrete polynomial phase transform} (DPT) first introduced by Peleg and Porat \cite{Peleg_DPT_1995}.  The transform enables each parameter to be be estimated iteratively using the Fast Fourier transform in a way analogous to the frequency estimator of Rife and Boorstyn \cite{Rife1974, Quinn2001}.  The DPT estimator is characterized by good statistical performance and computational efficiency.  Variants of the DPT have been suggested by numerous authors including Golden and Friedlander \cite{Golden_mod_dpt_1998} and O'Shea \cite{Oshea_iterative_1996}. The recent \emph{high-order phase function} estimators are based on similar ideas \cite{Farquharson_another_poly_est_2005}.  A substantial drawback of all of these estimators is that they only work correctly over a small set of parameters, generally much smaller than the identifiable region.  This property appears to have been first noticed in \cite{McKilliam2009IndentifiabliltyAliasingPolyphase, McKilliam2009asilomar_polyest_lattice} and has been analysed further in \cite[Ch. 10]{McKilliam2010thesis}.
%
%Another estimator, described by Kitchen \cite{Kitchen_polyphase_unwrapping_1994}, is based on the phase unwrapping approach to frequency estimation suggested by Kay \cite{Kay1989}.  The primary advantage of Kitchen's estimator is that it requires only $O(N)$ arithmetic operations to compute. The primary disadvantage is that it only works correctly when the signal to noise ratio (SNR) is high. It was also observed in \cite{McKilliam2009asilomar_polyest_lattice} that Kitchen's estimator does not work correctly for all coefficients in the identifiable region, but, the problem is far less severe than for estimators based on the DPT \cite[Sec. 10.3]{McKilliam2010thesis}.

An obvious estimator of the unknown coefficients is the least squares estimator (LSE).  This is also the maximum likelihood estimator (MLE) when the noise sequence $\{X_n\}$ is white and Gaussian.  When $m=0$ (phase estimation) or $m=1$ (frequency estimation) the LSE is an effective approach, being both computationally efficient and statistically accurate \cite{Quinn2009_dasp_phase_only_information_loss,Hannan1973,Quinn2001,McKilliam_mean_dir_est_sq_arc_length2010}\cite[Sec.~6.4~and~9.1]{McKilliam2010thesis}. When $m \geq 2$ the computational complexity of the LSE is large~\cite[Sec.~10.1]{McKilliam2010thesis}\cite{Abatzoglou_ml_chirp_1986}. For this reason many authors have considered alternative approaches to polynomial phase estimation. These can loosely be grouped into two classes, estimators based on polynomial phase transforms, such as the discrete polynomial phase transform (DPT)~\cite{Peleg_DPT_1995,Peleg1991_est_class_PPS_1991,Porat_asympt_HAF_DPT_1996} and the high order phase function~\cite{Farquharson_another_poly_est_2005}, and estimators based on phase unwrapping, such as Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, the estimator of Djuric and Kay~\cite{Djuric_phase_unwrap_chirp_1990}, and Morelande's Bayesian unwrapping estimator~\cite{Morelande_bayes_unwrapping_2009_tsp}.

In this paper we consider the estimator that results from unwrapping the phase in a least squares manner.  We call this the \emph{least squares unwrapping estimator} (LSU)~\cite{McKilliam2009asilomar_polyest_lattice, McKilliamFrequencyEstimationByPhaseUnwrapping2009}\cite[Chap. 8]{McKilliam2010thesis}.  It was shown in \cite{McKilliam2009asilomar_polyest_lattice, McKilliamFrequencyEstimationByPhaseUnwrapping2009} that the LSU estimator can be represented as a \emph{nearest lattice point problem}, and Monte-Carlo simulations were used to show the LSU estimator's favourable statistical performance. %Moreover, it was noticed that the LSU appears to work correctly for coefficients anywhere in the identifiable region.  A drawback of the LSU estimator is that computing a nearest lattice point is, in general, computationally difficult.  In \cite{McKilliam2009asilomar_polyest_lattice}, two standard techniques were considered, the \emph{sphere decoder} \cite{Pohst_sphere_decoder_1981}, and \emph{Babai's nearest plane algorithm} \cite{Babai1986}. The sphere decoder was observed to have excellent statistical performance but can only be computed efficiently for $N<50$ \cite{Jalden2005_sphere_decoding_complexity}.  Computing the Babai point requires only $O(N^2)$ operations, but its statistical performance is comparatively poor at low signal to noise ratio (SNR). A major point of interest is that the lattices considered are not \emph{random}, and therefore may admit fast nearest point algorithms. 
%This has been studied in~\cite[Sec.~4.2]{McKilliam2010thesis} where polynomial time algorithms where found that compute the nearest point exactly.  Unfortantely, although polynomial time, the algorithm are still computationally very slow in practice. 
% Two main questions were raised in \cite{McKilliam2009asilomar_polyest_lattice}:
% \begin{enumerate}
% \vspace{-0.1cm}
% \item What are the asymptotic statistical properties of the LSU estimator?
% \vspace{-0.18cm}
% \item Can the structure of the lattice be exploited to obtain fast (exact or approximate) nearest point algorithms and thereby a computationally efficient estimator?
% \vspace{-0.1cm}
% \end{enumerate}
% The purpose of this paper is to answer the first question. We prove that the LSU estimator is strongly consistent and asymptotically normally distributed. 
%We also apply another known approximate nearest point algorithm, called the $K$-best method \cite{Zhan2006_K_best_sphere_decoder}, that we show provides near sphere decoder performance, but can be computed in a reasonable amount of time if $N$ is less than about 300.
In this paper we derive the asymptotic properties of the LSU estimator, showing that it is strongly consistent and asymptotically normally distributed.  Similar results were stated without a complete proof in~\cite{McKilliam_polyphase_est_icassp_2011}.  Here, we give a proof.  The results here are also more general than in~\cite{McKilliam_polyphase_est_icassp_2011}, allowing for a wider class of noise distributions.

The paper is organised in the following way. Section~\ref{sec:lattice-theory} describes the required concepts from lattice theory.  In Section~\ref{sec:ident_aliasing} we briefly describe the identifiable region that was also derived in~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}.  These identifiability results are required in order to properly understand the statistical properties of polynomial phase estimators. In Section \ref{sec:least-squar-unwr} we describe the LSU estimator and derive its asymptotic statistical properties.
%and also consider techniques for computing the LSU as a nearest lattice point problem. 
Section \ref{sec:simulations} describes the results of Monte Carlo simulations that compare the performance of the LSU estimator with existing estimators.  These simulations agree with the derived asymptotic properties.

%\subsection{Notation}
%We write random variables using capital letters, such as $X$ and $Y$ and circular random variables using the capital Greek letters $\Theta$ and $\Phi$.  We use $\round{x}$ to denote the nearest integer to $x$ with half integers rounded up and use $\fracpart{x} = x - \round{x}$ to denote the \emph{centred} fractional part.  Both $\round{\cdot}$ and $\fractpart{\cdot}$ operate elementwise on vectors.

\section{Lattice Theory}\label{sec:lattice-theory}

A \term{lattice},  $\Lambda$, is a discrete subset of points in $\reals^n$ such that
\[
   \Lambda = \{\xbf = \Bbf\ubf \mid \ubf \in \ints^d \}
\]
where $\Bbf \in \reals^{n \times d}$ is an $n \times d$ matrix of rank $d$, called the \term{generator matrix} or \term{basis matrix} or simply generator or basis.  If $n = d$ the lattice is said to be \emph{full rank}.  Lattices are discrete Abelian groups under vector addition.  They are subgroups of the Euclidean group $\reals^n$.  Lattices naturally give rise to \emph{tessellations} of $\reals^n$ by the specification of a set of coset representatives for the quotient $\reals^n / \Lambda$.  One choice for a set of coset representatives is a \emph{fundamental parallelepiped}, i.e., the parallelepiped generated by the columns of a generator matrix.  Another choice is based on the \emph{Voronoi cell}, i.e., those points from $\reals^n$ nearest (with respect to a given norm) to the lattice point at the origin.  It is always possible to construct a \emph{rectangular} set of representatives, as the next proposition will show.  We will use these rectangular regions for describing the aliasing of polynomial phase signals in Section~\ref{sec:ident_aliasing}.  These rectangular regions will also be important for the derivation of the asymptotic properties of the LSU estimator in Section~\ref{sec:least-squar-unwr}.

\begin{proposition}\label{prop:lattice-theory-constructing_a_rectangular_tesselating_region}
Let  $\Lambda$ be an $n$ dimensional lattice and $\Bbf \in \reals^{n\times n}$ be a generator matrix for $\Lambda$. Let $\Bbf = \Qbf\Rbf$ where $\Qbf$ is orthonormal and $\Rbf$ is upper triangular with elements $r_{ij}$.  Then the rectangular prism $\Qbf P$ where
\[
P = \prod_{k=1}^{n}{\left[-\frac{r_{kk}}{2}, \frac{r_{kk}}{2}\right)}
\]
is a set of coset representatives for $\reals^n / \Lambda$.
\end{proposition}
\begin{IEEEproof}
This result is well known~\cite[Chapter IX, Theorem IV]{Cassels_geom_numbers_1997}~\cite[Proposition 2.1]{McKilliam2010thesis}.  This result is for lattices with full rank.  A result in the general case can be obtained similarly, but is not required here.  
\end{IEEEproof}

\begin{figure}[tp]
	\centering
		\includegraphics[width=\linewidth]{plots/tesselationfigures-2.mps}
		\caption{Rectangular tessellation constructed according to Proposition~\ref{prop:lattice-theory-constructing_a_rectangular_tesselating_region} where $\Lambda$ is a 2 dimensional lattice with generator matrix having columns $[1, 0.2]^\prime$ and $[0.2, 1]^\prime$. Any one of the boxes is a rectangular set of coset representatives for $\reals^2 / \Lambda$.  The shaded box centered at the origin is the one given by Proposition~\ref{prop:lattice-theory-constructing_a_rectangular_tesselating_region}.}
		\label{lattices:fig:tessellation2}
\end{figure}

Given a point $\ybf\in\reals^n$ and a lattice $\Lambda$ contained in~$\reals^n$, the \emph{nearest lattice point problem} is to find the point $\xbf \in \Lambda$ such that the distance (with respect to a given norm) between $\ybf$ and $\xbf$ is minimised.  Here, the Euclidean norm is assumed, so that the nearest point is
\[ 
\arg\min_{\xbf \in \Lambda} \| \xbf - \ybf \|^2.
\]
The nearest lattice point problem is known to be NP-hard under certain conditions when the lattice itself, or rather a basis thereof, is considered as an additional input parameter~\cite{micciancio_hardness_2001, Jalden2005_sphere_decoding_complexity}. Nevertheless, algorithms exist that can compute a nearest lattice point in reasonable time if the dimension is small. One such algorithm, introduced by Pohst~\cite{Pohst_sphere_decoder_1981}, was popularised in the signal processing and communications fields by Viterbo and Boutros~\cite{Viterbo_sphere_decoder_1999} and has since been called the \emph{sphere decoder}.

Approximate algorithms for computing the nearest point have also been studied.  A classic example is \emph{Babai's nearest plane algorithm}~\cite{Babai1986}, which requires $O(n^4)$ arithmetic operations in the worst case where $n$ is the dimension of the lattice and only $O(n^2)$ if the lattice basis is \emph{Lov\'asz reduced}~\cite{Lenstra1982}\index{Lov\'asz reduced}. Recently, various approximate techniques for solving the nearest lattice point problem have been motivated by applications to multiple-input multiple-output communications.  An example is the \emph{$K$-best algorithm}~\cite{Zhan2006_K_best_sphere_decoder} that works similarly to the sequential $M$-algorithm~\cite{Anderson1984_seq_coding_alg} used in coding theory. Yet another example is the \emph{fixed sphere decoder}~\cite{Jalden2009_error_prob_fixed_sphere_decoder,Barbero2008_fixed_sphere_decoder}.

In~\cite{McKilliam2009asilomar_polyest_lattice}, it was shown how the least squares phase unwrapping estimator of a polynomial phase signal can be computed by finding a nearest point in a lattice.  We will briefly describe this result in Section~\ref{sec:least-squar-unwr}.

\section{Identifiability and aliasing}\label{sec:ident_aliasing}

As discussed in the introduction, a polynomial phase signal of order $m$ is a complex valued function of the form $s(t) = e^{2\pi j y(t)}$ where $t$ is a real number and $y(t)$ is a polynomial of order $m$. We will often drop the $(t)$ and just write the polynomial as $y$ and the polynomial phase signal as $s$ whenever there is no chance of ambiguity. %In practice the signal obtained is typically \emph{sampled} at discrete points in time, $t$. In this thesis we only consider \term{uniform sampling}\index{uniform sampling}, that is, where the gap between consecutive samples is a constant. In this case we can always consider the samples to be taken at some set of consecutive integers and our sampled polynomial phase signal looks like
%\[
%s(n) = e^{2\pi j y(n)}
%\] 
%where $n$ is an integer. The phase of $s(n)$ is described by the sampled polynomial
%\[
%y(n) = \mu_0 + \mu_1 n + \mu_2 n^2 + \dots + \mu_m n^m.
%\]

Aliasing can occur when polynomial-phase signals are sampled.  That is, two or more distinct polynomial-phase signals can take exactly the same values at the sample points.  These aliasing results are also given in~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}, but the presentation here is different, and is better suited to studying the asymptotic properties of the LSU estimator.  A description of the aliasing properties of polynomial phase signals of order $2$ has been independently discovered in~\cite{Abatzoglou_ml_chirp_1986}~and~\cite{Angeby_PPS_aliasing_2000}.  The results here are for polynomial phase signals of any order.

Let $\mathcal{Z}$ be the set of polynomials of order at most $m$ that take integer values when evaluated at integers. That is, $\mathcal{Z}$ contains all polynomials $p$ such that $p(n)$ is an integer whenever $n$ is an integer.
Let $y$ and $z$ be two \emph{distinct} polynomials such that $z = y + p$ for some polynomial $p$ in $\mathcal{Z}$. The two polynomial phase signals
\[
s(t) = e^{2\pi j y(t)} \qquad \text{and} \qquad r(t) = e^{2\pi j z(t)}
\]
are distinct because $y$ and $z$ are distinct, but if we sample $s$ and $r$ at the integers  
\begin{align*}
s(n) &= e^{2\pi j y(n)} =  e^{2\pi j y(n)} e^{2\pi j p(n)} \\
&= e^{2\pi j (y(n) + p(n))} = e^{2\pi j z(n)} = r(n)
\end{align*}
because $p(n)$ is always an integer and therefore $e^{2\pi j p(n)} = 1$ for all $n \in \ints$. The polynomial phase signals $s$ and $r$ are equal at the integers, and although they are distinct, they are indistinguishable from their samples. We call such polynomial phase signals \term{aliases}\index{alias} and immediately obtain the following theorem.

\begin{theorem}\label{thm:circpolysampledthm}
Two polynomial phase signals $s(t) = e^{2\pi j y(t)}$  and $r(t) = e^{2\pi j z(t)}$  are aliases if and only if the polynomials that define their phase, $y$ and $z$, differ by a polynomial from the set $\mathcal{Z}$, that is, $y - z \in \mathcal{Z}$.
\end{theorem}

%\begin{corollary}\label{cor:circpolysampledexp}
%The polynomial phase signals $s(t)$ and $r(t)$ are aliases if and only if $s(t) = r(t)e^{2\pi j p(t)}$ where $p$ is a polynomial from $\mathcal{Z}$.
%\end{corollary}

%It may be helpful to observe Figures~\ref{fig:circstatplot_zero},~\ref{fig:circstatplot_line},~\ref{fig:circstatplot_quad} and~\ref{fig:circstatplot_cube}. In these, the phase (divided by $2\pi$) of two distinct polynomial phase signals is plotted on the left, and on the right the principal component of the phase is plotted. The circles display the samples at the integers. Note that the samples of the principal components intersect.  The corresponding polynomial phase signals are aliases.

It may be helpful to observe Figures~\ref{fig:circstatplot_line} and~\ref{fig:circstatplot_cube}.  In these, the phase (divided by $2\pi$) of two distinct polynomial phase signals is plotted on the left, and on the right the principal component of the phase is plotted. The circles display the samples at the integers. Note that the samples of the principal components intersect.  The corresponding polynomial phase signals are aliases.

We can derive an analogue of the theorem above in terms of the coefficients of the polynomials $y$ and $z$. This will be useful when we consider estimating the coefficients in Section~\ref{sec:least-squar-unwr}.  We first need the following family of polynomials.

\begin{definition} \emph{(Integer valued polynomials)} \label{def:intvaledpolys}
\\The integer valued polynomial of order $k$, denoted by $p_k$, is
\[
p_k(x) = \binom{x}{k} = \frac{x(x-1)(x-2)\dots(x-k+1)}{k!},
\]
where we define $p_0(x) = 1$.
\end{definition}

\begin{lemma}\label{lem:intvalpol}
  The integer valued polynomials $p_0,\dots,p_m$ are an integer basis for $\mathcal{Z}$.  That is, every polynomial in $\mathcal{Z}$ can be uniquely written as
\begin{equation} \label{eq:lem_polynomial}
c_0 p_0 + c_1 p_1 + \dots + c_m p_m,
\end{equation}
where the $c_i \in \ints$.
\end{lemma}
\begin{IEEEproof}
See~\citep[p. 2]{cahen_integer-valued_1997} or~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}. 
\end{IEEEproof}
% \begin{IEEEproof}
% Note that $x(x-1)(x-2)\dots(x-k+1)$ is divisible by all integers $1,2,\dots,k$ and so $p_k$ takes integer values for all $x\in\ints$.  Then any polynomial generated as in \eqref{eq:lem_polynomial} is an element in $\mathcal{Z}$.  The proof proceeds by induction.  Consider any polynomial $f \in \mathcal{Z}$.  Let $d < n$ and assume that $c_i \in \ints$ for all $i \leq d$.  Let $g$ be the polynomial
% \[
% g = f - \sum_{k=0}^{d}{c_k p_k}
% \]
% and note that $g \in \mathcal{Z}$. Then
% \begin{equation}
% g = c_{d+1}p_{d+1} + \dots + c_{m}p_{m}.
% \end{equation}
% Now $p_{d+1}(d+1) = 1$ and $p_k(d+1) = 0$ for all $k>d+1$.  Then $g(d+1) = c_{d+1}p_{d+1}(d+1)$ and therefore $c_{d+1} = g(d+1) \in \ints$.  The proof follows by induction because $f(0) = c_0 \in \ints$.
% \end{IEEEproof}

Given a polynomial $g(x) = a_0 + a_1x + \dots + a_m x^m$, let
\[
\coef(g) = \left[ \begin{array}{ccccc} a_0 & a_1 & a_2 & \dots & a_m, \end{array} \right]^\prime,
\]
where superscript $^\prime$ indicates the transpose, denote the column vector of length $m+1$ containing the coefficients of $g$.  If $y$ and $z$ differ by a polynomial from $\mathcal{Z}$ then we can write $y = z + p$ where $p \in \mathcal{Z}$ and then also $\coef(y) = \coef(z) + \coef(p)$.
%\footnote{In group theory terminology $\coef(\cdot)$ coupled with vector addition is called a \term{group homomorphism}.\index{group homomorphism}}. 
  Consider the set
\[
L_{m+1} = \{ \coef(p) \mid p \in \mathcal{Z} \},
\]
containing the coefficient vectors corresponding to the polynomials in $\mathcal{Z}$.  Since the integer valued polynomials are a basis for $\mathcal{Z}$,
\begin{align*}
L_{m+1} &= \{ \coef(c_0 p_0 + c_1p_1 + \dots + c_mp_m) \mid c_i \in \ints \} \\
&= \{ c_0 \coef(p_0) + \dots + c_m\coef(p_m) \mid c_i \in \ints \}.
\end{align*}
Let
\[
\Pbf = \left[ \begin{array}{cccc} \coef(p_0)& \coef(p_1)& \dots& \coef(p_m)  \end{array} \right]
\]
be the $m+1$ by $m+1$ matrix with columns given by the coefficients of the integer valued polynomials.  Then,
\[
L_{m+1} = \{ \xbf = \Pbf\ubf \mid \ubf \in \ints^{m+1} \}
\]
and it is clear that $L_{m+1}$ is an $m+1$ dimensional lattice.  That is, the set of coefficients of the polynomials from $\mathcal{Z}$ forms a lattice with generator matrix $\Pbf$. We can restate Theorem~\ref{thm:circpolysampledthm} as:
\begin{corollary}\label{cor:circpolysampledcoef}
Two polynomial phase signals $s(t) = e^{2\pi j y(t)}$  and $r(t) = e^{2\pi j z(t)}$ are aliases if and only if $\coef(y)$ and $\coef(z)$ differ by a lattice point in $L_{m+1}$.
\end{corollary}

%\begin{figure}[tp]
%	\centering
%		\includegraphics[width=\linewidth]{plots/circstatfigzero-1.mps}
%		\caption{The zeroth order polynomials $\tfrac{7}{10}$ (solid line) and $\tfrac{17}{10}$ (dashed line).}
%		\label{fig:circstatplot_zero}
%\end{figure}

\begin{figure}[tp]
	\centering
		\includegraphics[width=\linewidth]{plots/circstatfig-1.mps}
		\caption{The first order polynomials $\tfrac{1}{10}(3 + 8t)$ (solid) and $\tfrac{1}{10}(33 - 2t)$ (dashed line).}
		\label{fig:circstatplot_line}
\end{figure}

%\begin{figure}[tp]
%	\centering
%		\includegraphics[width=\linewidth]{plots/circstatfigquad-1.mps}
%		\caption{The quadratic polynomials $\tfrac{1}{10} (15 - 15 t + 4 t^2)$ (solid line) and $\tfrac{1}{10}(25 -  t^2)$  (dashed line).}
%		\label{fig:circstatplot_quad}
%\end{figure}

\begin{figure}[tp]
	\centering
		\includegraphics[width=\linewidth]{plots/circstatfigcube-1.mps}
		\caption{The cubic polynomials $\tfrac{1}{160} (174 + 85 t - 118 t^2 + 40 t^3)$ (solid line) and $\tfrac{1}{48} (84 + 19 t + 12 t^2 - 4 t^3)$  (dashed line).}
		\label{fig:circstatplot_cube}
\end{figure}

For the purpose of estimating the coefficients of a polynomial phase signal we must (in order to ensure identifiability) restrict the set of allowable coefficients so that no two polynomial phase signals are aliases of each other. In consideration of Corollary~\ref{cor:circpolysampledcoef} we require that the coefficients of $y(t)$, written in vector form $\mubf$, are contained in a set of coset representatives for the quotient $\reals^{m+1}/L_{m+1}$.  We call the chosen set of representatives the \term{identifiable region}\index{identifiable region}.

As an example consider the polynomial phase signal of order zero $e^{2\pi j \mu_0} = e^{2\pi j(\mu_0 + k)}$ for any integer $k$.  In order to ensure identifiability, we must restrict $\mu_0$ to some interval of length $1$.  A natural choice is the interval $[-\nicefrac{1}{2}, \nicefrac{1}{2})$. The lattice $L_1$ is the 1-dimensional integer lattice $\ints$ and the interval $[-\nicefrac{1}{2}, \nicefrac{1}{2})$ corresponds to the Voronoi cell of $L_1$. 
When $m=1$ it turns out that a natural choice of identifiable region is the square box $[-\nicefrac{1}{2}, \nicefrac{1}{2})^2$. This corresponds with the \term{Nyquist criterion}.  The lattice $L_2$ is equal to $\ints^2$ so the box $[-\nicefrac{1}{2}, \nicefrac{1}{2})^2$ corresponds with the Voronoi cell of $L_2$.  
When $m > 1$ the identifiable region becomes more complicated and $L_{m+1} \neq \ints^{m+1}$. %However, we can always construct an identifiable region by taking a tessellating region of lattice $L_{m+1}$.  %Two convenient tessellating regions are the Voronoi cell of $L$ and the rectangular tessellating region given by~\eqref{eq:rectangular_identifiable_region}.

In general there are infinitely many choices for the identifiable region. A natural choice is the Voronoi cell of $L_{m+1}$ used in~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}. Another potential choice is a fundamental parallelepiped of $L_{m+1}$. In this paper we will use the rectangular set constructed using Proposition~\ref{prop:lattice-theory-constructing_a_rectangular_tesselating_region}. Observe that $\Pbf$ is upper triangular with $k$th diagonal element equal to $\tfrac{1}{k!}$.  So this rectangular region is
\begin{equation}\label{eq:rectangular_identifiable_region}
B = \prod_{k=0}^{m}\left[ -\frac{0.5}{k!}, \frac{0.5}{k!}  \right).
\end{equation}
We will make use of this region when deriving the statistical properties of the LSU estimator in the next section.

Given vectors $\xbf$ and $\ybf$ in $\reals^{m+1}$ we say that $\xbf \equiv \ybf \bmod L_{m+1}$ if $\xbf$ and $\ybf$ differ by a lattice point in $L_{m+1}$.  We define the function $\dealias(\xbf)$ to take $\xbf$ to its coset representative inside $B$. That is, $\dealias(\xbf) = \zbf \in B$ where $\xbf - \zbf \in L_{m+1}$.  %The effect of $\dealias(\xbf)$ is to \emph{dealias} the polynomial phase coefficients. 
When $m = 0$ or $1$ then $\dealias(\xbf) = \fracpart{\xbf}$ where $\fracpart{\xbf} = \xbf - \round{\xbf}$ denotes the (centered) fractional part and $\round{\xbf}$ denotes the nearest integer to $\xbf$ with half integers rounded upwards and both $\fracpart{\cdot}$ and $\round{\cdot}$ operate on vectors elementwise.  For $m \geq 2$ the function $\dealias(\xbf)$ can be computed by a simple sequential algorithm \cite[Sec. 7.2.1]{McKilliam2010thesis}.

\section{The Least squares unwrapping estimator}\label{sec:least-squar-unwr}

We now describe the least squares unwrapping (LSU) estimator of the polynomial coefficients. Recall that we desire to estimate the coefficients $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ from the noisy samples $Y_1, \dots, Y_N$ given in~\eqref{eq:Y_nsamplednoisey}.  We take the complex argument of the $Y_n$ and divide by $2\pi$ to obtain
\begin{equation}\label{eq:noise_circ_poly}
\Theta_n = \frac{\angle{Y_n}}{2\pi} = \fracpart{ \Phi_n + y(n) }
\end{equation}
where $\angle{z}$ denotes the complex argument of the complex number $z$, and 
\[
\Phi_n = \frac{1}{2\pi}\angle(1 + \rho^{-1}s_n^{-1}X_n)
\] 
are random variables representing the phase noise induced by the $X_n$~\cite{Tretter1985,Quinn2009_dasp_phase_only_information_loss}.  If the distribution of $X_n$ is circularly symmetric (i.e., the angle $\angle X_n$ is uniformly distributed on $[-\pi, \pi)$ and is independent of the magnitude $\abs{X_n}$) then the distribution of $\Phi_n$ is the same as the distribution of $\tfrac{1}{2\pi}\angle(1 + \rho^{-1}X_n)$.  If the $X_1, \dots, X_N$ are circularly symmetric and identically distributed, then $\Phi_1, \dots, \Phi_n$ are also identically distributed.

Let $\mubf$ be the vector $[\mu_0, \mu_1, \dots, \mu_m]$ and put,
\begin{equation} \label{eq:sumofsquaresfunction}
SS(\mubf) = \sum_{n=1}^{N}\fracpart{  \Theta_{n} - \sum_{k = 0}^{m}{\mu_k n^k} }^{2}.
\end{equation}
The least squares unwrapping estimator is defined as those coefficients $\widehat{\mu}_0, \dots, \widehat{\mu}_m$ that minimise $SS$ over the identifiable region $B$. That is, the LSU estimator is,
\begin{equation}\label{eq:hatmubfLSUdefn}
\widehat{\mubf} = \arg \min_{\mubf \in B} SS(\mubf). 
\end{equation}

It is shown in~\cite{McKilliam2009asilomar_polyest_lattice} and \cite[Sec~8.1]{McKilliam2010thesis} how this minimisation problem can be posed as a nearest lattice point problem, and we briefly discuss this here.  Write $SS$ as
\[
SS(\mubf) = \sum_{n=1}^{N}{ \left( \Theta_{n} - W_n - \sum_{k = 0}^{m}{\mu_k n^k} \right)^2 },
\]
where $W_n = \round{\Theta_{n} - \sum_{k = 0}^{m}{\mu_k n^k} }$ for $n = 1, \dots, N$ are integers called \emph{wrapping variables}.  If we consider $W_1,\dots,W_N$ as nuisance parameters to be estimated then $SS$ can be written as a function of both $\mu_0,\dots,\mu_m$ and $W_1,\dots,W_N$.  The least squares unwrapping estimator is then found by jointly minimising over $\mu_0,\dots,\mu_m$ and $W_1,\dots,W_N$.   This joint minimisation problem can be solved by computing a nearest point in a lattice.  To see this write $SS$ as a function of both $\mu_0,\dots,\mu_m$ and $W_1,\dots,W_N$ using vectors,
\begin{equation}\label{eq:SS_in_vector_form_with_u_nuisance}
SS(\mubf, \wbf) = \| \thetabf - \Xbf\mubf - \wbf \|^2,
\end{equation}
where $\|\cdot\|^2$ denotes the squared Euclidean norm, and $\Xbf$ is the rectangular Vandermonde matrix
\begin{equation}\label{eq:Xvandermonde}
\Xbf = \left[ \begin{array}{ccccc} 
1 & 1 & 1 & \cdots & 1 \\
1 & 2 & 4 & \cdots & 2^m \\
\vdots & \vdots  & \vdots & \ddots & \vdots \\
1 & N & N^2 & \cdots & N^m \\
\end{array} \right],
\end{equation}
and where the column vectors $\wbf = [W_1, \dots, W_N]^\prime$ and $\thetabf = [\Theta_1, \dots, \Theta_N]^\prime$.   For fixed $\wbf$ the minimiser of $SS(\mubf, \wbf)$ with respect to $\mubf$ is
\begin{equation} \label{eq:hat_theta_cal}
\mubf(\wbf) = (\Xbf^\prime\Xbf)^{-1}\Xbf^\prime( \thetabf - \wbf) = \Xbf^+( \thetabf - \wbf),
\end{equation}
where $\Xbf^+ = (\Xbf^\prime\Xbf)^{-1}\Xbf^\prime$ is the Moore-Penrose pseudoinverse of $\Xbf$.

We may thus compute $\widehat{\mubf}$ as $\widehat{\mubf} = \mu(\widehat{\wbf})$, where $\widehat{\wbf}$ is the minimiser of
\[
SS(\wbf) = SS(\mubf(\wbf),\wbf) = \| \Qbf\thetabf - \Qbf\wbf \|^2,
\]
where $\Qbf = \Ibf - \Xbf\Xbf^+$ and $\Ibf$ is the $N \times N$ identity matrix\footnote{We have slightly abused notation here by reusing $SS$. This should not cause any confusion as $SS(\mubf)$ and $SS(\mubf,\wbf)$ and $SS(\wbf)$ are easily told apart by their inputs.}.

Let $\Lambda$ be the lattice generated by $\Qbf$, that is $\Lambda = \{\Qbf\xbf \mid \xbf \in \ints^N \}$\footnote{The $N \times N$ projection matrix $\Qbf$ does not have full rank, it has rank $N-m-1$, and for this reason it is not strictly a generator matrix.  However, a generator can be constructed by removing $m+1$ consecutive columns from $\Qbf$. In our implementation the last $m+1$ columns are removed.}.  It follows that $\Qbf \wbf$ is a lattice point in $\Lambda$ and that minimising $SS(\wbf)$ is equivalent to finding the nearest lattice point in $\Lambda$ to $\Qbf\thetabf$.  Denote this point by $\Qbf\widehat{\wbf}$.  Now the estimate $\widehat{\mubf}$ is given by substituting $\widehat{\wbf}$ for $\wbf$ in~\eqref{eq:hat_theta_cal}.  After this procedure it is possible that the $\widehat{\mubf}$ obtained is not in the identifiable region but is instead an aliased version of the desired estimate. This can be resolved by computing $\dealias(\widehat{\mubf})$.  There exist algorithms that can compute a nearest point in $\Lambda$ that require a number of operations that is polynomial in $N$~\cite[Sec 4.3]{McKilliam2010thesis}.  Although polynomial in complexity, these algorithms are not fast in practice.  The existence of practically fast nearest point algorithms for these lattices is an interesting open problem.

The next theorem describes the asymptotic properties of this estimator.  Before we give the proof it is necessary to understand some of the properties of the phase noise $\Phi_1,\dots,\Phi_N$, which are \emph{circular} random variables with support on $[-\nicefrac{1}{2}, \nicefrac{1}{2})$~\cite{McKilliam2010thesis,McKilliam_mean_dir_est_sq_arc_length2010,Mardia_directional_statistics,Fisher1993}.  Circular random variables are often considered modulo $2\pi$ and therefore have support $[-\pi, \pi)$ with $-\pi$ and $\pi$ being identified as equivalent.  Here we instead consider circular random variables modulo 1 with support $[-\nicefrac{1}{2}, \nicefrac{1}{2})$ and with $-\nicefrac{1}{2}$ and $\nicefrac{1}{2}$ being equivalent.  This is nonstandard but it allows us to use notation such as $\round{\cdot}$ for rounding and $\fracpart{\cdot}$ for the centered fractional part in a convenient way.   %The random variable $2\pi \Phi_n$ has support on $[-\pi, \pi)$ and can be identified as a \emph{circular random variable}~\cite{Mardia_directional_statistics,Fisher1993}.  %Let $f$ be the probability density function (pdf) of $\Phi_n$.  Then $f$ takes nonzero values only on $[-\nicefrac{1}{2}, \nicefrac{1}{2})$.  It is instructive to think of $f$ as describing a describing a distribution on a circle.  For example, when the noise term $X_n$ is complex Gaussian with independent and identically distributed real and imaginary parts, the distribution of $\Phi_n$ is the projected Guassian~\cite[Section 5.6.1]{McKilliam2010thesis}\cite[page 46]{Mardia_directional_statistics}.  It is instrutive the consider a periodic version of $f$, that we write as $f(\fracpart{x})$, which has period equal to one, because the fractional part function $\fracpart{x}$ has period one.

The \emph{intrinsic mean} or \emph{Fr\'{e}chet mean} of $\Phi_n$ is defined as~\cite{McKilliam_mean_dir_est_sq_arc_length2010,bwhk07a,Bhattacharya_int_ext_means_2003,Bhattacharya_int_ext_means_2005},
\begin{equation}\label{eq:intrmeandefn}
 \mu_{\text{intr}}  = \arg \min_{\mu \in [-\nicefrac{1}{2}, \nicefrac{1}{2})} \expect \fracpart{\Phi_n - \mu}^2, 
% %&= \arg \min_{\mu \in [-\nicefrac{1}{2}, \nicefrac{1}{2})}\int_{-\nicefrac{1}{2}}^{\nicefrac{1}{2}}\fracpart{\phi - \mu}^2f(\theta) d\theta. \label{eq:minunrappedmeandef}
\end{equation}
and the \emph{intrinsic variance} is
\[
\sigma_{\text{intr}}^2 = E\fracpart{\Theta - \mu_{\text{intr}}}^2 = \min_{\mu \in [-\nicefrac{1}{2}, \nicefrac{1}{2})} \expect \fracpart{\Phi_n - \mu}^2,
\]
where $\expect$ denotes the expected value.  Depending on the distribution of $\Phi_n$ the argument that minimises~\eqref{eq:intrmeandefn} may not be unique.  The set of minima is often called the~\emph{Fr\'{e}chet mean set}~\cite{Bhattacharya_int_ext_means_2003,Bhattacharya_int_ext_means_2005}.  If the minimiser is not unique we say that $\Phi_n$ has no intrinsic mean.  We are now equipped to state the asymptotic properties of the LSU estimator.

%\begin{figure}[tp]
% 	\centering 
% 		\includegraphics[width=\linewidth]{plots/distplots.class.distributions.circular.ProjectedNormalDistribution.var3.0-1.mps}
% 		\caption{The projected normal distribution.}
% 		\label{fig:circularuniformdist}
%\end{figure}

%BLERG NOTE: theorem should be rewritten to instead assume that the X_1, \dots X_N are circularly symmetric.  Then the continuity stuff for the f follows immediately.  You can then talk about how it could apply more generally, but gets complicated.

\begin{theorem} \label{thm:asymp_proof} 
Let $\widehat{\mubf}$ be defined by~\eqref{eq:hatmubfLSUdefn} and put $\widehat{\lambdabf}_N = \dealias(\tilde{\mubf} - \widehat{\mubf})$.  Denote the elements of $\widehat{\lambdabf}_N$ by $\widehat{\lambda}_{0,N}, \dots, \widehat{\lambda}_{m,N}$.  Suppose $\Phi_1, \dots, \Phi_N$ are independent and identically distributed with zero intrinsic mean, intrinsic variance $\sigma^2$, and pdf $f$, then: 
\begin{enumerate}
\item (Strong consistency) $N^k \widehat{\lambda}_{k,N}$ converges almost surely to $0$ as $N\rightarrow\infty$ for all $k = 0, 1, \dots, m$.
\item (Asymptotic normality) If $f(\fracpart{x})$ is continuous at $x = -\nicefrac{1}{2}$ and $f(-\nicefrac{1}{2}) < 1$ then the distribution of
\[
\left[
\begin{array}
[c]{cccc}%
\sqrt{N} \widehat{\lambda}_{0,N} & N \sqrt{N}\widehat{\lambda}_{1,N}  & \dots & N^m\sqrt{N} \widehat{\lambda}_{m,N}
\end{array}
\right]^\prime
\]
converges to the normal with zero mean and covariance
\[
\frac{\sigma^2}{\left(1-f( -\nicefrac{1}{2}) \right)^{2}} \Cbf^{-1},
\]
where $\Cbf$ is the $(m+1)\times (m+1)$ Hilbert matrix with elements $C_{ik} = 1/(i + k + 1)$ for $i,k = \{0, 1, \dots, m\}$.
\end{enumerate}
\end{theorem}
The proof of Theorem~\ref{thm:asymp_proof} is contained within the next two sections. Section~\ref{sec:strongconstproof} proves strong consistency and Section~\ref{sec:centlimitproof} proves asymptotic normality.  The proofs for the case of $m=0$ were given in~\cite{McKilliam_mean_dir_est_sq_arc_length2010} and for the case when $m=1$ were given in~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.  The proofs here take a similar approach, but require new techniques.  Before giving the proof we make the following remarks.

The theorem gives conditions on the \emph{dealiased} difference $\dealias(\tilde{\mubf} - \widehat{\mubf})$ between the true coefficients $\tilde{\mubf}$ and the estimated coefficients $\widehat{\mubf}$ rather than directly on the difference $\tilde{\mubf} - \widehat{\mubf}$.   To see why this makes sense, consider the case when $m=0$, $\tilde{\mu}_0 = -0.5$ and $\widehat{\mu}_0 = 0.49$, so that $\tilde{\mu}_0 - \widehat{\mu}_0 = -0.99$.  However, the two phases are obviously close, since the phase $\pm 0.5$ are actually the same.  In this case 
\[
\dealias(\tilde{\mu}_0 - \widehat{\mu}_0) = \fracpart{\tilde{\mu}_0 - \widehat{\mu}_0} = 0.01.
\] 
The same reasoning holds for $m > 0$.

The requirement that $\Phi_1, \dots, \Phi_N$ be identically distributed will typically hold only when the complex random variables $X_1, \dots, X_N$ are identically distributed and circularly symmetric.  It would be possible to drop the assumption that $\Phi_1, \dots, \Phi_N$ be identically distributed, but this complicates the theorem statement and the proof.  In the interest of simplicity we only consider the case when $\Phi_1, \dots, \Phi_N$ are identically distributed here.  If $X_n$ is circularly symmetric with density function nonincreasing with magnitude $\abs{X_n}$, then, the corresponding $\Phi_n$ necessarily has zero intrinsic mean~\cite[Theorem 5.2, page 78]{McKilliam2010thesis}.  Thus, our theorem covers commonly used distributions for $X_1, \dots, X_N$, such as the normal distribution.

The proof of asymptotic normality places requirements on the pdf $f$ of the phase noise.  The requirement that $\Phi_1, \dots, \Phi_N$ have zero intrinsic mean implies that $f(-\nicefrac{1}{2}) \leq 1$~\cite[Lemma~1]{McKilliam_mean_dir_est_sq_arc_length2010}, so the only case not handled is when equality holds, i.e., when $f(-\nicefrac{1}{2}) = 1$ or when $f(\fracpart{x})$ is discontinuous at $x = -\nicefrac{1}{2}$. In this exceptional case other expressions for the asymptotic variance can be found (similar to \cite[Theorem 3.1]{Hotz_circle_means_2011}), but this comes at a substantial increase in complexity and for this reason we have omitted them. %Distributions that do not satisfy these requirements are unlikely to be needed in practice.

%Finally note that $\Cbf$ is a \emph{Hilbert matrix}\index{Hilbert matrix} and the elements of the inverse $\Cbf^{-1}$ are given by
 %\[
 %C^{-1}_{ik} =\frac{i+k+1}{(-1)^{i+k}}{m+i+1 \choose m-k}{m+k+1 \choose m-i}{i+k \choose i}^2.
 %\]
%Hilbert matrices are particularly ill conditioned and difficult to numerically invert.  This problem can be avoided using the above formula. 
%I MIGHT HAVE AN ERROR IN THIS FORMULA!

\section{Proof of strong consistency}\label{sec:strongconstproof}
 Substituting \eqref{eq:noise_circ_poly} into $SS$ we obtain
 \begin{align*}
SS\left( \mubf \right) &= \sum_{n=1}^{N}\fracpart{ \fracpart{ \Phi_n + \sum_{k = 0}^{m}{\tilde{\mu}_k n^k} } - \sum_{k = 0}^{m}{\mu_k n^k} }^{2} \\
&= \sum_{n=1}^{N}\fracpart{  \Phi_n + \sum_{k = 0}^{m}{(\tilde{\mu}_k - \mu_k) n^k} }^{2}.
\end{align*}
Let $\lambdabf = \dealias(\tilde{\mubf} - \mubf) = \tilde{\mubf} - \mubf - \pbf$ where $\pbf$ is a lattice point from $L_{m+1}$. From the definition of $L_{m+1}$ we have $p_0 + p_1 n + \dots + p_{m} n^m$ an integer whenever $n$ is an integer, so
\begin{align*}
\fracpart{\sum_{k=0}^{m}\lambda_k n^k } &= \fracpart{\sum_{k=0}^{m}(\tilde{\mu}_k - \mu_k - p_k) n^k } \\
&= \fracpart{\sum_{k=0}^{m}(\tilde{\mu}_k - \mu_k) n^k }.
\end{align*}
Let
\[
SS\left( \mubf \right) = \sum_{n=1}^{N}\fracpart{  \Phi_n + \sum_{k = 0}^{m}{\lambda_k n^k} }  ^{2} = N S_{N}\left( \lambdabf \right).
 \]
From the definition of the $\dealias(\cdot)$ function $\lambdabf$ is inside $B$ so the elements of $\lambdabf$ satisfy
 \begin{equation} \label{eq:identifiability}
 -\frac{0.5}{k!} \leq \lambda_k < \frac{0.5}{k!}.
 \end{equation} 
Now $\widehat{\lambdabf}_N = \dealias(\tilde{\mubf} - \widehat{\mubf})$ is the minimiser of $S_{N}$ in $B$. We shall show that $N^k\widehat{\lambda}_{k,N}\rightarrow0$ almost surely as $N\rightarrow\infty$ for all $k = 0,1, \dots, m$ and from this the proof of strong consistency follows.

We first show, in Lemma~\ref{lem:vn}, that $S_N(\lambdabf)$ converges almost surely to its expectation, $\expect S_N(\lambdabf)$,  uniformly over $\lambdabf \in B$.  We use this to prove that $\widehat{\lambdabf}_N$, the minimiser of $S_N$, converges almost
surely to the minimiser of $\expect S_N$. %We can then reason about the minimiser of $S_N$ using the minimiser of $\expect S_N$.  
This method of proof is common in the literature and is connected with what is called the \emph{uniform law of large numbers}~\cite{Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.

Let
\[
V_N(\lambdabf) =  \expect S_N(\lambdabf) = \frac{1}{N}\sum_{n=1}^{N} \expect \fracpart{  \Phi_{n}+\sum_{k = 0}^{m}{\lambda_k n^k}}  ^{2}
\]
and
 \[
 D_{N}\left( \lambdabf \right)  =  S_N(\lambdabf) - V_N(\lambdabf).
 \]
%be the difference between $S_N$ and its expectation $V_N =  \expect S_N$.  
We use order notation in the standard way, that is, for functions $h$ and $g$, we write $h(N) = O(g(N))$ to mean that there exists a constant $K > 0$ and a finite $N_0$ such that $h(N) \leq K g(N)$ for all $N > N_0$.  We use $\prob$ to denote probability.

\begin{lemma} \label{lem:vn} 
For any $\epsilon > 0$ and $0 < c < 8$,
\[
\prob \left\{ \sup_{\lambdabf \in B}\abs{ D_N(\lambdabf) } > \epsilon \right\} = O(e^{-c \epsilon^2 N}).
\]
 \end{lemma}
\begin{IEEEproof}
Consider a rectangular grid of points spaced over the identifiable region $B$.  We use $\lambdabf[\rbf]$, where $\rbf \in \ints^{m+1}$, to denote the grid point 
 %\begin{align*}
\[ 
%\lambdabf[\rbf] = &\left[  \frac{r_0}{N^{b}} - \frac{1}{2}, \; \frac{r_1}{N^{b+1}} - \frac{1}{2}, \right.\\ 
%&\dots, \; \frac{r_k}{N^{b+k}} - \frac{1}{2(k!)}, \; \dots \\
%&\left. \dots, \; \frac{r_m}{N^{b+m}} - \frac{1}{2(m!) }\right]
 \lambdabf[\rbf] = \left[  \frac{r_0}{N^{b}} - \frac{1}{2}, \; \frac{r_1}{N^{b+1}} - \frac{1}{2}, \; \dots, \; \frac{r_m}{m!N^{b+m}} - \frac{1}{2(m!) }\right]
\]
 %\end{align*}
 for some constant $b>0$.  Adjacent grid points are separated by $\tfrac{1}{N^b}$ in the first coordinate, $\tfrac{1}{N^{b+1}}$ in the second coordinate and $\tfrac{1}{k!N^{b+k}}$ in the $k$th coordinate. Let
 \[
 B[\rbf]=\left\{  \xbf\in\reals^{m+1}  ; \frac{r_k}{N^{b+k}}\leq x_k + \frac{1}{2(k!)} < \frac{r_k + 1}{N^{b+k}} \right\}  .
 \]
and let $G$ be the finite set of grid points
\[
G = \left\{ \xbf \in \ints^{m+1} \mid x_k = 0,1,2\dots,N^{b+k}-1  \right\}.
\]
The total number of grid points is $|G| = N^{(m+1)(2b + m)/2}$, and the $B[\rbf]$ partition $B$, that is, $B = \cup_{\rbf \in G}B[\rbf]$.  Now 
\begin{align}
 &\sup_{\lambdabf \in B}\sabs{ D_N(\lambdabf) }  \nonumber \\
&= \sup_{\rbf \in G}\sup_{ \lambdabf \in B[\rbf]}\sabs{ D_N(\lambdabf[\rbf]) + D_N(\lambdabf) - D_N(\lambdabf[\rbf])  } \nonumber \\
 &\leq \sup_{\rbf \in G}\sabs{ D_N(\lambdabf[\rbf])} + \sup_{\rbf \in G}\sup_{\lambdabf\in B[\rbf]}\sabs{ D_N(\lambdabf) - D_N(\lambdabf[\rbf])  }. \label{eq:gridanddensespace}
 \end{align}
From Lemma~\ref{lem:supVjk} it will follow that
\[
\prob\left\{   \sup_{\rbf \in G }\abs{ D_N(\lambdabf[\rbf])  } > \frac{\epsilon}{2} \right\} = O(e^{-c \epsilon^2 N})
\]
for any real number $c < 2$.
%that is, with high probability $D_N$ is small on all of the grid points.  
In Lemma~\ref{lem:supBVn} we show that
\[
\sup_{\rbf \in G}\sup_{\lambdabf\in B[\rbf]}\abs{ D_N(\lambdabf) - D_N(\lambdabf[\rbf])} < 2\frac{m+1}{N^b}.
\]
%that is, $D_N$ does not change much within any of the regions $B[\rbf]$ neighbouring the grid points.  
Combining these results with~\eqref{eq:gridanddensespace}, we obtain
\[
\prob\left(\sup_{\lambdabf \in B }\sabs{ D_N(\lambdabf)} > \frac{\epsilon}{2} +  \frac{2(m+1)}{N^b} \right) = O(e^{-c \epsilon^2 N}),
\]
and for sufficiently large $N$, we have $\epsilon/2 + \frac{2(m+1)}{N^b} < \epsilon$ completing the proof.  It remains to prove Lemmas~\ref{lem:supVjk} and~\ref{lem:supBVn}.
\end{IEEEproof}

We first need the following special case of Hoeffding's inequality~\cite{Hoeffding_inequality_1963}.

\begin{lemma} \label{lem:zero_mean_indepent_sum_bound}
  Let $\bar{Z} = (Z_1 + \dots + Z_N)/N$ where $Z_1, \dots, Z_N$ are independent, zero-mean random variables with $|Z_n| \leq \tfrac{1}{4}$ for each $n$.  Then,
\[
  \prob\{ \sabs{\bar{Z}} > \epsilon \} \leq 2e^{ - 8 \epsilon^2 N}.
\]
\end{lemma}

\begin{lemma}\label{lem:supVjk}
For any $\epsilon > 0$ and $0 < c < 8$,
\[
\prob\cubr{   \sup_{\rbf \in G}\abs{ D_N(\lambdabf[\rbf])  } > \epsilon } = O(e^{-c \epsilon^2 N}).
\]
\end{lemma}
 \begin{IEEEproof}
Fix $\lambdabf$ and write
 \[
 D_N(\lambdabf) = \bar{Z} = \frac{1}{N}  \sum_{n=1}^{N}Z_{n},
 \]
 where
 \[
 Z_{n}=\fracpart{  \Phi_n+\sum_{k = 0}^{m}{\lambda_k n^k} }^{2} - E\fracpart{  \Phi_n + \sum_{k = 0}^{m}{\lambda_k n^k} }^{2}
 \]
 are independent with zero mean and $\abs{Z_n} \leq \tfrac{1}{4}$. From Lemma~\ref{lem:zero_mean_indepent_sum_bound},
 \begin{equation}
  \prob( \abs{D_N(\lambdabf)} > \epsilon ) \leq 2e^{ - 8 \epsilon^2 N}.
 \label{eq:markapp}%
 \end{equation}
But,
 \begin{align*}
 \prob\cubr{  \sup_{\rbf\in G}\left\vert D_N(\lambdabf[\rbf]) \right\vert >\epsilon }  &\leq \sum_{\rbf \in G} \prob\scubr{  \left\vert D_N(\lambdabf[\rbf])  \right\vert >\epsilon } \nonumber \\
&= 2 |G| e^{ - 8 \epsilon^2 N} \nonumber \\
&= O(e^{-c \epsilon^2 N}), \label{eq:betaprobconv}
\end{align*}
where $c$ is any real number less than $8$, since $|G| = N^{(m+1)(2b + m)/2}$ is polynomial in $N$.
\end{IEEEproof}

Before proving Lemma~\ref{lem:supBVn} we need the following result.

\begin{lemma}\label{lem:boundedsquarefracparts}
Let $x$ and $\delta$ be real numbers.  Then
\[
\fracpart{x}^2 - |\delta| \leq \fracpart{x + \delta}^2 \leq \fracpart{x}^2 + |\delta|.
\]
\end{lemma}
\begin{IEEEproof}
Since $\abs{\delta} \leq \abs{n + \delta}$ for all $\delta \in [-\nicefrac{1}{2}, \nicefrac{1}{2})$ and $n \in \ints$, the result will follow if we can show that it holds when both $x$ and $\delta$ are in $[-\nicefrac{1}{2}, \nicefrac{1}{2})$.  Also, for reasons of symmetry, we need only show that it holds when $\delta \geq 0$.  Now
\[
\fracpart{x + \delta}^2 - x^2 = \begin{cases}
2x\delta + \delta^2, & x \in [-\nicefrac{1}{2}, \nicefrac{1}{2} - \delta) \\
2x(\delta-1) + (\delta-1)^2, & x \in [\nicefrac{1}{2} - \delta, \nicefrac{1}{2})
\end{cases}
\] 
But, when $x \in [-\nicefrac{1}{2}, \nicefrac{1}{2} - \delta)$, 
\[
-1 \leq -1 + \delta \leq 2x + \delta < 1 - \delta \leq 1,
\]
and so
\[
-\delta \leq (-1 + \delta)\delta \leq (2x + \delta)\delta < (1-\delta)\delta \leq \delta.
\]
Also, when $x \in [\nicefrac{1}{2} - \delta, \nicefrac{1}{2})$,
\[
-\delta \leq 2x + \delta - 1 < \delta,
\]
and consequently
\[
-\delta \leq -\delta(1-\delta) \leq (2x + \delta - 1)(1 - \delta) \leq \delta(1 - \delta) \leq \delta.
\]
%and consequenly
%\[
%-\delta \leq (2x + \delta - 1)(\delta - 1) \leq \delta.
%\]

% In some sense this result is `obvious' because  the function $\fracpart{x}^2$ is continuous and piecewise differentiable and the derivative has magnitude less than one whenever it exists.  We will give a rigorous proof of the lemma when $\delta \geq 0$, the proof for $\delta \leq 0$ is similar. Note that $0 \leq \fracpart{x}^2 \leq \tfrac{1}{4}$ for all real numbers $x$. So if $\delta \geq \tfrac{1}{4}$ then the lemma is true because
% \[
% \fracpart{x}^2 - |\delta| \leq 0 \leq \fracpart{x + \delta}^2 \leq \tfrac{1}{4} \leq \fracpart{x}^2 + |\delta|.
% \]
% So we may assume that $\delta < \tfrac{1}{4}$. We consider two cases, firstly when $\fracpart{x} + \delta \in [-\nicefrac{1}{2}, \nicefrac{1}{2})$, and secondly when $\fracpart{x} + \delta \geq \nicefrac{1}{2}$. Assume that $\fracpart{x} + \delta \in [-\nicefrac{1}{2}, \nicefrac{1}{2}]$ and therefore
% \[
% -\tfrac{1}{2} \leq \fracpart{x} \leq \tfrac{1}{2} - \delta
% \]
% and also
% \[
% \fracpart{x + \delta}^2 = \fracpart{\fracpart{x} + \delta}^2 = (\fracpart{x} + \delta)^2 = \fracpart{x}^2 + 2\fracpart{x}\delta + \delta^2.
% \]
% Substituting the bounds on $\fracpart{x}$ above into the right hand side of this equation we obtain
% \[
% \fracpart{x}^2 - \delta(1 - \delta) \leq \fracpart{x + \delta}^2 \leq \fracpart{x}^2 + \delta(1 - \delta).
% \]
% The lemma is true because $\delta(1 - \delta) \leq |\delta|$ since $0 \leq \delta < \tfrac{1}{4}$. Now assume that $\fracpart{x} + \delta \geq \nicefrac{1}{2}$. We have
% \[
% \fracpart{x + \delta} = \fracpart{x} + \delta - 1
% \]
% and therefore
% \[
% \tfrac{1}{2} - \delta \leq \fracpart{x} \leq \tfrac{1}{2}
% \]
% and also 
% \[
% \fracpart{x + \delta}^2 = (\fracpart{x} + \delta - 1)^2 = \fracpart{x}^2 + 2\fracpart{x}(\delta  - 1) + (\delta - 1)^2.
% \]
% Substituting the bounds on $\fracpart{x}$ above into the right hand side of this equation we again obtain
% \[
% \fracpart{x}^2 - \delta(1 - \delta) \leq \fracpart{x + \delta}^2 \leq \fracpart{x}^2 + \delta(1 - \delta)
% \]
% and the lemma is true because $\delta(1 - \delta) \leq |\delta|$ since $0 \leq \delta < \tfrac{1}{4}$.
\end{IEEEproof}

 \begin{lemma}\label{lem:supBVn}
For all $N$,
\[
 \sup_{\rbf\in G}\sup_{ \lambdabf  \in B[\rbf] }\left\vert D_N\left(  \lambdabf \right) - D_N\left(  \lambdabf[\rbf] \right)\right\vert < 2\frac{m+1}{N^b}.
 \]
 \end{lemma}
 \begin{IEEEproof}
%The proof we give is not just almost surely but \emph{surely} as the convergence will be shown to occur irrespective of the values of the $\Phi_n$. 
For convenience put 
\[
b_n = \Phi_n + \sum_{k = 0}^{m}{\lambda_k n^k} \qquad \text{and} \qquad  a_n = \Phi_n + \sum_{k = 0}^{m}{\lambda[\rbf]_k n^k},
\] 
where $\lambda[\rbf]_k$ denotes the $k$th element of the grid point $\lambdabf[\rbf]$. For $\lambdabf \in B[\rbf]$,
\[
b_n = a_n + \delta_n,
\]
where 
\[
\sabs{\delta_n} \leq \sum_{k=0}^m\frac{n^k}{k! N^{b+k}} \leq \frac{m+1}{N^{b}}.
\] 
Taking fractional parts, squaring, and applying Lemma~\ref{lem:boundedsquarefracparts} we see that, for all real numbers $x$,
\[
 -|\delta_n| \leq \fracpart{x + b_n}^2 -  \fracpart{x + a_n}^2  \leq |\delta_n|,
\]
and $|\fracpart{x + b_n}^2 -  \fracpart{x + a_n}^2| \leq \tfrac{m+1}{N^b}$. Now
\[
S_N(\lambdabf) - S_N(\lambdabf[\rbf]) =  \frac{1}{N}\sum_{n=1}^{N}\big( \fracpart{\Phi_n + b_n}^2 -  \fracpart{\Phi_n + a_n}^2 \big)
\]
and therefore, for all $\lambdabf \in B[\rbf]$,
\[
\sabs{S_N(\lambdabf) - S_N(\lambdabf[\rbf])} \leq \frac{m+1}{N^b}.
\]
As this bound is independent of $\Phi_1 \dots \Phi_N$, we have
\[
\sabs{ V_N(\lambdabf) - V_N(\lambdabf[\rbf])} \leq \expect \sabs{ S_N(\lambdabf) - S_N(\lambdabf[\rbf])} \leq \frac{m+1}{N^b}
\]
by Jensen's inequality.  Therefore, for all $\lambdabf \in B[\rbf]$,
 \begin{align*}
 \sabs{ D_N(\lambdabf) - &D_N(\lambdabf[\rbf]) } \\
&= \sabs{ S_N(\lambdabf) - S_N(\lambdabf[\rbf]) + V_N(\lambdabf) - V_N(\lambdabf[\rbf]) } \\
&\leq \sabs{ S_N(\lambdabf) - S_N(\lambdabf[\rbf]) \vert + \vert V_N(\lambdabf) - V_N(\lambdabf[\rbf]) },  \\
&\leq 2\frac{m+1}{N^{b}}
\end{align*}
 and as this bound is independent of $\rbf$,
\[
\sup_{\rbf\in G}\sup_{ \lambdabf  \in B[\rbf] }\sabs{ D_N\left(  \lambdabf \right) - D_N\left(  \lambdabf[\rbf] \right) } \leq 2\frac{m+1}{N^{b}}.
\]
 \end{IEEEproof}

The proof of Lemma~\ref{lem:vn} is now complete.  We concentrate our attention on the minimiser of $V_N$. Because $\Phi_n$ has zero intrinsic mean 
\begin{equation}\label{eq:Efracpartmined}
\expect \fracpart{ \Phi_n + z }^{2}
\end{equation}
is minimised uniquely at $z = 0$ for $z \in [-\nicefrac{1}{2}, \nicefrac{1}{2})$.  Since the intrinsic variance of $\Phi_n$ is $\sigma^2$, when $z = 0$,
\begin{equation}\label{eq:Efracpartphi}
\expect\fracpart{\Phi_1+z}^{2} = \expect\fracpart{\Phi_1}^{2} = \sigma^2,
\end{equation}
and so the minimum attained value is $\sigma^2$.

\begin{lemma}\label{lem:ES_Nminimisedzero}
For $\lambdabf \in B$ the function $V_N(\lambdabf)$ is minimised uniquely at $\zerobf$, the vector of all zeros.  At this minimum $V_N(\zerobf) = \sigma^2$.
\end{lemma}
\begin{IEEEproof}
Let $z$ be the polynomial
\[
z(n) = \lambda_0 + \lambda_1 n + \lambda_2 n^2 + \dots + \lambda_m n^m.
\]
Then
\begin{align*}
V_N(\lambdabf) &= \frac{1}{N}\expect\sum_{n=1}^{N}\fracpart{ \Phi_n + \sum_{k=0}^m{\lambda_k n^k} }^2 \\
&= \frac{1}{N}\sum_{n=1}^{N}\expect\fracpart{ \Phi_n + \fracpart{z(n)} }^2.
\end{align*}
We know that $\expect\fracpart{ \Phi_n + \fracpart{z(n)} }^2$ is minimised uniquely when $\fracpart{z(n)} = 0$ at which point it takes the value $\sigma^2$. Now $\fracpart{z(n)}$ is equal to zero for all integers $n$ if and only if $z \in \mathcal{Z}$, or equivalently if $\coef(z)$ is a lattice point in $L_{m+1}$. By definition $B$ contains precisely one lattice point from $L_{m+1}$, this being the origin $\zerobf$. Therefore $V_N$ is minimised uniquely at $\zerobf$, at which point it takes the value $\sigma^2$.
\end{IEEEproof}

\begin{lemma} \label{lem:ESNconv}
For any $\epsilon > 0$ and $0 < c < 2$,
\[
\prob\left\{ \abs{V_N(\widehat{\lambdabf}_N) - \sigma^2} > \epsilon \right\} = O(e^{-c \epsilon^2 N}).
\]
\end{lemma}
\begin{IEEEproof}
By definition
\[
\widehat{\lambdabf}_N =\arg\min_{\lambdabf\in B} S_N(\lambdabf)
\]
and therefore
\[
 0 \leq S_{N}(\zerobf) - S_N(\widehat{\lambdabf}_N).
\]
Also, because $V_N$ is minimised at $\zerobf$,
\[
 0 \leq V_N(\widehat{\lambdabf}_N) - V_N(\zerobf).
\]
Thus,
\begin{align*}
0 &\leq V_N(\widehat{\lambdabf}_N) - V_N(\zerobf)  \\
&\leq V_N(\widehat{\lambdabf}_N) - V_N(\zerobf) + S_{N}(\zerobf) - S_N(\widehat{\lambdabf}_N)   \\
&\leq \sabs{ V_N(\widehat{\lambdabf}_N) - S_N(\widehat{\lambdabf}_N) } + \sabs{ S_{N}(\zerobf) - V_N(\zerobf) }\\
&= \sabs{ D_N(\widehat{\lambdabf}_N) } + \sabs{ D_N(\zerobf) },
\end{align*}
and from Lemma~\ref{lem:vn},
\begin{align*}
\prob&\left\{ \sabs{V_N(\widehat{\lambdabf}_N) - V_N(\zerobf))} > \epsilon \right\} \\
&\leq \prob\left\{ \sabs{ D_N(\widehat{\lambdabf}_N) } + \sabs{ D_N(\zerobf) } > \epsilon \right\} \\
&\leq \prob\left\{ \sabs{ D_N(\widehat{\lambdabf}_N) } > \frac{\epsilon}{2}\right\} + \prob\left\{ \sabs{ D_N(\zerobf) } > \frac{\epsilon}{2} \right\} \\
&= O(e^{-d \epsilon^2 N / 4}),
\end{align*}
where $d$ is any real number less than $8$.
%where the third line follows from Lemma~\ref{lem:ABprob}.
\end{IEEEproof}

\begin{corollary}  \label{cor:ESNconvcor}
$\sabs{V_N(\widehat{\lambdabf}_N) - \sigma^2}$ converges almost surely to zero as $N\rightarrow\infty$.
\end{corollary}
\begin{IEEEproof}
From Lemma~\ref{lem:ESNconv}, for any $\epsilon > 0$ and $0 < c < \tfrac{1}{2}$,
\[
\sum_{N=1}^{\infty}\prob\left\{ \abs{V_N(\widehat{\lambdabf}_N) - \sigma^2} > \epsilon \right\} = \sum_{N=1}^{\infty} O(e^{-c \epsilon^2 N}) < \infty.
\]
The corollary follows from the Borel-Cantelli lemma.
\end{IEEEproof}

% \begin{lemma}\label{lem:ABprob}
% Let $A$ and $B$ be positive random variables.  Then
% \[
% \prob( A + B > \epsilon ) \leq \prob( A > \tfrac{\epsilon}{2}) + \prob( A > \tfrac{\epsilon}{2}).
% \]
% \end{lemma}
% \begin{IEEEproof}
% Write 
% \begin{equation}\label{eq:A+B>epsilon}
% \prob( A + B > \epsilon ) = 1 - \prob( A + B \leq \epsilon ).
% \end{equation}
% If $A \leq \tfrac{\epsilon}{2}$ and $B \leq \tfrac{\epsilon}{2}$ then $A + B \leq \epsilon$, so
% \begin{align*}
% \prob( A + B \leq \epsilon ) &\geq \prob( A \leq \tfrac{\epsilon}{2} \;\; \text{and} \;\; B \leq \tfrac{\epsilon}{2} ) \\
% &= 1 - \prob( A > \tfrac{\epsilon}{2} \;\; \text{or} \;\; B > \tfrac{\epsilon}{2} ) \\
% &\geq 1 - \prob( A > \tfrac{\epsilon}{2}) - \prob( B > \tfrac{\epsilon}{2} ).
% \end{align*}
% Subtituting this into~\eqref{eq:A+B>epsilon} gives the required result.
% \end{IEEEproof}

We have now shown that $V_N$ is uniquely minimised at $\zerobf$, that $V_N(\zerobf) = \sigma^2$, and that $V_N(\widehat{\lambdabf}_N)$ converges almost surely to $\sigma^2$.  These results are enough to show that $\widehat{\lambdabf}_N$ converges almost surely to zero.  However, this tells us nothing about the rate at which the components of $\widehat{\lambdabf}_N$ approach zero as required by Theorem~\ref{thm:asymp_proof}.  To prove these stronger properties we need some preliminary results about arithmetic progressions, and from the calculus of finite differences.

 %\begin{equation}
 %E\left(  \sum_{n=1}^{N}z_{n}\right)  ^{2\beta}=\sum_{n_{1},\ldots,n_{2\beta
 %}=1}^{N}E\left(  \prod\limits_{j=1}^{2\beta}z_{n_{j}}\right)  .
 %\label{eq:markapp}%
 %\end{equation}
 %Since the terms%
 %\[
 %E\left(  \prod\limits_{j=1}^{2\beta}z_{n_{j}}\right)
 %\]
 %in this containing any $n_{j}$ only once will have $0$ mean, and since the
 %$z_{j}$ are bounded, it follows that%
 %\[
 %\frac{1}{N^{\beta}}E\left(  \sum_{n=1}^{N}z_{n}\right)  ^{2\beta}=O\left(
 %1\right)  ,
 %\]
 %as $N\rightarrow\infty,$ the dominant terms coming from the $z_{j}$ in
 %$\left(  \ref{eq:markapp}\right)  $ occurring in pairs.

Let $W = \{1,2,\dots, N\}$ and let $K$ be a subset of $W$.  For any integer $h$, let
\begin{equation} \label{eq:S(h,G)def} 
A(h,K) = \big\{ n \mid n + ih \in K \;\forall\; i \in \{0,1,\dots,m\} \big\}
\end{equation}
be the set containing all integers $n$ such that the arithmetic progression
\[
n, \,\, n + h, \,\, n + 2h, \,\, \dots, \,\, n + mh
\]
of length $m+1$ is contained in the subset $K$.  If $K$ is a small subset of $W$ then $A(h,K)$ might be empty, i.e., there may be no arithmetic progression $n, n+h, \dots$ of length $m+1$ in $K$. However, the next two lemmas and the following corollary will show that if $K$ is sufficiently large then it always contains at least one arithmetic progression (for all sufficiently small $h$) and therefore $A(h,K)$ is not empty. Denote by $K \backslash \{r\}$ the set $K$ with the element $r$ removed.

\begin{lemma} \label{lem:S(h,G/r)size}
Let $r \in K$.  For any $h$, removing $r$ from $K$ removes at most $m+1$ arithmetic progressions $n, n+h, \dots n+mh$ of length $m+1$.  That is,
\[
|A(h,K \backslash \{r\})| \geq |A(h,K)| - (m+1).
\]
\end{lemma}
\begin{IEEEproof}
The proof follows because there are at most $m+1$ integers, $n$, such that $n+ih = r$ for some $i \in \{0,1,\dots,m\}$.  That is, there are at most $m+1$ arithmetic progressions of type $n, n+h, \dots n+mh$ that contain $r$.
\end{IEEEproof}

 \begin{lemma} \label{lem:S(h,K)size}
 Let $K \subseteq W$.  Then 
\[
|A(h,K)| \geq N - mh - (N - |K|)(m+1).
\]
 \end{lemma}
 \begin{IEEEproof}
 Note that $|A(h,W)| = N - mh$.  The proof follows by starting with $A(h,W)$ and applying Lemma~\ref{lem:S(h,G/r)size} 
\[
|W|-|K| = N - |K|
\] 
times. That is, $K$ can be constructed by removing $N - |K|$ elements from $W$ and this removes at most $(N - |K|)(m+1)$ arithmetic progressions from $A(h,W)$.
 \end{IEEEproof}

 \begin{corollary} \label{cor:S(h,K)>0}
 Let $K \subseteq W$ such that $|K| > \frac{2m+1}{2m+2}N$. Then for all $h$ such that $1\leq h \leq\frac{N}{2m}$ the set $K$ contains at least one arithmetic progression $n, n+h, \dots, n+mh$ of length $m+1$. That is, $|A(h,K)| > 0$.
 \end{corollary}
 \begin{IEEEproof}
 By substituting the bounds $|K| > \frac{2m+1}{2m+2}N$ and $h \leq\frac{N}{2m}$ into the inequality from Lemma~\ref{lem:S(h,K)size} we immediately obtain $|A(h,K)| > 0$.
 \end{IEEEproof}

The next result we require comes from the calculus of finite differences. For any function $d(n)$ mapping $\reals$ to $\reals$, let 
\[
\Delta_h d(n) = d(n+h) - d(n)
\] 
denote the first difference with interval $h$. Let 
\begin{align*}
\Delta_h^2 d(n) &= \Delta_h d(n+h) - \Delta_h d(n) 
\\ &= d(n+2h) - 2d(n+h) + d(n)
\end{align*}
denote the second difference with interval $h$ and similarly let 
\begin{equation}\label{eq:mthdiffformula}
\begin{split}
\Delta_h^r d(n) &= \Delta_h^{r-1} d(n+h) - \Delta_h^{r-1} d(n)
\\ &= \sum_{k=0}^{r}\binom{r}{k}(-1)^{r-k}d(n+kh)
\end{split}
\end{equation}
denote the $r$th difference with interval $h$. It is well known that 
\[
\sum_{k=0}^{r}\binom{r}{k} = 2^r.
\] 
So $\Delta_h^r d(n)$ can be represented by adding and subtracting the 
\[
d(n), \,\, d(n+h), \,\, \dots, \,\, d(n+kh)
\] 
precisely $2^r$ times.

The operator $\Delta_h$ has special properties when applied to polynomials. If $d(n) = a_r n^r + \dots + a_0$ is a polynomial of order $r$ then
 \begin{equation} \label{eq:mfinitediffpoly}
 \Delta_h^r d(n) = h^r r! a_r. 
 \end{equation}
So, the $r$th difference of the polynomial is a constant depending on $h$, $r$ and the $r$th coefficient $a_r$~\cite[page 51]{Jordan_Calculus_of_finite_difference_1965}.  We can now continue the proof of strong consistency.  The next lemma is a key result.

\begin{lemma}\label{lem:moran2}
Suppose $\lambdabf_1, \lambdabf_2,\dots$ is a sequence of vectors from $B$ with $V_N(\lambdabf_N) - \sigma^2\rightarrow 0$ as $N\rightarrow\infty$. Then the elements $\lambda_{0,N}, \dots \lambda_{m,N}$ of $\lambdabf_N$ satisfy $N^k\lambda_{k, N}\rightarrow0$ as $N\rightarrow\infty$.
\end{lemma}
\begin{proof}
Define the function
\begin{equation}\label{eq:gz}
g(z) = \expect\fracpart{\Phi_1 + z}^2 - \sigma^2
\end{equation}
which is continuous in $z$. Because of~\eqref{eq:Efracpartmined} and~\eqref{eq:Efracpartphi}, $g(z) \geq 0$ with equality only at $z = 0$ for $z \in [-\nicefrac{1}{2}, \nicefrac{1}{2})$. Now
\[
V_N(\lambdabf_N) - \sigma^2 = \frac{1}{N}\sum_{n=1}^{N} g\left( \fracpart{ \sum_{k=0}^{m}{n^k \lambda_{k,N}} } \right) \rightarrow 0
\]
as $N \rightarrow \infty$. Let
\[
z_N(n) = \lambda_{0,N} + \lambda_{1,N} n + \lambda_{2,N} n^2 + \dots + \lambda_{m,N} n^m
\]
so that
\[
V_N(\lambdabf_N) - \sigma^2 = \frac{1}{N}\sum_{n=1}^{N} g\left( \fracpart{z_N(n)} \right) \rightarrow 0.
\]
%At this stage we can \emph{see} the proof. Note that, due to the definition of the identifiable region, $B$, the only circular polynomial $\phi$ with values that are all zeros (and hence sends the sum above to zero) is the circular polynomial whose coefficients are all zeros.  Obviously, this observation does not constitute a proof, and in particular gives us little indication as to the order of convergence of the coefficients as required by the theorem. To make the proof 
Choose constants 
\[
c = \frac{2m+1}{2m+2} \qquad \text{and} \qquad 0 < \delta < \frac{1}{2^{2m+1}}
\]
and define the set 
\[
K_{N}=\left\{  n\leq N \mid \sabs{\fracpart{z_N(n)}} < \delta \right\}.
\]
There exists $N_0$ such that for all $N > N_0$ the number of elements in $K_N$ is at least $cN$.  Too see this, suppose that $\sabs{K_N} < cN$, and let $\gamma$ be the minimum value of $g$ over $[-\nicefrac{1}{2},-\delta] \cup [\delta, \nicefrac{1}{2})$. Because $g(0) = 0$ is the unique minimiser of $g$, then $\gamma$ is strictly greater than $0$ and
\begin{align*}
V_N(\lambdabf_N) - \sigma^2 &= \frac{1}{N}\sum_{n=1}^{N} g\left( \fracpart{z_N(n)} \right) \\
&\geq \frac{1}{N}\sum_{n\in K_N} \gamma = (1-c)\gamma,
\end{align*}
violating that $V_N(\lambdabf_N) - \sigma^2$ converges to zero as $N \rightarrow \infty$.  We will assume $N > N_0$ in what follows.

From Corollary \ref{cor:S(h,K)>0} it follows that for all $h$ satisfying $1\leq h \leq\frac{N}{2m}$ the set $A(h,K_N)$ contains at least one element, that is, there exists $n' \in A(h,K_N)$ such that all the elements from the arithmetic progression $n', n'+h, \dots, n' + mh$ are in $K_N$ and therefore 
\[
\sabs{\fracpart{z_N(n')}},\,\, \sabs{\fracpart{z_N(n'+h)}}, \,\, \dots, \,\, \sabs{\fracpart{z_N(n'+mh)}} 
\] 
are all less than $\delta$.  Because the $m$th difference is a linear combination of $2^{m}$ elements (see~\eqref{eq:mthdiffformula}) from 
\[
\fracpart{z_N(n')}, \,\, \fracpart{z_N(n'+h)}, \,\, \dots, \,\, \fracpart{z_N(n'+mh)}
\]
all with magnitude less than $\delta$ we obtain, from Lemma~\ref{lem:fracpartsumanddelta},
\begin{equation}\label{eq:Deltazfracbound}
|\fracpart{\Delta_h^m z_N(n')}| \leq |\Delta_h^m \fracpart{ z_N(n')}| \leq \Delta_h^m |\fracpart{ z_N(n')}|< 2^{m}\delta.
\end{equation}
From \eqref{eq:mfinitediffpoly} it follows that the left hand side is equal to a constant involving $h$, $m$ and $\lambda_{m,N}$ giving the bound
\begin{equation}\label{eq:startiterativearg}
|\fracpart{ h^m m! \lambda_{m,N} }|  = |\fracpart{   \Delta_h^m z_N(n') }| < 2^m\delta
\end{equation}
for all $h$ satisfying $1\leq h \leq\frac{N}{2m}$. Setting $h = 1$ and recalling from~\eqref{eq:identifiability} that $\lambda_{m,N} \in [-\tfrac{0.5}{m!}, \tfrac{0.5}{m!})$, we have
 \[
 |\fracpart{ m! \lambda_{m,N} }| = | m! \lambda_{m,N} |< 2^m\delta.
 \]
Now, because we chose $\delta < \tfrac{1}{2^{2m}}$ it follows that 
\[
| \lambda_{m,N} |< \tfrac{2^m}{m!}\delta < \tfrac{1}{m! 2^{m+1}}.
\]
So, when $h = 2$, 
\[
|\fracpart{ 2^m m! \lambda_{m,N} }| = | 2^m m! \lambda_{m,N} |< 2^m\delta
\]
because $2^m m! \lambda_{m,N} \in [-0.5, 0.5)$. Therefore
\[
| \lambda_{m,N} |< \frac{1}{m!}\delta < \frac{1}{m! 2^{2m+1}}.
\]
Now, with $h = 4$, we similarly obtain 
\[
|\fracpart{ 4^m m! \lambda_{m,N} }| = | 4^m m! \lambda_{m,N} |< 2^m\delta
\]
and iterating this process we eventually obtain 
\[
| \lambda_{m,N} | < \frac{2^m}{2^{um} m!}\delta
\]
where $2^u$ is the largest power of 2 less than or equal to $\tfrac{N}{2m}$.  So $2^{u+1} > \frac{N}{2m}$ and substituting this into the inequality above gives 
 \[ 
 |\lambda_{m,N}| < \frac{2^m}{ m! \left(\frac{N}{4m}\right)^m}\delta
 \]
from which it follows that 
 \begin{equation}\label{eq:enditerativearg}
 N^m|\lambda_{m,N}| < \frac{2^{2m+m}m^m}{m!}\delta
 \end{equation}
for all $N > N_0$.  As $\delta$ is arbitrary, $N^m \lambda_{m,N} \rightarrow 0$ as $N\rightarrow \infty$.

We have now shown that the highest order coefficient $\lambda_{m,N}$ converges as required. The remaining coefficients will be shown to converge by induction.  Assume that $N^k \lambda_{k,N} \rightarrow 0$ for all $k=r+1, r+2, \dots, m$, that is, assume that the $m-r$ highest order coefficients all converge as required. Let
\[
z_{N,r}(n) = \lambda_{0,N} + \lambda_{1,N} n + \lambda_{2,N} n^2 + \dots + \lambda_{r,N} n^r.
\]
Because the $m-r$ highest order coefficients converge we can write $z_N(n) = z_{N,r}(n) + \gamma(n)$ where $\sup_{n\in\{1,\dots,N\}}\abs{\gamma(n)} \rightarrow 0$ as $N\rightarrow\infty$. Now the bound from \eqref{eq:Deltazfracbound}, but applied using the $r$th difference, gives
 \begin{equation}\label{eq:zrbound}
\begin{split}
 \left|\fracpart{  \Delta_h^r z_N(n')}\right| &= \left|\fracpart{  \Delta_h^r\gamma(n') + \Delta_h^r z_r(n') }\right| \\
&= |\fracpart{ \epsilon + h^r r! \lambda_{r,N} }| < 2^r\delta,
\end{split}
 \end{equation}
 where
\[
\epsilon = \Delta_h^r \gamma(n') \leq 2^r \sup_{n\in\{1,\dots,N\}}\abs{\gamma(n)} \rightarrow 0
\] 
as $N\rightarrow\infty$.  Choose $\delta$ and $\epsilon$ such that $2^r\delta < \tfrac{1}{4}$ and $|\epsilon| < \tfrac{1}{4}$.  Then, from \eqref{eq:zrbound} and from Lemma~\ref{lem:fracpartinternalsumlessdelta},
\[
\abs{\fracpart{ h^r r! \lambda_{r,N} }} < 2^r\delta + \abs{\epsilon}
\]
for all $h$ such that $1 \leq h \leq \tfrac{N}{2m}$.  Choosing $2^r\delta + |\epsilon| < \frac{1}{2^{2r+1}}$ and using the same iterative process as for the highest order coefficient $\lambda_{m,N}$  (see~\eqref{eq:startiterativearg}~to~\eqref{eq:enditerativearg}) we find that $N^r \lambda_{r,N} \rightarrow 0$ as $N\rightarrow\infty$.  The proof now follows by induction.
 \end{proof}

% \begin{lemma}\label{sec:proof-strong-cons}
% For any constants $0 \leq c < 1$ and $\delta>0$ there exists an $N_{0}$ such that for all $N > N_0$ the proportion of $\fracpart{z_N(n)}$ with magnitude less than $\delta$ is greater than $c$.  That is, the set
% \[
% K_{N}=\left\{  n\leq N \mid \vert \fracpart{z(n)} \vert < \delta \right\} 
% \]
% has more than $c N$ elements for all $N > N_0$.
% \end{lemma}
% \begin{proof}
% Assume not.  Then for every $N_0$ there exists an $N > N_0$ such that there are more than $(1-c)N$ integers from $1$ to $N$ with $|\fracpart{z(n)}| > \delta$.  Let $\gamma$ be the minimum value of $g$ from~(\ref{eq:gz}) over the interval given by the union $[-\nicefrac{1}{2},-\delta] \cup [\delta, \nicefrac{1}{2})$. Because $g$ is minimised uniquely at $0$ then $\gamma$ is strictly greater than $0$ and the sum
% \[
% \frac{1}{N}\sum_{n=1}^{N} g\left( \fracpart{z(n)} \right) \geq (1-c)\gamma
% \]
% with $(1-c)\gamma$ a positive constant. This violates the fact that $g$ converges to zero as $N \rightarrow \infty$ and the lemma is true by contradiction.
% \end{proof}

 \begin{lemma} \label{lem:fracpartsumanddelta}
 Let $a_1, a_2, \dots, a_r$ be $r$ real numbers such  that $\left|\fracpart{a_n}\right| < \delta$ for all $n = 1,2,\dots,r$.  Then $\left|\fracpart{\sum_{n=1}^r{a_n}}\right| < r\delta.$
 \end{lemma}
 \begin{IEEEproof}
 If $\delta > \tfrac{1}{2r}$ the proof is trivial as $\left|\fracpart{\sum_{n=1}^r{a_n}}\right| \leq \tfrac{1}{2}$ for all $a_n \in \reals$.  If $\delta \leq \tfrac{1}{2r}$ then $\fracpart{\sum_{n=1}^r{a_n}} = \sum_{n=1}^r{\fracpart{a_n}}$ and
 \[
 \left\vert\fracpart{\sum_{n=1}^r{a_n}}\right\vert = \left\vert\sum_{n=1}^r{\fracpart{a_n}}\right\vert \leq \sum_{n=1}^r{\left\vert\fracpart{a_n}\right\vert} < r\delta
 \]
 by the triangle inequality.
 \end{IEEEproof}

\begin{lemma} \label{lem:fracpartinternalsumlessdelta}
Let $\left|\fracpart{a + \epsilon}\right| < \delta$ where $|\epsilon| < \nicefrac{1}{4}$ and $0<\delta<\nicefrac{1}{4}$. Then $\left|\fracpart{a}\right| < \delta + |\epsilon|$.
\end{lemma}
\begin{IEEEproof}
By supposition, for some $n \in \ints$,
\[
n - \delta < a + \epsilon < n + \delta.
\]
Since 
\[
-\delta - \epsilon > -\tfrac{1}{2} \qquad \text{and} \qquad \delta - \epsilon < \tfrac{1}{2},
\]
it follows that
\[
n - \tfrac{1}{2} < n - \delta - \epsilon < a < n + \delta - \epsilon < n + \tfrac{1}{2}.
\]
Hence $\fracpart{a} = a - n$ and so
\[
-\delta - \abs{\epsilon} \leq -\delta - \epsilon < \fracpart{a} < \delta - \epsilon \leq \delta + \abs{\epsilon}
\]
and $\abs{\fracpart{a}} \leq \delta + \abs{\epsilon}$.

% The idea behind this proof is to show that under the conditions given $\fracpart{\fracpart{a} + \epsilon}$ does not wrap, i.e. that $\fracpart{a} + \epsilon \in \left[-\nicefrac{1}{2}, \nicefrac{1}{2}\right)$.  To start, assume that $\fracpart{a} + \epsilon \geq \nicefrac{1}{2}$.  Then $\nicefrac{1}{2} \leq \fracpart{a} + \epsilon < \nicefrac{3}{4}$ and
% \[
% -\nicefrac{1}{2} \leq \fracpart{\fracpart{a} + \epsilon} = \fracpart{a + \epsilon} < -\nicefrac{1}{4}
% \] 
% and therefore $|\fracpart{a + \epsilon}| > \nicefrac{1}{4} > \delta$, a contradiction.  Similarly, assume that $\fracpart{a} + \epsilon < -\nicefrac{1}{2}$.  Then $-\nicefrac{1}{2} > \fracpart{a} + \epsilon > -\nicefrac{3}{4}$ and $\nicefrac{1}{2} > \fracpart{a + \epsilon} > \nicefrac{1}{4}$ and therefore $\left|\fracpart{a + \epsilon}\right| > \nicefrac{1}{4} > \delta$, a contradiction.  So, $\fracpart{a} + \epsilon \in \left[-\nicefrac{1}{2}, \nicefrac{1}{2}\right)$ and 
% \[
% |\fracpart{a + \epsilon}| = \left|\fracpart{a} + \epsilon\right| < \delta
% \]
% from which it follows that $\left|\fracpart{a}\right| < \delta + \left|\epsilon\right|$.
\end{IEEEproof}

We are now in a position to complete the proof of strong consistency.  Let $A$ be the subset of the sample space on which 
\[
V_N(\widehat{\lambdabf}_N) - \sigma^2 \rightarrow 0 \qquad \text{as $N\rightarrow\infty$}.
\]
From Corollary~\ref{cor:ESNconvcor}, the $\prob\{A\} =1$.  Let $A'$ be the subset of the sample space on which 
\[
N^k\widehat{\lambda}_{k,N} \rightarrow 0 \qquad \text{for $k=0,\dots,m$}
\] 
as $N\rightarrow\infty$.  As a result of Lemma~\ref{lem:moran2}, $A \subseteq A'$, and so $\prob\{A'\} \geq \prob\{A\} = 1$.  Strong consistency follows.

\section{Proof of asymptotic normality}\label{sec:centlimitproof}

Let $\psibf$ be the vector with $k$th component $\psi_k = N^k \lambda_k$, $k=0, \dots, m$ and let 
\[
T_{N}(\psibf) = S_N(\lambdabf) = \frac{1}{N} \sum_{n=1}^{N} \fracpart{ \Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \psi_k }^2.
\]
Let $\widehat{\psibf}_N$ be the vector with elements $\widehat{\psi}_{k,N} = N^k \widehat{\lambda}_{k,N}$ so that $\widehat{\psibf}_N$ is the minimiser of $T_N$.  Because each of $N^k \widehat{\lambda}_{k,N}$ converges almost surely to zero as $N \rightarrow \infty$, then $\widehat{\psibf}_{N}$ converges almost surely to $\zerobf$ as $N \rightarrow \infty$.  We want to find the asymptotic distribution of
\[
\sqrt{N}\widehat{\psibf}_{N} = 
\left[
\begin{array}
[c]{c}%
\sqrt{N} \widehat{\psi}_{0,N} \\ \sqrt{N}\widehat{\psi}_{1,N}  \\ \vdots \\ \sqrt{N} \widehat{\psi}_{m,N}
\end{array}
\right]
=
\left[
\begin{array}
[c]{c}%
\sqrt{N} \widehat{\lambda}_{0,N} \\ N\sqrt{N}\widehat{\lambda}_{1,N} \\ \vdots \\ N^m\sqrt{N} \widehat{\lambda}_{m,N}
\end{array}
\right].
\]
The proof is complicated by the fact that $T_N$ is not differentiable everywhere as $\fracpart{x}^2$ is not differentiable when $\fracpart{x} = \tfrac{1}{2}$.  This precludes the use of ``standard approaches'' to proving asymptotic normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.  However, we show in Lemma~\ref{lem:diffathatpsi} that all the partial derivatives $\frac{\partial T_N}{\partial \psi_\ell}$ for $\ell = 0, \dots, m$ exist, and are equal to zero, at the minimiser $\widehat{\psibf}_N$.  Thus, putting
\begin{equation}\label{eq:Wn}
W_{n} = \round{\Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \widehat{\psi}_{k,N}},
\end{equation}
we have, for each $\ell = 0, \dots, m$,
\begin{align*}
0 & = \frac{\partial T_N}{\partial \psi_\ell}(\widehat{\psibf}_N)\\
&= \frac{2}{N}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\left( \Phi_n - W_{n}+ \sum_{k=0}^{m}(\tfrac{n}{N})^k \widehat{\psi}_{k,N}  \right),
\end{align*}
so that
\[
D_{\ell,N} = K_{\ell,N},
\]
where
\[
D_{\ell,N} = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}(\tfrac{n}{N})^\ell \Phi_n,
\]
and
\begin{equation}\label{eq:KellN}
K_{\ell,N} = \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\left(W_{n}- \sum_{k=0}^{m}(\tfrac{n}{N})^k \widehat{\psi}_{k,N}  \right).
\end{equation}
Lemma~\ref{lem:Kconvfhalf} shows that, for all $\ell = 0, \dots m$,
\begin{equation}\label{eq:KellNconv}
K_{\ell,N} =  (h - 1) \sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big( C_{\ell k} + o_P(1) \big) + o_P(1)
\end{equation}
where $C_{\ell k} =  \tfrac{1}{\ell + k + 1}$, and $h = f(-\nicefrac{1}{2})$, and $o_P(1)$ denotes a random variable converging in probability to zero as $N\rightarrow\infty$.

It is now convenient to write in vector form.  Let 
\begin{equation}\label{eq:dnandkn}
\kbf_N = 
\left[
\begin{array}
[c]{c}%
K_{0,N}  \\ \vdots \\ K_{m,N}
\end{array}
\right]
\qquad \text{and} \qquad 
\dbf_N = \left[
\begin{array}
[c]{c}%
D_{0,N}  \\ \vdots \\ D_{m,N}
\end{array}
\right].
\end{equation}
From~\eqref{eq:KellNconv},
\[
\dbf_N = \kbf_N = \sqrt{N} (h - 1)(\Cbf + o_P(1)) \widehat{\psibf}_N + o_P(1)
\]
where $o_P(1)$ here means a vector or matrix of the appropriate dimension with every element converging in probability to zero as $N\rightarrow\infty$.  Thus $\sqrt{N}\widehat{\psibf}_N$ has the same asymptotic distribution as $(h - 1)^{-1}\Cbf^{-1} \dbf_N$.  Lemma~\ref{eq:convdn} shows that $\dbf_N$ is asymptotically normally distributed with zero mean and covariance matrix $\sigma^2\Cbf$.  Hence $\sqrt{N}\widehat{\psibf}_N$ is asymptotically normal with zero mean and covariance matrix
\[
\frac{\sigma^2\Cbf^{-1}\Cbf(\Cbf^{-1})^\prime}{(1 - h)^2} = 
\frac{\sigma^2\Cbf^{-1}}{(1 - h)^2}.
\] 
It remains to prove Lemmas~\ref{lem:diffathatpsi},~\ref{lem:Kconvfhalf} and~\ref{eq:convdn}.

\begin{lemma}\label{lem:diffathatpsi}
For all $\ell = 0, \dots, m$ the partial derivatives $\frac{\partial T_N}{\partial \psi_\ell}$ exist, and are equal to zero, at the minimiser $\widehat{\psibf}_N$.  That is,
\[
\frac{\partial T_N}{\partial \psi_\ell}(\widehat{\psibf}_N) = 0 \qquad \text{for each $\ell = 0, \dots, m$.}
\]
\end{lemma}
\begin{IEEEproof}
The function $\fracpart{x}^2$ is differentiable everywhere except if $\fracpart{x} \neq -\nicefrac{1}{2}$.  Recalling that
\[
T_{N}(\psibf) = \frac{1}{N} \sum_{n=1}^{N} \fracpart{ \Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \psi_k }^2
\]
we see that $T_N$ will be differentiable with respect to $\psibf$ at $\widehat{\psibf}_N$ if 
\[
\fracpart{\Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \widehat{\psi}_{k,N}} \neq -\nicefrac{1}{2} \qquad \text{for all $n = 1, \dots, N$.}
\] 
This is proved in Lemma~\ref{lem:boundonhatlambda}.  So the partial derivatives $\frac{\partial T_N}{\partial \psi_\ell}$ exist for all $\ell = 0, \dots, m$ at $\widehat{\psibf}_N$.  That each of the partial derivatives is equal to zero at $\widehat{\psibf}_N$ follows immediately from the fact that $\widehat{\psibf}_N$ is a minimiser of $T_N$.
\end{IEEEproof}

\begin{lemma}\label{lem:boundonhatlambda} $\sabs{\sfracpart{\Phi_n + \sum_{k=0}^m (\nicefrac{n}{N})^k \widehat{\psi}_{k,N}}} \leq \frac{1}{2} - \frac{1}{2N}$ for all $n = 1, \dots, N$.
\end{lemma}
\begin{IEEEproof}
To simplify our notation let 
\[
B_n = \Phi_n + \sum_{k=1}^m (\nicefrac{n}{N})^k \widehat{\psi}_{k,N}
\]
so that we now require to prove $\abs{\fracpart{B_n + \widehat{\psi}_{0,N}}} \leq \frac{1}{2} - \frac{1}{2N}$
for all $n = 1, \dots, N$.  From~\eqref{eq:Wn}, $W_n = \round{B_n + \widehat{\psi}_{0,N}}$, and 
\begin{align*}
T_N(\widehat{\psibf}_N) &= \frac{1}{N}\sum_{n=1}^N \fracpart{B_n + \widehat{\psi}_{0,N}}^2
\\ &= \frac{1}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N} - W_n)^2.
\end{align*}
Since $\widehat{\psi}_{0,N}$ is the minimiser of the quadratic above,
\begin{equation}\label{eq:lamsum}
\widehat{\psi}_{0,N} = -\frac{1}{N}\sum_{n=1}^N(B_n - W_n).
\end{equation}
The proof now proceeds by contradiction.  Assume that for some $k$,
\begin{equation}\label{eq:philamfraccontra}
\fracpart{B_k + \widehat{\psi}_{0,N}} > \frac{1}{2} - \frac{1}{2N}.
\end{equation}
Let $F_n = W_n$ for all $n \neq k$ and $F_k = W_k + 1$, and let
\[
\phi = -\frac{1}{N}\sum_{n=1}^N(B_n - F_n) = \widehat{\psi}_{0,N} + \frac{1}{N}.
\]
Then,
\begin{align}
(B_k + &\phi - F_k)^2 = (B_k + \phi - W_k - 1)^2 \nonumber \\
&= (B_k + \phi - W_k)^2 - 2(B_k + \phi - W_k) + 1 \nonumber \\
&= (B_k + \phi - W_k)^2 - 2(B_k + \widehat{\psi}_{0,N} - W_k) + 1 - \frac{2}{N} \nonumber \\
&= (B_k + \phi - W_k)^2 - 2\fracpart{B_k + \widehat{\psi}_{0,N}} + 1 - \frac{2}{N} \nonumber \\
&< (B_k + \phi - W_k)^2 - \frac{1}{N}, \label{eq:Bkineq}
\end{align}
where the inequality in the last line follows from~\eqref{eq:philamfraccontra}. Let
\[
\bbf = \left[
\begin{array}
[c]{cccc}%
\phi & \widehat{\psi}_{1,N} & \cdots & \widehat{\psi}_{m,N}
\end{array}
\right]^\prime
\]
be the vector of length $m+1$ with components $b_0 = \phi$ and $b_\ell = \widehat{\psi}_{\ell,N}$ for $\ell = 1 , \dots m$.  Now,
\[
N T_N(\bbf) = \sum_{n=1}^N \fracpart{B_n + \phi }^2 \leq  \sum_{n=1}^N (B_n + \phi  - F_n)^2, 
\]
and using the inequality from~\eqref{eq:Bkineq},
\begin{align*}
N &T_N(\bbf) < - \frac{1}{N} + \sum_{n=1}^N (B_n + \phi  - W_n)^2 \\
&= - \frac{1}{N} + \sum_{n=1}^N (B_n + \widehat{\psi}_{0,N} + \frac{1}{N}  - W_n)^2 \\
&= \sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n)^2 +  \frac{2}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n)\\
%&= N T_N(\widehat{\psibf}_N) + \frac{2}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n) \\
&= N T_N(\widehat{\psibf}_N),
\end{align*}
because $\frac{2}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n) = 0$ as a result of~\eqref{eq:lamsum}.  But, now $T_N(\bbf) < T_N(\widehat{\psibf}_N)$ violating the fact that $\widehat{\psibf}_N$ is a minimiser of $T_N$.  So~\eqref{eq:philamfraccontra} is false by contradiction.

If $\fracpart{B_k + \widehat{\psi}_{0,N}} < -\frac{1}{2} + \frac{1}{2N}$ for some $k$, we set $F_k = W_k - 1$ and using the same procedure as before obtain $T_N(\bbf) < T_N(\widehat{\psibf}_N)$ again.  The proof follows.
\end{IEEEproof}

\begin{lemma}\label{lem:Kconvfhalf}
With $K_{\ell,N}$ defined as in~\eqref{eq:KellN} and $h = f(-\nicefrac{1}{2})$,
\[
K_{\ell,N} = (h - 1) \sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big( C_{\ell k} + o_P(1) \big) + o_P(1) 
\] 
for all $\ell = 0, \dots, m$, where $C_{\ell k} =  \frac{1}{\ell + k + 1}$.% is an element of the $m+1$ by $m+1$ Hilbert matrix, and $o_P(1)$ denotes a random variable converging in probability to zero as $N\rightarrow\infty$.
\end{lemma}
\begin{IEEEproof}
Care must be taken since $\widehat{\psibf}_N$ depends on the sequence $\{ \Phi_n \}$.  For $n = 1, \dots, N$ and positive $N$, let
\begin{equation}
p_{nN}(\psibf) = \sum_{k=0}^{m}\left( \tfrac{n}{N}\right)^k \psi_k,
\end{equation}
and put
\[
q_{n}(x) = \round{\Phi_n + x}, \qquad Q(x) = \expect q_{n}(x) =  \expect q_{1}(x).
\]
Let
\begin{equation}\label{eq:GNdef}
G_N(\psibf) = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}\left( \tfrac{n}{N}\right)^\ell  \big(q_n(p_{nN}(\psibf)) - Q(p_{nN}(\psibf)) \big).
\end{equation}
and put
\begin{equation}\label{eq:hatpnN}
\widehat{p}_{nN} = p_{nN}(\widehat{\psibf}_N) = \sum_{k=0}^m \left(\tfrac{n}{N}\right)^k \widehat{\psi}_{k,N}.
\end{equation}
Observe that $G_N$ depends on $\ell$ and we could write $G_{\ell,N}$ but have suppressed the subscript $\ell$ for notational simplicity.  Now $W_n$ from~\eqref{eq:Wn} can be written as
\[
W_n = \round{\Phi_n + \widehat{p}_{nN}} = q_{n}(\widehat{p}_{nN})
\]
and $K_{\ell,N}$ from~\eqref{eq:KellN} can be written as
\begin{align*}
K_{\ell,N} &= \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\big( q_{n}(\widehat{p}_{nN}) - \widehat{p}_{nN}  \big) \\
&= \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\big(q_{n}(\widehat{p}_{nN}) - \widehat{p}_{nN} + Q(\widehat{p}_{nN}) - Q(\widehat{p}_{nN}) \big) \\
&= G_N(\widehat{\psibf}_N) + H_{\ell,N},
\end{align*}
where 
\begin{equation}\label{eq:HellNdef}
H_{\ell,N} = \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell \big( Q(\widehat{p}_{nN}) - \widehat{p}_{nN} \big).
\end{equation}
Lemma~\ref{lem:unifprobG} in the Appendix shows that for any $\delta >0$ and $\nu > 0$ there exists an $\epsilon > 0$ such that
\[
\prob\left\{ \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf) } > \delta   \right\} < \nu
\]
for all positive integers $N$, where $\|\psibf\|_\infty = \sup_{k} \abs{\psi_k}$.  Since $\widehat{\psibf}_N$ converges almost surely to zero, it follows that for any $\epsilon > 0$,
\[
\lim_{N\rightarrow\infty}\prob\left\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} = 0
\] 
and therefore $\prob\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \} < \nu$ for all sufficiently large $N$.  Now
\begin{align*}
  \prob&\left\{\abs{ G_N(\widehat{\psibf}_N) } > \delta \right\} \\
&= \prob\left\{ \abs{G_N(\widehat{\psibf}_N)} > \delta \;, \; \|\widehat{\psibf}_N\|_{\infty} < \epsilon \right\} \\
&\hspace{0.7cm} + \prob\left\{ \abs{G_N(\widehat{\psibf}_N)} > \delta  \;, \; \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} \\
&\leq \prob\left\{  \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf) } > \delta \right\} + \prob\left\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} \\
&\leq 2\nu
\end{align*}
for all sufficiently large $N$.  Since $\nu$ and $\delta$ can be chosen arbitrarily small, it follows that $G_N(\widehat{\psibf}_N)$ converges in probability to zero as $N\rightarrow\infty$, and therefore $K_{\ell,N} = H_{\ell,N} + o_P(1)$.  Lemma~\ref{lem:sumpNhilb} shows that
\begin{align*}
H_{\ell,N} =  (h-1)\sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big(C_{\ell k} + o_P(1)\big).
\end{align*}
The proof follows.
\end{IEEEproof}

\begin{lemma}\label{lem:sumpNhilb}
With $H_{\ell,N}$ defined in~\eqref{eq:HellNdef} and $\widehat{p}_{nN}$ defined in~\eqref{eq:hatpnN}, and with $h = f(-\nicefrac{1}{2})$,
\[
H_{\ell,N} = (h-1)\sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big(C_{\ell k} + o_P(1)\big),
\]
where $C_{\ell k} = \tfrac{1}{\ell + k + 1}$.
\end{lemma}
\begin{IEEEproof}
If $\abs{x} < 1$, then
\[
q_n(x)  = \round{\Phi_n + x} = \begin{cases}
1, & \Phi_n + x  \geq \nicefrac{1}{2} \\
-1, & \Phi_n + x  < -\nicefrac{1}{2} \\
0, & \text{otherwise},
\end{cases}
\]
and,
\begin{align*}
Q(x) = Eq_1(x) &=  \begin{cases}
\int_{\nicefrac{1}{2} -x }^{\nicefrac{1}{2}}f(t)\,dt, &   x \geq 0 \\
-\int_{-\nicefrac{1}{2}}^{-\nicefrac{1}{2} - x }f(t)\,dt, &  x  < 0.
\end{cases} 
% \\
% &= \begin{cases}
% 1 - F(\nicefrac{1}{2} - x ), & x  \geq 0 \\
% -F(-\nicefrac{1}{2} - x ), &  x < 0,
% \end{cases} 
\end{align*}
Because $f(\fracpart{x})$ is continuous at $-\nicefrac{1}{2}$ it follows from the mean value theorem that
\[
Q(x) = x \big( h + \zeta(x) \big),
\] 
where $\zeta(x)$ is a function that converges to zero as $x$ converges to zero.  
%The equation above is still valid when $\abs{x} \geq 1$ by an appropriate choice of the function $\zeta(x)$.  
Observe that $\abs{\widehat{p}_{nN}} \leq \sum_{k=0}^m\sabs{ \widehat{\psi}_{k,N}}$ and, since each of the $\widehat{\psi}_{k,N} \rightarrow 0$ almost surely as $N\rightarrow\infty$, it follows that $\widehat{p}_{nN} \rightarrow 0$ almost surely uniformly in $n = 1, \dots, N$ as $N\rightarrow\infty$.  Thus, $\zeta(\widehat{p}_{nN}) \rightarrow 0$ almost surely (and therefore also in probability) uniformly in $n = 1, \dots, N$ as $N\rightarrow\infty$.  Now,
\begin{align*}
Q(\widehat{p}_{nN}) - \widehat{p}_{nN} &= \widehat{p}_{nN}\big( h - 1 + \zeta(\widehat{p}_{nN}) \big) \\
&= \widehat{p}_{nN}\big( h - 1 + o_P(1) \big), 
\end{align*}
and so,
\begin{align*}
H_{\ell,N} &= \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell \widehat{p}_{nN}\big( h - 1 + o_P(1) \big)  \\
&=  \frac{1}{\sqrt{N}} \sum_{n=1}^{N}(\tfrac{n}{N})^\ell \sum_{k=0}^m (\tfrac{n}{N})^k \widehat{\psi}_{k,N} \big( h - 1 + o_P(1) \big)   \\
&= \sqrt{N}\sum_{k=0}^m \widehat{\psi}_{k,N} \frac{1}{N} \sum_{n=1}^{N} \frac{n^{\ell + k}}{N^{\ell + k + 1}}\big( h - 1 + o_P(1)  \big).
\end{align*}
The Riemann sum
\[
\frac{1}{N} \sum_{n=1}^{N} \frac{n^{\ell + k}}{N^{\ell + k + 1}} = \int_{0}^{1}x^{k+\ell+1}dx + o_P(1),
\]
and since the integral above evaluates to $\frac{1}{k+\ell+1}$, we have
\[
H_{\ell,N} = (h - 1)\sqrt{N}\sum_{k=0}^m \widehat{\psi}_{k,N}\left( \frac{1}{k+\ell+1} + o_P(1)  \right)
\]
as required.
% Put
% \[
% C_{\ell + k,N} = \sum_{n=1}^{N}\frac{n^{\ell + k}}{N^{\ell + k+1}}.
% \]
% It is well known (see for example~\cite[eq. 21]{Peleg1991_CRB_PPS_1991}) that,
% \begin{equation}\label{eq:ClkNHilbert}
% C_{\ell + k,N} = C_{\ell k} + O(N^{-1}).
% \end{equation}
% Since $\zeta(\widehat{p}_{nN}) = o_P(1)$ for all $n = 1, \dots, N$ then
% \[
% \sum_{n=1}^{N}\frac{n^{\ell + k}}{N^{\ell + k + 1}}\zeta(\widehat{p}_{nN}) 
% \]
% converges in probability to zero as $N\rightarrow\infty$.  So,
% \begin{align*}
% \sum_{n=1}^{N}\frac{n^{\ell + k}}{N^{\ell + k + 1}}\big( h - 1 +  \zeta(\widehat{p}_{nN}) \big) = (h-1)\big( C_{\ell k} + o_P(1) \big).
% \end{align*}
% The proof follows by substituting this into~\eqref{eq:HellNre}.
\end{IEEEproof}

\begin{lemma}\label{eq:convdn}
The distribution of the vector $\dbf_N$, defined in~\eqref{eq:dnandkn}, converges to the multivariate normal with zero mean and covariance matrix $\sigma^2\Cbf$.
\end{lemma}
\begin{IEEEproof}
For any constant vector $\alphabf$, let
\[
z_N = \alphabf^\prime \dbf_N = \frac{1}{\sqrt{N}} \sum_{n=1}^{N} \Phi_n \sum_{\ell=0}^{m}\alpha_\ell\left( \frac{n}{N} \right)^\ell.
\]
By Lypanov's central limit theorem $\zbf_N$ is asymptotically normally distributed with zero mean and variance
\[
\lim_{N\rightarrow\infty} \sigma^2 \frac{1}{N} \sum_{n=1}^{N}\left( \sum_{\ell=0}^{m}\alpha_\ell\left( \frac{n}{N} \right)^\ell \right)^2 = \sigma^2 \alphabf^\prime \Cbf \alphabf.
\]
By the Cramer-Wold theorem it follows that $\dbf_N$ is asymptotically normally distributed with zero mean and covariance $\sigma^2 \Cbf$. 
% Recall that the $\ell$th element of $\dbf_N$ is
% \[
% D_{\ell,N} = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}(\tfrac{n}{N})^\ell \Phi_n.
% \]
% Since the $\Phi_1, \dots, \Phi_N$ are independent, each with variance $\sigma^2$, the distribution of $D_{\ell,N}$ converges to the normal with zero mean and variance
% \[
% \lim_{N\rightarrow\infty}\var D_{\ell,N} = \sigma^2 \lim_{N\rightarrow\infty} C_{2\ell,N} = \sigma^2 C_{\ell\ell}
% \]
% by Lypanov's central limit theorem and by using~\eqref{eq:ClkNHilbert}.  The covariance between $D_{\ell,N}$ and $D_{k,N}$ is
% \[
% \covar(D_{\ell,N}, D_{k,N}) = \sigma^2 C_{\ell+k,N} \rightarrow \sigma^2 C_{\ell k}
% \]
% as $N \rightarrow \infty$, again using~\eqref{eq:ClkNHilbert}.  The proof follows.
\end{IEEEproof}

\begin{figure*}[tb] 
   	\centering 
  		\includegraphics{code/lseplot-1.mps} 
   		\caption{Performance of the LSE and LSU estimator for a polynomial phase signal of order two, i.e. $m=2$.  (Left) Mean square error (MSE) in the second order parameter $\mu_2$ for $N=10,50$ and $200$ versus the variance $\sigma_c^2$ of the $X_1, \dots, X_N$. The $X_1, \dots, X_N$ are independent and identically distributed (i.i.d.) complex Gaussian random variables.  (Right) MSE in second order parameter $\mu_2$ versus the variance $\sigma^2$ of the $\Phi_1, \dots, \Phi_N$. The $\Phi_1, \dots, \Phi_N$ are i.i.d. uniform random variables.} 
   		\label{plot:lsetest} 
  \end{figure*}

\section{Simulations}\label{sec:simulations}

This section describes the results of Monte-Carlo simulations involving the least squares unwrapping estimator (LSU), the least squares estimator (LSE), Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, and the discrete polynomial phase transform (DPT)~\cite{Peleg_DPT_1995}.  In all simulations the sample sizes considered are $N = 10, 50$ and $200$ and the unknown amplitude is $\rho = 1$.  The number of replications of each experiment is 2000 and the $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ are distributed uniformly randomly over the identifiable region.

Figure~\ref{plot:lsetest} shows the performance of the LSU estimator and LSE when $m=2$ (linear chirp). Only results for the highest order parameter $\widehat{\mu}_2$ are displayed, the results for the lower order parameters leading to similar conclusions.  The LSE is computed using the approach described in~\cite[Sec.~10.1]{McKilliam2010thesis} which is based on that of Abatzoglou~\cite{Abatzoglou_ml_chirp_1986}. The LSE is computationally intensive.  We were able to compute it within a reasonable amount of time only with $N$ less than $50$ when $m=2$. Results for the exact LSU estimator computed using the sphere decoder and the approximate LSU estimator computed using Babai's algorithm are displayed. The sphere decoder is computationally intractable when $N = 200$. It is evident that a significant performance penalty is suffered by using Babai's algorithm when the noise variance is large.  We have also displayed results using another approximate nearest point algorithm called the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder}. Our implementation requires $O(N^3\log{N})$ operations. This is more computationally expensive than Babai's algorithm but, as can be seen, the $K$-best method is more statistically accurate.

For the simulations displayed in the left hand plot of Figure~\ref{plot:lsetest} the signal model used is that from~\eqref{eq:Y_nsamplednoisey}.  The noise terms $X_1, \dots, X_N$ are independent and identically distributed zero mean complex Gaussian random variables with independent real and imaginary parts having variance $\sigma_c^2$.  In this case the LSE is also the maximum likelihood estimator.  The \term{Cram\`{e}r-Rao lower bound} (CRB) for the variance of unbiased estimators of $\tilde{\mu}_j$ is given by Peleg and Porat~\cite{Peleg1991_CRB_PPS_1991} to be
\[ 
\frac{\sigma_c^2}{4\pi^2 N^{2j + 1}} C^{-1}_{jj} + o(N^{-2j-1})
\]
where $C^{-1}_{jj}$ is the $j$th diagonal element of the inverse of the $m+1$ by $m+1$ Hilbert matrix $\Cbf$.  Observe that the performance of the exact LSU estimator (computed using the sphere decoder) is close to the maximum likelihood LSE.  The asymptotic Cramer-Rao bound (CRB) (solid line) and the asymptotic variance predicted in Theorem~\ref{thm:asymp_proof} (dashed line) are also shown. Provided the noise variance is small enough (so that the `threshold' is avoided) the LSE variance is close to the CRB (as is expected) and the LSU variance is close to that predicted by Theorem~\ref{thm:asymp_proof}. 
%Note the gap between the CRB and the variance predicted by Theorem~\ref{thm:asymp_proof}.  In practice this gap can be overcome using a numerical technique, such as Newton's method, with the LSU estimate as a starting point.  We have not done this here in order to show the accuracy of Theorem~\ref{thm:asymp_proof}.  The other noticeable gap is that between the threshold behaviour of the LSU estimator and the LSE.  When $N=50$ the threshold \emph{kicks in} a little earlier for the LSE than it does for the LSU. We suspect that this gap will actually \emph{close} as $N$ increases. This hypothesis is supported by the fact that this behaviour is observed in the case of frequency estimation (when $m=1$)~\cite[Sec. 9.5]{McKilliam2010thesis}\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}. Unfortunately it is not feasible to test this hypothesis because of the computational complexity of the LSE. 
For the simulations displayed in the right hand plot of Figure~\ref{plot:lsetest} the signal model used is that from~\eqref{eq:noise_circ_poly}.  The phase noise terms $\Phi_1, \dots, \Phi_N$ are uniform random variables with variance $\sigma^2$.  In this case, the LSU estimator performs slightly better than the LSE.  This result agrees with similar observations made in circular statistics~\cite{McKilliam_mean_dir_est_sq_arc_length2010}.  The asymptotic variance predicted by Theorem~\ref{thm:asymp_proof} again accurately predicts the performance of the LSU estimator.

Figure~\ref{plot:polyest} displays the performance of the LSU estimator, the DPT and Kitchen's estimator when $m=3$. Both the DPT and Kitchen's estimator perform poorly because they do not work correctly for some parameters in the identifiable region $B$. This behaviour has been observed previously~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase,McKilliam_polyphase_est_icassp_2011}. Again the performance of the LSU is well predicted by Theorem~\ref{thm:asymp_proof} provided that the noise variance is small enough to avoid the threshold effect and provided that $N$ is sufficiently large.  The asymptotic CRB is also plotted.

\begin{figure*}[p] 
   	\centering 
  		\includegraphics{code/gaussianplot4-1.mps} 
  		\caption{Mean square error (MSE) of polynomial phase estimators for $N=10,50$ and $200$ versus the variance $\sigma_c^2$ of the complex noise terms $X_1, \dots X_N$. The polynomial phase signal has order 3, and the $X_1, \dots, X_N$ are independent and identically distributed complex Gaussian random variables. (Top left) MSE in the phase coefficient $\mu_0$.  (Top right) MSE in the frequency coefficient $\mu_1$.  (Bottom left) MSE in the quadratic coefficient $\mu_2$.  (Bottom right) MSE in the cubic coefficient $\mu_3$.} 
  		\label{plot:polyest} 
 \end{figure*}

\section{Computational considerations}

When $m=0$, polynomial phase estimation is equivalent to the problem of \emph{direction estimation} from circular statistics, and the least squares estimator corresponds with the \emph{extrinsic mean} (or \emph{sample circular mean})~\cite{McKilliam_mean_dir_est_sq_arc_length2010,Fisher1993,Mardia_directional_statistics}.  In this case the LSE can be computed efficiently in $O(N)$ by averaging $N$ complex numbers.  When $m=1$, polynomial phase estimation is equivalent to single frequency estimation, and the least squares estimator can be computed efficiently in $O(N\log N)$ operations by maximising the \emph{periodogram} using the fast Fourier transform followed by an appropriate numerical optimisation technique~\cite{Quinn2008maximizing_the_periodogram,Rife1974,Quinn2001}.  When $m=2$, the least squares estimator can be computed in a reasonable amount of time when $N$ is approximately less than $50$.  The LSE is computationally intensive for any $N$ when $m > 2$. By comparison the exact LSU estimator implemented using the sphere decoder can be computed quickly for any $m$ with $N$ approximately less than $50$.  In situations where $N$ is small, but high statistical accuracy is required, the sphere decoder is computationally a better choice than the LSE for polynomial phase signals of order greater than or equal to $2$.

Babai's nearest plane algorithm requires $O(N^2)$ operations and can be run in a reasonable amount of time for any $m$ and reasonably large $N$. It should be noted that Babai's nearest plane algorithm requires a Lov\`as reduced lattice basis~\cite{Lenstra1982}.  This requires $O(N^4)$ operations to compute, but, the Lov\`as reduced basis can be computed once offline. It does not need to be computed each time the estimator is run. The $K$-best algorithm runs in a reasonable amount of time for any $m$ and when $N$ is less than about $300$. If the noise variance is small, then both the sphere decoder and the $K$-best algorithm are actually fast.  This is because both of these estimators begin with the lattice point found by Babai's nearest plane algorithm.  If this point is close to (or \emph{is}) the nearest point then both the sphere decoder and the $K$-best algorithm terminate quickly.  This is in contrast to the least squares estimator that is slow regardless of the noise variance.

Both Kitchen's estimator and the DPT are fast to compute.  Kitchen's estimator requires only $O(N)$ arithmetic operations and the DPT requires $O(N \log N)$ operations.  However, these estimators are not as accurate as the LSE or the LSU.  They also fail to operate correctly over the entire identifiable region of polynomial phase coefficients~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}~\cite[Ch.~10]{McKilliam2010thesis}.

%\section{Discussion}

\section{Conclusion} \label{sec:conclusion}

This paper has considered the estimation of the coefficients of a noisy polynomial phase signal by least squares phase unwrapping (LSU). It has been shown that the LSU estimator is strongly consistent and asymptotically normally distributed. %Simulations are used to show that the LSU has near maximum likelihood performance in the case that the noise is additive white and Gaussian.  
The LSU estimator can be computed by finding a nearest lattice point in a lattice~\cite{McKilliam2009asilomar_polyest_lattice}.  %Here, we consider using the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder} to achieve this in reasonable time when the number of observations, $N$, is less than about 300.  
Polynomial time nearest point algorithms for these lattices exist~\cite[Sec 4.3]{McKilliam2010thesis}, but these algorithms are not fast in practice.  Instead, we considered some general purpose algorithms, the \term{sphere decoder}, \term{Babai's nearest plane algorithm} and the \term{$K$-best algorithm}. The sphere decoder and $K$-best algorithm result in accurate estimators.

The major outstanding question is whether faster nearest point algorithms exist for these specific lattices.  Considering the excellent statistical performance (both theoretically and practically) of the LSU estimator, even fast \emph{approximate} nearest point algorithms are likely to prove useful for the estimation of polynomial phase signals.

%\begin{figure}[p] 
%   	\centering 
%  		\includegraphics[width=\linewidth]{polyfig/polyphase4plot.4} 
%  		\caption{MSE in the third order coefficient $\mu_3$ versus $\var{X_n} = \sigma^2$. The true coefficients are uniformly spread in the identifiable region.} 
%  		\label{plot:polyest_projnorm_mu3} 
% \end{figure}

% \begin{figure}[p] 
%  	\centering 
% 		\includegraphics[width=\linewidth]{polyfig/polyphase4plot_unfair.4} 
% 		\caption{MSE in the third order coefficient $\mu_3$ versus $\var{X_n} = \sigma_c^2$. The coefficient have been restricted for Kitchen's estimator and the DPT.} 
% 		\label{plot:polyest_projnorm_mu3_unfair} 
%\end{figure} 
% 
% \begin{figure}[p] 
%  	\centering 
% 		\includegraphics[width=\linewidth]{polyfig/polyphase4plot_fair.4} 
% 		\caption{MSE in the third order coefficient $\mu_3$ versus $\var{X_n} = \sigma_c^2$. The DPT estimator runs at the higher sampling rate $\delta$ so that the volumes $V_{DPT}(\delta) = V_m$.} 
% 		\label{plot:polyest_projnorm_mu3_fair} 
%\end{figure}

%\addtolength{\parskip}{-0.2cm} 
%\bibliographystyle{../../bib/IEEEbib} 
\small 
\bibliography{bib}

\normalsize
\appendix

%\section{Tricks with fractional parts}
%Throughout the paper, and particularly in  have made use of a number of results involving the fractional part function

\subsection{An empirical process result}

During the proof of asymptotic normality in Lemma~\ref{lem:Kconvfhalf} we made use of the following result regarding the function,
\[
G_N(\psibf) = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}\left( \tfrac{n}{N}\right)^\ell  \big(q_n(p_{nN}(\psibf)) - Q(p_{nN}(\psibf)) \big),
\]
where the functions $q_n$, $Q$ and $p_{nN}$ are defined above~\eqref{eq:GNdef} and $\ell \in \{0, 1, \dots, m\}$.

\begin{lemma}\label{lem:unifprobG}
For any $\delta >0$ and $\nu > 0$ there exists $\epsilon > 0$ such that
\[
\prob\left\{ \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf) } > \delta   \right\} < \nu %\qquad \text{for all $N > N_0$}
\]
for all positive integers $N$.
\end{lemma}

This result is related to what is called \emph{tightness} in the literature on empirical processes and weak convergence on metric spaces~\cite{Billingsley1999_convergence_of_probability_measures,Dudley_unif_central_lim_th_1999,Shorak_emp_proc_stat_2009}.  The lemma is different from what is usually proved in the literature because the function 
\[
p_{nN}(\psibf) = \sum_{k=0}^m\left( \tfrac{n}{N}\right)^k\psi_k
\]
depends on $n$.  Nevertheless, we find that the methods of proof from the literature can still be used. The proof is based on a technique called \emph{symmetrisation} and another technique called \emph{chaining} (also known as \emph{bracketing})~\cite{Pollard_asymp_empi_proc_1989}.  During the proof we treat $\ell$ as a fixed constant.  The proof holds for any nonegative integer $\ell$.

\begin{proof}
Define the function 
\begin{align*}
f_{nN}(\psibf, \Phi_n) &= \left( \tfrac{n}{N}\right)^\ell q_n(p_{nN}(\psibf)) \\
%&= \left( \tfrac{n}{N}\right)^\ell \round{\Phi_n + p_{nN}(\psibf)} \\
&= \left( \tfrac{n}{N}\right)^\ell \round{\Phi_n + \sum_{k=0}^m\left( \tfrac{n}{N}\right)^k\psi_k}
\end{align*}
%and observe that 
%\[
%\expect f_{nN}(\psibf, \Phi_n) = \expect f_{nN}(\psibf, \Phi_1) = \left( \tfrac{n}{N}\right)^\ell Q(p_{nN}(\psibf))
%\]
so that $G_N$ can be written as
\[
G_N(\psibf) = \frac{1}{\sqrt{N}} \sum_{n=1}^{N} \big( f_{nN}(\psibf, \Phi_n) - \expect f_{nN}(\psibf, \Phi_n) \big).
\]
Let $\{g_n\}$ be a sequence of independent standard normal random variables, independent of the phase noise sequence $\{\Phi_n\}$.  Lemma~\ref{lem:symmetrisation} shows that
\[
\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} \leq \sqrt{2\pi} \; \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ Z_N(\psibf) },
\]
where 
\begin{equation}\label{eq:ZpsiCondGaussProc}
Z_N(\psibf) = \frac{1}{\sqrt{N}} \sum_{n=1}^{N} g_n f_{nN}(\psibf, \Phi_n),
\end{equation}
and where $\expect$ runs over both $\{g_n\}$ and $\{\Phi_n\}$.  Conditionally on $\{\Phi_n\}$, $\{Z_N(\psibf), \psibf \in \reals^{m+1} \}$ is a \emph{Gaussian process}, and numerous techniques exist for its analysis.  Lemma~\ref{lem:chaining} shows that for any $\kappa > 0$ there exists an $\epsilon > 0$ such that
\[
\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ Z_N(\psibf) } < \kappa.
\]
It follows that,
\[
\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)}  <  \sqrt{2\pi} \; \kappa,
\]
and by Markov's inequality,
\[
\prob \cubr{  \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} > \delta } \leq  \sqrt{2\pi} \; \frac{\kappa}{\delta},
\]
for any $\delta > 0$.  The proof follows with $\nu =  \sqrt{2\pi} \kappa/\delta$.  It remains to prove Lemmas~\ref{lem:symmetrisation}~and~\ref{lem:chaining}.

\end{proof}

The proof of the next lemma is based on a technique called \emph{symmetrisation}~\cite{Gine_Zinn_symmetrisation_1984} and the proof we give closely follows that of Pollard~\cite[Section 4]{Pollard_asymp_empi_proc_1989}.

\begin{lemma} \label{lem:symmetrisation}(Symmetrisation)
\[
\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} \leq \sqrt{2\pi} \; \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ Z_N(\psibf) }.
\]
\end{lemma}
\begin{IEEEproof}
Let $\{\Phi_n^\prime\}$ be a sequence of random variables distributed identically to $\{\Phi_n\}$ and independent of both $\{\Phi_n\}$ and $\{g_n\}$.  Write $\expect_\Phi$ to denote expectation conditional on $\{\Phi_n\}$.  %That is, expectation taken over $\{g_n\}$ and $\{\Phi_n^\prime\}$ treating $\{\Phi_n\}$ as fixed.  
Since $\Phi_n$ and $\Phi_n^\prime$ are identically distributed,
\[
\expect f_{nN}(\psibf, \Phi_n) = \expect_\Phi f_{nN}(\psibf, \Phi_n^\prime),
\]
so that,
\begin{align*}
G_N(\psibf) &= \frac{1}{\sqrt{N}} \sum_{n=1}^{N} \big( f_{nN}(\psibf, \Phi_n) - \expect_\Phi f_{nN}(\psibf, \Phi_n^\prime) \big)  \\
&= \expect_\Phi \frac{1}{\sqrt{N}} \sum_{n=1}^{N} ( f_{nN} - f_{nN}^\prime ),
\end{align*}
where, for notational convenience, we put 
\[
f_{nN} = f_{nN}(\psibf, \Phi_n) \qquad \text{and} \qquad  f_{nN}^\prime = f_{nN}(\psibf, \Phi_n^\prime).
\] 
Taking absolute values followed by supremums,
\begin{align*}
 \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} &= \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{\frac{1}{\sqrt{N}} \expect_\Phi \sum_{n=1}^{N} (f_{nN} - f_{nN}^\prime )} \\
&\leq \sup_{\|\psibf\|_{\infty} < \epsilon} \expect_\Phi  \abs{ \frac{1}{\sqrt{N}} \sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) },
\end{align*}
the upper bound following from Jensen's inequality.  Since $\sup \expect \abs{\dots} \leq \expect \sup \abs{\dots}$, %the bound can be increased by moving $\expect_{\Phi}$ outside the supremum,
\begin{equation}\label{eq:Eoutsidesup}
 \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} \leq \expect_\Phi \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} ( f_{nN} - f_{nN}^\prime ) }.
\end{equation}
Applying $\expect$ to both sides gives the inequality
\begin{equation}\label{eq:Gninequfin}
 \expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} \leq  \expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) },
\end{equation}
since $\expect \expect_\Phi$ is equivalent to $\expect$.   Let
\[
\sigma_n = \frac{g_n}{\abs{g_n}},
\]
and put $\sigma_n = 1$ if $g_n = 0$.  The $\sigma_n$ are thus independent with 
\[
\prob\{ \sigma_n = -1 \} = \tfrac{1}{2} \qquad \text{and} \qquad  \prob\{ \sigma_n = 1 \} = \tfrac{1}{2}.
\]  
As $\Phi_n$ and $\Phi_n^\prime$ are independent and identically distributed, the random variable $f_{nN} - f_{nN}^\prime$ is symmetrically distributed about zero, and is therefore distributed identically to $\sigma_n( f_{nN} - f_{nN}^\prime)$.  Thus,
%\[
%\sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) \qquad \text{and} \qquad \sum_{n=1}^{N}  \sigma_n(f_{nN} - f_{nN}^\prime)
%\]
%are identically distributed, and therefore
\begin{align*}
 \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  &\abs{  \frac{1}{\sqrt{N}}\sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) } \\
&= \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{  \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n ( f_{nN} - f_{nN}^\prime) }.
\end{align*}
By the triangle inequality
\[
\abs{ \sum_{n=1}^{N} \sigma_n (f_{nN} - f_{nN}^\prime)} \leq \abs{ \sum_{n=1}^{N} \sigma_n f_{nN} } + \abs{ \sum_{n=1}^{N}  \sigma_n f_{nN}^\prime },
\]
and it follows from~\eqref{eq:Gninequfin}, that
\begin{align*}
 \expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} &\leq \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} } \\
&\hspace{1cm} + \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{\frac{1}{\sqrt{N}} \sum_{n=1}^{N}  \sigma_n  f_{nN}^\prime } \\
&= 2 \; \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} }.
\end{align*}
Since $g_n$ is a standard normal random variable, $\expect \abs{g_n} = \expect_\Phi \abs{g_n} = \sqrt{2/\pi}$,
and
\begin{align*}
\expect \sup_{\|\psibf\|_{\infty} < \epsilon}  &\abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} } \\
&= \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} \sqrt{\frac{\pi}{2}}\expect_\Phi \abs{g_n}} \\
&\leq \sqrt{\frac{\pi}{2}} \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{ \sqrt{N}}\sum_{n=1}^{N} \sigma_n \abs{g_n}  f_{nN}},
\end{align*}
the last line following from Jensen's inequality and by moving $\expect_{\Phi}$ outside the supremum similarly to~\eqref{eq:Eoutsidesup}. As $g_n = \sigma_n \abs{g_n}$, it follows that
\begin{align*}
\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{G_N(\psibf)}  &\leq 2\sqrt{\frac{\pi}{2}} \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} g_n  f_{nN}} \\
&= \sqrt{2\pi} \; \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ Z_N(\psibf) } 
\end{align*}
as required.
\end{IEEEproof}

\begin{lemma} \label{lem:chaining}
For any $\kappa > 0$ there exists $\epsilon > 0$ such that
\[
\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ Z_N(\psibf) } < \kappa,
\]
where $Z_N(\psibf)$ is defined by~\eqref{eq:ZpsiCondGaussProc}.
\end{lemma}
\begin{IEEEproof}
Without loss of generality, assume that $\epsilon < \frac{1}{m+1}$.  Lemma~\ref{lem:chaining2} shows that
\begin{equation}\label{eq:supZCK1} 
\expect_\Phi \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ Z_N(\psibf) } \leq K_1 \sqrt{C_\epsilon(\{\Phi_n\})}
\end{equation}
where $K_1$ is a finite, positive constant, and $C_\epsilon(\{\Phi_n\})$ is the average number of times $\abs{\Phi_1}, \dots, \abs{\Phi_N}$ is greater than or equal to $\nicefrac{1}{2} - (m+1)\epsilon$.  That is,
\begin{equation}\label{eq:Cedefn}
C_\epsilon(\{\Phi_n\}) = \frac{1}{N} \sum_{n=1}^{N} I_\epsilon(\abs{\Phi_n})
\end{equation}
where $I_\epsilon(\abs{\Phi_n})$ is 1 when $\abs{\Phi_n} \geq \nicefrac{1}{2} - (m+1)\epsilon$ and zero otherwise.  Recall that $f$ is the probability density function of $\Phi_n$, and (by assumption in Theorem~\ref{thm:asymp_proof}) that $f(\fracpart{x})$ is continuous at $x = -\nicefrac{1}{2}$.  Because of this, the expected value of $C_\epsilon(\{\Phi_n\})$ is small when $\epsilon$ is small, since
\begin{align*}
\expect C_\epsilon(\{\Phi\}) &= \frac{1}{N} \sum_{n=1}^{N} \expect I_\epsilon(\abs{\Phi_n}) \\
&= \prob\cubr{\abs{\Phi_1} \geq \nicefrac{1}{2} - (m+1)\epsilon} \\
&= \int_{-1/2}^{-1/2 + (m+1)\epsilon} f(\phi) d\phi + \int_{1/2 - (m+1)\epsilon}^{1/2} f(\phi) d\phi \\
&= \int_{-1/2 + (m+1)\epsilon}^{1/2 - (m+1)\epsilon} f(\fracpart{\phi}) d\phi \\
&= 2(m+1)\epsilon(f(-\nicefrac{1}{2}) + o(1))
\end{align*}
where $o(1)$ goes to zero as $\epsilon$ goes to zero, by the mean value theorem.  Since $\sqrt{\cdot}$ is a concave function on the positive real line, and since $C_{\epsilon}(\{\Phi_n\})$ is always positive,
\[
\expect \sqrt{C_\epsilon(\{\Phi_n\})} \leq  \sqrt{\expect  C_\epsilon(\{\Phi_n\} )} < \sqrt{K_2 \epsilon }
\]
by Jensen's inequality, for some constant $K_2$.  Applying $\expect$ to both sides of~\eqref{eq:supZCK1},
\[
\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ Z_N(\psibf) } \leq K_1 \sqrt{\expect C_\epsilon(\{\Phi_n\})} < K_1 \sqrt{K_2 \epsilon}.
\]
Choosing $\epsilon = \kappa^2/(K_1^2 K_2)$ completes the proof.  It remains to prove Lemma~\ref{lem:chaining2}.
\end{IEEEproof}

The proofs of Lemmas~\ref{lem:chaining2}~and~\ref{lem:chaining3} are based on a technique called \emph{chaining} (or \emph{bracketing})~\cite{Dudley_unif_central_lim_th_1999,Ossiander_clt_bracketing_1984,Pollard_asymp_empi_proc_1989,Pollard_new_ways_clts_1986}.  The proofs here follow those of Pollard~\cite[Section 3]{Pollard_asymp_empi_proc_1989}.  Within the remaining proofs, we are interested in expectations conditional on the phase noise sequence $\{\Phi_n\}$, so for the purpose of these lemmas, $\{\Phi_n\}$ is treated as a fixed realisation.  For notational convenience we use the abbreviations 
\[
C_\epsilon = C_\epsilon(\{\Phi_n\}) \qquad \text{and} \qquad f_{nN}(\psibf) = f_{nN}(\psibf, \Phi_n)
\]
in what follows.  As in Lemma~\ref{lem:chaining} we assume, without loss of generality, that $\epsilon < \tfrac{1}{m+1}$.  %This comes at no loss since Lemma~\ref{lem:chaining} only asks for the existence of an $\epsilon > 0$.

\begin{lemma}\label{lem:chaining2}
There exists a positive constant $K_1$ such that
\[
\expect_\Phi \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ Z_N(\psibf) } \leq K_1 \sqrt{C_\epsilon}.
\]
\end{lemma}
\begin{IEEEproof}
Let
\[
B_\epsilon = \{\xbf \in \reals \mid \|\xbf\|_\infty < \epsilon \}.
\]
%We are now required to prove that
%\[
%\expect_\Phi \sup_{\psibf \in B_\epsilon} \abs{ Z_N(\psibf) } \leq K_1 \sqrt{C_\epsilon}.
%\]
For each non negative integer $k$, let $T_\epsilon(k)$ be a discrete subset of $\reals^{m+1}$ with the property that for every $\psibf \in B_\epsilon$ there exists some $\psibf^* \in T_\epsilon(k)$ such that the pseudometric
\[
d(\psibf, \psibf^*) = \sum_{n = 1}^N \big( f_{nN}(\psibf) - f_{nN}(\psibf^*) \big)^2 \leq 2^{-k} C_\epsilon N.
\]
We specifically define $T_\epsilon(0)$ to contain a single point, that being the origin $\zerobf$.  Defined this way, $T_\epsilon(0)$ satisfies the inequality above because
\[
d(\psibf, \zerobf) = \sum_{n = 1}^N f_{nN}(\psibf)^2  \leq C_\epsilon N ,
\]
for all $\psibf \in B_\epsilon$, as a result of Lemma~\ref{lem:epslmlemma}.

The existence of $T_\epsilon(k)$ for each positive integer $k$ will be proved in Lemma~\ref{lem:metricentropy}.  It is worth giving some intuition regarding $T_\epsilon(k)$.  If we place a `ball' of radius $2^{-k} C_\epsilon N$  with respect to the pseudometric $d(\cdot, \cdot)$ around each point in $T_\epsilon(k)$, then, by definition, the union of these balls is a superset of $B_\epsilon$.  The balls are said to \emph{cover} $B_\epsilon$ and $T_\epsilon(k)$ is said to form a \emph{covering} of $B_\epsilon$~\cite[Section 1.2]{Dudley_unif_central_lim_th_1999}.  The minimum number of such balls required to cover $B_\epsilon$ is called a \emph{covering number} of $B_\epsilon$.  In Lemma~\ref{lem:metricentropy} we show that no more than $K_3 2^{(m+1)k}$ balls of radius $2^{-k} C_\epsilon N$ are required to cover $B_\epsilon$, where $K_3$ is a constant, independent of $N$ and $\epsilon$.  We make use of this upper bound in Lemma~\ref{lem:chaining3}.

We now continue the proof.  For all $\psibf \in \reals^{m+1}$, 
\[
f_{nN}(\psibf) = \left( \tfrac{n}{N}\right)^\ell \round{\Phi_n + p_{nN}(\psibf)}
\]
is a multiple of $N^{-\ell}$ and so $d(\psibf, \psibf^*)$ is a multiple of $N^{-2\ell}$.  When $2^k > C_\epsilon N^{1+2\ell}$ we have 
\[
0 \leq d(\psibf, \psibf^*) < 2^{-k} C_\epsilon N < N^{-2\ell},
\]
and so we must have $d(\psibf, \psibf^*) = 0$.  Consequently $f_{nN}(\psibf) = f_{nN}(\psibf^*)$  for every $n = 1, \dots, N$, and therefore $Z_N(\psibf) - Z_N(\psibf^*) = 0$.  Thus,
\[
\sup_{\|\psibf\|_{\infty} < \epsilon } \abs{ Z_N(\psibf) } = \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) }
\]
for all $k$ large enough that $2^{k} > C_\epsilon N^{1+2\ell}$.  So, to analyse the supremum of $Z_N(\psibf)$ over the continuous interval $B_\epsilon$ it is enough to analyse the supremum over the discrete set $T_\epsilon(k)$ for large $k$.  Lemma~\ref{lem:chaining3} shows that 
\[
\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) } \leq \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{\sqrt{ i A_1 + A_2}}{2^{i/2}} < \infty
\]
for every positive integer $k$, where $A_1$ and $A_2$ are the constants,
\begin{equation}\label{eq:A1andA2}
A_1 = 18(m+1)\log 2\ \qquad \text{and} \qquad A_2 = 18\log K_3
\end{equation}
and $\log(\cdot)$ is the natural logarithm.  The lemma holds with $K_1 =  \sum_{i=1}^{\infty}2^{-i/2}\sqrt{ i A_1 + A_2 }$.
\end{IEEEproof}

\begin{lemma}\label{lem:epslmlemma}
For $\epsilon < \frac{1}{m+1}$ and all $\psibf \in B_\epsilon$ and $n = 1, \dots, N$,
\[
f_{nN}(\psibf)^2 \leq \abs{f_{nN}(\psibf)} \leq I_\epsilon(\abs{\Phi_n})
\]
and consequently,
\begin{equation}\label{eq:sumf2absfCe}
\sum_{n=1}^{N}f_{nN}(\psibf)^2 \leq \sum_{n=1}^{N}\abs{f_{nN}(\psibf)} \leq N C_\epsilon.
\end{equation}
\end{lemma}
\begin{IEEEproof}
Recall that
\[
f_{nN}(\psibf) = \left( \tfrac{n}{N}\right)^\ell \round{\Phi_n + p_{nN}(\psibf)}.
\]
Because $\abs{\psi_i} < \epsilon$ for all $i = 0, \dots, m$, 
\begin{equation}\label{eq:pnNsmall}
\abs{p_{nN}(\psibf)} = \abs{\sum_{i=0}^{m}\big(\tfrac{n}{N}\big)^i \psi_i} \leq (m+1)\epsilon < 1.
\end{equation}
%for all $n = 1, \dots, N$.  
Since $\Phi_n \in [-\nicefrac{1}{2}, \nicefrac{1}{2})$, it follows that $f_{nN}(\psibf)$ equals either $-(\tfrac{n}{N})^\ell$, $(\tfrac{n}{N})^\ell$ or $0$ and so
\begin{equation}\label{eq:fnNleqfnNs}
   f_{nN}(\psibf)^2 \leq \abs{f_{nN}(\psibf)} \leq 1.
\end{equation}
Whenever $f_{nN}(\psibf) \neq 0$ we must have 
\[
\abs{\Phi_n} \geq \nicefrac{1}{2} - \abs{p_{nN}(\psibf)} \geq \nicefrac{1}{2} - (m+1)\epsilon
\]
and $I_\epsilon(\abs{\Phi_n}) = 1$.  Thus, for all $n = 1, \dots, N$,
\[
f_{nN}(\psibf)^2 \leq \abs{f_{nN}(\psibf)} \leq I_\epsilon(\abs{\Phi_n}). %\qquad \text{for all $n = 1, \dots, N$.}
\]
Summing the terms in this inequality over $n = 1, \dots, N$ and using~\eqref{eq:Cedefn} gives~\eqref{eq:sumf2absfCe}.
\end{IEEEproof}

\begin{lemma}\label{lem:chaining3}(Chaining)
For all positive integers $k$,
\[
\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) } \leq \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{\sqrt{ i A_1 + A_2 }}{2^{i/2}},
\]
where the constants $A_1$ and $A_2$ are defined in~\eqref{eq:A1andA2}.
\end{lemma}
\begin{IEEEproof}
Let $b_k$ be a function that maps each $\psibf \in T_\epsilon(k)$ to $b_k(\psibf) \in T_\epsilon(k-1)$ such that 
\[
d(\psibf,b_k(\psibf)) \leq 2^{1-k} C_\epsilon N.
\]
The existence of the function $b_k$ is guaranteed by the definition of $T_\epsilon(k)$.  By the triangle inequality,
\[
\abs{ Z_N(\psibf) } \leq \abs{ Z_N(b_k(\psibf)) } + \abs{ Z_N(\psibf) - Z_N(b_k(\psibf))  },
\]
and by taking supremums on both sides,
\begin{align}
&\sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) } \nonumber \\ %\leq \sup_{\psibf \in T_\epsilon(k) } \big( \abs{ Z_N(\psibf^*) } + \abs{ Z_N(\psibf) - Z_N(\psibf^*)  } \big) \nonumber \\
&\leq \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(b_k(\psibf)) } + \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) - Z_N(b_k(\psibf))  } \nonumber \\
&\leq \sup_{\psibf \in T_\epsilon(k-1)} \sabs{ Z_N(\psibf) } + \sup_{\psibf \in T_\epsilon(k)} \sabs{ Z_N(\psibf) - Z_N(b_k(\psibf)) }, \label{eq:supZchain}
\end{align}
the last line following since $b_k(\psibf) \in T_\epsilon(k-1)$ and so 
\[
\sup_{\psibf \in T_\epsilon(k)} \abs{ Z_N(b_k(\psibf)) } \leq \sup_{\psibf \in T_\epsilon(k-1)} \abs{ Z_N(\psibf) }.
\]
Conditional on $\Phi_1, \Phi_2, \dots$ the random variable
\[
X(\psibf) = Z_N(\psibf) - Z_N(b_k(\psibf))
\]
has zero mean, and is normally distributed with variance
\begin{align*}
\sigma_X^2 &= \expect_{\Phi} \frac{1}{N} \sum_{n=1}^N g_n^2 ( f_{nN}(\psibf) - f_{nN}(b_k(\psibf)) )^2 \\
&= \frac{1}{N} \sum_{n=1}^N ( f_{nN}(\psibf) - f_{nN}(b_k(\psibf)) )^2 \\
&= d(\psibf, b_k(\psibf)) \leq 2^{1-k} C_\epsilon,
\end{align*}
since $\expect_{\Phi} g_n^2 = 1$.  Using Lemma~\ref{lem:maxineq}, 
\begin{align*}
\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k)} \sabs{X(\psibf)} &\leq 3\sqrt{2^{1-k} C_\epsilon \log\sabs{T_\epsilon(k)}} \\
&\leq \sqrt{C_\epsilon}\frac{\sqrt{  k A_1 + A_2 }}{2^{k/2}}
\end{align*}
because $\log\abs{T_\epsilon(k)} \leq k(m+1)\log 2 +  \log K_3$.  Taking expectations on both side of~\eqref{eq:supZchain},
\begin{align*}
\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) } &\leq \expect_{\Phi} \sup_{\psibf \in T_\epsilon(k-1)} \sabs{ Z_N(\psibf) } \\
&\hspace{1cm} + \sqrt{C_\epsilon}\frac{\sqrt{  k A_1 + A_2 }}{2^{k/2}},
\end{align*}
which involves a recursion in $k$.  By unravelling the recursion, and using the fact $T_\epsilon(0)$ contains only the origin, and therefore 
\[
\expect_{\Phi}\sup_{\psibf \in T_\epsilon(0)} \abs{ Z_N(\psibf) } = \expect_{\Phi}\abs{ Z_N(0) } = 0,
\]
we obtain
\begin{align*}
\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) } < \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{\sqrt{  k A_1 + A_2 }}{2^{k/2}}
\end{align*}
as required.
\end{IEEEproof}

\begin{lemma}\label{lem:maxineq}(Maximal inequalities)
Suppose $X_1, \dots, X_N$ are zero mean Gaussian random variables each with variance less than some positive constant $K$, then
\[
\expect \sup_{n = 1, \dots, N} \abs{X_n} \leq 3  \sqrt{K \log N}
\]
where $\log N$ is the natural logarithm of $N$.
\end{lemma}
\begin{IEEEproof}
This result is well known, see for example~\cite[Section 3]{Pollard_asymp_empi_proc_1989}  
\end{IEEEproof}

\begin{lemma}\label{lem:metricentropy}(Covering numbers)
For $k \in \ints$ there exists a discrete set $T_\epsilon(k) \subset \reals^{m+1}$ with the property that, for every $\psibf \in B_\epsilon$, there is a $\psibf^* \in T_\epsilon(k)$ such that,
\[
d(\psibf,\psibf^*) = \sum_{n = 1}^N \big( f_{nN}(\psibf) - f_{nN}(\psibf^*) \big)^2 \leq \frac{C_\epsilon N}{2^k}.
\]
The number of elements in $T_\epsilon(k)$ is no more than $K_3 2^{(m+1)k}$ where $K_3$ is a positive constant, independent of $N$, $\epsilon$ and $k$.
\end{lemma}

Before we give the proof of this lemma we need some results from the literature on hyperplane arrangements and what are called \emph{$\epsilon$-cuttings}~\cite{Chazelle_discrepency_method_2000,Matousek_lect_disc_geom_2002}.  Let $H$ be a set of $m$-dimensional affine hyperplanes lying in $\reals^{m+1}$.  By \emph{affine} it is meant that  the hyperplanes need not pass through the origin.  For each hyperplane $h \in H$ let $D(h)$ and its complement $\bar{D}(h)$ be the corresponding half spaces of $\mathbb{R}^{m+1}$.  For a point $\xbf \in \mathbb{R}^{m+1}$, let
\begin{equation}\label{eq:bhsi}
b(h,\xbf) = \begin{cases} 
1 & \xbf \in D(h) \\
0 & \xbf \in \bar{D}(h).
\end{cases}
\end{equation}
Note that $b(h,\xbf)$ is piecewise constant in $\xbf$.  Two points $\xbf$ and $\ybf$ from $\reals^{m+1}$ are in the same halfspace of $h$ if and only if $b(h, \xbf) = b(h, \ybf)$.  Then the pseudometric
\[
\sigma(\xbf,\ybf) = \sum_{h \in H} |b(h,\xbf) - b(h,\ybf)|
\]
is the number of hyperplanes in $H$ that pass between the points $\xbf$ and $\ybf$.  %Observe that $\sigma(\xbf, \ybf)$ is peicewise constant in both $\xbf$ and $\ybf$.

The next theorem considers the partitioning of $\reals^{m+1}$ into subsets so that not too many hyperplanes intersect with any subset.  Proofs can be found in Theorem 5.1 on page 206 of~\cite{Chazelle_discrepency_method_2000} and also Theorem 6.5.3 on page 144 of~\cite{Matousek_lect_disc_geom_2002}.

\begin{theorem}\label{thm:epscutting}
There exists a constant $K$, independent of the set of hyperplanes $H$, such that for any positive real number $r$, we can partition $\reals^{m+1}$ into $K r^{m+1}$ generalised $(m+1)$-dimensional simplices with the property that no more than $\abs{H}/r$ hyperplanes from $H$ pass through the interior of any simplex.
\end{theorem}

By the phrase `There exists a constant $K$, independent of the set of hyperplanes $H$', it is meant that the constant $K$ is valid for every possible set of hyperplanes in $\reals^{m+1}$, regardless of the number of hyperplanes or their position and orientation.  A \emph{generalised} $(m+1)$-dimensional simplex is the region defined by the intersection of $m+2$ half spaces in $\reals^{m+1}$.  Note that a generalised simplex (unlike an ordinary simplex) can be unbounded.  For our purposes Theorem~\ref{thm:epscutting} is important because of the following corollary.

\begin{corollary}\label{cor:epscutting}
There exists a constant $K$, independent of the set of hyperplanes $H$, such that for every positive real number $r$ there is a discrete subset $T \subset \reals^{m+1}$ containing no more than $K r^{m+1}$ elements with the property that for every $\xbf \in \reals^{m+1}$ there exists $\ybf \in T$ with $\sigma(\xbf, \ybf) \leq \abs{H}/r$.
\end{corollary}
\begin{IEEEproof}
Let $C$ be the set of generalised simplices constructed according to Theorem~\ref{thm:epscutting}.  Define $T$ as a set containing precisely one point from the interior of each simplex in $C$.  Let $\xbf \in \reals^{m+1}$.  Since $b(h, \xbf)$ is piecewise constant for each $h \in H$, and since $C$ partitions $\reals^{m+1}$, there must exist a simplex $c \in C$ with a point $\zbf$ in its interior such that $b(h,\zbf) = b(h, \xbf)$ for all $h \in H$, and correspondingly $\sigma(\zbf, \xbf) = 0$.  Let $\ybf$ be the element from $T$ that is in the interior of $c$.  Since at most $\abs{H}/r$ hyperplanes cross the interior of $c$ there can be at most $\abs{H}/r$ hyperplanes between $\zbf$ and $\ybf$, and so $\sigma(\zbf, \ybf) \leq \abs{H}/r$.  Now, 
\[
\sigma(\xbf, \ybf) < \sigma(\zbf, \xbf) + \sigma(\zbf, \ybf) \leq \frac{\abs{H}}{r}
\]
follows from the triangle inequality.
\end{IEEEproof}

The previous corollary ensures that we can \emph{cover} $\reals^{m+1}$ using $K r^{m+1}$ `balls' of radius $\abs{H}/r$ with respect to the pseudometric $\sigma(\xbf, \ybf)$.  The balls are placed at the positions defined by points in the set $T$, and these points could be anywhere in $\reals^{m+1}$.  The next corollary asserts that we can cover a subset of $\reals^{m+1}$, by placing the balls at points only within this subset.

\begin{corollary}\label{cor:epscuttingsubset}
Let $B$ be a subset of $\reals^{m+1}$.  There exists a constant $K$, independent of the set of hyperplanes $H$, such that for every positive real number $r$ there is a discrete subset $T_B \subset B$ containing no more than $K r^{m+1}$ elements with the property that for every $\xbf \in B$ there is a $\ybf \in T_B$ with $\sigma(\xbf, \ybf) \leq \abs{H}/r$.
\end{corollary}
\begin{IEEEproof}
Let $C$ be the set of generalised simplices constructed according to Theorem~\ref{thm:epscutting} and let $C_B$ be the subset of those indices that intersect $B$.  Let $T_B$ contain a point from $c \cap B$ for each simplex $c \in C_B$.  The proof now follows similarly to Corollary~\ref{cor:epscutting}.
\end{IEEEproof}

We are now ready to prove Lemma~\ref{lem:metricentropy}.

\begin{IEEEproof} (Lemma~\ref{lem:metricentropy})
Put 
\[
g_{nN}(\psibf) = (\tfrac{N}{n})^\ell f_{nN}(\psibf) = \round{ \Phi_n + p_{nN}(\psibf)},
\]
and let
\[
d_g(\psibf, \psibf^*) = \sum_{n=1}^{N} \big( g_{nN}(\psibf) - g_{nN}(\psibf^*) \big)^2.
\]
We have $d(\psibf, \psibf^*) \leq d_g(\psibf, \psibf^*)$, and so it suffices to prove the Lemma with $d$ replaced by $d_g$.  
%It will be more convenient to work with $d_g$ rather than $d$.

From~\eqref{eq:pnNsmall}, 
\[
\abs{p_{nN}(\psibf)} \leq (m+1)\epsilon < 1.
\]
Since $\Phi_n \in [-\nicefrac{1}{2}, \nicefrac{1}{2})$, when $\Phi_n \geq 0$,
\[
g_{nN}(\psibf) = \begin{cases}
1 & p_{nN}(\psibf) \geq \nicefrac{1}{2} - \Phi_n \\
0 & \text{otherwise},
\end{cases}
\]
and when $\Phi_n < 0$,
\[
g_{nN}(\psibf) = \begin{cases}
-1 & p_{nN}(\psibf) < -\nicefrac{1}{2} - \Phi_n \\
0 & \text{otherwise}.
\end{cases}
\]
Thus, $( g_{nN}(\psibf) - g_{nN}(\psibf^*) )^2$ is either equal to one when $g_{nN}(\psibf) \neq g_{nN}(\psibf^*)$ or zero when $g_{nN}(\psibf) = g_{nN}(\psibf^*)$.  Now $g_{nN}(\psibf) \neq 0$ only if 
\[
\abs{\Phi_n} \geq \nicefrac{1}{2} - \abs{p_{nN}(\psibf)} \geq \nicefrac{1}{2} - (m+1)\epsilon,
\]
that is, only if $I_\epsilon(\Phi_n) = 1$.  Let
\[
A = \{ n \in \{1, \dots, N\} \mid I_\epsilon(\Phi_n) = 1 \}
\]
be the subset of the indices where $I_\epsilon(\Phi_n) = 1$.  By definition the number of elements in $A$ is $C_\epsilon N$ (see~\eqref{eq:Cedefn}).  If both $\psibf$ and $\psibf^*$ are in $B_\epsilon$, then 
\[
( g_{nN}(\psibf) - g_{nN}(\psibf^*) )^2 \neq 0 
\]
only if $n \notin A$.  Thus,
\begin{align*}
d_g(\psibf, \psibf^*) &= \sum_{n=1}^{N} \big( g_{nN}(\psibf) - g_{nN}(\psibf^*) \big)^2 \\
&= \sum_{n \in A} \big( g_{nN}(\psibf) - g_{nN}(\psibf^*) \big)^2.
\end{align*}

We now use Corollary~\ref{cor:epscuttingsubset}.  Let $h_n$ be the $m$ dimensional hyperplane in $\reals^{m+1}$ satisfying
\[
p_{nN}(\psibf) =  \sum_{i=0}^{m} \big(\tfrac{n}{N}\big)^i \psi_i = \tfrac{1}{2} \sign{\Phi_n} - \Phi_n
\]
where $\sign{\Phi_n}$ is equal to $1$ when $\Phi_n \geq 0$ and $-1$ otherwise.  The hyperplane $h_n$ divides $\reals^{m+1}$ into two halfspaces, $D(h_n)$ and its complement $\bar{D}(h_n)$.  If $\psibf$ and $\psibf^*$ are in the same halfspace, then $\abs{ b(h_n,\psibf) - b(h_n,\psibf^*) } = 0$ and $g_{nN}(\psibf) = g_{nN}(\psibf^*)$, and therefore $( g_{nN}(\psibf) - g_{nN}(\psibf^*) )^2 = 0$.  Otherwise, if $\psibf$ and $\psibf^*$ are in different halfspaces, then $\abs{ b(h_n,\psibf) - b(h_n,\psibf^*) } = 1$ and $g_{nN}(\psibf) \neq g_{nN}(\psibf^*)$, and therefore $( g_{nN}(\psibf) - g_{nN}(\psibf^*) )^2 = 1$.  Thus, 
\[
(g_{nN}(\psibf) - g_{nN}(\psibf^*) )^2 = \abs{ b(h_n,\psibf) - b(h_n,\psibf^*) }
\]
for $n = 1, \dots,  N$.  Let $H$ be the finite set of hyperplanes $\{ h_n, n \in A\}$ and observe that the number of hyperplanes is $\abs{H} = \abs{A} = C_\epsilon N$.  When both $\psibf$ and $\psibf^*$ are inside $B_\epsilon$, $d_g$ can be written as
\begin{align*}
d_g(\psibf, \psibf^*) &= \sum_{n\in A} \abs{ b(h_n,\psibf) - b(h_n,\psibf^*) } \\
&=  \sum_{h \in H} \abs{ b(h,\psibf) - b(h,\psibf^*) } = \sigma(\psibf, \psibf^*).
\end{align*}
That is, when both $\psibf, \psibf^* \in B_\epsilon$, $d_g(\psibf, \psibf^*)$ is the number of hyperplanes from $H$ that pass between the points $\psibf$ and $\psibf^*$.

It follows from Corollary~\ref{cor:epscuttingsubset} that for any positive $r$ there exists a finite subset $T_B$ of $B_\epsilon$ containing at most $K_3 r^{m+1}$ elements, such that for every $\psibf \in B_\epsilon$ there is a $\psibf^* \in T_B$ with 
\[
d_g(\psibf, \psibf^*) = \sigma(\psibf, \psibf^*) \leq \frac{\abs{H}}{r} = \frac{\abs{A}}{r} = \frac{C_\epsilon N}{r}.
\]
Putting $r = 2^k$ and choosing $T_\epsilon(k) = T_B$ completes the proof.
\end{IEEEproof}

%\begin{figure}[p] 
%   	\centering 
%  		\includegraphics[width=\linewidth]{polyfig/polyphase4plot.4} 
%  		\caption{MSE in the third order coefficient $\mu_3$ versus $\var{X_n} = \sigma^2$. The true coefficients are uniformly spread in the identifiable region.} 
%  		\label{plot:polyest_projnorm_mu3} 
% \end{figure}

% \begin{figure}[p] 
%  	\centering 
% 		\includegraphics[width=\linewidth]{polyfig/polyphase4plot_unfair.4} 
% 		\caption{MSE in the third order coefficient $\mu_3$ versus $\var{X_n} = \sigma_c^2$. The coefficient have been restricted for Kitchen's estimator and the DPT.} 
% 		\label{plot:polyest_projnorm_mu3_unfair} 
%\end{figure} 
% 
% \begin{figure}[p] 
%  	\centering 
% 		\includegraphics[width=\linewidth]{polyfig/polyphase4plot_fair.4} 
% 		\caption{MSE in the third order coefficient $\mu_3$ versus $\var{X_n} = \sigma_c^2$. The DPT estimator runs at the higher sampling rate $\delta$ so that the volumes $V_{DPT}(\delta) = V_m$.} 
% 		\label{plot:polyest_projnorm_mu3_fair} 
%\end{figure}

\end{document}

### 2012-10-12 build.sh

@@ -1,0 +1,25 @@
+#batch file to compile all the figures from 
+#my thesis. If you haven't go rubypost
+#installed you will want to uncomment
+#the gem line below
+
+#gem install rubypost
+
+cd plots 
+ruby1.9.1 circstatfigcube.rb
+ruby1.9.1 circstatfigquad.rb
+ruby1.9.1 circstatfig.rb
+ruby1.9.1 circstatfigzero.rb
+ruby1.9.1 tesselationfigures.rb
+mpost -interaction=nonstopmode distplots.class.distributions.circular.ProjectedNormalDistribution.var3.0.mp
+cd .. 
+
+cd code 
+mpost -interaction=nonstopmode gaussianplot4.mp 
+mpost -interaction=nonstopmode lseplot.mp 
+cd .. 
+
+pdflatex paper.tex
+bibtex paper.aux
+pdflatex paper.tex
+pdflatex paper.tex

### 2012-10-12 code/runsim.sh

@@ -1,0 +1,2 @@
+scala -nocompdaemon -cp PubSim.jar:Jama-1.0.2.jar:flanagan.jar:colt.jar:RngPack.jar simgaussian.scala
+scala -nocompdaemon -cp PubSim.jar:Jama-1.0.2.jar:flanagan.jar:colt.jar:RngPack.jar simwrappedu.scala

### 2012-10-12 code/simgaussian.scala

@@ -1,0 +1,115 @@
+/**
+ * Run simulations for the sparse noisy period estimation problem.
+ * Various estimators are available.
+ */
+import pubsim.poly.PolynomialPhaseSignal
+import pubsim.poly.KitchenEstimator
+import pubsim.poly.DPTEstimator
+import pubsim.poly.MbestEstimator
+import pubsim.poly.BabaiEstimator
+import pubsim.poly.MbestEstimator
+import pubsim.poly.MaximumLikelihood
+import pubsim.poly.SphereDecoderEstimator
+import pubsim.distributions.ContinuousRandomVariable
+import pubsim.poly.PolynomialPhaseEstimator
+import pubsim.poly.bounds.AngularLeastSquaresVariance
+import pubsim.poly.bounds.GaussianCRB
+import pubsim.distributions.GaussianNoise
+import pubsim.distributions.circular.ProjectedNormalDistribution
+
+val iters = 2000 //number of Monte-Carlo trials.
+val Ns = List(10,50,200) //values of N we will generate curves for
+val ms = List(2,3) //order of our polynomial phase signals
+
+def gp( a : Double, r : Double, n : Int ) : Double = a * scala.math.pow(r,n)
+
+//returns an array of noise distributions with a logarithmic scale
+def noises = (0 to 21).map( n => gp(0.0065,1.3,n) ).map( v => new GaussianNoise(0,v) )
+
+//Returns a list of functions that return estimators we will run (factory patern to enable parallelism)
+def estfactory(m : Int, N : Int) : List[() => PolynomialPhaseEstimator] = {
+  var ret = List( 
+    () => new KitchenEstimator(m,N),
+    () => new DPTEstimator(m,N),
+    () => new BabaiEstimator(m,N),
+    () => new MbestEstimator(m,N,4*N) 
+  )
+  //add the sphere decoder and Least squares estimators if N and m are small
+  if( N < 60 ) ret = ret :+ ( () => new SphereDecoderEstimator(m,N) )
+  if( N < 60 && m <= 2 ) ret = ret :+ ( () => new MaximumLikelihood(m,N) )
+  return ret
+}
+
+val starttime = (new java.util.Date).getTime
+
+//for all the the values of N and m.
+for( N <-  Ns; m <- ms ) {
+
+  for(estf <- estfactory(m,N) ){
+    
+    val estname = estf().getClass.getSimpleName + "N" + N.toString + "m" + m + "WU"
+    print("Running " + estname)
[…]

### 2012-10-12 code/simwrappedu.scala

@@ -1,0 +1,98 @@
+/**
+ * Run simulations for the sparse noisy period estimation problem.
+ * Various estimators are available.
+ */
+import pubsim.poly.CircularNoisePolynomialPhaseSignal
+import pubsim.poly.KitchenEstimator
+import pubsim.poly.DPTEstimator
+import pubsim.poly.MbestEstimator
+import pubsim.poly.BabaiEstimator
+import pubsim.poly.MbestEstimator
+import pubsim.poly.MaximumLikelihood
+import pubsim.poly.SphereDecoderEstimator
+import pubsim.distributions.ContinuousRandomVariable
+import pubsim.poly.PolynomialPhaseEstimator
+import pubsim.poly.bounds.AngularLeastSquaresVariance
+import pubsim.distributions.circular.WrappedUniform
+
+val iters = 2000 //number of Monte-Carlo trials.
+val Ns = List(10,50,200) //values of N we will generate curves for
+val ms = List(2,3) //order of our polynomial phase signals
+
+def gp( a : Double, r : Double, n : Int ) : Double = a * scala.math.pow(r,n)
+
+//returns an array of noise distributions with a logarithmic scale
+def noises = (0 to 25).map( n => gp(0.002,1.15,n) ).map( v => new WrappedUniform(0,v) )
+
+//Returns a list of functions that return estimators we will run (factory patern to enable parallelism)
+def estfactory(m : Int, N : Int) : List[() => PolynomialPhaseEstimator] = {
+  var ret = List( 
+    () => new KitchenEstimator(m,N),
+    () => new DPTEstimator(m,N),
+    () => new BabaiEstimator(m,N),
+    () => new MbestEstimator(m,N,4*N) 
+  )
+  //add the sphere decoder and Least squares estimators if N and m are small
+  if( N < 60 ) ret = ret :+ ( () => new SphereDecoderEstimator(m,N) )
+  if( N < 60 && m <= 2 ) ret = ret :+ ( () => new MaximumLikelihood(m,N) )
+  return ret
+}
+
+val starttime = (new java.util.Date).getTime
+
+//for all the the values of N and m.
+for( N <-  Ns; m <- ms ) {
+
+  for(estf <- estfactory(m,N) ){
+    
+    val estname = estf().getClass.getSimpleName + "N" + N.toString + "m" + m + "WU"
+    print("Running " + estname)
+    val eststarttime = (new java.util.Date).getTime
+
+    //for all the noise distributions (in parallel threads)
[…]

### 2012-10-12 plots/circstatfig.rb

@@ -1,0 +1,102 @@
+require 'rubypost'
+
+require 'matrix'
+require 'mathn'
+
+include RubyPost
+
+file = RubyPost::File.new('circstatfig')
+
+#def dround(d) return (10000.0*d).round/10000.0 end
+def dround(d) return d end
+
+def drawpoly(e, s, func)
+	polypath = Path.new
+	t = 0
+	while(t < e) do
+		polypath.add_pair(Pair.new(dround(t).cm, dround(func.call(t)).cm))
+		t+=s
+	end
+	return polypath
+end
+
+def drawsamplepoly(e, s, func, picture)
+	t = 0
+	while(t < e) do
+		picture.add_drawable(Draw.new(Circle.new()).scale(0.15.cm).translate(dround(t).cm, dround(func.call(t)).cm))
+		t+=s
+	end
+	return picture
+end
+
+def drawcircpoly(e, s, func)
+	polypath = Path.new.curved
+	t = 0
+	while(t < e) do
+		a = 2*Math::PI*func.call(t)
+		x = dround(t*Math.cos(a)/2.0)
+		y= dround(t*Math.sin(a)/2.0)		
+		polypath.add_pair(Pair.new(x.cm, y.cm))
+		t+=s
+	end
+	return polypath
+end
+
+def drawsamplecircpoly(e, s, func, picture)
+	t = 0
+	while(t < e) do
+    a = 2*Math::PI*func.call(t)
+		x = dround(t*Math.cos(a)/2.0)
+		y= dround(t*Math.sin(a)/2.0)		
+    picture.add_drawable(Draw.new(Circle.new()).scale(0.12.cm).translate(x.cm, y.cm))
+		t+=s
+	end
+	return picture
+end
+
+poly = lambda{ |t| 0.8*t + 0.3 } 
+poly2 = lambda{ |t| -0.2*t + 3.3 } 
+
+polypic = Picture.new('polypic')
+polypic.add_drawable('drawdblarrow (5.4cm,0)--(0,0)--(0,4.8cm);')
+polypic.add_drawable('label.rt(btex $y$ etex, (0,4.8cm));')
+polypic.add_drawable('label.top(btex $t$ etex, (5.4cm,0));')
+polypic.add_drawable('label.lft(btex $0.5$ etex, (0,0.5cm));')
+polypic.add_drawable('draw (-0.05cm, 0.5cm)--(0.05cm,0.5cm);')
[…]

### 2012-10-12 plots/circstatfigcube.rb

@@ -1,0 +1,103 @@
+require 'rubypost'
+
+require 'matrix'
+require 'mathn'
+
+include RubyPost
+
+file = RubyPost::File.new('circstatfigcube')
+
+#def dround(d) return (1000.0*d).round/1000.0 end
+def dround(d) return d end
+
+def drawpoly(e, s, func)
+	polypath = Path.new
+	t = 0
+	while(t < e) do
+		polypath.add_pair(Pair.new(dround(t).cm, dround(func.call(t)).cm))
+		t+=s
+	end
+	return polypath
+end
+
+def drawsamplepoly(e, s, func, picture)
+	t = 0
+	while(t < e) do
+		picture.add_drawable(Draw.new(Circle.new()).scale(0.15.cm).translate(dround(t).cm, dround(func.call(t)).cm))
+		t+=s
+	end
+	return picture
+end
+
+def drawcircpoly(e, s, func)
+	polypath = Path.new.curved
+	t = 0
+	while(t < e) do
+		a = 2*Math::PI*func.call(t)
+		x = dround(t*Math.cos(a)/2.0)
+		y= dround(t*Math.sin(a)/2.0)		
+		polypath.add_pair(Pair.new(x.cm, y.cm))
+		t+=s
+	end
+	return polypath
+end
+
+def drawsamplecircpoly(e, s, func, picture)
+	t = 0
+	while(t < e) do
+    a = 2*Math::PI*func.call(t)
+		x = dround(t*Math.cos(a)/2.0)
+		y= dround(t*Math.sin(a)/2.0)		
+    picture.add_drawable(Draw.new(Circle.new()).scale(0.12.cm).translate(x.cm, y.cm))
+		t+=s
+	end
+	return picture
+end
+
+# 1/4*t^3- 7/4*t^2 + 49/16*t + 3/40
+poly = lambda{ |t| 0.25*(t-0.5)*(t-2)*(t-4.5) + 1.2 } 
+poly2 = lambda{ |t| -1/12.0 * t**3 + 1/4.0*t**2 + (49.0/16 - 2.0/3 - 2)*t + 1 + 3.0/40 } 
+
+polypic = Picture.new('polypic')
+polypic.add_drawable('drawdblarrow (5.4cm,0)--(0,0)--(0,4.8cm);')
+polypic.add_drawable('label.rt(btex $y$ etex, (0,4.8cm));')
+polypic.add_drawable('label.top(btex $t$ etex, (5.4cm,0));')
[…]

### 2012-10-12 plots/circstatfigquad.rb

@@ -1,0 +1,102 @@
+require 'rubypost'
+
+require 'matrix'
+require 'mathn'
+
+include RubyPost
+
+file = RubyPost::File.new('circstatfigquad')
+
+#def dround(d) return (10000.0*d).round/10000.0 end
+def dround(d) return d end
+
+def drawpoly(e, s, func)
+	polypath = Path.new
+	t = 0
+	while(t < e) do
+		polypath.add_pair(Pair.new(dround(t).cm, dround(func.call(t)).cm))
+		t+=s
+	end
+	return polypath
+end
+
+def drawsamplepoly(e, s, func, picture)
+	t = 0
+	while(t < e) do
+		picture.add_drawable(Draw.new(Circle.new()).scale(0.15.cm).translate(dround(t).cm, dround(func.call(t)).cm))
+		t+=s
+	end
+	return picture
+end
+
+def drawcircpoly(e, s, func)
+	polypath = Path.new.curved
+	t = 0
+	while(t < e) do
+		a = 2*Math::PI*func.call(t)
+		x = dround(t*Math.cos(a)/2.0)
+		y= dround(t*Math.sin(a)/2.0)		
+		polypath.add_pair(Pair.new(x.cm, y.cm))
+		t+=s
+	end
+	return polypath
+end
+
+def drawsamplecircpoly(e, s, func, picture)
+	t = 0
+	while(t < e) do
+    a = 2*Math::PI*func.call(t)
+		x = dround(t*Math.cos(a)/2.0)
+		y= dround(t*Math.sin(a)/2.0)		
+    picture.add_drawable(Draw.new(Circle.new()).scale(0.12.cm).translate(x.cm, y.cm))
+		t+=s
+	end
+	return picture
+end
+
+poly = lambda{ |t| 0.4*t**2 - 1.5*t+ 1.5 } 
+poly2 = lambda{ |t| -0.1*t**2 - 0*t+ 2.5 } 
+
+polypic = Picture.new('polypic')
+polypic.add_drawable('drawdblarrow (5.4cm,0)--(0,0)--(0,4.8cm);')
+polypic.add_drawable('label.rt(btex $y$ etex, (0,4.8cm));')
+polypic.add_drawable('label.top(btex $t$ etex, (5.4cm,0));')
+polypic.add_drawable('label.lft(btex $0.5$ etex, (0,0.5cm));')
[…]

### 2012-10-12 plots/circstatfigzero.rb

@@ -1,0 +1,103 @@
+require 'rubypost'
+
+require 'matrix'
+require 'mathn'
+
+include RubyPost
+
+file = RubyPost::File.new('circstatfigzero')
+
+#def dround(d) return (10000.0*d).round/10000.0 end
+def dround(d) return d end
+
+def drawpoly(e, s, func)
+	polypath = Path.new
+	t = 0
+	while(t < e) do
+		polypath.add_pair(Pair.new(dround(t).cm, dround(func.call(t)).cm))
+		t+=s
+	end
+	return polypath
+end
+
+def drawsamplepoly(e, s, func, picture)
+	t = 0
+	while(t < e) do
+		picture.add_drawable(Draw.new(Circle.new()).scale(0.15.cm).translate(dround(t).cm, dround(func.call(t)).cm))
+		t+=s
+	end
+	return picture
+end
+
+def drawcircpoly(e, s, func)
+	polypath = Path.new.curved
+	t = 0
+	while(t < e) do
+		a = 2*Math::PI*func.call(t)
+		x = dround(t*Math.cos(a)/2.0)
+		y= dround(t*Math.sin(a)/2.0)		
+		polypath.add_pair(Pair.new(x.cm, y.cm))
+		t+=s
+	end
+	return polypath
+end
+
+def drawsamplecircpoly(e, s, func, picture)
+	t = 0
+	while(t < e) do
+    a = 2*Math::PI*func.call(t)
+		x = dround(t*Math.cos(a)/2.0)
+		y= dround(t*Math.sin(a)/2.0)		
+    picture.add_drawable(Draw.new(Circle.new()).scale(0.12.cm).translate(x.cm, y.cm))
+		t+=s
+	end
+	return picture
+end
+
+
+poly = lambda{ |t| 0.7 } 
+poly2 = lambda{ |t| 1.7 } 
+
+polypic = Picture.new('polypic')
+polypic.add_drawable('drawdblarrow (5.4cm,0)--(0,0)--(0,4.8cm);')
+polypic.add_drawable('label.rt(btex $y$ etex, (0,4.8cm));')
+polypic.add_drawable('label.top(btex $t$ etex, (5.4cm,0));')
+polypic.add_drawable('label.lft(btex $0.5$ etex, (0,1cm));')
+polypic.add_drawable('draw (-0.05cm, 1cm)--(0.05cm,1cm);')
[…]

### 2012-10-12 plots/tesselationfigures.rb

@@ -1,0 +1,134 @@
+require 'rubypost'
+
+require 'matrix'
+require 'mathn'
+
+include RubyPost
+
+file = RubyPost::File.new('tesselationfigures')
+
+#create an axis peicture
+axes = Picture.new('axes')
+
+#axes.add_drawable('draw (0.15cm, 0)--(-0.15cm, 0);')
+#axes.add_drawable('draw (0, 0.15cm)--(0, -0.15cm);')
+
+#~ xaxis = Path.new
+#~ xaxis.add_pair(Pair.new(-3.95.cm, 0))
+#~ xaxis.add_pair(Pair.new(3.95.cm, 0))
+#~ yaxis = Path.new
+#~ yaxis.add_pair(Pair.new(0, -3.95.cm))
+#~ yaxis.add_pair(Pair.new(0, 3.95.cm))
+
+#~ axes.add_drawable(DoubleArrow.new(xaxis))
+#~ axes.add_drawable(DoubleArrow.new(yaxis))
+#axes.add_drawable(Label.new(latex("$y$"), Pair.new(0,3.95.cm)).bottom_right)
+#axes.add_drawable(Label.new(latex("$x$"), Pair.new(3.95.cm, 0)).top_left)
+
+
+
+pic = Picture.new('picL')
+pic.add_drawable(Draw.new(axes))
+
+matscale = 1.2
+
+M = Matrix[ [1,0.2], [0.2, 1] ]*(matscale)
+(-5..5).each do |x|
+	(-5..5).each do |y|
+		v = M*Vector[x,y]
+		pic.add_drawable(Fill.new(Circle.new()).scale(0.07.cm).translate(v[0].cm,v[1].cm).colour(0,0,0))
+	end
+end
+
+#####################
+# Figure with disconnected triangular tesselating region
+#####################
+v10 = M*Vector[1,0]
+v01 = M*Vector[0,1]
+v10n = M*Vector[-2,-1]
+v01n = M*Vector[-1,-2]
+v11n = M*Vector[-1,-1]
+v00 = M*Vector[0,0]
+tricyc1 = Path.new
+tricyc1.add_pair(Pair.new(0,0))
+tricyc1.add_pair(Pair.new(v10[0].cm,v10[1].cm))
+tricyc1.add_pair(Pair.new(v01[0].cm,v01[1].cm))
+tricyc1.add_pair('cycle')
+tricyc2 = Path.new
+tricyc2.add_pair(Pair.new(v11n[0].cm,v11n[0].cm))
[…]

### 2012-10-12 build.sh
edited readme.

@@ -1,17 +1,13 @@
-#batch file to compile all the figures from 
-#my thesis. If you haven't go rubypost
-#installed you will want to uncomment
-#the gem line below
-
-#gem install rubypost
+#batch file to compile figures and latex
 
 cd plots 
-ruby1.9.1 circstatfigcube.rb
-ruby1.9.1 circstatfigquad.rb
-ruby1.9.1 circstatfig.rb
-ruby1.9.1 circstatfigzero.rb
-ruby1.9.1 tesselationfigures.rb
-mpost -interaction=nonstopmode distplots.class.distributions.circular.ProjectedNormalDistribution.var3.0.mp
+mpost -interaction=nonstopmode circstatfigcube.mp 
+mpost -interaction=nonstopmode circstatfigquad.mp 
+mpost -interaction=nonstopmode circstatfig.mp 
+mpost -interaction=nonstopmode circstatfigzero.mp 
+mpost -interaction=nonstopmode tesselationfigures.mp 
+mpost -interaction=nonstopmode 
+distplots.class.distributions.circular.ProjectedNormalDistribution.var3.0.mp 
 cd .. 
 
 cd code 
@@ -19,7 +15,7 @@ mpost -interaction=nonstopmode gaussianplot4.mp
 mpost -interaction=nonstopmode lseplot.mp 
 cd .. 
 
-pdflatex paper.tex
-bibtex paper.aux
-pdflatex paper.tex
-pdflatex paper.tex
+pdflatex paper.tex 
+bibtex paper.aux 
+pdflatex paper.tex 
+pdflatex paper.tex 

### 2012-10-12 code/runsim.sh

@@ -1,2 +1,4 @@
+#export JAVA_OPTS="-Xprof -server -Xms1g -Xmx1g"
+export JAVA_OPTS="-d64 -server -Xms1g -Xmx1g"
 scala -nocompdaemon -cp PubSim.jar:Jama-1.0.2.jar:flanagan.jar:colt.jar:RngPack.jar simgaussian.scala
 scala -nocompdaemon -cp PubSim.jar:Jama-1.0.2.jar:flanagan.jar:colt.jar:RngPack.jar simwrappedu.scala

### 2012-10-12 build.sh
Added aop templates

@@ -16,6 +16,6 @@ mpost -interaction=nonstopmode lseplot.mp
 cd .. 
 
 pdflatex paper.tex 
-bibtex paper.aux 
+bibtex paper 
 pdflatex paper.tex 
 pdflatex paper.tex 

### 2012-10-13 paper.tex
filled in ims template
moved ims template to paper.

@@
-%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
-\documentclass[journal]{IEEEtran}
-
+% Template for the submission to:
+%   The Annals of Probability           [aop]
+%   The Annals of Applied Probability   [aap]
+%   The Annals of Statistics            [aos] 
+%   The Annals of Applied Statistics    [aoas]
+%   Stochastic Systems                  [ssy]
+%
+%Author: In this template, the places where you need to add information
+%        (or delete line) are indicated by {???}.  Mostly the information
+%        required is obvious, but some explanations are given in lines starting
+%Author:
+%All other lines should be ignored.  After editing, there should be
+%no instances of ??? after this line.
+
+% use option [preprint] to remove info line at bottom
+% journal options: aop,aap,aos,aoas,ssy
+% natbib option: authoryear
+\documentclass[aap]{imsart}
+
+\usepackage{amsthm,amsmath,amsfonts,amssymb}
+\usepackage[square,comma,numbers,sort&compress]{natbib}
+%\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
+\usepackage{units}
+\usepackage{ifpdf}
+\ifpdf
+  \usepackage[pdftex]{graphicx}
+\else
+  \usepackage{graphicx}% standard graphics package for inclusion images and eps files into LaTeX document
+\fi
+
+% provide arXiv number if available:
+%\arxiv{arXiv:0000.0000}
+
+% put your definitions there:
+\startlocaldefs
+\newcommand{\reals}{{\mathbb R}}
+\newcommand{\ints}{{\mathbb Z}}
+\newcommand{\term}{\emph}
+\newcommand{\var}{\operatorname{var}}
+\newcommand{\covar}{\operatorname{covar}}
+%\newcommand{\prob}{{\mathbb P}}
+\newcommand{\prob}{\operatorname{Pr}}
+\newcommand{\expect}{{\mathbb E}}
+\newcommand{\dealias}{\operatorname{dealias}}
+\renewcommand{\mid}{\; ; \;}
+\newcommand{\fracpart}[1]{\left\langle #1 \right\rangle}
+\newcommand{\sfracpart}[1]{\langle #1 \rangle}
+\newcommand{\abs}[1]{\left\vert #1 \right\vert}
+\newcommand{\sabs}[1]{\vert #1 \vert}
+\newcommand{\floor}[1]{{\left\lfloor #1 \right\rfloor}}
+\newcommand{\ceiling}[1]{{\left\lceil #1 \right\rceil}}
+\newcommand{\ceil}[1]{\lceil #1 \rceil}
+\newcommand{\round}[1]{{\left\lceil #1 \right\rfloor}}
 \usepackage{mathbf-abbrevs}
-
-%\newcommand{\dealias}{\operatorname{dealias}}
-
-\input{defs}
-
-%\newcommand{\term}{\emph}
-%\renewcommand{\index}[1]{}
-
-%\title{Polynomial phase estimation by phase unwrapping 1: Identifiability and strong consistency}
+\newtheorem{proposition}{Proposition}
+\newtheorem{corollary}{Corollary}
+\newtheorem{theorem}{Theorem}
+\newtheorem{lemma}{Lemma}
+\newtheorem{definition}{Definition}
+\newcommand{\coef}{\operatorname{coef}}
+\newcommand{\cubr}[1]{{\left\{ #1 \right\}}}
+\newcommand{\scubr}[1]{{\{ #1 \}}}
+\newcommand{\sign}[1]{\operatorname{sgn}\left( #1 \right)}
+\endlocaldefs
+
+\begin{document}
+
+\begin{frontmatter}
+
+% "Title of the paper"
 \title{Polynomial phase estimation by phase unwrapping}
-
-\author{Robby~G.~McKilliam, Barry~G.~Quinn, I.~Vaughan~L.~Clarkson, Bill~Moran and Badri~N.~Vellambi%
-    \thanks{%A preliminary version of some of this material is contained in Part 3 of Robby McKilliam's PhD thesis \cite{McKilliam2010thesis}. 
-Robby~McKilliam and Badri Vellambi are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.
-Barry~Quinn is with the Department of Statistics, Macquarie University, Sydney, NSW, 2109, Australia.
-Vaughan~Clarkson is with the School of Information Technology \& Electrical Engineering, The University of Queensland, QLD., 4072, Australia.
-B.
-Moran is with the Department of Electrical Engineering and Computer
-Science, Melbourne Systems Lab, Dept of Elec \& Electronic Eng, Uni of Melbourne, Vic. 3010, Australia.}}
-% The paper headers 
-\markboth{Polynomial phase estimation by phase unwrapping}{DRAFT \today}
-
-% make the title area
-
-\begin{document} 
-\maketitle
+\runtitle{Polynomial phase estimation by phase unwrapping}
+
+% indicate corresponding author with \corref{}
+% \author{\fnms{John} \snm{Smith}\corref{}\ead[label=e1]{smith@foo.com}\thanksref{t1}}
+% \thankstext{t1}{Thanks to somebody} 
+% \address{line 1\\ line 2\\ printead{e1}}
+% \affiliation{Some University}
+
+\begin{aug}
+\author{\fnms{Robby G.} \snm{McKilliam}\thanksref{m1}\ead[label=e1]{robby.mckilliam@unisa.edu.au}},
+\author{\fnms{Barry G.} \snm{Quinn}\thanksref{m2}\ead[label=e2]{barry.quinn@mq.edu.au}},
+\author{\fnms{I. Vaughan L.} \snm{Clarkson}\thanksref{m3}\ead[label=e3]{v.clarkson@uq.edu.au}},
+\author{\fnms{Bill} \snm{Moran}\thanksref{m4}\ead[label=e4]{wmoran@unimelb.edu.au}},
+\and
+\author{\fnms{Badri N.} \snm{Vellambi}\thanksref{m1}\ead[label=e5]{badri.vellambi@unisa.edu.au}}
+\runauthor{R. McKilliam et al.}
+
+\affiliation{University of South Australia\thanksmark{m1}, Macquarie University\thanksmark{m2}, University of Queensland\thanksmark{m3} and Melbourne University\thanksmark{m4}}
+
+\address{R. G. McKilliam and B. N. Vellambi\\
+Institute for Telecommunications Research\\
+University of South Australia, Adelaide\\
+SA, 5095, Australia\\
+\printead{e1}\\
+\phantom{E-mail:\ }\printead*{e5}}
+
+\address{B. G. Quinn\\
+Department of Statistics\\
+Macquarie University, Sydney,\\
+NSW, 2109, Australia\\
+\printead{e3}}
+
+\address{I. V. L. Clarkson\\
+School of Information Technology \& Electrical\\
+Engineering, University of Queensland,\\
+Brisbane, QLD, 4072, Australia\\
+\printead{e3}}
+
+\address{B. Moran\\
+Department of Electrical \& Electronic\\
+Engineering, Melbourne University,\\
+Melbourne, VIC, 3010, Australia\\
+\printead{e4}}
+
+\end{aug}
 
@@
 
-\begin{keywords}
-Polynomial phase signals, phase unwrapping, asymptotic properties, nearest lattice point problem
-\end{keywords}
-
-%\bibliographystyle{unsrt}
+%\begin{keyword}[class=AMS]
+%\kwd[Primary ]{}
+%\kwd{}
+%\kwd[; secondary ]{}
+%\end{keyword}
+
+\begin{keyword}
+\kwd{Polynomial phase signals, phase unwrapping, asymptotic statistics}
+\end{keyword}
+
+\end{frontmatter}
+
+% AOS,AOAS: If there are supplements please fill:
+%\begin{supplement}[id=suppA]
+%  \sname{Supplement A}
+%  \stitle{Title}
+%  \slink[url]{http://lib.stat.cmu.edu/aoas/???/???}
+%  \sdescription{Some text}
+%\end{supplement}
 
@@
 In radar and sonar applications polynomial phase signals arise when acquiring radial velocity and acceleration (and higher order motion descriptors) of a target from a reflected signal, and also in continuous wave radar and low probability of intercept radar~\cite{Levanon_Radar_signals_2004}.
-In biology, polynomial phase signals are used to describe the sounds emitted by bats and dolphins for echo location \citep{Suga_1975_bats_echolocation, Moss_2005echolocation}.
+In biology, polynomial phase signals are used to describe the sounds emitted by bats and dolphins for echo location \cite{Suga_1975_bats_echolocation, Moss_2005echolocation}.
 
@@
 In this paper we only consider uniform sampling, where the gap between consecutive samples is constant.
-In this case we can always consider the samples to be taken at some set of consecutive integers and our sampled polynomial phase signal looks like
+In this case we can always consider the samples to be taken at some set of consecutive integers and our sampled polynomial phase signal is
 \[
@@
 
-%\subsection{Notation}
-%We write random variables using capital letters, such as $X$ and $Y$ and circular random variables using the capital Greek letters $\Theta$ and $\Phi$.
-We use $\round{x}$ to denote the nearest integer to $x$ with half integers rounded up and use $\fracpart{x} = x - \round{x}$ to denote the \emph{centred} fractional part.
-Both $\round{\cdot}$ and $\fractpart{\cdot}$ operate elementwise on vectors.
-
-\section{Lattice Theory}\label{sec:lattice-theory}
+\section{Lattices}\label{sec:lattice-theory}
 
@@
 \end{proposition}
-\begin{IEEEproof}
+\begin{proof}
 This result is well known~\cite[Chapter IX, Theorem IV]{Cassels_geom_numbers_1997}~\cite[Proposition 2.1]{McKilliam2010thesis}.
@@
 A result in the general case can be obtained similarly, but is not required here.
-\end{IEEEproof}
+\end{proof}
 
@@
 	\centering
-		\includegraphics[width=\linewidth]{plots/tesselationfigures-2.mps}
+		\includegraphics[width=0.8\linewidth]{plots/tesselationfigures-2.mps}
 		\caption{Rectangular tessellation constructed according to Proposition~\ref{prop:lattice-theory-constructing_a_rectangular_tesselating_region} where $\Lambda$ is a 2 dimensional lattice with generator matrix having columns $[1, 0.2]^\prime$ and $[0.2, 1]^\prime$.
@@
 \end{lemma}
-\begin{IEEEproof}
+\begin{proof}
 See~\citep[p. 2]{cahen_integer-valued_1997} or~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}. 
-\end{IEEEproof}
-% \begin{IEEEproof}
+\end{proof}
+% \begin{proof}
 % Note that $x(x-1)(x-2)\dots(x-k+1)$ is divisible by all integers $1,2,\dots,k$ and so $p_k$ takes integer values for all $x\in\ints$.
@@
 The proof follows by induction because $f(0) = c_0 \in \ints$.
-% \end{IEEEproof}
+% \end{proof}
 
@@
 	\centering
-		\includegraphics[width=\linewidth]{plots/circstatfig-1.mps}
+		\includegraphics[width=0.8\linewidth]{plots/circstatfig-1.mps}
 		\caption{The first order polynomials $\tfrac{1}{10}(3 + 8t)$ (solid) and $\tfrac{1}{10}(33 - 2t)$ (dashed line).}
@@
 
-%\begin{figure}[tp]
-%	\centering
-%		\includegraphics[width=\linewidth]{plots/circstatfigquad-1.mps}
-%		\caption{The quadratic polynomials $\tfrac{1}{10} (15 - 15 t + 4 t^2)$ (solid line) and $\tfrac{1}{10}(25 -  t^2)$  (dashed line).}
-%		\label{fig:circstatplot_quad}
-%\end{figure}
-
 \begin{figure}[tp]
 	\centering
-		\includegraphics[width=\linewidth]{plots/circstatfigcube-1.mps}
+		\includegraphics[width=0.8\linewidth]{plots/circstatfigquad-1.mps}
+		\caption{The quadratic polynomials $\tfrac{1}{10} (15 - 15 t + 4 t^2)$ (solid line) and $\tfrac{1}{10}(25 -  t^2)$  (dashed line).}
+		\label{fig:circstatplot_quad}
+\end{figure}
+
+\begin{figure}[tp]
+	\centering
+		\includegraphics[width=0.8\linewidth]{plots/circstatfigcube-1.mps}
 		\caption{The cubic polynomials $\tfrac{1}{160} (174 + 85 t - 118 t^2 + 40 t^3)$ (solid line) and $\tfrac{1}{48} (84 + 19 t + 12 t^2 - 4 t^3)$  (dashed line).}
@@
 
-%Finally note that $\Cbf$ is a \emph{Hilbert matrix}\index{Hilbert matrix} and the elements of the inverse $\Cbf^{-1}$ are given by
- %\[
- %C^{-1}_{ik} =\frac{i+k+1}{(-1)^{i+k}}{m+i+1 \choose m-k}{m+k+1 \choose m-i}{i+k \choose i}^2.
- %\]
-%Hilbert matrices are particularly ill conditioned and difficult to numerically invert.  This problem can be avoided using the above formula. 
-%I MIGHT HAVE AN ERROR IN THIS FORMULA!
-
 \section{Proof of strong consistency}\label{sec:strongconstproof}
@@
  \end{lemma}
-\begin{IEEEproof}
+\begin{proof}
 Consider a rectangular grid of points spaced over the identifiable region $B$.  We use $\lambdabf[\rbf]$, where $\rbf \in \ints^{m+1}$, to denote the grid point 
@@
 and for sufficiently large $N$, we have $\epsilon/2 + \frac{2(m+1)}{N^b} < \epsilon$ completing the proof.  It remains to prove Lemmas~\ref{lem:supVjk} and~\ref{lem:supBVn}.
-\end{IEEEproof}
+\end{proof}
 
@@
 \end{lemma}
- \begin{IEEEproof}
+ \begin{proof}
 Fix $\lambdabf$ and write
@@
 where $c$ is any real number less than $8$, since $|G| = N^{(m+1)(2b + m)/2}$ is polynomial in $N$.
-\end{IEEEproof}
+\end{proof}
 
@@
 \end{lemma}
-\begin{IEEEproof}
+\begin{proof}
 Since $\abs{\delta} \leq \abs{n + \delta}$ for all $\delta \in [-\nicefrac{1}{2}, \nicefrac{1}{2})$ and $n \in \ints$, the result will follow if we can show that it holds when both $x$ and $\delta$ are in $[-\nicefrac{1}{2}, \nicefrac{1}{2})$.
@@
 % and the lemma is true because $\delta(1 - \delta) \leq |\delta|$ since $0 \leq \delta < \tfrac{1}{4}$.
-\end{IEEEproof}
+\end{proof}
 
@@
  \end{lemma}
- \begin{IEEEproof}
+ \begin{proof}
 %The proof we give is not just almost surely but \emph{surely} as the convergence will be shown to occur irrespective of the values of the $\Phi_n$. 
@@
 \]
- \end{IEEEproof}
+ \end{proof}
 
@@
 \end{lemma}
-\begin{IEEEproof}
+\begin{proof}
 Let $z$ be the polynomial
@@
 Therefore $V_N$ is minimised uniquely at $\zerobf$, at which point it takes the value $\sigma^2$.
-\end{IEEEproof}
+\end{proof}
 
@@
 \end{lemma}
-\begin{IEEEproof}
+\begin{proof}
 By definition
@@
 %where the third line follows from Lemma~\ref{lem:ABprob}.
-\end{IEEEproof}
+\end{proof}
 
@@
 \end{corollary}
-\begin{IEEEproof}
+\begin{proof}
 From Lemma~\ref{lem:ESNconv}, for any $\epsilon > 0$ and $0 < c < \tfrac{1}{2}$,
@@
 The corollary follows from the Borel-Cantelli lemma.
-\end{IEEEproof}
+\end{proof}
 
@@
 % \end{lemma}
-% \begin{IEEEproof}
+% \begin{proof}
 % Write 
@@
 % Subtituting this into~\eqref{eq:A+B>epsilon} gives the required result.
-% \end{IEEEproof}
+% \end{proof}
 
@@
 \end{lemma}
-\begin{IEEEproof}
+\begin{proof}
 The proof follows because there are at most $m+1$ integers, $n$, such that $n+ih = r$ for some $i \in \{0,1,\dots,m\}$.
 That is, there are at most $m+1$ arithmetic progressions of type $n, n+h, \dots n+mh$ that contain $r$.
-\end{IEEEproof}
+\end{proof}
 
@@
  \end{lemma}
- \begin{IEEEproof}
+ \begin{proof}
  Note that $|A(h,W)| = N - mh$.  The proof follows by starting with $A(h,W)$ and applying Lemma~\ref{lem:S(h,G/r)size} 
@@
 times. That is, $K$ can be constructed by removing $N - |K|$ elements from $W$ and this removes at most $(N - |K|)(m+1)$ arithmetic progressions from $A(h,W)$.
- \end{IEEEproof}
+ \end{proof}
 
@@
  \end{corollary}
- \begin{IEEEproof}
+ \begin{proof}
  By substituting the bounds $|K| > \frac{2m+1}{2m+2}N$ and $h \leq\frac{N}{2m}$ into the inequality from Lemma~\ref{lem:S(h,K)size} we immediately obtain $|A(h,K)| > 0$.
- \end{IEEEproof}
+ \end{proof}
 
@@
  \end{lemma}
- \begin{IEEEproof}
+ \begin{proof}
  If $\delta > \tfrac{1}{2r}$ the proof is trivial as $\left|\fracpart{\sum_{n=1}^r{a_n}}\right| \leq \tfrac{1}{2}$ for all $a_n \in \reals$.
@@
  by the triangle inequality.
- \end{IEEEproof}
+ \end{proof}
 
@@
 \end{lemma}
-\begin{IEEEproof}
+\begin{proof}
 By supposition, for some $n \in \ints$,
@@
 % from which it follows that $\left|\fracpart{a}\right| < \delta + \left|\epsilon\right|$.
-\end{IEEEproof}
+\end{proof}
 
@@
 \end{lemma}
-\begin{IEEEproof}
+\begin{proof}
 The function $\fracpart{x}^2$ is differentiable everywhere except if $\fracpart{x} \neq -\nicefrac{1}{2}$.  Recalling that
@@
 That each of the partial derivatives is equal to zero at $\widehat{\psibf}_N$ follows immediately from the fact that $\widehat{\psibf}_N$ is a minimiser of $T_N$.
-\end{IEEEproof}
+\end{proof}
 
@@
 \end{lemma}
-\begin{IEEEproof}
+\begin{proof}
 To simplify our notation let 
@@
 The proof follows.
-\end{IEEEproof}
+\end{proof}
 
@@
 \end{lemma}
-\begin{IEEEproof}
+\begin{proof}
 Care must be taken since $\widehat{\psibf}_N$ depends on the sequence $\{ \Phi_n \}$.  For $n = 1, \dots, N$ and positive $N$, let
@@
 \end{lemma}
-\begin{IEEEproof}
+\begin{proof}
 If $\abs{x} < 1$, then
@@
 % The proof follows by substituting this into~\eqref{eq:HellNre}.
-\end{IEEEproof}
+\end{proof}
 
@@
 \end{lemma}
-\begin{IEEEproof}
+\begin{proof}
 For any constant vector $\alphabf$, let
@@
 % as $N \rightarrow \infty$, again using~\eqref{eq:ClkNHilbert}.  The proof follows.
-\end{IEEEproof}
-
-\begin{figure*}[tb] 
-   	\centering 
-  		\includegraphics{code/lseplot-1.mps} 
-   		\caption{Performance of the LSE and LSU estimator for a polynomial phase signal of order two, i.e.
-$m=2$.
-(Left) Mean square error (MSE) in the second order parameter $\mu_2$ for $N=10,50$ and $200$ versus the variance $\sigma_c^2$ of the $X_1, \dots, X_N$.
-The $X_1, \dots, X_N$ are independent and identically distributed (i.i.d.) complex Gaussian random variables.
-(Right) MSE in second order parameter $\mu_2$ versus the variance $\sigma^2$ of the $\Phi_1, \dots, \Phi_N$.
-The $\Phi_1, \dots, \Phi_N$ are i.i.d.
-uniform random variables.}
-   		\label{plot:lsetest} 
-  \end{figure*}
+\end{proof}
 
@@
 
-\begin{figure*}[p] 
+\begin{figure}[tp] 
    	\centering 
+  		%\includegraphics[width=\linewidth]{code/lseplot-1.mps} 
+               \includegraphics{code/lseplot-1.mps}
+   		\caption{Performance of the LSE and LSU estimator for a polynomial phase signal of order two, i.e.
+$m=2$.
+(Left) Mean square error (MSE) in the second order parameter $\mu_2$ for $N=10,50$ and $200$ versus the variance $\sigma_c^2$ of the $X_1, \dots, X_N$.
+The $X_1, \dots, X_N$ are independent and identically distributed (i.i.d.) complex Gaussian random variables.
+(Right) MSE in second order parameter $\mu_2$ versus the variance $\sigma^2$ of the $\Phi_1, \dots, \Phi_N$.
+The $\Phi_1, \dots, \Phi_N$ are i.i.d.
+uniform random variables.}
+   		\label{plot:lsetest} 
+  \end{figure}
+
+\begin{figure}[p] 
+   	\centering 
+  		%\includegraphics[width=\linewidth]{code/gaussianplot4-1.mps}
                 \includegraphics{code/gaussianplot4-1.mps}
@@
   		\label{plot:polyest} 
- \end{figure*}
+ \end{figure}
 
@@
 
-%\section{Discussion}
-
 \section{Conclusion} \label{sec:conclusion}
@@
 
-%\begin{figure}[p] 
-%   	\centering 
-%  		\includegraphics[width=\linewidth]{polyfig/polyphase4plot.4} 
-%  		\caption{MSE in the third order coefficient $\mu_3$ versus $\var{X_n} = \sigma^2$. The true coefficients are uniformly spread in the identifiable region.} 
-%  		\label{plot:polyest_projnorm_mu3} 
-% \end{figure}
-
-% \begin{figure}[p] 
-%  	\centering 
-% 		\includegraphics[width=\linewidth]{polyfig/polyphase4plot_unfair.4} 
-% 		\caption{MSE in the third order coefficient $\mu_3$ versus $\var{X_n} = \sigma_c^2$. The coefficient have been restricted for Kitchen's estimator and the DPT.} 
-% 		\label{plot:polyest_projnorm_mu3_unfair} 
-%\end{figure} 
-% 
-% \begin{figure}[p] 
-%  	\centering 
-% 		\includegraphics[width=\linewidth]{polyfig/polyphase4plot_fair.4} 
-% 		\caption{MSE in the third order coefficient $\mu_3$ versus $\var{X_n} = \sigma_c^2$. The DPT estimator runs at the higher sampling rate $\delta$ so that the volumes $V_{DPT}(\delta) = V_m$.} 
-% 		\label{plot:polyest_projnorm_mu3_fair} 
-%\end{figure}
-
-%\addtolength{\parskip}{-0.2cm} 
-%\bibliographystyle{../../bib/IEEEbib} 
-\small 
-\bibliography{bib}
-
-\normalsize
 \appendix
@@
 \end{lemma}
-\begin{IEEEproof}
+\begin{proof}
 Let $\{\Phi_n^\prime\}$ be a sequence of random variables distributed identically to $\{\Phi_n\}$ and independent of both $\{\Phi_n\}$ and $\{g_n\}$.
@@
 as required.
-\end{IEEEproof}
+\end{proof}
 
@@
 \end{lemma}
-\begin{IEEEproof}
+\begin{proof}
 Without loss of generality, assume that $\epsilon < \frac{1}{m+1}$.  Lemma~\ref{lem:chaining2} shows that
@@
 Choosing $\epsilon = \kappa^2/(K_1^2 K_2)$ completes the proof.  It remains to prove Lemma~\ref{lem:chaining2}.
-\end{IEEEproof}
+\end{proof}
 
@@
 \end{lemma}
-\begin{IEEEproof}
+\begin{proof}
 Let
@@
 and $\log(\cdot)$ is the natural logarithm.  The lemma holds with $K_1 =  \sum_{i=1}^{\infty}2^{-i/2}\sqrt{ i A_1 + A_2 }$.
-\end{IEEEproof}
+\end{proof}
 
@@
 \end{lemma}
-\begin{IEEEproof}
+\begin{proof}
 Recall that
@@
 Summing the terms in this inequality over $n = 1, \dots, N$ and using~\eqref{eq:Cedefn} gives~\eqref{eq:sumf2absfCe}.
-\end{IEEEproof}
+\end{proof}
 
@@
 \end{lemma}
-\begin{IEEEproof}
+\begin{proof}
 Let $b_k$ be a function that maps each $\psibf \in T_\epsilon(k)$ to $b_k(\psibf) \in T_\epsilon(k-1)$ such that 
@@
 \end{lemma}
-\begin{IEEEproof}
+\begin{proof}
 This result is well known, see for example~\cite[Section 3]{Pollard_asymp_empi_proc_1989}  
-\end{IEEEproof}
+\end{proof}
 
@@
 \end{corollary}
-\begin{IEEEproof}
+\begin{proof}
 Let $C$ be the set of generalised simplices constructed according to Theorem~\ref{thm:epscutting}.
@@
 follows from the triangle inequality.
-\end{IEEEproof}
+\end{proof}
 
@@
 \end{corollary}
-\begin{IEEEproof}
+\begin{proof}
 Let $C$ be the set of generalised simplices constructed according to Theorem~\ref{thm:epscutting} and let $C_B$ be the subset of those indices that intersect $B$.
@@
 The proof now follows similarly to Corollary~\ref{cor:epscutting}.
-\end{IEEEproof}
+\end{proof}
 
@@
 
-\begin{IEEEproof} (Lemma~\ref{lem:metricentropy})
+\begin{proof} (Lemma~\ref{lem:metricentropy})
 Put 
@@
 Putting $r = 2^k$ and choosing $T_\epsilon(k) = T_B$ completes the proof.
-\end{IEEEproof}
-
-%\begin{figure}[p] 
-%   	\centering 
-%  		\includegraphics[width=\linewidth]{polyfig/polyphase4plot.4} 
-%  		\caption{MSE in the third order coefficient $\mu_3$ versus $\var{X_n} = \sigma^2$. The true coefficients are uniformly spread in the identifiable region.} 
-%  		\label{plot:polyest_projnorm_mu3} 
-% \end{figure}
-
-% \begin{figure}[p] 
-%  	\centering 
-% 		\includegraphics[width=\linewidth]{polyfig/polyphase4plot_unfair.4} 
-% 		\caption{MSE in the third order coefficient $\mu_3$ versus $\var{X_n} = \sigma_c^2$. The coefficient have been restricted for Kitchen's estimator and the DPT.} 
-% 		\label{plot:polyest_projnorm_mu3_unfair} 
-%\end{figure} 
-% 
-% \begin{figure}[p] 
-%  	\centering 
-% 		\includegraphics[width=\linewidth]{polyfig/polyphase4plot_fair.4} 
-% 		\caption{MSE in the third order coefficient $\mu_3$ versus $\var{X_n} = \sigma_c^2$. The DPT estimator runs at the higher sampling rate $\delta$ so that the volumes $V_{DPT}(\delta) = V_m$.} 
-% 		\label{plot:polyest_projnorm_mu3_fair} 
-%\end{figure}
+\end{proof}
+
+\bibliographystyle{acmtrans-ims}
+\bibliography{bib}
 

### 2012-10-13 paper.tex
modifed some of the plots.  I like the look of this, though it probably needs an edit for the purpose of a non signal processing audience.

@@
 
-\begin{figure}[tp]
+\begin{figure}[t]
 	\centering
@@
 \begin{align*}
-s(n) &= e^{2\pi j y(n)} =  e^{2\pi j y(n)} e^{2\pi j p(n)} \\
-&= e^{2\pi j (y(n) + p(n))} = e^{2\pi j z(n)} = r(n)
+s(n) = e^{2\pi j y(n)} =  e^{2\pi j y(n)} e^{2\pi j p(n)} = e^{2\pi j (y(n) + p(n))} = e^{2\pi j z(n)} = r(n)
 \end{align*}
@@
 
-\begin{figure}[tp]
+\begin{figure}[p]
 	\centering
-		\includegraphics[width=0.8\linewidth]{plots/circstatfig-1.mps}
+		\includegraphics[width=0.85\linewidth]{plots/circstatfig-1.mps}
 		\caption{The first order polynomials $\tfrac{1}{10}(3 + 8t)$ (solid) and $\tfrac{1}{10}(33 - 2t)$ (dashed line).}
@@
 
-\begin{figure}[tp]
+\begin{figure}[p]
 	\centering
-		\includegraphics[width=0.8\linewidth]{plots/circstatfigquad-1.mps}
+		\includegraphics[width=0.85\linewidth]{plots/circstatfigquad-1.mps}
 		\caption{The quadratic polynomials $\tfrac{1}{10} (15 - 15 t + 4 t^2)$ (solid line) and $\tfrac{1}{10}(25 -  t^2)$  (dashed line).}
@@
 
-\begin{figure}[tp]
+\begin{figure}[p]
 	\centering
-		\includegraphics[width=0.8\linewidth]{plots/circstatfigcube-1.mps}
+		\includegraphics[width=0.85\linewidth]{plots/circstatfigcube-1.mps}
 		\caption{The cubic polynomials $\tfrac{1}{160} (174 + 85 t - 118 t^2 + 40 t^3)$ (solid line) and $\tfrac{1}{48} (84 + 19 t + 12 t^2 - 4 t^3)$  (dashed line).}
@@
 \begin{align*}
-\fracpart{\sum_{k=0}^{m}\lambda_k n^k } &= \fracpart{\sum_{k=0}^{m}(\tilde{\mu}_k - \mu_k - p_k) n^k } \\
-&= \fracpart{\sum_{k=0}^{m}(\tilde{\mu}_k - \mu_k) n^k }.
+\fracpart{\sum_{k=0}^{m}\lambda_k n^k } = \fracpart{\sum_{k=0}^{m}(\tilde{\mu}_k - \mu_k - p_k) n^k } = \fracpart{\sum_{k=0}^{m}(\tilde{\mu}_k - \mu_k) n^k }.
 \end{align*}
@@
 \begin{align}
- &\sup_{\lambdabf \in B}\sabs{ D_N(\lambdabf) }  \nonumber \\
-&= \sup_{\rbf \in G}\sup_{ \lambdabf \in B[\rbf]}\sabs{ D_N(\lambdabf[\rbf]) + D_N(\lambdabf) - D_N(\lambdabf[\rbf])  } \nonumber \\
+ \sup_{\lambdabf \in B}\sabs{ D_N(\lambdabf) }  &= \sup_{\rbf \in G}\sup_{ \lambdabf \in B[\rbf]}\sabs{ D_N(\lambdabf[\rbf]) + D_N(\lambdabf) - D_N(\lambdabf[\rbf])  } \nonumber \\
  &\leq \sup_{\rbf \in G}\sabs{ D_N(\lambdabf[\rbf])} + \sup_{\rbf \in G}\sup_{\lambdabf\in B[\rbf]}\sabs{ D_N(\lambdabf) - D_N(\lambdabf[\rbf])  }. \label{eq:gridanddensespace}
@@
  \begin{proof}
-Fix $\lambdabf$ and write
- \[
- D_N(\lambdabf) = \bar{Z} = \frac{1}{N}  \sum_{n=1}^{N}Z_{n},
- \]
- where
+Fix $\lambdabf$ and write $D_N(\lambdabf) = \bar{Z} = \frac{1}{N}  \sum_{n=1}^{N}Z_{n}$, where
  \[
@@
  \prob\cubr{  \sup_{\rbf\in G}\left\vert D_N(\lambdabf[\rbf]) \right\vert >\epsilon }  &\leq \sum_{\rbf \in G} \prob\scubr{  \left\vert D_N(\lambdabf[\rbf])  \right\vert >\epsilon } \nonumber \\
-&= 2 |G| e^{ - 8 \epsilon^2 N} \nonumber \\
-&= O(e^{-c \epsilon^2 N}), \label{eq:betaprobconv}
+&= 2 |G| e^{ - 8 \epsilon^2 N} = O(e^{-c \epsilon^2 N}), \label{eq:betaprobconv}
 \end{align*}
@@
 
- \begin{lemma}\label{lem:supBVn}
-For all $N$,
-\[
- \sup_{\rbf\in G}\sup_{ \lambdabf  \in B[\rbf] }\left\vert D_N\left(  \lambdabf \right) - D_N\left(  \lambdabf[\rbf] \right)\right\vert < 2\frac{m+1}{N^b}.
- \]
+ \begin{lemma}\label{lem:supBVn} $\sup_{\rbf\in G}\sup_{ \lambdabf  \in B[\rbf] }\left\vert D_N\left(  \lambdabf \right) - D_N\left(  \lambdabf[\rbf] \right)\right\vert < 2\frac{m+1}{N^b}$ for all $N$.
  \end{lemma}
@@
 \] 
-where $\lambda[\rbf]_k$ denotes the $k$th element of the grid point $\lambdabf[\rbf]$. For $\lambdabf \in B[\rbf]$,
-\[
-b_n = a_n + \delta_n,
-\]
+where $\lambda[\rbf]_k$ denotes the $k$th element of the grid point $\lambdabf[\rbf]$. For $\lambdabf \in B[\rbf]$ we have $b_n = a_n + \delta_n$,
 where 
@@
 \] 
-Taking fractional parts, squaring, and applying Lemma~\ref{lem:boundedsquarefracparts} we see that, for all real numbers $x$,
-\[
- -|\delta_n| \leq \fracpart{x + b_n}^2 -  \fracpart{x + a_n}^2  \leq |\delta_n|,
-\]
-and $|\fracpart{x + b_n}^2 -  \fracpart{x + a_n}^2| \leq \tfrac{m+1}{N^b}$. Now
+From Lemma~\ref{lem:boundedsquarefracparts} it follows that $-|\delta_n| \leq \fracpart{x + b_n}^2 -  \fracpart{x + a_n}^2  \leq |\delta_n|$,
+and consequently $|\fracpart{x + b_n}^2 -  \fracpart{x + a_n}^2| \leq \tfrac{m+1}{N^b}$ for all real numbers $x$. Now
 \[
@@
  \begin{align*}
- \sabs{ D_N(\lambdabf) - &D_N(\lambdabf[\rbf]) } \\
-&= \sabs{ S_N(\lambdabf) - S_N(\lambdabf[\rbf]) + V_N(\lambdabf) - V_N(\lambdabf[\rbf]) } \\
+ \sabs{ D_N(\lambdabf) - D_N(\lambdabf[\rbf]) } &= \sabs{ S_N(\lambdabf) - S_N(\lambdabf[\rbf]) + V_N(\lambdabf) - V_N(\lambdabf[\rbf]) } \\
 &\leq \sabs{ S_N(\lambdabf) - S_N(\lambdabf[\rbf]) \vert + \vert V_N(\lambdabf) - V_N(\lambdabf[\rbf]) },  \\
@@
 \begin{align*}
-V_N(\lambdabf) &= \frac{1}{N}\expect\sum_{n=1}^{N}\fracpart{ \Phi_n + \sum_{k=0}^m{\lambda_k n^k} }^2 \\
-&= \frac{1}{N}\sum_{n=1}^{N}\expect\fracpart{ \Phi_n + \fracpart{z(n)} }^2.
+V_N(\lambdabf) = \frac{1}{N}\expect\sum_{n=1}^{N}\fracpart{ \Phi_n + \sum_{k=0}^m{\lambda_k n^k} }^2 = \frac{1}{N}\sum_{n=1}^{N}\expect\fracpart{ \Phi_n + \fracpart{z(n)} }^2.
 \end{align*}
@@
 \]
-and therefore
-\[
- 0 \leq S_{N}(\zerobf) - S_N(\widehat{\lambdabf}_N).
-\]
-Also, because $V_N$ is minimised at $\zerobf$,
-\[
- 0 \leq V_N(\widehat{\lambdabf}_N) - V_N(\zerobf).
-\]
-Thus,
+and therefore $0 \leq S_{N}(\zerobf) - S_N(\widehat{\lambdabf}_N)$.  Also, because $V_N$ is minimised at $\zerobf$, it follows that $0 \leq V_N(\widehat{\lambdabf}_N) - V_N(\zerobf)$.  Thus,
 \begin{align*}
@@
 \begin{align*}
-\prob&\left\{ \sabs{V_N(\widehat{\lambdabf}_N) - V_N(\zerobf))} > \epsilon \right\} \\
-&\leq \prob\left\{ \sabs{ D_N(\widehat{\lambdabf}_N) } + \sabs{ D_N(\zerobf) } > \epsilon \right\} \\
+\prob\left\{ \sabs{V_N(\widehat{\lambdabf}_N) - V_N(\zerobf))} > \epsilon \right\} &\leq \prob\left\{ \sabs{ D_N(\widehat{\lambdabf}_N) } + \sabs{ D_N(\zerobf) } > \epsilon \right\} \\
 &\leq \prob\left\{ \sabs{ D_N(\widehat{\lambdabf}_N) } > \frac{\epsilon}{2}\right\} + \prob\left\{ \sabs{ D_N(\zerobf) } > \frac{\epsilon}{2} \right\} \\
@@
 \begin{align*}
-\Delta_h^2 d(n) &= \Delta_h d(n+h) - \Delta_h d(n) 
-\\ &= d(n+2h) - 2d(n+h) + d(n)
+\Delta_h^2 d(n) = \Delta_h d(n+h) - \Delta_h d(n) = d(n+2h) - 2d(n+h) + d(n)
 \end{align*}
@@
 \begin{equation}\label{eq:mthdiffformula}
-\begin{split}
-\Delta_h^r d(n) &= \Delta_h^{r-1} d(n+h) - \Delta_h^{r-1} d(n)
-\\ &= \sum_{k=0}^{r}\binom{r}{k}(-1)^{r-k}d(n+kh)
-\end{split}
-\end{equation}
-denote the $r$th difference with interval $h$. It is well known that 
-\[
-\sum_{k=0}^{r}\binom{r}{k} = 2^r.
-\] 
-So $\Delta_h^r d(n)$ can be represented by adding and subtracting the 
+\Delta_h^r d(n) = \Delta_h^{r-1} d(n+h) - \Delta_h^{r-1} d(n) = \sum_{k=0}^{r}\binom{r}{k}(-1)^{r-k}d(n+kh)
+\end{equation}
+denote the $r$th difference with interval $h$. Since $\sum_{k=0}^{r}\binom{r}{k} = 2^r$ it follows that $\Delta_h^r d(n)$ can be represented by adding and subtracting the 
 \[
@@
 \]
-so that
-\[
-V_N(\lambdabf_N) - \sigma^2 = \frac{1}{N}\sum_{n=1}^{N} g\left( \fracpart{z_N(n)} \right) \rightarrow 0.
-\]
+so that $V_N(\lambdabf_N) - \sigma^2 = \frac{1}{N}\sum_{n=1}^{N} g\left( \fracpart{z_N(n)} \right) \rightarrow 0$.
 %At this stage we can \emph{see} the proof.
@@
 \]
-and define the set 
-\[
-K_{N}=\left\{  n\leq N \mid \sabs{\fracpart{z_N(n)}} < \delta \right\}.
-\]
+and define the set $K_{N}=\left\{  n\leq N \mid \sabs{\fracpart{z_N(n)}} < \delta \right\}$.
 There exists $N_0$ such that for all $N > N_0$ the number of elements in $K_N$ is at least $cN$.
@@
 \begin{align*}
-V_N(\lambdabf_N) - \sigma^2 &= \frac{1}{N}\sum_{n=1}^{N} g\left( \fracpart{z_N(n)} \right) \\
-&\geq \frac{1}{N}\sum_{n\in K_N} \gamma = (1-c)\gamma,
+V_N(\lambdabf_N) - \sigma^2 = \frac{1}{N}\sum_{n=1}^{N} g\left( \fracpart{z_N(n)} \right) \geq \frac{1}{N}\sum_{n\in K_N} \gamma = (1-c)\gamma,
 \end{align*}
@@
  \begin{equation}\label{eq:zrbound}
-\begin{split}
- \left|\fracpart{  \Delta_h^r z_N(n')}\right| &= \left|\fracpart{  \Delta_h^r\gamma(n') + \Delta_h^r z_r(n') }\right| \\
-&= |\fracpart{ \epsilon + h^r r! \lambda_{r,N} }| < 2^r\delta,
-\end{split}
+\left|\fracpart{  \Delta_h^r z_N(n')}\right| = \left|\fracpart{  \Delta_h^r\gamma(n') + \Delta_h^r z_r(n') }\right| = |\fracpart{ \epsilon + h^r r! \lambda_{r,N} }| < 2^r\delta,
  \end{equation}
@@
 \begin{proof}
-By supposition, for some $n \in \ints$,
-\[
-n - \delta < a + \epsilon < n + \delta.
-\]
-Since 
-\[
--\delta - \epsilon > -\tfrac{1}{2} \qquad \text{and} \qquad \delta - \epsilon < \tfrac{1}{2},
-\]
-it follows that
+By supposition $n - \delta < a + \epsilon < n + \delta$ for some $n \in \ints$.  Since $-\delta - \epsilon > -\tfrac{1}{2}$ and $\delta - \epsilon < \tfrac{1}{2}$, it follows that
 \[
@@
 \end{equation}
-we have, for each $\ell = 0, \dots, m$,
-\begin{align*}
-0 & = \frac{\partial T_N}{\partial \psi_\ell}(\widehat{\psibf}_N)\\
-&= \frac{2}{N}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\left( \Phi_n - W_{n}+ \sum_{k=0}^{m}(\tfrac{n}{N})^k \widehat{\psi}_{k,N}  \right),
-\end{align*}
-so that
-\[
-D_{\ell,N} = K_{\ell,N},
-\]
-where
+we have,
+\begin{align*}
+0  = \frac{\partial T_N}{\partial \psi_\ell}(\widehat{\psibf}_N) = \frac{2}{N}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\left( \Phi_n - W_{n}+ \sum_{k=0}^{m}(\tfrac{n}{N})^k \widehat{\psi}_{k,N}  \right),
+\end{align*}
+for each $\ell = 0, \dots, m$, so that $D_{\ell,N} = K_{\ell,N}$, where
 \[
@@
 \right]
-\qquad \text{and} \qquad 
+%\qquad \text{and} \qquad 
+=
 \dbf_N = \left[
@@
 \begin{lemma}\label{lem:diffathatpsi}
-For all $\ell = 0, \dots, m$ the partial derivatives $\frac{\partial T_N}{\partial \psi_\ell}$ exist, and are equal to zero, at the minimiser $\widehat{\psibf}_N$.  That is,
-\[
-\frac{\partial T_N}{\partial \psi_\ell}(\widehat{\psibf}_N) = 0 \qquad \text{for each $\ell = 0, \dots, m$.}
-\]
+For all $\ell = 0, \dots, m$ the partial derivatives $\frac{\partial T_N}{\partial \psi_\ell}$ exist, and are equal to zero, at the minimiser $\widehat{\psibf}_N$.
+That is $\frac{\partial T_N}{\partial \psi_\ell}(\widehat{\psibf}_N) = 0$ for each $\ell = 0, \dots, m$.
 \end{lemma}
@@
 \begin{proof}
-To simplify our notation let 
-\[
-B_n = \Phi_n + \sum_{k=1}^m (\nicefrac{n}{N})^k \widehat{\psi}_{k,N}
-\]
-so that we now require to prove $\abs{\fracpart{B_n + \widehat{\psi}_{0,N}}} \leq \frac{1}{2} - \frac{1}{2N}$
-for all $n = 1, \dots, N$.  From~\eqref{eq:Wn}, $W_n = \round{B_n + \widehat{\psi}_{0,N}}$, and 
-\begin{align*}
-T_N(\widehat{\psibf}_N) &= \frac{1}{N}\sum_{n=1}^N \fracpart{B_n + \widehat{\psi}_{0,N}}^2
-\\ &= \frac{1}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N} - W_n)^2.
+To simplify our notation let $B_n = \Phi_n + \sum_{k=1}^m (\nicefrac{n}{N})^k \widehat{\psi}_{k,N}$ so that we now require to prove $\abs{\fracpart{B_n + \widehat{\psi}_{0,N}}} \leq \frac{1}{2} - \frac{1}{2N}$ for all $n = 1, \dots, N$.
+From~\eqref{eq:Wn}, $W_n = \round{B_n + \widehat{\psi}_{0,N}}$, and
+\begin{align*}
+T_N(\widehat{\psibf}_N) = \frac{1}{N}\sum_{n=1}^N \fracpart{B_n + \widehat{\psi}_{0,N}}^2= \frac{1}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N} - W_n)^2.
 \end{align*}
@@
 \begin{align}
-(B_k + &\phi - F_k)^2 = (B_k + \phi - W_k - 1)^2 \nonumber \\
+(B_k + \phi - F_k)^2 &= (B_k + \phi - W_k - 1)^2 \nonumber \\
 &= (B_k + \phi - W_k)^2 - 2(B_k + \phi - W_k) + 1 \nonumber \\
@@
 \end{align}
-where the inequality in the last line follows from~\eqref{eq:philamfraccontra}. Let
-\[
-\bbf = \left[
-\begin{array}
-[c]{cccc}%
-\phi & \widehat{\psi}_{1,N} & \cdots & \widehat{\psi}_{m,N}
-\end{array}
-\right]^\prime
-\]
+where the inequality in the last line follows from~\eqref{eq:philamfraccontra}. Let $\bbf = [ \phi,  \widehat{\psi}_{1,N} , \dots , \widehat{\psi}_{m,N}]$
+% \[
+% \bbf = \left[
+% \begin{array}
+% [c]{cccc}%
+% \phi & \widehat{\psi}_{1,N} & \cdots & \widehat{\psi}_{m,N}
+% \end{array}
+% \right]^\prime
+% \]
 be the vector of length $m+1$ with components $b_0 = \phi$ and $b_\ell = \widehat{\psi}_{\ell,N}$ for $\ell = 1 , \dots m$.  Now,
@@
 \begin{align*}
-N &T_N(\bbf) < - \frac{1}{N} + \sum_{n=1}^N (B_n + \phi  - W_n)^2 \\
+N T_N(\bbf) &< - \frac{1}{N} + \sum_{n=1}^N (B_n + \phi  - W_n)^2 \\
 &= - \frac{1}{N} + \sum_{n=1}^N (B_n + \widehat{\psi}_{0,N} + \frac{1}{N}  - W_n)^2 \\
@@
 \end{equation}
-and put
-\[
-q_{n}(x) = \round{\Phi_n + x}, \qquad Q(x) = \expect q_{n}(x) =  \expect q_{1}(x).
-\]
-Let
+and put $q_{n}(x) = \round{\Phi_n + x}$ and $Q(x) = \expect q_{n}(x) =  \expect q_{1}(x)$.  Let
 \begin{equation}\label{eq:GNdef}
@@
 \end{equation}
-Observe that $G_N$ depends on $\ell$ and we could write $G_{\ell,N}$ but have suppressed the subscript $\ell$ for notational simplicity.  Now $W_n$ from~\eqref{eq:Wn} can be written as
-\[
-W_n = \round{\Phi_n + \widehat{p}_{nN}} = q_{n}(\widehat{p}_{nN})
-\]
-and $K_{\ell,N}$ from~\eqref{eq:KellN} can be written as
+Observe that $G_N$ depends on $\ell$ and we could write $G_{\ell,N}$ but have suppressed the subscript $\ell$ for notational simplicity.
+Now $W_n$ from~\eqref{eq:Wn} can be written as $W_n = \round{\Phi_n + \widehat{p}_{nN}} = q_{n}(\widehat{p}_{nN})$ and $K_{\ell,N}$ from~\eqref{eq:KellN} can be written as
 \begin{align*}
@@
 \]
-for all positive integers $N$, where $\|\psibf\|_\infty = \sup_{k} \abs{\psi_k}$.  Since $\widehat{\psibf}_N$ converges almost surely to zero, it follows that for any $\epsilon > 0$,
-\[
-\lim_{N\rightarrow\infty}\prob\left\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} = 0
-\] 
-and therefore $\prob\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \} < \nu$ for all sufficiently large $N$.  Now
-\begin{align*}
-  \prob&\left\{\abs{ G_N(\widehat{\psibf}_N) } > \delta \right\} \\
-&= \prob\left\{ \abs{G_N(\widehat{\psibf}_N)} > \delta \;, \; \|\widehat{\psibf}_N\|_{\infty} < \epsilon \right\} \\
-&\hspace{0.7cm} + \prob\left\{ \abs{G_N(\widehat{\psibf}_N)} > \delta  \;, \; \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} \\
+for all positive integers $N$, where $\|\psibf\|_\infty = \sup_{k} \abs{\psi_k}$.
+Since $\widehat{\psibf}_N$ converges almost surely to zero, it follows that $\lim_{N\rightarrow\infty}\prob\left\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} = 0$ for any $\epsilon > 0$, and therefore $\prob\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \} < \nu$ for all sufficiently large $N$.
+Now
+\begin{align*}
+  \prob\left\{\abs{ G_N(\widehat{\psibf}_N) } > \delta \right\} &= \prob\left\{ \abs{G_N(\widehat{\psibf}_N)} > \delta \;, \; \|\widehat{\psibf}_N\|_{\infty} < \epsilon \right\} \\
+&\hspace{1cm} + \prob\left\{ \abs{G_N(\widehat{\psibf}_N)} > \delta  \;, \; \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} \\
 &\leq \prob\left\{  \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf) } > \delta \right\} + \prob\left\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} \\
@@
 \begin{lemma}\label{lem:sumpNhilb}
-With $H_{\ell,N}$ defined in~\eqref{eq:HellNdef} and $\widehat{p}_{nN}$ defined in~\eqref{eq:hatpnN}, and with $h = f(-\nicefrac{1}{2})$,
-\[
-H_{\ell,N} = (h-1)\sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big(C_{\ell k} + o_P(1)\big),
-\]
-where $C_{\ell k} = \tfrac{1}{\ell + k + 1}$.
+With $H_{\ell,N}$ defined in~\eqref{eq:HellNdef} and $\widehat{p}_{nN}$ defined in~\eqref{eq:hatpnN}, and with $h = f(-\nicefrac{1}{2})$, we have $H_{\ell,N} = (h-1)\sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big(C_{\ell k} + o_P(1)\big)$, where $C_{\ell k} = \tfrac{1}{\ell + k + 1}$.
 \end{lemma}
@@
 \begin{align*}
-Q(\widehat{p}_{nN}) - \widehat{p}_{nN} &= \widehat{p}_{nN}\big( h - 1 + \zeta(\widehat{p}_{nN}) \big) \\
-&= \widehat{p}_{nN}\big( h - 1 + o_P(1) \big), 
+Q(\widehat{p}_{nN}) - \widehat{p}_{nN} = \widehat{p}_{nN}\big( h - 1 + \zeta(\widehat{p}_{nN}) \big) = \widehat{p}_{nN}\big( h - 1 + o_P(1) \big), 
 \end{align*}
@@
    	\centering 
-  		%\includegraphics[width=\linewidth]{code/lseplot-1.mps} 
-               \includegraphics{code/lseplot-1.mps}
+  		\includegraphics[width=\linewidth]{code/lseplot-1.mps} 
+               %\includegraphics{code/lseplot-1.mps}
    		\caption{Performance of the LSE and LSU estimator for a polynomial phase signal of order two, i.e.
 $m=2$.
-(Left) Mean square error (MSE) in the second order parameter $\mu_2$ for $N=10,50$ and $200$ versus the variance $\sigma_c^2$ of the $X_1, \dots, X_N$.
+(Top) Mean square error (MSE) in the second order parameter $\mu_2$ for $N=10,50$ and $200$ versus the variance $\sigma_c^2$ of the $X_1, \dots, X_N$.
 The $X_1, \dots, X_N$ are independent and identically distributed (i.i.d.) complex Gaussian random variables.
-(Right) MSE in second order parameter $\mu_2$ versus the variance $\sigma^2$ of the $\Phi_1, \dots, \Phi_N$.
+(Bottom) MSE in second order parameter $\mu_2$ versus the variance $\sigma^2$ of the $\Phi_1, \dots, \Phi_N$.
 The $\Phi_1, \dots, \Phi_N$ are i.i.d.
@@
    	\centering 
-  		%\includegraphics[width=\linewidth]{code/gaussianplot4-1.mps}
-                \includegraphics{code/gaussianplot4-1.mps}
+  		\includegraphics[width=\linewidth]{code/gaussianplot4-1.mps}
+                %\includegraphics{code/gaussianplot4-1.mps}
   		\caption{Mean square error (MSE) of polynomial phase estimators for $N=10,50$ and $200$ versus the variance $\sigma_c^2$ of the complex noise terms $X_1, \dots X_N$.
@@
 
-\section{Computational considerations}
-
-When $m=0$, polynomial phase estimation is equivalent to the problem of \emph{direction estimation} from circular statistics, and the least squares estimator corresponds with the \emph{extrinsic mean} (or \emph{sample circular mean})~\cite{McKilliam_mean_dir_est_sq_arc_length2010,Fisher1993,Mardia_directional_statistics}.
+% \section{Computational considerations}
+
+% When $m=0$, polynomial phase estimation is equivalent to the problem of \emph{direction estimation} from circular statistics, and the least squares estimator corresponds with the \emph{extrinsic mean} (or \emph{sample circular mean})~\cite{McKilliam_mean_dir_est_sq_arc_length2010,Fisher1993,Mardia_directional_statistics}.
 In this case the LSE can be computed efficiently in $O(N)$ by averaging $N$ complex numbers.
@@
 
-Babai's nearest plane algorithm requires $O(N^2)$ operations and can be run in a reasonable amount of time for any $m$ and reasonably large $N$.
+% Babai's nearest plane algorithm requires $O(N^2)$ operations and can be run in a reasonable amount of time for any $m$ and reasonably large $N$.
 It should be noted that Babai's nearest plane algorithm requires a Lov\`as reduced lattice basis~\cite{Lenstra1982}.
@@
 
-Both Kitchen's estimator and the DPT are fast to compute.
+% Both Kitchen's estimator and the DPT are fast to compute.
 Kitchen's estimator requires only $O(N)$ arithmetic operations and the DPT requires $O(N \log N)$ operations.
@@
 
-\subsection{An empirical process result}
+\section{An empirical process result}
 
@@
 \begin{align*}
-f_{nN}(\psibf, \Phi_n) &= \left( \tfrac{n}{N}\right)^\ell q_n(p_{nN}(\psibf)) \\
-%&= \left( \tfrac{n}{N}\right)^\ell \round{\Phi_n + p_{nN}(\psibf)} \\
-&= \left( \tfrac{n}{N}\right)^\ell \round{\Phi_n + \sum_{k=0}^m\left( \tfrac{n}{N}\right)^k\psi_k}
+f_{nN}(\psibf, \Phi_n) = \left( \tfrac{n}{N}\right)^\ell q_n(p_{nN}(\psibf)) = \left( \tfrac{n}{N}\right)^\ell \round{\Phi_n + \sum_{k=0}^m\left( \tfrac{n}{N}\right)^k\psi_k}
 \end{align*}
@@
 %That is, expectation taken over $\{g_n\}$ and $\{\Phi_n^\prime\}$ treating $\{\Phi_n\}$ as fixed.
-Since $\Phi_n$ and $\Phi_n^\prime$ are identically distributed,
-\[
-\expect f_{nN}(\psibf, \Phi_n) = \expect_\Phi f_{nN}(\psibf, \Phi_n^\prime),
-\]
-so that,
+Since $\Phi_n$ and $\Phi_n^\prime$ are identically distributed $\expect f_{nN}(\psibf, \Phi_n) = \expect_\Phi f_{nN}(\psibf, \Phi_n^\prime)$ so that,
 \begin{align*}
@@
 \end{align*}
-where, for notational convenience, we put 
-\[
-f_{nN} = f_{nN}(\psibf, \Phi_n) \qquad \text{and} \qquad  f_{nN}^\prime = f_{nN}(\psibf, \Phi_n^\prime).
-\] 
+where we put $f_{nN} = f_{nN}(\psibf, \Phi_n)$ and  $f_{nN}^\prime = f_{nN}(\psibf, \Phi_n^\prime)$  for notational convenience.
 Taking absolute values followed by supremums,
@@
 \end{equation}
-since $\expect \expect_\Phi$ is equivalent to $\expect$.   Let
-\[
-\sigma_n = \frac{g_n}{\abs{g_n}},
-\]
-and put $\sigma_n = 1$ if $g_n = 0$.  The $\sigma_n$ are thus independent with 
-\[
-\prob\{ \sigma_n = -1 \} = \tfrac{1}{2} \qquad \text{and} \qquad  \prob\{ \sigma_n = 1 \} = \tfrac{1}{2}.
-\]  
+since $\expect \expect_\Phi$ is equivalent to $\expect$.
+Let $\sigma_n = \frac{g_n}{\abs{g_n}}$ and put $\sigma_n = 1$ if $g_n = 0$.
+The $\sigma_n$ are thus independent with $\prob\{ \sigma_n = -1 \} = \tfrac{1}{2}$ and $\prob\{ \sigma_n = 1 \} = \tfrac{1}{2}$.
 As $\Phi_n$ and $\Phi_n^\prime$ are independent and identically distributed, the random variable $f_{nN} - f_{nN}^\prime$ is symmetrically distributed about zero, and is therefore distributed identically to $\sigma_n( f_{nN} - f_{nN}^\prime)$.
@@
 \begin{align*}
- \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  &\abs{  \frac{1}{\sqrt{N}}\sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) } \\
-&= \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{  \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n ( f_{nN} - f_{nN}^\prime) }.
+ \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{  \frac{1}{\sqrt{N}}\sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) } = \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{  \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n ( f_{nN} - f_{nN}^\prime) }.
 \end{align*}
@@
 \begin{align*}
- \expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} &\leq \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} } \\
-&\hspace{1cm} + \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{\frac{1}{\sqrt{N}} \sum_{n=1}^{N}  \sigma_n  f_{nN}^\prime } \\
+ \expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} &\leq \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} } + \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{\frac{1}{\sqrt{N}} \sum_{n=1}^{N}  \sigma_n  f_{nN}^\prime } \\
 &= 2 \; \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} }.
@@
 \begin{align*}
-\expect \sup_{\|\psibf\|_{\infty} < \epsilon}  &\abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} } \\
-&= \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} \sqrt{\frac{\pi}{2}}\expect_\Phi \abs{g_n}} \\
+\expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} } &= \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} \sqrt{\frac{\pi}{2}}\expect_\Phi \abs{g_n}} \\
 &\leq \sqrt{\frac{\pi}{2}} \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{ \sqrt{N}}\sum_{n=1}^{N} \sigma_n \abs{g_n}  f_{nN}},
@@
 \]
-for every positive integer $k$, where $A_1$ and $A_2$ are the constants,
-\begin{equation}\label{eq:A1andA2}
-A_1 = 18(m+1)\log 2\ \qquad \text{and} \qquad A_2 = 18\log K_3
-\end{equation}
-and $\log(\cdot)$ is the natural logarithm.  The lemma holds with $K_1 =  \sum_{i=1}^{\infty}2^{-i/2}\sqrt{ i A_1 + A_2 }$.
+for every positive integer $k$, where $A_1 = 18(m+1)\log 2$ and $A_2 = 18\log K_3$ are constants and $\log(\cdot)$ is the natural logarithm.
+The lemma holds with $K_1 =  \sum_{i=1}^{\infty}2^{-i/2}\sqrt{ i A_1 + A_2 }$.
 \end{proof}
@@
 \begin{proof}
-Recall that
-\[
-f_{nN}(\psibf) = \left( \tfrac{n}{N}\right)^\ell \round{\Phi_n + p_{nN}(\psibf)}.
-\]
-Because $\abs{\psi_i} < \epsilon$ for all $i = 0, \dots, m$, 
+Recall that $f_{nN}(\psibf) = \left( \tfrac{n}{N}\right)^\ell \round{\Phi_n + p_{nN}(\psibf)}$.  Because $\abs{\psi_i} < \epsilon$ for all $i = 0, \dots, m$, 
 \begin{equation}\label{eq:pnNsmall}
@@
 \]
-where the constants $A_1$ and $A_2$ are defined in~\eqref{eq:A1andA2}.
+where the constants $A_1 = 18(m+1)\log 2$ and $A_2 = 18\log K_3$.
 \end{lemma}
@@
 \begin{align}
-&\sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) } \nonumber \\ %\leq \sup_{\psibf \in T_\epsilon(k) } \big( \abs{ Z_N(\psibf^*) } + \abs{ Z_N(\psibf) - Z_N(\psibf^*)  } \big) \nonumber \\
-&\leq \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(b_k(\psibf)) } + \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) - Z_N(b_k(\psibf))  } \nonumber \\
+\sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) } &\leq \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(b_k(\psibf)) } + \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) - Z_N(b_k(\psibf))  } \nonumber \\
 &\leq \sup_{\psibf \in T_\epsilon(k-1)} \sabs{ Z_N(\psibf) } + \sup_{\psibf \in T_\epsilon(k)} \sabs{ Z_N(\psibf) - Z_N(b_k(\psibf)) }, \label{eq:supZchain}
@@
 \]
-Conditional on $\Phi_1, \Phi_2, \dots$ the random variable
-\[
-X(\psibf) = Z_N(\psibf) - Z_N(b_k(\psibf))
-\]
-has zero mean, and is normally distributed with variance
+Conditional on $\Phi_1, \Phi_2, \dots$ the random variable $X(\psibf) = Z_N(\psibf) - Z_N(b_k(\psibf))$ has zero mean, and is normally distributed with variance
 \begin{align*}
 \sigma_X^2 &= \expect_{\Phi} \frac{1}{N} \sum_{n=1}^N g_n^2 ( f_{nN}(\psibf) - f_{nN}(b_k(\psibf)) )^2 \\
-&= \frac{1}{N} \sum_{n=1}^N ( f_{nN}(\psibf) - f_{nN}(b_k(\psibf)) )^2 \\
-&= d(\psibf, b_k(\psibf)) \leq 2^{1-k} C_\epsilon,
+&= \frac{1}{N} \sum_{n=1}^N ( f_{nN}(\psibf) - f_{nN}(b_k(\psibf)) )^2 = d(\psibf, b_k(\psibf)) \leq 2^{1-k} C_\epsilon,
 \end{align*}
@@
 \begin{align*}
-\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k)} \sabs{X(\psibf)} &\leq 3\sqrt{2^{1-k} C_\epsilon \log\sabs{T_\epsilon(k)}} \\
-&\leq \sqrt{C_\epsilon}\frac{\sqrt{  k A_1 + A_2 }}{2^{k/2}}
-\end{align*}
-because $\log\abs{T_\epsilon(k)} \leq k(m+1)\log 2 +  \log K_3$.  Taking expectations on both side of~\eqref{eq:supZchain},
-\begin{align*}
-\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) } &\leq \expect_{\Phi} \sup_{\psibf \in T_\epsilon(k-1)} \sabs{ Z_N(\psibf) } \\
-&\hspace{1cm} + \sqrt{C_\epsilon}\frac{\sqrt{  k A_1 + A_2 }}{2^{k/2}},
+\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k)} \sabs{X(\psibf)} \leq 3\sqrt{2^{1-k} C_\epsilon \log\sabs{T_\epsilon(k)}} \leq \sqrt{C_\epsilon}\frac{\sqrt{  k A_1 + A_2 }}{2^{k/2}}
+\end{align*}
+because $\log\abs{T_\epsilon(k)} \leq k(m+1)\log 2 +  \log K_3$.  Taking expectations on both sides of~\eqref{eq:supZchain},
+\begin{align*}
+\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) } \leq \expect_{\Phi} \sup_{\psibf \in T_\epsilon(k-1)} \sabs{ Z_N(\psibf) }  + \sqrt{C_\epsilon}\frac{\sqrt{  k A_1 + A_2 }}{2^{k/2}},
 \end{align*}
@@
 \]
-we obtain
-\begin{align*}
-\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) } < \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{\sqrt{  k A_1 + A_2 }}{2^{k/2}}
-\end{align*}
+we obtain $\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) } < \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{\sqrt{  k A_1 + A_2 }}{2^{k/2}}$
 as required.
@@
 %It will be more convenient to work with $d_g$ rather than $d$.
-
-From~\eqref{eq:pnNsmall}, 
-\[
-\abs{p_{nN}(\psibf)} \leq (m+1)\epsilon < 1.
-\]
-Since $\Phi_n \in [-\nicefrac{1}{2}, \nicefrac{1}{2})$, when $\Phi_n \geq 0$,
+From~\eqref{eq:pnNsmall} it follows that $\abs{p_{nN}(\psibf)} \leq (m+1)\epsilon < 1$.  Since $\Phi_n \in [-\nicefrac{1}{2}, \nicefrac{1}{2})$, when $\Phi_n \geq 0$,
 \[
@@
 \begin{align*}
-d_g(\psibf, \psibf^*) &= \sum_{n=1}^{N} \big( g_{nN}(\psibf) - g_{nN}(\psibf^*) \big)^2 \\
-&= \sum_{n \in A} \big( g_{nN}(\psibf) - g_{nN}(\psibf^*) \big)^2.
+d_g(\psibf, \psibf^*) = \sum_{n=1}^{N} \big( g_{nN}(\psibf) - g_{nN}(\psibf^*) \big)^2 = \sum_{n \in A} \big( g_{nN}(\psibf) - g_{nN}(\psibf^*) \big)^2.
 \end{align*}

### 2012-10-13 plots/circstatfigcube.rb

@@ -95,7 +95,7 @@ drawsamplecircpoly(5.02,1, lambda { |t| poly2.call(t) } , spiralpic)
 
 fig1 = Figure.new
 fig1.add_drawable(Draw.new(polypic))
-fig1.add_drawable(Draw.new(polycircpic).translate((6.2).cm,2.5.cm))
+fig1.add_drawable(Draw.new(polycircpic).translate((6.5).cm,2.5.cm))
 #fig1.add_drawable(Draw.new(spiralpic).translate(2.2.cm,(-9.3).cm))
 file.add_figure(fig1)
 

### 2012-10-13 plots/circstatfigquad.rb

@@ -94,7 +94,7 @@ drawsamplecircpoly(5.02,1, lambda { |t| poly2.call(t) } , spiralpic)
 
 fig1 = Figure.new
 fig1.add_drawable(Draw.new(polypic))
-fig1.add_drawable(Draw.new(polycircpic).translate((6.2).cm,2.5.cm))
+fig1.add_drawable(Draw.new(polycircpic).translate((6.5).cm,2.5.cm))
 #fig1.add_drawable(Draw.new(spiralpic).translate(2.2.cm,(-9.3).cm))
 file.add_figure(fig1)
 

### 2012-10-14 paper.tex
modified plots.  edits.

@@
 
-Polynomial phase signals arise in numerous fields including radar, sonar, geophysics, biology, and radio communication \cite{Angeby_estimating_2000}.
-In radar and sonar applications polynomial phase signals arise when acquiring radial velocity and acceleration (and higher order motion descriptors) of a target from a reflected signal, and also in continuous wave radar and low probability of intercept radar~\cite{Levanon_Radar_signals_2004}.
-In biology, polynomial phase signals are used to describe the sounds emitted by bats and dolphins for echo location \cite{Suga_1975_bats_echolocation, Moss_2005echolocation}.
+Polynomial phase signals arise in numerous fields including radar, sonar, geophysics, speech analysis, biology, and radio communication~\cite{Hlawatsch_lin_quad_time_freq_spmag_1992,Ridleyspeechpolyphase1989, Suga_1975_bats_echolocation, Moss_2005echolocation}.
+In radar and sonar applications polynomial phase signals arise when acquiring radial velocity and acceleration (and higher order motion descriptors) of a target from a reflected signal, and also in continuous wave radar and low probability of intercept radar.
+In biology, polynomial phase signals are used to describe the sounds emitted by bats and dolphins for echo location.
 
@@
 In this paper we only consider uniform sampling, where the gap between consecutive samples is constant.
-In this case we can always consider the samples to be taken at some set of consecutive integers and our sampled polynomial phase signal is
-\[
-s_n = s(n) = e^{2\pi j y(n)},
-\] 
-where $n$ is an integer.  Of practical importance is the estimation of the coefficients $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ from a number, say $N$, of observations of the noisy sampled signal
+In this case we can always consider the samples to be taken at some set of consecutive integers and our sampled polynomial phase signal is $s_n = s(n) = e^{2\pi j y(n)}$, where $n$ is an integer.
+Of practical importance is the estimation of the coefficients $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ from a number, say $N$, of consequetive observations of the noisy sampled signal
 \begin{equation}\label{eq:Y_nsamplednoisey}
@@
 
-An obvious estimator of the unknown coefficients is the least squares estimator (LSE).
-This is also the maximum likelihood estimator (MLE) when the noise sequence $\{X_n\}$ is white and Gaussian.
-When $m=0$ (phase estimation) or $m=1$ (frequency estimation) the LSE is an effective approach, being both computationally efficient and statistically accurate \cite{Quinn2009_dasp_phase_only_information_loss,Hannan1973,Quinn2001,McKilliam_mean_dir_est_sq_arc_length2010}\cite[Sec.~6.4~and~9.1]{McKilliam2010thesis}.
-When $m \geq 2$ the computational complexity of the LSE is large~\cite[Sec.~10.1]{McKilliam2010thesis}\cite{Abatzoglou_ml_chirp_1986}.
-For this reason many authors have considered alternative approaches to polynomial phase estimation.
-These can loosely be grouped into two classes, estimators based on polynomial phase transforms, such as the discrete polynomial phase transform (DPT)~\cite{Peleg_DPT_1995,Peleg1991_est_class_PPS_1991,Porat_asympt_HAF_DPT_1996} and the high order phase function~\cite{Farquharson_another_poly_est_2005}, and estimators based on phase unwrapping, such as Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, the estimator of Djuric and Kay~\cite{Djuric_phase_unwrap_chirp_1990}, and Morelande's Bayesian unwrapping estimator~\cite{Morelande_bayes_unwrapping_2009_tsp}.
+An obvious estimator of the unknown coefficients is the least squares estimator.  
+%This is also the maximum likelihood estimator (MLE) when the noise sequence $\{X_n\}$ is white and Gaussian.  
+When $m=0$ (phase estimation) or $m=1$ (frequency estimation) the least squares estimator is an effective approach, being both computationally efficient and statistically accurate~\cite{Hannan1973,Quinn2001,McKilliam_mean_dir_est_sq_arc_length2010}.
+When $m \geq 2$ the computational complexity of the least squares estimator is large, and for this reason authors have considered alternative estimators.
+These can loosely be grouped into two classes; estimators based on polynomial phase transforms, such as the discrete polynomial phase transform~\cite{Peleg_DPT_1995,Porat_asympt_HAF_DPT_1996} and the high order phase function~\cite{Farquharson_another_poly_est_2005}; and estimators based on phase unwrapping, such as Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, and Morelande's Bayesian unwrapping estimator~\cite{Morelande_bayes_unwrapping_2009_tsp}.
 
 In this paper we consider the estimator that results from unwrapping the phase in a least squares manner.
-We call this the \emph{least squares unwrapping estimator} (LSU)~\cite{McKilliam2009asilomar_polyest_lattice, McKilliamFrequencyEstimationByPhaseUnwrapping2009}\cite[Chap.
-8]{McKilliam2010thesis}.
-It was shown in \cite{McKilliam2009asilomar_polyest_lattice, McKilliamFrequencyEstimationByPhaseUnwrapping2009} that the LSU estimator can be represented as a \emph{nearest lattice point problem}, and Monte-Carlo simulations were used to show the LSU estimator's favourable statistical performance.
+We call this the \emph{least squares unwrapping estimator} (LSU).
+It was shown in \cite{McKilliam2009asilomar_polyest_lattice} that the LSU estimator can be represented as a \emph{nearest lattice point problem}, and Monte-Carlo simulations were used to show the LSU estimator's favourable statistical performance.
 %Moreover, it was noticed that the LSU appears to work correctly for coefficients anywhere in the identifiable region.
@@
 In this paper we derive the asymptotic properties of the LSU estimator, showing that it is strongly consistent and asymptotically normally distributed.
-Similar results were stated without a complete proof in~\cite{McKilliam_polyphase_est_icassp_2011}.
+Similar results were stated without proof in~\cite{McKilliam_polyphase_est_icassp_2011}.
 Here, we give a proof.
@@
 
+An interesting property is that the estimator of the $k$th polynomial phase coefficient converges almost surely to $\tilde{\mu}_k$ at rate $N^{-k}$.
+This is perhaps not surprising, since it is the same rate observed in polynomial regression.
+However, asserting that convergence at this rate occurs is not trivial.
+For this purpose we make use of elementary results about the number of arithmetic progressions that fit inside subsets of $\{1,2,\dots,N\}$.
+This proof technique appears to be novel, and may be useful in other problems.
+The proof of asymptotic normality is complicated by the fact that the objective function corresponding with the LSU estimator is not differentiable everywhere.
+Techniques developed by Huber~\cite{Huberrobustest1964,Huberbehavmaxnonstandard1967} and refined by Pollard~\cite{Pollard_new_ways_clts_1986,Pollard_asymp_empi_proc_1989} become useful here.
+We believe the LSU estimator to be an interesting new application of these techniques.
+We are hopeful that the proof techniques developed here will be useful for purposes other than polynomial phase estimation, and in particularly other applications involving data that is `wrapped' in some sense.
+Potential candidates are the phase wrapped images observed in synthetic aperture radar and magnetic resonance imaging~\cite{Nico_phaseunwrappingSAR_2000}.
+
 The paper is organised in the following way.
-Section~\ref{sec:lattice-theory} describes the required concepts from lattice theory.
-In Section~\ref{sec:ident_aliasing} we briefly describe the identifiable region that was also derived in~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}.
+Section~\ref{sec:lattice-theory} describes some preliminary concepts from lattice theory, and in Section~\ref{sec:ident_aliasing} we use these results to describe an identifiable region for the set of polynomial phase coefficients.
 These identifiability results are required in order to properly understand the statistical properties of polynomial phase estimators.
-In Section \ref{sec:least-squar-unwr} we describe the LSU estimator and derive its asymptotic statistical properties.
+In Section \ref{sec:least-squar-unwr} we describe the LSU estimator and state its asymptotic statistical properties.
+Section~\ref{sec:strongconstproof} gives the proof of strong consistency and Section~\ref{sec:centlimitproof} gives the proof of asymptotic normality.
 %and also consider techniques for computing the LSU as a nearest lattice point problem. 
-Section \ref{sec:simulations} describes the results of Monte Carlo simulations that compare the performance of the LSU estimator with existing estimators.
-These simulations agree with the derived asymptotic properties.
+Section~\ref{sec:simulations} describes the results of Monte Carlo simulations with the LSU estimator.  These simulations agree with the derived asymptotic properties.
 
@@
 Another choice is based on the \emph{Voronoi cell}, i.e., those points from $\reals^n$ nearest (with respect to a given norm) to the lattice point at the origin.
+In what follows when we refer to the Voronoi cell, we will also mean the Vornoi cell with respect to the Euclidean norm.
 It is always possible to construct a \emph{rectangular} set of representatives, as the next proposition will show.
 We will use these rectangular regions for describing the aliasing of polynomial phase signals in Section~\ref{sec:ident_aliasing}.
-These rectangular regions will also be important for the derivation of the asymptotic properties of the LSU estimator in Section~\ref{sec:least-squar-unwr}.
+These rectangular regions will be important for the derivation of the asymptotic properties of the LSU estimator in Section~\ref{sec:least-squar-unwr}.
 
@@
 
-Given a point $\ybf\in\reals^n$ and a lattice $\Lambda$ contained in~$\reals^n$, the \emph{nearest lattice point problem} is to find the point $\xbf \in \Lambda$ such that the distance (with respect to a given norm) between $\ybf$ and $\xbf$ is minimised.
+% Given a point $\ybf\in\reals^n$ and a lattice $\Lambda$ contained in~$\reals^n$, the \emph{nearest lattice point problem} is to find the point $\xbf \in \Lambda$ such that the distance (with respect to a given norm) between $\ybf$ and $\xbf$ is minimised.
 Here, the Euclidean norm is assumed, so that the nearest point is
-\[ 
-\arg\min_{\xbf \in \Lambda} \| \xbf - \ybf \|^2.
-\]
-The nearest lattice point problem is known to be NP-hard under certain conditions when the lattice itself, or rather a basis thereof, is considered as an additional input parameter~\cite{micciancio_hardness_2001, Jalden2005_sphere_decoding_complexity}.
+% \[ 
+% \arg\min_{\xbf \in \Lambda} \| \xbf - \ybf \|^2.
+% \]
+% The nearest lattice point problem is known to be NP-hard under certain conditions when the lattice itself, or rather a basis thereof, is considered as an additional input parameter~\cite{micciancio_hardness_2001, Jalden2005_sphere_decoding_complexity}.
 Nevertheless, algorithms exist that can compute a nearest lattice point in reasonable time if the dimension is small.
@@
 
-Approximate algorithms for computing the nearest point have also been studied.
+% Approximate algorithms for computing the nearest point have also been studied.
 A classic example is \emph{Babai's nearest plane algorithm}~\cite{Babai1986}, which requires $O(n^4)$ arithmetic operations in the worst case where $n$ is the dimension of the lattice and only $O(n^2)$ if the lattice basis is \emph{Lov\'asz reduced}~\cite{Lenstra1982}\index{Lov\'asz reduced}.
@@
 
-In~\cite{McKilliam2009asilomar_polyest_lattice}, it was shown how the least squares phase unwrapping estimator of a polynomial phase signal can be computed by finding a nearest point in a lattice.
+% In~\cite{McKilliam2009asilomar_polyest_lattice}, it was shown how the least squares phase unwrapping estimator of a polynomial phase signal can be computed by finding a nearest point in a lattice.
 We will briefly describe this result in Section~\ref{sec:least-squar-unwr}.
@@
 
-\section{The Least squares unwrapping estimator}\label{sec:least-squar-unwr}
+\section{The least squares unwrapping estimator}\label{sec:least-squar-unwr}
 
@@
 
-Figure~\ref{plot:polyest} displays the performance of the LSU estimator, the DPT and Kitchen's estimator when $m=3$.
+Figures~\ref{plot:polyest1}~and~~\ref{plot:polyest2} displays the performance of the LSU estimator, the DPT and Kitchen's estimator when $m=3$.
 Both the DPT and Kitchen's estimator perform poorly because they do not work correctly for some parameters in the identifiable region $B$.
@@
 
-\begin{figure}[tp] 
+\begin{figure}[p] 
    	\centering 
@@
 (Bottom right) MSE in the cubic coefficient $\mu_3$.}
-  		\label{plot:polyest} 
+  		\label{plot:polyest1} 
+ \end{figure}
+
+\begin{figure}[p] 
+   	\centering 
+  		\includegraphics[width=\linewidth]{code/gaussianplot4-2.mps}
+                %\includegraphics{code/gaussianplot4-1.mps}
+  		\caption{Mean square error (MSE) of polynomial phase estimators for $N=10,50$ and $200$ versus the variance $\sigma_c^2$ of the complex noise terms $X_1, \dots X_N$.
+The polynomial phase signal has order 3, and the $X_1, \dots, X_N$ are independent and identically distributed complex Gaussian random variables.
+(Top left) MSE in the phase coefficient $\mu_0$.
+(Top right) MSE in the frequency coefficient $\mu_1$.
+(Bottom left) MSE in the quadratic coefficient $\mu_2$.
+(Bottom right) MSE in the cubic coefficient $\mu_3$.}
+  		\label{plot:polyest2} 
  \end{figure}
@@
 Let $\sigma_n = \frac{g_n}{\abs{g_n}}$ and put $\sigma_n = 1$ if $g_n = 0$.
-The $\sigma_n$ are thus independent with $\prob\{ \sigma_n = -1 \} = \tfrac{1}{2}$ and $\prob\{ \sigma_n = 1 \} = \tfrac{1}{2}$.
+The $\sigma_n$ are independent with $\prob\{ \sigma_n = -1 \} = \tfrac{1}{2}$ and $\prob\{ \sigma_n = 1 \} = \tfrac{1}{2}$.
 As $\Phi_n$ and $\Phi_n^\prime$ are independent and identically distributed, the random variable $f_{nN} - f_{nN}^\prime$ is symmetrically distributed about zero, and is therefore distributed identically to $\sigma_n( f_{nN} - f_{nN}^\prime)$.

### 2012-10-14 paper.tex
minor

@@
 We call this the \emph{least squares unwrapping estimator} (LSU).
-It was shown in \cite{McKilliam2009asilomar_polyest_lattice} that the LSU estimator can be represented as a \emph{nearest lattice point problem}, and Monte-Carlo simulations were used to show the LSU estimator's favourable statistical performance.
+It was shown in \cite{McKilliam2009asilomar_polyest_lattice} that the LSU estimator can be computed by finding a nearest point in a lattice~\cite{Agrell2002}, and Monte-Carlo simulations were used to show the LSU estimator's favourable statistical performance.
 %Moreover, it was noticed that the LSU appears to work correctly for coefficients anywhere in the identifiable region.
@@
 %We also apply another known approximate nearest point algorithm, called the $K$-best method \cite{Zhan2006_K_best_sphere_decoder}, that we show provides near sphere decoder performance, but can be computed in a reasonable amount of time if $N$ is less than about 300.
-In this paper we derive the asymptotic properties of the LSU estimator, showing that it is strongly consistent and asymptotically normally distributed.
+In this paper we derive the asymptotic properties of the LSU estimator.
+Under some assumptions about the distribution of the noise $X_1, \dots, X_N$, we show the estimator to be strongly consistent and asymptotically normally distributed.
 Similar results were stated without proof in~\cite{McKilliam_polyphase_est_icassp_2011}.
@@
 We believe the LSU estimator to be an interesting new application of these techniques.
-We are hopeful that the proof techniques developed here will be useful for purposes other than polynomial phase estimation, and in particularly other applications involving data that is `wrapped' in some sense.
+We are hopeful that the proof techniques developed here will be useful for purposes other than polynomial phase estimation, and in particular other applications involving data that is `wrapped' in some sense.
 Potential candidates are the phase wrapped images observed in synthetic aperture radar and magnetic resonance imaging~\cite{Nico_phaseunwrappingSAR_2000}.
@@
 %\]
-
 Aliasing can occur when polynomial-phase signals are sampled.
 That is, two or more distinct polynomial-phase signals can take exactly the same values at the sample points.
-These aliasing results are also given in~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}, but the presentation here is different, and is better suited to studying the asymptotic properties of the LSU estimator.
-A description of the aliasing properties of polynomial phase signals of order $2$ has been independently discovered in~\cite{Abatzoglou_ml_chirp_1986}~and~\cite{Angeby_PPS_aliasing_2000}.
+Understanding how aliasing occurs is crucial to understanding the behaviour of the polynomial phase estimators.
+The aliasing properties are described in~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}, but, here we present the properties in a way that is better suited to studying the LSU estimator.
+%A description of the aliasing properties of polynomial phase signals of order $2$ has been independently discovered in~\cite{Abatzoglou_ml_chirp_1986}~and~\cite{Angeby_PPS_aliasing_2000}.
 The results here are for polynomial phase signals of any order.
@@
 That is, $\mathcal{Z}$ contains all polynomials $p$ such that $p(n)$ is an integer whenever $n$ is an integer.
-Let $y$ and $z$ be two \emph{distinct} polynomials such that $z = y + p$ for some polynomial $p$ in $\mathcal{Z}$. The two polynomial phase signals
-\[
-s(t) = e^{2\pi j y(t)} \qquad \text{and} \qquad r(t) = e^{2\pi j z(t)}
-\]
-are distinct because $y$ and $z$ are distinct, but if we sample $s$ and $r$ at the integers  
+Let $y$ and $z$ be two distinct polynomials such that $z = y + p$ for some polynomial $p$ in $\mathcal{Z}$.
+The two polynomial phase signals $s(t) = e^{2\pi j y(t)}$ and $r(t) = e^{2\pi j z(t)}$ are distinct because $y$ and $z$ are distinct, but if we sample $s$ and $r$ at the integers
 \begin{align*}
@@
 
-It may be helpful to observe Figures~\ref{fig:circstatplot_line} and~\ref{fig:circstatplot_cube}.
+It may be helpful to observe Figures~\ref{fig:circstatplot_line},~\ref{fig:circstatplot_quad}~and~\ref{fig:circstatplot_cube}.
 In these, the phase (divided by $2\pi$) of two distinct polynomial phase signals is plotted on the left, and on the right the principal component of the phase is plotted.
@@
 \begin{equation} \label{eq:lem_polynomial}
-c_0 p_0 + c_1 p_1 + \dots + c_m p_m,
-\end{equation}
-where the $c_i \in \ints$.
+c_0 p_0 + c_1 p_1 + \dots + c_m p_m, \qquad c_0,c_1,\dots,c_m \in \ints.
+\end{equation}
 \end{lemma}

### 2012-10-15 paper.tex
edited paper.  Simplified plots.

@@
 \author{\fnms{Badri N.} \snm{Vellambi}\thanksref{m1}\ead[label=e5]{badri.vellambi@unisa.edu.au}}
-\runauthor{R. McKilliam et al.}
+\runauthor{R. McKilliam et. al.}
 
@@
 \begin{abstract}
-Estimating the coefficients of a noisy polynomial phase signal is important in many fields including radar, biology and radio communications.
+Estimating the coefficients of a noisy polynomial phase signal is important in fields including radar, biology and radio communications.
 One approach attempts to perform polynomial regression on the phase of the signal.
@@
 
-Polynomial phase signals arise in numerous fields including radar, sonar, geophysics, speech analysis, biology, and radio communication~\cite{Hlawatsch_lin_quad_time_freq_spmag_1992,Ridleyspeechpolyphase1989, Suga_1975_bats_echolocation, Moss_2005echolocation}.
+Polynomial phase signals arise in fields including radar, sonar, geophysics, speech analysis, biology, and radio communication~\cite{Hlawatsch_lin_quad_time_freq_spmag_1992,Ridleyspeechpolyphase1989, Suga_1975_bats_echolocation, Moss_2005echolocation}.
 In radar and sonar applications polynomial phase signals arise when acquiring radial velocity and acceleration (and higher order motion descriptors) of a target from a reflected signal, and also in continuous wave radar and low probability of intercept radar.
@@
 In this case we can always consider the samples to be taken at some set of consecutive integers and our sampled polynomial phase signal is $s_n = s(n) = e^{2\pi j y(n)}$, where $n$ is an integer.
-Of practical importance is the estimation of the coefficients $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ from a number, say $N$, of consequetive observations of the noisy sampled signal
+Of practical importance is the estimation of the coefficients $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ from a number, say $N$, of consecutive observations of the noisy sampled signal
 \begin{equation}\label{eq:Y_nsamplednoisey}
@@
 \end{equation}
-where $\rho$ is a real number greater than zero representing the (usually unknown) signal amplitude and $\{X_n, n \in \ints\}$ is a sequence complex noise variables.
+where $\rho$ is a real number greater than zero representing the (usually unknown) signal amplitude and $\{X_n, n \in \ints\}$ is a sequence of complex noise variables.
 In order to ensure identifiability it is necessary to restrict the $m+1$ coefficients to a region of $m+1$ dimensional Euclidean space $\reals^{m+1}$ called an \emph{identifiable region}.
-It was shown in \cite{McKilliam2009IndentifiabliltyAliasingPolyphase} that an identifiable region tessellates a particular $m+1$ dimensional lattice.
+It was shown by some of the authors \cite{McKilliam2009IndentifiabliltyAliasingPolyphase} that an identifiable region tessellates a particular $m+1$ dimensional lattice.
 We discuss this in Section \ref{sec:ident_aliasing}.
@@
 In this paper we consider the estimator that results from unwrapping the phase in a least squares manner.
-We call this the \emph{least squares unwrapping estimator} (LSU).
-It was shown in \cite{McKilliam2009asilomar_polyest_lattice} that the LSU estimator can be computed by finding a nearest point in a lattice~\cite{Agrell2002}, and Monte-Carlo simulations were used to show the LSU estimator's favourable statistical performance.
+We call this the \emph{least squares unwrapping} (LSU) estimator.
+It was shown by some of the authors \cite[Sec.
+8.1]{McKilliam2010thesis}\cite{McKilliam2009asilomar_polyest_lattice} that the LSU estimator can be computed by finding a nearest point in a lattice~\cite{Agrell2002}, and Monte-Carlo simulations were used to show the LSU estimator's favourable statistical performance.
 %Moreover, it was noticed that the LSU appears to work correctly for coefficients anywhere in the identifiable region.
@@
 This is perhaps not surprising, since it is the same rate observed in polynomial regression.
-However, asserting that convergence at this rate occurs is not trivial.
-For this purpose we make use of elementary results about the number of arithmetic progressions that fit inside subsets of $\{1,2,\dots,N\}$.
-This proof technique appears to be novel, and may be useful in other problems.
+However, asserting that convergence at this rate occurs in the polynomial phase setting is not trivial.
+For this purpose we make use of an elementary result about the number of arithmetic progressions contained inside subsets of $\{1,2,\dots,N\}$.
+%This proof technique appears to be novel, and may be useful in other problems.
 The proof of asymptotic normality is complicated by the fact that the objective function corresponding with the LSU estimator is not differentiable everywhere.
-Techniques developed by Huber~\cite{Huberrobustest1964,Huberbehavmaxnonstandard1967} and refined by Pollard~\cite{Pollard_new_ways_clts_1986,Pollard_asymp_empi_proc_1989} become useful here.
-We believe the LSU estimator to be an interesting new application of these techniques.
+Techniques developed by Huber~\cite{Huberrobustest1964,Huberbehavmaxnonstandard1967} and refined by Pollard~\cite{Pollard_new_ways_clts_1986,Pollard_asymp_empi_proc_1989} and results from the literature on hyperplane arrangements~\cite{Chazelle_discrepency_method_2000,Matousek_lect_disc_geom_2002} become useful here.
 We are hopeful that the proof techniques developed here will be useful for purposes other than polynomial phase estimation, and in particular other applications involving data that is `wrapped' in some sense.
@@
 One choice for a set of coset representatives is a \emph{fundamental parallelepiped}, i.e., the parallelepiped generated by the columns of a generator matrix.
-Another choice is based on the \emph{Voronoi cell}, i.e., those points from $\reals^n$ nearest (with respect to a given norm) to the lattice point at the origin.
-In what follows when we refer to the Voronoi cell, we will also mean the Vornoi cell with respect to the Euclidean norm.
+Another choice is based on the \emph{Voronoi cell}, i.e., those points from $\reals^n$ nearest (with respect to the Euclidean norm here ) to the lattice point at the origin.
 It is always possible to construct a \emph{rectangular} set of representatives, as the next proposition will show.
-We will use these rectangular regions for describing the aliasing of polynomial phase signals in Section~\ref{sec:ident_aliasing}.
+We will use these rectangular regions for describing the aliasing properties of polynomial phase signals in Section~\ref{sec:ident_aliasing}.
 These rectangular regions will be important for the derivation of the asymptotic properties of the LSU estimator in Section~\ref{sec:least-squar-unwr}.
@@
 Let $\Bbf = \Qbf\Rbf$ where $\Qbf$ is orthonormal and $\Rbf$ is upper triangular with elements $r_{ij}$.
-Then the rectangular prism $\Qbf P$ where
-\[
-P = \prod_{k=1}^{n}{\left[-\frac{r_{kk}}{2}, \frac{r_{kk}}{2}\right)}
-\]
-is a set of coset representatives for $\reals^n / \Lambda$.
+Then the rectangular prism $\Qbf P$ where $P = \prod_{k=1}^{n}{[-\tfrac{r_{kk}}{2}, \frac{r_{kk}}{2})}$ is a set of coset representatives for $\reals^n / \Lambda$.
 \end{proposition}
@@
 \[
-\coef(g) = \left[ \begin{array}{ccccc} a_0 & a_1 & a_2 & \dots & a_m, \end{array} \right]^\prime,
-\]
-where superscript $^\prime$ indicates the transpose, denote the column vector of length $m+1$ containing the coefficients of $g$.
-If $y$ and $z$ differ by a polynomial from $\mathcal{Z}$ then we can write $y = z + p$ where $p \in \mathcal{Z}$ and then also $\coef(y) = \coef(z) + \coef(p)$.
+\coef(g) = \left[ \begin{array}{ccccc} a_0 & a_1 & a_2 & \dots & a_m, \end{array} \right]^\prime
+\]
+denote the column vector of length $m+1$ containing the coefficients of $g$.
+We use superscript $^\prime$ to indicate the vector or matrix transpose.
+If $y$ and $z$ differ by a polynomial from $\mathcal{Z}$ then $y = z + p$ where $p \in \mathcal{Z}$ and also $\coef(y) = \coef(z) + \coef(p)$.
 %\footnote{In group theory terminology $\coef(\cdot)$ coupled with vector addition is called a \term{group homomorphism}.\index{group homomorphism}}. 
@@
 \[
-L_{m+1} = \{ \coef(p) \mid p \in \mathcal{Z} \},
+L_{m+1} = \{ \coef(p) \mid p \in \mathcal{Z} \}
 \]
@@
 \end{equation}
-where $\angle{z}$ denotes the complex argument of the complex number $z$, and 
+where $\angle$ denotes the complex argument (or phase), and 
 \[
@@
 \] 
-are random variables representing the phase noise induced by the $X_n$~\cite{Tretter1985,Quinn2009_dasp_phase_only_information_loss}.
-If the distribution of $X_n$ is circularly symmetric (i.e., the angle $\angle X_n$ is uniformly distributed on $[-\pi, \pi)$ and is independent of the magnitude $\abs{X_n}$) then the distribution of $\Phi_n$ is the same as the distribution of $\tfrac{1}{2\pi}\angle(1 + \rho^{-1}X_n)$.
-If the $X_1, \dots, X_N$ are circularly symmetric and identically distributed, then $\Phi_1, \dots, \Phi_n$ are also identically distributed.
+are random variables representing the `phase noise' induced by the $X_n$.
+If the distribution of $X_n$ is circularly symmetric (the phase $\angle X_n$ is uniformly distributed on $[-\pi, \pi)$ and is independent of the magnitude $\abs{X_n}$) then the distribution of $\Phi_n$ is the same as the distribution of $\tfrac{1}{2\pi}\angle(1 + \rho^{-1}X_n)$.
+If $X_1, \dots, X_N$ are circularly symmetric and identically distributed, then $\Phi_1, \dots, \Phi_n$ are also identically distributed.
 
@@
 \end{equation}
-The least squares unwrapping estimator is defined as those coefficients $\widehat{\mu}_0, \dots, \widehat{\mu}_m$ that minimise $SS$ over the identifiable region $B$. That is, the LSU estimator is,
+The least squares unwrapping estimator is defined as those coefficients $\widehat{\mubf} = [\widehat{\mu}_0, \dots, \widehat{\mu}_m]$ that minimise $SS$ over the identifiable region $B$, i.e., the LSU estimator is,
 \begin{equation}\label{eq:hatmubfLSUdefn}
@@
 
-It is shown in~\cite{McKilliam2009asilomar_polyest_lattice} and \cite[Sec~8.1]{McKilliam2010thesis} how this minimisation problem can be posed as a nearest lattice point problem, and we briefly discuss this here.
-Write $SS$ as
-\[
-SS(\mubf) = \sum_{n=1}^{N}{ \left( \Theta_{n} - W_n - \sum_{k = 0}^{m}{\mu_k n^k} \right)^2 },
-\]
-where $W_n = \round{\Theta_{n} - \sum_{k = 0}^{m}{\mu_k n^k} }$ for $n = 1, \dots, N$ are integers called \emph{wrapping variables}.
-If we consider $W_1,\dots,W_N$ as nuisance parameters to be estimated then $SS$ can be written as a function of both $\mu_0,\dots,\mu_m$ and $W_1,\dots,W_N$.
-The least squares unwrapping estimator is then found by jointly minimising over $\mu_0,\dots,\mu_m$ and $W_1,\dots,W_N$.
-This joint minimisation problem can be solved by computing a nearest point in a lattice.
-To see this write $SS$ as a function of both $\mu_0,\dots,\mu_m$ and $W_1,\dots,W_N$ using vectors,
-\begin{equation}\label{eq:SS_in_vector_form_with_u_nuisance}
-SS(\mubf, \wbf) = \| \thetabf - \Xbf\mubf - \wbf \|^2,
-\end{equation}
-where $\|\cdot\|^2$ denotes the squared Euclidean norm, and $\Xbf$ is the rectangular Vandermonde matrix
-\begin{equation}\label{eq:Xvandermonde}
-\Xbf = \left[ \begin{array}{ccccc} 
-1 & 1 & 1 & \cdots & 1 \\
-1 & 2 & 4 & \cdots & 2^m \\
-\vdots & \vdots  & \vdots & \ddots & \vdots \\
-1 & N & N^2 & \cdots & N^m \\
-\end{array} \right],
-\end{equation}
-and where the column vectors $\wbf = [W_1, \dots, W_N]^\prime$ and $\thetabf = [\Theta_1, \dots, \Theta_N]^\prime$.   For fixed $\wbf$ the minimiser of $SS(\mubf, \wbf)$ with respect to $\mubf$ is
-\begin{equation} \label{eq:hat_theta_cal}
-\mubf(\wbf) = (\Xbf^\prime\Xbf)^{-1}\Xbf^\prime( \thetabf - \wbf) = \Xbf^+( \thetabf - \wbf),
-\end{equation}
-where $\Xbf^+ = (\Xbf^\prime\Xbf)^{-1}\Xbf^\prime$ is the Moore-Penrose pseudoinverse of $\Xbf$.
-
-We may thus compute $\widehat{\mubf}$ as $\widehat{\mubf} = \mu(\widehat{\wbf})$, where $\widehat{\wbf}$ is the minimiser of
-\[
-SS(\wbf) = SS(\mubf(\wbf),\wbf) = \| \Qbf\thetabf - \Qbf\wbf \|^2,
-\]
-where $\Qbf = \Ibf - \Xbf\Xbf^+$ and $\Ibf$ is the $N \times N$ identity matrix\footnote{We have slightly abused notation here by reusing $SS$.
-This should not cause any confusion as $SS(\mubf)$ and $SS(\mubf,\wbf)$ and $SS(\wbf)$ are easily told apart by their inputs.}.
-
-Let $\Lambda$ be the lattice generated by $\Qbf$, that is $\Lambda = \{\Qbf\xbf \mid \xbf \in \ints^N \}$\footnote{The $N \times N$ projection matrix $\Qbf$ does not have full rank, it has rank $N-m-1$, and for this reason it is not strictly a generator matrix.
-However, a generator can be constructed by removing $m+1$ consecutive columns from $\Qbf$.
-In our implementation the last $m+1$ columns are removed.}.
-It follows that $\Qbf \wbf$ is a lattice point in $\Lambda$ and that minimising $SS(\wbf)$ is equivalent to finding the nearest lattice point in $\Lambda$ to $\Qbf\thetabf$.
-Denote this point by $\Qbf\widehat{\wbf}$.
-Now the estimate $\widehat{\mubf}$ is given by substituting $\widehat{\wbf}$ for $\wbf$ in~\eqref{eq:hat_theta_cal}.
-After this procedure it is possible that the $\widehat{\mubf}$ obtained is not in the identifiable region but is instead an aliased version of the desired estimate.
-This can be resolved by computing $\dealias(\widehat{\mubf})$.
-There exist algorithms that can compute a nearest point in $\Lambda$ that require a number of operations that is polynomial in $N$~\cite[Sec 4.3]{McKilliam2010thesis}.
+It is shown in~\cite{McKilliam2009asilomar_polyest_lattice} how this minimisation problem can be posed as that of computing a nearest lattice point in a particular lattice.
+Polynomial time algorithms that compute the nearest point are described in~\cite[Sec 4.3]{McKilliam2010thesis}.
 Although polynomial in complexity, these algorithms are not fast in practice.
 The existence of practically fast nearest point algorithms for these lattices is an interesting open problem.
+We focus on the asymptotic statistical properties of the LSU estimator, rather than computational aspects in this paper.
 
@@
 
-The \emph{intrinsic mean} or \emph{Fr\'{e}chet mean} of $\Phi_n$ is defined as~\cite{McKilliam_mean_dir_est_sq_arc_length2010,bwhk07a,Bhattacharya_int_ext_means_2003,Bhattacharya_int_ext_means_2005},
+The \emph{intrinsic mean} or \emph{Fr\'{e}chet mean} of $\Phi_n$ is defined as~\cite{McKilliam_mean_dir_est_sq_arc_length2010,Bhattacharya_int_ext_means_2003,Bhattacharya_int_ext_means_2005},
 \begin{equation}\label{eq:intrmeandefn}
@@
 \[
-\sigma_{\text{intr}}^2 = E\fracpart{\Theta - \mu_{\text{intr}}}^2 = \min_{\mu \in [-\nicefrac{1}{2}, \nicefrac{1}{2})} \expect \fracpart{\Phi_n - \mu}^2,
+\sigma_{\text{intr}}^2 = \expect\fracpart{\Theta - \mu_{\text{intr}}}^2 = \min_{\mu \in [-\nicefrac{1}{2}, \nicefrac{1}{2})} \expect \fracpart{\Phi_n - \mu}^2,
 \]
@@
 Denote the elements of $\widehat{\lambdabf}_N$ by $\widehat{\lambda}_{0,N}, \dots, \widehat{\lambda}_{m,N}$.
-Suppose $\Phi_1, \dots, \Phi_N$ are independent and identically distributed with zero intrinsic mean, intrinsic variance $\sigma^2$, and pdf $f$, then:
+Suppose $\Phi_1, \dots, \Phi_N$ are independent and identically distributed with zero intrinsic mean, intrinsic variance $\sigma^2$, and probability density function $f$, then:
 \begin{enumerate}
 \item (Strong consistency) $N^k \widehat{\lambda}_{k,N}$ converges almost surely to $0$ as $N\rightarrow\infty$ for all $k = 0, 1, \dots, m$.
-\item (Asymptotic normality) If $f(\fracpart{x})$ is continuous at $x = -\nicefrac{1}{2}$ and $f(-\nicefrac{1}{2}) < 1$ then the distribution of
+\item (Asymptotic normality) If $f(\fracpart{x})$ is continuous at $x = -\nicefrac{1}{2}$ and if $f(-\nicefrac{1}{2}) < 1$ then the distribution of the vector
 \[
@@
 \]
-where $\Cbf$ is the $(m+1)\times (m+1)$ Hilbert matrix with elements $C_{ik} = 1/(i + k + 1)$ for $i,k = \{0, 1, \dots, m\}$.
+where $\Cbf$ is the $m+1$ by $m+1$ Hilbert matrix with elements $C_{ik} = 1/(i + k + 1)$ for $i,k \in \{0, 1, \dots, m\}$.
 \end{enumerate}
@@
 Section~\ref{sec:strongconstproof} proves strong consistency and Section~\ref{sec:centlimitproof} proves asymptotic normality.
-The proofs for the case of $m=0$ were given in~\cite{McKilliam_mean_dir_est_sq_arc_length2010} and for the case when $m=1$ were given in~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.
+Proofs for the case when $m=0$ were given in~\cite{McKilliam_mean_dir_est_sq_arc_length2010} and for the case when $m=1$ were given in~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.
 The proofs here take a similar approach, but require new techniques.
-Before giving the proof we make the following remarks.
-
-The theorem gives conditions on the \emph{dealiased} difference $\dealias(\tilde{\mubf} - \widehat{\mubf})$ between the true coefficients $\tilde{\mubf}$ and the estimated coefficients $\widehat{\mubf}$ rather than directly on the difference $\tilde{\mubf} - \widehat{\mubf}$.
+The proof of asymptotic normality in~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009} contains a subtle error that is amened by the proof given in this paper.
+%Before begining the proof we first discuss the assumptions made in the theorem statement.
+
+The theorem gives conditions on $\dealias(\tilde{\mubf} - \widehat{\mubf})$ rather than directly on the difference $\tilde{\mubf} - \widehat{\mubf}$.
 To see why this makes sense, consider the case when $m=0$, $\tilde{\mu}_0 = -0.5$ and $\widehat{\mu}_0 = 0.49$, so that $\tilde{\mu}_0 - \widehat{\mu}_0 = -0.99$.
-However, the two phases are obviously close, since the phase $\pm 0.5$ are actually the same.
-In this case
-\[
-\dealias(\tilde{\mu}_0 - \widehat{\mu}_0) = \fracpart{\tilde{\mu}_0 - \widehat{\mu}_0} = 0.01.
-\] 
+However, the two phases are obviously close, since the phases $\pm 0.5$ are actually the same.
+In this case $\dealias(\tilde{\mu}_0 - \widehat{\mu}_0) = \fracpart{\tilde{\mu}_0 - \widehat{\mu}_0} = 0.01$.
 The same reasoning holds for $m > 0$.
@@
 In the interest of simplicity we only consider the case when $\Phi_1, \dots, \Phi_N$ are identically distributed here.
-If $X_n$ is circularly symmetric with density function nonincreasing with magnitude $\abs{X_n}$, then, the corresponding $\Phi_n$ necessarily has zero intrinsic mean~\cite[Theorem 5.2, page 78]{McKilliam2010thesis}.
+If $X_n$ is circularly symmetric with density function nonincreasing with magnitude $\abs{X_n}$, then, the corresponding $\Phi_n$ necessarily has zero intrinsic mean~\cite[Theorem 5.2]{McKilliam2010thesis}.
 Thus, our theorem covers commonly used distributions for $X_1, \dots, X_N$, such as the normal distribution.
 
-The proof of asymptotic normality places requirements on the pdf $f$ of the phase noise.
-The requirement that $\Phi_1, \dots, \Phi_N$ have zero intrinsic mean implies that $f(-\nicefrac{1}{2}) \leq 1$~\cite[Lemma~1]{McKilliam_mean_dir_est_sq_arc_length2010}, so the only case not handled is when equality holds, i.e., when $f(-\nicefrac{1}{2}) = 1$ or when $f(\fracpart{x})$ is discontinuous at $x = -\nicefrac{1}{2}$.
-In this exceptional case other expressions for the asymptotic variance can be found (similar to \cite[Theorem 3.1]{Hotz_circle_means_2011}), but this comes at a substantial increase in complexity and for this reason we have omitted them.
+The proof of asymptotic normality places requirements on the probability density function $f$ of the phase noise $\Phi_1, \dots, \Phi_N$.
+The requirement that $\Phi_1, \dots, \Phi_N$ have zero intrinsic mean implies $f(-\nicefrac{1}{2}) \leq 1$~\cite[Lemma~1]{McKilliam_mean_dir_est_sq_arc_length2010}, so the only case not handled is when equality holds, i.e., when $f(-\nicefrac{1}{2}) = 1$ or when $f(\fracpart{x})$ is discontinuous at $x = -\nicefrac{1}{2}$.
+In this exceptional case other expressions for the asymptotic variance can be found (similar to \cite[Theorem 3.1]{Hotz_circle_means_2011}), but this comes at a substantial increase in complexity and we have omitted them for this reason.
 %Distributions that do not satisfy these requirements are unlikely to be needed in practice.
@@
 We shall show that $N^k\widehat{\lambda}_{k,N}\rightarrow0$ almost surely as $N\rightarrow\infty$ for all $k = 0,1, \dots, m$ and from this the proof of strong consistency follows.
-
-We first show, in Lemma~\ref{lem:vn}, that $S_N(\lambdabf)$ converges almost surely to its expectation, $\expect S_N(\lambdabf)$,  uniformly over $\lambdabf \in B$.
-We use this to prove that $\widehat{\lambdabf}_N$, the minimiser of $S_N$, converges almost
-surely to the minimiser of $\expect S_N$. %We can then reason about the minimiser of $S_N$ using the minimiser of $\expect S_N$.  
-This method of proof is common in the literature and is connected with what is called the \emph{uniform law of large numbers}~\cite{Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.
-
 Let
 \[
-V_N(\lambdabf) =  \expect S_N(\lambdabf) = \frac{1}{N}\sum_{n=1}^{N} \expect \fracpart{  \Phi_{n}+\sum_{k = 0}^{m}{\lambda_k n^k}}  ^{2}
-\]
-and
- \[
- D_{N}\left( \lambdabf \right)  =  S_N(\lambdabf) - V_N(\lambdabf).
- \]
-%be the difference between $S_N$ and its expectation $V_N =  \expect S_N$.  
-We use order notation in the standard way, that is, for functions $h$ and $g$, we write $h(N) = O(g(N))$ to mean that there exists a constant $K > 0$ and a finite $N_0$ such that $h(N) \leq K g(N)$ for all $N > N_0$.
-We use $\prob$ to denote probability.
+V_N(\lambdabf) =  \expect S_N(\lambdabf) = \frac{1}{N}\sum_{n=1}^{N} \expect \fracpart{  \Phi_{n}+\sum_{k = 0}^{m}{\lambda_k n^k}}  ^{2}.
+\]
+It will follow that $\sabs{S_N(\lambdabf) - V_N(\lambdabf)}$ converges almost sure to zero as $N \rightarrow \infty$ uniformly in $B$, that is, 
+ \begin{equation}\label{eq:SNVNunifmlln}
+\sup_{\lambdabf \in B}\sabs{S_N(\lambdabf) - V_N(\lambdabf)} \rightarrow 0
+ \end{equation}
+almost surely as $N\rightarrow\infty$.
+This type of result has been called a uniform law of large numbers~\cite{Pollard_conv_stat_proc_1984} and follows from standard techniques.
+We give a full proof of~\eqref{eq:SNVNunifmlln} in Appendix~\ref{app:uniform-law-large}.
+We now concentrate attention on the minimiser of $V_N$.
+Because $\Phi_n$ has zero intrinsic mean
+\begin{equation}\label{eq:Efracpartmined}
+\expect \fracpart{ \Phi_n + z }^{2}
+\end{equation}
+is minimised uniquely at $z = 0$ for $z \in [-\nicefrac{1}{2}, \nicefrac{1}{2})$.  Since the intrinsic variance of $\Phi_n$ is $\sigma^2$, when $z = 0$,
+\begin{equation}\label{eq:Efracpartphi}
+\expect\fracpart{\Phi_1+z}^{2} = \expect\fracpart{\Phi_1}^{2} = \sigma^2,
+\end{equation}
+and so the minimum attained value is $\sigma^2$.
+
+\begin{lemma}\label{lem:ES_Nminimisedzero}
+For $\lambdabf \in B$ the function $V_N(\lambdabf)$ is minimised uniquely at $\zerobf$, the vector of all zeros.  At this minimum $V_N(\zerobf) = \sigma^2$.
+\end{lemma}
+\begin{proof}
+Let $z$ be the polynomial
+\[
+z(n) = \lambda_0 + \lambda_1 n + \lambda_2 n^2 + \dots + \lambda_m n^m.
+\]
+Then
+\begin{align*}
+V_N(\lambdabf) = \frac{1}{N}\expect\sum_{n=1}^{N}\fracpart{ \Phi_n + \sum_{k=0}^m{\lambda_k n^k} }^2 = \frac{1}{N}\sum_{n=1}^{N}\expect\fracpart{ \Phi_n + \fracpart{z(n)} }^2.
+\end{align*}
+We know that $\expect\fracpart{ \Phi_n + \fracpart{z(n)} }^2$ is minimised uniquely when $\fracpart{z(n)} = 0$ at which point it takes the value $\sigma^2$.
+Now $\fracpart{z(n)}$ is equal to zero for all integers $n$ if and only if $z \in \mathcal{Z}$, or equivalently if $\coef(z)$ is a lattice point in $L_{m+1}$.
+By definition $B$ contains precisely one lattice point from $L_{m+1}$, this being the origin $\zerobf$.
+Therefore $V_N$ is minimised uniquely at $\zerobf$, at which point it takes the value $\sigma^2$.
+\end{proof}
+
+\begin{lemma} \label{lem:ESNconv}
+$\sabs{V_N(\widehat{\lambdabf}_N) - \sigma^2} \rightarrow 0$ almost surely as $N \rightarrow \infty$.
+\end{lemma}
+\begin{proof}
+By definition $\widehat{\lambdabf}_N =\arg\min_{\lambdabf\in B} S_N(\lambdabf)$ so $0 \leq S_{N}(\zerobf) - S_N(\widehat{\lambdabf}_N)$.
+Also, because $V_N$ is minimised at $\zerobf$, it follows that $0 \leq V_N(\widehat{\lambdabf}_N) - V_N(\zerobf)$.
+Thus,
+\begin{align*}
+0 &\leq V_N(\widehat{\lambdabf}_N) - V_N(\zerobf) \\
+ &\leq V_N(\widehat{\lambdabf}_N) - V_N(\zerobf) + S_{N}(\zerobf) - S_N(\widehat{\lambdabf}_N)   \\
+&\leq \sabs{ V_N(\widehat{\lambdabf}_N) - S_N(\widehat{\lambdabf}_N) } + \sabs{ S_{N}(\zerobf) - V_N(\zerobf) }
+\end{align*}
+which converges almost surely to zero as $N\rightarrow\infty$ as a result of~\eqref{eq:SNVNunifmlln}.
+\end{proof}
+
+% \begin{lemma}\label{lem:ABprob}
+% Let $A$ and $B$ be positive random variables.  Then
+% \[
+% \prob( A + B > \epsilon ) \leq \prob( A > \tfrac{\epsilon}{2}) + \prob( A > \tfrac{\epsilon}{2}).
+% \]
+% \end{lemma}
+% \begin{proof}
+% Write 
+% \begin{equation}\label{eq:A+B>epsilon}
+% \prob( A + B > \epsilon ) = 1 - \prob( A + B \leq \epsilon ).
+% \end{equation}
+% If $A \leq \tfrac{\epsilon}{2}$ and $B \leq \tfrac{\epsilon}{2}$ then $A + B \leq \epsilon$, so
+% \begin{align*}
+% \prob( A + B \leq \epsilon ) &\geq \prob( A \leq \tfrac{\epsilon}{2} \;\; \text{and} \;\; B \leq \tfrac{\epsilon}{2} ) \\
+% &= 1 - \prob( A > \tfrac{\epsilon}{2} \;\; \text{or} \;\; B > \tfrac{\epsilon}{2} ) \\
+% &\geq 1 - \prob( A > \tfrac{\epsilon}{2}) - \prob( B > \tfrac{\epsilon}{2} ).
+% \end{align*}
+% Subtituting this into~\eqref{eq:A+B>epsilon} gives the required result.
+% \end{proof}
+
+We have now shown that $V_N$ is uniquely minimised at $\zerobf$, that $V_N(\zerobf) = \sigma^2$, and that $V_N(\widehat{\lambdabf}_N)$ converges almost surely to $\sigma^2$.
+These results are enough to show that $\widehat{\lambdabf}_N$ converges almost surely to zero.
+However, this tells us nothing about the rate at which the components of $\widehat{\lambdabf}_N$ approach zero as required by Theorem~\ref{thm:asymp_proof}.
+To prove these stronger properties we need some preliminary results about arithmetic progressions, and from the calculus of finite differences.
+
+ %\begin{equation}
+ %E\left(  \sum_{n=1}^{N}z_{n}\right)  ^{2\beta}=\sum_{n_{1},\ldots,n_{2\beta
+ %}=1}^{N}E\left(  \prod\limits_{j=1}^{2\beta}z_{n_{j}}\right)  .
+ %\label{eq:markapp}%
+ %\end{equation}
+ %Since the terms%
+ %\[
+ %E\left(  \prod\limits_{j=1}^{2\beta}z_{n_{j}}\right)
+ %\]
+ %in this containing any $n_{j}$ only once will have $0$ mean, and since the
+ %$z_{j}$ are bounded, it follows that%
+ %\[
+ %\frac{1}{N^{\beta}}E\left(  \sum_{n=1}^{N}z_{n}\right)  ^{2\beta}=O\left(
+ %1\right)  ,
+ %\]
+ %as $N\rightarrow\infty,$ the dominant terms coming from the $z_{j}$ in
+ %$\left(  \ref{eq:markapp}\right)  $ occurring in pairs.
+
+Let $W = \{1,2,\dots, N\}$ and let $K$ be a subset of $W$.  For any integer $h$, let
+\begin{equation} \label{eq:S(h,G)def} 
+A(h,K) = \big\{ n \mid n + ih \in K \;\forall\; i \in \{0,1,\dots,m\} \big\}
+\end{equation}
+be the set containing all integers $n$ such that the arithmetic progression
+\[
+n, \,\, n + h, \,\, n + 2h, \,\, \dots, \,\, n + mh
+\]
+of length $m+1$ is contained in the subset $K$.
+If $K$ is a small subset of $W$ then $A(h,K)$ might be empty, i.e., there may be no arithmetic progression $n, n+h, \dots$ of length $m+1$ in $K$.
+However, the next two lemmas and the following corollary will show that if $K$ is sufficiently large then it always contains at least one arithmetic progression (for all sufficiently small $h$) and therefore $A(h,K)$ is not empty.
+Denote by $K \backslash \{r\}$ the set $K$ with the element $r$ removed.
+
+\begin{lemma} \label{lem:S(h,G/r)size}
+Let $r \in K$.  For any $h$, removing $r$ from $K$ removes at most $m+1$ arithmetic progressions $n, n+h, \dots n+mh$ of length $m+1$.  That is,
+\[
+|A(h,K \backslash \{r\})| \geq |A(h,K)| - (m+1).
+\]
+\end{lemma}
+\begin{proof}
+The proof follows because there are at most $m+1$ integers, $n$, such that $n+ih = r$ for some $i \in \{0,1,\dots,m\}$.
+That is, there are at most $m+1$ arithmetic progressions of type $n, n+h, \dots n+mh$ that contain $r$.
+\end{proof}
+
+ \begin{lemma} \label{lem:S(h,K)size}
+ Let $K \subseteq W$.  Then $|A(h,K)| \geq N - mh - (N - |K|)(m+1)$.
+ \end{lemma}
+ \begin{proof}
+ Note that $|A(h,W)| = N - mh$.  The proof follows by starting with $A(h,W)$ and applying Lemma~\ref{lem:S(h,G/r)size} 
+\[
+|W|-|K| = N - |K|
+\] 
+times. That is, $K$ can be constructed by removing $N - |K|$ elements from $W$ and this removes at most $(N - |K|)(m+1)$ arithmetic progressions from $A(h,W)$.
+ \end{proof}
+
+ \begin{corollary} \label{cor:S(h,K)>0}
+ Let $K \subseteq W$ such that $|K| > \frac{2m+1}{2m+2}N$.
+Then for all $h$ such that $1\leq h \leq\frac{N}{2m}$ the set $K$ contains at least one arithmetic progression $n, n+h, \dots, n+mh$ of length $m+1$.
+That is, $|A(h,K)| > 0$.
+ \end{corollary}
+ \begin{proof}
+ By substituting the bounds $|K| > \frac{2m+1}{2m+2}N$ and $h \leq\frac{N}{2m}$ into the inequality from Lemma~\ref{lem:S(h,K)size} we immediately obtain $|A(h,K)| > 0$.
+ \end{proof}
+
+The next result we require comes from the calculus of finite differences. For any function $d(n)$ mapping $\reals$ to $\reals$, let 
+\[
+\Delta_h d(n) = d(n+h) - d(n)
+\] 
+denote the first difference with interval $h$. Let 
+\begin{align*}
+\Delta_h^2 d(n) = \Delta_h d(n+h) - \Delta_h d(n) = d(n+2h) - 2d(n+h) + d(n)
+\end{align*}
+denote the second difference with interval $h$ and similarly let 
+\begin{equation}\label{eq:mthdiffformula}
+\Delta_h^r d(n) = \Delta_h^{r-1} d(n+h) - \Delta_h^{r-1} d(n) = \sum_{k=0}^{r}\binom{r}{k}(-1)^{r-k}d(n+kh)
+\end{equation}
+denote the $r$th difference with interval $h$. Since $\sum_{k=0}^{r}\binom{r}{k} = 2^r$ it follows that $\Delta_h^r d(n)$ can be represented by adding and subtracting the 
+\[
+d(n), \,\, d(n+h), \,\, \dots, \,\, d(n+kh)
+\] 
+precisely $2^r$ times.
+
+The operator $\Delta_h$ has special properties when applied to polynomials. If $d(n) = a_r n^r + \dots + a_0$ is a polynomial of order $r$ then
+ \begin{equation} \label{eq:mfinitediffpoly}
+ \Delta_h^r d(n) = h^r r! a_r. 
+ \end{equation}
+So, the $r$th difference of the polynomial is a constant depending on $h$, $r$ and the $r$th coefficient $a_r$~\cite[page 51]{Jordan_Calculus_of_finite_difference_1965}.
+We can now continue the proof of strong consistency.
+The next lemma is a key result.
+
+\begin{lemma}\label{lem:moran2}
+Suppose $\lambdabf_1, \lambdabf_2,\dots$ is a sequence of vectors from $B$ with $V_N(\lambdabf_N) - \sigma^2\rightarrow 0$ as $N\rightarrow\infty$.
+Then the elements $\lambda_{0,N}, \dots \lambda_{m,N}$ of $\lambdabf_N$ satisfy $N^k\lambda_{k, N}\rightarrow0$ as $N\rightarrow\infty$.
+\end{lemma}
+\begin{proof}
+Define the function
+\begin{equation}\label{eq:gz}
+g(z) = \expect\fracpart{\Phi_1 + z}^2 - \sigma^2
+\end{equation}
+which is continuous in $z$. Because of~\eqref{eq:Efracpartmined} and~\eqref{eq:Efracpartphi}, $g(z) \geq 0$ with equality only at $z = 0$ for $z \in [-\nicefrac{1}{2}, \nicefrac{1}{2})$. Now
+\[
+V_N(\lambdabf_N) - \sigma^2 = \frac{1}{N}\sum_{n=1}^{N} g\left( \fracpart{ \sum_{k=0}^{m}{n^k \lambda_{k,N}} } \right) \rightarrow 0
+\]
+as $N \rightarrow \infty$. Let
+\[
+z_N(n) = \lambda_{0,N} + \lambda_{1,N} n + \lambda_{2,N} n^2 + \dots + \lambda_{m,N} n^m
+\]
+so that $V_N(\lambdabf_N) - \sigma^2 = \frac{1}{N}\sum_{n=1}^{N} g\left( \fracpart{z_N(n)} \right) \rightarrow 0$.
+%At this stage we can \emph{see} the proof.
+Note that, due to the definition of the identifiable region, $B$, the only circular polynomial $\phi$ with values that are all zeros (and hence sends the sum above to zero) is the circular polynomial whose coefficients are all zeros.
+Obviously, this observation does not constitute a proof, and in particular gives us little indication as to the order of convergence of the coefficients as required by the theorem.
+To make the proof
+Choose constants 
+\[
+c = \frac{2m+1}{2m+2} \qquad \text{and} \qquad 0 < \delta < \frac{1}{2^{2m+1}}
+\]
+and define the set $K_{N}=\left\{  n\leq N \mid \sabs{\fracpart{z_N(n)}} < \delta \right\}$.
+There exists $N_0$ such that for all $N > N_0$ the number of elements in $K_N$ is at least $cN$.
+Too see this, suppose that $\sabs{K_N} < cN$, and let $\gamma$ be the minimum value of $g$ over $[-\nicefrac{1}{2},-\delta] \cup [\delta, \nicefrac{1}{2})$.
+Because $g(0) = 0$ is the unique minimiser of $g$, then $\gamma$ is strictly greater than $0$ and
+\begin{align*}
+V_N(\lambdabf_N) - \sigma^2 = \frac{1}{N}\sum_{n=1}^{N} g\left( \fracpart{z_N(n)} \right) \geq \frac{1}{N}\sum_{n\in K_N} \gamma = (1-c)\gamma,
+\end{align*}
+violating that $V_N(\lambdabf_N) - \sigma^2$ converges to zero as $N \rightarrow \infty$.  We will assume $N > N_0$ in what follows.
+
+From Corollary \ref{cor:S(h,K)>0} it follows that for all $h$ satisfying $1\leq h \leq\frac{N}{2m}$ the set $A(h,K_N)$ contains at least one element, that is, there exists $n' \in A(h,K_N)$ such that all the elements from the arithmetic progression $n', n'+h, \dots, n' + mh$ are in $K_N$ and therefore
+\[
+\sabs{\fracpart{z_N(n')}},\,\, \sabs{\fracpart{z_N(n'+h)}}, \,\, \dots, \,\, \sabs{\fracpart{z_N(n'+mh)}} 
+\] 
+are all less than $\delta$.  Because the $m$th difference is a linear combination of $2^{m}$ elements (see~\eqref{eq:mthdiffformula}) from 
+\[
+\fracpart{z_N(n')}, \,\, \fracpart{z_N(n'+h)}, \,\, \dots, \,\, \fracpart{z_N(n'+mh)}
+\]
+all with magnitude less than $\delta$ we obtain, from Lemma~\ref{lem:fracpartsumanddelta},
+\begin{equation}\label{eq:Deltazfracbound}
+|\fracpart{\Delta_h^m z_N(n')}| \leq |\Delta_h^m \fracpart{ z_N(n')}| \leq \Delta_h^m |\fracpart{ z_N(n')}|< 2^{m}\delta.
+\end{equation}
+From \eqref{eq:mfinitediffpoly} it follows that the left hand side is equal to a constant involving $h$, $m$ and $\lambda_{m,N}$ giving the bound
+\begin{equation}\label{eq:startiterativearg}
+|\fracpart{ h^m m! \lambda_{m,N} }|  = |\fracpart{   \Delta_h^m z_N(n') }| < 2^m\delta
+\end{equation}
+for all $h$ satisfying $1\leq h \leq\frac{N}{2m}$. Setting $h = 1$ and recalling from~\eqref{eq:identifiability} that $\lambda_{m,N} \in [-\tfrac{0.5}{m!}, \tfrac{0.5}{m!})$, we have
+ \[
+ |\fracpart{ m! \lambda_{m,N} }| = | m! \lambda_{m,N} |< 2^m\delta.
+ \]
+Now, because we chose $\delta < \tfrac{1}{2^{2m}}$ it follows that 
+\[
+| \lambda_{m,N} |< \tfrac{2^m}{m!}\delta < \tfrac{1}{m! 2^{m+1}}.
+\]
+So, when $h = 2$, 
+\[
+|\fracpart{ 2^m m! \lambda_{m,N} }| = | 2^m m! \lambda_{m,N} |< 2^m\delta
+\]
+because $2^m m! \lambda_{m,N} \in [-0.5, 0.5)$. Therefore
+\[
+| \lambda_{m,N} |< \frac{1}{m!}\delta < \frac{1}{m! 2^{2m+1}}.
+\]
+Now, with $h = 4$, we similarly obtain 
+\[
+|\fracpart{ 4^m m! \lambda_{m,N} }| = | 4^m m! \lambda_{m,N} |< 2^m\delta
+\]
+and iterating this process we eventually obtain 
+\[
+| \lambda_{m,N} | < \frac{2^m}{2^{um} m!}\delta
+\]
+where $2^u$ is the largest power of 2 less than or equal to $\tfrac{N}{2m}$.  So $2^{u+1} > \frac{N}{2m}$ and substituting this into the inequality above gives 
+ \[ 
+ |\lambda_{m,N}| < \frac{2^m}{ m! \left(\frac{N}{4m}\right)^m}\delta
+ \]
+from which it follows that 
+ \begin{equation}\label{eq:enditerativearg}
+ N^m|\lambda_{m,N}| < \frac{2^{2m+m}m^m}{m!}\delta
+ \end{equation}
+for all $N > N_0$.  As $\delta$ is arbitrary, $N^m \lambda_{m,N} \rightarrow 0$ as $N\rightarrow \infty$.
+
+We have now shown that the highest order coefficient $\lambda_{m,N}$ converges as required.
+The remaining coefficients will be shown to converge by induction.
+Assume that $N^k \lambda_{k,N} \rightarrow 0$ for all $k=r+1, r+2, \dots, m$, that is, assume that the $m-r$ highest order coefficients all converge as required.
+Let
+\[
+z_{N,r}(n) = \lambda_{0,N} + \lambda_{1,N} n + \lambda_{2,N} n^2 + \dots + \lambda_{r,N} n^r.
+\]
+Because the $m-r$ highest order coefficients converge we can write $z_N(n) = z_{N,r}(n) + \gamma(n)$ where $\sup_{n\in\{1,\dots,N\}}\abs{\gamma(n)} \rightarrow 0$ as $N\rightarrow\infty$.
+Now the bound from \eqref{eq:Deltazfracbound}, but applied using the $r$th difference, gives
+ \begin{equation}\label{eq:zrbound}
+\left|\fracpart{  \Delta_h^r z_N(n')}\right| = \left|\fracpart{  \Delta_h^r\gamma(n') + \Delta_h^r z_r(n') }\right| = |\fracpart{ \epsilon + h^r r! \lambda_{r,N} }| < 2^r\delta,
+ \end{equation}
+ where
+\[
+\epsilon = \Delta_h^r \gamma(n') \leq 2^r \sup_{n\in\{1,\dots,N\}}\abs{\gamma(n)} \rightarrow 0
+\] 
+as $N\rightarrow\infty$.
+Choose $\delta$ and $\epsilon$ such that $2^r\delta < \tfrac{1}{4}$ and $|\epsilon| < \tfrac{1}{4}$.
+Then, from \eqref{eq:zrbound} and from Lemma~\ref{lem:fracpartinternalsumlessdelta},
+\[
+\abs{\fracpart{ h^r r! \lambda_{r,N} }} < 2^r\delta + \abs{\epsilon}
+\]
+for all $h$ such that $1 \leq h \leq \tfrac{N}{2m}$.
+Choosing $2^r\delta + |\epsilon| < \frac{1}{2^{2r+1}}$ and using the same iterative process as for the highest order coefficient $\lambda_{m,N}$  (see~\eqref{eq:startiterativearg}~to~\eqref{eq:enditerativearg}) we find that $N^r \lambda_{r,N} \rightarrow 0$ as $N\rightarrow\infty$.
+The proof now follows by induction.
+ \end{proof}
+
+% \begin{lemma}\label{sec:proof-strong-cons}
+% For any constants $0 \leq c < 1$ and $\delta>0$ there exists an $N_{0}$ such that for all $N > N_0$ the proportion of $\fracpart{z_N(n)}$ with magnitude less than $\delta$ is greater than $c$.
+That is, the set
+% \[
+% K_{N}=\left\{  n\leq N \mid \vert \fracpart{z(n)} \vert < \delta \right\} 
+% \]
+% has more than $c N$ elements for all $N > N_0$.
+% \end{lemma}
+% \begin{proof}
+% Assume not.
+Then for every $N_0$ there exists an $N > N_0$ such that there are more than $(1-c)N$ integers from $1$ to $N$ with $|\fracpart{z(n)}| > \delta$.
+Let $\gamma$ be the minimum value of $g$ from~(\ref{eq:gz}) over the interval given by the union $[-\nicefrac{1}{2},-\delta] \cup [\delta, \nicefrac{1}{2})$.
+Because $g$ is minimised uniquely at $0$ then $\gamma$ is strictly greater than $0$ and the sum
+% \[
+% \frac{1}{N}\sum_{n=1}^{N} g\left( \fracpart{z(n)} \right) \geq (1-c)\gamma
+% \]
+% with $(1-c)\gamma$ a positive constant. This violates the fact that $g$ converges to zero as $N \rightarrow \infty$ and the lemma is true by contradiction.
+% \end{proof}
+
+ \begin{lemma} \label{lem:fracpartsumanddelta}
+ Let $a_1, a_2, \dots, a_r$ be $r$ real numbers such  that $\left|\fracpart{a_n}\right| < \delta$ for all $n = 1,2,\dots,r$.  Then $\left|\fracpart{\sum_{n=1}^r{a_n}}\right| < r\delta.$
+ \end{lemma}
+ \begin{proof}
+ If $\delta > \tfrac{1}{2r}$ the proof is trivial as $\left|\fracpart{\sum_{n=1}^r{a_n}}\right| \leq \tfrac{1}{2}$ for all $a_n \in \reals$.
+If $\delta \leq \tfrac{1}{2r}$ then $\fracpart{\sum_{n=1}^r{a_n}} = \sum_{n=1}^r{\fracpart{a_n}}$ and
+ \[
+ \left\vert\fracpart{\sum_{n=1}^r{a_n}}\right\vert = \left\vert\sum_{n=1}^r{\fracpart{a_n}}\right\vert \leq \sum_{n=1}^r{\left\vert\fracpart{a_n}\right\vert} < r\delta
+ \]
+ by the triangle inequality.
+ \end{proof}
+
+\begin{lemma} \label{lem:fracpartinternalsumlessdelta}
+Let $\left|\fracpart{a + \epsilon}\right| < \delta$ where $|\epsilon| < \nicefrac{1}{4}$ and $0<\delta<\nicefrac{1}{4}$. Then $\left|\fracpart{a}\right| < \delta + |\epsilon|$.
+\end{lemma}
+\begin{proof}
+By supposition $n - \delta < a + \epsilon < n + \delta$ for some $n \in \ints$.  Since $-\delta - \epsilon > -\tfrac{1}{2}$ and $\delta - \epsilon < \tfrac{1}{2}$, it follows that
+\[
+n - \tfrac{1}{2} < n - \delta - \epsilon < a < n + \delta - \epsilon < n + \tfrac{1}{2}.
+\]
+Hence $\fracpart{a} = a - n$ and so
+\[
+-\delta - \abs{\epsilon} \leq -\delta - \epsilon < \fracpart{a} < \delta - \epsilon \leq \delta + \abs{\epsilon}
+\]
+and $\abs{\fracpart{a}} \leq \delta + \abs{\epsilon}$.
+
+% The idea behind this proof is to show that under the conditions given $\fracpart{\fracpart{a} + \epsilon}$ does not wrap, i.e.
+that $\fracpart{a} + \epsilon \in \left[-\nicefrac{1}{2}, \nicefrac{1}{2}\right)$.
+To start, assume that $\fracpart{a} + \epsilon \geq \nicefrac{1}{2}$.
+Then $\nicefrac{1}{2} \leq \fracpart{a} + \epsilon < \nicefrac{3}{4}$ and
+% \[
+% -\nicefrac{1}{2} \leq \fracpart{\fracpart{a} + \epsilon} = \fracpart{a + \epsilon} < -\nicefrac{1}{4}
+% \] 
+% and therefore $|\fracpart{a + \epsilon}| > \nicefrac{1}{4} > \delta$, a contradiction.
+Similarly, assume that $\fracpart{a} + \epsilon < -\nicefrac{1}{2}$.
+Then $-\nicefrac{1}{2} > \fracpart{a} + \epsilon > -\nicefrac{3}{4}$ and $\nicefrac{1}{2} > \fracpart{a + \epsilon} > \nicefrac{1}{4}$ and therefore $\left|\fracpart{a + \epsilon}\right| > \nicefrac{1}{4} > \delta$, a contradiction.
+So, $\fracpart{a} + \epsilon \in \left[-\nicefrac{1}{2}, \nicefrac{1}{2}\right)$ and
+% \[
+% |\fracpart{a + \epsilon}| = \left|\fracpart{a} + \epsilon\right| < \delta
+% \]
+% from which it follows that $\left|\fracpart{a}\right| < \delta + \left|\epsilon\right|$.
+\end{proof}
+
+We are now in a position to complete the proof of strong consistency.
+Let $A$ be the subset of the sample space on which  $V_N(\widehat{\lambdabf}_N) - \sigma^2 \rightarrow 0$ as $N\rightarrow\infty$.
+From Lemma~\ref{lem:ESNconv}, the $\prob\{A\} =1$.
+Let $A'$ be the subset of the sample space on which $N^k\widehat{\lambda}_{k,N} \rightarrow 0$ for $k=0,\dots,m$ as $N\rightarrow\infty$.
+As a result of Lemma~\ref{lem:moran2}, $A \subseteq A'$, and so $\prob\{A'\} \geq \prob\{A\} = 1$.
+Strong consistency follows.
+
+\section{Proof of asymptotic normality}\label{sec:centlimitproof}
+
+Let $\psibf$ be the vector with $k$th component $\psi_k = N^k \lambda_k$, $k=0, \dots, m$ and let 
+\[
+T_{N}(\psibf) = S_N(\lambdabf) = \frac{1}{N} \sum_{n=1}^{N} \fracpart{ \Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \psi_k }^2.
+\]
+Let $\widehat{\psibf}_N$ be the vector with elements $\widehat{\psi}_{k,N} = N^k \widehat{\lambda}_{k,N}$ so that $\widehat{\psibf}_N$ is the minimiser of $T_N$.
+Because each of $N^k \widehat{\lambda}_{k,N}$ converges almost surely to zero as $N \rightarrow \infty$, then $\widehat{\psibf}_{N}$ converges almost surely to $\zerobf$ as $N \rightarrow \infty$.
+We want to find the asymptotic distribution of
+\[
+\sqrt{N}\widehat{\psibf}_{N} = 
+\left[
+\begin{array}
+[c]{c}%
+\sqrt{N} \widehat{\psi}_{0,N} \\ \sqrt{N}\widehat{\psi}_{1,N}  \\ \vdots \\ \sqrt{N} \widehat{\psi}_{m,N}
+\end{array}
+\right]
+=
+\left[
+\begin{array}
+[c]{c}%
+\sqrt{N} \widehat{\lambda}_{0,N} \\ N\sqrt{N}\widehat{\lambda}_{1,N} \\ \vdots \\ N^m\sqrt{N} \widehat{\lambda}_{m,N}
+\end{array}
+\right].
+\]
+The proof is complicated by the fact that $T_N$ is not differentiable everywhere as $\fracpart{x}^2$ is not differentiable when $\fracpart{x} = \tfrac{1}{2}$.
+This precludes the use of `standard approaches' to proving asymptotic normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.
+However, we show in Lemma~\ref{lem:diffathatpsi} that all the partial derivatives $\frac{\partial T_N}{\partial \psi_\ell}$ for $\ell = 0, \dots, m$ exist, and are equal to zero, at the minimiser $\widehat{\psibf}_N$.
+Thus, putting
+\begin{equation}\label{eq:Wn}
+W_{n} = \round{\Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \widehat{\psi}_{k,N}},
+\end{equation}
+so that 
+\[
+T_{N}(\psibf) = \frac{1}{N} \sum_{n=1}^{N} \big( \Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \psi_k  - W_n \big)^2
+\] 
+we have,
+\begin{align*}
+0  = \frac{\partial T_N}{\partial \psi_\ell}(\widehat{\psibf}_N) = \frac{2}{N}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\left( \Phi_n - W_{n}+ \sum_{k=0}^{m}(\tfrac{n}{N})^k \widehat{\psi}_{k,N}  \right),
+\end{align*}
+for each $\ell = 0, \dots, m$.  Now $D_{\ell,N} = K_{\ell,N}$, where $D_{\ell,N} = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}(\tfrac{n}{N})^\ell \Phi_n$, and
+\begin{equation}\label{eq:KellN}
+K_{\ell,N} = \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\left(W_{n}- \sum_{k=0}^{m}(\tfrac{n}{N})^k \widehat{\psi}_{k,N}  \right).
+\end{equation}
+Lemma~\ref{lem:Kconvfhalf} shows that,
+\begin{equation}\label{eq:KellNconv}
+K_{\ell,N} =  (h - 1) \sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big( C_{\ell k} + o_P(1) \big) + o_P(1),
+\end{equation}
+for all $\ell = 0, \dots m$, where $C_{\ell k} =  \tfrac{1}{\ell + k + 1}$, and $h = f(-\nicefrac{1}{2})$, and $o_P(1)$ denotes a random variable converging in probability to zero as $N\rightarrow\infty$.
+
+It is now convenient to write in vector form.  Let 
+\begin{equation}\label{eq:dnandkn}
+\kbf_N = 
+\left[
+\begin{array}{ccc}%
+K_{0,N}  & \cdots & K_{m,N}
+\end{array}
+\right]^\prime
+%\qquad \text{and} \qquad 
+=
+\dbf_N = \left[
+\begin{array}{ccc}%
+D_{0,N}  & \cdots & D_{m,N}
+\end{array}
+\right]^\prime.
+\end{equation}
+From~\eqref{eq:KellNconv},
+\[
+\dbf_N = \kbf_N = \sqrt{N} (h - 1)(\Cbf + o_P(1)) \widehat{\psibf}_N + o_P(1)
+\]
+where $o_P(1)$ here means a vector or matrix of the appropriate dimension with every element converging in probability to zero as $N\rightarrow\infty$.
+Thus $\sqrt{N}\widehat{\psibf}_N$ has the same asymptotic distribution as $(h - 1)^{-1}\Cbf^{-1} \dbf_N$.
+Lemma~\ref{eq:convdn} shows that $\dbf_N$ is asymptotically normally distributed with zero mean and covariance matrix $\sigma^2\Cbf$.
+Hence $\sqrt{N}\widehat{\psibf}_N$ is asymptotically normal with zero mean and covariance matrix
+\[
+\frac{\sigma^2\Cbf^{-1}\Cbf(\Cbf^{-1})^\prime}{(1 - h)^2} = 
+\frac{\sigma^2\Cbf^{-1}}{(1 - h)^2}.
+\] 
+It remains to prove Lemmas~\ref{lem:diffathatpsi},~\ref{lem:Kconvfhalf} and~\ref{eq:convdn}.
+
+\begin{lemma}\label{lem:diffathatpsi}
+For all $\ell = 0, \dots, m$ the partial derivatives $\frac{\partial T_N}{\partial \psi_\ell}$ exist, and are equal to zero, at the minimiser $\widehat{\psibf}_N$.
+That is $\frac{\partial T_N}{\partial \psi_\ell}(\widehat{\psibf}_N) = 0$ for each $\ell = 0, \dots, m$.
+\end{lemma}
+\begin{proof}
+The function $\fracpart{x}^2$ is differentiable everywhere except if $\fracpart{x} \neq -\frac{1}{2}$.
+Recalling that $T_{N}(\psibf) = \frac{1}{N} \sum_{n=1}^{N} \sfracpart{ \Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \psi_k }^2$ we see that $T_N$ will be differentiable with respect to $\psibf$ at $\widehat{\psibf}_N$ if $\sfracpart{\Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \widehat{\psi}_{k,N}} \neq -\tfrac{1}{2}$ for all $n = 1, \dots, N$.
+This is proved in Lemma~\ref{lem:boundonhatlambda}.
+So the partial derivatives $\frac{\partial T_N}{\partial \psi_\ell}$ exist for all $\ell = 0, \dots, m$ at $\widehat{\psibf}_N$.
+That each of the partial derivatives is equal to zero at $\widehat{\psibf}_N$ follows immediately from the fact that $\widehat{\psibf}_N$ is a minimiser of $T_N$.
+\end{proof}
+
+\begin{lemma}\label{lem:boundonhatlambda} $\sabs{\sfracpart{\Phi_n + \sum_{k=0}^m (\nicefrac{n}{N})^k \widehat{\psi}_{k,N}}} \leq \frac{1}{2} - \frac{1}{2N}$ for all $n = 1, \dots, N$.
+\end{lemma}
+\begin{proof}
+To simplify our notation let $B_n = \Phi_n + \sum_{k=1}^m (\nicefrac{n}{N})^k \widehat{\psi}_{k,N}$ so that we now require to prove $\abs{\fracpart{B_n + \widehat{\psi}_{0,N}}} \leq \frac{1}{2} - \frac{1}{2N}$ for all $n = 1, \dots, N$.
+From~\eqref{eq:Wn}, $W_n = \round{B_n + \widehat{\psi}_{0,N}}$, and
+\begin{align*}
+T_N(\widehat{\psibf}_N) = \frac{1}{N}\sum_{n=1}^N \fracpart{B_n + \widehat{\psi}_{0,N}}^2= \frac{1}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N} - W_n)^2.
+\end{align*}
+Since $\widehat{\psi}_{0,N}$ is the minimiser of the quadratic above,
+\begin{equation}\label{eq:lamsum}
+\widehat{\psi}_{0,N} = -\frac{1}{N}\sum_{n=1}^N(B_n - W_n).
+\end{equation}
+The proof now proceeds by contradiction.  Assume that for some $k$,
+\begin{equation}\label{eq:philamfraccontra}
+\fracpart{B_k + \widehat{\psi}_{0,N}} > \frac{1}{2} - \frac{1}{2N}.
+\end{equation}
+Let $F_n = W_n$ for all $n \neq k$ and $F_k = W_k + 1$, and let
+\[
+\phi = -\frac{1}{N}\sum_{n=1}^N(B_n - F_n) = \widehat{\psi}_{0,N} + \frac{1}{N}.
+\]
+Then,
+\begin{align}
+(B_k + \phi - F_k)^2 &= (B_k + \phi - W_k - 1)^2 \nonumber \\
+&= (B_k + \phi - W_k)^2 - 2(B_k + \phi - W_k) + 1 \nonumber \\
+&= (B_k + \phi - W_k)^2 - 2(B_k + \widehat{\psi}_{0,N} - W_k) + 1 - \frac{2}{N} \nonumber \\
+&= (B_k + \phi - W_k)^2 - 2\fracpart{B_k + \widehat{\psi}_{0,N}} + 1 - \frac{2}{N} \nonumber \\
+&< (B_k + \phi - W_k)^2 - \frac{1}{N}, \label{eq:Bkineq}
+\end{align}
+where the inequality in the last line follows from~\eqref{eq:philamfraccontra}. Let $\bbf = [ \phi,  \widehat{\psi}_{1,N} , \dots , \widehat{\psi}_{m,N}]$
+% \[
+% \bbf = \left[
+% \begin{array}
+% [c]{cccc}%
+% \phi & \widehat{\psi}_{1,N} & \cdots & \widehat{\psi}_{m,N}
+% \end{array}
+% \right]^\prime
+% \]
+be the vector of length $m+1$ with components $b_0 = \phi$ and $b_\ell = \widehat{\psi}_{\ell,N}$ for $\ell = 1 , \dots m$.  Now,
+\[
+N T_N(\bbf) = \sum_{n=1}^N \fracpart{B_n + \phi }^2 \leq  \sum_{n=1}^N (B_n + \phi  - F_n)^2, 
+\]
+and using the inequality from~\eqref{eq:Bkineq},
+\begin{align*}
+N T_N(\bbf) &< - \frac{1}{N} + \sum_{n=1}^N (B_n + \phi  - W_n)^2 \\
+&= - \frac{1}{N} + \sum_{n=1}^N (B_n + \widehat{\psi}_{0,N} + \frac{1}{N}  - W_n)^2 \\
+&= \sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n)^2 +  \frac{2}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n)\\
+%&= N T_N(\widehat{\psibf}_N) + \frac{2}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n) \\
+&= N T_N(\widehat{\psibf}_N),
+\end{align*}
+because $\frac{2}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n) = 0$ as a result of~\eqref{eq:lamsum}.
+But, now $T_N(\bbf) < T_N(\widehat{\psibf}_N)$ violating the fact that $\widehat{\psibf}_N$ is a minimiser of $T_N$.
+So~\eqref{eq:philamfraccontra} is false by contradiction.
+
+If $\fracpart{B_k + \widehat{\psi}_{0,N}} < -\frac{1}{2} + \frac{1}{2N}$ for some $k$, we set $F_k = W_k - 1$ and using the same procedure as before obtain $T_N(\bbf) < T_N(\widehat{\psibf}_N)$ again.
+The proof follows.
+\end{proof}
+
+\begin{lemma}\label{lem:Kconvfhalf}
+With $K_{\ell,N}$ defined as in~\eqref{eq:KellN} and $h = f(-\nicefrac{1}{2})$,
+\[
+K_{\ell,N} = (h - 1) \sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big( C_{\ell k} + o_P(1) \big) + o_P(1) 
+\] 
+for all $\ell = 0, \dots, m$, where $C_{\ell k} =  \frac{1}{\ell + k + 1}$.% is an element of the $m+1$ by $m+1$ Hilbert matrix, and $o_P(1)$ denotes a random variable converging in probability to zero as $N\rightarrow\infty$.
+\end{lemma}
+\begin{proof}
+Care must be taken since $\widehat{\psibf}_N$ depends on the sequence $\{ \Phi_n \}$.  For $n = 1, \dots, N$ and positive $N$, let
+\begin{equation}
+p_{nN}(\psibf) = \sum_{k=0}^{m}\left( \tfrac{n}{N}\right)^k \psi_k,
+\end{equation}
+and put $q_{n}(x) = \round{\Phi_n + x}$ and $Q(x) = \expect q_{n}(x) =  \expect q_{1}(x)$.  Let
+\begin{equation}\label{eq:GNdef}
+G_N(\psibf) = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}\left( \tfrac{n}{N}\right)^\ell  \big(q_n(p_{nN}(\psibf)) - Q(p_{nN}(\psibf)) \big),
+\end{equation}
+and put
+\begin{equation}\label{eq:hatpnN}
+\widehat{p}_{nN} = p_{nN}(\widehat{\psibf}_N) = \sum_{k=0}^m \left(\tfrac{n}{N}\right)^k \widehat{\psi}_{k,N}.
+\end{equation}
+Observe that $G_N$ depends on $\ell$ and we could write $G_{\ell,N}$ but have suppressed the subscript $\ell$ for notational simplicity.
+Now $W_n$ from~\eqref{eq:Wn} can be written as $W_n = \round{\Phi_n + \widehat{p}_{nN}} = q_{n}(\widehat{p}_{nN})$ and $K_{\ell,N}$ from~\eqref{eq:KellN} can be written as
+\begin{align*}
+K_{\ell,N} &= \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\big( q_{n}(\widehat{p}_{nN}) - \widehat{p}_{nN}  \big) \\
+&= \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\big(q_{n}(\widehat{p}_{nN}) - \widehat{p}_{nN} + Q(\widehat{p}_{nN}) - Q(\widehat{p}_{nN}) \big) \\
+&= G_N(\widehat{\psibf}_N) + H_{\ell,N},
+\end{align*}
+where 
+\begin{equation}\label{eq:HellNdef}
+H_{\ell,N} = \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell \big( Q(\widehat{p}_{nN}) - \widehat{p}_{nN} \big).
+\end{equation}
+Lemma~\ref{lem:unifprobG} in the Appendix shows that for any $\delta >0$ and $\nu > 0$ there exists an $\epsilon > 0$ such that
+\[
+\prob\left\{ \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf) } > \delta   \right\} < \nu
+\]
+for all positive integers $N$, where $\|\psibf\|_\infty = \sup_{k} \abs{\psi_k}$.
+Since $\widehat{\psibf}_N$ converges almost surely to zero, it follows that $\lim_{N\rightarrow\infty}\prob\left\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} = 0$ for any $\epsilon > 0$, and therefore $\prob\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \} < \nu$ for all sufficiently large $N$.
+Now
+\begin{align*}
+  \prob\left\{\abs{ G_N(\widehat{\psibf}_N) } > \delta \right\} &= \prob\left\{ \abs{G_N(\widehat{\psibf}_N)} > \delta \;, \; \|\widehat{\psibf}_N\|_{\infty} < \epsilon \right\} \\
+&\hspace{1cm} + \prob\left\{ \abs{G_N(\widehat{\psibf}_N)} > \delta  \;, \; \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} \\
+&\leq \prob\left\{  \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf) } > \delta \right\} + \prob\left\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} \\
+&\leq 2\nu
+\end{align*}
+for all sufficiently large $N$.
+Since $\nu$ and $\delta$ can be chosen arbitrarily small, it follows that $G_N(\widehat{\psibf}_N)$ converges in probability to zero as $N\rightarrow\infty$, and therefore $K_{\ell,N} = H_{\ell,N} + o_P(1)$.
+Lemma~\ref{lem:sumpNhilb} shows that
+\begin{align*}
+H_{\ell,N} =  (h-1)\sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big(C_{\ell k} + o_P(1)\big).
+\end{align*}
+The proof follows.
+\end{proof}
+
+\begin{lemma}\label{lem:sumpNhilb}
+With $H_{\ell,N}$ defined in~\eqref{eq:HellNdef} and with $h = f(-\nicefrac{1}{2})$, we have $H_{\ell,N} = (h-1)\sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big(C_{\ell k} + o_P(1)\big)$, where $C_{\ell k} = \tfrac{1}{\ell + k + 1}$.
+\end{lemma}
+\begin{proof}
+If $\abs{x} < 1$, then
+\[
+q_n(x)  = \round{\Phi_n + x} = \begin{cases}
+1, & \Phi_n + x  \geq \nicefrac{1}{2} \\
+-1, & \Phi_n + x  < -\nicefrac{1}{2} \\
+0, & \text{otherwise},
+\end{cases}
+\]
+and,
+\begin{align*}
+Q(x) = Eq_1(x) &=  \begin{cases}
+\int_{\nicefrac{1}{2} -x }^{\nicefrac{1}{2}}f(t)\,dt, &   x \geq 0 \\
+-\int_{-\nicefrac{1}{2}}^{-\nicefrac{1}{2} - x }f(t)\,dt, &  x  < 0.
+\end{cases} 
+% \\
+% &= \begin{cases}
+% 1 - F(\nicefrac{1}{2} - x ), & x  \geq 0 \\
+% -F(-\nicefrac{1}{2} - x ), &  x < 0,
+% \end{cases} 
+\end{align*}
+Because $f(\fracpart{x})$ is continuous at $-\nicefrac{1}{2}$ it follows from the mean value theorem that $Q(x) = x \big( h + \zeta(x) \big)$
+where $\zeta(x)$ is a function that converges to zero as $x$ converges to zero.  
+%The equation above is still valid when $\abs{x} \geq 1$ by an appropriate choice of the function $\zeta(x)$.  
+Observe that $\abs{\widehat{p}_{nN}} \leq \sum_{k=0}^m\sabs{ \widehat{\psi}_{k,N}}$ and, since each of the $\widehat{\psi}_{k,N} \rightarrow 0$ almost surely as $N\rightarrow\infty$, it follows that $\widehat{p}_{nN} \rightarrow 0$ almost surely uniformly in $n = 1, \dots, N$ as $N\rightarrow\infty$.
+Thus, $\zeta(\widehat{p}_{nN}) \rightarrow 0$ almost surely (and therefore also in probability) uniformly in $n = 1, \dots, N$ as $N\rightarrow\infty$.
+Now,
+\begin{align*}
+Q(\widehat{p}_{nN}) - \widehat{p}_{nN} = \widehat{p}_{nN}\big( h - 1 + \zeta(\widehat{p}_{nN}) \big) = \widehat{p}_{nN}\big( h - 1 + o_P(1) \big), 
+\end{align*}
+and, using~\eqref{eq:HellNdef},
+\begin{align*}
+H_{\ell,N} &= \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell \widehat{p}_{nN}\big( h - 1 + o_P(1) \big)  \\
+&=  \frac{1}{\sqrt{N}} \sum_{n=1}^{N}(\tfrac{n}{N})^\ell \sum_{k=0}^m (\tfrac{n}{N})^k \widehat{\psi}_{k,N} \big( h - 1 + o_P(1) \big)   \\
+&= \sqrt{N}\sum_{k=0}^m \widehat{\psi}_{k,N} \frac{1}{N} \sum_{n=1}^{N} \frac{n^{\ell + k}}{N^{\ell + k + 1}}\big( h - 1 + o_P(1)  \big).
+\end{align*}
+The Riemann sum
+\[
+\frac{1}{N} \sum_{n=1}^{N} \frac{n^{\ell + k}}{N^{\ell + k + 1}} = \int_{0}^{1}x^{k+\ell+1}dx + o_P(1),
+\]
+and since the integral above evaluates to $\frac{1}{k+\ell+1}$, we have
+\[
+H_{\ell,N} = (h - 1)\sqrt{N}\sum_{k=0}^m \widehat{\psi}_{k,N}\left( \frac{1}{k+\ell+1} + o_P(1)  \right)
+\]
+as required.
+% Put
+% \[
+% C_{\ell + k,N} = \sum_{n=1}^{N}\frac{n^{\ell + k}}{N^{\ell + k+1}}.
+% \]
+% It is well known (see for example~\cite[eq. 21]{Peleg1991_CRB_PPS_1991}) that,
+% \begin{equation}\label{eq:ClkNHilbert}
+% C_{\ell + k,N} = C_{\ell k} + O(N^{-1}).
+% \end{equation}
+% Since $\zeta(\widehat{p}_{nN}) = o_P(1)$ for all $n = 1, \dots, N$ then
+% \[
+% \sum_{n=1}^{N}\frac{n^{\ell + k}}{N^{\ell + k + 1}}\zeta(\widehat{p}_{nN}) 
+% \]
+% converges in probability to zero as $N\rightarrow\infty$.  So,
+% \begin{align*}
+% \sum_{n=1}^{N}\frac{n^{\ell + k}}{N^{\ell + k + 1}}\big( h - 1 +  \zeta(\widehat{p}_{nN}) \big) = (h-1)\big( C_{\ell k} + o_P(1) \big).
+% \end{align*}
+% The proof follows by substituting this into~\eqref{eq:HellNre}.
+\end{proof}
+
+\begin{lemma}\label{eq:convdn}
+The distribution of the vector $\dbf_N$, defined in~\eqref{eq:dnandkn}, converges to the multivariate normal with zero mean and covariance matrix $\sigma^2\Cbf$.
+\end{lemma}
+\begin{proof}
+For any constant vector $\alphabf$, let
+\[
+z_N = \alphabf^\prime \dbf_N = \frac{1}{\sqrt{N}} \sum_{n=1}^{N} \Phi_n \sum_{\ell=0}^{m}\alpha_\ell\left( \frac{n}{N} \right)^\ell.
+\]
+By Lypanov's central limit theorem $z_N$ is asymptotically normally distributed with zero mean and variance
+\[
+\lim_{N\rightarrow\infty} \sigma^2 \frac{1}{N} \sum_{n=1}^{N}\left( \sum_{\ell=0}^{m}\alpha_\ell\left( \frac{n}{N} \right)^\ell \right)^2 = \sigma^2 \alphabf^\prime \Cbf \alphabf.
+\]
+By the Cramer-Wold theorem it follows that $\dbf_N$ is asymptotically normally distributed with zero mean and covariance $\sigma^2 \Cbf$. 
+% Recall that the $\ell$th element of $\dbf_N$ is
+% \[
+% D_{\ell,N} = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}(\tfrac{n}{N})^\ell \Phi_n.
+% \]
+% Since the $\Phi_1, \dots, \Phi_N$ are independent, each with variance $\sigma^2$, the distribution of $D_{\ell,N}$ converges to the normal with zero mean and variance
+% \[
+% \lim_{N\rightarrow\infty}\var D_{\ell,N} = \sigma^2 \lim_{N\rightarrow\infty} C_{2\ell,N} = \sigma^2 C_{\ell\ell}
+% \]
+% by Lypanov's central limit theorem and by using~\eqref{eq:ClkNHilbert}.  The covariance between $D_{\ell,N}$ and $D_{k,N}$ is
+% \[
+% \covar(D_{\ell,N}, D_{k,N}) = \sigma^2 C_{\ell+k,N} \rightarrow \sigma^2 C_{\ell k}
+% \]
+% as $N \rightarrow \infty$, again using~\eqref{eq:ClkNHilbert}.  The proof follows.
+\end{proof}
+
+\section{Simulations}\label{sec:simulations}
+
+This section describes the results of Monte-Carlo simulations with the least squares unwrapping (LSU) estimator.
+The sample sizes considered are $N = 10, 50$ and $200$ and the unknown amplitude is $\rho = 1$.
+The $X_1, \dots, X_N$ are pseudorandomly generated independent and identically distributed circularly symmetric complex Gaussian random variables with variance $\sigma_c^2$.
+The coefficients $\tilde{\mubf} = [\tilde{\mu}_0, \dots, \tilde{\mu}_m]$ are distributed uniformly randomly in the identifiable region $B$.
+The number of replications of each experiment is $T = 2000$ to obtain estimates $\widehat{\mubf}_1, \dots, \widehat{\mubf}_T$ and the corresponding dealiased errors $\widehat{\lambdabf}_t = \dealias(\widehat{\mubf}_T - \tilde{\mubf})$ are computed.
+The sample mean square error (MSE) of the $k$th coefficient is computed according to $\tfrac{1}{T}\sum_{t=1}^T \widehat{\lambda}_{k,t}^2$ where $\widehat{\lambda}_{k,t}$ is the $k$th element of $\widehat{\lambdabf}_t$.
+
+Figures~\ref{plot:polyest1}~and~\ref{plot:polyest2} show the sample MSEs obtained for a polynomial phase signal of order $m=3$.
+Figure~\ref{plot:polyest1} displays results for the zeroth and first order coefficients $\widehat{\mu}_0$ and $\widehat{\mu}_1$ and Figure~\ref{plot:polyest2} displays results for the second and third order coefficients $\widehat{\mu}_0$ and $\widehat{\mu}_1$.
+When $N = 10$ and $50$ the LSU estimator can be computed exactly using a general purpose algorithm for finding nearest lattice points called the \emph{sphere~decoder}~\cite{Pohst_sphere_decoder_1981,Agrell2002,Viterbo_sphere_decoder_1999}.
+This is displayed be the circles in the figures.
+When $N=200$ the sphere decoder is intractable and we instead use an approximate nearest point algorithm called the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder}.
+This is displayed by the dots.
+For the purpose of comparison we have also plotted the results for the $K$-best method when $N = 10$ and $50$.
+The asymptotic variance predicted in Theorem~\ref{thm:asymp_proof} is displayed by the dashed line.
+Provided the noise variance is small enough (so that the `threshold' is avoided) the sample MSE of the LSU estimator is close to that predicted by Theorem~\ref{thm:asymp_proof}.
+The Cram\`{e}r-Rao lower bound for the variance of unbiased polynomial phase estimators in Gaussian noise is also plotted with the solid line~\cite{McKilliam_polycrbs_2012,Peleg1991_CRB_PPS_1991}.
+When the noise variance is small the asymptotic variance of the LSU estimator is close to the Cram\`{e}r-Rao lower bound.
+
+%Note the gap between the CRB and the variance predicted by Theorem~\ref{thm:asymp_proof}.
+In practice this gap can be overcome using a numerical technique, such as Newton's method, with the LSU estimate as a starting point.
+We have not done this here in order to show the accuracy of Theorem~\ref{thm:asymp_proof}.
+The other noticeable gap is that between the threshold behaviour of the LSU estimator and the LSE.
+When $N=50$ the threshold \emph{kicks in} a little earlier for the LSE than it does for the LSU.
+We suspect that this gap will actually \emph{close} as $N$ increases.
+This hypothesis is supported by the fact that this behaviour is observed in the case of frequency estimation (when $m=1$)~\cite[Sec.
+9.5]{McKilliam2010thesis}\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.
+Unfortunately it is not feasible to test this hypothesis because of the computational complexity of the LSE.
+
+% \begin{figure}[p] 
+ %   	\centering 
+ %  		\includegraphics[width=\linewidth]{code/lseplot-1.mps} 
+ %               %\includegraphics{code/lseplot-1.mps}
+ %   		\caption{Performance of the LSE and LSU estimator for a polynomial phase signal of order two, i.e.
+$m=2$.
+(Top) Mean square error (MSE) in the second order parameter $\mu_2$ for $N=10,50$ and $200$ versus the variance $\sigma_c^2$ of the $X_1, \dots, X_N$.
+The $X_1, \dots, X_N$ are independent and identically distributed (i.i.d.) complex Gaussian random variables.
+(Bottom) MSE in second order parameter $\mu_2$ versus the variance $\sigma^2$ of the $\Phi_1, \dots, \Phi_N$.
+The $\Phi_1, \dots, \Phi_N$ are i.i.d.
+uniform random variables.}
+ %   		\label{plot:lsetest} 
+ %  \end{figure}
+
+\begin{figure}[p] 
+   	\centering 
+  		\includegraphics[width=\linewidth]{code/gaussianplot4-1.mps}
+                %\includegraphics{code/gaussianplot4-1.mps}
+  		\caption{Sample mean square error (MSE) of the least squares unwrapping estimator for $N=10,50$ and $200$ for a polynomial phase signal of order $m=3$.
+(Top) MSE of the phase coefficient $\mu_0$.
+(Bottom) MSE of the frequency coefficient $\mu_1$.}
+  		\label{plot:polyest1} 
+ \end{figure}
+
+\begin{figure}[p] 
+   	\centering 
+  		\includegraphics[width=\linewidth]{code/gaussianplot4-2.mps}
+                %\includegraphics{code/gaussianplot4-1.mps}
+  		\caption{Sample mean square error (MSE) of the least squares unwrapping estimator for $N=10,50$ and $200$ for a polynomial phase signal of order $m=3$.
+(Top) MSE of the quadratic coefficient $\mu_2$.
+(Bottom) MSE of the cubic coefficient $\mu_3$.}
+  		\label{plot:polyest2} 
+ \end{figure}
+
+% \section{Computational considerations}
+
+% When $m=0$, polynomial phase estimation is equivalent to the problem of \emph{direction estimation} from circular statistics, and the least squares estimator corresponds with the \emph{extrinsic mean} (or \emph{sample circular mean})~\cite{McKilliam_mean_dir_est_sq_arc_length2010,Fisher1993,Mardia_directional_statistics}.
+In this case the LSE can be computed efficiently in $O(N)$ by averaging $N$ complex numbers.
+When $m=1$, polynomial phase estimation is equivalent to single frequency estimation, and the least squares estimator can be computed efficiently in $O(N\log N)$ operations by maximising the \emph{periodogram} using the fast Fourier transform followed by an appropriate numerical optimisation technique~\cite{Quinn2008maximizing_the_periodogram,Rife1974,Quinn2001}.
+When $m=2$, the least squares estimator can be computed in a reasonable amount of time when $N$ is approximately less than $50$.
+The LSE is computationally intensive for any $N$ when $m > 2$.
+By comparison the exact LSU estimator implemented using the sphere decoder can be computed quickly for any $m$ with $N$ approximately less than $50$.
+In situations where $N$ is small, but high statistical accuracy is required, the sphere decoder is computationally a better choice than the LSE for polynomial phase signals of order greater than or equal to $2$.
+
+% Babai's nearest plane algorithm requires $O(N^2)$ operations and can be run in a reasonable amount of time for any $m$ and reasonably large $N$.
+It should be noted that Babai's nearest plane algorithm requires a Lov\`as reduced lattice basis~\cite{Lenstra1982}.
+This requires $O(N^4)$ operations to compute, but, the Lov\`as reduced basis can be computed once offline.
+It does not need to be computed each time the estimator is run.
+The $K$-best algorithm runs in a reasonable amount of time for any $m$ and when $N$ is less than about $300$.
+If the noise variance is small, then both the sphere decoder and the $K$-best algorithm are actually fast.
+This is because both of these estimators begin with the lattice point found by Babai's nearest plane algorithm.
+If this point is close to (or \emph{is}) the nearest point then both the sphere decoder and the $K$-best algorithm terminate quickly.
+This is in contrast to the least squares estimator that is slow regardless of the noise variance.
+
+% Both Kitchen's estimator and the DPT are fast to compute.
+Kitchen's estimator requires only $O(N)$ arithmetic operations and the DPT requires $O(N \log N)$ operations.
+However, these estimators are not as accurate as the LSE or the LSU.
+They also fail to operate correctly over the entire identifiable region of polynomial phase coefficients~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}~\cite[Ch.~10]{McKilliam2010thesis}.
+
+\section{Conclusion} \label{sec:conclusion}
+
+This paper has considered the estimation of the coefficients of a noisy polynomial phase signal by least squares phase unwrapping (LSU).
+It has been shown that the LSU estimator is strongly consistent and asymptotically normally distributed.
+%Simulations are used to show that the LSU has near maximum likelihood performance in the case that the noise is additive white and Gaussian.
+It is shown in~\cite{McKilliam2009asilomar_polyest_lattice} how the LSU estimator can be computed by finding a nearest lattice point in a particular lattice.
+Polynomial time algorithms that compute the nearest point are described in~\cite[Sec 4.3]{McKilliam2010thesis}, but these are slow algorithms in practice.
+A significant outstanding question is whether practically fast nearest point algorithms exist for these specific lattices.
+Considering the excellent statistical performance (both theoretically and practically) of the LSU estimator, even fast \emph{approximate} nearest point algorithms are likely to prove useful for the estimation of polynomial phase signals.
+
+\appendix
+
+%\section{Tricks with fractional parts}
+%Throughout the paper, and particularly in  have made use of a number of results involving the fractional part function
+
+\section{A uniform law of large numbers} \label{app:uniform-law-large}
+
+During the proof of strong consistency we made use of the fact that
+\begin{equation}~\label{eq:SNVNunifmlln2}
+\sup_{\lambdabf \in B}\sabs{S_N(\lambdabf) - V_N(\lambdabf)} \rightarrow 0
+\end{equation}
+almost surely as $N\rightarrow\infty$, where $V_N(\lambdabf) = \expect S_N(\lambdabf)$.  We prove this result here. Put $D_N(\lambdabf) = S_N(\lambdabf) - V_N(\lambdabf)$.  By Lemma~\ref{lem:vn},
+\[
+\sum_{N=1}^\infty \prob \left\{ \sup_{\lambdabf \in B}\abs{ D_N(\lambdabf) } > \epsilon \right\} < \infty
+\]
+for any $\epsilon > 0$.
+Thus~\eqref{eq:SNVNunifmlln2} follows from the Borel-Cantelli lemma.
+In what follows we use order notation in the standard way, that is, for functions $h$ and $g$, we write $h(N) = O(g(N))$ to mean that there exists a constant $K > 0$ and a finite $N_0$ such that $h(N) \leq K g(N)$ for all $N > N_0$.
 
@@
  \sabs{ D_N(\lambdabf) - D_N(\lambdabf[\rbf]) } &= \sabs{ S_N(\lambdabf) - S_N(\lambdabf[\rbf]) + V_N(\lambdabf) - V_N(\lambdabf[\rbf]) } \\
-&\leq \sabs{ S_N(\lambdabf) - S_N(\lambdabf[\rbf]) \vert + \vert V_N(\lambdabf) - V_N(\lambdabf[\rbf]) },  \\
-&\leq 2\frac{m+1}{N^{b}}
-\end{align*}
- and as this bound is independent of $\rbf$,
-\[
-\sup_{\rbf\in G}\sup_{ \lambdabf  \in B[\rbf] }\sabs{ D_N\left(  \lambdabf \right) - D_N\left(  \lambdabf[\rbf] \right) } \leq 2\frac{m+1}{N^{b}}.
-\]
- \end{proof}
-
-The proof of Lemma~\ref{lem:vn} is now complete.  We concentrate our attention on the minimiser of $V_N$. Because $\Phi_n$ has zero intrinsic mean 
-\begin{equation}\label{eq:Efracpartmined}
-\expect \fracpart{ \Phi_n + z }^{2}
-\end{equation}
-is minimised uniquely at $z = 0$ for $z \in [-\nicefrac{1}{2}, \nicefrac{1}{2})$.  Since the intrinsic variance of $\Phi_n$ is $\sigma^2$, when $z = 0$,
-\begin{equation}\label{eq:Efracpartphi}
-\expect\fracpart{\Phi_1+z}^{2} = \expect\fracpart{\Phi_1}^{2} = \sigma^2,
-\end{equation}
-and so the minimum attained value is $\sigma^2$.
-
-\begin{lemma}\label{lem:ES_Nminimisedzero}
-For $\lambdabf \in B$ the function $V_N(\lambdabf)$ is minimised uniquely at $\zerobf$, the vector of all zeros.  At this minimum $V_N(\zerobf) = \sigma^2$.
-\end{lemma}
-\begin{proof}
-Let $z$ be the polynomial
-\[
-z(n) = \lambda_0 + \lambda_1 n + \lambda_2 n^2 + \dots + \lambda_m n^m.
-\]
-Then
-\begin{align*}
-V_N(\lambdabf) = \frac{1}{N}\expect\sum_{n=1}^{N}\fracpart{ \Phi_n + \sum_{k=0}^m{\lambda_k n^k} }^2 = \frac{1}{N}\sum_{n=1}^{N}\expect\fracpart{ \Phi_n + \fracpart{z(n)} }^2.
-\end{align*}
-We know that $\expect\fracpart{ \Phi_n + \fracpart{z(n)} }^2$ is minimised uniquely when $\fracpart{z(n)} = 0$ at which point it takes the value $\sigma^2$.
-Now $\fracpart{z(n)}$ is equal to zero for all integers $n$ if and only if $z \in \mathcal{Z}$, or equivalently if $\coef(z)$ is a lattice point in $L_{m+1}$.
-By definition $B$ contains precisely one lattice point from $L_{m+1}$, this being the origin $\zerobf$.
-Therefore $V_N$ is minimised uniquely at $\zerobf$, at which point it takes the value $\sigma^2$.
-\end{proof}
-
-\begin{lemma} \label{lem:ESNconv}
-For any $\epsilon > 0$ and $0 < c < 2$,
-\[
-\prob\left\{ \abs{V_N(\widehat{\lambdabf}_N) - \sigma^2} > \epsilon \right\} = O(e^{-c \epsilon^2 N}).
-\]
-\end{lemma}
-\begin{proof}
-By definition 
-\[
-\widehat{\lambdabf}_N =\arg\min_{\lambdabf\in B} S_N(\lambdabf)
-\]
-and therefore $0 \leq S_{N}(\zerobf) - S_N(\widehat{\lambdabf}_N)$.  Also, because $V_N$ is minimised at $\zerobf$, it follows that $0 \leq V_N(\widehat{\lambdabf}_N) - V_N(\zerobf)$.  Thus,
-\begin{align*}
-0 &\leq V_N(\widehat{\lambdabf}_N) - V_N(\zerobf) \\
- &\leq V_N(\widehat{\lambdabf}_N) - V_N(\zerobf) + S_{N}(\zerobf) - S_N(\widehat{\lambdabf}_N)   \\
-&\leq \sabs{ V_N(\widehat{\lambdabf}_N) - S_N(\widehat{\lambdabf}_N) } + \sabs{ S_{N}(\zerobf) - V_N(\zerobf) }\\
-&= \sabs{ D_N(\widehat{\lambdabf}_N) } + \sabs{ D_N(\zerobf) },
-\end{align*}
-and from Lemma~\ref{lem:vn},
-\begin{align*}
-\prob\left\{ \sabs{V_N(\widehat{\lambdabf}_N) - V_N(\zerobf))} > \epsilon \right\} &\leq \prob\left\{ \sabs{ D_N(\widehat{\lambdabf}_N) } + \sabs{ D_N(\zerobf) } > \epsilon \right\} \\
-&\leq \prob\left\{ \sabs{ D_N(\widehat{\lambdabf}_N) } > \frac{\epsilon}{2}\right\} + \prob\left\{ \sabs{ D_N(\zerobf) } > \frac{\epsilon}{2} \right\} \\
-&= O(e^{-d \epsilon^2 N / 4}),
-\end{align*}
-where $d$ is any real number less than $8$.
-%where the third line follows from Lemma~\ref{lem:ABprob}.
-\end{proof}
-
-\begin{corollary}  \label{cor:ESNconvcor}
-$\sabs{V_N(\widehat{\lambdabf}_N) - \sigma^2}$ converges almost surely to zero as $N\rightarrow\infty$.
-\end{corollary}
-\begin{proof}
-From Lemma~\ref{lem:ESNconv}, for any $\epsilon > 0$ and $0 < c < \tfrac{1}{2}$,
-\[
-\sum_{N=1}^{\infty}\prob\left\{ \abs{V_N(\widehat{\lambdabf}_N) - \sigma^2} > \epsilon \right\} = \sum_{N=1}^{\infty} O(e^{-c \epsilon^2 N}) < \infty.
-\]
-The corollary follows from the Borel-Cantelli lemma.
-\end{proof}
-
-% \begin{lemma}\label{lem:ABprob}
-% Let $A$ and $B$ be positive random variables.  Then
-% \[
-% \prob( A + B > \epsilon ) \leq \prob( A > \tfrac{\epsilon}{2}) + \prob( A > \tfrac{\epsilon}{2}).
-% \]
-% \end{lemma}
-% \begin{proof}
-% Write 
-% \begin{equation}\label{eq:A+B>epsilon}
-% \prob( A + B > \epsilon ) = 1 - \prob( A + B \leq \epsilon ).
-% \end{equation}
-% If $A \leq \tfrac{\epsilon}{2}$ and $B \leq \tfrac{\epsilon}{2}$ then $A + B \leq \epsilon$, so
-% \begin{align*}
-% \prob( A + B \leq \epsilon ) &\geq \prob( A \leq \tfrac{\epsilon}{2} \;\; \text{and} \;\; B \leq \tfrac{\epsilon}{2} ) \\
-% &= 1 - \prob( A > \tfrac{\epsilon}{2} \;\; \text{or} \;\; B > \tfrac{\epsilon}{2} ) \\
-% &\geq 1 - \prob( A > \tfrac{\epsilon}{2}) - \prob( B > \tfrac{\epsilon}{2} ).
-% \end{align*}
-% Subtituting this into~\eqref{eq:A+B>epsilon} gives the required result.
-% \end{proof}
-
-We have now shown that $V_N$ is uniquely minimised at $\zerobf$, that $V_N(\zerobf) = \sigma^2$, and that $V_N(\widehat{\lambdabf}_N)$ converges almost surely to $\sigma^2$.
-These results are enough to show that $\widehat{\lambdabf}_N$ converges almost surely to zero.
-However, this tells us nothing about the rate at which the components of $\widehat{\lambdabf}_N$ approach zero as required by Theorem~\ref{thm:asymp_proof}.
-To prove these stronger properties we need some preliminary results about arithmetic progressions, and from the calculus of finite differences.
-
- %\begin{equation}
- %E\left(  \sum_{n=1}^{N}z_{n}\right)  ^{2\beta}=\sum_{n_{1},\ldots,n_{2\beta
- %}=1}^{N}E\left(  \prod\limits_{j=1}^{2\beta}z_{n_{j}}\right)  .
- %\label{eq:markapp}%
- %\end{equation}
- %Since the terms%
+&\leq \sabs{ S_N(\lambdabf) - S_N(\lambdabf[\rbf]) \vert + \vert V_N(\lambdabf) - V_N(\lambdabf[\rbf]) } \\
+&\leq 2\frac{m+1}{N^{b}},
+\end{align*}
+and the lemma follows because this bound is independent of $\rbf$,
 %\[
- %E\left(  \prod\limits_{j=1}^{2\beta}z_{n_{j}}\right)
+%\sup_{\rbf\in G}\sup_{ \lambdabf  \in B[\rbf] }\sabs{ D_N\left(  \lambdabf \right) - D_N\left(  \lambdabf[\rbf] \right) } \leq 2\frac{m+1}{N^{b}}.
 %\]
- %in this containing any $n_{j}$ only once will have $0$ mean, and since the
- %$z_{j}$ are bounded, it follows that%
- %\[
- %\frac{1}{N^{\beta}}E\left(  \sum_{n=1}^{N}z_{n}\right)  ^{2\beta}=O\left(
- %1\right)  ,
- %\]
- %as $N\rightarrow\infty,$ the dominant terms coming from the $z_{j}$ in
- %$\left(  \ref{eq:markapp}\right)  $ occurring in pairs.
-
-Let $W = \{1,2,\dots, N\}$ and let $K$ be a subset of $W$.  For any integer $h$, let
-\begin{equation} \label{eq:S(h,G)def} 
-A(h,K) = \big\{ n \mid n + ih \in K \;\forall\; i \in \{0,1,\dots,m\} \big\}
-\end{equation}
-be the set containing all integers $n$ such that the arithmetic progression
-\[
-n, \,\, n + h, \,\, n + 2h, \,\, \dots, \,\, n + mh
-\]
-of length $m+1$ is contained in the subset $K$.
-If $K$ is a small subset of $W$ then $A(h,K)$ might be empty, i.e., there may be no arithmetic progression $n, n+h, \dots$ of length $m+1$ in $K$.
-However, the next two lemmas and the following corollary will show that if $K$ is sufficiently large then it always contains at least one arithmetic progression (for all sufficiently small $h$) and therefore $A(h,K)$ is not empty.
-Denote by $K \backslash \{r\}$ the set $K$ with the element $r$ removed.
-
-\begin{lemma} \label{lem:S(h,G/r)size}
-Let $r \in K$.  For any $h$, removing $r$ from $K$ removes at most $m+1$ arithmetic progressions $n, n+h, \dots n+mh$ of length $m+1$.  That is,
-\[
-|A(h,K \backslash \{r\})| \geq |A(h,K)| - (m+1).
-\]
-\end{lemma}
-\begin{proof}
-The proof follows because there are at most $m+1$ integers, $n$, such that $n+ih = r$ for some $i \in \{0,1,\dots,m\}$.
-That is, there are at most $m+1$ arithmetic progressions of type $n, n+h, \dots n+mh$ that contain $r$.
-\end{proof}
-
- \begin{lemma} \label{lem:S(h,K)size}
- Let $K \subseteq W$.  Then 
-\[
-|A(h,K)| \geq N - mh - (N - |K|)(m+1).
-\]
- \end{lemma}
- \begin{proof}
- Note that $|A(h,W)| = N - mh$.  The proof follows by starting with $A(h,W)$ and applying Lemma~\ref{lem:S(h,G/r)size} 
-\[
-|W|-|K| = N - |K|
-\] 
-times. That is, $K$ can be constructed by removing $N - |K|$ elements from $W$ and this removes at most $(N - |K|)(m+1)$ arithmetic progressions from $A(h,W)$.
- \end{proof}
-
- \begin{corollary} \label{cor:S(h,K)>0}
- Let $K \subseteq W$ such that $|K| > \frac{2m+1}{2m+2}N$.
-Then for all $h$ such that $1\leq h \leq\frac{N}{2m}$ the set $K$ contains at least one arithmetic progression $n, n+h, \dots, n+mh$ of length $m+1$.
-That is, $|A(h,K)| > 0$.
- \end{corollary}
- \begin{proof}
- By substituting the bounds $|K| > \frac{2m+1}{2m+2}N$ and $h \leq\frac{N}{2m}$ into the inequality from Lemma~\ref{lem:S(h,K)size} we immediately obtain $|A(h,K)| > 0$.
- \end{proof}
-
-The next result we require comes from the calculus of finite differences. For any function $d(n)$ mapping $\reals$ to $\reals$, let 
-\[
-\Delta_h d(n) = d(n+h) - d(n)
-\] 
-denote the first difference with interval $h$. Let 
-\begin{align*}
-\Delta_h^2 d(n) = \Delta_h d(n+h) - \Delta_h d(n) = d(n+2h) - 2d(n+h) + d(n)
-\end{align*}
-denote the second difference with interval $h$ and similarly let 
-\begin{equation}\label{eq:mthdiffformula}
-\Delta_h^r d(n) = \Delta_h^{r-1} d(n+h) - \Delta_h^{r-1} d(n) = \sum_{k=0}^{r}\binom{r}{k}(-1)^{r-k}d(n+kh)
-\end{equation}
-denote the $r$th difference with interval $h$. Since $\sum_{k=0}^{r}\binom{r}{k} = 2^r$ it follows that $\Delta_h^r d(n)$ can be represented by adding and subtracting the 
-\[
-d(n), \,\, d(n+h), \,\, \dots, \,\, d(n+kh)
-\] 
-precisely $2^r$ times.
-
-The operator $\Delta_h$ has special properties when applied to polynomials. If $d(n) = a_r n^r + \dots + a_0$ is a polynomial of order $r$ then
- \begin{equation} \label{eq:mfinitediffpoly}
- \Delta_h^r d(n) = h^r r! a_r. 
- \end{equation}
-So, the $r$th difference of the polynomial is a constant depending on $h$, $r$ and the $r$th coefficient $a_r$~\cite[page 51]{Jordan_Calculus_of_finite_difference_1965}.
-We can now continue the proof of strong consistency.
-The next lemma is a key result.
-
-\begin{lemma}\label{lem:moran2}
-Suppose $\lambdabf_1, \lambdabf_2,\dots$ is a sequence of vectors from $B$ with $V_N(\lambdabf_N) - \sigma^2\rightarrow 0$ as $N\rightarrow\infty$.
-Then the elements $\lambda_{0,N}, \dots \lambda_{m,N}$ of $\lambdabf_N$ satisfy $N^k\lambda_{k, N}\rightarrow0$ as $N\rightarrow\infty$.
-\end{lemma}
-\begin{proof}
-Define the function
-\begin{equation}\label{eq:gz}
-g(z) = \expect\fracpart{\Phi_1 + z}^2 - \sigma^2
-\end{equation}
-which is continuous in $z$. Because of~\eqref{eq:Efracpartmined} and~\eqref{eq:Efracpartphi}, $g(z) \geq 0$ with equality only at $z = 0$ for $z \in [-\nicefrac{1}{2}, \nicefrac{1}{2})$. Now
-\[
-V_N(\lambdabf_N) - \sigma^2 = \frac{1}{N}\sum_{n=1}^{N} g\left( \fracpart{ \sum_{k=0}^{m}{n^k \lambda_{k,N}} } \right) \rightarrow 0
-\]
-as $N \rightarrow \infty$. Let
-\[
-z_N(n) = \lambda_{0,N} + \lambda_{1,N} n + \lambda_{2,N} n^2 + \dots + \lambda_{m,N} n^m
-\]
-so that $V_N(\lambdabf_N) - \sigma^2 = \frac{1}{N}\sum_{n=1}^{N} g\left( \fracpart{z_N(n)} \right) \rightarrow 0$.
-%At this stage we can \emph{see} the proof.
-Note that, due to the definition of the identifiable region, $B$, the only circular polynomial $\phi$ with values that are all zeros (and hence sends the sum above to zero) is the circular polynomial whose coefficients are all zeros.
-Obviously, this observation does not constitute a proof, and in particular gives us little indication as to the order of convergence of the coefficients as required by the theorem.
-To make the proof
-Choose constants 
-\[
-c = \frac{2m+1}{2m+2} \qquad \text{and} \qquad 0 < \delta < \frac{1}{2^{2m+1}}
-\]
-and define the set $K_{N}=\left\{  n\leq N \mid \sabs{\fracpart{z_N(n)}} < \delta \right\}$.
-There exists $N_0$ such that for all $N > N_0$ the number of elements in $K_N$ is at least $cN$.
-Too see this, suppose that $\sabs{K_N} < cN$, and let $\gamma$ be the minimum value of $g$ over $[-\nicefrac{1}{2},-\delta] \cup [\delta, \nicefrac{1}{2})$.
-Because $g(0) = 0$ is the unique minimiser of $g$, then $\gamma$ is strictly greater than $0$ and
-\begin{align*}
-V_N(\lambdabf_N) - \sigma^2 = \frac{1}{N}\sum_{n=1}^{N} g\left( \fracpart{z_N(n)} \right) \geq \frac{1}{N}\sum_{n\in K_N} \gamma = (1-c)\gamma,
-\end{align*}
-violating that $V_N(\lambdabf_N) - \sigma^2$ converges to zero as $N \rightarrow \infty$.  We will assume $N > N_0$ in what follows.
-
-From Corollary \ref{cor:S(h,K)>0} it follows that for all $h$ satisfying $1\leq h \leq\frac{N}{2m}$ the set $A(h,K_N)$ contains at least one element, that is, there exists $n' \in A(h,K_N)$ such that all the elements from the arithmetic progression $n', n'+h, \dots, n' + mh$ are in $K_N$ and therefore
-\[
-\sabs{\fracpart{z_N(n')}},\,\, \sabs{\fracpart{z_N(n'+h)}}, \,\, \dots, \,\, \sabs{\fracpart{z_N(n'+mh)}} 
-\] 
-are all less than $\delta$.  Because the $m$th difference is a linear combination of $2^{m}$ elements (see~\eqref{eq:mthdiffformula}) from 
-\[
-\fracpart{z_N(n')}, \,\, \fracpart{z_N(n'+h)}, \,\, \dots, \,\, \fracpart{z_N(n'+mh)}
-\]
-all with magnitude less than $\delta$ we obtain, from Lemma~\ref{lem:fracpartsumanddelta},
-\begin{equation}\label{eq:Deltazfracbound}
-|\fracpart{\Delta_h^m z_N(n')}| \leq |\Delta_h^m \fracpart{ z_N(n')}| \leq \Delta_h^m |\fracpart{ z_N(n')}|< 2^{m}\delta.
-\end{equation}
-From \eqref{eq:mfinitediffpoly} it follows that the left hand side is equal to a constant involving $h$, $m$ and $\lambda_{m,N}$ giving the bound
-\begin{equation}\label{eq:startiterativearg}
-|\fracpart{ h^m m! \lambda_{m,N} }|  = |\fracpart{   \Delta_h^m z_N(n') }| < 2^m\delta
-\end{equation}
-for all $h$ satisfying $1\leq h \leq\frac{N}{2m}$. Setting $h = 1$ and recalling from~\eqref{eq:identifiability} that $\lambda_{m,N} \in [-\tfrac{0.5}{m!}, \tfrac{0.5}{m!})$, we have
- \[
- |\fracpart{ m! \lambda_{m,N} }| = | m! \lambda_{m,N} |< 2^m\delta.
- \]
-Now, because we chose $\delta < \tfrac{1}{2^{2m}}$ it follows that 
-\[
-| \lambda_{m,N} |< \tfrac{2^m}{m!}\delta < \tfrac{1}{m! 2^{m+1}}.
-\]
-So, when $h = 2$, 
-\[
-|\fracpart{ 2^m m! \lambda_{m,N} }| = | 2^m m! \lambda_{m,N} |< 2^m\delta
-\]
-because $2^m m! \lambda_{m,N} \in [-0.5, 0.5)$. Therefore
-\[
-| \lambda_{m,N} |< \frac{1}{m!}\delta < \frac{1}{m! 2^{2m+1}}.
-\]
-Now, with $h = 4$, we similarly obtain 
-\[
-|\fracpart{ 4^m m! \lambda_{m,N} }| = | 4^m m! \lambda_{m,N} |< 2^m\delta
-\]
-and iterating this process we eventually obtain 
-\[
-| \lambda_{m,N} | < \frac{2^m}{2^{um} m!}\delta
-\]
-where $2^u$ is the largest power of 2 less than or equal to $\tfrac{N}{2m}$.  So $2^{u+1} > \frac{N}{2m}$ and substituting this into the inequality above gives 
- \[ 
- |\lambda_{m,N}| < \frac{2^m}{ m! \left(\frac{N}{4m}\right)^m}\delta
- \]
-from which it follows that 
- \begin{equation}\label{eq:enditerativearg}
- N^m|\lambda_{m,N}| < \frac{2^{2m+m}m^m}{m!}\delta
- \end{equation}
-for all $N > N_0$.  As $\delta$ is arbitrary, $N^m \lambda_{m,N} \rightarrow 0$ as $N\rightarrow \infty$.
-
-We have now shown that the highest order coefficient $\lambda_{m,N}$ converges as required.
-The remaining coefficients will be shown to converge by induction.
-Assume that $N^k \lambda_{k,N} \rightarrow 0$ for all $k=r+1, r+2, \dots, m$, that is, assume that the $m-r$ highest order coefficients all converge as required.
-Let
-\[
-z_{N,r}(n) = \lambda_{0,N} + \lambda_{1,N} n + \lambda_{2,N} n^2 + \dots + \lambda_{r,N} n^r.
-\]
-Because the $m-r$ highest order coefficients converge we can write $z_N(n) = z_{N,r}(n) + \gamma(n)$ where $\sup_{n\in\{1,\dots,N\}}\abs{\gamma(n)} \rightarrow 0$ as $N\rightarrow\infty$.
-Now the bound from \eqref{eq:Deltazfracbound}, but applied using the $r$th difference, gives
- \begin{equation}\label{eq:zrbound}
-\left|\fracpart{  \Delta_h^r z_N(n')}\right| = \left|\fracpart{  \Delta_h^r\gamma(n') + \Delta_h^r z_r(n') }\right| = |\fracpart{ \epsilon + h^r r! \lambda_{r,N} }| < 2^r\delta,
- \end{equation}
- where
-\[
-\epsilon = \Delta_h^r \gamma(n') \leq 2^r \sup_{n\in\{1,\dots,N\}}\abs{\gamma(n)} \rightarrow 0
-\] 
-as $N\rightarrow\infty$.
-Choose $\delta$ and $\epsilon$ such that $2^r\delta < \tfrac{1}{4}$ and $|\epsilon| < \tfrac{1}{4}$.
-Then, from \eqref{eq:zrbound} and from Lemma~\ref{lem:fracpartinternalsumlessdelta},
-\[
-\abs{\fracpart{ h^r r! \lambda_{r,N} }} < 2^r\delta + \abs{\epsilon}
-\]
-for all $h$ such that $1 \leq h \leq \tfrac{N}{2m}$.
-Choosing $2^r\delta + |\epsilon| < \frac{1}{2^{2r+1}}$ and using the same iterative process as for the highest order coefficient $\lambda_{m,N}$  (see~\eqref{eq:startiterativearg}~to~\eqref{eq:enditerativearg}) we find that $N^r \lambda_{r,N} \rightarrow 0$ as $N\rightarrow\infty$.
-The proof now follows by induction.
- \end{proof}
-
-% \begin{lemma}\label{sec:proof-strong-cons}
-% For any constants $0 \leq c < 1$ and $\delta>0$ there exists an $N_{0}$ such that for all $N > N_0$ the proportion of $\fracpart{z_N(n)}$ with magnitude less than $\delta$ is greater than $c$.
-That is, the set
-% \[
-% K_{N}=\left\{  n\leq N \mid \vert \fracpart{z(n)} \vert < \delta \right\} 
-% \]
-% has more than $c N$ elements for all $N > N_0$.
-% \end{lemma}
-% \begin{proof}
-% Assume not.
-Then for every $N_0$ there exists an $N > N_0$ such that there are more than $(1-c)N$ integers from $1$ to $N$ with $|\fracpart{z(n)}| > \delta$.
-Let $\gamma$ be the minimum value of $g$ from~(\ref{eq:gz}) over the interval given by the union $[-\nicefrac{1}{2},-\delta] \cup [\delta, \nicefrac{1}{2})$.
-Because $g$ is minimised uniquely at $0$ then $\gamma$ is strictly greater than $0$ and the sum
-% \[
-% \frac{1}{N}\sum_{n=1}^{N} g\left( \fracpart{z(n)} \right) \geq (1-c)\gamma
-% \]
-% with $(1-c)\gamma$ a positive constant. This violates the fact that $g$ converges to zero as $N \rightarrow \infty$ and the lemma is true by contradiction.
-% \end{proof}
-
- \begin{lemma} \label{lem:fracpartsumanddelta}
- Let $a_1, a_2, \dots, a_r$ be $r$ real numbers such  that $\left|\fracpart{a_n}\right| < \delta$ for all $n = 1,2,\dots,r$.  Then $\left|\fracpart{\sum_{n=1}^r{a_n}}\right| < r\delta.$
- \end{lemma}
- \begin{proof}
- If $\delta > \tfrac{1}{2r}$ the proof is trivial as $\left|\fracpart{\sum_{n=1}^r{a_n}}\right| \leq \tfrac{1}{2}$ for all $a_n \in \reals$.
-If $\delta \leq \tfrac{1}{2r}$ then $\fracpart{\sum_{n=1}^r{a_n}} = \sum_{n=1}^r{\fracpart{a_n}}$ and
- \[
- \left\vert\fracpart{\sum_{n=1}^r{a_n}}\right\vert = \left\vert\sum_{n=1}^r{\fracpart{a_n}}\right\vert \leq \sum_{n=1}^r{\left\vert\fracpart{a_n}\right\vert} < r\delta
- \]
- by the triangle inequality.
- \end{proof}
-
-\begin{lemma} \label{lem:fracpartinternalsumlessdelta}
-Let $\left|\fracpart{a + \epsilon}\right| < \delta$ where $|\epsilon| < \nicefrac{1}{4}$ and $0<\delta<\nicefrac{1}{4}$. Then $\left|\fracpart{a}\right| < \delta + |\epsilon|$.
-\end{lemma}
-\begin{proof}
-By supposition $n - \delta < a + \epsilon < n + \delta$ for some $n \in \ints$.  Since $-\delta - \epsilon > -\tfrac{1}{2}$ and $\delta - \epsilon < \tfrac{1}{2}$, it follows that
-\[
-n - \tfrac{1}{2} < n - \delta - \epsilon < a < n + \delta - \epsilon < n + \tfrac{1}{2}.
-\]
-Hence $\fracpart{a} = a - n$ and so
-\[
--\delta - \abs{\epsilon} \leq -\delta - \epsilon < \fracpart{a} < \delta - \epsilon \leq \delta + \abs{\epsilon}
-\]
-and $\abs{\fracpart{a}} \leq \delta + \abs{\epsilon}$.
-
-% The idea behind this proof is to show that under the conditions given $\fracpart{\fracpart{a} + \epsilon}$ does not wrap, i.e.
-that $\fracpart{a} + \epsilon \in \left[-\nicefrac{1}{2}, \nicefrac{1}{2}\right)$.
-To start, assume that $\fracpart{a} + \epsilon \geq \nicefrac{1}{2}$.
-Then $\nicefrac{1}{2} \leq \fracpart{a} + \epsilon < \nicefrac{3}{4}$ and
-% \[
-% -\nicefrac{1}{2} \leq \fracpart{\fracpart{a} + \epsilon} = \fracpart{a + \epsilon} < -\nicefrac{1}{4}
-% \] 
-% and therefore $|\fracpart{a + \epsilon}| > \nicefrac{1}{4} > \delta$, a contradiction.
-Similarly, assume that $\fracpart{a} + \epsilon < -\nicefrac{1}{2}$.
-Then $-\nicefrac{1}{2} > \fracpart{a} + \epsilon > -\nicefrac{3}{4}$ and $\nicefrac{1}{2} > \fracpart{a + \epsilon} > \nicefrac{1}{4}$ and therefore $\left|\fracpart{a + \epsilon}\right| > \nicefrac{1}{4} > \delta$, a contradiction.
-So, $\fracpart{a} + \epsilon \in \left[-\nicefrac{1}{2}, \nicefrac{1}{2}\right)$ and
-% \[
-% |\fracpart{a + \epsilon}| = \left|\fracpart{a} + \epsilon\right| < \delta
-% \]
-% from which it follows that $\left|\fracpart{a}\right| < \delta + \left|\epsilon\right|$.
-\end{proof}
-
-We are now in a position to complete the proof of strong consistency.  Let $A$ be the subset of the sample space on which 
-\[
-V_N(\widehat{\lambdabf}_N) - \sigma^2 \rightarrow 0 \qquad \text{as $N\rightarrow\infty$}.
-\]
-From Corollary~\ref{cor:ESNconvcor}, the $\prob\{A\} =1$.  Let $A'$ be the subset of the sample space on which 
-\[
-N^k\widehat{\lambda}_{k,N} \rightarrow 0 \qquad \text{for $k=0,\dots,m$}
-\] 
-as $N\rightarrow\infty$.  As a result of Lemma~\ref{lem:moran2}, $A \subseteq A'$, and so $\prob\{A'\} \geq \prob\{A\} = 1$.  Strong consistency follows.
-
-\section{Proof of asymptotic normality}\label{sec:centlimitproof}
-
-Let $\psibf$ be the vector with $k$th component $\psi_k = N^k \lambda_k$, $k=0, \dots, m$ and let 
-\[
-T_{N}(\psibf) = S_N(\lambdabf) = \frac{1}{N} \sum_{n=1}^{N} \fracpart{ \Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \psi_k }^2.
-\]
-Let $\widehat{\psibf}_N$ be the vector with elements $\widehat{\psi}_{k,N} = N^k \widehat{\lambda}_{k,N}$ so that $\widehat{\psibf}_N$ is the minimiser of $T_N$.
-Because each of $N^k \widehat{\lambda}_{k,N}$ converges almost surely to zero as $N \rightarrow \infty$, then $\widehat{\psibf}_{N}$ converges almost surely to $\zerobf$ as $N \rightarrow \infty$.
-We want to find the asymptotic distribution of
-\[
-\sqrt{N}\widehat{\psibf}_{N} = 
-\left[
-\begin{array}
-[c]{c}%
-\sqrt{N} \widehat{\psi}_{0,N} \\ \sqrt{N}\widehat{\psi}_{1,N}  \\ \vdots \\ \sqrt{N} \widehat{\psi}_{m,N}
-\end{array}
-\right]
-=
-\left[
-\begin{array}
-[c]{c}%
-\sqrt{N} \widehat{\lambda}_{0,N} \\ N\sqrt{N}\widehat{\lambda}_{1,N} \\ \vdots \\ N^m\sqrt{N} \widehat{\lambda}_{m,N}
-\end{array}
-\right].
-\]
-The proof is complicated by the fact that $T_N$ is not differentiable everywhere as $\fracpart{x}^2$ is not differentiable when $\fracpart{x} = \tfrac{1}{2}$.
-This precludes the use of ``standard approaches'' to proving asymptotic normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.
-However, we show in Lemma~\ref{lem:diffathatpsi} that all the partial derivatives $\frac{\partial T_N}{\partial \psi_\ell}$ for $\ell = 0, \dots, m$ exist, and are equal to zero, at the minimiser $\widehat{\psibf}_N$.
-Thus, putting
-\begin{equation}\label{eq:Wn}
-W_{n} = \round{\Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \widehat{\psi}_{k,N}},
-\end{equation}
-we have,
-\begin{align*}
-0  = \frac{\partial T_N}{\partial \psi_\ell}(\widehat{\psibf}_N) = \frac{2}{N}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\left( \Phi_n - W_{n}+ \sum_{k=0}^{m}(\tfrac{n}{N})^k \widehat{\psi}_{k,N}  \right),
-\end{align*}
-for each $\ell = 0, \dots, m$, so that $D_{\ell,N} = K_{\ell,N}$, where
-\[
-D_{\ell,N} = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}(\tfrac{n}{N})^\ell \Phi_n,
-\]
-and
-\begin{equation}\label{eq:KellN}
-K_{\ell,N} = \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\left(W_{n}- \sum_{k=0}^{m}(\tfrac{n}{N})^k \widehat{\psi}_{k,N}  \right).
-\end{equation}
-Lemma~\ref{lem:Kconvfhalf} shows that, for all $\ell = 0, \dots m$,
-\begin{equation}\label{eq:KellNconv}
-K_{\ell,N} =  (h - 1) \sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big( C_{\ell k} + o_P(1) \big) + o_P(1)
-\end{equation}
-where $C_{\ell k} =  \tfrac{1}{\ell + k + 1}$, and $h = f(-\nicefrac{1}{2})$, and $o_P(1)$ denotes a random variable converging in probability to zero as $N\rightarrow\infty$.
-
-It is now convenient to write in vector form.  Let 
-\begin{equation}\label{eq:dnandkn}
-\kbf_N = 
-\left[
-\begin{array}
-[c]{c}%
-K_{0,N}  \\ \vdots \\ K_{m,N}
-\end{array}
-\right]
-%\qquad \text{and} \qquad 
-=
-\dbf_N = \left[
-\begin{array}
-[c]{c}%
-D_{0,N}  \\ \vdots \\ D_{m,N}
-\end{array}
-\right].
-\end{equation}
-From~\eqref{eq:KellNconv},
-\[
-\dbf_N = \kbf_N = \sqrt{N} (h - 1)(\Cbf + o_P(1)) \widehat{\psibf}_N + o_P(1)
-\]
-where $o_P(1)$ here means a vector or matrix of the appropriate dimension with every element converging in probability to zero as $N\rightarrow\infty$.
-Thus $\sqrt{N}\widehat{\psibf}_N$ has the same asymptotic distribution as $(h - 1)^{-1}\Cbf^{-1} \dbf_N$.
-Lemma~\ref{eq:convdn} shows that $\dbf_N$ is asymptotically normally distributed with zero mean and covariance matrix $\sigma^2\Cbf$.
-Hence $\sqrt{N}\widehat{\psibf}_N$ is asymptotically normal with zero mean and covariance matrix
-\[
-\frac{\sigma^2\Cbf^{-1}\Cbf(\Cbf^{-1})^\prime}{(1 - h)^2} = 
-\frac{\sigma^2\Cbf^{-1}}{(1 - h)^2}.
-\] 
-It remains to prove Lemmas~\ref{lem:diffathatpsi},~\ref{lem:Kconvfhalf} and~\ref{eq:convdn}.
-
-\begin{lemma}\label{lem:diffathatpsi}
-For all $\ell = 0, \dots, m$ the partial derivatives $\frac{\partial T_N}{\partial \psi_\ell}$ exist, and are equal to zero, at the minimiser $\widehat{\psibf}_N$.
-That is $\frac{\partial T_N}{\partial \psi_\ell}(\widehat{\psibf}_N) = 0$ for each $\ell = 0, \dots, m$.
-\end{lemma}
-\begin{proof}
-The function $\fracpart{x}^2$ is differentiable everywhere except if $\fracpart{x} \neq -\nicefrac{1}{2}$.  Recalling that
-\[
-T_{N}(\psibf) = \frac{1}{N} \sum_{n=1}^{N} \fracpart{ \Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \psi_k }^2
-\]
-we see that $T_N$ will be differentiable with respect to $\psibf$ at $\widehat{\psibf}_N$ if 
-\[
-\fracpart{\Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \widehat{\psi}_{k,N}} \neq -\nicefrac{1}{2} \qquad \text{for all $n = 1, \dots, N$.}
-\] 
-This is proved in Lemma~\ref{lem:boundonhatlambda}.
-So the partial derivatives $\frac{\partial T_N}{\partial \psi_\ell}$ exist for all $\ell = 0, \dots, m$ at $\widehat{\psibf}_N$.
-That each of the partial derivatives is equal to zero at $\widehat{\psibf}_N$ follows immediately from the fact that $\widehat{\psibf}_N$ is a minimiser of $T_N$.
-\end{proof}
-
-\begin{lemma}\label{lem:boundonhatlambda} $\sabs{\sfracpart{\Phi_n + \sum_{k=0}^m (\nicefrac{n}{N})^k \widehat{\psi}_{k,N}}} \leq \frac{1}{2} - \frac{1}{2N}$ for all $n = 1, \dots, N$.
-\end{lemma}
-\begin{proof}
-To simplify our notation let $B_n = \Phi_n + \sum_{k=1}^m (\nicefrac{n}{N})^k \widehat{\psi}_{k,N}$ so that we now require to prove $\abs{\fracpart{B_n + \widehat{\psi}_{0,N}}} \leq \frac{1}{2} - \frac{1}{2N}$ for all $n = 1, \dots, N$.
-From~\eqref{eq:Wn}, $W_n = \round{B_n + \widehat{\psi}_{0,N}}$, and
-\begin{align*}
-T_N(\widehat{\psibf}_N) = \frac{1}{N}\sum_{n=1}^N \fracpart{B_n + \widehat{\psi}_{0,N}}^2= \frac{1}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N} - W_n)^2.
-\end{align*}
-Since $\widehat{\psi}_{0,N}$ is the minimiser of the quadratic above,
-\begin{equation}\label{eq:lamsum}
-\widehat{\psi}_{0,N} = -\frac{1}{N}\sum_{n=1}^N(B_n - W_n).
-\end{equation}
-The proof now proceeds by contradiction.  Assume that for some $k$,
-\begin{equation}\label{eq:philamfraccontra}
-\fracpart{B_k + \widehat{\psi}_{0,N}} > \frac{1}{2} - \frac{1}{2N}.
-\end{equation}
-Let $F_n = W_n$ for all $n \neq k$ and $F_k = W_k + 1$, and let
-\[
-\phi = -\frac{1}{N}\sum_{n=1}^N(B_n - F_n) = \widehat{\psi}_{0,N} + \frac{1}{N}.
-\]
-Then,
-\begin{align}
-(B_k + \phi - F_k)^2 &= (B_k + \phi - W_k - 1)^2 \nonumber \\
-&= (B_k + \phi - W_k)^2 - 2(B_k + \phi - W_k) + 1 \nonumber \\
-&= (B_k + \phi - W_k)^2 - 2(B_k + \widehat{\psi}_{0,N} - W_k) + 1 - \frac{2}{N} \nonumber \\
-&= (B_k + \phi - W_k)^2 - 2\fracpart{B_k + \widehat{\psi}_{0,N}} + 1 - \frac{2}{N} \nonumber \\
-&< (B_k + \phi - W_k)^2 - \frac{1}{N}, \label{eq:Bkineq}
-\end{align}
-where the inequality in the last line follows from~\eqref{eq:philamfraccontra}. Let $\bbf = [ \phi,  \widehat{\psi}_{1,N} , \dots , \widehat{\psi}_{m,N}]$
-% \[
-% \bbf = \left[
-% \begin{array}
-% [c]{cccc}%
-% \phi & \widehat{\psi}_{1,N} & \cdots & \widehat{\psi}_{m,N}
-% \end{array}
-% \right]^\prime
-% \]
-be the vector of length $m+1$ with components $b_0 = \phi$ and $b_\ell = \widehat{\psi}_{\ell,N}$ for $\ell = 1 , \dots m$.  Now,
-\[
-N T_N(\bbf) = \sum_{n=1}^N \fracpart{B_n + \phi }^2 \leq  \sum_{n=1}^N (B_n + \phi  - F_n)^2, 
-\]
-and using the inequality from~\eqref{eq:Bkineq},
-\begin{align*}
-N T_N(\bbf) &< - \frac{1}{N} + \sum_{n=1}^N (B_n + \phi  - W_n)^2 \\
-&= - \frac{1}{N} + \sum_{n=1}^N (B_n + \widehat{\psi}_{0,N} + \frac{1}{N}  - W_n)^2 \\
-&= \sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n)^2 +  \frac{2}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n)\\
-%&= N T_N(\widehat{\psibf}_N) + \frac{2}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n) \\
-&= N T_N(\widehat{\psibf}_N),
-\end{align*}
-because $\frac{2}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n) = 0$ as a result of~\eqref{eq:lamsum}.
-But, now $T_N(\bbf) < T_N(\widehat{\psibf}_N)$ violating the fact that $\widehat{\psibf}_N$ is a minimiser of $T_N$.
-So~\eqref{eq:philamfraccontra} is false by contradiction.
-
-If $\fracpart{B_k + \widehat{\psi}_{0,N}} < -\frac{1}{2} + \frac{1}{2N}$ for some $k$, we set $F_k = W_k - 1$ and using the same procedure as before obtain $T_N(\bbf) < T_N(\widehat{\psibf}_N)$ again.
-The proof follows.
-\end{proof}
-
-\begin{lemma}\label{lem:Kconvfhalf}
-With $K_{\ell,N}$ defined as in~\eqref{eq:KellN} and $h = f(-\nicefrac{1}{2})$,
-\[
-K_{\ell,N} = (h - 1) \sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big( C_{\ell k} + o_P(1) \big) + o_P(1) 
-\] 
-for all $\ell = 0, \dots, m$, where $C_{\ell k} =  \frac{1}{\ell + k + 1}$.% is an element of the $m+1$ by $m+1$ Hilbert matrix, and $o_P(1)$ denotes a random variable converging in probability to zero as $N\rightarrow\infty$.
-\end{lemma}
-\begin{proof}
-Care must be taken since $\widehat{\psibf}_N$ depends on the sequence $\{ \Phi_n \}$.  For $n = 1, \dots, N$ and positive $N$, let
-\begin{equation}
-p_{nN}(\psibf) = \sum_{k=0}^{m}\left( \tfrac{n}{N}\right)^k \psi_k,
-\end{equation}
-and put $q_{n}(x) = \round{\Phi_n + x}$ and $Q(x) = \expect q_{n}(x) =  \expect q_{1}(x)$.  Let
-\begin{equation}\label{eq:GNdef}
-G_N(\psibf) = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}\left( \tfrac{n}{N}\right)^\ell  \big(q_n(p_{nN}(\psibf)) - Q(p_{nN}(\psibf)) \big).
-\end{equation}
-and put
-\begin{equation}\label{eq:hatpnN}
-\widehat{p}_{nN} = p_{nN}(\widehat{\psibf}_N) = \sum_{k=0}^m \left(\tfrac{n}{N}\right)^k \widehat{\psi}_{k,N}.
-\end{equation}
-Observe that $G_N$ depends on $\ell$ and we could write $G_{\ell,N}$ but have suppressed the subscript $\ell$ for notational simplicity.
-Now $W_n$ from~\eqref{eq:Wn} can be written as $W_n = \round{\Phi_n + \widehat{p}_{nN}} = q_{n}(\widehat{p}_{nN})$ and $K_{\ell,N}$ from~\eqref{eq:KellN} can be written as
-\begin{align*}
-K_{\ell,N} &= \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\big( q_{n}(\widehat{p}_{nN}) - \widehat{p}_{nN}  \big) \\
-&= \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\big(q_{n}(\widehat{p}_{nN}) - \widehat{p}_{nN} + Q(\widehat{p}_{nN}) - Q(\widehat{p}_{nN}) \big) \\
-&= G_N(\widehat{\psibf}_N) + H_{\ell,N},
-\end{align*}
-where 
-\begin{equation}\label{eq:HellNdef}
-H_{\ell,N} = \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell \big( Q(\widehat{p}_{nN}) - \widehat{p}_{nN} \big).
-\end{equation}
-Lemma~\ref{lem:unifprobG} in the Appendix shows that for any $\delta >0$ and $\nu > 0$ there exists an $\epsilon > 0$ such that
-\[
-\prob\left\{ \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf) } > \delta   \right\} < \nu
-\]
-for all positive integers $N$, where $\|\psibf\|_\infty = \sup_{k} \abs{\psi_k}$.
-Since $\widehat{\psibf}_N$ converges almost surely to zero, it follows that $\lim_{N\rightarrow\infty}\prob\left\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} = 0$ for any $\epsilon > 0$, and therefore $\prob\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \} < \nu$ for all sufficiently large $N$.
-Now
-\begin{align*}
-  \prob\left\{\abs{ G_N(\widehat{\psibf}_N) } > \delta \right\} &= \prob\left\{ \abs{G_N(\widehat{\psibf}_N)} > \delta \;, \; \|\widehat{\psibf}_N\|_{\infty} < \epsilon \right\} \\
-&\hspace{1cm} + \prob\left\{ \abs{G_N(\widehat{\psibf}_N)} > \delta  \;, \; \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} \\
-&\leq \prob\left\{  \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf) } > \delta \right\} + \prob\left\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} \\
-&\leq 2\nu
-\end{align*}
-for all sufficiently large $N$.
-Since $\nu$ and $\delta$ can be chosen arbitrarily small, it follows that $G_N(\widehat{\psibf}_N)$ converges in probability to zero as $N\rightarrow\infty$, and therefore $K_{\ell,N} = H_{\ell,N} + o_P(1)$.
-Lemma~\ref{lem:sumpNhilb} shows that
-\begin{align*}
-H_{\ell,N} =  (h-1)\sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big(C_{\ell k} + o_P(1)\big).
-\end{align*}
-The proof follows.
-\end{proof}
-
-\begin{lemma}\label{lem:sumpNhilb}
-With $H_{\ell,N}$ defined in~\eqref{eq:HellNdef} and $\widehat{p}_{nN}$ defined in~\eqref{eq:hatpnN}, and with $h = f(-\nicefrac{1}{2})$, we have $H_{\ell,N} = (h-1)\sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big(C_{\ell k} + o_P(1)\big)$, where $C_{\ell k} = \tfrac{1}{\ell + k + 1}$.
-\end{lemma}
-\begin{proof}
-If $\abs{x} < 1$, then
-\[
-q_n(x)  = \round{\Phi_n + x} = \begin{cases}
-1, & \Phi_n + x  \geq \nicefrac{1}{2} \\
--1, & \Phi_n + x  < -\nicefrac{1}{2} \\
-0, & \text{otherwise},
-\end{cases}
-\]
-and,
-\begin{align*}
-Q(x) = Eq_1(x) &=  \begin{cases}
-\int_{\nicefrac{1}{2} -x }^{\nicefrac{1}{2}}f(t)\,dt, &   x \geq 0 \\
--\int_{-\nicefrac{1}{2}}^{-\nicefrac{1}{2} - x }f(t)\,dt, &  x  < 0.
-\end{cases} 
-% \\
-% &= \begin{cases}
-% 1 - F(\nicefrac{1}{2} - x ), & x  \geq 0 \\
-% -F(-\nicefrac{1}{2} - x ), &  x < 0,
-% \end{cases} 
-\end{align*}
-Because $f(\fracpart{x})$ is continuous at $-\nicefrac{1}{2}$ it follows from the mean value theorem that
-\[
-Q(x) = x \big( h + \zeta(x) \big),
-\] 
-where $\zeta(x)$ is a function that converges to zero as $x$ converges to zero.  
-%The equation above is still valid when $\abs{x} \geq 1$ by an appropriate choice of the function $\zeta(x)$.  
-Observe that $\abs{\widehat{p}_{nN}} \leq \sum_{k=0}^m\sabs{ \widehat{\psi}_{k,N}}$ and, since each of the $\widehat{\psi}_{k,N} \rightarrow 0$ almost surely as $N\rightarrow\infty$, it follows that $\widehat{p}_{nN} \rightarrow 0$ almost surely uniformly in $n = 1, \dots, N$ as $N\rightarrow\infty$.
-Thus, $\zeta(\widehat{p}_{nN}) \rightarrow 0$ almost surely (and therefore also in probability) uniformly in $n = 1, \dots, N$ as $N\rightarrow\infty$.
-Now,
-\begin{align*}
-Q(\widehat{p}_{nN}) - \widehat{p}_{nN} = \widehat{p}_{nN}\big( h - 1 + \zeta(\widehat{p}_{nN}) \big) = \widehat{p}_{nN}\big( h - 1 + o_P(1) \big), 
-\end{align*}
-and so,
-\begin{align*}
-H_{\ell,N} &= \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell \widehat{p}_{nN}\big( h - 1 + o_P(1) \big)  \\
-&=  \frac{1}{\sqrt{N}} \sum_{n=1}^{N}(\tfrac{n}{N})^\ell \sum_{k=0}^m (\tfrac{n}{N})^k \widehat{\psi}_{k,N} \big( h - 1 + o_P(1) \big)   \\
-&= \sqrt{N}\sum_{k=0}^m \widehat{\psi}_{k,N} \frac{1}{N} \sum_{n=1}^{N} \frac{n^{\ell + k}}{N^{\ell + k + 1}}\big( h - 1 + o_P(1)  \big).
-\end{align*}
-The Riemann sum
-\[
-\frac{1}{N} \sum_{n=1}^{N} \frac{n^{\ell + k}}{N^{\ell + k + 1}} = \int_{0}^{1}x^{k+\ell+1}dx + o_P(1),
-\]
-and since the integral above evaluates to $\frac{1}{k+\ell+1}$, we have
-\[
-H_{\ell,N} = (h - 1)\sqrt{N}\sum_{k=0}^m \widehat{\psi}_{k,N}\left( \frac{1}{k+\ell+1} + o_P(1)  \right)
-\]
-as required.
-% Put
-% \[
-% C_{\ell + k,N} = \sum_{n=1}^{N}\frac{n^{\ell + k}}{N^{\ell + k+1}}.
-% \]
-% It is well known (see for example~\cite[eq. 21]{Peleg1991_CRB_PPS_1991}) that,
-% \begin{equation}\label{eq:ClkNHilbert}
-% C_{\ell + k,N} = C_{\ell k} + O(N^{-1}).
-% \end{equation}
-% Since $\zeta(\widehat{p}_{nN}) = o_P(1)$ for all $n = 1, \dots, N$ then
-% \[
-% \sum_{n=1}^{N}\frac{n^{\ell + k}}{N^{\ell + k + 1}}\zeta(\widehat{p}_{nN}) 
-% \]
-% converges in probability to zero as $N\rightarrow\infty$.  So,
-% \begin{align*}
-% \sum_{n=1}^{N}\frac{n^{\ell + k}}{N^{\ell + k + 1}}\big( h - 1 +  \zeta(\widehat{p}_{nN}) \big) = (h-1)\big( C_{\ell k} + o_P(1) \big).
-% \end{align*}
-% The proof follows by substituting this into~\eqref{eq:HellNre}.
-\end{proof}
-
-\begin{lemma}\label{eq:convdn}
-The distribution of the vector $\dbf_N$, defined in~\eqref{eq:dnandkn}, converges to the multivariate normal with zero mean and covariance matrix $\sigma^2\Cbf$.
-\end{lemma}
-\begin{proof}
-For any constant vector $\alphabf$, let
-\[
-z_N = \alphabf^\prime \dbf_N = \frac{1}{\sqrt{N}} \sum_{n=1}^{N} \Phi_n \sum_{\ell=0}^{m}\alpha_\ell\left( \frac{n}{N} \right)^\ell.
-\]
-By Lypanov's central limit theorem $\zbf_N$ is asymptotically normally distributed with zero mean and variance
-\[
-\lim_{N\rightarrow\infty} \sigma^2 \frac{1}{N} \sum_{n=1}^{N}\left( \sum_{\ell=0}^{m}\alpha_\ell\left( \frac{n}{N} \right)^\ell \right)^2 = \sigma^2 \alphabf^\prime \Cbf \alphabf.
-\]
-By the Cramer-Wold theorem it follows that $\dbf_N$ is asymptotically normally distributed with zero mean and covariance $\sigma^2 \Cbf$. 
-% Recall that the $\ell$th element of $\dbf_N$ is
-% \[
-% D_{\ell,N} = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}(\tfrac{n}{N})^\ell \Phi_n.
-% \]
-% Since the $\Phi_1, \dots, \Phi_N$ are independent, each with variance $\sigma^2$, the distribution of $D_{\ell,N}$ converges to the normal with zero mean and variance
-% \[
-% \lim_{N\rightarrow\infty}\var D_{\ell,N} = \sigma^2 \lim_{N\rightarrow\infty} C_{2\ell,N} = \sigma^2 C_{\ell\ell}
-% \]
-% by Lypanov's central limit theorem and by using~\eqref{eq:ClkNHilbert}.  The covariance between $D_{\ell,N}$ and $D_{k,N}$ is
-% \[
-% \covar(D_{\ell,N}, D_{k,N}) = \sigma^2 C_{\ell+k,N} \rightarrow \sigma^2 C_{\ell k}
-% \]
-% as $N \rightarrow \infty$, again using~\eqref{eq:ClkNHilbert}.  The proof follows.
-\end{proof}
-
-\section{Simulations}\label{sec:simulations}
-
-This section describes the results of Monte-Carlo simulations involving the least squares unwrapping estimator (LSU), the least squares estimator (LSE), Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, and the discrete polynomial phase transform (DPT)~\cite{Peleg_DPT_1995}.
-In all simulations the sample sizes considered are $N = 10, 50$ and $200$ and the unknown amplitude is $\rho = 1$.
-The number of replications of each experiment is 2000 and the $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ are distributed uniformly randomly over the identifiable region.
-
-Figure~\ref{plot:lsetest} shows the performance of the LSU estimator and LSE when $m=2$ (linear chirp).
-Only results for the highest order parameter $\widehat{\mu}_2$ are displayed, the results for the lower order parameters leading to similar conclusions.
-The LSE is computed using the approach described in~\cite[Sec.~10.1]{McKilliam2010thesis} which is based on that of Abatzoglou~\cite{Abatzoglou_ml_chirp_1986}.
-The LSE is computationally intensive.
-We were able to compute it within a reasonable amount of time only with $N$ less than $50$ when $m=2$.
-Results for the exact LSU estimator computed using the sphere decoder and the approximate LSU estimator computed using Babai's algorithm are displayed.
-The sphere decoder is computationally intractable when $N = 200$.
-It is evident that a significant performance penalty is suffered by using Babai's algorithm when the noise variance is large.
-We have also displayed results using another approximate nearest point algorithm called the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder}.
-Our implementation requires $O(N^3\log{N})$ operations.
-This is more computationally expensive than Babai's algorithm but, as can be seen, the $K$-best method is more statistically accurate.
-
-For the simulations displayed in the left hand plot of Figure~\ref{plot:lsetest} the signal model used is that from~\eqref{eq:Y_nsamplednoisey}.
-The noise terms $X_1, \dots, X_N$ are independent and identically distributed zero mean complex Gaussian random variables with independent real and imaginary parts having variance $\sigma_c^2$.
-In this case the LSE is also the maximum likelihood estimator.
-The \term{Cram\`{e}r-Rao lower bound} (CRB) for the variance of unbiased estimators of $\tilde{\mu}_j$ is given by Peleg and Porat~\cite{Peleg1991_CRB_PPS_1991} to be
-\[ 
-\frac{\sigma_c^2}{4\pi^2 N^{2j + 1}} C^{-1}_{jj} + o(N^{-2j-1})
-\]
-where $C^{-1}_{jj}$ is the $j$th diagonal element of the inverse of the $m+1$ by $m+1$ Hilbert matrix $\Cbf$.
-Observe that the performance of the exact LSU estimator (computed using the sphere decoder) is close to the maximum likelihood LSE.
-The asymptotic Cramer-Rao bound (CRB) (solid line) and the asymptotic variance predicted in Theorem~\ref{thm:asymp_proof} (dashed line) are also shown.
-Provided the noise variance is small enough (so that the `threshold' is avoided) the LSE variance is close to the CRB (as is expected) and the LSU variance is close to that predicted by Theorem~\ref{thm:asymp_proof}.
-%Note the gap between the CRB and the variance predicted by Theorem~\ref{thm:asymp_proof}.
-In practice this gap can be overcome using a numerical technique, such as Newton's method, with the LSU estimate as a starting point.
-We have not done this here in order to show the accuracy of Theorem~\ref{thm:asymp_proof}.
-The other noticeable gap is that between the threshold behaviour of the LSU estimator and the LSE.
-When $N=50$ the threshold \emph{kicks in} a little earlier for the LSE than it does for the LSU.
-We suspect that this gap will actually \emph{close} as $N$ increases.
-This hypothesis is supported by the fact that this behaviour is observed in the case of frequency estimation (when $m=1$)~\cite[Sec.
-9.5]{McKilliam2010thesis}\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.
-Unfortunately it is not feasible to test this hypothesis because of the computational complexity of the LSE.
-For the simulations displayed in the right hand plot of Figure~\ref{plot:lsetest} the signal model used is that from~\eqref{eq:noise_circ_poly}.
-The phase noise terms $\Phi_1, \dots, \Phi_N$ are uniform random variables with variance $\sigma^2$.
-In this case, the LSU estimator performs slightly better than the LSE.
-This result agrees with similar observations made in circular statistics~\cite{McKilliam_mean_dir_est_sq_arc_length2010}.
-The asymptotic variance predicted by Theorem~\ref{thm:asymp_proof} again accurately predicts the performance of the LSU estimator.
-
-Figures~\ref{plot:polyest1}~and~~\ref{plot:polyest2} displays the performance of the LSU estimator, the DPT and Kitchen's estimator when $m=3$.
-Both the DPT and Kitchen's estimator perform poorly because they do not work correctly for some parameters in the identifiable region $B$.
-This behaviour has been observed previously~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase,McKilliam_polyphase_est_icassp_2011}.
-Again the performance of the LSU is well predicted by Theorem~\ref{thm:asymp_proof} provided that the noise variance is small enough to avoid the threshold effect and provided that $N$ is sufficiently large.
-The asymptotic CRB is also plotted.
-
-\begin{figure}[p] 
-   	\centering 
-  		\includegraphics[width=\linewidth]{code/lseplot-1.mps} 
-               %\includegraphics{code/lseplot-1.mps}
-   		\caption{Performance of the LSE and LSU estimator for a polynomial phase signal of order two, i.e.
-$m=2$.
-(Top) Mean square error (MSE) in the second order parameter $\mu_2$ for $N=10,50$ and $200$ versus the variance $\sigma_c^2$ of the $X_1, \dots, X_N$.
-The $X_1, \dots, X_N$ are independent and identically distributed (i.i.d.) complex Gaussian random variables.
-(Bottom) MSE in second order parameter $\mu_2$ versus the variance $\sigma^2$ of the $\Phi_1, \dots, \Phi_N$.
-The $\Phi_1, \dots, \Phi_N$ are i.i.d.
-uniform random variables.}
-   		\label{plot:lsetest} 
-  \end{figure}
-
-\begin{figure}[p] 
-   	\centering 
-  		\includegraphics[width=\linewidth]{code/gaussianplot4-1.mps}
-                %\includegraphics{code/gaussianplot4-1.mps}
-  		\caption{Mean square error (MSE) of polynomial phase estimators for $N=10,50$ and $200$ versus the variance $\sigma_c^2$ of the complex noise terms $X_1, \dots X_N$.
-The polynomial phase signal has order 3, and the $X_1, \dots, X_N$ are independent and identically distributed complex Gaussian random variables.
-(Top left) MSE in the phase coefficient $\mu_0$.
-(Top right) MSE in the frequency coefficient $\mu_1$.
-(Bottom left) MSE in the quadratic coefficient $\mu_2$.
-(Bottom right) MSE in the cubic coefficient $\mu_3$.}
-  		\label{plot:polyest1} 
- \end{figure}
-
-\begin{figure}[p] 
-   	\centering 
-  		\includegraphics[width=\linewidth]{code/gaussianplot4-2.mps}
-                %\includegraphics{code/gaussianplot4-1.mps}
-  		\caption{Mean square error (MSE) of polynomial phase estimators for $N=10,50$ and $200$ versus the variance $\sigma_c^2$ of the complex noise terms $X_1, \dots X_N$.
-The polynomial phase signal has order 3, and the $X_1, \dots, X_N$ are independent and identically distributed complex Gaussian random variables.
-(Top left) MSE in the phase coefficient $\mu_0$.
-(Top right) MSE in the frequency coefficient $\mu_1$.
-(Bottom left) MSE in the quadratic coefficient $\mu_2$.
-(Bottom right) MSE in the cubic coefficient $\mu_3$.}
-  		\label{plot:polyest2} 
- \end{figure}
-
-% \section{Computational considerations}
-
-% When $m=0$, polynomial phase estimation is equivalent to the problem of \emph{direction estimation} from circular statistics, and the least squares estimator corresponds with the \emph{extrinsic mean} (or \emph{sample circular mean})~\cite{McKilliam_mean_dir_est_sq_arc_length2010,Fisher1993,Mardia_directional_statistics}.
-In this case the LSE can be computed efficiently in $O(N)$ by averaging $N$ complex numbers.
-When $m=1$, polynomial phase estimation is equivalent to single frequency estimation, and the least squares estimator can be computed efficiently in $O(N\log N)$ operations by maximising the \emph{periodogram} using the fast Fourier transform followed by an appropriate numerical optimisation technique~\cite{Quinn2008maximizing_the_periodogram,Rife1974,Quinn2001}.
-When $m=2$, the least squares estimator can be computed in a reasonable amount of time when $N$ is approximately less than $50$.
-The LSE is computationally intensive for any $N$ when $m > 2$.
-By comparison the exact LSU estimator implemented using the sphere decoder can be computed quickly for any $m$ with $N$ approximately less than $50$.
-In situations where $N$ is small, but high statistical accuracy is required, the sphere decoder is computationally a better choice than the LSE for polynomial phase signals of order greater than or equal to $2$.
-
-% Babai's nearest plane algorithm requires $O(N^2)$ operations and can be run in a reasonable amount of time for any $m$ and reasonably large $N$.
-It should be noted that Babai's nearest plane algorithm requires a Lov\`as reduced lattice basis~\cite{Lenstra1982}.
-This requires $O(N^4)$ operations to compute, but, the Lov\`as reduced basis can be computed once offline.
-It does not need to be computed each time the estimator is run.
-The $K$-best algorithm runs in a reasonable amount of time for any $m$ and when $N$ is less than about $300$.
-If the noise variance is small, then both the sphere decoder and the $K$-best algorithm are actually fast.
-This is because both of these estimators begin with the lattice point found by Babai's nearest plane algorithm.
-If this point is close to (or \emph{is}) the nearest point then both the sphere decoder and the $K$-best algorithm terminate quickly.
-This is in contrast to the least squares estimator that is slow regardless of the noise variance.
-
-% Both Kitchen's estimator and the DPT are fast to compute.
-Kitchen's estimator requires only $O(N)$ arithmetic operations and the DPT requires $O(N \log N)$ operations.
-However, these estimators are not as accurate as the LSE or the LSU.
-They also fail to operate correctly over the entire identifiable region of polynomial phase coefficients~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}~\cite[Ch.~10]{McKilliam2010thesis}.
-
-\section{Conclusion} \label{sec:conclusion}
-
-This paper has considered the estimation of the coefficients of a noisy polynomial phase signal by least squares phase unwrapping (LSU).
-It has been shown that the LSU estimator is strongly consistent and asymptotically normally distributed.
-%Simulations are used to show that the LSU has near maximum likelihood performance in the case that the noise is additive white and Gaussian.
-The LSU estimator can be computed by finding a nearest lattice point in a lattice~\cite{McKilliam2009asilomar_polyest_lattice}.
-%Here, we consider using the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder} to achieve this in reasonable time when the number of observations, $N$, is less than about 300.
-Polynomial time nearest point algorithms for these lattices exist~\cite[Sec 4.3]{McKilliam2010thesis}, but these algorithms are not fast in practice.
-Instead, we considered some general purpose algorithms, the \term{sphere decoder}, \term{Babai's nearest plane algorithm} and the \term{$K$-best algorithm}.
-The sphere decoder and $K$-best algorithm result in accurate estimators.
-
-The major outstanding question is whether faster nearest point algorithms exist for these specific lattices.
-Considering the excellent statistical performance (both theoretically and practically) of the LSU estimator, even fast \emph{approximate} nearest point algorithms are likely to prove useful for the estimation of polynomial phase signals.
-
-\appendix
-
-%\section{Tricks with fractional parts}
-%Throughout the paper, and particularly in  have made use of a number of results involving the fractional part function
-
-\section{An empirical process result}
+ \end{proof}
+
+\section{A tightness result}\label{app:tightness-result}
 
@@
 depends on $n$.
-Nevertheless, we find that the methods of proof from the literature can still be used.
-The proof is based on a technique called \emph{symmetrisation} and another technique called \emph{chaining} (also known as \emph{bracketing})~\cite{Pollard_asymp_empi_proc_1989}.
+Nevertheless, we find that the methods of proof from the literature can still be used if we include an extra result from the literature on hyperplane arrangements~\cite[Ch.
+5]{Chazelle_discrepency_method_2000}\cite[Ch.
+6]{Matousek_lect_disc_geom_2002}.
+Our proof is based on a technique called \emph{symmetrisation} and another technique called \emph{chaining} (also known as \emph{bracketing})~\cite{Pollard_asymp_empi_proc_1989}.
 During the proof we treat $\ell$ as a fixed constant.
@@
 \]
-It follows that,
-\[
-\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)}  <  \sqrt{2\pi} \; \kappa,
-\]
-and by Markov's inequality,
+It follows that $\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)}  <  \sqrt{2\pi} \; \kappa$, and by Markov's inequality,
 \[
@@
 
-The proof of the next lemma is based on a technique called \emph{symmetrisation}~\cite{Gine_Zinn_symmetrisation_1984} and the proof we give closely follows that of Pollard~\cite[Section 4]{Pollard_asymp_empi_proc_1989}.
+The proof of the next lemma is based on a technique called \emph{symmetrisation}~\cite{Gine_Zinn_symmetrisation_1984} and the proof we give closely follows Pollard~\cite[Sec.~4]{Pollard_asymp_empi_proc_1989}.
 
@@
 \end{equation}
-Note that $b(h,\xbf)$ is piecewise constant in $\xbf$.
-Two points $\xbf$ and $\ybf$ from $\reals^{m+1}$ are in the same halfspace of $h$ if and only if $b(h, \xbf) = b(h, \ybf)$.
-Then the pseudometric
+Note that $b(h,\xbf)$ is piecewise constant in $\xbf$.  Two points $\xbf$ and $\ybf$ from $\reals^{m+1}$ are in the same halfspace of $h$ if and only if $b(h, \xbf) = b(h, \ybf)$.  So the pseudometric
 \[

### 2012-10-15 build.sh

@@ -16,6 +16,6 @@ mpost -interaction=nonstopmode lseplot.mp
 cd .. 
 
 pdflatex paper.tex 
-bibtex paper 
+bibtex paper.aux 
 pdflatex paper.tex 
 pdflatex paper.tex 

### 2012-10-15 paper.tex
minor edits to paper and bib.  I think this is starting to look really nice.

@@
 The proof of asymptotic normality is complicated by the fact that the objective function corresponding with the LSU estimator is not differentiable everywhere.
-Techniques developed by Huber~\cite{Huberrobustest1964,Huberbehavmaxnonstandard1967} and refined by Pollard~\cite{Pollard_new_ways_clts_1986,Pollard_asymp_empi_proc_1989} and results from the literature on hyperplane arrangements~\cite{Chazelle_discrepency_method_2000,Matousek_lect_disc_geom_2002} become useful here.
+Techniques developed by Huber~\cite{Huberrobustest1964,Huberbehavmaxnonstandard1967} and Pollard~\cite{Pollard_new_ways_clts_1986,Pollard_asymp_empi_proc_1989} and results from the literature on hyperplane arrangements~\cite{Chazelle_discrepency_method_2000,Matousek_lect_disc_geom_2002} become useful here.
 We are hopeful that the proof techniques developed here will be useful for purposes other than polynomial phase estimation, and in particular other applications involving data that is `wrapped' in some sense.
@@
 \]
-where $\Bbf \in \reals^{n \times d}$ is an $n \times d$ matrix of rank $d$, called the \term{generator matrix} or \term{basis matrix} or simply generator or basis.
-If $n = d$ the lattice is said to be \emph{full rank}.
+where $\Bbf \in \reals^{n \times d}$ is an $n \times d$ matrix of rank $d$, called the generator matrix.
+If $n = d$ the lattice is said to be full rank.
 Lattices are discrete Abelian groups under vector addition.
 They are subgroups of the Euclidean group $\reals^n$.
-Lattices naturally give rise to \emph{tessellations} of $\reals^n$ by the specification of a set of coset representatives for the quotient $\reals^n / \Lambda$.
-One choice for a set of coset representatives is a \emph{fundamental parallelepiped}, i.e., the parallelepiped generated by the columns of a generator matrix.
-Another choice is based on the \emph{Voronoi cell}, i.e., those points from $\reals^n$ nearest (with respect to the Euclidean norm here ) to the lattice point at the origin.
-It is always possible to construct a \emph{rectangular} set of representatives, as the next proposition will show.
+Lattices naturally give rise to tessellations of $\reals^n$ by the specification of a set of coset representatives for the quotient $\reals^n / \Lambda$.
+One choice for a set of coset representatives is a fundamental parallelepiped; the parallelepiped generated by the columns of a generator matrix.
+Another choice is based on the Voronoi cell, i.e., those points from $\reals^n$ nearest (with respect to the Euclidean norm here) to the lattice point at the origin.
+It is always possible to construct a rectangular set of representatives, as the next proposition will show.
 We will use these rectangular regions for describing the aliasing properties of polynomial phase signals in Section~\ref{sec:ident_aliasing}.
@@
 It may be helpful to observe Figures~\ref{fig:circstatplot_line},~\ref{fig:circstatplot_quad}~and~\ref{fig:circstatplot_cube}.
-In these, the phase (divided by $2\pi$) of two distinct polynomial phase signals is plotted on the left, and on the right the principal component of the phase is plotted.
+In these, the phase (divided by $2\pi$) of two distinct polynomial phase signals is plotted on the left, and on the right the principal component of the phase (also divided by $2\pi$) is plotted.
 The circles display the samples at the integers.
@@
 \[
-\coef(g) = \left[ \begin{array}{ccccc} a_0 & a_1 & a_2 & \dots & a_m, \end{array} \right]^\prime
+\coef(g) = \left[ \begin{array}{ccccc} a_0 & a_1 & a_2 & \dots & a_m \end{array} \right]^\prime
 \]
@@
 
-As an example consider the polynomial phase signal of order zero $e^{2\pi j \mu_0} = e^{2\pi j(\mu_0 + k)}$ for any integer $k$.
-In order to ensure identifiability, we must restrict $\mu_0$ to some interval of length $1$.
+As an example consider the polynomial phase signal of order zero $e^{2\pi j \mu_0}$.
+Since $e^{2\pi j \mu_0} = e^{2\pi j(\mu_0 + k)}$ for any integer $k$ we must, in order to ensure identifiability, restrict $\mu_0$ to some interval of length $1$.
 A natural choice is the interval $[-\nicefrac{1}{2}, \nicefrac{1}{2})$.
@@
 
-Given vectors $\xbf$ and $\ybf$ in $\reals^{m+1}$ we say that $\xbf \equiv \ybf \bmod L_{m+1}$ if $\xbf$ and $\ybf$ differ by a lattice point in $L_{m+1}$.
-We define the function $\dealias(\xbf)$ to take $\xbf$ to its coset representative inside $B$.
+We define the function $\dealias(\xbf)$ to take $\xbf\in\reals^{m+1}$ to its coset representative inside $B$.
 That is, $\dealias(\xbf) = \zbf \in B$ where $\xbf - \zbf \in L_{m+1}$.
 %The effect of $\dealias(\xbf)$ is to \emph{dealias} the polynomial phase coefficients.
-When $m = 0$ or $1$ then $\dealias(\xbf) = \fracpart{\xbf}$ where $\fracpart{\xbf} = \xbf - \round{\xbf}$ denotes the (centered) fractional part and $\round{\xbf}$ denotes the nearest integer to $\xbf$ with half integers rounded upwards and both $\fracpart{\cdot}$ and $\round{\cdot}$ operate on vectors elementwise.
+When $m = 0$ or $1$ $\dealias(\xbf) = \fracpart{\xbf}$ where $\fracpart{\xbf} = \xbf - \round{\xbf}$ denotes the (centered) fractional part and $\round{\xbf}$ denotes the nearest integer to $\xbf$ with half integers rounded upwards and both $\fracpart{\cdot}$ and $\round{\cdot}$ operate on vectors elementwise.
 For $m \geq 2$ the function $\dealias(\xbf)$ can be computed by a simple sequential algorithm \cite[Sec.
@@
 Recall that we desire to estimate the coefficients $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ from the noisy samples $Y_1, \dots, Y_N$ given in~\eqref{eq:Y_nsamplednoisey}.
-We take the complex argument of the $Y_n$ and divide by $2\pi$ to obtain
+We take the complex argument of $Y_n$ and divide by $2\pi$ to obtain
 \begin{equation}\label{eq:noise_circ_poly}
@@
 \] 
-are random variables representing the `phase noise' induced by the $X_n$.
+is a random variable representing the `phase noise' induced by $X_n$.
 If the distribution of $X_n$ is circularly symmetric (the phase $\angle X_n$ is uniformly distributed on $[-\pi, \pi)$ and is independent of the magnitude $\abs{X_n}$) then the distribution of $\Phi_n$ is the same as the distribution of $\tfrac{1}{2\pi}\angle(1 + \rho^{-1}X_n)$.
@@
 The existence of practically fast nearest point algorithms for these lattices is an interesting open problem.
-We focus on the asymptotic statistical properties of the LSU estimator, rather than computational aspects in this paper.
-
-The next theorem describes the asymptotic properties of this estimator.
-Before we give the proof it is necessary to understand some of the properties of the phase noise $\Phi_1,\dots,\Phi_N$, which are \emph{circular} random variables with support on $[-\nicefrac{1}{2}, \nicefrac{1}{2})$~\cite{McKilliam2010thesis,McKilliam_mean_dir_est_sq_arc_length2010,Mardia_directional_statistics,Fisher1993}.
+In this paper we focus on the asymptotic statistical properties of the LSU estimator, rather than computational aspects.
+
+The next theorem describes the asymptotic properties of the LSU estimator.
+Before we state the theorem it is necessary to understand some of the properties of the phase noise $\Phi_1,\dots,\Phi_N$, which are \emph{circular} random variables with support on $[-\nicefrac{1}{2}, \nicefrac{1}{2})$~\cite{McKilliam2010thesis,McKilliam_mean_dir_est_sq_arc_length2010,Mardia_directional_statistics,Fisher1993}.
 Circular random variables are often considered modulo $2\pi$ and therefore have support $[-\pi, \pi)$ with $-\pi$ and $\pi$ being identified as equivalent.
@@
 
-The \emph{intrinsic mean} or \emph{Fr\'{e}chet mean} of $\Phi_n$ is defined as~\cite{McKilliam_mean_dir_est_sq_arc_length2010,Bhattacharya_int_ext_means_2003,Bhattacharya_int_ext_means_2005},
+The \emph{intrinsic mean} (or \emph{Fr\'{e}chet mean}) of $\Phi_n$ is defined as~\cite{McKilliam_mean_dir_est_sq_arc_length2010,Bhattacharya_int_ext_means_2003,Bhattacharya_int_ext_means_2005},
 \begin{equation}\label{eq:intrmeandefn}
@@
 However, the two phases are obviously close, since the phases $\pm 0.5$ are actually the same.
-In this case $\dealias(\tilde{\mu}_0 - \widehat{\mu}_0) = \fracpart{\tilde{\mu}_0 - \widehat{\mu}_0} = 0.01$.
+In this case $\dealias(\tilde{\mu}_0 - \widehat{\mu}_0) = \fracpart{\tilde{\mu}_0 - \widehat{\mu}_0} = 0.01$ as expected.
 The same reasoning holds for $m > 0$.
@@
 In the interest of simplicity we only consider the case when $\Phi_1, \dots, \Phi_N$ are identically distributed here.
-If $X_n$ is circularly symmetric with density function nonincreasing with magnitude $\abs{X_n}$, then, the corresponding $\Phi_n$ necessarily has zero intrinsic mean~\cite[Theorem 5.2]{McKilliam2010thesis}.
+If $X_n$ is circularly symmetric with density function nonincreasing with magnitude $\abs{X_n}$, then the corresponding $\Phi_n$ necessarily has zero intrinsic mean~\cite[Theorem 5.2]{McKilliam2010thesis}.
 Thus, our theorem covers commonly used distributions for $X_1, \dots, X_N$, such as the normal distribution.
@@
 The proof of asymptotic normality places requirements on the probability density function $f$ of the phase noise $\Phi_1, \dots, \Phi_N$.
-The requirement that $\Phi_1, \dots, \Phi_N$ have zero intrinsic mean implies $f(-\nicefrac{1}{2}) \leq 1$~\cite[Lemma~1]{McKilliam_mean_dir_est_sq_arc_length2010}, so the only case not handled is when equality holds, i.e., when $f(-\nicefrac{1}{2}) = 1$ or when $f(\fracpart{x})$ is discontinuous at $x = -\nicefrac{1}{2}$.
+The requirement that $\Phi_1, \dots, \Phi_N$ have zero intrinsic mean implies $f(-\nicefrac{1}{2}) \leq 1$~\cite[Lemma~1]{McKilliam_mean_dir_est_sq_arc_length2010}, so the only case not handled is when $f(-\nicefrac{1}{2}) = 1$ or when $f(\fracpart{x})$ is discontinuous at $x = -\nicefrac{1}{2}$.
 In this exceptional case other expressions for the asymptotic variance can be found (similar to \cite[Theorem 3.1]{Hotz_circle_means_2011}), but this comes at a substantial increase in complexity and we have omitted them for this reason.
@@
  \]
-From the definition of the $\dealias(\cdot)$ function $\lambdabf$ is inside $B$ so the elements of $\lambdabf$ satisfy
+From the definition of the $\dealias(\cdot)$ function $\lambdabf \in B$ so the elements of $\lambdabf$ satisfy
  \begin{equation} \label{eq:identifiability}
@@
 \]
-It will follow that $\sabs{S_N(\lambdabf) - V_N(\lambdabf)}$ converges almost sure to zero as $N \rightarrow \infty$ uniformly in $B$, that is, 
+It will follow that
  \begin{equation}\label{eq:SNVNunifmlln}
-\sup_{\lambdabf \in B}\sabs{S_N(\lambdabf) - V_N(\lambdabf)} \rightarrow 0
- \end{equation}
-almost surely as $N\rightarrow\infty$.
-This type of result has been called a uniform law of large numbers~\cite{Pollard_conv_stat_proc_1984} and follows from standard techniques.
+\sup_{\lambdabf \in B}\sabs{S_N(\lambdabf) - V_N(\lambdabf)} \rightarrow 0 \qquad \text{almost surely as $N\rightarrow\infty$}.  
+ \end{equation}
+This type of result has been called a \emph{uniform law of large numbers} and follows from standard techniques~\cite{Pollard_conv_stat_proc_1984}.
 We give a full proof of~\eqref{eq:SNVNunifmlln} in Appendix~\ref{app:uniform-law-large}.
@@
 \begin{proof}
-Let $z$ be the polynomial
-\[
-z(n) = \lambda_0 + \lambda_1 n + \lambda_2 n^2 + \dots + \lambda_m n^m.
-\]
-Then
+Put $z(n) = \lambda_0 + \lambda_1 n + \dots + \lambda_m n^m$.  Then
 \begin{align*}
@@
 of length $m+1$ is contained in the subset $K$.
-If $K$ is a small subset of $W$ then $A(h,K)$ might be empty, i.e., there may be no arithmetic progression $n, n+h, \dots$ of length $m+1$ in $K$.
+If $K$ is a small subset of $W$ then $A(h,K)$ might be empty.
 However, the next two lemmas and the following corollary will show that if $K$ is sufficiently large then it always contains at least one arithmetic progression (for all sufficiently small $h$) and therefore $A(h,K)$ is not empty.
@@
  \begin{lemma} \label{lem:S(h,K)size}
- Let $K \subseteq W$.  Then $|A(h,K)| \geq N - mh - (N - |K|)(m+1)$.
- \end{lemma}
- \begin{proof}
- Note that $|A(h,W)| = N - mh$.  The proof follows by starting with $A(h,W)$ and applying Lemma~\ref{lem:S(h,G/r)size} 
-\[
-|W|-|K| = N - |K|
-\] 
-times. That is, $K$ can be constructed by removing $N - |K|$ elements from $W$ and this removes at most $(N - |K|)(m+1)$ arithmetic progressions from $A(h,W)$.
+ $|A(h,K)| \geq N - mh - (N - |K|)(m+1)$.
+ \end{lemma}
+ \begin{proof}
+ Note that $|A(h,W)| = N - mh$.
+The proof follows by starting with $A(h,W)$ and applying Lemma~\ref{lem:S(h,G/r)size} precisely $|W|-|K|=N-|K|$ times.
+That is, $K$ can be constructed by removing $N - |K|$ elements from $W$ and this removes at most $(N - |K|)(m+1)$ arithmetic progressions from $A(h,W)$.
  \end{proof}
@@
  Let $K \subseteq W$ such that $|K| > \frac{2m+1}{2m+2}N$.
-Then for all $h$ such that $1\leq h \leq\frac{N}{2m}$ the set $K$ contains at least one arithmetic progression $n, n+h, \dots, n+mh$ of length $m+1$.
+For all $h$ such that $1\leq h \leq\frac{N}{2m}$ the set $K$ contains at least one arithmetic progression $n, n+h, \dots, n+mh$ of length $m+1$.
 That is, $|A(h,K)| > 0$.

### 2012-10-15 paper.tex
edits

@@
 \]
-so that $V_N(\lambdabf_N) - \sigma^2 = \frac{1}{N}\sum_{n=1}^{N} g\left( \fracpart{z_N(n)} \right) \rightarrow 0$.
+so that $V_N(\lambdabf_N) - \sigma^2 = \frac{1}{N}\sum_{n=1}^{N} g\left( \fracpart{z_N(n)} \right) \rightarrow 0$ as $N \rightarrow \infty$.
 %At this stage we can \emph{see} the proof.
@@
 \begin{equation}\label{eq:Deltazfracbound}
-|\fracpart{\Delta_h^m z_N(n')}| \leq |\Delta_h^m \fracpart{ z_N(n')}| \leq \Delta_h^m |\fracpart{ z_N(n')}|< 2^{m}\delta.
+|\fracpart{\Delta_h^m z_N(n')}| \leq |\Delta_h^m \fracpart{ z_N(n')}| < 2^{m}\delta.
 \end{equation}
@@
 \[
-| \lambda_{m,N} |< \tfrac{2^m}{m!}\delta < \tfrac{1}{m! 2^{m+1}}.
+| \lambda_{m,N} |< \frac{2^m}{m!}\delta < \frac{1}{m! 2^{m+1}}.
 \]
@@
 \]
-where $2^u$ is the largest power of 2 less than or equal to $\tfrac{N}{2m}$.  So $2^{u+1} > \frac{N}{2m}$ and substituting this into the inequality above gives 
- \[ 
- |\lambda_{m,N}| < \frac{2^m}{ m! \left(\frac{N}{4m}\right)^m}\delta
- \]
-from which it follows that 
+where $2^u$ is the largest power of 2 less than or equal to $\tfrac{N}{2m}$.  %So $2^{u+1} > \frac{N}{2m}$ and substituting this into the inequality above gives 
+%\[
+%|\lambda_{m,N}| < \frac{2^m}{ m! \left(\frac{N}{4m}\right)^m}\delta
+%\] 
+%from which it follows that 
+By substituting $2^{u+1} > \frac{N}{2m}$ it follows that
  \begin{equation}\label{eq:enditerativearg}
@@
 \]
-Because the $m-r$ highest order coefficients converge we can write $z_N(n) = z_{N,r}(n) + \gamma(n)$ where $\sup_{n\in\{1,\dots,N\}}\abs{\gamma(n)} \rightarrow 0$ as $N\rightarrow\infty$.
+Because the $m-r$ highest order coefficients converge we can write $z_N(n) = z_{N,r}(n) + \gamma_N(n)$ where $\sup_{n\in\{1,\dots,N\}}\abs{\gamma_N(n)} \rightarrow 0$ as $N\rightarrow\infty$.
 Now the bound from \eqref{eq:Deltazfracbound}, but applied using the $r$th difference, gives
  \begin{equation}\label{eq:zrbound}
-\left|\fracpart{  \Delta_h^r z_N(n')}\right| = \left|\fracpart{  \Delta_h^r\gamma(n') + \Delta_h^r z_r(n') }\right| = |\fracpart{ \epsilon + h^r r! \lambda_{r,N} }| < 2^r\delta,
+\left|\fracpart{  \Delta_h^r z_N(n')}\right| = \left|\fracpart{  \Delta_h^r\gamma_N(n') + \Delta_h^r z_r(n') }\right| = |\fracpart{ \epsilon + h^r r! \lambda_{r,N} }| < 2^r\delta,
  \end{equation}
@@
 \[
-\epsilon = \Delta_h^r \gamma(n') \leq 2^r \sup_{n\in\{1,\dots,N\}}\abs{\gamma(n)} \rightarrow 0
+\epsilon = \Delta_h^r \gamma_N(n') \leq 2^r \sup_{n\in\{1,\dots,N\}}\abs{\gamma_N(n)} \rightarrow 0
 \] 
@@
  \[
- \left\vert\fracpart{\sum_{n=1}^r{a_n}}\right\vert = \left\vert\sum_{n=1}^r{\fracpart{a_n}}\right\vert \leq \sum_{n=1}^r{\left\vert\fracpart{a_n}\right\vert} < r\delta
- \]
- by the triangle inequality.
+ \left\vert\fracpart{\sum_{n=1}^r{a_n}}\right\vert = \left\vert\sum_{n=1}^r{\fracpart{a_n}}\right\vert \leq \sum_{n=1}^r{\left\vert\fracpart{a_n}\right\vert} < r\delta.
+ \]
  \end{proof}
@@
 \]
-The proof is complicated by the fact that $T_N$ is not differentiable everywhere as $\fracpart{x}^2$ is not differentiable when $\fracpart{x} = \tfrac{1}{2}$.
+The proof is complicated by the fact that $T_N$ is not differentiable everywhere because $\fracpart{x}^2$ is not differentiable when $\fracpart{x} = \tfrac{1}{2}$.
 This precludes the use of `standard approaches' to proving asymptotic normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.
@@
 \[
-T_{N}(\psibf) = \frac{1}{N} \sum_{n=1}^{N} \big( \Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \psi_k  - W_n \big)^2
+T_{N}(\psibf) = \frac{1}{N} \sum_{n=1}^{N} \big( \Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \psi_k  - W_n \big)^2,
 \] 
@@
 Lemma~\ref{eq:convdn} shows that $\dbf_N$ is asymptotically normally distributed with zero mean and covariance matrix $\sigma^2\Cbf$.
-Hence $\sqrt{N}\widehat{\psibf}_N$ is asymptotically normal with zero mean and covariance matrix
+Thus $\sqrt{N}\widehat{\psibf}_N$ is asymptotically normal with zero mean and covariance matrix
 \[
@@
 \begin{proof}
-The function $\fracpart{x}^2$ is differentiable everywhere except if $\fracpart{x} \neq -\frac{1}{2}$.
-Recalling that $T_{N}(\psibf) = \frac{1}{N} \sum_{n=1}^{N} \sfracpart{ \Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \psi_k }^2$ we see that $T_N$ will be differentiable with respect to $\psibf$ at $\widehat{\psibf}_N$ if $\sfracpart{\Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \widehat{\psi}_{k,N}} \neq -\tfrac{1}{2}$ for all $n = 1, \dots, N$.
+The function $\fracpart{x}^2$ is differentiable everywhere except if $\fracpart{x} \neq -\frac{1}{2}$, and so $T_N$ is differentiable with respect to $\psibf$ at $\widehat{\psibf}_N$ if $\sfracpart{\Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \widehat{\psi}_{k,N}} \neq -\tfrac{1}{2}$ for all $n = 1, \dots, N$.
 This is proved in Lemma~\ref{lem:boundonhatlambda}.
 So the partial derivatives $\frac{\partial T_N}{\partial \psi_\ell}$ exist for all $\ell = 0, \dots, m$ at $\widehat{\psibf}_N$.
-That each of the partial derivatives is equal to zero at $\widehat{\psibf}_N$ follows immediately from the fact that $\widehat{\psibf}_N$ is a minimiser of $T_N$.
+That each of the partial derivatives is equal to zero at $\widehat{\psibf}_N$ follows since $\widehat{\psibf}_N$ is a minimiser of $T_N$.
 \end{proof}
@@
 \]
-Then,
+Now,
 \begin{align}
@@
 \begin{lemma}\label{lem:Kconvfhalf}
-With $K_{\ell,N}$ defined as in~\eqref{eq:KellN} and $h = f(-\nicefrac{1}{2})$,
-\[
-K_{\ell,N} = (h - 1) \sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big( C_{\ell k} + o_P(1) \big) + o_P(1) 
-\] 
-for all $\ell = 0, \dots, m$, where $C_{\ell k} =  \frac{1}{\ell + k + 1}$.% is an element of the $m+1$ by $m+1$ Hilbert matrix, and $o_P(1)$ denotes a random variable converging in probability to zero as $N\rightarrow\infty$.
+With $K_{\ell,N}$ defined in~\eqref{eq:KellN}, $h = f(-\nicefrac{1}{2})$, and $C_{\ell k} =  \frac{1}{\ell + k + 1}$ we have $K_{\ell,N} = (h - 1) \sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big( C_{\ell k} + o_P(1) \big) + o_P(1)$ for all $\ell = 0, \dots, m$.% is an element of the $m+1$ by $m+1$ Hilbert matrix, and $o_P(1)$ denotes a random variable converging in probability to zero as $N\rightarrow\infty$.
 \end{lemma}
@@
 \begin{equation}\label{eq:GNdef}
-G_N(\psibf) = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}\left( \tfrac{n}{N}\right)^\ell  \big(q_n(p_{nN}(\psibf)) - Q(p_{nN}(\psibf)) \big),
+G_{\ell,N}(\psibf) = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}\left( \tfrac{n}{N}\right)^\ell  \big(q_n(p_{nN}(\psibf)) - Q(p_{nN}(\psibf)) \big),
 \end{equation}
@@
 \end{equation}
-Observe that $G_N$ depends on $\ell$ and we could write $G_{\ell,N}$ but have suppressed the subscript $\ell$ for notational simplicity.
 Now $W_n$ from~\eqref{eq:Wn} can be written as $W_n = \round{\Phi_n + \widehat{p}_{nN}} = q_{n}(\widehat{p}_{nN})$ and $K_{\ell,N}$ from~\eqref{eq:KellN} can be written as
@@
 &= \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\big(q_{n}(\widehat{p}_{nN}) - \widehat{p}_{nN} + Q(\widehat{p}_{nN}) - Q(\widehat{p}_{nN}) \big) \\
-&= G_N(\widehat{\psibf}_N) + H_{\ell,N},
+&= G_{\ell,N}(\widehat{\psibf}_N) + H_{\ell,N},
 \end{align*}
@@
 \[
-\prob\left\{ \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf) } > \delta   \right\} < \nu
-\]
-for all positive integers $N$, where $\|\psibf\|_\infty = \sup_{k} \abs{\psi_k}$.
+\prob\left\{ \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_{\ell,N}(\psibf) } > \delta   \right\} < \nu
+\]
+for all positive integers $N$ and all $\ell = 0,\dots,m$, where $\|\psibf\|_\infty = \sup_{k} \abs{\psi_k}$.
 Since $\widehat{\psibf}_N$ converges almost surely to zero, it follows that $\lim_{N\rightarrow\infty}\prob\left\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} = 0$ for any $\epsilon > 0$, and therefore $\prob\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \} < \nu$ for all sufficiently large $N$.
@@
 \begin{align*}
-  \prob\left\{\abs{ G_N(\widehat{\psibf}_N) } > \delta \right\} &= \prob\left\{ \abs{G_N(\widehat{\psibf}_N)} > \delta \;, \; \|\widehat{\psibf}_N\|_{\infty} < \epsilon \right\} \\
-&\hspace{1cm} + \prob\left\{ \abs{G_N(\widehat{\psibf}_N)} > \delta  \;, \; \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} \\
-&\leq \prob\left\{  \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf) } > \delta \right\} + \prob\left\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} \\
+  \prob\left\{\abs{ G_{\ell,N}(\widehat{\psibf}_N) } > \delta \right\} &= \prob\left\{ \abs{G_{\ell,N}(\widehat{\psibf}_N)} > \delta \;, \; \|\widehat{\psibf}_N\|_{\infty} < \epsilon \right\} \\
+&\hspace{1cm} + \prob\left\{ \abs{G_{\ell,N}(\widehat{\psibf}_N)} > \delta  \;, \; \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} \\
+&\leq \prob\left\{  \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_{\ell,N}(\psibf) } > \delta \right\} + \prob\left\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} \\
 &\leq 2\nu
@@
 for all sufficiently large $N$.
-Since $\nu$ and $\delta$ can be chosen arbitrarily small, it follows that $G_N(\widehat{\psibf}_N)$ converges in probability to zero as $N\rightarrow\infty$, and therefore $K_{\ell,N} = H_{\ell,N} + o_P(1)$.
+Since $\nu$ and $\delta$ can be chosen arbitrarily small, it follows that $G_{\ell,N}(\widehat{\psibf}_N)$ converges in probability to zero as $N\rightarrow\infty$, and therefore $K_{\ell,N} = H_{\ell,N} + o_P(1)$.
 Lemma~\ref{lem:sumpNhilb} shows that
@@
 \end{align*}
-The proof follows.
+%The proof follows.
 \end{proof}
@@
 \begin{lemma}\label{lem:sumpNhilb}
-With $H_{\ell,N}$ defined in~\eqref{eq:HellNdef} and with $h = f(-\nicefrac{1}{2})$, we have $H_{\ell,N} = (h-1)\sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big(C_{\ell k} + o_P(1)\big)$, where $C_{\ell k} = \tfrac{1}{\ell + k + 1}$.
+With $H_{\ell,N}$ defined in~\eqref{eq:HellNdef}, $h = f(-\nicefrac{1}{2})$, and  $C_{\ell k} = \tfrac{1}{\ell + k + 1}$ we have $H_{\ell,N} = (h-1)\sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big(C_{\ell k} + o_P(1)\big)$.
 \end{lemma}
@@
 \]
-and since the integral above evaluates to $\frac{1}{k+\ell+1}$, we have
-\[
-H_{\ell,N} = (h - 1)\sqrt{N}\sum_{k=0}^m \widehat{\psi}_{k,N}\left( \frac{1}{k+\ell+1} + o_P(1)  \right)
-\]
-as required.
+and since the integral above evaluates to $C_{\ell k} = \frac{1}{k+\ell+1}$, we have
+\[
+H_{\ell,N} = (h - 1)\sqrt{N}\sum_{k=0}^m \widehat{\psi}_{k,N}\big( C_{\ell k} + o_P(1)  \big).C_{\ell k}
+\]
 % Put
@@
 This section describes the results of Monte-Carlo simulations with the least squares unwrapping (LSU) estimator.
-The sample sizes considered are $N = 10, 50$ and $200$ and the unknown amplitude is $\rho = 1$.
+The sample sizes considered are $N = 10, 50, 200$ and the unknown amplitude is $\rho = 1$.
 The $X_1, \dots, X_N$ are pseudorandomly generated independent and identically distributed circularly symmetric complex Gaussian random variables with variance $\sigma_c^2$.
@@
 Figure~\ref{plot:polyest1} displays results for the zeroth and first order coefficients $\widehat{\mu}_0$ and $\widehat{\mu}_1$ and Figure~\ref{plot:polyest2} displays results for the second and third order coefficients $\widehat{\mu}_0$ and $\widehat{\mu}_1$.
-When $N = 10$ and $50$ the LSU estimator can be computed exactly using a general purpose algorithm for finding nearest lattice points called the \emph{sphere~decoder}~\cite{Pohst_sphere_decoder_1981,Agrell2002,Viterbo_sphere_decoder_1999}.
+When $N = 10$ and $50$ the LSU estimator can be computed exactly using a general purpose algorithm for finding nearest lattice points called the sphere~decoder~\cite{Pohst_sphere_decoder_1981,Agrell2002,Viterbo_sphere_decoder_1999}.
 This is displayed be the circles in the figures.
-When $N=200$ the sphere decoder is intractable and we instead use an approximate nearest point algorithm called the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder}.
+When $N=200$ the sphere decoder is computationally intractable and we instead use an approximate nearest point algorithm called the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder}.
 This is displayed by the dots.
@@
 Provided the noise variance is small enough (so that the `threshold' is avoided) the sample MSE of the LSU estimator is close to that predicted by Theorem~\ref{thm:asymp_proof}.
-The Cram\`{e}r-Rao lower bound for the variance of unbiased polynomial phase estimators in Gaussian noise is also plotted with the solid line~\cite{McKilliam_polycrbs_2012,Peleg1991_CRB_PPS_1991}.
+The Cram\`{e}r-Rao lower bound for the variance of unbiased polynomial phase estimators in Gaussian noise is also plotted using the solid line~\cite{McKilliam_polycrbs_2012,Peleg1991_CRB_PPS_1991}.
 When the noise variance is small the asymptotic variance of the LSU estimator is close to the Cram\`{e}r-Rao lower bound.
@@
 \[
-G_N(\psibf) = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}\left( \tfrac{n}{N}\right)^\ell  \big(q_n(p_{nN}(\psibf)) - Q(p_{nN}(\psibf)) \big),
+G_{\ell,N}(\psibf) = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}\left( \tfrac{n}{N}\right)^\ell  \big(q_n(p_{nN}(\psibf)) - Q(p_{nN}(\psibf)) \big),
 \]
 where the functions $q_n$, $Q$ and $p_{nN}$ are defined above~\eqref{eq:GNdef} and $\ell \in \{0, 1, \dots, m\}$.
+To simplify notation we drop the subscript $\ell$ and write $G_{\ell,N}$ as $G_N$ in what follows.
+The proofs we will give hold for any nonegative integer $\ell$.
 
@@
 Our proof is based on a technique called \emph{symmetrisation} and another technique called \emph{chaining} (also known as \emph{bracketing})~\cite{Pollard_asymp_empi_proc_1989}.
-During the proof we treat $\ell$ as a fixed constant.
-The proof holds for any nonegative integer $\ell$.
 
@@
 The proofs here follow those of Pollard~\cite[Section 3]{Pollard_asymp_empi_proc_1989}.
-Within the remaining proofs, we are interested in expectations conditional on the phase noise sequence $\{\Phi_n\}$, so for the purpose of these lemmas, $\{\Phi_n\}$ is treated as a fixed realisation.
+Within the remaining proofs we are interested in expectations conditional on the phase noise sequence $\{\Phi_n\}$, and so $\{\Phi_n\}$ is treated as a fixed realisation.
 For notational convenience we use the abbreviations

### 2012-10-16 paper.tex
minor edits.

@@
 \newcommand{\sign}[1]{\operatorname{sgn}\left( #1 \right)}
+\newcommand{\bfracpart}[1]{\big\langle #1 \big\rangle}
+\newcommand{\babs}[1]{\big\vert #1 \big\vert}
 \endlocaldefs
@@
 We are hopeful that the proof techniques developed here will be useful for purposes other than polynomial phase estimation, and in particular other applications involving data that is `wrapped' in some sense.
-Potential candidates are the phase wrapped images observed in synthetic aperture radar and magnetic resonance imaging~\cite{Nico_phaseunwrappingSAR_2000}.
+Potential candidates are the phase wrapped images observed in modern radar and medical imaging devices such as synthetic aperture radar and magnetic resonance imaging~\cite{Nico_phaseunwrappingSAR_2000}.
 
@@
 
-It is shown in~\cite{McKilliam2009asilomar_polyest_lattice} how this minimisation problem can be posed as that of computing a nearest lattice point in a particular lattice.
-Polynomial time algorithms that compute the nearest point are described in~\cite[Sec 4.3]{McKilliam2010thesis}.
+It is shown in~\cite[Sec~8.1]{McKilliam2010thesis}\cite{McKilliam2009asilomar_polyest_lattice} how this minimisation problem can be posed as that of computing a nearest lattice point in a particular lattice.
+Polynomial time algorithms that compute the nearest point are described in~\cite[Sec.~4.3]{McKilliam2010thesis}.
 Although polynomial in complexity, these algorithms are not fast in practice.
@@
 To prove these stronger properties we need some preliminary results about arithmetic progressions, and from the calculus of finite differences.
-
- %\begin{equation}
- %E\left(  \sum_{n=1}^{N}z_{n}\right)  ^{2\beta}=\sum_{n_{1},\ldots,n_{2\beta
- %}=1}^{N}E\left(  \prod\limits_{j=1}^{2\beta}z_{n_{j}}\right)  .
- %\label{eq:markapp}%
- %\end{equation}
- %Since the terms%
- %\[
- %E\left(  \prod\limits_{j=1}^{2\beta}z_{n_{j}}\right)
- %\]
- %in this containing any $n_{j}$ only once will have $0$ mean, and since the
- %$z_{j}$ are bounded, it follows that%
- %\[
- %\frac{1}{N^{\beta}}E\left(  \sum_{n=1}^{N}z_{n}\right)  ^{2\beta}=O\left(
- %1\right)  ,
- %\]
- %as $N\rightarrow\infty,$ the dominant terms coming from the $z_{j}$ in
- %$\left(  \ref{eq:markapp}\right)  $ occurring in pairs.
 
@@
 %Simulations are used to show that the LSU has near maximum likelihood performance in the case that the noise is additive white and Gaussian.
-It is shown in~\cite{McKilliam2009asilomar_polyest_lattice} how the LSU estimator can be computed by finding a nearest lattice point in a particular lattice.
-Polynomial time algorithms that compute the nearest point are described in~\cite[Sec 4.3]{McKilliam2010thesis}, but these are slow algorithms in practice.
+It is shown in~\cite[Sec~8.1]{McKilliam2010thesis}\cite{McKilliam2009asilomar_polyest_lattice} how the LSU estimator can be computed by finding a nearest lattice point in a particular lattice.
+Polynomial time algorithms that compute the nearest point are described in~\cite[Sec.~4.3]{McKilliam2010thesis}, but these are slow algorithms in practice.
 A significant outstanding question is whether practically fast nearest point algorithms exist for these specific lattices.
@@
 \end{equation}
-almost surely as $N\rightarrow\infty$, where $V_N(\lambdabf) = \expect S_N(\lambdabf)$.  We prove this result here. Put $D_N(\lambdabf) = S_N(\lambdabf) - V_N(\lambdabf)$.  By Lemma~\ref{lem:vn},
-\[
-\sum_{N=1}^\infty \prob \left\{ \sup_{\lambdabf \in B}\abs{ D_N(\lambdabf) } > \epsilon \right\} < \infty
-\]
-for any $\epsilon > 0$.
-Thus~\eqref{eq:SNVNunifmlln2} follows from the Borel-Cantelli lemma.
+almost surely as $N\rightarrow\infty$, where $V_N(\lambdabf) = \expect S_N(\lambdabf)$.
+We prove this result here.
 In what follows we use order notation in the standard way, that is, for functions $h$ and $g$, we write $h(N) = O(g(N))$ to mean that there exists a constant $K > 0$ and a finite $N_0$ such that $h(N) \leq K g(N)$ for all $N > N_0$.
-
-\begin{lemma} \label{lem:vn} 
-For any $\epsilon > 0$ and $0 < c < 8$,
-\[
-\prob \left\{ \sup_{\lambdabf \in B}\abs{ D_N(\lambdabf) } > \epsilon \right\} = O(e^{-c \epsilon^2 N}).
-\]
+Put $D_N(\lambdabf) = S_N(\lambdabf) - V_N(\lambdabf)$.
+Now $\sum_{N=1}^\infty \prob \left\{ \sup_{\lambdabf \in B}\abs{ D_N(\lambdabf) } > \epsilon \right\} < \infty$ for any $\epsilon > 0$ by Lemma~\ref{lem:vn}, and~\eqref{eq:SNVNunifmlln2} follows from the Borel-Cantelli lemma.
+
+\begin{lemma} \label{lem:vn} $\prob \left\{ \sup_{\lambdabf \in B}\abs{ D_N(\lambdabf) } > \epsilon \right\} = O(e^{-c \epsilon^2 N})$ for any $\epsilon > 0$ and $c < 2$.
  \end{lemma}
@@
  for some constant $b>0$.
-Adjacent grid points are separated by $\tfrac{1}{N^b}$ in the first coordinate, $\tfrac{1}{N^{b+1}}$ in the second coordinate and $\tfrac{1}{k!N^{b+k}}$ in the $k$th coordinate.
+Adjacent grid points are separated by $\tfrac{1}{N^b}$ in the zeroth coordinate, $\tfrac{1}{N^{b+1}}$ in the first coordinate and $\tfrac{1}{k!N^{b+k}}$ in the $k$th coordinate.
 Let
@@
 \]
-for any real number $c < 2$.
-%that is, with high probability $D_N$ is small on all of the grid points.  
-In Lemma~\ref{lem:supBVn} we show that
+for any $\epsilon > 0$ and $c < 2$.  In Lemma~\ref{lem:supBVn} we show that
 \[
@@
 \begin{lemma}\label{lem:supVjk}
-For any $\epsilon > 0$ and $0 < c < 8$,
-\[
-\prob\cubr{   \sup_{\rbf \in G}\abs{ D_N(\lambdabf[\rbf])  } > \epsilon } = O(e^{-c \epsilon^2 N}).
-\]
+$\prob\cubr{   \sup_{\rbf \in G}\abs{ D_N(\lambdabf[\rbf])  } > \epsilon } = O(e^{-c \epsilon^2 N})$ for any $\epsilon > 0$ and $c < 8$.
 \end{lemma}
@@
  \[
- Z_{n}=\fracpart{  \Phi_n+\sum_{k = 0}^{m}{\lambda_k n^k} }^{2} - E\fracpart{  \Phi_n + \sum_{k = 0}^{m}{\lambda_k n^k} }^{2}
+ Z_{n}=\fracpart{  \Phi_n+\sum_{k = 0}^{m}{\lambda_k n^k} }^{2} - \expect \fracpart{  \Phi_n + \sum_{k = 0}^{m}{\lambda_k n^k} }^{2}
  \]
  are independent with zero mean and $\abs{Z_n} \leq \tfrac{1}{4}$. From Lemma~\ref{lem:zero_mean_indepent_sum_bound},
- \begin{equation}
-  \prob( \abs{D_N(\lambdabf)} > \epsilon ) \leq 2e^{ - 8 \epsilon^2 N}.
- \label{eq:markapp}%
- \end{equation}
-But,
+ \[
+  \prob\{ \abs{D_N(\lambdabf)} > \epsilon \} \leq 2e^{ - 8 \epsilon^2 N},
+ \]
+and so,
  \begin{align*}
@@
 \begin{lemma}\label{lem:boundedsquarefracparts}
-Let $x$ and $\delta$ be real numbers.  Then
-\[
-\fracpart{x}^2 - |\delta| \leq \fracpart{x + \delta}^2 \leq \fracpart{x}^2 + |\delta|.
-\]
+$\fracpart{x}^2 - |\delta| \leq \fracpart{x + \delta}^2 \leq \fracpart{x}^2 + |\delta|$ for all real numbers $x$ and $\delta$.
 \end{lemma}
@@
 \]
-Also, when $x \in [\nicefrac{1}{2} - \delta, \nicefrac{1}{2})$,
-\[
--\delta \leq 2x + \delta - 1 < \delta,
-\]
-and consequently
+Also, when $x \in [\nicefrac{1}{2} - \delta, \nicefrac{1}{2})$ we  have $-\delta \leq 2x + \delta - 1 < \delta$, and consequently
 \[
@@
 \]
-and therefore, for all $\lambdabf \in B[\rbf]$,
-\[
-\sabs{S_N(\lambdabf) - S_N(\lambdabf[\rbf])} \leq \frac{m+1}{N^b}.
-\]
-As this bound is independent of $\Phi_1 \dots \Phi_N$, we have
+and therefore $\sabs{S_N(\lambdabf) - S_N(\lambdabf[\rbf])} \leq \frac{m+1}{N^b}$ for all $\lambdabf \in B[\rbf]$.  As this bound is independent of $\Phi_1 \dots \Phi_N$, we have
 \[
@@
 \end{align*}
-and the lemma follows because this bound is independent of $\rbf$,
+and the lemma follows because this bound is independent of $\rbf$.
 %\[
@@
 To simplify notation we drop the subscript $\ell$ and write $G_{\ell,N}$ as $G_N$ in what follows.
-The proofs we will give hold for any nonegative integer $\ell$.
+The proof we will give holds for any nonegative integers $\ell$.
 
@@
 depends on $n$.
-Nevertheless, we find that the methods of proof from the literature can still be used if we include an extra result from the literature on hyperplane arrangements~\cite[Ch.
+Nevertheless, we find that the methods of proof from the literature can still be used if we make use of a known result about hyperplane arrangements~\cite[Ch.
 5]{Chazelle_discrepency_method_2000}\cite[Ch.
@@
 and where $\expect$ runs over both $\{g_n\}$ and $\{\Phi_n\}$.
-Conditionally on $\{\Phi_n\}$, $\{Z_N(\psibf), \psibf \in \reals^{m+1} \}$ is a \emph{Gaussian process}, and numerous techniques exist for its analysis.
+Conditionally on $\{\Phi_n\}$, the process $\{Z_N(\psibf), \psibf \in \reals^{m+1} \}$ is a \emph{Gaussian process}, and numerous techniques exist for its analysis.
 Lemma~\ref{lem:chaining} shows that for any $\kappa > 0$ there exists an $\epsilon > 0$ such that
@@
 Taking absolute values followed by supremums,
-\begin{align*}
- \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} &= \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{\frac{1}{\sqrt{N}} \expect_\Phi \sum_{n=1}^{N} (f_{nN} - f_{nN}^\prime )} \\
-&\leq \sup_{\|\psibf\|_{\infty} < \epsilon} \expect_\Phi  \abs{ \frac{1}{\sqrt{N}} \sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) },
-\end{align*}
-the upper bound following from Jensen's inequality.  Since $\sup \expect \abs{\dots} \leq \expect \sup \abs{\dots}$, %the bound can be increased by moving $\expect_{\Phi}$ outside the supremum,
-\begin{equation}\label{eq:Eoutsidesup}
- \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} \leq \expect_\Phi \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} ( f_{nN} - f_{nN}^\prime ) }.
-\end{equation}
-Applying $\expect$ to both sides gives the inequality
+\begin{align}
+ \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} &= \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{\frac{1}{\sqrt{N}} \expect_\Phi \sum_{n=1}^{N} (f_{nN} - f_{nN}^\prime )} \nonumber \\
+&\leq \sup_{\|\psibf\|_{\infty} < \epsilon} \expect_\Phi  \abs{ \frac{1}{\sqrt{N}} \sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) } \nonumber  \\
+&\leq \expect_\Phi \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} ( f_{nN} - f_{nN}^\prime ) }, \label{eq:Eoutsidesup}
+\end{align}
+the second line following from Jensen's inequality and the third line since $\sup \expect \abs{\dots} \leq \expect \sup \abs{\dots}$.  Applying $\expect$ to both sides gives the inequality
 \begin{equation}\label{eq:Gninequfin}
@@
 \end{equation}
-since $\expect \expect_\Phi$ is equivalent to $\expect$.
+because $\expect \expect_\Phi$ is equivalent to $\expect$.
 Let $\sigma_n = \frac{g_n}{\abs{g_n}}$ and put $\sigma_n = 1$ if $g_n = 0$.
 The $\sigma_n$ are independent with $\prob\{ \sigma_n = -1 \} = \tfrac{1}{2}$ and $\prob\{ \sigma_n = 1 \} = \tfrac{1}{2}$.
+The symmetry of the distribtution of $g_n$ implies that $g_n$ and $\sigma_n$ are independent.
 As $\Phi_n$ and $\Phi_n^\prime$ are independent and identically distributed, the random variable $f_{nN} - f_{nN}^\prime$ is symmetrically distributed about zero, and is therefore distributed identically to $\sigma_n( f_{nN} - f_{nN}^\prime)$.
@@
 \end{align*}
-Since $g_n$ is a standard normal random variable, $\expect \abs{g_n} = \expect_\Phi \abs{g_n} = \sqrt{2/\pi}$,
-and
-\begin{align*}
-\expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} } &= \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} \sqrt{\frac{\pi}{2}}\expect_\Phi \abs{g_n}} \\
+Write $\expect_{\Phi\sigma}$ to denote expectation conditional on $\{\Phi_n\}$ and $\{\sigma_n\}$.
+Since $g_n$ is a standard normal random variable, $\expect \abs{g_n} = \expect_{\Phi\sigma} \abs{g_n} = \sqrt{2/\pi}$, and
+\begin{align*}
+\expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} } &= \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} \sqrt{\frac{\pi}{2}}\expect_{\Phi\sigma} \abs{g_n}} \\
 &\leq \sqrt{\frac{\pi}{2}} \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{ \sqrt{N}}\sum_{n=1}^{N} \sigma_n \abs{g_n}  f_{nN}},
 \end{align*}
-the last line following from Jensen's inequality and by moving $\expect_{\Phi}$ outside the supremum similarly to~\eqref{eq:Eoutsidesup}. As $g_n = \sigma_n \abs{g_n}$, it follows that
+the last line following from Jensen's inequality and by moving $\expect_{\Phi\sigma}$ outside the supremum similarly to~\eqref{eq:Eoutsidesup}. As $g_n = \sigma_n \abs{g_n}$, it follows that
 \begin{align*}
 \expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{G_N(\psibf)}  &\leq 2\sqrt{\frac{\pi}{2}} \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} g_n  f_{nN}} \\
-&= \sqrt{2\pi} \; \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ Z_N(\psibf) } 
-\end{align*}
-as required.
+&= \sqrt{2\pi} \; \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ Z_N(\psibf) } .
+\end{align*}
 \end{proof}
@@
 &= \int_{-1/2 + (m+1)\epsilon}^{1/2 - (m+1)\epsilon} f(\fracpart{\phi}) d\phi \\
-&= 2(m+1)\epsilon(f(-\nicefrac{1}{2}) + o(1))
-\end{align*}
-where $o(1)$ goes to zero as $\epsilon$ goes to zero, by the mean value theorem.
-Since $\sqrt{\cdot}$ is a concave function on the positive real line, and since $C_{\epsilon}(\{\Phi_n\})$ is always positive,
-\[
-\expect \sqrt{C_\epsilon(\{\Phi_n\})} \leq  \sqrt{\expect  C_\epsilon(\{\Phi_n\} )} < \sqrt{K_2 \epsilon }
-\]
-by Jensen's inequality, for some constant $K_2$.  Applying $\expect$ to both sides of~\eqref{eq:supZCK1},
+&= 2(m+1)\epsilon \big(f(-\nicefrac{1}{2}) + o(1)\big),
+\end{align*}
+where $o(1)$ goes to zero as $\epsilon$ goes to zero.
+Since $\sqrt{\cdot}$ is a concave function on the positive real line and $C_{\epsilon}(\{\Phi_n\})$ nonegative, it follows from Jensen's inequality that $\expect \sqrt{C_\epsilon(\{\Phi_n\})} \leq  \sqrt{\expect  C_\epsilon(\{\Phi_n\} )} < \sqrt{K_2 \epsilon }$ for some constant $K_2$.
+Applying $\expect$ to both sides of~\eqref{eq:supZCK1} gives
 \[
@@
 \begin{proof}
-Let
-\[
-B_\epsilon = \{\xbf \in \reals \mid \|\xbf\|_\infty < \epsilon \}.
-\]
+Let $B_\epsilon = \{\xbf \in \reals \mid \|\xbf\|_\infty < \epsilon \}$.
 %We are now required to prove that
@@
 
-We now continue the proof.  For all $\psibf \in \reals^{m+1}$, 
-\[
-f_{nN}(\psibf) = \left( \tfrac{n}{N}\right)^\ell \round{\Phi_n + p_{nN}(\psibf)}
-\]
-is a multiple of $N^{-\ell}$ and so $d(\psibf, \psibf^*)$ is a multiple of $N^{-2\ell}$.  When $2^k > C_\epsilon N^{1+2\ell}$ we have 
+We now continue the proof.
+Since $f_{nN}(\psibf) = \left( \tfrac{n}{N}\right)^\ell \round{\Phi_n + p_{nN}(\psibf)}$ is a multiple of $N^{-\ell}$ for all $\psibf \in \reals^{m+1}$ it follows that $d(\psibf, \psibf^*)$ is a multiple of $N^{-2\ell}$.
+When $2^k > C_\epsilon N^{1+2\ell}$ we have
 \[
@@
 \]
-and so we must have $d(\psibf, \psibf^*) = 0$.  Consequently $f_{nN}(\psibf) = f_{nN}(\psibf^*)$  for every $n = 1, \dots, N$, and therefore $Z_N(\psibf) - Z_N(\psibf^*) = 0$.  Thus,
+and so we must have $d(\psibf, \psibf^*) = 0$, and consequently $f_{nN}(\psibf) = f_{nN}(\psibf^*)$  for every $n = 1, \dots, N$ and $Z_N(\psibf) - Z_N(\psibf^*) = 0$.  Thus,
 \[
@@
 \begin{proof}
-Let $b_k$ be a function that maps each $\psibf \in T_\epsilon(k)$ to $b_k(\psibf) \in T_\epsilon(k-1)$ such that 
-\[
-d(\psibf,b_k(\psibf)) \leq 2^{1-k} C_\epsilon N.
-\]
-The existence of the function $b_k$ is guaranteed by the definition of $T_\epsilon(k)$.  By the triangle inequality,
+Let $b_k$ be a function that maps each $\psibf \in T_\epsilon(k)$ to $b_k(\psibf) \in T_\epsilon(k-1)$ such that  $d(\psibf,b_k(\psibf)) \leq 2^{1-k} C_\epsilon N$.
+The existence of the function $b_k$ is guaranteed by the definition of $T_\epsilon(k)$.
+By the triangle inequality,
 \[
@@
 \begin{align*}
-\sigma_X^2 &= \expect_{\Phi} \frac{1}{N} \sum_{n=1}^N g_n^2 ( f_{nN}(\psibf) - f_{nN}(b_k(\psibf)) )^2 \\
-&= \frac{1}{N} \sum_{n=1}^N ( f_{nN}(\psibf) - f_{nN}(b_k(\psibf)) )^2 = d(\psibf, b_k(\psibf)) \leq 2^{1-k} C_\epsilon,
-\end{align*}
-since $\expect_{\Phi} g_n^2 = 1$.  Using Lemma~\ref{lem:maxineq}, 
+\sigma_X^2 &= \expect_{\Phi} \frac{1}{N} \sum_{n=1}^N g_n^2 \big( f_{nN}(\psibf) - f_{nN}(b_k(\psibf)) \big)^2 \\
+&= \frac{1}{N} \sum_{n=1}^N \big( f_{nN}(\psibf) - f_{nN}(b_k(\psibf)) \big)^2 = d(\psibf, b_k(\psibf)) \leq 2^{1-k} C_\epsilon,
+\end{align*}
+because $\expect_{\Phi} g_n^2 = 1$.  Using Lemma~\ref{lem:maxineq}, 
 \begin{align*}
@@
 \end{align*}
-because $\log\abs{T_\epsilon(k)} \leq k(m+1)\log 2 +  \log K_3$.  Taking expectations on both sides of~\eqref{eq:supZchain},
+because $\log\abs{T_\epsilon(k)} \leq k(m+1)\log 2 +  \log K_3$.  Taking expectations on both sides of~\eqref{eq:supZchain} gives
 \begin{align*}
@@
 \begin{lemma}\label{lem:maxineq}(Maximal inequalities)
-Suppose $X_1, \dots, X_N$ are zero mean Gaussian random variables each with variance less than some positive constant $K$, then
-\[
-\expect \sup_{n = 1, \dots, N} \abs{X_n} \leq 3  \sqrt{K \log N}
-\]
-where $\log N$ is the natural logarithm of $N$.
+Suppose $X_1, \dots, X_N$ are zero mean Gaussian random variables each with variance less than some positive constant $K$, then $\expect \sup_{n = 1, \dots, N} \abs{X_n} \leq 3  \sqrt{K \log N}$ where $\log N$ is the natural logarithm of $N$.
 \end{lemma}
@@
 Since at most $\abs{H}/r$ hyperplanes cross the interior of $c$ there can be at most $\abs{H}/r$ hyperplanes between $\zbf$ and $\ybf$, and so $\sigma(\zbf, \ybf) \leq \abs{H}/r$.
-Now,
-\[
-\sigma(\xbf, \ybf) < \sigma(\zbf, \xbf) + \sigma(\zbf, \ybf) \leq \frac{\abs{H}}{r}
-\]
-follows from the triangle inequality.
+Now $\sigma(\xbf, \ybf) < \sigma(\zbf, \xbf) + \sigma(\zbf, \ybf) \leq \frac{\abs{H}}{r}$ follows from the triangle inequality.
 \end{proof}
@@
 Let $B$ be a subset of $\reals^{m+1}$.
-There exists a constant $K$, independent of the set of hyperplanes $H$, such that for every positive real number $r$ there is a discrete subset $T_B \subset B$ containing no more than $K r^{m+1}$ elements with the property that for every $\xbf \in B$ there is a $\ybf \in T_B$ with $\sigma(\xbf, \ybf) \leq \abs{H}/r$.
+There exists a constant $K$, independent of the set of hyperplanes $H$, such that for every positive real number $r$ there is a discrete subset $T_B \subset B$ containing no more than $K r^{m+1}$ elements with the property that for every $\xbf \in B$ there exists $\ybf \in T_B$ with $\sigma(\xbf, \ybf) \leq \abs{H}/r$.
 \end{corollary}
@@
 \begin{proof} (Lemma~\ref{lem:metricentropy})
-Put 
-\[
-g_{nN}(\psibf) = (\tfrac{N}{n})^\ell f_{nN}(\psibf) = \round{ \Phi_n + p_{nN}(\psibf)},
-\]
-and let
+Put $g_{nN}(\psibf) = (\tfrac{N}{n})^\ell f_{nN}(\psibf) = \round{ \Phi_n + p_{nN}(\psibf)}$, and let
 \[
@@
 \]
-We have $d(\psibf, \psibf^*) \leq d_g(\psibf, \psibf^*)$, and so it suffices to prove the Lemma with $d$ replaced by $d_g$.  
+We have $d(\psibf, \psibf^*) \leq d_g(\psibf, \psibf^*)$, and so it suffices to prove the lemma with $d$ replaced by $d_g$.  
 %It will be more convenient to work with $d_g$ rather than $d$.
@@
 \]
-that is, only if $I_\epsilon(\Phi_n) = 1$.  Let
-\[
-A = \{ n \in \{1, \dots, N\} \mid I_\epsilon(\Phi_n) = 1 \}
-\]
-be the subset of the indices where $I_\epsilon(\Phi_n) = 1$.
+that is, only if $I_\epsilon(\Phi_n) = 1$.
+Let $A = \{ n \in \{1, \dots, N\} \mid I_\epsilon(\Phi_n) = 1 \}$ be the subset of the indices where $I_\epsilon(\Phi_n) = 1$.
 By definition the number of elements in $A$ is $C_\epsilon N$ (see~\eqref{eq:Cedefn}).
@@
 \[
-( g_{nN}(\psibf) - g_{nN}(\psibf^*) )^2 \neq 0 
+(g_{nN}(\psibf) - g_{nN}(\psibf^*) )^2 \neq 0 
 \]

### 2012-10-16 paper.tex
spell checked.

@@
 To simplify notation we drop the subscript $\ell$ and write $G_{\ell,N}$ as $G_N$ in what follows.
-The proof we will give holds for any nonegative integers $\ell$.
+The proof we will give holds for any nonnegative integers $\ell$.
 
@@
 The $\sigma_n$ are independent with $\prob\{ \sigma_n = -1 \} = \tfrac{1}{2}$ and $\prob\{ \sigma_n = 1 \} = \tfrac{1}{2}$.
-The symmetry of the distribtution of $g_n$ implies that $g_n$ and $\sigma_n$ are independent.
+The symmetry of the distribution of $g_n$ implies that $g_n$ and $\sigma_n$ are independent.
 As $\Phi_n$ and $\Phi_n^\prime$ are independent and identically distributed, the random variable $f_{nN} - f_{nN}^\prime$ is symmetrically distributed about zero, and is therefore distributed identically to $\sigma_n( f_{nN} - f_{nN}^\prime)$.
@@
 where $o(1)$ goes to zero as $\epsilon$ goes to zero.
-Since $\sqrt{\cdot}$ is a concave function on the positive real line and $C_{\epsilon}(\{\Phi_n\})$ nonegative, it follows from Jensen's inequality that $\expect \sqrt{C_\epsilon(\{\Phi_n\})} \leq  \sqrt{\expect  C_\epsilon(\{\Phi_n\} )} < \sqrt{K_2 \epsilon }$ for some constant $K_2$.
+Since $\sqrt{\cdot}$ is a concave function on the positive real line and $C_{\epsilon}(\{\Phi_n\})$ nonnegative, it follows from Jensen's inequality that $\expect \sqrt{C_\epsilon(\{\Phi_n\})} \leq  \sqrt{\expect  C_\epsilon(\{\Phi_n\} )} < \sqrt{K_2 \epsilon }$ for some constant $K_2$.
 Applying $\expect$ to both sides of~\eqref{eq:supZCK1} gives

### 2012-10-16 paper.tex
minor

@@
 When $m \geq 2$ the computational complexity of the least squares estimator is large, and for this reason authors have considered alternative estimators.
-These can loosely be grouped into two classes; estimators based on polynomial phase transforms, such as the discrete polynomial phase transform~\cite{Peleg_DPT_1995,Porat_asympt_HAF_DPT_1996} and the high order phase function~\cite{Farquharson_another_poly_est_2005}; and estimators based on phase unwrapping, such as Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, and Morelande's Bayesian unwrapping estimator~\cite{Morelande_bayes_unwrapping_2009_tsp}.
+These can loosely be grouped into two classes; estimators based on polynomial phase transforms, such as the discrete polynomial phase transform~\cite{Peleg_DPT_1995} and the high order phase function~\cite{Farquharson_another_poly_est_2005,Porat_asympt_HAF_DPT_1996}; and estimators based on phase unwrapping, such as Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, and Morelande's Bayesian unwrapping estimator~\cite{Morelande_bayes_unwrapping_2009_tsp}.
 

### 2012-10-17 paper.tex
paper mods
minor test.
minor

@@
 That is, two or more distinct polynomial-phase signals can take exactly the same values at the sample points.
-Understanding how aliasing occurs is crucial to understanding the behaviour of the polynomial phase estimators.
+Understanding how aliasing occurs is crucial to understanding the behaviour of polynomial phase estimators.
 The aliasing properties are described in~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}, but, here we present the properties in a way that is better suited to studying the LSU estimator.
@@
 To simplify notation we drop the subscript $\ell$ and write $G_{\ell,N}$ as $G_N$ in what follows.
-The proof we will give holds for any nonnegative integers $\ell$.
+The proof we will give holds for any nonnegative integer $\ell$.
 

### 2012-10-17 paper.tex
minor

@@
 \]
-for all positive integers $N$ and all $\ell = 0,\dots,m$, where $\|\psibf\|_\infty = \sup_{k} \abs{\psi_k}$.
-Since $\widehat{\psibf}_N$ converges almost surely to zero, it follows that $\lim_{N\rightarrow\infty}\prob\left\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} = 0$ for any $\epsilon > 0$, and therefore $\prob\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \} < \nu$ for all sufficiently large $N$.
-Now
+for all positive integers $N$ and all $\ell = 0,\dots,m$, where $\|\psibf\|_\infty = \sup_{k} \abs{\psi_k}$.  Since $\widehat{\psibf}_N$ converges almost surely to zero, it follows that 
+\[
+\lim_{N\rightarrow\infty}\prob\left\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} = 0
+\]
+for any $\epsilon > 0$, and therefore $\prob\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \} < \nu$ for all sufficiently large $N$.  Now
 \begin{align*}

### 2012-10-18 paper.tex
more edits to paper.

@@
 We are hopeful that the proof techniques developed here will be useful for purposes other than polynomial phase estimation, and in particular other applications involving data that is `wrapped' in some sense.
-Potential candidates are the phase wrapped images observed in modern radar and medical imaging devices such as synthetic aperture radar and magnetic resonance imaging~\cite{Nico_phaseunwrappingSAR_2000}.
+Potential candidates are the phase wrapped images observed in modern radar and medical imaging devices such as synthetic aperture radar and magnetic resonance imaging~\cite{Nico_phaseunwrappingSAR_2000,Friedlander_PD_phaseunwrapping_1996}.
 

### 2012-10-19 paper.tex
minor mods

@@
 The proofs here take a similar approach, but require new techniques.
-The proof of asymptotic normality in~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009} contains a subtle error that is amened by the proof given in this paper.
+%The proof of asymptotic normality in~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009} contains a subtle error that is amended by the proof given in this paper.
 %Before begining the proof we first discuss the assumptions made in the theorem statement.
-
 The theorem gives conditions on $\dealias(\tilde{\mubf} - \widehat{\mubf})$ rather than directly on the difference $\tilde{\mubf} - \widehat{\mubf}$.
@@
 \] 
-denote the first difference with interval $h$. Let 
-\begin{align*}
-\Delta_h^2 d(n) = \Delta_h d(n+h) - \Delta_h d(n) = d(n+2h) - 2d(n+h) + d(n)
-\end{align*}
-denote the second difference with interval $h$ and similarly let 
+denote the first difference with interval $h$, and let
+% \begin{align*}
+% \Delta_h^2 d(n) = \Delta_h d(n+h) - \Delta_h d(n) = d(n+2h) - 2d(n+h) + d(n)
+% \end{align*}
+% denote the second difference with interval $h$ and similarly let 
 \begin{equation}\label{eq:mthdiffformula}
@@
 \]
-By the Cramer-Wold theorem it follows that $\dbf_N$ is asymptotically normally distributed with zero mean and covariance $\sigma^2 \Cbf$. 
+By the Cram\`{e}r-Wold theorem it follows that $\dbf_N$ is asymptotically normally distributed with zero mean and covariance $\sigma^2 \Cbf$. 
 % Recall that the $\ell$th element of $\dbf_N$ is
@@
 Figures~\ref{plot:polyest1}~and~\ref{plot:polyest2} show the sample MSEs obtained for a polynomial phase signal of order $m=3$.
-Figure~\ref{plot:polyest1} displays results for the zeroth and first order coefficients $\widehat{\mu}_0$ and $\widehat{\mu}_1$ and Figure~\ref{plot:polyest2} displays results for the second and third order coefficients $\widehat{\mu}_0$ and $\widehat{\mu}_1$.
+Figure~\ref{plot:polyest1} displays results for the zeroth and first order coefficients $\widehat{\mu}_0$ and $\widehat{\mu}_1$ and Figure~\ref{plot:polyest2} displays results for the second and third order coefficients $\widehat{\mu}_2$ and $\widehat{\mu}_3$.
 When $N = 10$ and $50$ the LSU estimator can be computed exactly using a general purpose algorithm for finding nearest lattice points called the sphere~decoder~\cite{Pohst_sphere_decoder_1981,Agrell2002,Viterbo_sphere_decoder_1999}.
-This is displayed be the circles in the figures.
+This is displayed by the circles in the figures.
 When $N=200$ the sphere decoder is computationally intractable and we instead use an approximate nearest point algorithm called the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder}.
@@
 Provided the noise variance is small enough (so that the `threshold' is avoided) the sample MSE of the LSU estimator is close to that predicted by Theorem~\ref{thm:asymp_proof}.
-The Cram\`{e}r-Rao lower bound for the variance of unbiased polynomial phase estimators in Gaussian noise is also plotted using the solid line~\cite{McKilliam_polycrbs_2012,Peleg1991_CRB_PPS_1991}.
+The Cram\`{e}r-Rao lower bound for the variance of unbiased polynomial phase estimators in Gaussian noise is also plotted using the solid line~\cite{Peleg1991_CRB_PPS_1991}.
 When the noise variance is small the asymptotic variance of the LSU estimator is close to the Cram\`{e}r-Rao lower bound.
@@
 %Simulations are used to show that the LSU has near maximum likelihood performance in the case that the noise is additive white and Gaussian.
-It is shown in~\cite[Sec~8.1]{McKilliam2010thesis}\cite{McKilliam2009asilomar_polyest_lattice} how the LSU estimator can be computed by finding a nearest lattice point in a particular lattice.
-Polynomial time algorithms that compute the nearest point are described in~\cite[Sec.~4.3]{McKilliam2010thesis}, but these are slow algorithms in practice.
-A significant outstanding question is whether practically fast nearest point algorithms exist for these specific lattices.
-Considering the excellent statistical performance (both theoretically and practically) of the LSU estimator, even fast \emph{approximate} nearest point algorithms are likely to prove useful for the estimation of polynomial phase signals.
+%It is shown in~\cite[Sec~8.1]{McKilliam2010thesis}\cite{McKilliam2009asilomar_polyest_lattice} how the LSU estimator can be computed by finding a nearest lattice point in a particular lattice.  
+Polynomial time algorithms that compute the LSU estimator are described in~\cite{McKilliam2010thesis}, but these are slow algorithms in practice.
+A significant outstanding question is whether practically fast algorithms exist.
+Considering the excellent statistical performance (both theoretically and practically) of the LSU estimator, even fast \emph{approximate} algorithms are likely to prove useful for the estimation of polynomial phase signals.
 

### 2012-10-21 paper.tex
mofified the plots.  Cut back appendix B proofs

@@
 \end{align*}
-Because $f(\fracpart{x})$ is continuous at $-\nicefrac{1}{2}$ it follows from the mean value theorem that $Q(x) = x \big( h + \zeta(x) \big)$
+Because $f(\fracpart{x})$ is continuous at $-\nicefrac{1}{2}$ it follows that $Q(x) = x \big( h + \zeta(x) \big)$
 where $\zeta(x)$ is a function that converges to zero as $x$ converges to zero.  
@@
 
-Figures~\ref{plot:polyest1}~and~\ref{plot:polyest2} show the sample MSEs obtained for a polynomial phase signal of order $m=3$.
-Figure~\ref{plot:polyest1} displays results for the zeroth and first order coefficients $\widehat{\mu}_0$ and $\widehat{\mu}_1$ and Figure~\ref{plot:polyest2} displays results for the second and third order coefficients $\widehat{\mu}_2$ and $\widehat{\mu}_3$.
+Figure~\ref{plot:polyest1} shows the sample MSEs obtained for a polynomial phase signal of order $m=3$.
+Results are displayed results for the zeroth and third order coefficients $\widehat{\mu}_1$ and $\widehat{\mu}_3$.
+The results for $\widehat{\mu}_1$ and $\widehat{\mu}_2$ lead to similar conclusions.
 When $N = 10$ and $50$ the LSU estimator can be computed exactly using a general purpose algorithm for finding nearest lattice points called the sphere~decoder~\cite{Pohst_sphere_decoder_1981,Agrell2002,Viterbo_sphere_decoder_1999}.
@@
    	\centering 
-  		\includegraphics[width=\linewidth]{code/gaussianplot4-1.mps}
+  		\includegraphics[width=\linewidth]{code/gaussianplot4-3.mps}
                 %\includegraphics{code/gaussianplot4-1.mps}
   		\caption{Sample mean square error (MSE) of the least squares unwrapping estimator for $N=10,50$ and $200$ for a polynomial phase signal of order $m=3$.
-(Top) MSE of the phase coefficient $\mu_0$.
-(Bottom) MSE of the frequency coefficient $\mu_1$.}
+(Top) MSE of the frequency coefficient $\mu_0$.
+(Bottom) MSE of the cubic coefficient $\mu_3$.}
   		\label{plot:polyest1} 
@@
 
-\begin{figure}[p] 
-   	\centering 
-  		\includegraphics[width=\linewidth]{code/gaussianplot4-2.mps}
-                %\includegraphics{code/gaussianplot4-1.mps}
-  		\caption{Sample mean square error (MSE) of the least squares unwrapping estimator for $N=10,50$ and $200$ for a polynomial phase signal of order $m=3$.
+%\begin{figure}[p] 
+%   	\centering 
+%  		\includegraphics[width=\linewidth]{code/gaussianplot4-2.mps}
+%                %\includegraphics{code/gaussianplot4-1.mps}
+%  		\caption{Sample mean square error (MSE) of the least squares unwrapping estimator for $N=10,50$ and $200$ for a polynomial phase signal of order $m=3$.
 (Top) MSE of the quadratic coefficient $\mu_2$.
 (Bottom) MSE of the cubic coefficient $\mu_3$.}
-  		\label{plot:polyest2} 
- \end{figure}
+%  		\label{plot:polyest2} 
+% \end{figure}
 
@@
 
-We first need the following special case of Hoeffding's inequality~\cite{Hoeffding_inequality_1963}.
-
-\begin{lemma} \label{lem:zero_mean_indepent_sum_bound}
-  Let $\bar{Z} = (Z_1 + \dots + Z_N)/N$ where $Z_1, \dots, Z_N$ are independent, zero-mean random variables with $|Z_n| \leq \tfrac{1}{4}$ for each $n$.  Then,
-\[
-  \prob\{ \sabs{\bar{Z}} > \epsilon \} \leq 2e^{ - 8 \epsilon^2 N}.
-\]
-\end{lemma}
+% We first need the following special case of Hoeffding's inequality~\cite{Hoeffding_inequality_1963}.
+
+% \begin{lemma} \label{lem:zero_mean_indepent_sum_bound}
+%   Let $\bar{Z} = (Z_1 + \dots + Z_N)/N$ where $Z_1, \dots, Z_N$ are independent, zero-mean random variables with $|Z_n| \leq \tfrac{1}{4}$ for each $n$.  Then,
+% \[
+%   \prob\{ \sabs{\bar{Z}} > \epsilon \} \leq 2e^{ - 8 \epsilon^2 N}.
+% \]
+% \end{lemma}
 
@@
  \]
- are independent with zero mean and $\abs{Z_n} \leq \tfrac{1}{4}$. From Lemma~\ref{lem:zero_mean_indepent_sum_bound},
+ are independent with zero mean and $\abs{Z_n} \leq \tfrac{1}{4}$. It follows from Hoeffding's inequality~\cite{Hoeffding_inequality_1963} that,%From Lemma~\ref{lem:zero_mean_indepent_sum_bound},
  \[
@@
 To simplify notation we drop the subscript $\ell$ and write $G_{\ell,N}$ as $G_N$ in what follows.
-The proof we will give holds for any nonnegative integer $\ell$.
+The proof we will give holds for any nonnegative integers $\ell$.
 
@@
 This result is related to what is called \emph{tightness} in the literature on empirical processes and weak convergence on metric spaces~\cite{Billingsley1999_convergence_of_probability_measures,Dudley_unif_central_lim_th_1999,Shorak_emp_proc_stat_2009}.
-The lemma is different from what is usually proved in the literature because the function
-\[
-p_{nN}(\psibf) = \sum_{k=0}^m\left( \tfrac{n}{N}\right)^k\psi_k
-\]
+The lemma is different from what is usually proved in the literature because the function $p_{nN}(\psibf) = \sum_{k=0}^m\left( \tfrac{n}{N}\right)^k\psi_k$
 depends on $n$.
-Nevertheless, we find that the methods of proof from the literature can still be used if we make use of a known result about hyperplane arrangements~\cite[Ch.
+Nevertheless, the methods of proof from the literature can be used if we include a known result about hyperplane arrangements~\cite[Ch.
 5]{Chazelle_discrepency_method_2000}\cite[Ch.
@@
 \]
-Let $\{g_n\}$ be a sequence of independent standard normal random variables, independent of the phase noise sequence $\{\Phi_n\}$.  Lemma~\ref{lem:symmetrisation} shows that
+Let $\{g_n\}$ be a sequence of independent standard normal random variables, independent of the phase noise sequence $\{\Phi_n\}$.
+The symmetrisation argument~\cite[Sec.~4]{Pollard_asymp_empi_proc_1989}\cite{Gine_Zinn_symmetrisation_1984} can be used to show that
+%Lemma~\ref{lem:symmetrisation} shows that
 \[
@@
 Conditionally on $\{\Phi_n\}$, the process $\{Z_N(\psibf), \psibf \in \reals^{m+1} \}$ is a \emph{Gaussian process}, and numerous techniques exist for its analysis.
-Lemma~\ref{lem:chaining} shows that for any $\kappa > 0$ there exists an $\epsilon > 0$ such that
-\[
-\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ Z_N(\psibf) } < \kappa.
-\]
-It follows that $\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)}  <  \sqrt{2\pi} \; \kappa$, and by Markov's inequality,
+Lemma~\ref{lem:chaining} shows that for any $\kappa > 0$ there exists an $\epsilon > 0$ such that $\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ Z_N(\psibf) } < \kappa.$  Thus $\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)}  <  \sqrt{2\pi} \; \kappa$, and by Markov's inequality,
 \[
@@
 \]
-for any $\delta > 0$.  The proof follows with $\nu =  \sqrt{2\pi} \kappa/\delta$.  It remains to prove Lemmas~\ref{lem:symmetrisation}~and~\ref{lem:chaining}.
-
-\end{proof}
-
-The proof of the next lemma is based on a technique called \emph{symmetrisation}~\cite{Gine_Zinn_symmetrisation_1984} and the proof we give closely follows Pollard~\cite[Sec.~4]{Pollard_asymp_empi_proc_1989}.
-
-\begin{lemma} \label{lem:symmetrisation}(Symmetrisation)
-\[
-\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} \leq \sqrt{2\pi} \; \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ Z_N(\psibf) }.
-\]
-\end{lemma}
-\begin{proof}
-Let $\{\Phi_n^\prime\}$ be a sequence of random variables distributed identically to $\{\Phi_n\}$ and independent of both $\{\Phi_n\}$ and $\{g_n\}$.
+for any $\delta > 0$.  The proof follows with $\nu =  \sqrt{2\pi} \kappa/\delta$.  It remains to prove Lemma~\ref{lem:chaining}.
+%Lemmas~\ref{lem:symmetrisation}~and~\ref{lem:chaining}.
+
+\end{proof}
+
+% The proof of the next lemma is based on a technique called \emph{symmetrisation}~\cite{Gine_Zinn_symmetrisation_1984} and the proof we give closely follows Pollard~\cite[Sec.~4]{Pollard_asymp_empi_proc_1989}.
+
+% \begin{lemma} \label{lem:symmetrisation}(Symmetrisation)
+% \[
+% \expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} \leq \sqrt{2\pi} \; \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ Z_N(\psibf) }.
+% \]
+% \end{lemma}
+% \begin{proof}
+% Let $\{\Phi_n^\prime\}$ be a sequence of random variables distributed identically to $\{\Phi_n\}$ and independent of both $\{\Phi_n\}$ and $\{g_n\}$.
 Write $\expect_\Phi$ to denote expectation conditional on $\{\Phi_n\}$.
 %That is, expectation taken over $\{g_n\}$ and $\{\Phi_n^\prime\}$ treating $\{\Phi_n\}$ as fixed.
-Since $\Phi_n$ and $\Phi_n^\prime$ are identically distributed $\expect f_{nN}(\psibf, \Phi_n) = \expect_\Phi f_{nN}(\psibf, \Phi_n^\prime)$ so that,
-\begin{align*}
-G_N(\psibf) &= \frac{1}{\sqrt{N}} \sum_{n=1}^{N} \big( f_{nN}(\psibf, \Phi_n) - \expect_\Phi f_{nN}(\psibf, \Phi_n^\prime) \big)  \\
-&= \expect_\Phi \frac{1}{\sqrt{N}} \sum_{n=1}^{N} ( f_{nN} - f_{nN}^\prime ),
-\end{align*}
-where we put $f_{nN} = f_{nN}(\psibf, \Phi_n)$ and  $f_{nN}^\prime = f_{nN}(\psibf, \Phi_n^\prime)$  for notational convenience.
-Taking absolute values followed by supremums,
-\begin{align}
- \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} &= \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{\frac{1}{\sqrt{N}} \expect_\Phi \sum_{n=1}^{N} (f_{nN} - f_{nN}^\prime )} \nonumber \\
-&\leq \sup_{\|\psibf\|_{\infty} < \epsilon} \expect_\Phi  \abs{ \frac{1}{\sqrt{N}} \sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) } \nonumber  \\
-&\leq \expect_\Phi \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} ( f_{nN} - f_{nN}^\prime ) }, \label{eq:Eoutsidesup}
-\end{align}
-the second line following from Jensen's inequality and the third line since $\sup \expect \abs{\dots} \leq \expect \sup \abs{\dots}$.  Applying $\expect$ to both sides gives the inequality
-\begin{equation}\label{eq:Gninequfin}
- \expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} \leq  \expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) },
-\end{equation}
-because $\expect \expect_\Phi$ is equivalent to $\expect$.
+% Since $\Phi_n$ and $\Phi_n^\prime$ are identically distributed $\expect f_{nN}(\psibf, \Phi_n) = \expect_\Phi f_{nN}(\psibf, \Phi_n^\prime)$ so that,
+% \begin{align*}
+% G_N(\psibf) &= \frac{1}{\sqrt{N}} \sum_{n=1}^{N} \big( f_{nN}(\psibf, \Phi_n) - \expect_\Phi f_{nN}(\psibf, \Phi_n^\prime) \big)  \\
+% &= \expect_\Phi \frac{1}{\sqrt{N}} \sum_{n=1}^{N} ( f_{nN} - f_{nN}^\prime ),
+% \end{align*}
+% where we put $f_{nN} = f_{nN}(\psibf, \Phi_n)$ and  $f_{nN}^\prime = f_{nN}(\psibf, \Phi_n^\prime)$  for notational convenience.
+% Taking absolute values followed by supremums,
+% \begin{align}
+%  \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} &= \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{\frac{1}{\sqrt{N}} \expect_\Phi \sum_{n=1}^{N} (f_{nN} - f_{nN}^\prime )} \nonumber \\
+% &\leq \sup_{\|\psibf\|_{\infty} < \epsilon} \expect_\Phi  \abs{ \frac{1}{\sqrt{N}} \sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) } \nonumber  \\
+% &\leq \expect_\Phi \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} ( f_{nN} - f_{nN}^\prime ) }, \label{eq:Eoutsidesup}
+% \end{align}
+% the second line following from Jensen's inequality and the third line since $\sup \expect \abs{\dots} \leq \expect \sup \abs{\dots}$.  Applying $\expect$ to both sides gives the inequality
+% \begin{equation}\label{eq:Gninequfin}
+%  \expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} \leq  \expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) },
+% \end{equation}
+% because $\expect \expect_\Phi$ is equivalent to $\expect$.
 Let $\sigma_n = \frac{g_n}{\abs{g_n}}$ and put $\sigma_n = 1$ if $g_n = 0$.
@@
 Thus,
-%\[
-%\sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) \qquad \text{and} \qquad \sum_{n=1}^{N}  \sigma_n(f_{nN} - f_{nN}^\prime)
-%\]
-%are identically distributed, and therefore
-\begin{align*}
- \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{  \frac{1}{\sqrt{N}}\sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) } = \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{  \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n ( f_{nN} - f_{nN}^\prime) }.
-\end{align*}
-By the triangle inequality
-\[
-\abs{ \sum_{n=1}^{N} \sigma_n (f_{nN} - f_{nN}^\prime)} \leq \abs{ \sum_{n=1}^{N} \sigma_n f_{nN} } + \abs{ \sum_{n=1}^{N}  \sigma_n f_{nN}^\prime },
-\]
-and it follows from~\eqref{eq:Gninequfin}, that
-\begin{align*}
- \expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} &\leq \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} } + \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{\frac{1}{\sqrt{N}} \sum_{n=1}^{N}  \sigma_n  f_{nN}^\prime } \\
-&= 2 \; \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} }.
-\end{align*}
-Write $\expect_{\Phi\sigma}$ to denote expectation conditional on $\{\Phi_n\}$ and $\{\sigma_n\}$.
+% %\[
+% %\sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) \qquad \text{and} \qquad \sum_{n=1}^{N}  \sigma_n(f_{nN} - f_{nN}^\prime)
+% %\]
+% %are identically distributed, and therefore
+% \begin{align*}
+%  \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{  \frac{1}{\sqrt{N}}\sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) } = \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{  \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n ( f_{nN} - f_{nN}^\prime) }.
+% \end{align*}
+% By the triangle inequality
+% \[
+% \abs{ \sum_{n=1}^{N} \sigma_n (f_{nN} - f_{nN}^\prime)} \leq \abs{ \sum_{n=1}^{N} \sigma_n f_{nN} } + \abs{ \sum_{n=1}^{N}  \sigma_n f_{nN}^\prime },
+% \]
+% and it follows from~\eqref{eq:Gninequfin}, that 
+% \begin{align*}
+%  \expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} &\leq \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} } + \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{\frac{1}{\sqrt{N}} \sum_{n=1}^{N}  \sigma_n  f_{nN}^\prime } \\
+% &= 2 \; \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} }.
+% \end{align*}
+% Write $\expect_{\Phi\sigma}$ to denote expectation conditional on $\{\Phi_n\}$ and $\{\sigma_n\}$.
 Since $g_n$ is a standard normal random variable, $\expect \abs{g_n} = \expect_{\Phi\sigma} \abs{g_n} = \sqrt{2/\pi}$, and
-\begin{align*}
-\expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} } &= \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} \sqrt{\frac{\pi}{2}}\expect_{\Phi\sigma} \abs{g_n}} \\
-&\leq \sqrt{\frac{\pi}{2}} \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{ \sqrt{N}}\sum_{n=1}^{N} \sigma_n \abs{g_n}  f_{nN}},
-\end{align*}
-the last line following from Jensen's inequality and by moving $\expect_{\Phi\sigma}$ outside the supremum similarly to~\eqref{eq:Eoutsidesup}. As $g_n = \sigma_n \abs{g_n}$, it follows that
-\begin{align*}
-\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{G_N(\psibf)}  &\leq 2\sqrt{\frac{\pi}{2}} \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} g_n  f_{nN}} \\
-&= \sqrt{2\pi} \; \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ Z_N(\psibf) } .
-\end{align*}
-\end{proof}
+% \begin{align*}
+% \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} } &= \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} \sqrt{\frac{\pi}{2}}\expect_{\Phi\sigma} \abs{g_n}} \\
+% &\leq \sqrt{\frac{\pi}{2}} \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{ \sqrt{N}}\sum_{n=1}^{N} \sigma_n \abs{g_n}  f_{nN}},
+% \end{align*}
+% the last line following from Jensen's inequality and by moving $\expect_{\Phi\sigma}$ outside the supremum similarly to~\eqref{eq:Eoutsidesup}. As $g_n = \sigma_n \abs{g_n}$, it follows that
+% \begin{align*}
+% \expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{G_N(\psibf)}  &\leq 2\sqrt{\frac{\pi}{2}} \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} g_n  f_{nN}} \\
+% &= \sqrt{2\pi} \; \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ Z_N(\psibf) } .
+% \end{align*}
+% \end{proof}
 
@@
 \[
-\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ Z_N(\psibf) } < \kappa,
-\]
-where $Z_N(\psibf)$ is defined by~\eqref{eq:ZpsiCondGaussProc}.
+\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ Z_N(\psibf) } < \kappa.
+\]
+%where $Z_N(\psibf)$ is defined by~\eqref{eq:ZpsiCondGaussProc}.
 \end{lemma}
@@
 \begin{equation}\label{eq:supZCK1} 
-\expect_\Phi \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ Z_N(\psibf) } \leq K_1 \sqrt{C_\epsilon(\{\Phi_n\})}
+\expect_\Phi \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ Z_N(\psibf) } \leq K_1 \sqrt{C_\epsilon(\{\Phi_n\})},
 \end{equation}
@@
 \begin{equation}\label{eq:Cedefn}
-C_\epsilon(\{\Phi_n\}) = \frac{1}{N} \sum_{n=1}^{N} I_\epsilon(\abs{\Phi_n})
+C_\epsilon(\{\Phi_n\}) = \frac{1}{N} \sum_{n=1}^{N} I_\epsilon(\abs{\Phi_n}),
 \end{equation}
@@
 \begin{align*}
-\expect C_\epsilon(\{\Phi\}) &= \frac{1}{N} \sum_{n=1}^{N} \expect I_\epsilon(\abs{\Phi_n}) \\
+\expect C_\epsilon(\{\Phi_n\}) &= \frac{1}{N} \sum_{n=1}^{N} \expect I_\epsilon(\abs{\Phi_n}) \\
 &= \prob\cubr{\abs{\Phi_1} \geq \nicefrac{1}{2} - (m+1)\epsilon} \\
@@
 where $o(1)$ goes to zero as $\epsilon$ goes to zero.
-Since $\sqrt{\cdot}$ is a concave function on the positive real line and $C_{\epsilon}(\{\Phi_n\})$ nonnegative, it follows from Jensen's inequality that $\expect \sqrt{C_\epsilon(\{\Phi_n\})} \leq  \sqrt{\expect  C_\epsilon(\{\Phi_n\} )} < \sqrt{K_2 \epsilon }$ for some constant $K_2$.
+Since $\sqrt{\cdot}$ is a concave function on the positive real line and $C_{\epsilon}(\{\Phi_n\})$ is nonnegative, it follows from Jensen's inequality that $\expect \sqrt{C_\epsilon(\{\Phi_n\})} \leq  \sqrt{\expect  C_\epsilon(\{\Phi_n\} )} < \sqrt{K_2 \epsilon }$ for some constant $K_2$.
 Applying $\expect$ to both sides of~\eqref{eq:supZCK1} gives
@@
 
-The proofs of Lemmas~\ref{lem:chaining2}~and~\ref{lem:chaining3} are based on a technique called \emph{chaining} (or \emph{bracketing})~\cite{Dudley_unif_central_lim_th_1999,Ossiander_clt_bracketing_1984,Pollard_asymp_empi_proc_1989,Pollard_new_ways_clts_1986}.
+%The proofs of Lemmas~\ref{lem:chaining2}~and~\ref{lem:chaining3} are based on a technique called \emph{chaining} (or \emph{bracketing})~\cite{Dudley_unif_central_lim_th_1999,Ossiander_clt_bracketing_1984,Pollard_asymp_empi_proc_1989,Pollard_new_ways_clts_1986}.
 The proofs here follow those of Pollard~\cite[Section 3]{Pollard_asymp_empi_proc_1989}.
-Within the remaining proofs we are interested in expectations conditional on the phase noise sequence $\{\Phi_n\}$, and so $\{\Phi_n\}$ is treated as a fixed realisation.
-For notational convenience we use the abbreviations
-\[
-C_\epsilon = C_\epsilon(\{\Phi_n\}) \qquad \text{and} \qquad f_{nN}(\psibf) = f_{nN}(\psibf, \Phi_n)
-\]
-in what follows.
+The remaining lemmas consider a fixed realisation $\{\Phi_n\}$, and we use the abbreviations $C_\epsilon = C_\epsilon(\{\Phi_n\})$ and $f_{nN}(\psibf) = f_{nN}(\psibf, \Phi_n)$.
 As in Lemma~\ref{lem:chaining} we assume, without loss of generality, that $\epsilon < \tfrac{1}{m+1}$.
@@
 \]
-We specifically define $T_\epsilon(0)$ to contain a single point, that being the origin $\zerobf$.  Defined this way, $T_\epsilon(0)$ satisfies the inequality above because
-\[
-d(\psibf, \zerobf) = \sum_{n = 1}^N f_{nN}(\psibf)^2  \leq C_\epsilon N ,
-\]
-for all $\psibf \in B_\epsilon$, as a result of Lemma~\ref{lem:epslmlemma}.
-
+We define $T_\epsilon(0)$ to contain a single point, the origin $\zerobf$.
+Defined this way, $T_\epsilon(0)$ satisfies the inequality above because $d(\psibf, \zerobf) = \sum_{n = 1}^N f_{nN}(\psibf)^2  \leq C_\epsilon N$ for all $\psibf \in B_\epsilon$, as a result of Lemma~\ref{lem:epslmlemma}.
 The existence of $T_\epsilon(k)$ for each positive integer $k$ will be proved in Lemma~\ref{lem:metricentropy}.
-It is worth giving some intuition regarding $T_\epsilon(k)$.
+%It is worth giving some intuition regarding $T_\epsilon(k)$.
 If we place a `ball' of radius $2^{-k} C_\epsilon N$  with respect to the pseudometric $d(\cdot, \cdot)$ around each point in $T_\epsilon(k)$, then, by definition, the union of these balls is a superset of $B_\epsilon$.
@@
 The minimum number of such balls required to cover $B_\epsilon$ is called a \emph{covering number} of $B_\epsilon$.
-In Lemma~\ref{lem:metricentropy} we show that no more than $K_3 2^{(m+1)k}$ balls of radius $2^{-k} C_\epsilon N$ are required to cover $B_\epsilon$, where $K_3$ is a constant, independent of $N$ and $\epsilon$.
+%In Lemma~\ref{lem:metricentropy} we show that no more than $K_3 2^{(m+1)k}$ balls of radius $2^{-k} C_\epsilon N$ are required to cover $B_\epsilon$, where $K_3$ is a constant, independent of $N$ and $\epsilon$.
 We make use of this upper bound in Lemma~\ref{lem:chaining3}.
-
-We now continue the proof.
-Since $f_{nN}(\psibf) = \left( \tfrac{n}{N}\right)^\ell \round{\Phi_n + p_{nN}(\psibf)}$ is a multiple of $N^{-\ell}$ for all $\psibf \in \reals^{m+1}$ it follows that $d(\psibf, \psibf^*)$ is a multiple of $N^{-2\ell}$.
-When $2^k > C_\epsilon N^{1+2\ell}$ we have
-\[
-0 \leq d(\psibf, \psibf^*) < 2^{-k} C_\epsilon N < N^{-2\ell},
-\]
-and so we must have $d(\psibf, \psibf^*) = 0$, and consequently $f_{nN}(\psibf) = f_{nN}(\psibf^*)$  for every $n = 1, \dots, N$ and $Z_N(\psibf) - Z_N(\psibf^*) = 0$.  Thus,
+The lemma also upper bounds the number of elements by $\abs{T_\epsilon(k)} < K_3 2^{(m+1)k}$ where $K_3$ is a constant independent of $N$ and $\epsilon$.
+
+Since $f_{nN}(\psibf) = \left( \tfrac{n}{N}\right)^\ell \round{\Phi_n + p_{nN}(\psibf)}$ is a multiple of $N^{-\ell}$ for all $\psibf \in \reals^{m+1}$, it follows that $d(\psibf, \psibf^*)$ is a multiple of $N^{-2\ell}$.
+When $2^k > C_\epsilon N^{1+2\ell}$ we have $0 \leq d(\psibf, \psibf^*) < 2^{-k} C_\epsilon N < N^{-2\ell}$, and so $d(\psibf, \psibf^*) = 0$, and consequently $f_{nN}(\psibf) = f_{nN}(\psibf^*)$  for every $n = 1, \dots, N$ and $Z_N(\psibf) = Z_N(\psibf^*)$.
+Thus,
 \[
@@
 So, to analyse the supremum of $Z_N(\psibf)$ over the continuous interval $B_\epsilon$ it is enough to analyse the supremum over the discrete set $T_\epsilon(k)$ for large $k$.
-Lemma~\ref{lem:chaining3} shows that
-\[
-\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) } \leq \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{\sqrt{ i A_1 + A_2}}{2^{i/2}} < \infty
-\]
-for every positive integer $k$, where $A_1 = 18(m+1)\log 2$ and $A_2 = 18\log K_3$ are constants and $\log(\cdot)$ is the natural logarithm.
+The chaining argument~\cite[Sec.~3]{Pollard_asymp_empi_proc_1989} can  be used to show that
+\[
+\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) } \leq \expect_{\Phi} \sup_{\psibf \in T_\epsilon(k-1)} \sabs{ Z_N(\psibf) }  + \sqrt{C_\epsilon}\frac{6 \sqrt{\log\sabs{T_\epsilon(k)}}}{2^{k/2}},
+\]
+where $\log(\cdot)$ is the natural logarithm.
+Since $T_\epsilon(0)$ contains only the origin, $\expect_{\Phi}\sup_{\psibf \in T_\epsilon(0)} \abs{ Z_N(\psibf) } = \expect_{\Phi}\abs{ Z_N(0) } = 0$ and, %by unvravelling the recursion above, we obtain
+\[
+\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) } < \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{6\sqrt{\log\abs{T_\epsilon(i)}}}{2^{i/2}}.
+\]
+Now $\log\abs{T_\epsilon(i)} \leq \log(K_3) + i(m+1)\log(2)$ by Lemma~\ref{lem:metricentropy}, and so the sum above converges as $k\rightarrow\infty$.
+The lemma holds with $K_1 =  \sum_{i=1}^{\infty}\frac{6\sqrt{\log\abs{T_\epsilon(i)}}}{2^{i/2}}$.
+% as required.
+
+% Lemma~\ref{lem:chaining3} shows that 
+% \[
+% \expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) } \leq \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{\sqrt{ i A_1 + A_2}}{2^{i/2}} < \infty
+% \]
+% for every positive integer $k$, where $A_1 = 18(m+1)\log 2$ and $A_2 = 18\log K_3$ are constants and $\log(\cdot)$ is the natural logarithm.
 The lemma holds with $K_1 =  \sum_{i=1}^{\infty}2^{-i/2}\sqrt{ i A_1 + A_2 }$.
@@
 
-\begin{lemma}\label{lem:chaining3}(Chaining)
-For all positive integers $k$,
-\[
-\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) } \leq \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{\sqrt{ i A_1 + A_2 }}{2^{i/2}},
-\]
-where the constants $A_1 = 18(m+1)\log 2$ and $A_2 = 18\log K_3$.
-\end{lemma}
-\begin{proof}
-Let $b_k$ be a function that maps each $\psibf \in T_\epsilon(k)$ to $b_k(\psibf) \in T_\epsilon(k-1)$ such that  $d(\psibf,b_k(\psibf)) \leq 2^{1-k} C_\epsilon N$.
+% \begin{lemma}\label{lem:chaining3}(Chaining)
+% For all positive integers $k$,
+% \[
+% \expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) } \leq \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{\sqrt{ i A_1 + A_2 }}{2^{i/2}},
+% \]
+% where the constants $A_1 = 18(m+1)\log 2$ and $A_2 = 18\log K_3$.
+% \end{lemma}
+% \begin{proof}
+% Let $b_k$ be a function that maps each $\psibf \in T_\epsilon(k)$ to $b_k(\psibf) \in T_\epsilon(k-1)$ such that  $d(\psibf,b_k(\psibf)) \leq 2^{1-k} C_\epsilon N$.
 The existence of the function $b_k$ is guaranteed by the definition of $T_\epsilon(k)$.
 By the triangle inequality,
-\[
-\abs{ Z_N(\psibf) } \leq \abs{ Z_N(b_k(\psibf)) } + \abs{ Z_N(\psibf) - Z_N(b_k(\psibf))  },
-\]
-and by taking supremums on both sides,
-\begin{align}
-\sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) } &\leq \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(b_k(\psibf)) } + \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) - Z_N(b_k(\psibf))  } \nonumber \\
-&\leq \sup_{\psibf \in T_\epsilon(k-1)} \sabs{ Z_N(\psibf) } + \sup_{\psibf \in T_\epsilon(k)} \sabs{ Z_N(\psibf) - Z_N(b_k(\psibf)) }, \label{eq:supZchain}
-\end{align}
-the last line following since $b_k(\psibf) \in T_\epsilon(k-1)$ and so 
-\[
-\sup_{\psibf \in T_\epsilon(k)} \abs{ Z_N(b_k(\psibf)) } \leq \sup_{\psibf \in T_\epsilon(k-1)} \abs{ Z_N(\psibf) }.
-\]
-Conditional on $\Phi_1, \Phi_2, \dots$ the random variable $X(\psibf) = Z_N(\psibf) - Z_N(b_k(\psibf))$ has zero mean, and is normally distributed with variance
-\begin{align*}
-\sigma_X^2 &= \expect_{\Phi} \frac{1}{N} \sum_{n=1}^N g_n^2 \big( f_{nN}(\psibf) - f_{nN}(b_k(\psibf)) \big)^2 \\
-&= \frac{1}{N} \sum_{n=1}^N \big( f_{nN}(\psibf) - f_{nN}(b_k(\psibf)) \big)^2 = d(\psibf, b_k(\psibf)) \leq 2^{1-k} C_\epsilon,
-\end{align*}
-because $\expect_{\Phi} g_n^2 = 1$.  Using Lemma~\ref{lem:maxineq}, 
-\begin{align*}
-\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k)} \sabs{X(\psibf)} \leq 3\sqrt{2^{1-k} C_\epsilon \log\sabs{T_\epsilon(k)}} \leq \sqrt{C_\epsilon}\frac{\sqrt{  k A_1 + A_2 }}{2^{k/2}}
-\end{align*}
-because $\log\abs{T_\epsilon(k)} \leq k(m+1)\log 2 +  \log K_3$.  Taking expectations on both sides of~\eqref{eq:supZchain} gives
-\begin{align*}
-\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) } \leq \expect_{\Phi} \sup_{\psibf \in T_\epsilon(k-1)} \sabs{ Z_N(\psibf) }  + \sqrt{C_\epsilon}\frac{\sqrt{  k A_1 + A_2 }}{2^{k/2}},
-\end{align*}
-which involves a recursion in $k$.  By unravelling the recursion, and using the fact $T_\epsilon(0)$ contains only the origin, and therefore 
-\[
-\expect_{\Phi}\sup_{\psibf \in T_\epsilon(0)} \abs{ Z_N(\psibf) } = \expect_{\Phi}\abs{ Z_N(0) } = 0,
-\]
-we obtain $\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) } < \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{\sqrt{  k A_1 + A_2 }}{2^{k/2}}$
-as required.
-\end{proof}
-
-\begin{lemma}\label{lem:maxineq}(Maximal inequalities)
-Suppose $X_1, \dots, X_N$ are zero mean Gaussian random variables each with variance less than some positive constant $K$, then $\expect \sup_{n = 1, \dots, N} \abs{X_n} \leq 3  \sqrt{K \log N}$ where $\log N$ is the natural logarithm of $N$.
-\end{lemma}
-\begin{proof}
-This result is well known, see for example~\cite[Section 3]{Pollard_asymp_empi_proc_1989}  
-\end{proof}
-
-\begin{lemma}\label{lem:metricentropy}(Covering numbers)
+% \[
+% \abs{ Z_N(\psibf) } \leq \abs{ Z_N(b_k(\psibf)) } + \abs{ Z_N(\psibf) - Z_N(b_k(\psibf))  },
+% \]
+% and by taking supremums on both sides,
+% \begin{align}
+% \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) } &\leq \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(b_k(\psibf)) } + \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) - Z_N(b_k(\psibf))  } \nonumber \\
+% &\leq \sup_{\psibf \in T_\epsilon(k-1)} \sabs{ Z_N(\psibf) } + \sup_{\psibf \in T_\epsilon(k)} \sabs{ Z_N(\psibf) - Z_N(b_k(\psibf)) }, \label{eq:supZchain}
+% \end{align}
+% the last line following since $b_k(\psibf) \in T_\epsilon(k-1)$ and so 
+% \[
+% \sup_{\psibf \in T_\epsilon(k)} \abs{ Z_N(b_k(\psibf)) } \leq \sup_{\psibf \in T_\epsilon(k-1)} \abs{ Z_N(\psibf) }.
+% \]
+% Conditional on $\Phi_1, \Phi_2, \dots$ the random variable $X(\psibf) = Z_N(\psibf) - Z_N(b_k(\psibf))$ has zero mean, and is normally distributed with variance
+% \begin{align*}
+% \sigma_X^2 &= \expect_{\Phi} \frac{1}{N} \sum_{n=1}^N g_n^2 \big( f_{nN}(\psibf) - f_{nN}(b_k(\psibf)) \big)^2 \\
+% &= \frac{1}{N} \sum_{n=1}^N \big( f_{nN}(\psibf) - f_{nN}(b_k(\psibf)) \big)^2 = d(\psibf, b_k(\psibf)) \leq 2^{1-k} C_\epsilon,
+% \end{align*}
+% because $\expect_{\Phi} g_n^2 = 1$.  Using Lemma~\ref{lem:maxineq}, 
+% \begin{align*}
+% \expect_{\Phi} \sup_{\psibf \in T_\epsilon(k)} \sabs{X(\psibf)} \leq 3\sqrt{2^{1-k} C_\epsilon \log\sabs{T_\epsilon(k)}} \leq \sqrt{C_\epsilon}\frac{\sqrt{  k A_1 + A_2 }}{2^{k/2}}
+% \end{align*}
+% because $\log\abs{T_\epsilon(k)} \leq k(m+1)\log 2 +  \log K_3$.  Taking expectations on both sides of~\eqref{eq:supZchain} gives
+% \begin{align*}
+% \expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) } \leq \expect_{\Phi} \sup_{\psibf \in T_\epsilon(k-1)} \sabs{ Z_N(\psibf) }  + \sqrt{C_\epsilon}\frac{\sqrt{  k A_1 + A_2 }}{2^{k/2}},
+% \end{align*}
+% which involves a recursion in $k$.  By unravelling the recursion, and using the fact $T_\epsilon(0)$ contains only the origin, and therefore 
+% \[
+% \expect_{\Phi}\sup_{\psibf \in T_\epsilon(0)} \abs{ Z_N(\psibf) } = \expect_{\Phi}\abs{ Z_N(0) } = 0,
+% \]
+% we obtain $\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) } < \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{\sqrt{  k A_1 + A_2 }}{2^{k/2}}$
+% as required.
+% \end{proof}
+
+% \begin{lemma}\label{lem:maxineq}(Maximal inequalities)
+% Suppose $X_1, \dots, X_N$ are zero mean Gaussian random variables each with variance less than some positive constant $K$, then $\expect \sup_{n = 1, \dots, N} \abs{X_n} \leq 3  \sqrt{K \log N}$ where $\log N$ is the natural logarithm of $N$.
+% \end{lemma}
+% \begin{proof}
+% This result is well known, see for example~\cite[Section 3]{Pollard_asymp_empi_proc_1989}  
+% \end{proof}
+
+\begin{lemma}\label{lem:metricentropy} %(Covering numbers)
 For $k \in \ints$ there exists a discrete set $T_\epsilon(k) \subset \reals^{m+1}$ with the property that, for every $\psibf \in B_\epsilon$, there is a $\psibf^* \in T_\epsilon(k)$ such that,

### 2012-10-22 paper.tex
edits to paper. Removed the symmetrisation proof, which was very close to Pollards

@@
+%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGIN AOAP START %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 % Template for the submission to:
@@
 % AOS,AOAS: If there are supplements please fill:
-%\begin{supplement}[id=suppA]
+% \begin{supplement}[id=suppA]
 %  \sname{Supplement A}
@@
 %  \sdescription{Some text}
-%\end{supplement}
+% \end{supplement}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%% END AOAP START %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGIN IEEE START %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% %\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
+% \documentclass[journal]{IEEEtran}
+
+% \usepackage{mathbf-abbrevs}
+
+% %\newcommand{\dealias}{\operatorname{dealias}}
+
+% \input{defs}
+
+% %\newcommand{\term}{\emph}
+% %\renewcommand{\index}[1]{}
+
+% %\title{Polynomial phase estimation by phase unwrapping 1: Identifiability and strong consistency}
+% \title{Polynomial phase estimation by phase unwrapping}
+
+% \author{Robby~G.~McKilliam, Barry~G.~Quinn, I.~Vaughan~L.~Clarkson, Bill~Moran and Badri~N.~Vellambi%
+%     \thanks{%A preliminary version of some of this material is contained in Part 3 of Robby McKilliam's PhD thesis \cite{McKilliam2010thesis}. 
+% Robby~McKilliam and Badri Vellambi are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.
+Barry~Quinn is with the Department of Statistics, Macquarie University, Sydney, NSW, 2109, Australia.
+Vaughan~Clarkson is with the School of Information Technology \& Electrical Engineering, The University of Queensland, QLD., 4072, Australia.
+B.
+Moran is with the Department of Electrical Engineering and Computer
+% Science, Melbourne Systems Lab, Dept of Elec \& Electronic Eng, Uni of Melbourne, Vic. 3010, Australia.}}
+% % The paper headers 
+% \markboth{Polynomial phase estimation by phase unwrapping}{DRAFT \today}
+
+% % make the title area
+
+% \begin{document} 
+% \maketitle
+
+% \begin{abstract}
+% Estimating the coefficients of a noisy polynomial phase signal is important in many fields including radar, biology and radio communications.
+One approach attempts to perform polynomial regression on the phase of the signal.
+This is complicated by the fact that the phase is \emph{wrapped} modulo $2\pi$ and must be \emph{unwrapped} before regression can be performed.
+%A recent approach suggested by the authors is to perform the unwrapping in a least squares manner.
+%It was shown how the resulting estimator could be posed as a \emph{nearest lattice point problem} in a specific lattice and could be solved using existing algorithms.
+% %It was shown by Monte Carlo simulation that this produces a remarkably accurate estimator.  
+% In this paper we consider an estimator that performs phase unwrapping in a least squares manner.
+We describe the asymptotic properties of this estimator, showing that it is strongly consistent and asymptotically normally distributed.
+%We hypothesise that the estimator produces very near maximum likelihood performance.
+% \end{abstract}
+
+% \begin{keywords}
+% Polynomial phase signals, phase unwrapping, asymptotic properties, nearest lattice point problem
+% \end{keywords}
+
+% %\bibliographystyle{unsrt}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%% END IEEE START %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 When $m=0$ (phase estimation) or $m=1$ (frequency estimation) the least squares estimator is an effective approach, being both computationally efficient and statistically accurate~\cite{Hannan1973,Quinn2001,McKilliam_mean_dir_est_sq_arc_length2010}.
-When $m \geq 2$ the computational complexity of the least squares estimator is large, and for this reason authors have considered alternative estimators.
+When $m \geq 2$ the computational complexity of the least squares estimator is large, and alternative estimators have been considered for this reason.
 These can loosely be grouped into two classes; estimators based on polynomial phase transforms, such as the discrete polynomial phase transform~\cite{Peleg_DPT_1995} and the high order phase function~\cite{Farquharson_another_poly_est_2005,Porat_asympt_HAF_DPT_1996}; and estimators based on phase unwrapping, such as Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, and Morelande's Bayesian unwrapping estimator~\cite{Morelande_bayes_unwrapping_2009_tsp}.
@@
 
-An interesting property is that the estimator of the $k$th polynomial phase coefficient converges almost surely to $\tilde{\mu}_k$ at rate $N^{-k}$.
+An interesting property is that the estimator of the $k$th polynomial phase coefficient converges almost surely to $\tilde{\mu}_k$ at rate $o(N^{-k})$.
 This is perhaps not surprising, since it is the same rate observed in polynomial regression.
 However, asserting that convergence at this rate occurs in the polynomial phase setting is not trivial.
-For this purpose we make use of an elementary result about the number of arithmetic progressions contained inside subsets of $\{1,2,\dots,N\}$.
+For this purpose we make use of an elementary result about the number of arithmetic progressions contained inside subsets of $\{1,2,\dots,N\}$~\cite{Erdos_on_some_sequence_of_integers1936,Szemeredi_setint_no_k_arth1975,Gowers_new_proof2001}.
 %This proof technique appears to be novel, and may be useful in other problems.
@@
 One choice for a set of coset representatives is a fundamental parallelepiped; the parallelepiped generated by the columns of a generator matrix.
-Another choice is based on the Voronoi cell, i.e., those points from $\reals^n$ nearest (with respect to the Euclidean norm here) to the lattice point at the origin.
+Another choice is based on the Voronoi cell those points from $\reals^n$ nearest (with respect to the Euclidean norm here) to the lattice point at the origin.
 It is always possible to construct a rectangular set of representatives, as the next proposition will show.
@@
 Observe that $\Pbf$ is upper triangular with $k$th diagonal element equal to $\tfrac{1}{k!}$.
-So this rectangular region is
+So this rectangular set is
 \begin{equation}\label{eq:rectangular_identifiable_region}
@@
 \end{equation}
-We will make use of this region when deriving the statistical properties of the LSU estimator in the next section.
+We will make use of this set when deriving the statistical properties of the LSU estimator in the next section.
 
@@
 However, the next two lemmas and the following corollary will show that if $K$ is sufficiently large then it always contains at least one arithmetic progression (for all sufficiently small $h$) and therefore $A(h,K)$ is not empty.
+We do not wish to claim any novelty here, the study of arithmetic progressions within subsets of $W$ has a considerable history~\cite{Erdos_on_some_sequence_of_integers1936,Szemeredi_setint_no_k_arth1975,Gowers_new_proof2001}.
+In particular, Gower's~\cite[Theorem 1.3]{Gowers_new_proof2001} gives a result far stronger than we require here.
 Denote by $K \backslash \{r\}$ the set $K$ with the element $r$ removed.
@@
 We prove this result here.
-In what follows we use order notation in the standard way, that is, for functions $h$ and $g$, we write $h(N) = O(g(N))$ to mean that there exists a constant $K > 0$ and a finite $N_0$ such that $h(N) \leq K g(N)$ for all $N > N_0$.
 Put $D_N(\lambdabf) = S_N(\lambdabf) - V_N(\lambdabf)$.
 Now $\sum_{N=1}^\infty \prob \left\{ \sup_{\lambdabf \in B}\abs{ D_N(\lambdabf) } > \epsilon \right\} < \infty$ for any $\epsilon > 0$ by Lemma~\ref{lem:vn}, and~\eqref{eq:SNVNunifmlln2} follows from the Borel-Cantelli lemma.
+In what follows we use order notation in the standard way, that is, for functions $h$ and $g$, we write $h(N) = O(g(N))$ to mean that there exists a constant $K > 0$ and a finite $N_0$ such that $h(N) \leq K g(N)$ for all $N > N_0$.
 
@@
  \end{align}
-From Lemma~\ref{lem:supVjk} it will follow that
-\[
-\prob\left\{   \sup_{\rbf \in G }\abs{ D_N(\lambdabf[\rbf])  } > \frac{\epsilon}{2} \right\} = O(e^{-c \epsilon^2 N})
-\]
+From Lemma~\ref{lem:supVjk} it will follow that $\prob\left\{   \sup_{\rbf \in G }\abs{ D_N(\lambdabf[\rbf])  } > \frac{\epsilon}{2} \right\} = O(e^{-c \epsilon^2 N})$
 for any $\epsilon > 0$ and $c < 2$.  In Lemma~\ref{lem:supBVn} we show that
@@
  \]
- are independent with zero mean and $\abs{Z_n} \leq \tfrac{1}{4}$. It follows from Hoeffding's inequality~\cite{Hoeffding_inequality_1963} that,%From Lemma~\ref{lem:zero_mean_indepent_sum_bound},
- \[
-  \prob\{ \abs{D_N(\lambdabf)} > \epsilon \} \leq 2e^{ - 8 \epsilon^2 N},
- \]
-and so,
+ are independent with zero mean and $\abs{Z_n} \leq \tfrac{1}{4}$.
+It follows from Hoeffding's inequality~\cite{Hoeffding_inequality_1963} that,  $\prob\{ \abs{D_N(\lambdabf)} > \epsilon \} \leq 2e^{ - 8 \epsilon^2 N}$, and so,
  \begin{align*}
@@
 \begin{lemma}\label{lem:boundedsquarefracparts}
-$\fracpart{x}^2 - |\delta| \leq \fracpart{x + \delta}^2 \leq \fracpart{x}^2 + |\delta|$ for all real numbers $x$ and $\delta$.
+$\fracpart{x}^2 - |\delta| \leq \fracpart{x + \delta}^2 \leq \fracpart{x}^2 + |\delta|$ for all $x, \delta \in \reals$.
 \end{lemma}
@@
 %The proof we give is not just almost surely but \emph{surely} as the convergence will be shown to occur irrespective of the values of the $\Phi_n$. 
-For convenience put 
-\[
-b_n = \Phi_n + \sum_{k = 0}^{m}{\lambda_k n^k} \qquad \text{and} \qquad  a_n = \Phi_n + \sum_{k = 0}^{m}{\lambda[\rbf]_k n^k},
-\] 
-where $\lambda[\rbf]_k$ denotes the $k$th element of the grid point $\lambdabf[\rbf]$. For $\lambdabf \in B[\rbf]$ we have $b_n = a_n + \delta_n$,
-where 
-\[
-\sabs{\delta_n} \leq \sum_{k=0}^m\frac{n^k}{k! N^{b+k}} \leq \frac{m+1}{N^{b}}.
-\] 
+Put $b_n = \Phi_n + \sum_{k = 0}^{m}{\lambda_k n^k}$ and $a_n = \Phi_n + \sum_{k = 0}^{m}{\lambda[\rbf]_k n^k}$, where $\lambda[\rbf]_k$ denotes the $k$th element of the grid point $\lambdabf[\rbf]$.
+For $\lambdabf \in B[\rbf]$ we have $b_n = a_n + \delta_n$,
+where $\sabs{\delta_n} \leq \sum_{k=0}^m\frac{n^k}{k!
+N^{b+k}} \leq \frac{m+1}{N^{b}}$.
 From Lemma~\ref{lem:boundedsquarefracparts} it follows that $-|\delta_n| \leq \fracpart{x + b_n}^2 -  \fracpart{x + a_n}^2  \leq |\delta_n|$,
-and consequently $|\fracpart{x + b_n}^2 -  \fracpart{x + a_n}^2| \leq \tfrac{m+1}{N^b}$ for all real numbers $x$. Now
+and consequently $|\fracpart{x + b_n}^2 -  \fracpart{x + a_n}^2| \leq \tfrac{m+1}{N^b}$ for all $x \in \reals$. Now
 \[
@@
 
-%The proofs of Lemmas~\ref{lem:chaining2}~and~\ref{lem:chaining3} are based on a technique called \emph{chaining} (or \emph{bracketing})~\cite{Dudley_unif_central_lim_th_1999,Ossiander_clt_bracketing_1984,Pollard_asymp_empi_proc_1989,Pollard_new_ways_clts_1986}.
-The proofs here follow those of Pollard~\cite[Section 3]{Pollard_asymp_empi_proc_1989}.
-The remaining lemmas consider a fixed realisation $\{\Phi_n\}$, and we use the abbreviations $C_\epsilon = C_\epsilon(\{\Phi_n\})$ and $f_{nN}(\psibf) = f_{nN}(\psibf, \Phi_n)$.
+The proofs of Lemmas~\ref{lem:chaining2}~and~\ref{lem:chaining3} are based on a technique called \emph{chaining} (or \emph{bracketing})~\cite{Dudley_unif_central_lim_th_1999,Ossiander_clt_bracketing_1984,Pollard_asymp_empi_proc_1989,Pollard_new_ways_clts_1986}.
+The proofs here follow those of Pollard~\cite{Pollard_asymp_empi_proc_1989}.
+In the remaining lemmas we consider expectation conditional on $\{\Phi_n\}$ and treat $\{\Phi_n\}$ as a fixed realisation.
+We consequently use the abbreviations $C_\epsilon = C_\epsilon(\{\Phi_n\})$ and $f_{nN}(\psibf) = f_{nN}(\psibf, \Phi_n)$.
 As in Lemma~\ref{lem:chaining} we assume, without loss of generality, that $\epsilon < \tfrac{1}{m+1}$.
@@
 \end{lemma}
-\begin{proof}
-Let $B_\epsilon = \{\xbf \in \reals \mid \|\xbf\|_\infty < \epsilon \}$.
+\begin{proof}  Let $B_\epsilon = \{\xbf \in \reals \mid \|\xbf\|_\infty < \epsilon \}$.
 %We are now required to prove that
@@
 We define $T_\epsilon(0)$ to contain a single point, the origin $\zerobf$.
-Defined this way, $T_\epsilon(0)$ satisfies the inequality above because $d(\psibf, \zerobf) = \sum_{n = 1}^N f_{nN}(\psibf)^2  \leq C_\epsilon N$ for all $\psibf \in B_\epsilon$, as a result of Lemma~\ref{lem:epslmlemma}.
+Defined this way $T_\epsilon(0)$ satisfies the inequality above because $d(\psibf, \zerobf) = \sum_{n = 1}^N f_{nN}(\psibf)^2  \leq C_\epsilon N$ for all $\psibf \in B_\epsilon$, as a result of Lemma~\ref{lem:epslmlemma}.
+
 The existence of $T_\epsilon(k)$ for each positive integer $k$ will be proved in Lemma~\ref{lem:metricentropy}.
-%It is worth giving some intuition regarding $T_\epsilon(k)$.
+It is worth giving some intuition regarding $T_\epsilon(k)$.
 If we place a `ball' of radius $2^{-k} C_\epsilon N$  with respect to the pseudometric $d(\cdot, \cdot)$ around each point in $T_\epsilon(k)$, then, by definition, the union of these balls is a superset of $B_\epsilon$.
@@
 The minimum number of such balls required to cover $B_\epsilon$ is called a \emph{covering number} of $B_\epsilon$.
-%In Lemma~\ref{lem:metricentropy} we show that no more than $K_3 2^{(m+1)k}$ balls of radius $2^{-k} C_\epsilon N$ are required to cover $B_\epsilon$, where $K_3$ is a constant, independent of $N$ and $\epsilon$.
-We make use of this upper bound in Lemma~\ref{lem:chaining3}.
-The lemma also upper bounds the number of elements by $\abs{T_\epsilon(k)} < K_3 2^{(m+1)k}$ where $K_3$ is a constant independent of $N$ and $\epsilon$.
+In Lemma~\ref{lem:metricentropy} we show that no more than $K_3 2^{(m+1)k}$ balls of radius $2^{-k} C_\epsilon N$ are required to cover $B_\epsilon$, that is $\abs{T_\epsilon(k)} \leq K_3 2^{(m+1)k}$, where $K_3$ is a constant, independent of $N$ and $\epsilon$.
+%We make use of this upper bound now.
+%The lemma also upper bounds the number of elements by $\abs{T_\epsilon(k)} < K_3 2^{(m+1)k}$ where $K_3$ is a constant independent of $N$ and $\epsilon$.
 
@@
 So, to analyse the supremum of $Z_N(\psibf)$ over the continuous interval $B_\epsilon$ it is enough to analyse the supremum over the discrete set $T_\epsilon(k)$ for large $k$.
-The chaining argument~\cite[Sec.~3]{Pollard_asymp_empi_proc_1989} can  be used to show that
-\[
-\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) } \leq \expect_{\Phi} \sup_{\psibf \in T_\epsilon(k-1)} \sabs{ Z_N(\psibf) }  + \sqrt{C_\epsilon}\frac{6 \sqrt{\log\sabs{T_\epsilon(k)}}}{2^{k/2}},
-\]
-where $\log(\cdot)$ is the natural logarithm.
-Since $T_\epsilon(0)$ contains only the origin, $\expect_{\Phi}\sup_{\psibf \in T_\epsilon(0)} \abs{ Z_N(\psibf) } = \expect_{\Phi}\abs{ Z_N(0) } = 0$ and, %by unvravelling the recursion above, we obtain
-\[
-\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) } < \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{6\sqrt{\log\abs{T_\epsilon(i)}}}{2^{i/2}}.
-\]
-Now $\log\abs{T_\epsilon(i)} \leq \log(K_3) + i(m+1)\log(2)$ by Lemma~\ref{lem:metricentropy}, and so the sum above converges as $k\rightarrow\infty$.
+% The chaining argument~\cite[Sec.~3]{Pollard_asymp_empi_proc_1989} can  be used to show that
+% \[
+% \expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) } < \expect_{\Phi} \sup_{\psibf \in T_\epsilon(k-1)} \sabs{ Z_N(\psibf) }  + \sqrt{C_\epsilon}\frac{6 \sqrt{\log\sabs{T_\epsilon(k)}}}{2^{k/2}},
+% \]
+% which involves a recursion in $k$, where $\log(\cdot)$ is the natural logarithm.
+Since $T_\epsilon(0)$ contains only the origin, $\expect_{\Phi}\sup_{\psibf \in T_\epsilon(0)} \abs{ Z_N(\psibf) } = \expect_{\Phi}\abs{ Z_N(0) } = 0$ and, by unravelling the recursion, we obtain
+% \[
+% \expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) } < \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{6\sqrt{\log\abs{T_\epsilon(i)}}}{2^{i/2}}.
+% \]
+% Now $\log\abs{T_\epsilon(i)} \leq \log(K_3) + i(m+1)\log(2)$ by Lemma~\ref{lem:metricentropy}, and so the sum above converges as $k\rightarrow\infty$.
 The lemma holds with $K_1 =  \sum_{i=1}^{\infty}\frac{6\sqrt{\log\abs{T_\epsilon(i)}}}{2^{i/2}}$.
 % as required.
-
-% Lemma~\ref{lem:chaining3} shows that 
-% \[
-% \expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) } \leq \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{\sqrt{ i A_1 + A_2}}{2^{i/2}} < \infty
-% \]
-% for every positive integer $k$, where $A_1 = 18(m+1)\log 2$ and $A_2 = 18\log K_3$ are constants and $\log(\cdot)$ is the natural logarithm.
+Lemma~\ref{lem:chaining3} shows that 
+\[
+\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) } \leq \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{\sqrt{ i A_1 + A_2}}{2^{i/2}} < \infty
+\]
+for every positive integer $k$, where $A_1 = 18(m+1)\log 2$ and $A_2 = 18\log K_3$ are constants and $\log(\cdot)$ is the natural logarithm.
 The lemma holds with $K_1 =  \sum_{i=1}^{\infty}2^{-i/2}\sqrt{ i A_1 + A_2 }$.
@@
 
-% \begin{lemma}\label{lem:chaining3}(Chaining)
-% For all positive integers $k$,
-% \[
-% \expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) } \leq \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{\sqrt{ i A_1 + A_2 }}{2^{i/2}},
-% \]
-% where the constants $A_1 = 18(m+1)\log 2$ and $A_2 = 18\log K_3$.
-% \end{lemma}
-% \begin{proof}
-% Let $b_k$ be a function that maps each $\psibf \in T_\epsilon(k)$ to $b_k(\psibf) \in T_\epsilon(k-1)$ such that  $d(\psibf,b_k(\psibf)) \leq 2^{1-k} C_\epsilon N$.
+\begin{lemma}\label{lem:chaining3}(Chaining)
+For all positive integers $k$,
+\[
+\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) } \leq \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{\sqrt{ i A_1 + A_2 }}{2^{i/2}},
+\]
+where the constants $A_1 = 18(m+1)\log 2$ and $A_2 = 18\log K_3$.
+\end{lemma}
+\begin{proof}
+Let $b_k$ be a function that maps each $\psibf \in T_\epsilon(k)$ to $b_k(\psibf) \in T_\epsilon(k-1)$ such that  $d(\psibf,b_k(\psibf)) \leq 2^{1-k} C_\epsilon N$.
 The existence of the function $b_k$ is guaranteed by the definition of $T_\epsilon(k)$.
 By the triangle inequality,
-% \[
-% \abs{ Z_N(\psibf) } \leq \abs{ Z_N(b_k(\psibf)) } + \abs{ Z_N(\psibf) - Z_N(b_k(\psibf))  },
-% \]
-% and by taking supremums on both sides,
-% \begin{align}
-% \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) } &\leq \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(b_k(\psibf)) } + \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) - Z_N(b_k(\psibf))  } \nonumber \\
-% &\leq \sup_{\psibf \in T_\epsilon(k-1)} \sabs{ Z_N(\psibf) } + \sup_{\psibf \in T_\epsilon(k)} \sabs{ Z_N(\psibf) - Z_N(b_k(\psibf)) }, \label{eq:supZchain}
-% \end{align}
-% the last line following since $b_k(\psibf) \in T_\epsilon(k-1)$ and so 
-% \[
-% \sup_{\psibf \in T_\epsilon(k)} \abs{ Z_N(b_k(\psibf)) } \leq \sup_{\psibf \in T_\epsilon(k-1)} \abs{ Z_N(\psibf) }.
-% \]
-% Conditional on $\Phi_1, \Phi_2, \dots$ the random variable $X(\psibf) = Z_N(\psibf) - Z_N(b_k(\psibf))$ has zero mean, and is normally distributed with variance
-% \begin{align*}
-% \sigma_X^2 &= \expect_{\Phi} \frac{1}{N} \sum_{n=1}^N g_n^2 \big( f_{nN}(\psibf) - f_{nN}(b_k(\psibf)) \big)^2 \\
-% &= \frac{1}{N} \sum_{n=1}^N \big( f_{nN}(\psibf) - f_{nN}(b_k(\psibf)) \big)^2 = d(\psibf, b_k(\psibf)) \leq 2^{1-k} C_\epsilon,
-% \end{align*}
-% because $\expect_{\Phi} g_n^2 = 1$.  Using Lemma~\ref{lem:maxineq}, 
-% \begin{align*}
-% \expect_{\Phi} \sup_{\psibf \in T_\epsilon(k)} \sabs{X(\psibf)} \leq 3\sqrt{2^{1-k} C_\epsilon \log\sabs{T_\epsilon(k)}} \leq \sqrt{C_\epsilon}\frac{\sqrt{  k A_1 + A_2 }}{2^{k/2}}
-% \end{align*}
-% because $\log\abs{T_\epsilon(k)} \leq k(m+1)\log 2 +  \log K_3$.  Taking expectations on both sides of~\eqref{eq:supZchain} gives
-% \begin{align*}
-% \expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) } \leq \expect_{\Phi} \sup_{\psibf \in T_\epsilon(k-1)} \sabs{ Z_N(\psibf) }  + \sqrt{C_\epsilon}\frac{\sqrt{  k A_1 + A_2 }}{2^{k/2}},
-% \end{align*}
-% which involves a recursion in $k$.  By unravelling the recursion, and using the fact $T_\epsilon(0)$ contains only the origin, and therefore 
-% \[
-% \expect_{\Phi}\sup_{\psibf \in T_\epsilon(0)} \abs{ Z_N(\psibf) } = \expect_{\Phi}\abs{ Z_N(0) } = 0,
-% \]
-% we obtain $\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) } < \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{\sqrt{  k A_1 + A_2 }}{2^{k/2}}$
-% as required.
-% \end{proof}
-
-% \begin{lemma}\label{lem:maxineq}(Maximal inequalities)
-% Suppose $X_1, \dots, X_N$ are zero mean Gaussian random variables each with variance less than some positive constant $K$, then $\expect \sup_{n = 1, \dots, N} \abs{X_n} \leq 3  \sqrt{K \log N}$ where $\log N$ is the natural logarithm of $N$.
-% \end{lemma}
-% \begin{proof}
-% This result is well known, see for example~\cite[Section 3]{Pollard_asymp_empi_proc_1989}  
-% \end{proof}
-
-\begin{lemma}\label{lem:metricentropy} %(Covering numbers)
+\[
+\abs{ Z_N(\psibf) } \leq \abs{ Z_N(b_k(\psibf)) } + \abs{ Z_N(\psibf) - Z_N(b_k(\psibf))  },
+\]
+and by taking supremums on both sides,
+\begin{align}
+\sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) } &\leq \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(b_k(\psibf)) } + \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) - Z_N(b_k(\psibf))  } \nonumber \\
+&\leq \sup_{\psibf \in T_\epsilon(k-1)} \sabs{ Z_N(\psibf) } + \sup_{\psibf \in T_\epsilon(k)} \sabs{ Z_N(\psibf) - Z_N(b_k(\psibf)) }, \label{eq:supZchain}
+\end{align}
+the last line following since $b_k(\psibf) \in T_\epsilon(k-1)$ and so 
+\[
+\sup_{\psibf \in T_\epsilon(k)} \abs{ Z_N(b_k(\psibf)) } \leq \sup_{\psibf \in T_\epsilon(k-1)} \abs{ Z_N(\psibf) }.
+\]
+Conditional on $\Phi_1, \Phi_2, \dots$ the random variable $X(\psibf) = Z_N(\psibf) - Z_N(b_k(\psibf))$ has zero mean, and is normally distributed with variance
+\begin{align*}
+\sigma_X^2 &= \expect_{\Phi} \frac{1}{N} \sum_{n=1}^N g_n^2 \big( f_{nN}(\psibf) - f_{nN}(b_k(\psibf)) \big)^2 \\
+&= \frac{1}{N} \sum_{n=1}^N \big( f_{nN}(\psibf) - f_{nN}(b_k(\psibf)) \big)^2 = d(\psibf, b_k(\psibf)) \leq 2^{1-k} C_\epsilon,
+\end{align*}
+because $\expect_{\Phi} g_n^2 = 1$.  Using Lemma~\ref{lem:maxineq}, 
+\begin{align*}
+\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k)} \sabs{X(\psibf)} \leq 3\sqrt{2^{1-k} C_\epsilon \log\sabs{T_\epsilon(k)}} \leq \sqrt{C_\epsilon}\frac{\sqrt{  k A_1 + A_2 }}{2^{k/2}}
+\end{align*}
+because $\log\abs{T_\epsilon(k)} \leq k(m+1)\log 2 +  \log K_3$.  Taking expectations on both sides of~\eqref{eq:supZchain} gives
+\begin{align*}
+\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) } \leq \expect_{\Phi} \sup_{\psibf \in T_\epsilon(k-1)} \sabs{ Z_N(\psibf) }  + \sqrt{C_\epsilon}\frac{\sqrt{  k A_1 + A_2 }}{2^{k/2}},
+\end{align*}
+which involves a recursion in $k$.  By unravelling the recursion, and using the fact $T_\epsilon(0)$ contains only the origin, and therefore 
+\[
+\expect_{\Phi}\sup_{\psibf \in T_\epsilon(0)} \abs{ Z_N(\psibf) } = \expect_{\Phi}\abs{ Z_N(0) } = 0,
+\]
+we obtain $\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) } < \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{\sqrt{  k A_1 + A_2 }}{2^{k/2}}$
+as required.
+\end{proof}
+
+\begin{lemma}\label{lem:maxineq}(Maximal inequality)
+Suppose $X_1, \dots, X_N$ are zero mean Gaussian random variables each with variance less than some positive constant $K$, then $\expect \sup_{n = 1, \dots, N} \abs{X_n} \leq 3  \sqrt{K \log N}$ where $\log N$ is the natural logarithm of $N$.
+\end{lemma}
+\begin{proof}
+This result is well known, see for example~\cite[Section 3]{Pollard_asymp_empi_proc_1989}  
+\end{proof}
+
+\begin{lemma}\label{lem:metricentropy} (Covering numbers)
 For $k \in \ints$ there exists a discrete set $T_\epsilon(k) \subset \reals^{m+1}$ with the property that, for every $\psibf \in B_\epsilon$, there is a $\psibf^* \in T_\epsilon(k)$ such that,

### 2012-10-25 paper.tex
minor edits to aop?

@@
 The proof of asymptotic normality is complicated by the fact that the objective function corresponding with the LSU estimator is not differentiable everywhere.
-Techniques developed by Huber~\cite{Huberrobustest1964,Huberbehavmaxnonstandard1967} and Pollard~\cite{Pollard_new_ways_clts_1986,Pollard_asymp_empi_proc_1989} and results from the literature on hyperplane arrangements~\cite{Chazelle_discrepency_method_2000,Matousek_lect_disc_geom_2002} become useful here.
+Empirical process techniques~\cite{Pollard_new_ways_clts_1986,Pollard_asymp_empi_proc_1989,van2009empirical,Dudley_unif_central_lim_th_1999} and results from the literature on hyperplane arrangements~\cite{Chazelle_discrepency_method_2000,Matousek_lect_disc_geom_2002} become useful here.
 We are hopeful that the proof techniques developed here will be useful for purposes other than polynomial phase estimation, and in particular other applications involving data that is `wrapped' in some sense.

### 2012-10-26 paper.tex
minor

--- 2012-10-22 paper.tex
@@
 
-% Template for the submission to:
-%   The Annals of Probability           [aop]
-%   The Annals of Applied Probability   [aap]
-%   The Annals of Statistics            [aos] 
-%   The Annals of Applied Statistics    [aoas]
-%   Stochastic Systems                  [ssy]
-%
-%Author: In this template, the places where you need to add information
-%        (or delete line) are indicated by {???}.  Mostly the information
-%        required is obvious, but some explanations are given in lines starting
-%Author:
-%All other lines should be ignored.  After editing, there should be
-%no instances of ??? after this line.
-
-% use option [preprint] to remove info line at bottom
-% journal options: aop,aap,aos,aoas,ssy
-% natbib option: authoryear
-\documentclass[aap]{imsart}
-
-\usepackage{amsthm,amsmath,amsfonts,amssymb}
-\usepackage[square,comma,numbers,sort&compress]{natbib}
-%\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
-\usepackage{units}
-\usepackage{ifpdf}
-\ifpdf
-  \usepackage[pdftex]{graphicx}
-\else
-  \usepackage{graphicx}% standard graphics package for inclusion images and eps files into LaTeX document
-\fi
-
-% provide arXiv number if available:
-%\arxiv{arXiv:0000.0000}
-
-% put your definitions there:
-\startlocaldefs
-\newcommand{\reals}{{\mathbb R}}
-\newcommand{\ints}{{\mathbb Z}}
-\newcommand{\term}{\emph}
-\newcommand{\var}{\operatorname{var}}
-\newcommand{\covar}{\operatorname{covar}}
-%\newcommand{\prob}{{\mathbb P}}
-\newcommand{\prob}{\operatorname{Pr}}
-\newcommand{\expect}{{\mathbb E}}
-\newcommand{\dealias}{\operatorname{dealias}}
-\renewcommand{\mid}{\; ; \;}
-\newcommand{\fracpart}[1]{\left\langle #1 \right\rangle}
-\newcommand{\sfracpart}[1]{\langle #1 \rangle}
-\newcommand{\abs}[1]{\left\vert #1 \right\vert}
-\newcommand{\sabs}[1]{\vert #1 \vert}
-\newcommand{\floor}[1]{{\left\lfloor #1 \right\rfloor}}
-\newcommand{\ceiling}[1]{{\left\lceil #1 \right\rceil}}
-\newcommand{\ceil}[1]{\lceil #1 \rceil}
-\newcommand{\round}[1]{{\left\lceil #1 \right\rfloor}}
+% % Template for the submission to:
+% %   The Annals of Probability           [aop]
+% %   The Annals of Applied Probability   [aap]
+% %   The Annals of Statistics            [aos] 
+% %   The Annals of Applied Statistics    [aoas]
+% %   Stochastic Systems                  [ssy]
+% %
+% %Author: In this template, the places where you need to add information
+% %        (or delete line) are indicated by {???}.  Mostly the information
+% %        required is obvious, but some explanations are given in lines starting
+% %Author:
+% %All other lines should be ignored.  After editing, there should be
+% %no instances of ??? after this line.
+
+% % use option [preprint] to remove info line at bottom
+% % journal options: aop,aap,aos,aoas,ssy
+% % natbib option: authoryear
+% \documentclass[aap]{imsart}
+
+% \usepackage{amsthm,amsmath,amsfonts,amssymb}
+% \usepackage[square,comma,numbers,sort&compress]{natbib}
+% %\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
+% \usepackage{units}
+% \usepackage{ifpdf}
+% \ifpdf
+%   \usepackage[pdftex]{graphicx}
+% \else
+%   \usepackage{graphicx}% standard graphics package for inclusion images and eps files into LaTeX document
+% \fi
+
+% % provide arXiv number if available:
+% %\arxiv{arXiv:0000.0000}
+
+% % put your definitions there:
+% \startlocaldefs
+% \newcommand{\reals}{{\mathbb R}}
+% \newcommand{\ints}{{\mathbb Z}}
+% \newcommand{\term}{\emph}
+% \newcommand{\var}{\operatorname{var}}
+% \newcommand{\covar}{\operatorname{covar}}
+% %\newcommand{\prob}{{\mathbb P}}
+% \newcommand{\prob}{\operatorname{Pr}}
+% \newcommand{\expect}{{\mathbb E}}
+% \newcommand{\dealias}{\operatorname{dealias}}
+% \renewcommand{\mid}{\; ; \;}
+% \newcommand{\fracpart}[1]{\left\langle #1 \right\rangle}
+% \newcommand{\sfracpart}[1]{\langle #1 \rangle}
+% \newcommand{\abs}[1]{\left\vert #1 \right\vert}
+% \newcommand{\sabs}[1]{\vert #1 \vert}
+% \newcommand{\floor}[1]{{\left\lfloor #1 \right\rfloor}}
+% \newcommand{\ceiling}[1]{{\left\lceil #1 \right\rceil}}
+% \newcommand{\ceil}[1]{\lceil #1 \rceil}
+% \newcommand{\round}[1]{{\left\lceil #1 \right\rfloor}}
+% \usepackage{mathbf-abbrevs}
+% \newtheorem{proposition}{Proposition}
+% \newtheorem{corollary}{Corollary}
+% \newtheorem{theorem}{Theorem}
+% \newtheorem{lemma}{Lemma}
+% \newtheorem{definition}{Definition}
+% \newcommand{\coef}{\operatorname{coef}}
+% \newcommand{\cubr}[1]{{\left\{ #1 \right\}}}
+% \newcommand{\scubr}[1]{{\{ #1 \}}}
+% \newcommand{\sign}[1]{\operatorname{sgn}\left( #1 \right)}
+% \newcommand{\bfracpart}[1]{\big\langle #1 \big\rangle}
+% \newcommand{\babs}[1]{\big\vert #1 \big\vert}
+% \endlocaldefs
+
+% \begin{document}
+
+% \begin{frontmatter}
+
+% % "Title of the paper"
+% \title{Polynomial phase estimation by phase unwrapping}
+% \runtitle{Polynomial phase estimation by phase unwrapping}
+
+% % indicate corresponding author with \corref{}
+% % \author{\fnms{John} \snm{Smith}\corref{}\ead[label=e1]{smith@foo.com}\thanksref{t1}}
+% % \thankstext{t1}{Thanks to somebody} 
+% % \address{line 1\\ line 2\\ printead{e1}}
+% % \affiliation{Some University}
+
+% \begin{aug}
+% \author{\fnms{Robby G.} \snm{McKilliam}\thanksref{m1}\ead[label=e1]{robby.mckilliam@unisa.edu.au}},
+% \author{\fnms{Barry G.} \snm{Quinn}\thanksref{m2}\ead[label=e2]{barry.quinn@mq.edu.au}},
+% \author{\fnms{I. Vaughan L.} \snm{Clarkson}\thanksref{m3}\ead[label=e3]{v.clarkson@uq.edu.au}},
+% \author{\fnms{Bill} \snm{Moran}\thanksref{m4}\ead[label=e4]{wmoran@unimelb.edu.au}},
+% \and
+% \author{\fnms{Badri N.} \snm{Vellambi}\thanksref{m1}\ead[label=e5]{badri.vellambi@unisa.edu.au}}
+% \runauthor{R. McKilliam et. al.}
+
+% \affiliation{University of South Australia\thanksmark{m1}, Macquarie University\thanksmark{m2}, University of Queensland\thanksmark{m3} and Melbourne University\thanksmark{m4}}
+
+% \address{R. G. McKilliam and B. N. Vellambi\\
+% Institute for Telecommunications Research\\
+% University of South Australia, Adelaide\\
+% SA, 5095, Australia\\
+% \printead{e1}\\
+% \phantom{E-mail:\ }\printead*{e5}}
+
+% \address{B. G. Quinn\\
+% Department of Statistics\\
+% Macquarie University, Sydney,\\
+% NSW, 2109, Australia\\
+% \printead{e3}}
+
+% \address{I. V. L. Clarkson\\
+% School of Information Technology \& Electrical\\
+% Engineering, University of Queensland,\\
+% Brisbane, QLD, 4072, Australia\\
+% \printead{e3}}
+
+% \address{B. Moran\\
+% Department of Electrical \& Electronic\\
+% Engineering, Melbourne University,\\
+% Melbourne, VIC, 3010, Australia\\
+% \printead{e4}}
+
+% \end{aug}
+
+% \begin{abstract}
+% Estimating the coefficients of a noisy polynomial phase signal is important in fields including radar, biology and radio communications.
+One approach attempts to perform polynomial regression on the phase of the signal.
+This is complicated by the fact that the phase is \emph{wrapped} modulo $2\pi$ and must be \emph{unwrapped} before regression can be performed.
+%A recent approach suggested by the authors is to perform the unwrapping in a least squares manner.
+%It was shown how the resulting estimator could be posed as a \emph{nearest lattice point problem} in a specific lattice and could be solved using existing algorithms.
+% %It was shown by Monte Carlo simulation that this produces a remarkably accurate estimator.  
+% In this paper we consider an estimator that performs phase unwrapping in a least squares manner.
+We describe the asymptotic properties of this estimator, showing that it is strongly consistent and asymptotically normally distributed.
+%We hypothesise that the estimator produces very near maximum likelihood performance.
+% \end{abstract}
+
+% %\begin{keyword}[class=AMS]
+% %\kwd[Primary ]{}
+% %\kwd{}
+% %\kwd[; secondary ]{}
+% %\end{keyword}
+
+% \begin{keyword}
+% \kwd{Polynomial phase signals, phase unwrapping, asymptotic statistics}
+% \end{keyword}
+
+% \end{frontmatter}
+
+% % AOS,AOAS: If there are supplements please fill:
+% % \begin{supplement}[id=suppA]
+% %  \sname{Supplement A}
+% %  \stitle{Title}
+% %  \slink[url]{http://lib.stat.cmu.edu/aoas/???/???}
+% %  \sdescription{Some text}
+% % \end{supplement}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%% END AOAP START %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGIN IEEE START %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
+\documentclass[journal]{IEEEtran}
+
 \usepackage{mathbf-abbrevs}
-\newtheorem{proposition}{Proposition}
-\newtheorem{corollary}{Corollary}
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}{Lemma}
-\newtheorem{definition}{Definition}
-\newcommand{\coef}{\operatorname{coef}}
-\newcommand{\cubr}[1]{{\left\{ #1 \right\}}}
-\newcommand{\scubr}[1]{{\{ #1 \}}}
-\newcommand{\sign}[1]{\operatorname{sgn}\left( #1 \right)}
-\newcommand{\bfracpart}[1]{\big\langle #1 \big\rangle}
-\newcommand{\babs}[1]{\big\vert #1 \big\vert}
-\endlocaldefs
+
+%\newcommand{\dealias}{\operatorname{dealias}}
+
+\input{defs}
+
+%\newcommand{\term}{\emph}
+%\renewcommand{\index}[1]{}
+
+%\title{Polynomial phase estimation by phase unwrapping 1: Identifiability and strong consistency}
+\title{Polynomial phase estimation by phase unwrapping}
+
+\author{Robby~G.~McKilliam, Barry~G.~Quinn, I.~Vaughan~L.~Clarkson, Bill~Moran and Badri~N.~Vellambi%
+    \thanks{%A preliminary version of some of this material is contained in Part 3 of Robby McKilliam's PhD thesis \cite{McKilliam2010thesis}. 
+Robby~McKilliam and Badri Vellambi are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.
+Barry~Quinn is with the Department of Statistics, Macquarie University, Sydney, NSW, 2109, Australia.
+Vaughan~Clarkson is with the School of Information Technology \& Electrical Engineering, The University of Queensland, QLD., 4072, Australia.
+B.
+Moran is with the Department of Electrical Engineering and Computer
+Science, Melbourne Systems Lab, Dept of Elec \& Electronic Eng, Uni of Melbourne, Vic. 3010, Australia.}}
+% The paper headers 
+\markboth{Polynomial phase estimation by phase unwrapping}{DRAFT \today}
+
+% make the title area
 
 \begin{document} 
-
-\begin{frontmatter}
-
-% "Title of the paper"
-\title{Polynomial phase estimation by phase unwrapping}
-\runtitle{Polynomial phase estimation by phase unwrapping}
-
-% indicate corresponding author with \corref{}
-% \author{\fnms{John} \snm{Smith}\corref{}\ead[label=e1]{smith@foo.com}\thanksref{t1}}
-% \thankstext{t1}{Thanks to somebody} 
-% \address{line 1\\ line 2\\ printead{e1}}
-% \affiliation{Some University}
-
-\begin{aug}
-\author{\fnms{Robby G.} \snm{McKilliam}\thanksref{m1}\ead[label=e1]{robby.mckilliam@unisa.edu.au}},
-\author{\fnms{Barry G.} \snm{Quinn}\thanksref{m2}\ead[label=e2]{barry.quinn@mq.edu.au}},
-\author{\fnms{I. Vaughan L.} \snm{Clarkson}\thanksref{m3}\ead[label=e3]{v.clarkson@uq.edu.au}},
-\author{\fnms{Bill} \snm{Moran}\thanksref{m4}\ead[label=e4]{wmoran@unimelb.edu.au}},
-\and
-\author{\fnms{Badri N.} \snm{Vellambi}\thanksref{m1}\ead[label=e5]{badri.vellambi@unisa.edu.au}}
-\runauthor{R. McKilliam et. al.}
-
-\affiliation{University of South Australia\thanksmark{m1}, Macquarie University\thanksmark{m2}, University of Queensland\thanksmark{m3} and Melbourne University\thanksmark{m4}}
-
-\address{R. G. McKilliam and B. N. Vellambi\\
-Institute for Telecommunications Research\\
-University of South Australia, Adelaide\\
-SA, 5095, Australia\\
-\printead{e1}\\
-\phantom{E-mail:\ }\printead*{e5}}
-
-\address{B. G. Quinn\\
-Department of Statistics\\
-Macquarie University, Sydney,\\
-NSW, 2109, Australia\\
-\printead{e3}}
-
-\address{I. V. L. Clarkson\\
-School of Information Technology \& Electrical\\
-Engineering, University of Queensland,\\
-Brisbane, QLD, 4072, Australia\\
-\printead{e3}}
-
-\address{B. Moran\\
-Department of Electrical \& Electronic\\
-Engineering, Melbourne University,\\
-Melbourne, VIC, 3010, Australia\\
-\printead{e4}}
-
-\end{aug}
+\maketitle
 
 \begin{abstract}
-Estimating the coefficients of a noisy polynomial phase signal is important in fields including radar, biology and radio communications.
+Estimating the coefficients of a noisy polynomial phase signal is important in many fields including radar, biology and radio communications.
 One approach attempts to perform polynomial regression on the phase of the signal.
@@
 
-%\begin{keyword}[class=AMS]
-%\kwd[Primary ]{}
-%\kwd{}
-%\kwd[; secondary ]{}
-%\end{keyword}
-
-\begin{keyword}
-\kwd{Polynomial phase signals, phase unwrapping, asymptotic statistics}
-\end{keyword}
-
-\end{frontmatter}
-
-% AOS,AOAS: If there are supplements please fill:
-% \begin{supplement}[id=suppA]
-%  \sname{Supplement A}
-%  \stitle{Title}
-%  \slink[url]{http://lib.stat.cmu.edu/aoas/???/???}
-%  \sdescription{Some text}
-% \end{supplement}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%% END AOAP START %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGIN IEEE START %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% %\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
-% \documentclass[journal]{IEEEtran}
-
-% \usepackage{mathbf-abbrevs}
-
-% %\newcommand{\dealias}{\operatorname{dealias}}
-
-% \input{defs}
-
-% %\newcommand{\term}{\emph}
-% %\renewcommand{\index}[1]{}
-
-% %\title{Polynomial phase estimation by phase unwrapping 1: Identifiability and strong consistency}
-% \title{Polynomial phase estimation by phase unwrapping}
-
-% \author{Robby~G.~McKilliam, Barry~G.~Quinn, I.~Vaughan~L.~Clarkson, Bill~Moran and Badri~N.~Vellambi%
-%     \thanks{%A preliminary version of some of this material is contained in Part 3 of Robby McKilliam's PhD thesis \cite{McKilliam2010thesis}. 
-% Robby~McKilliam and Badri Vellambi are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.
-Barry~Quinn is with the Department of Statistics, Macquarie University, Sydney, NSW, 2109, Australia.
-Vaughan~Clarkson is with the School of Information Technology \& Electrical Engineering, The University of Queensland, QLD., 4072, Australia.
-B.
-Moran is with the Department of Electrical Engineering and Computer
-% Science, Melbourne Systems Lab, Dept of Elec \& Electronic Eng, Uni of Melbourne, Vic. 3010, Australia.}}
-% % The paper headers 
-% \markboth{Polynomial phase estimation by phase unwrapping}{DRAFT \today}
-
-% % make the title area
-
-% \begin{document} 
-% \maketitle
-
-% \begin{abstract}
-% Estimating the coefficients of a noisy polynomial phase signal is important in many fields including radar, biology and radio communications.
-One approach attempts to perform polynomial regression on the phase of the signal.
-This is complicated by the fact that the phase is \emph{wrapped} modulo $2\pi$ and must be \emph{unwrapped} before regression can be performed.
-%A recent approach suggested by the authors is to perform the unwrapping in a least squares manner.
-%It was shown how the resulting estimator could be posed as a \emph{nearest lattice point problem} in a specific lattice and could be solved using existing algorithms.
-% %It was shown by Monte Carlo simulation that this produces a remarkably accurate estimator.  
-% In this paper we consider an estimator that performs phase unwrapping in a least squares manner.
-We describe the asymptotic properties of this estimator, showing that it is strongly consistent and asymptotically normally distributed.
-%We hypothesise that the estimator produces very near maximum likelihood performance.
-% \end{abstract}
-
-% \begin{keywords}
-% Polynomial phase signals, phase unwrapping, asymptotic properties, nearest lattice point problem
-% \end{keywords}
-
-% %\bibliographystyle{unsrt}
+\begin{keywords}
+Polynomial phase signals, phase unwrapping, asymptotic properties, nearest lattice point problem
+\end{keywords}
+
+%\bibliographystyle{unsrt}
 

### 2012-10-26 paper.tex
minor

--- 2012-10-25 paper.tex
@@
 
-This result is related to what is called \emph{tightness} in the literature on empirical processes and weak convergence on metric spaces~\cite{Billingsley1999_convergence_of_probability_measures,Dudley_unif_central_lim_th_1999,Shorak_emp_proc_stat_2009}.
+This result is related to what is called \emph{tightness} or \emph{asymptotic continuity} in the literature on empirical processes and weak convergence on metric spaces~\cite{Billingsley1999_convergence_of_probability_measures,Dudley_unif_central_lim_th_1999,Shorak_emp_proc_stat_2009,van2009empirical}.
 The lemma is different from what is usually proved in the literature because the function $p_{nN}(\psibf) = \sum_{k=0}^m\left( \tfrac{n}{N}\right)^k\psi_k$
@@
 6]{Matousek_lect_disc_geom_2002}.
-Our proof is based on a technique called \emph{symmetrisation} and another technique called \emph{chaining} (also known as \emph{bracketing})~\cite{Pollard_asymp_empi_proc_1989}.
+Our proof is based on a technique called \emph{symmetrisation} and another technique called \emph{chaining} (also known as \emph{bracketing})~\cite{Pollard_asymp_empi_proc_1989,van2009empirical}.
 
@@
 Let $\{g_n\}$ be a sequence of independent standard normal random variables, independent of the phase noise sequence $\{\Phi_n\}$.
-The symmetrisation argument~\cite[Sec.~4]{Pollard_asymp_empi_proc_1989}\cite{Gine_Zinn_symmetrisation_1984} can be used to show that
+The symmetrisation argument~\cite[Sec.~4]{Pollard_asymp_empi_proc_1989}\cite{Gine_Zinn_symmetrisation_1984,van2009empirical} can be used to show that
 %Lemma~\ref{lem:symmetrisation} shows that
@@
 
-The proofs of Lemmas~\ref{lem:chaining2}~and~\ref{lem:chaining3} are based on a technique called \emph{chaining} (or \emph{bracketing})~\cite{Dudley_unif_central_lim_th_1999,Ossiander_clt_bracketing_1984,Pollard_asymp_empi_proc_1989,Pollard_new_ways_clts_1986}.
+The proofs of Lemmas~\ref{lem:chaining2}~and~\ref{lem:chaining3} are based on a technique called \emph{chaining} (or \emph{bracketing})~\cite{Dudley_unif_central_lim_th_1999,Ossiander_clt_bracketing_1984,Pollard_asymp_empi_proc_1989,Pollard_new_ways_clts_1986,van2009empirical}.
 The proofs here follow those of Pollard~\cite{Pollard_asymp_empi_proc_1989}.
@@
 
-\bibliographystyle{acmtrans-ims}
+%\bibliographystyle{acmtrans-ims}
+\bibliographystyle{IEEEbib}
 \bibliography{bib}

### 2012-10-26 paper.tex
added to paper

--- 2012-10-26 paper.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGIN AOAP START %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% % Template for the submission to:
-% %   The Annals of Probability           [aop]
-% %   The Annals of Applied Probability   [aap]
-% %   The Annals of Statistics            [aos] 
-% %   The Annals of Applied Statistics    [aoas]
-% %   Stochastic Systems                  [ssy]
-% %
-% %Author: In this template, the places where you need to add information
-% %        (or delete line) are indicated by {???}.  Mostly the information
-% %        required is obvious, but some explanations are given in lines starting
-% %Author:
-% %All other lines should be ignored.  After editing, there should be
-% %no instances of ??? after this line.
-
-% % use option [preprint] to remove info line at bottom
-% % journal options: aop,aap,aos,aoas,ssy
-% % natbib option: authoryear
-% \documentclass[aap]{imsart}
-
-% \usepackage{amsthm,amsmath,amsfonts,amssymb}
-% \usepackage[square,comma,numbers,sort&compress]{natbib}
-% %\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
-% \usepackage{units}
-% \usepackage{ifpdf}
-% \ifpdf
-%   \usepackage[pdftex]{graphicx}
-% \else
-%   \usepackage{graphicx}% standard graphics package for inclusion images and eps files into LaTeX document
-% \fi
-
-% % provide arXiv number if available:
-% %\arxiv{arXiv:0000.0000}
-
-% % put your definitions there:
-% \startlocaldefs
-% \newcommand{\reals}{{\mathbb R}}
-% \newcommand{\ints}{{\mathbb Z}}
-% \newcommand{\term}{\emph}
-% \newcommand{\var}{\operatorname{var}}
-% \newcommand{\covar}{\operatorname{covar}}
-% %\newcommand{\prob}{{\mathbb P}}
-% \newcommand{\prob}{\operatorname{Pr}}
-% \newcommand{\expect}{{\mathbb E}}
-% \newcommand{\dealias}{\operatorname{dealias}}
-% \renewcommand{\mid}{\; ; \;}
-% \newcommand{\fracpart}[1]{\left\langle #1 \right\rangle}
-% \newcommand{\sfracpart}[1]{\langle #1 \rangle}
-% \newcommand{\abs}[1]{\left\vert #1 \right\vert}
-% \newcommand{\sabs}[1]{\vert #1 \vert}
-% \newcommand{\floor}[1]{{\left\lfloor #1 \right\rfloor}}
-% \newcommand{\ceiling}[1]{{\left\lceil #1 \right\rceil}}
-% \newcommand{\ceil}[1]{\lceil #1 \rceil}
-% \newcommand{\round}[1]{{\left\lceil #1 \right\rfloor}}
-% \usepackage{mathbf-abbrevs}
-% \newtheorem{proposition}{Proposition}
-% \newtheorem{corollary}{Corollary}
-% \newtheorem{theorem}{Theorem}
-% \newtheorem{lemma}{Lemma}
-% \newtheorem{definition}{Definition}
-% \newcommand{\coef}{\operatorname{coef}}
-% \newcommand{\cubr}[1]{{\left\{ #1 \right\}}}
-% \newcommand{\scubr}[1]{{\{ #1 \}}}
-% \newcommand{\sign}[1]{\operatorname{sgn}\left( #1 \right)}
-% \newcommand{\bfracpart}[1]{\big\langle #1 \big\rangle}
-% \newcommand{\babs}[1]{\big\vert #1 \big\vert}
-% \endlocaldefs
-
-% \begin{document}
-
-% \begin{frontmatter}
-
-% % "Title of the paper"
-% \title{Polynomial phase estimation by phase unwrapping}
-% \runtitle{Polynomial phase estimation by phase unwrapping}
-
-% % indicate corresponding author with \corref{}
-% % \author{\fnms{John} \snm{Smith}\corref{}\ead[label=e1]{smith@foo.com}\thanksref{t1}}
-% % \thankstext{t1}{Thanks to somebody} 
-% % \address{line 1\\ line 2\\ printead{e1}}
-% % \affiliation{Some University}
-
-% \begin{aug}
-% \author{\fnms{Robby G.} \snm{McKilliam}\thanksref{m1}\ead[label=e1]{robby.mckilliam@unisa.edu.au}},
-% \author{\fnms{Barry G.} \snm{Quinn}\thanksref{m2}\ead[label=e2]{barry.quinn@mq.edu.au}},
-% \author{\fnms{I. Vaughan L.} \snm{Clarkson}\thanksref{m3}\ead[label=e3]{v.clarkson@uq.edu.au}},
-% \author{\fnms{Bill} \snm{Moran}\thanksref{m4}\ead[label=e4]{wmoran@unimelb.edu.au}},
-% \and
-% \author{\fnms{Badri N.} \snm{Vellambi}\thanksref{m1}\ead[label=e5]{badri.vellambi@unisa.edu.au}}
-% \runauthor{R. McKilliam et. al.}
-
-% \affiliation{University of South Australia\thanksmark{m1}, Macquarie University\thanksmark{m2}, University of Queensland\thanksmark{m3} and Melbourne University\thanksmark{m4}}
-
-% \address{R. G. McKilliam and B. N. Vellambi\\
-% Institute for Telecommunications Research\\
-% University of South Australia, Adelaide\\
-% SA, 5095, Australia\\
-% \printead{e1}\\
-% \phantom{E-mail:\ }\printead*{e5}}
-
-% \address{B. G. Quinn\\
-% Department of Statistics\\
-% Macquarie University, Sydney,\\
-% NSW, 2109, Australia\\
-% \printead{e3}}
-
-% \address{I. V. L. Clarkson\\
-% School of Information Technology \& Electrical\\
-% Engineering, University of Queensland,\\
-% Brisbane, QLD, 4072, Australia\\
-% \printead{e3}}
-
-% \address{B. Moran\\
-% Department of Electrical \& Electronic\\
-% Engineering, Melbourne University,\\
-% Melbourne, VIC, 3010, Australia\\
-% \printead{e4}}
-
-% \end{aug}
-
-% \begin{abstract}
-% Estimating the coefficients of a noisy polynomial phase signal is important in fields including radar, biology and radio communications.
-One approach attempts to perform polynomial regression on the phase of the signal.
-This is complicated by the fact that the phase is \emph{wrapped} modulo $2\pi$ and must be \emph{unwrapped} before regression can be performed.
-%A recent approach suggested by the authors is to perform the unwrapping in a least squares manner.
-%It was shown how the resulting estimator could be posed as a \emph{nearest lattice point problem} in a specific lattice and could be solved using existing algorithms.
-% %It was shown by Monte Carlo simulation that this produces a remarkably accurate estimator.  
-% In this paper we consider an estimator that performs phase unwrapping in a least squares manner.
-We describe the asymptotic properties of this estimator, showing that it is strongly consistent and asymptotically normally distributed.
-%We hypothesise that the estimator produces very near maximum likelihood performance.
-% \end{abstract}
-
-% %\begin{keyword}[class=AMS]
-% %\kwd[Primary ]{}
-% %\kwd{}
-% %\kwd[; secondary ]{}
-% %\end{keyword}
-
-% \begin{keyword}
-% \kwd{Polynomial phase signals, phase unwrapping, asymptotic statistics}
-% \end{keyword}
-
-% \end{frontmatter}
-
-% % AOS,AOAS: If there are supplements please fill:
-% % \begin{supplement}[id=suppA]
-% %  \sname{Supplement A}
-% %  \stitle{Title}
-% %  \slink[url]{http://lib.stat.cmu.edu/aoas/???/???}
-% %  \sdescription{Some text}
-% % \end{supplement}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%% END AOAP START %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGIN IEEE START %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 %\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
@@
 
-%\title{Polynomial phase estimation by phase unwrapping 1: Identifiability and strong consistency}
-\title{Polynomial phase estimation by phase unwrapping}
+\title{Polynomial phase estimation by phase unwrapping 1: Identifiability and strong consistency}
+%\title{Polynomial phase estimation by phase unwrapping}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%% END IEEE START %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \section{Introduction} \label{intro}
 
-Polynomial phase signals arise in fields including radar, sonar, geophysics, speech analysis, biology, and radio communication~\cite{Hlawatsch_lin_quad_time_freq_spmag_1992,Ridleyspeechpolyphase1989, Suga_1975_bats_echolocation, Moss_2005echolocation}.
-In radar and sonar applications polynomial phase signals arise when acquiring radial velocity and acceleration (and higher order motion descriptors) of a target from a reflected signal, and also in continuous wave radar and low probability of intercept radar.
-In biology, polynomial phase signals are used to describe the sounds emitted by bats and dolphins for echo location.
+Polynomial phase signals arise in fields including radar, sonar, geophysics, biology, and radio communication \cite{Angeby_estimating_2000}.
+In radar and sonar applications polynomial phase signals arise when acquiring radial velocity and acceleration (and higher order motion descriptors) of a target from a reflected signal, and also in continuous wave radar and low probability of intercept radar~\cite{Levanon_Radar_signals_2004}.
+In biology, polynomial phase signals are used to describe the sounds emitted by bats and dolphins for echo location \citep{Suga_1975_bats_echolocation, Moss_2005echolocation}.
 
@@
 In this paper we only consider uniform sampling, where the gap between consecutive samples is constant.
-In this case we can always consider the samples to be taken at some set of consecutive integers and our sampled polynomial phase signal is $s_n = s(n) = e^{2\pi j y(n)}$, where $n$ is an integer.
-Of practical importance is the estimation of the coefficients $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ from a number, say $N$, of consecutive observations of the noisy sampled signal
+In this case we can always consider the samples to be taken at some set of consecutive integers and our sampled polynomial phase signal looks like
+\[
+s_n = s(n) = e^{2\pi j y(n)},
+\] 
+where $n$ is an integer.  Of practical importance is the estimation of the coefficients $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ from a number, say $N$, of observations of the noisy sampled signal
 \begin{equation}\label{eq:Y_nsamplednoisey}
@@
 \end{equation}
-where $\rho$ is a real number greater than zero representing the (usually unknown) signal amplitude and $\{X_n, n \in \ints\}$ is a sequence of complex noise variables.
+where $\rho$ is a real number greater than zero representing the (usually unknown) signal amplitude and $\{X_n, n \in \ints\}$ is a sequence complex noise variables.
 In order to ensure identifiability it is necessary to restrict the $m+1$ coefficients to a region of $m+1$ dimensional Euclidean space $\reals^{m+1}$ called an \emph{identifiable region}.
-It was shown by some of the authors \cite{McKilliam2009IndentifiabliltyAliasingPolyphase} that an identifiable region tessellates a particular $m+1$ dimensional lattice.
+It was shown in \cite{McKilliam2009IndentifiabliltyAliasingPolyphase} that an identifiable region tessellates a particular $m+1$ dimensional lattice.
 We discuss this in Section \ref{sec:ident_aliasing}.
@@
 
-An obvious estimator of the unknown coefficients is the least squares estimator.  
-%This is also the maximum likelihood estimator (MLE) when the noise sequence $\{X_n\}$ is white and Gaussian.  
-When $m=0$ (phase estimation) or $m=1$ (frequency estimation) the least squares estimator is an effective approach, being both computationally efficient and statistically accurate~\cite{Hannan1973,Quinn2001,McKilliam_mean_dir_est_sq_arc_length2010}.
-When $m \geq 2$ the computational complexity of the least squares estimator is large, and alternative estimators have been considered for this reason.
-These can loosely be grouped into two classes; estimators based on polynomial phase transforms, such as the discrete polynomial phase transform~\cite{Peleg_DPT_1995} and the high order phase function~\cite{Farquharson_another_poly_est_2005,Porat_asympt_HAF_DPT_1996}; and estimators based on phase unwrapping, such as Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, and Morelande's Bayesian unwrapping estimator~\cite{Morelande_bayes_unwrapping_2009_tsp}.
+An obvious estimator of the unknown coefficients is the least squares estimator (LSE).
+This is also the maximum likelihood estimator (MLE) when the noise sequence $\{X_n\}$ is white and Gaussian.
+When $m=0$ (phase estimation) or $m=1$ (frequency estimation) the LSE is an effective approach, being both computationally efficient and statistically accurate \cite{Quinn2009_dasp_phase_only_information_loss,Hannan1973,Quinn2001,McKilliam_mean_dir_est_sq_arc_length2010}\cite[Sec.~6.4~and~9.1]{McKilliam2010thesis}.
+When $m \geq 2$ the computational complexity of the LSE is large~\cite[Sec.~10.1]{McKilliam2010thesis}\cite{Abatzoglou_ml_chirp_1986}.
+For this reason many authors have considered alternative approaches to polynomial phase estimation.
+These can loosely be grouped into two classes, estimators based on polynomial phase transforms, such as the discrete polynomial phase transform (DPT)~\cite{Peleg_DPT_1995,Peleg1991_est_class_PPS_1991,Porat_asympt_HAF_DPT_1996} and the high order phase function~\cite{Farquharson_another_poly_est_2005}, and estimators based on phase unwrapping, such as Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, the estimator of Djuric and Kay~\cite{Djuric_phase_unwrap_chirp_1990}, and Morelande's Bayesian unwrapping estimator~\cite{Morelande_bayes_unwrapping_2009_tsp}.
 
 In this paper we consider the estimator that results from unwrapping the phase in a least squares manner.
-We call this the \emph{least squares unwrapping} (LSU) estimator.
-It was shown by some of the authors \cite[Sec.
-8.1]{McKilliam2010thesis}\cite{McKilliam2009asilomar_polyest_lattice} that the LSU estimator can be computed by finding a nearest point in a lattice~\cite{Agrell2002}, and Monte-Carlo simulations were used to show the LSU estimator's favourable statistical performance.
+We call this the \emph{least squares unwrapping estimator} (LSU)~\cite{McKilliam2009asilomar_polyest_lattice, McKilliamFrequencyEstimationByPhaseUnwrapping2009}\cite[Chap.
+8]{McKilliam2010thesis}.
+It was shown in \cite{McKilliam2009asilomar_polyest_lattice, McKilliamFrequencyEstimationByPhaseUnwrapping2009} that the LSU estimator can be represented as a \emph{nearest lattice point problem}, and Monte-Carlo simulations were used to show the LSU estimator's favourable statistical performance.
 %Moreover, it was noticed that the LSU appears to work correctly for coefficients anywhere in the identifiable region.
@@
 %We also apply another known approximate nearest point algorithm, called the $K$-best method \cite{Zhan2006_K_best_sphere_decoder}, that we show provides near sphere decoder performance, but can be computed in a reasonable amount of time if $N$ is less than about 300.
-In this paper we derive the asymptotic properties of the LSU estimator.
-Under some assumptions about the distribution of the noise $X_1, \dots, X_N$, we show the estimator to be strongly consistent and asymptotically normally distributed.
-Similar results were stated without proof in~\cite{McKilliam_polyphase_est_icassp_2011}.
+In this paper we derive the asymptotic properties of the LSU estimator, showing that it is strongly consistent and asymptotically normally distributed.
+Similar results were stated without a complete proof in~\cite{McKilliam_polyphase_est_icassp_2011}.
 Here, we give a proof.
@@
 The paper is organised in the following way.
-Section~\ref{sec:lattice-theory} describes some preliminary concepts from lattice theory, and in Section~\ref{sec:ident_aliasing} we use these results to describe an identifiable region for the set of polynomial phase coefficients.
+Section~\ref{sec:lattice-theory} describes the required concepts from lattice theory.
+In Section~\ref{sec:ident_aliasing} we briefly describe the identifiable region that was also derived in~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}.
 These identifiability results are required in order to properly understand the statistical properties of polynomial phase estimators.
-In Section \ref{sec:least-squar-unwr} we describe the LSU estimator and state its asymptotic statistical properties.
-Section~\ref{sec:strongconstproof} gives the proof of strong consistency and Section~\ref{sec:centlimitproof} gives the proof of asymptotic normality.
+In Section \ref{sec:least-squar-unwr} we describe the LSU estimator and derive its asymptotic statistical properties.
 %and also consider techniques for computing the LSU as a nearest lattice point problem. 
-Section~\ref{sec:simulations} describes the results of Monte Carlo simulations with the LSU estimator.  These simulations agree with the derived asymptotic properties.
-
-\section{Lattices}\label{sec:lattice-theory}
+Section \ref{sec:simulations} describes the results of Monte Carlo simulations that compare the performance of the LSU estimator with existing estimators.
+These simulations agree with the derived asymptotic properties.
+
+%\subsection{Notation}
+%We write random variables using capital letters, such as $X$ and $Y$ and circular random variables using the capital Greek letters $\Theta$ and $\Phi$.
+We use $\round{x}$ to denote the nearest integer to $x$ with half integers rounded up and use $\fracpart{x} = x - \round{x}$ to denote the \emph{centred} fractional part.
+Both $\round{\cdot}$ and $\fractpart{\cdot}$ operate elementwise on vectors.
+
+\section{Lattice Theory}\label{sec:lattice-theory}
 
@@
 Let $\Bbf = \Qbf\Rbf$ where $\Qbf$ is orthonormal and $\Rbf$ is upper triangular with elements $r_{ij}$.
-Then the rectangular prism $\Qbf P$ where $P = \prod_{k=1}^{n}{[-\tfrac{r_{kk}}{2}, \frac{r_{kk}}{2})}$ is a set of coset representatives for $\reals^n / \Lambda$.
+Then the rectangular prism $\Qbf P$ where
+\[
+P = \prod_{k=1}^{n}{\left[-\frac{r_{kk}}{2}, \frac{r_{kk}}{2}\right)}
+\]
+is a set of coset representatives for $\reals^n / \Lambda$.
 \end{proposition}
-\begin{proof}
+\begin{IEEEproof}
 This result is well known~\cite[Chapter IX, Theorem IV]{Cassels_geom_numbers_1997}~\cite[Proposition 2.1]{McKilliam2010thesis}.
@@
 A result in the general case can be obtained similarly, but is not required here.
-\end{proof}
-
-\begin{figure}[t]
+\end{IEEEproof}
+
+\begin{figure}[tp]
 	\centering
-		\includegraphics[width=0.8\linewidth]{plots/tesselationfigures-2.mps}
+		\includegraphics[width=\linewidth]{plots/tesselationfigures-2.mps}
 		\caption{Rectangular tessellation constructed according to Proposition~\ref{prop:lattice-theory-constructing_a_rectangular_tesselating_region} where $\Lambda$ is a 2 dimensional lattice with generator matrix having columns $[1, 0.2]^\prime$ and $[0.2, 1]^\prime$.
@@
 \end{figure}
-
-% Given a point $\ybf\in\reals^n$ and a lattice $\Lambda$ contained in~$\reals^n$, the \emph{nearest lattice point problem} is to find the point $\xbf \in \Lambda$ such that the distance (with respect to a given norm) between $\ybf$ and $\xbf$ is minimised.
-Here, the Euclidean norm is assumed, so that the nearest point is
-% \[ 
-% \arg\min_{\xbf \in \Lambda} \| \xbf - \ybf \|^2.
-% \]
-% The nearest lattice point problem is known to be NP-hard under certain conditions when the lattice itself, or rather a basis thereof, is considered as an additional input parameter~\cite{micciancio_hardness_2001, Jalden2005_sphere_decoding_complexity}.
-Nevertheless, algorithms exist that can compute a nearest lattice point in reasonable time if the dimension is small.
-One such algorithm, introduced by Pohst~\cite{Pohst_sphere_decoder_1981}, was popularised in the signal processing and communications fields by Viterbo and Boutros~\cite{Viterbo_sphere_decoder_1999} and has since been called the \emph{sphere decoder}.
-
-% Approximate algorithms for computing the nearest point have also been studied.
-A classic example is \emph{Babai's nearest plane algorithm}~\cite{Babai1986}, which requires $O(n^4)$ arithmetic operations in the worst case where $n$ is the dimension of the lattice and only $O(n^2)$ if the lattice basis is \emph{Lov\'asz reduced}~\cite{Lenstra1982}\index{Lov\'asz reduced}.
-Recently, various approximate techniques for solving the nearest lattice point problem have been motivated by applications to multiple-input multiple-output communications.
-An example is the \emph{$K$-best algorithm}~\cite{Zhan2006_K_best_sphere_decoder} that works similarly to the sequential $M$-algorithm~\cite{Anderson1984_seq_coding_alg} used in coding theory.
-Yet another example is the \emph{fixed sphere decoder}~\cite{Jalden2009_error_prob_fixed_sphere_decoder,Barbero2008_fixed_sphere_decoder}.
-
-% In~\cite{McKilliam2009asilomar_polyest_lattice}, it was shown how the least squares phase unwrapping estimator of a polynomial phase signal can be computed by finding a nearest point in a lattice.
-We will briefly describe this result in Section~\ref{sec:least-squar-unwr}.
 
@@
 That is, two or more distinct polynomial-phase signals can take exactly the same values at the sample points.
-Understanding how aliasing occurs is crucial to understanding the behaviour of polynomial phase estimators.
-The aliasing properties are described in~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}, but, here we present the properties in a way that is better suited to studying the LSU estimator.
-%A description of the aliasing properties of polynomial phase signals of order $2$ has been independently discovered in~\cite{Abatzoglou_ml_chirp_1986}~and~\cite{Angeby_PPS_aliasing_2000}.
+These aliasing results are also given in~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}, but the presentation here is different, and is better suited to studying the asymptotic properties of the LSU estimator.
+A description of the aliasing properties of polynomial phase signals of order $2$ has been independently discovered in~\cite{Abatzoglou_ml_chirp_1986}~and~\cite{Angeby_PPS_aliasing_2000}.
 The results here are for polynomial phase signals of any order.
@@
 That is, $\mathcal{Z}$ contains all polynomials $p$ such that $p(n)$ is an integer whenever $n$ is an integer.
-Let $y$ and $z$ be two distinct polynomials such that $z = y + p$ for some polynomial $p$ in $\mathcal{Z}$.
-The two polynomial phase signals $s(t) = e^{2\pi j y(t)}$ and $r(t) = e^{2\pi j z(t)}$ are distinct because $y$ and $z$ are distinct, but if we sample $s$ and $r$ at the integers
+Let $y$ and $z$ be two \emph{distinct} polynomials such that $z = y + p$ for some polynomial $p$ in $\mathcal{Z}$. The two polynomial phase signals
+\[
+s(t) = e^{2\pi j y(t)} \qquad \text{and} \qquad r(t) = e^{2\pi j z(t)}
+\]
+are distinct because $y$ and $z$ are distinct, but if we sample $s$ and $r$ at the integers  
 \begin{align*}
-s(n) = e^{2\pi j y(n)} =  e^{2\pi j y(n)} e^{2\pi j p(n)} = e^{2\pi j (y(n) + p(n))} = e^{2\pi j z(n)} = r(n)
+s(n) &= e^{2\pi j y(n)} =  e^{2\pi j y(n)} e^{2\pi j p(n)} \\
+&= e^{2\pi j (y(n) + p(n))} = e^{2\pi j z(n)} = r(n)
 \end{align*}
@@
 
-It may be helpful to observe Figures~\ref{fig:circstatplot_line},~\ref{fig:circstatplot_quad}~and~\ref{fig:circstatplot_cube}.
-In these, the phase (divided by $2\pi$) of two distinct polynomial phase signals is plotted on the left, and on the right the principal component of the phase (also divided by $2\pi$) is plotted.
+It may be helpful to observe Figures~\ref{fig:circstatplot_zero} to~\ref{fig:circstatplot_cube}.
+In these, the phase (divided by $2\pi$) of two distinct polynomial phase signals is plotted on the left, and on the right the principal component of the phase is plotted.
 The circles display the samples at the integers.
@@
 \begin{equation} \label{eq:lem_polynomial}
-c_0 p_0 + c_1 p_1 + \dots + c_m p_m, \qquad c_0,c_1,\dots,c_m \in \ints.
-\end{equation}
+c_0 p_0 + c_1 p_1 + \dots + c_m p_m,
+\end{equation}
+where the $c_i \in \ints$.
 \end{lemma}
-\begin{proof}
+\begin{IEEEproof}
 See~\citep[p. 2]{cahen_integer-valued_1997} or~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}. 
-\end{proof}
-% \begin{proof}
+\end{IEEEproof}
+% \begin{IEEEproof}
 % Note that $x(x-1)(x-2)\dots(x-k+1)$ is divisible by all integers $1,2,\dots,k$ and so $p_k$ takes integer values for all $x\in\ints$.
@@
 The proof follows by induction because $f(0) = c_0 \in \ints$.
-% \end{proof}
+% \end{IEEEproof}
 
@@
 \[
-\coef(g) = \left[ \begin{array}{ccccc} a_0 & a_1 & a_2 & \dots & a_m \end{array} \right]^\prime
-\]
-denote the column vector of length $m+1$ containing the coefficients of $g$.
-We use superscript $^\prime$ to indicate the vector or matrix transpose.
-If $y$ and $z$ differ by a polynomial from $\mathcal{Z}$ then $y = z + p$ where $p \in \mathcal{Z}$ and also $\coef(y) = \coef(z) + \coef(p)$.
+\coef(g) = \left[ \begin{array}{ccccc} a_0 & a_1 & a_2 & \dots & a_m, \end{array} \right]^\prime,
+\]
+where superscript $^\prime$ indicates the transpose, denote the column vector of length $m+1$ containing the coefficients of $g$.
+If $y$ and $z$ differ by a polynomial from $\mathcal{Z}$ then we can write $y = z + p$ where $p \in \mathcal{Z}$ and then also $\coef(y) = \coef(z) + \coef(p)$.
 %\footnote{In group theory terminology $\coef(\cdot)$ coupled with vector addition is called a \term{group homomorphism}.\index{group homomorphism}}. 
@@
 \[
-L_{m+1} = \{ \coef(p) \mid p \in \mathcal{Z} \}
+L_{m+1} = \{ \coef(p) \mid p \in \mathcal{Z} \},
 \]
@@
 
-%\begin{figure}[tp]
-%	\centering
-%		\includegraphics[width=\linewidth]{plots/circstatfigzero-1.mps}
-%		\caption{The zeroth order polynomials $\tfrac{7}{10}$ (solid line) and $\tfrac{17}{10}$ (dashed line).}
-%		\label{fig:circstatplot_zero}
-%\end{figure}
-
 \begin{figure}[p]
 	\centering
-		\includegraphics[width=0.85\linewidth]{plots/circstatfig-1.mps}
+		\includegraphics[width=\linewidth]{plots/circstatfigzero-1.mps}
+		\caption{The zeroth order polynomials $\tfrac{7}{10}$ (solid line) and $\tfrac{17}{10}$ (dashed line).}		
+\label{fig:circstatplot_zero}
+\end{figure}
+
+\begin{figure}[p]
+	\centering
+		\includegraphics[width=\linewidth]{plots/circstatfig-1.mps}
 		\caption{The first order polynomials $\tfrac{1}{10}(3 + 8t)$ (solid) and $\tfrac{1}{10}(33 - 2t)$ (dashed line).}
@@
 	\centering
-		\includegraphics[width=0.85\linewidth]{plots/circstatfigquad-1.mps}
+		\includegraphics[width=\linewidth]{plots/circstatfigquad-1.mps}
 		\caption{The quadratic polynomials $\tfrac{1}{10} (15 - 15 t + 4 t^2)$ (solid line) and $\tfrac{1}{10}(25 -  t^2)$  (dashed line).}
@@
 	\centering
-		\includegraphics[width=0.85\linewidth]{plots/circstatfigcube-1.mps}
+		\includegraphics[width=\linewidth]{plots/circstatfigcube-1.mps}
 		\caption{The cubic polynomials $\tfrac{1}{160} (174 + 85 t - 118 t^2 + 40 t^3)$ (solid line) and $\tfrac{1}{48} (84 + 19 t + 12 t^2 - 4 t^3)$  (dashed line).}
@@
 
-As an example consider the polynomial phase signal of order zero $e^{2\pi j \mu_0}$.
-Since $e^{2\pi j \mu_0} = e^{2\pi j(\mu_0 + k)}$ for any integer $k$ we must, in order to ensure identifiability, restrict $\mu_0$ to some interval of length $1$.
+As an example consider the polynomial phase signal of order zero $e^{2\pi j \mu_0} = e^{2\pi j(\mu_0 + k)}$ for any integer $k$.
+In order to ensure identifiability, we must restrict $\mu_0$ to some interval of length $1$.
 A natural choice is the interval $[-\nicefrac{1}{2}, \nicefrac{1}{2})$.
@@
 Observe that $\Pbf$ is upper triangular with $k$th diagonal element equal to $\tfrac{1}{k!}$.
-So this rectangular set is
+So this rectangular region is
 \begin{equation}\label{eq:rectangular_identifiable_region}
@@
 \end{equation}
-We will make use of this set when deriving the statistical properties of the LSU estimator in the next section.
-
-We define the function $\dealias(\xbf)$ to take $\xbf\in\reals^{m+1}$ to its coset representative inside $B$.
+We will make use of this region when deriving the statistical properties of the LSU estimator in the next section.
+
+Given vectors $\xbf$ and $\ybf$ in $\reals^{m+1}$ we say that $\xbf \equiv \ybf \bmod L_{m+1}$ if $\xbf$ and $\ybf$ differ by a lattice point in $L_{m+1}$.
+We define the function $\dealias(\xbf)$ to take $\xbf$ to its coset representative inside $B$.
 That is, $\dealias(\xbf) = \zbf \in B$ where $\xbf - \zbf \in L_{m+1}$.
 %The effect of $\dealias(\xbf)$ is to \emph{dealias} the polynomial phase coefficients.
-When $m = 0$ or $1$ $\dealias(\xbf) = \fracpart{\xbf}$ where $\fracpart{\xbf} = \xbf - \round{\xbf}$ denotes the (centered) fractional part and $\round{\xbf}$ denotes the nearest integer to $\xbf$ with half integers rounded upwards and both $\fracpart{\cdot}$ and $\round{\cdot}$ operate on vectors elementwise.
+When $m = 0$ or $1$ then $\dealias(\xbf) = \fracpart{\xbf}$ where $\fracpart{\xbf} = \xbf - \round{\xbf}$ denotes the (centered) fractional part and $\round{\xbf}$ denotes the nearest integer to $\xbf$ with half integers rounded upwards and both $\fracpart{\cdot}$ and $\round{\cdot}$ operate on vectors elementwise.
 For $m \geq 2$ the function $\dealias(\xbf)$ can be computed by a simple sequential algorithm \cite[Sec.
@@
 Recall that we desire to estimate the coefficients $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ from the noisy samples $Y_1, \dots, Y_N$ given in~\eqref{eq:Y_nsamplednoisey}.
-We take the complex argument of $Y_n$ and divide by $2\pi$ to obtain
+We take the complex argument of the $Y_n$ and divide by $2\pi$ to obtain
 \begin{equation}\label{eq:noise_circ_poly}
@@
 \end{equation}
-where $\angle$ denotes the complex argument (or phase), and 
+where $\angle{z}$ denotes the complex argument of the complex number $z$, and 
 \[
@@
 \] 
-is a random variable representing the `phase noise' induced by $X_n$.
-If the distribution of $X_n$ is circularly symmetric (the phase $\angle X_n$ is uniformly distributed on $[-\pi, \pi)$ and is independent of the magnitude $\abs{X_n}$) then the distribution of $\Phi_n$ is the same as the distribution of $\tfrac{1}{2\pi}\angle(1 + \rho^{-1}X_n)$.
-If $X_1, \dots, X_N$ are circularly symmetric and identically distributed, then $\Phi_1, \dots, \Phi_n$ are also identically distributed.
+are random variables representing the phase noise induced by the $X_n$~\cite{Tretter1985,Quinn2009_dasp_phase_only_information_loss}.
+If the distribution of $X_n$ is circularly symmetric (i.e., the angle $\angle X_n$ is uniformly distributed on $[-\pi, \pi)$ and is independent of the magnitude $\abs{X_n}$) then the distribution of $\Phi_n$ is the same as the distribution of $\tfrac{1}{2\pi}\angle(1 + \rho^{-1}X_n)$.
+If the $X_1, \dots, X_N$ are circularly symmetric and identically distributed, then $\Phi_1, \dots, \Phi_n$ are also identically distributed.
 
@@
 \end{equation}
-The least squares unwrapping estimator is defined as those coefficients $\widehat{\mubf} = [\widehat{\mu}_0, \dots, \widehat{\mu}_m]$ that minimise $SS$ over the identifiable region $B$, i.e., the LSU estimator is,
+The least squares unwrapping estimator is defined as those coefficients $\widehat{\mu}_0, \dots, \widehat{\mu}_m$ that minimise $SS$ over the identifiable region $B$. That is, the LSU estimator is,
 \begin{equation}\label{eq:hatmubfLSUdefn}
@@
 
-The next theorem describes the asymptotic properties of the LSU estimator.
-Before we state the theorem it is necessary to understand some of the properties of the phase noise $\Phi_1,\dots,\Phi_N$, which are \emph{circular} random variables with support on $[-\nicefrac{1}{2}, \nicefrac{1}{2})$~\cite{McKilliam2010thesis,McKilliam_mean_dir_est_sq_arc_length2010,Mardia_directional_statistics,Fisher1993}.
+The next theorem describes the asymptotic properties of this estimator.
+Before we give the proof it is necessary to understand some of the properties of the phase noise $\Phi_1,\dots,\Phi_N$, which are \emph{circular} random variables with support on $[-\nicefrac{1}{2}, \nicefrac{1}{2})$~\cite{McKilliam2010thesis,McKilliam_mean_dir_est_sq_arc_length2010,Mardia_directional_statistics,Fisher1993}.
 Circular random variables are often considered modulo $2\pi$ and therefore have support $[-\pi, \pi)$ with $-\pi$ and $\pi$ being identified as equivalent.
@@
 
-The \emph{intrinsic mean} (or \emph{Fr\'{e}chet mean}) of $\Phi_n$ is defined as~\cite{McKilliam_mean_dir_est_sq_arc_length2010,Bhattacharya_int_ext_means_2003,Bhattacharya_int_ext_means_2005},
+The \emph{intrinsic mean} or \emph{Fr\'{e}chet mean} of $\Phi_n$ is defined as~\cite{McKilliam_mean_dir_est_sq_arc_length2010,bwhk07a,Bhattacharya_int_ext_means_2003,Bhattacharya_int_ext_means_2005},
 \begin{equation}\label{eq:intrmeandefn}
@@
 \[
-\sigma_{\text{intr}}^2 = \expect\fracpart{\Theta - \mu_{\text{intr}}}^2 = \min_{\mu \in [-\nicefrac{1}{2}, \nicefrac{1}{2})} \expect \fracpart{\Phi_n - \mu}^2,
+\sigma_{\text{intr}}^2 = E\fracpart{\Theta - \mu_{\text{intr}}}^2 = \min_{\mu \in [-\nicefrac{1}{2}, \nicefrac{1}{2})} \expect \fracpart{\Phi_n - \mu}^2,
 \]
@@
 Denote the elements of $\widehat{\lambdabf}_N$ by $\widehat{\lambda}_{0,N}, \dots, \widehat{\lambda}_{m,N}$.
-Suppose $\Phi_1, \dots, \Phi_N$ are independent and identically distributed with zero intrinsic mean, intrinsic variance $\sigma^2$, and probability density function $f$, then:
+Suppose $\Phi_1, \dots, \Phi_N$ are independent and identically distributed with zero intrinsic mean, intrinsic variance $\sigma^2$, and pdf $f$, then:
 \begin{enumerate}
 \item (Strong consistency) $N^k \widehat{\lambda}_{k,N}$ converges almost surely to $0$ as $N\rightarrow\infty$ for all $k = 0, 1, \dots, m$.
-\item (Asymptotic normality) If $f(\fracpart{x})$ is continuous at $x = -\nicefrac{1}{2}$ and if $f(-\nicefrac{1}{2}) < 1$ then the distribution of the vector
+\item (Asymptotic normality) If $f(\fracpart{x})$ is continuous at $x = -\nicefrac{1}{2}$ and $f(-\nicefrac{1}{2}) < 1$ then the distribution of
 \[
@@
 \]
-where $\Cbf$ is the $m+1$ by $m+1$ Hilbert matrix with elements $C_{ik} = 1/(i + k + 1)$ for $i,k \in \{0, 1, \dots, m\}$.
+where $\Cbf$ is the $(m+1)\times (m+1)$ Hilbert matrix with elements $C_{ik} = 1/(i + k + 1)$ for $i,k = \{0, 1, \dots, m\}$.
 \end{enumerate}
@@
 Section~\ref{sec:strongconstproof} proves strong consistency and Section~\ref{sec:centlimitproof} proves asymptotic normality.
-Proofs for the case when $m=0$ were given in~\cite{McKilliam_mean_dir_est_sq_arc_length2010} and for the case when $m=1$ were given in~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.
+The proofs for the case of $m=0$ were given in~\cite{McKilliam_mean_dir_est_sq_arc_length2010} and for the case when $m=1$ were given in~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.
 The proofs here take a similar approach, but require new techniques.
-%The proof of asymptotic normality in~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009} contains a subtle error that is amended by the proof given in this paper.
-%Before begining the proof we first discuss the assumptions made in the theorem statement.
-The theorem gives conditions on $\dealias(\tilde{\mubf} - \widehat{\mubf})$ rather than directly on the difference $\tilde{\mubf} - \widehat{\mubf}$.
+%Before giving the proof we make the following remarks.
+The theorem gives conditions on the \emph{dealiased} difference $\dealias(\tilde{\mubf} - \widehat{\mubf})$ between the true coefficients $\tilde{\mubf}$ and the estimated coefficients $\widehat{\mubf}$ rather than directly on the difference $\tilde{\mubf} - \widehat{\mubf}$.
 To see why this makes sense, consider the case when $m=0$, $\tilde{\mu}_0 = -0.5$ and $\widehat{\mu}_0 = 0.49$, so that $\tilde{\mu}_0 - \widehat{\mu}_0 = -0.99$.
-However, the two phases are obviously close, since the phases $\pm 0.5$ are actually the same.
-In this case $\dealias(\tilde{\mu}_0 - \widehat{\mu}_0) = \fracpart{\tilde{\mu}_0 - \widehat{\mu}_0} = 0.01$ as expected.
+However, the two phases are obviously close, since the phase $\pm 0.5$ are actually the same.
+In this case
+\[
+\dealias(\tilde{\mu}_0 - \widehat{\mu}_0) = \fracpart{\tilde{\mu}_0 - \widehat{\mu}_0} = 0.01.
+\] 
 The same reasoning holds for $m > 0$.
@@
 In the interest of simplicity we only consider the case when $\Phi_1, \dots, \Phi_N$ are identically distributed here.
-If $X_n$ is circularly symmetric with density function nonincreasing with magnitude $\abs{X_n}$, then the corresponding $\Phi_n$ necessarily has zero intrinsic mean~\cite[Theorem 5.2]{McKilliam2010thesis}.
+If $X_n$ is circularly symmetric with density function nonincreasing with magnitude $\abs{X_n}$, then, the corresponding $\Phi_n$ necessarily has zero intrinsic mean~\cite[Theorem 5.2, page 78]{McKilliam2010thesis}.
 Thus, our theorem covers commonly used distributions for $X_1, \dots, X_N$, such as the normal distribution.
 
-The proof of asymptotic normality places requirements on the probability density function $f$ of the phase noise $\Phi_1, \dots, \Phi_N$.
-The requirement that $\Phi_1, \dots, \Phi_N$ have zero intrinsic mean implies $f(-\nicefrac{1}{2}) \leq 1$~\cite[Lemma~1]{McKilliam_mean_dir_est_sq_arc_length2010}, so the only case not handled is when $f(-\nicefrac{1}{2}) = 1$ or when $f(\fracpart{x})$ is discontinuous at $x = -\nicefrac{1}{2}$.
-In this exceptional case other expressions for the asymptotic variance can be found (similar to \cite[Theorem 3.1]{Hotz_circle_means_2011}), but this comes at a substantial increase in complexity and we have omitted them for this reason.
+The proof of asymptotic normality places requirements on the pdf $f$ of the phase noise.
+The requirement that $\Phi_1, \dots, \Phi_N$ have zero intrinsic mean implies that $f(-\nicefrac{1}{2}) \leq 1$~\cite[Lemma~1]{McKilliam_mean_dir_est_sq_arc_length2010}, so the only case not handled is when equality holds, i.e., when $f(-\nicefrac{1}{2}) = 1$ or when $f(\fracpart{x})$ is discontinuous at $x = -\nicefrac{1}{2}$.
+In this exceptional case other expressions for the asymptotic variance can be found (similar to \cite[Theorem 3.1]{Hotz_circle_means_2011}), but this comes at a substantial increase in complexity and for this reason we have omitted them.
 %Distributions that do not satisfy these requirements are unlikely to be needed in practice.
+
+%Finally note that $\Cbf$ is a \emph{Hilbert matrix}\index{Hilbert matrix} and the elements of the inverse $\Cbf^{-1}$ are given by
+ %\[
+ %C^{-1}_{ik} =\frac{i+k+1}{(-1)^{i+k}}{m+i+1 \choose m-k}{m+k+1 \choose m-i}{i+k \choose i}^2.
+ %\]
+%Hilbert matrices are particularly ill conditioned and difficult to numerically invert.  This problem can be avoided using the above formula. 
+%I MIGHT HAVE AN ERROR IN THIS FORMULA!
 
@@
 \begin{align*}
-\fracpart{\sum_{k=0}^{m}\lambda_k n^k } = \fracpart{\sum_{k=0}^{m}(\tilde{\mu}_k - \mu_k - p_k) n^k } = \fracpart{\sum_{k=0}^{m}(\tilde{\mu}_k - \mu_k) n^k }.
+\fracpart{\sum_{k=0}^{m}\lambda_k n^k } &= \fracpart{\sum_{k=0}^{m}(\tilde{\mu}_k - \mu_k - p_k) n^k } \\
+&= \fracpart{\sum_{k=0}^{m}(\tilde{\mu}_k - \mu_k) n^k }.
 \end{align*}
@@
 \begin{align*}
-V_N(\lambdabf) = \frac{1}{N}\expect\sum_{n=1}^{N}\fracpart{ \Phi_n + \sum_{k=0}^m{\lambda_k n^k} }^2 = \frac{1}{N}\sum_{n=1}^{N}\expect\fracpart{ \Phi_n + \fracpart{z(n)} }^2.
+V_N(\lambdabf) &= \frac{1}{N}\expect\sum_{n=1}^{N}\fracpart{ \Phi_n + \sum_{k=0}^m{\lambda_k n^k} }^2 \\
+&= \frac{1}{N}\sum_{n=1}^{N}\expect\fracpart{ \Phi_n + \fracpart{z(n)} }^2.
 \end{align*}
@@
 \begin{equation}\label{eq:mthdiffformula}
-\Delta_h^r d(n) = \Delta_h^{r-1} d(n+h) - \Delta_h^{r-1} d(n) = \sum_{k=0}^{r}\binom{r}{k}(-1)^{r-k}d(n+kh)
+\begin{split}
+\Delta_h^r d(n) &= \Delta_h^{r-1} d(n+h) - \Delta_h^{r-1} d(n) \\
+&= \sum_{k=0}^{r}\binom{r}{k}(-1)^{r-k}d(n+kh)
+\end{split}
 \end{equation}
@@
 \begin{align*}
-V_N(\lambdabf_N) - \sigma^2 = \frac{1}{N}\sum_{n=1}^{N} g\left( \fracpart{z_N(n)} \right) \geq \frac{1}{N}\sum_{n\in K_N} \gamma = (1-c)\gamma,
+V_N(\lambdabf_N) - \sigma^2 &= \frac{1}{N}\sum_{n=1}^{N} g\left( \fracpart{z_N(n)} \right) \\
+&\geq \frac{1}{N}\sum_{n\in K_N} \gamma = (1-c)\gamma,
 \end{align*}
@@
  \begin{equation}\label{eq:zrbound}
-\left|\fracpart{  \Delta_h^r z_N(n')}\right| = \left|\fracpart{  \Delta_h^r\gamma_N(n') + \Delta_h^r z_r(n') }\right| = |\fracpart{ \epsilon + h^r r! \lambda_{r,N} }| < 2^r\delta,
+\begin{split}
+\left|\fracpart{  \Delta_h^r z_N(n')}\right| &= \left|\fracpart{  \Delta_h^r\gamma_N(n') + \Delta_h^r z_r(n') }\right| 
+\\ &= |\fracpart{ \epsilon + h^r r! \lambda_{r,N} }| < 2^r\delta,
+\end{split}
  \end{equation}
@@
 
-\section{Proof of asymptotic normality}\label{sec:centlimitproof}
-
-Let $\psibf$ be the vector with $k$th component $\psi_k = N^k \lambda_k$, $k=0, \dots, m$ and let 
-\[
-T_{N}(\psibf) = S_N(\lambdabf) = \frac{1}{N} \sum_{n=1}^{N} \fracpart{ \Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \psi_k }^2.
-\]
-Let $\widehat{\psibf}_N$ be the vector with elements $\widehat{\psi}_{k,N} = N^k \widehat{\lambda}_{k,N}$ so that $\widehat{\psibf}_N$ is the minimiser of $T_N$.
-Because each of $N^k \widehat{\lambda}_{k,N}$ converges almost surely to zero as $N \rightarrow \infty$, then $\widehat{\psibf}_{N}$ converges almost surely to $\zerobf$ as $N \rightarrow \infty$.
-We want to find the asymptotic distribution of
-\[
-\sqrt{N}\widehat{\psibf}_{N} = 
-\left[
-\begin{array}
-[c]{c}%
-\sqrt{N} \widehat{\psi}_{0,N} \\ \sqrt{N}\widehat{\psi}_{1,N}  \\ \vdots \\ \sqrt{N} \widehat{\psi}_{m,N}
-\end{array}
-\right]
-=
-\left[
-\begin{array}
-[c]{c}%
-\sqrt{N} \widehat{\lambda}_{0,N} \\ N\sqrt{N}\widehat{\lambda}_{1,N} \\ \vdots \\ N^m\sqrt{N} \widehat{\lambda}_{m,N}
-\end{array}
-\right].
-\]
-The proof is complicated by the fact that $T_N$ is not differentiable everywhere because $\fracpart{x}^2$ is not differentiable when $\fracpart{x} = \tfrac{1}{2}$.
-This precludes the use of `standard approaches' to proving asymptotic normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.
-However, we show in Lemma~\ref{lem:diffathatpsi} that all the partial derivatives $\frac{\partial T_N}{\partial \psi_\ell}$ for $\ell = 0, \dots, m$ exist, and are equal to zero, at the minimiser $\widehat{\psibf}_N$.
-Thus, putting
-\begin{equation}\label{eq:Wn}
-W_{n} = \round{\Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \widehat{\psi}_{k,N}},
-\end{equation}
-so that 
-\[
-T_{N}(\psibf) = \frac{1}{N} \sum_{n=1}^{N} \big( \Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \psi_k  - W_n \big)^2,
-\] 
-we have,
-\begin{align*}
-0  = \frac{\partial T_N}{\partial \psi_\ell}(\widehat{\psibf}_N) = \frac{2}{N}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\left( \Phi_n - W_{n}+ \sum_{k=0}^{m}(\tfrac{n}{N})^k \widehat{\psi}_{k,N}  \right),
-\end{align*}
-for each $\ell = 0, \dots, m$.  Now $D_{\ell,N} = K_{\ell,N}$, where $D_{\ell,N} = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}(\tfrac{n}{N})^\ell \Phi_n$, and
-\begin{equation}\label{eq:KellN}
-K_{\ell,N} = \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\left(W_{n}- \sum_{k=0}^{m}(\tfrac{n}{N})^k \widehat{\psi}_{k,N}  \right).
-\end{equation}
-Lemma~\ref{lem:Kconvfhalf} shows that,
-\begin{equation}\label{eq:KellNconv}
-K_{\ell,N} =  (h - 1) \sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big( C_{\ell k} + o_P(1) \big) + o_P(1),
-\end{equation}
-for all $\ell = 0, \dots m$, where $C_{\ell k} =  \tfrac{1}{\ell + k + 1}$, and $h = f(-\nicefrac{1}{2})$, and $o_P(1)$ denotes a random variable converging in probability to zero as $N\rightarrow\infty$.
-
-It is now convenient to write in vector form.  Let 
-\begin{equation}\label{eq:dnandkn}
-\kbf_N = 
-\left[
-\begin{array}{ccc}%
-K_{0,N}  & \cdots & K_{m,N}
-\end{array}
-\right]^\prime
-%\qquad \text{and} \qquad 
-=
-\dbf_N = \left[
-\begin{array}{ccc}%
-D_{0,N}  & \cdots & D_{m,N}
-\end{array}
-\right]^\prime.
-\end{equation}
-From~\eqref{eq:KellNconv},
-\[
-\dbf_N = \kbf_N = \sqrt{N} (h - 1)(\Cbf + o_P(1)) \widehat{\psibf}_N + o_P(1)
-\]
-where $o_P(1)$ here means a vector or matrix of the appropriate dimension with every element converging in probability to zero as $N\rightarrow\infty$.
-Thus $\sqrt{N}\widehat{\psibf}_N$ has the same asymptotic distribution as $(h - 1)^{-1}\Cbf^{-1} \dbf_N$.
-Lemma~\ref{eq:convdn} shows that $\dbf_N$ is asymptotically normally distributed with zero mean and covariance matrix $\sigma^2\Cbf$.
-Thus $\sqrt{N}\widehat{\psibf}_N$ is asymptotically normal with zero mean and covariance matrix
-\[
-\frac{\sigma^2\Cbf^{-1}\Cbf(\Cbf^{-1})^\prime}{(1 - h)^2} = 
-\frac{\sigma^2\Cbf^{-1}}{(1 - h)^2}.
-\] 
-It remains to prove Lemmas~\ref{lem:diffathatpsi},~\ref{lem:Kconvfhalf} and~\ref{eq:convdn}.
-
-\begin{lemma}\label{lem:diffathatpsi}
-For all $\ell = 0, \dots, m$ the partial derivatives $\frac{\partial T_N}{\partial \psi_\ell}$ exist, and are equal to zero, at the minimiser $\widehat{\psibf}_N$.
-That is $\frac{\partial T_N}{\partial \psi_\ell}(\widehat{\psibf}_N) = 0$ for each $\ell = 0, \dots, m$.
-\end{lemma}
-\begin{proof}
-The function $\fracpart{x}^2$ is differentiable everywhere except if $\fracpart{x} \neq -\frac{1}{2}$, and so $T_N$ is differentiable with respect to $\psibf$ at $\widehat{\psibf}_N$ if $\sfracpart{\Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \widehat{\psi}_{k,N}} \neq -\tfrac{1}{2}$ for all $n = 1, \dots, N$.
-This is proved in Lemma~\ref{lem:boundonhatlambda}.
-So the partial derivatives $\frac{\partial T_N}{\partial \psi_\ell}$ exist for all $\ell = 0, \dots, m$ at $\widehat{\psibf}_N$.
-That each of the partial derivatives is equal to zero at $\widehat{\psibf}_N$ follows since $\widehat{\psibf}_N$ is a minimiser of $T_N$.
-\end{proof}
-
-\begin{lemma}\label{lem:boundonhatlambda} $\sabs{\sfracpart{\Phi_n + \sum_{k=0}^m (\nicefrac{n}{N})^k \widehat{\psi}_{k,N}}} \leq \frac{1}{2} - \frac{1}{2N}$ for all $n = 1, \dots, N$.
-\end{lemma}
-\begin{proof}
-To simplify our notation let $B_n = \Phi_n + \sum_{k=1}^m (\nicefrac{n}{N})^k \widehat{\psi}_{k,N}$ so that we now require to prove $\abs{\fracpart{B_n + \widehat{\psi}_{0,N}}} \leq \frac{1}{2} - \frac{1}{2N}$ for all $n = 1, \dots, N$.
-From~\eqref{eq:Wn}, $W_n = \round{B_n + \widehat{\psi}_{0,N}}$, and
-\begin{align*}
-T_N(\widehat{\psibf}_N) = \frac{1}{N}\sum_{n=1}^N \fracpart{B_n + \widehat{\psi}_{0,N}}^2= \frac{1}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N} - W_n)^2.
-\end{align*}
-Since $\widehat{\psi}_{0,N}$ is the minimiser of the quadratic above,
-\begin{equation}\label{eq:lamsum}
-\widehat{\psi}_{0,N} = -\frac{1}{N}\sum_{n=1}^N(B_n - W_n).
-\end{equation}
-The proof now proceeds by contradiction.  Assume that for some $k$,
-\begin{equation}\label{eq:philamfraccontra}
-\fracpart{B_k + \widehat{\psi}_{0,N}} > \frac{1}{2} - \frac{1}{2N}.
-\end{equation}
-Let $F_n = W_n$ for all $n \neq k$ and $F_k = W_k + 1$, and let
-\[
-\phi = -\frac{1}{N}\sum_{n=1}^N(B_n - F_n) = \widehat{\psi}_{0,N} + \frac{1}{N}.
-\]
-Now,
-\begin{align}
-(B_k + \phi - F_k)^2 &= (B_k + \phi - W_k - 1)^2 \nonumber \\
-&= (B_k + \phi - W_k)^2 - 2(B_k + \phi - W_k) + 1 \nonumber \\
-&= (B_k + \phi - W_k)^2 - 2(B_k + \widehat{\psi}_{0,N} - W_k) + 1 - \frac{2}{N} \nonumber \\
-&= (B_k + \phi - W_k)^2 - 2\fracpart{B_k + \widehat{\psi}_{0,N}} + 1 - \frac{2}{N} \nonumber \\
-&< (B_k + \phi - W_k)^2 - \frac{1}{N}, \label{eq:Bkineq}
-\end{align}
-where the inequality in the last line follows from~\eqref{eq:philamfraccontra}. Let $\bbf = [ \phi,  \widehat{\psi}_{1,N} , \dots , \widehat{\psi}_{m,N}]$
-% \[
-% \bbf = \left[
-% \begin{array}
-% [c]{cccc}%
-% \phi & \widehat{\psi}_{1,N} & \cdots & \widehat{\psi}_{m,N}
-% \end{array}
-% \right]^\prime
-% \]
-be the vector of length $m+1$ with components $b_0 = \phi$ and $b_\ell = \widehat{\psi}_{\ell,N}$ for $\ell = 1 , \dots m$.  Now,
-\[
-N T_N(\bbf) = \sum_{n=1}^N \fracpart{B_n + \phi }^2 \leq  \sum_{n=1}^N (B_n + \phi  - F_n)^2, 
-\]
-and using the inequality from~\eqref{eq:Bkineq},
-\begin{align*}
-N T_N(\bbf) &< - \frac{1}{N} + \sum_{n=1}^N (B_n + \phi  - W_n)^2 \\
-&= - \frac{1}{N} + \sum_{n=1}^N (B_n + \widehat{\psi}_{0,N} + \frac{1}{N}  - W_n)^2 \\
-&= \sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n)^2 +  \frac{2}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n)\\
-%&= N T_N(\widehat{\psibf}_N) + \frac{2}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n) \\
-&= N T_N(\widehat{\psibf}_N),
-\end{align*}
-because $\frac{2}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n) = 0$ as a result of~\eqref{eq:lamsum}.
-But, now $T_N(\bbf) < T_N(\widehat{\psibf}_N)$ violating the fact that $\widehat{\psibf}_N$ is a minimiser of $T_N$.
-So~\eqref{eq:philamfraccontra} is false by contradiction.
-
-If $\fracpart{B_k + \widehat{\psi}_{0,N}} < -\frac{1}{2} + \frac{1}{2N}$ for some $k$, we set $F_k = W_k - 1$ and using the same procedure as before obtain $T_N(\bbf) < T_N(\widehat{\psibf}_N)$ again.
-The proof follows.
-\end{proof}
-
-\begin{lemma}\label{lem:Kconvfhalf}
-With $K_{\ell,N}$ defined in~\eqref{eq:KellN}, $h = f(-\nicefrac{1}{2})$, and $C_{\ell k} =  \frac{1}{\ell + k + 1}$ we have $K_{\ell,N} = (h - 1) \sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big( C_{\ell k} + o_P(1) \big) + o_P(1)$ for all $\ell = 0, \dots, m$.% is an element of the $m+1$ by $m+1$ Hilbert matrix, and $o_P(1)$ denotes a random variable converging in probability to zero as $N\rightarrow\infty$.
-\end{lemma}
-\begin{proof}
-Care must be taken since $\widehat{\psibf}_N$ depends on the sequence $\{ \Phi_n \}$.  For $n = 1, \dots, N$ and positive $N$, let
-\begin{equation}
-p_{nN}(\psibf) = \sum_{k=0}^{m}\left( \tfrac{n}{N}\right)^k \psi_k,
-\end{equation}
-and put $q_{n}(x) = \round{\Phi_n + x}$ and $Q(x) = \expect q_{n}(x) =  \expect q_{1}(x)$.  Let
-\begin{equation}\label{eq:GNdef}
-G_{\ell,N}(\psibf) = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}\left( \tfrac{n}{N}\right)^\ell  \big(q_n(p_{nN}(\psibf)) - Q(p_{nN}(\psibf)) \big),
-\end{equation}
-and put
-\begin{equation}\label{eq:hatpnN}
-\widehat{p}_{nN} = p_{nN}(\widehat{\psibf}_N) = \sum_{k=0}^m \left(\tfrac{n}{N}\right)^k \widehat{\psi}_{k,N}.
-\end{equation}
-Now $W_n$ from~\eqref{eq:Wn} can be written as $W_n = \round{\Phi_n + \widehat{p}_{nN}} = q_{n}(\widehat{p}_{nN})$ and $K_{\ell,N}$ from~\eqref{eq:KellN} can be written as
-\begin{align*}
-K_{\ell,N} &= \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\big( q_{n}(\widehat{p}_{nN}) - \widehat{p}_{nN}  \big) \\
-&= \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\big(q_{n}(\widehat{p}_{nN}) - \widehat{p}_{nN} + Q(\widehat{p}_{nN}) - Q(\widehat{p}_{nN}) \big) \\
-&= G_{\ell,N}(\widehat{\psibf}_N) + H_{\ell,N},
-\end{align*}
-where 
-\begin{equation}\label{eq:HellNdef}
-H_{\ell,N} = \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell \big( Q(\widehat{p}_{nN}) - \widehat{p}_{nN} \big).
-\end{equation}
-Lemma~\ref{lem:unifprobG} in the Appendix shows that for any $\delta >0$ and $\nu > 0$ there exists an $\epsilon > 0$ such that
-\[
-\prob\left\{ \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_{\ell,N}(\psibf) } > \delta   \right\} < \nu
-\]
-for all positive integers $N$ and all $\ell = 0,\dots,m$, where $\|\psibf\|_\infty = \sup_{k} \abs{\psi_k}$.  Since $\widehat{\psibf}_N$ converges almost surely to zero, it follows that 
-\[
-\lim_{N\rightarrow\infty}\prob\left\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} = 0
-\]
-for any $\epsilon > 0$, and therefore $\prob\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \} < \nu$ for all sufficiently large $N$.  Now
-\begin{align*}
-  \prob\left\{\abs{ G_{\ell,N}(\widehat{\psibf}_N) } > \delta \right\} &= \prob\left\{ \abs{G_{\ell,N}(\widehat{\psibf}_N)} > \delta \;, \; \|\widehat{\psibf}_N\|_{\infty} < \epsilon \right\} \\
-&\hspace{1cm} + \prob\left\{ \abs{G_{\ell,N}(\widehat{\psibf}_N)} > \delta  \;, \; \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} \\
-&\leq \prob\left\{  \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_{\ell,N}(\psibf) } > \delta \right\} + \prob\left\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} \\
-&\leq 2\nu
-\end{align*}
-for all sufficiently large $N$.
-Since $\nu$ and $\delta$ can be chosen arbitrarily small, it follows that $G_{\ell,N}(\widehat{\psibf}_N)$ converges in probability to zero as $N\rightarrow\infty$, and therefore $K_{\ell,N} = H_{\ell,N} + o_P(1)$.
-Lemma~\ref{lem:sumpNhilb} shows that
-\begin{align*}
-H_{\ell,N} =  (h-1)\sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big(C_{\ell k} + o_P(1)\big).
-\end{align*}
-%The proof follows.
-\end{proof}
-
-\begin{lemma}\label{lem:sumpNhilb}
-With $H_{\ell,N}$ defined in~\eqref{eq:HellNdef}, $h = f(-\nicefrac{1}{2})$, and  $C_{\ell k} = \tfrac{1}{\ell + k + 1}$ we have $H_{\ell,N} = (h-1)\sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big(C_{\ell k} + o_P(1)\big)$.
-\end{lemma}
-\begin{proof}
-If $\abs{x} < 1$, then
-\[
-q_n(x)  = \round{\Phi_n + x} = \begin{cases}
-1, & \Phi_n + x  \geq \nicefrac{1}{2} \\
--1, & \Phi_n + x  < -\nicefrac{1}{2} \\
-0, & \text{otherwise},
-\end{cases}
-\]
-and,
-\begin{align*}
-Q(x) = Eq_1(x) &=  \begin{cases}
-\int_{\nicefrac{1}{2} -x }^{\nicefrac{1}{2}}f(t)\,dt, &   x \geq 0 \\
--\int_{-\nicefrac{1}{2}}^{-\nicefrac{1}{2} - x }f(t)\,dt, &  x  < 0.
-\end{cases} 
-% \\
-% &= \begin{cases}
-% 1 - F(\nicefrac{1}{2} - x ), & x  \geq 0 \\
-% -F(-\nicefrac{1}{2} - x ), &  x < 0,
-% \end{cases} 
-\end{align*}
-Because $f(\fracpart{x})$ is continuous at $-\nicefrac{1}{2}$ it follows that $Q(x) = x \big( h + \zeta(x) \big)$
-where $\zeta(x)$ is a function that converges to zero as $x$ converges to zero.  
-%The equation above is still valid when $\abs{x} \geq 1$ by an appropriate choice of the function $\zeta(x)$.  
-Observe that $\abs{\widehat{p}_{nN}} \leq \sum_{k=0}^m\sabs{ \widehat{\psi}_{k,N}}$ and, since each of the $\widehat{\psi}_{k,N} \rightarrow 0$ almost surely as $N\rightarrow\infty$, it follows that $\widehat{p}_{nN} \rightarrow 0$ almost surely uniformly in $n = 1, \dots, N$ as $N\rightarrow\infty$.
-Thus, $\zeta(\widehat{p}_{nN}) \rightarrow 0$ almost surely (and therefore also in probability) uniformly in $n = 1, \dots, N$ as $N\rightarrow\infty$.
-Now,
-\begin{align*}
-Q(\widehat{p}_{nN}) - \widehat{p}_{nN} = \widehat{p}_{nN}\big( h - 1 + \zeta(\widehat{p}_{nN}) \big) = \widehat{p}_{nN}\big( h - 1 + o_P(1) \big), 
-\end{align*}
-and, using~\eqref{eq:HellNdef},
-\begin{align*}
-H_{\ell,N} &= \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell \widehat{p}_{nN}\big( h - 1 + o_P(1) \big)  \\
-&=  \frac{1}{\sqrt{N}} \sum_{n=1}^{N}(\tfrac{n}{N})^\ell \sum_{k=0}^m (\tfrac{n}{N})^k \widehat{\psi}_{k,N} \big( h - 1 + o_P(1) \big)   \\
-&= \sqrt{N}\sum_{k=0}^m \widehat{\psi}_{k,N} \frac{1}{N} \sum_{n=1}^{N} \frac{n^{\ell + k}}{N^{\ell + k + 1}}\big( h - 1 + o_P(1)  \big).
-\end{align*}
-The Riemann sum
-\[
-\frac{1}{N} \sum_{n=1}^{N} \frac{n^{\ell + k}}{N^{\ell + k + 1}} = \int_{0}^{1}x^{k+\ell+1}dx + o_P(1),
-\]
-and since the integral above evaluates to $C_{\ell k} = \frac{1}{k+\ell+1}$, we have
-\[
-H_{\ell,N} = (h - 1)\sqrt{N}\sum_{k=0}^m \widehat{\psi}_{k,N}\big( C_{\ell k} + o_P(1)  \big).C_{\ell k}
-\]
-% Put
-% \[
-% C_{\ell + k,N} = \sum_{n=1}^{N}\frac{n^{\ell + k}}{N^{\ell + k+1}}.
-% \]
-% It is well known (see for example~\cite[eq. 21]{Peleg1991_CRB_PPS_1991}) that,
-% \begin{equation}\label{eq:ClkNHilbert}
-% C_{\ell + k,N} = C_{\ell k} + O(N^{-1}).
-% \end{equation}
-% Since $\zeta(\widehat{p}_{nN}) = o_P(1)$ for all $n = 1, \dots, N$ then
-% \[
-% \sum_{n=1}^{N}\frac{n^{\ell + k}}{N^{\ell + k + 1}}\zeta(\widehat{p}_{nN}) 
-% \]
-% converges in probability to zero as $N\rightarrow\infty$.  So,
-% \begin{align*}
-% \sum_{n=1}^{N}\frac{n^{\ell + k}}{N^{\ell + k + 1}}\big( h - 1 +  \zeta(\widehat{p}_{nN}) \big) = (h-1)\big( C_{\ell k} + o_P(1) \big).
-% \end{align*}
-% The proof follows by substituting this into~\eqref{eq:HellNre}.
-\end{proof}
-
-\begin{lemma}\label{eq:convdn}
-The distribution of the vector $\dbf_N$, defined in~\eqref{eq:dnandkn}, converges to the multivariate normal with zero mean and covariance matrix $\sigma^2\Cbf$.
-\end{lemma}
-\begin{proof}
-For any constant vector $\alphabf$, let
-\[
-z_N = \alphabf^\prime \dbf_N = \frac{1}{\sqrt{N}} \sum_{n=1}^{N} \Phi_n \sum_{\ell=0}^{m}\alpha_\ell\left( \frac{n}{N} \right)^\ell.
-\]
-By Lypanov's central limit theorem $z_N$ is asymptotically normally distributed with zero mean and variance
-\[
-\lim_{N\rightarrow\infty} \sigma^2 \frac{1}{N} \sum_{n=1}^{N}\left( \sum_{\ell=0}^{m}\alpha_\ell\left( \frac{n}{N} \right)^\ell \right)^2 = \sigma^2 \alphabf^\prime \Cbf \alphabf.
-\]
-By the Cram\`{e}r-Wold theorem it follows that $\dbf_N$ is asymptotically normally distributed with zero mean and covariance $\sigma^2 \Cbf$. 
-% Recall that the $\ell$th element of $\dbf_N$ is
-% \[
-% D_{\ell,N} = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}(\tfrac{n}{N})^\ell \Phi_n.
-% \]
-% Since the $\Phi_1, \dots, \Phi_N$ are independent, each with variance $\sigma^2$, the distribution of $D_{\ell,N}$ converges to the normal with zero mean and variance
-% \[
-% \lim_{N\rightarrow\infty}\var D_{\ell,N} = \sigma^2 \lim_{N\rightarrow\infty} C_{2\ell,N} = \sigma^2 C_{\ell\ell}
-% \]
-% by Lypanov's central limit theorem and by using~\eqref{eq:ClkNHilbert}.  The covariance between $D_{\ell,N}$ and $D_{k,N}$ is
-% \[
-% \covar(D_{\ell,N}, D_{k,N}) = \sigma^2 C_{\ell+k,N} \rightarrow \sigma^2 C_{\ell k}
-% \]
-% as $N \rightarrow \infty$, again using~\eqref{eq:ClkNHilbert}.  The proof follows.
-\end{proof}
-
-\section{Simulations}\label{sec:simulations}
-
-This section describes the results of Monte-Carlo simulations with the least squares unwrapping (LSU) estimator.
-The sample sizes considered are $N = 10, 50, 200$ and the unknown amplitude is $\rho = 1$.
-The $X_1, \dots, X_N$ are pseudorandomly generated independent and identically distributed circularly symmetric complex Gaussian random variables with variance $\sigma_c^2$.
-The coefficients $\tilde{\mubf} = [\tilde{\mu}_0, \dots, \tilde{\mu}_m]$ are distributed uniformly randomly in the identifiable region $B$.
-The number of replications of each experiment is $T = 2000$ to obtain estimates $\widehat{\mubf}_1, \dots, \widehat{\mubf}_T$ and the corresponding dealiased errors $\widehat{\lambdabf}_t = \dealias(\widehat{\mubf}_T - \tilde{\mubf})$ are computed.
-The sample mean square error (MSE) of the $k$th coefficient is computed according to $\tfrac{1}{T}\sum_{t=1}^T \widehat{\lambda}_{k,t}^2$ where $\widehat{\lambda}_{k,t}$ is the $k$th element of $\widehat{\lambdabf}_t$.
-
-Figure~\ref{plot:polyest1} shows the sample MSEs obtained for a polynomial phase signal of order $m=3$.
-Results are displayed results for the zeroth and third order coefficients $\widehat{\mu}_1$ and $\widehat{\mu}_3$.
-The results for $\widehat{\mu}_1$ and $\widehat{\mu}_2$ lead to similar conclusions.
-When $N = 10$ and $50$ the LSU estimator can be computed exactly using a general purpose algorithm for finding nearest lattice points called the sphere~decoder~\cite{Pohst_sphere_decoder_1981,Agrell2002,Viterbo_sphere_decoder_1999}.
-This is displayed by the circles in the figures.
-When $N=200$ the sphere decoder is computationally intractable and we instead use an approximate nearest point algorithm called the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder}.
-This is displayed by the dots.
-For the purpose of comparison we have also plotted the results for the $K$-best method when $N = 10$ and $50$.
-The asymptotic variance predicted in Theorem~\ref{thm:asymp_proof} is displayed by the dashed line.
-Provided the noise variance is small enough (so that the `threshold' is avoided) the sample MSE of the LSU estimator is close to that predicted by Theorem~\ref{thm:asymp_proof}.
-The Cram\`{e}r-Rao lower bound for the variance of unbiased polynomial phase estimators in Gaussian noise is also plotted using the solid line~\cite{Peleg1991_CRB_PPS_1991}.
-When the noise variance is small the asymptotic variance of the LSU estimator is close to the Cram\`{e}r-Rao lower bound.
-
-%Note the gap between the CRB and the variance predicted by Theorem~\ref{thm:asymp_proof}.
-In practice this gap can be overcome using a numerical technique, such as Newton's method, with the LSU estimate as a starting point.
-We have not done this here in order to show the accuracy of Theorem~\ref{thm:asymp_proof}.
-The other noticeable gap is that between the threshold behaviour of the LSU estimator and the LSE.
-When $N=50$ the threshold \emph{kicks in} a little earlier for the LSE than it does for the LSU.
-We suspect that this gap will actually \emph{close} as $N$ increases.
-This hypothesis is supported by the fact that this behaviour is observed in the case of frequency estimation (when $m=1$)~\cite[Sec.
-9.5]{McKilliam2010thesis}\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.
-Unfortunately it is not feasible to test this hypothesis because of the computational complexity of the LSE.
-
-% \begin{figure}[p] 
- %   	\centering 
- %  		\includegraphics[width=\linewidth]{code/lseplot-1.mps} 
- %               %\includegraphics{code/lseplot-1.mps}
- %   		\caption{Performance of the LSE and LSU estimator for a polynomial phase signal of order two, i.e.
-$m=2$.
-(Top) Mean square error (MSE) in the second order parameter $\mu_2$ for $N=10,50$ and $200$ versus the variance $\sigma_c^2$ of the $X_1, \dots, X_N$.
-The $X_1, \dots, X_N$ are independent and identically distributed (i.i.d.) complex Gaussian random variables.
-(Bottom) MSE in second order parameter $\mu_2$ versus the variance $\sigma^2$ of the $\Phi_1, \dots, \Phi_N$.
-The $\Phi_1, \dots, \Phi_N$ are i.i.d.
-uniform random variables.}
- %   		\label{plot:lsetest} 
- %  \end{figure}
-
-\begin{figure}[p] 
-   	\centering 
-  		\includegraphics[width=\linewidth]{code/gaussianplot4-3.mps}
-                %\includegraphics{code/gaussianplot4-1.mps}
-  		\caption{Sample mean square error (MSE) of the least squares unwrapping estimator for $N=10,50$ and $200$ for a polynomial phase signal of order $m=3$.
-(Top) MSE of the frequency coefficient $\mu_0$.
-(Bottom) MSE of the cubic coefficient $\mu_3$.}
-  		\label{plot:polyest1} 
- \end{figure}
-
-%\begin{figure}[p] 
-%   	\centering 
-%  		\includegraphics[width=\linewidth]{code/gaussianplot4-2.mps}
-%                %\includegraphics{code/gaussianplot4-1.mps}
-%  		\caption{Sample mean square error (MSE) of the least squares unwrapping estimator for $N=10,50$ and $200$ for a polynomial phase signal of order $m=3$.
-(Top) MSE of the quadratic coefficient $\mu_2$.
-(Bottom) MSE of the cubic coefficient $\mu_3$.}
-%  		\label{plot:polyest2} 
-% \end{figure}
-
-% \section{Computational considerations}
-
-% When $m=0$, polynomial phase estimation is equivalent to the problem of \emph{direction estimation} from circular statistics, and the least squares estimator corresponds with the \emph{extrinsic mean} (or \emph{sample circular mean})~\cite{McKilliam_mean_dir_est_sq_arc_length2010,Fisher1993,Mardia_directional_statistics}.
-In this case the LSE can be computed efficiently in $O(N)$ by averaging $N$ complex numbers.
-When $m=1$, polynomial phase estimation is equivalent to single frequency estimation, and the least squares estimator can be computed efficiently in $O(N\log N)$ operations by maximising the \emph{periodogram} using the fast Fourier transform followed by an appropriate numerical optimisation technique~\cite{Quinn2008maximizing_the_periodogram,Rife1974,Quinn2001}.
-When $m=2$, the least squares estimator can be computed in a reasonable amount of time when $N$ is approximately less than $50$.
-The LSE is computationally intensive for any $N$ when $m > 2$.
-By comparison the exact LSU estimator implemented using the sphere decoder can be computed quickly for any $m$ with $N$ approximately less than $50$.
-In situations where $N$ is small, but high statistical accuracy is required, the sphere decoder is computationally a better choice than the LSE for polynomial phase signals of order greater than or equal to $2$.
-
-% Babai's nearest plane algorithm requires $O(N^2)$ operations and can be run in a reasonable amount of time for any $m$ and reasonably large $N$.
-It should be noted that Babai's nearest plane algorithm requires a Lov\`as reduced lattice basis~\cite{Lenstra1982}.
-This requires $O(N^4)$ operations to compute, but, the Lov\`as reduced basis can be computed once offline.
-It does not need to be computed each time the estimator is run.
-The $K$-best algorithm runs in a reasonable amount of time for any $m$ and when $N$ is less than about $300$.
-If the noise variance is small, then both the sphere decoder and the $K$-best algorithm are actually fast.
-This is because both of these estimators begin with the lattice point found by Babai's nearest plane algorithm.
-If this point is close to (or \emph{is}) the nearest point then both the sphere decoder and the $K$-best algorithm terminate quickly.
-This is in contrast to the least squares estimator that is slow regardless of the noise variance.
-
-% Both Kitchen's estimator and the DPT are fast to compute.
-Kitchen's estimator requires only $O(N)$ arithmetic operations and the DPT requires $O(N \log N)$ operations.
-However, these estimators are not as accurate as the LSE or the LSU.
-They also fail to operate correctly over the entire identifiable region of polynomial phase coefficients~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}~\cite[Ch.~10]{McKilliam2010thesis}.
+%\section{Discussion}
 
@@
 %Simulations are used to show that the LSU has near maximum likelihood performance in the case that the noise is additive white and Gaussian.
-%It is shown in~\cite[Sec~8.1]{McKilliam2010thesis}\cite{McKilliam2009asilomar_polyest_lattice} how the LSU estimator can be computed by finding a nearest lattice point in a particular lattice.  
-Polynomial time algorithms that compute the LSU estimator are described in~\cite{McKilliam2010thesis}, but these are slow algorithms in practice.
-A significant outstanding question is whether practically fast algorithms exist.
-Considering the excellent statistical performance (both theoretically and practically) of the LSU estimator, even fast \emph{approximate} algorithms are likely to prove useful for the estimation of polynomial phase signals.
-
+The LSU estimator can be computed by finding a nearest lattice point in a lattice~\cite{McKilliam2009asilomar_polyest_lattice}.
+%Here, we consider using the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder} to achieve this in reasonable time when the number of observations, $N$, is less than about 300.
+Polynomial time nearest point algorithms for these lattices exist~\cite[Sec 4.3]{McKilliam2010thesis}, but these algorithms are not fast in practice.
+Instead, we considered some general purpose algorithms, the \term{sphere decoder}, \term{Babai's nearest plane algorithm} and the \term{$K$-best algorithm}.
+The sphere decoder and $K$-best algorithm result in accurate estimators.
+
+The major outstanding question is whether faster nearest point algorithms exist for these specific lattices.
+Considering the excellent statistical performance (both theoretically and practically) of the LSU estimator, even fast \emph{approximate} nearest point algorithms are likely to prove useful for the estimation of polynomial phase signals.
+
+%\begin{figure}[p] 
+%   	\centering 
+%  		\includegraphics[width=\linewidth]{polyfig/polyphase4plot.4} 
+%  		\caption{MSE in the third order coefficient $\mu_3$ versus $\var{X_n} = \sigma^2$. The true coefficients are uniformly spread in the identifiable region.} 
+%  		\label{plot:polyest_projnorm_mu3} 
+% \end{figure}
+
+% \begin{figure}[p] 
+%  	\centering 
+% 		\includegraphics[width=\linewidth]{polyfig/polyphase4plot_unfair.4} 
+% 		\caption{MSE in the third order coefficient $\mu_3$ versus $\var{X_n} = \sigma_c^2$. The coefficient have been restricted for Kitchen's estimator and the DPT.} 
+% 		\label{plot:polyest_projnorm_mu3_unfair} 
+%\end{figure} 
+% 
+% \begin{figure}[p] 
+%  	\centering 
+% 		\includegraphics[width=\linewidth]{polyfig/polyphase4plot_fair.4} 
+% 		\caption{MSE in the third order coefficient $\mu_3$ versus $\var{X_n} = \sigma_c^2$. The DPT estimator runs at the higher sampling rate $\delta$ so that the volumes $V_{DPT}(\delta) = V_m$.} 
+% 		\label{plot:polyest_projnorm_mu3_fair} 
+%\end{figure}
+
+%\addtolength{\parskip}{-0.2cm} 
+%\bibliographystyle{../../bib/IEEEbib} 
+\small 
+\bibliography{bib}
+
+\normalsize
 \appendix
@@
 
-\section{A uniform law of large numbers} \label{app:uniform-law-large}
+\subsection{A uniform law of large numbers} \label{app:uniform-law-large}
 
@@
 \end{equation}
-almost surely as $N\rightarrow\infty$, where $V_N(\lambdabf) = \expect S_N(\lambdabf)$.
-We prove this result here.
-Put $D_N(\lambdabf) = S_N(\lambdabf) - V_N(\lambdabf)$.
-Now $\sum_{N=1}^\infty \prob \left\{ \sup_{\lambdabf \in B}\abs{ D_N(\lambdabf) } > \epsilon \right\} < \infty$ for any $\epsilon > 0$ by Lemma~\ref{lem:vn}, and~\eqref{eq:SNVNunifmlln2} follows from the Borel-Cantelli lemma.
+almost surely as $N\rightarrow\infty$, where $V_N(\lambdabf) = \expect S_N(\lambdabf)$.  We prove this result here.  Put
+\[
+D_N(\lambdabf) = S_N(\lambdabf) - V_N(\lambdabf).
+\]  
+Now, for any $\epsilon > 0$,
+\[
+\sum_{N=1}^\infty \prob \left\{ \sup_{\lambdabf \in B}\abs{ D_N(\lambdabf) } > \epsilon \right\} < \infty
+\]
+by Lemma~\ref{lem:vn}.
+So~\eqref{eq:SNVNunifmlln2} follows from the Borel-Cantelli lemma.
 In what follows we use order notation in the standard way, that is, for functions $h$ and $g$, we write $h(N) = O(g(N))$ to mean that there exists a constant $K > 0$ and a finite $N_0$ such that $h(N) \leq K g(N)$ for all $N > N_0$.
 
-\begin{lemma} \label{lem:vn} $\prob \left\{ \sup_{\lambdabf \in B}\abs{ D_N(\lambdabf) } > \epsilon \right\} = O(e^{-c \epsilon^2 N})$ for any $\epsilon > 0$ and $c < 2$.
+\begin{lemma} \label{lem:vn} 
+For any $\epsilon > 0$ and $c < 2$,
+\[
+\prob \left\{ \sup_{\lambdabf \in B}\abs{ D_N(\lambdabf) } > \epsilon \right\} = O(e^{-c \epsilon^2 N}).
+\]
  \end{lemma}
@@
 \begin{align}
- \sup_{\lambdabf \in B}\sabs{ D_N(\lambdabf) }  &= \sup_{\rbf \in G}\sup_{ \lambdabf \in B[\rbf]}\sabs{ D_N(\lambdabf[\rbf]) + D_N(\lambdabf) - D_N(\lambdabf[\rbf])  } \nonumber \\
+ &\sup_{\lambdabf \in B}\sabs{ D_N(\lambdabf) }  \nonumber \\
+&= \sup_{\rbf \in G}\sup_{ \lambdabf \in B[\rbf]}\sabs{ D_N(\lambdabf[\rbf]) + D_N(\lambdabf) - D_N(\lambdabf[\rbf])  } \nonumber \\
  &\leq \sup_{\rbf \in G}\sabs{ D_N(\lambdabf[\rbf])} + \sup_{\rbf \in G}\sup_{\lambdabf\in B[\rbf]}\sabs{ D_N(\lambdabf) - D_N(\lambdabf[\rbf])  }. \label{eq:gridanddensespace}
  \end{align}
-From Lemma~\ref{lem:supVjk} it will follow that $\prob\left\{   \sup_{\rbf \in G }\abs{ D_N(\lambdabf[\rbf])  } > \frac{\epsilon}{2} \right\} = O(e^{-c \epsilon^2 N})$
+From Lemma~\ref{lem:supVjk} it will follow that 
+\[
+\prob\left\{   \sup_{\rbf \in G }\abs{ D_N(\lambdabf[\rbf])  } > \frac{\epsilon}{2} \right\} = O(e^{-c \epsilon^2 N})
+\]
 for any $\epsilon > 0$ and $c < 2$.  In Lemma~\ref{lem:supBVn} we show that
@@
 \begin{lemma}\label{lem:supVjk}
-$\prob\cubr{   \sup_{\rbf \in G}\abs{ D_N(\lambdabf[\rbf])  } > \epsilon } = O(e^{-c \epsilon^2 N})$ for any $\epsilon > 0$ and $c < 8$.
+For any $\epsilon > 0$ and $c < 8$,
+\[
+\prob\cubr{   \sup_{\rbf \in G}\abs{ D_N(\lambdabf[\rbf])  } > \epsilon } = O(e^{-c \epsilon^2 N}).
+\]
 \end{lemma}
  \begin{proof}
-Fix $\lambdabf$ and write $D_N(\lambdabf) = \bar{Z} = \frac{1}{N}  \sum_{n=1}^{N}Z_{n}$, where
+Fix $\lambdabf$ and write 
+\[
+D_N(\lambdabf) = \bar{Z} = \frac{1}{N}  \sum_{n=1}^{N}Z_{n},
+\] 
+where
  \[
@@
  \]
- are independent with zero mean and $\abs{Z_n} \leq \tfrac{1}{4}$.
-It follows from Hoeffding's inequality~\cite{Hoeffding_inequality_1963} that,  $\prob\{ \abs{D_N(\lambdabf)} > \epsilon \} \leq 2e^{ - 8 \epsilon^2 N}$, and so,
+ are independent with zero mean and $\abs{Z_n} \leq \tfrac{1}{4}$. It follows from Hoeffding's inequality~\cite{Hoeffding_inequality_1963} that,  
+\[
+\prob\{ \abs{D_N(\lambdabf)} > \epsilon \} \leq 2e^{ - 8 \epsilon^2 N},
+\] 
+and so,
  \begin{align*}
@@
 
-\begin{lemma}\label{lem:boundedsquarefracparts}
-$\fracpart{x}^2 - |\delta| \leq \fracpart{x + \delta}^2 \leq \fracpart{x}^2 + |\delta|$ for all $x, \delta \in \reals$.
+\begin{lemma}\label{lem:boundedsquarefracparts} For real numbers $x$ and $delta$,
+\[
+\fracpart{x}^2 - |\delta| \leq \fracpart{x + \delta}^2 \leq \fracpart{x}^2 + |\delta|.
+\]
 \end{lemma}
@@
 
- \begin{lemma}\label{lem:supBVn} $\sup_{\rbf\in G}\sup_{ \lambdabf  \in B[\rbf] }\left\vert D_N\left(  \lambdabf \right) - D_N\left(  \lambdabf[\rbf] \right)\right\vert < 2\frac{m+1}{N^b}$ for all $N$.
+ \begin{lemma}\label{lem:supBVn} 
+For all positive integers $N$,
+\[
+\sup_{\rbf\in G}\sup_{ \lambdabf  \in B[\rbf] }\left\vert D_N\left(  \lambdabf \right) - D_N\left(  \lambdabf[\rbf] \right)\right\vert < 2\frac{m+1}{N^b}.
+\]
  \end{lemma}
@@
 %The proof we give is not just almost surely but \emph{surely} as the convergence will be shown to occur irrespective of the values of the $\Phi_n$. 
-Put $b_n = \Phi_n + \sum_{k = 0}^{m}{\lambda_k n^k}$ and $a_n = \Phi_n + \sum_{k = 0}^{m}{\lambda[\rbf]_k n^k}$, where $\lambda[\rbf]_k$ denotes the $k$th element of the grid point $\lambdabf[\rbf]$.
-For $\lambdabf \in B[\rbf]$ we have $b_n = a_n + \delta_n$,
-where $\sabs{\delta_n} \leq \sum_{k=0}^m\frac{n^k}{k!
-N^{b+k}} \leq \frac{m+1}{N^{b}}$.
-From Lemma~\ref{lem:boundedsquarefracparts} it follows that $-|\delta_n| \leq \fracpart{x + b_n}^2 -  \fracpart{x + a_n}^2  \leq |\delta_n|$,
-and consequently $|\fracpart{x + b_n}^2 -  \fracpart{x + a_n}^2| \leq \tfrac{m+1}{N^b}$ for all $x \in \reals$. Now
+Put 
+\[
+b_n = \Phi_n + \sum_{k = 0}^{m}{\lambda_k n^k} \;\;\; \text{and}  \;\;\; a_n = \Phi_n + \sum_{k = 0}^{m}{\lambda[\rbf]_k n^k},
+\]
+where $\lambda[\rbf]_k$ denotes the $k$th element of the grid point $\lambdabf[\rbf]$. For $\lambdabf \in B[\rbf]$ we have $b_n = a_n + \delta_n$,
+where 
+\[
+\sabs{\delta_n} \leq \sum_{k=0}^m\frac{n^k}{k! N^{b+k}} \leq \frac{m+1}{N^{b}}.
+\]
+From Lemma~\ref{lem:boundedsquarefracparts} it follows that 
+\[
+-|\delta_n| \leq \fracpart{x + b_n}^2 -  \fracpart{x + a_n}^2  \leq |\delta_n|,
+\]
+and consequently 
+\[
+|\fracpart{x + b_n}^2 -  \fracpart{x + a_n}^2| \leq \tfrac{m+1}{N^b}
+\] 
+for all $x \in \reals$. Now
 \[
@@
 \]
-and therefore $\sabs{S_N(\lambdabf) - S_N(\lambdabf[\rbf])} \leq \frac{m+1}{N^b}$ for all $\lambdabf \in B[\rbf]$.  As this bound is independent of $\Phi_1 \dots \Phi_N$, we have
+and therefore 
+\[
+\sabs{S_N(\lambdabf) - S_N(\lambdabf[\rbf])} \leq \frac{m+1}{N^b}
+\] 
+for all $\lambdabf \in B[\rbf]$.  As this bound is independent of $\Phi_1 \dots \Phi_N$, we have
 \[
@@
 \]
-by Jensen's inequality.  Therefore, for all $\lambdabf \in B[\rbf]$,
+by Jensen's inequality.  Therefore,
  \begin{align*}
- \sabs{ D_N(\lambdabf) - D_N(\lambdabf[\rbf]) } &= \sabs{ S_N(\lambdabf) - S_N(\lambdabf[\rbf]) + V_N(\lambdabf) - V_N(\lambdabf[\rbf]) } \\
+ &\sabs{ D_N(\lambdabf) - D_N(\lambdabf[\rbf]) } \\
+&= \sabs{ S_N(\lambdabf) - S_N(\lambdabf[\rbf]) + V_N(\lambdabf) - V_N(\lambdabf[\rbf]) } \\
 &\leq \sabs{ S_N(\lambdabf) - S_N(\lambdabf[\rbf]) \vert + \vert V_N(\lambdabf) - V_N(\lambdabf[\rbf]) } \\
@@
 \end{align*}
-and the lemma follows because this bound is independent of $\rbf$.
+for all $\lambdabf \in B[\rbf]$.  The lemma follows because this bound is independent of $\rbf$.
 %\[
@@
 
-\section{A tightness result}\label{app:tightness-result}
-
-During the proof of asymptotic normality in Lemma~\ref{lem:Kconvfhalf} we made use of the following result regarding the function,
-\[
-G_{\ell,N}(\psibf) = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}\left( \tfrac{n}{N}\right)^\ell  \big(q_n(p_{nN}(\psibf)) - Q(p_{nN}(\psibf)) \big),
-\]
-where the functions $q_n$, $Q$ and $p_{nN}$ are defined above~\eqref{eq:GNdef} and $\ell \in \{0, 1, \dots, m\}$.
-To simplify notation we drop the subscript $\ell$ and write $G_{\ell,N}$ as $G_N$ in what follows.
-The proof we will give holds for any nonnegative integers $\ell$.
-
-\begin{lemma}\label{lem:unifprobG}
-For any $\delta >0$ and $\nu > 0$ there exists $\epsilon > 0$ such that
-\[
-\prob\left\{ \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf) } > \delta   \right\} < \nu %\qquad \text{for all $N > N_0$}
-\]
-for all positive integers $N$.
-\end{lemma}
-
-This result is related to what is called \emph{tightness} in the literature on empirical processes and weak convergence on metric spaces~\cite{Billingsley1999_convergence_of_probability_measures,Dudley_unif_central_lim_th_1999,Shorak_emp_proc_stat_2009}.
-The lemma is different from what is usually proved in the literature because the function $p_{nN}(\psibf) = \sum_{k=0}^m\left( \tfrac{n}{N}\right)^k\psi_k$
-depends on $n$.
-Nevertheless, the methods of proof from the literature can be used if we include a known result about hyperplane arrangements~\cite[Ch.
-5]{Chazelle_discrepency_method_2000}\cite[Ch.
-6]{Matousek_lect_disc_geom_2002}.
-Our proof is based on a technique called \emph{symmetrisation} and another technique called \emph{chaining} (also known as \emph{bracketing})~\cite{Pollard_asymp_empi_proc_1989}.
-
-\begin{proof}
-Define the function 
-\begin{align*}
-f_{nN}(\psibf, \Phi_n) = \left( \tfrac{n}{N}\right)^\ell q_n(p_{nN}(\psibf)) = \left( \tfrac{n}{N}\right)^\ell \round{\Phi_n + \sum_{k=0}^m\left( \tfrac{n}{N}\right)^k\psi_k}
-\end{align*}
-%and observe that 
-%\[
-%\expect f_{nN}(\psibf, \Phi_n) = \expect f_{nN}(\psibf, \Phi_1) = \left( \tfrac{n}{N}\right)^\ell Q(p_{nN}(\psibf))
-%\]
-so that $G_N$ can be written as
-\[
-G_N(\psibf) = \frac{1}{\sqrt{N}} \sum_{n=1}^{N} \big( f_{nN}(\psibf, \Phi_n) - \expect f_{nN}(\psibf, \Phi_n) \big).
-\]
-Let $\{g_n\}$ be a sequence of independent standard normal random variables, independent of the phase noise sequence $\{\Phi_n\}$.
-The symmetrisation argument~\cite[Sec.~4]{Pollard_asymp_empi_proc_1989}\cite{Gine_Zinn_symmetrisation_1984} can be used to show that
-%Lemma~\ref{lem:symmetrisation} shows that
-\[
-\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} \leq \sqrt{2\pi} \; \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ Z_N(\psibf) },
-\]
-where 
-\begin{equation}\label{eq:ZpsiCondGaussProc}
-Z_N(\psibf) = \frac{1}{\sqrt{N}} \sum_{n=1}^{N} g_n f_{nN}(\psibf, \Phi_n),
-\end{equation}
-and where $\expect$ runs over both $\{g_n\}$ and $\{\Phi_n\}$.
-Conditionally on $\{\Phi_n\}$, the process $\{Z_N(\psibf), \psibf \in \reals^{m+1} \}$ is a \emph{Gaussian process}, and numerous techniques exist for its analysis.
-Lemma~\ref{lem:chaining} shows that for any $\kappa > 0$ there exists an $\epsilon > 0$ such that $\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ Z_N(\psibf) } < \kappa.$  Thus $\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)}  <  \sqrt{2\pi} \; \kappa$, and by Markov's inequality,
-\[
-\prob \cubr{  \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} > \delta } \leq  \sqrt{2\pi} \; \frac{\kappa}{\delta},
-\]
-for any $\delta > 0$.  The proof follows with $\nu =  \sqrt{2\pi} \kappa/\delta$.  It remains to prove Lemma~\ref{lem:chaining}.
-%Lemmas~\ref{lem:symmetrisation}~and~\ref{lem:chaining}.
-
-\end{proof}
-
-% The proof of the next lemma is based on a technique called \emph{symmetrisation}~\cite{Gine_Zinn_symmetrisation_1984} and the proof we give closely follows Pollard~\cite[Sec.~4]{Pollard_asymp_empi_proc_1989}.
-
-% \begin{lemma} \label{lem:symmetrisation}(Symmetrisation)
-% \[
-% \expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} \leq \sqrt{2\pi} \; \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ Z_N(\psibf) }.
-% \]
-% \end{lemma}
-% \begin{proof}
-% Let $\{\Phi_n^\prime\}$ be a sequence of random variables distributed identically to $\{\Phi_n\}$ and independent of both $\{\Phi_n\}$ and $\{g_n\}$.
-Write $\expect_\Phi$ to denote expectation conditional on $\{\Phi_n\}$.
-%That is, expectation taken over $\{g_n\}$ and $\{\Phi_n^\prime\}$ treating $\{\Phi_n\}$ as fixed.
-% Since $\Phi_n$ and $\Phi_n^\prime$ are identically distributed $\expect f_{nN}(\psibf, \Phi_n) = \expect_\Phi f_{nN}(\psibf, \Phi_n^\prime)$ so that,
-% \begin{align*}
-% G_N(\psibf) &= \frac{1}{\sqrt{N}} \sum_{n=1}^{N} \big( f_{nN}(\psibf, \Phi_n) - \expect_\Phi f_{nN}(\psibf, \Phi_n^\prime) \big)  \\
-% &= \expect_\Phi \frac{1}{\sqrt{N}} \sum_{n=1}^{N} ( f_{nN} - f_{nN}^\prime ),
-% \end{align*}
-% where we put $f_{nN} = f_{nN}(\psibf, \Phi_n)$ and  $f_{nN}^\prime = f_{nN}(\psibf, \Phi_n^\prime)$  for notational convenience.
-% Taking absolute values followed by supremums,
-% \begin{align}
-%  \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} &= \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{\frac{1}{\sqrt{N}} \expect_\Phi \sum_{n=1}^{N} (f_{nN} - f_{nN}^\prime )} \nonumber \\
-% &\leq \sup_{\|\psibf\|_{\infty} < \epsilon} \expect_\Phi  \abs{ \frac{1}{\sqrt{N}} \sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) } \nonumber  \\
-% &\leq \expect_\Phi \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} ( f_{nN} - f_{nN}^\prime ) }, \label{eq:Eoutsidesup}
-% \end{align}
-% the second line following from Jensen's inequality and the third line since $\sup \expect \abs{\dots} \leq \expect \sup \abs{\dots}$.  Applying $\expect$ to both sides gives the inequality
-% \begin{equation}\label{eq:Gninequfin}
-%  \expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} \leq  \expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) },
-% \end{equation}
-% because $\expect \expect_\Phi$ is equivalent to $\expect$.
-Let $\sigma_n = \frac{g_n}{\abs{g_n}}$ and put $\sigma_n = 1$ if $g_n = 0$.
-The $\sigma_n$ are independent with $\prob\{ \sigma_n = -1 \} = \tfrac{1}{2}$ and $\prob\{ \sigma_n = 1 \} = \tfrac{1}{2}$.
-The symmetry of the distribution of $g_n$ implies that $g_n$ and $\sigma_n$ are independent.
-As $\Phi_n$ and $\Phi_n^\prime$ are independent and identically distributed, the random variable $f_{nN} - f_{nN}^\prime$ is symmetrically distributed about zero, and is therefore distributed identically to $\sigma_n( f_{nN} - f_{nN}^\prime)$.
-Thus,
-% %\[
-% %\sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) \qquad \text{and} \qquad \sum_{n=1}^{N}  \sigma_n(f_{nN} - f_{nN}^\prime)
-% %\]
-% %are identically distributed, and therefore
-% \begin{align*}
-%  \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{  \frac{1}{\sqrt{N}}\sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) } = \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{  \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n ( f_{nN} - f_{nN}^\prime) }.
-% \end{align*}
-% By the triangle inequality
-% \[
-% \abs{ \sum_{n=1}^{N} \sigma_n (f_{nN} - f_{nN}^\prime)} \leq \abs{ \sum_{n=1}^{N} \sigma_n f_{nN} } + \abs{ \sum_{n=1}^{N}  \sigma_n f_{nN}^\prime },
-% \]
-% and it follows from~\eqref{eq:Gninequfin}, that 
-% \begin{align*}
-%  \expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} &\leq \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} } + \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{\frac{1}{\sqrt{N}} \sum_{n=1}^{N}  \sigma_n  f_{nN}^\prime } \\
-% &= 2 \; \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} }.
-% \end{align*}
-% Write $\expect_{\Phi\sigma}$ to denote expectation conditional on $\{\Phi_n\}$ and $\{\sigma_n\}$.
-Since $g_n$ is a standard normal random variable, $\expect \abs{g_n} = \expect_{\Phi\sigma} \abs{g_n} = \sqrt{2/\pi}$, and
-% \begin{align*}
-% \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} } &= \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} \sqrt{\frac{\pi}{2}}\expect_{\Phi\sigma} \abs{g_n}} \\
-% &\leq \sqrt{\frac{\pi}{2}} \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{ \sqrt{N}}\sum_{n=1}^{N} \sigma_n \abs{g_n}  f_{nN}},
-% \end{align*}
-% the last line following from Jensen's inequality and by moving $\expect_{\Phi\sigma}$ outside the supremum similarly to~\eqref{eq:Eoutsidesup}. As $g_n = \sigma_n \abs{g_n}$, it follows that
-% \begin{align*}
-% \expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{G_N(\psibf)}  &\leq 2\sqrt{\frac{\pi}{2}} \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} g_n  f_{nN}} \\
-% &= \sqrt{2\pi} \; \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ Z_N(\psibf) } .
-% \end{align*}
-% \end{proof}
-
-\begin{lemma} \label{lem:chaining}
-For any $\kappa > 0$ there exists $\epsilon > 0$ such that
-\[
-\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ Z_N(\psibf) } < \kappa.
-\]
-%where $Z_N(\psibf)$ is defined by~\eqref{eq:ZpsiCondGaussProc}.
-\end{lemma}
-\begin{proof}
-Without loss of generality, assume that $\epsilon < \frac{1}{m+1}$.  Lemma~\ref{lem:chaining2} shows that
-\begin{equation}\label{eq:supZCK1} 
-\expect_\Phi \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ Z_N(\psibf) } \leq K_1 \sqrt{C_\epsilon(\{\Phi_n\})},
-\end{equation}
-where $K_1$ is a finite, positive constant, and $C_\epsilon(\{\Phi_n\})$ is the average number of times $\abs{\Phi_1}, \dots, \abs{\Phi_N}$ is greater than or equal to $\nicefrac{1}{2} - (m+1)\epsilon$.
-That is,
-\begin{equation}\label{eq:Cedefn}
-C_\epsilon(\{\Phi_n\}) = \frac{1}{N} \sum_{n=1}^{N} I_\epsilon(\abs{\Phi_n}),
-\end{equation}
-where $I_\epsilon(\abs{\Phi_n})$ is 1 when $\abs{\Phi_n} \geq \nicefrac{1}{2} - (m+1)\epsilon$ and zero otherwise.
-Recall that $f$ is the probability density function of $\Phi_n$, and (by assumption in Theorem~\ref{thm:asymp_proof}) that $f(\fracpart{x})$ is continuous at $x = -\nicefrac{1}{2}$.
-Because of this, the expected value of $C_\epsilon(\{\Phi_n\})$ is small when $\epsilon$ is small, since
-\begin{align*}
-\expect C_\epsilon(\{\Phi_n\}) &= \frac{1}{N} \sum_{n=1}^{N} \expect I_\epsilon(\abs{\Phi_n}) \\
-&= \prob\cubr{\abs{\Phi_1} \geq \nicefrac{1}{2} - (m+1)\epsilon} \\
-&= \int_{-1/2}^{-1/2 + (m+1)\epsilon} f(\phi) d\phi + \int_{1/2 - (m+1)\epsilon}^{1/2} f(\phi) d\phi \\
-&= \int_{-1/2 + (m+1)\epsilon}^{1/2 - (m+1)\epsilon} f(\fracpart{\phi}) d\phi \\
-&= 2(m+1)\epsilon \big(f(-\nicefrac{1}{2}) + o(1)\big),
-\end{align*}
-where $o(1)$ goes to zero as $\epsilon$ goes to zero.
-Since $\sqrt{\cdot}$ is a concave function on the positive real line and $C_{\epsilon}(\{\Phi_n\})$ is nonnegative, it follows from Jensen's inequality that $\expect \sqrt{C_\epsilon(\{\Phi_n\})} \leq  \sqrt{\expect  C_\epsilon(\{\Phi_n\} )} < \sqrt{K_2 \epsilon }$ for some constant $K_2$.
-Applying $\expect$ to both sides of~\eqref{eq:supZCK1} gives
-\[
-\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ Z_N(\psibf) } \leq K_1 \sqrt{\expect C_\epsilon(\{\Phi_n\})} < K_1 \sqrt{K_2 \epsilon}.
-\]
-Choosing $\epsilon = \kappa^2/(K_1^2 K_2)$ completes the proof.  It remains to prove Lemma~\ref{lem:chaining2}.
-\end{proof}
-
-The proofs of Lemmas~\ref{lem:chaining2}~and~\ref{lem:chaining3} are based on a technique called \emph{chaining} (or \emph{bracketing})~\cite{Dudley_unif_central_lim_th_1999,Ossiander_clt_bracketing_1984,Pollard_asymp_empi_proc_1989,Pollard_new_ways_clts_1986}.
-The proofs here follow those of Pollard~\cite{Pollard_asymp_empi_proc_1989}.
-In the remaining lemmas we consider expectation conditional on $\{\Phi_n\}$ and treat $\{\Phi_n\}$ as a fixed realisation.
-We consequently use the abbreviations $C_\epsilon = C_\epsilon(\{\Phi_n\})$ and $f_{nN}(\psibf) = f_{nN}(\psibf, \Phi_n)$.
-As in Lemma~\ref{lem:chaining} we assume, without loss of generality, that $\epsilon < \tfrac{1}{m+1}$.
-%This comes at no loss since Lemma~\ref{lem:chaining} only asks for the existence of an $\epsilon > 0$.
-
-\begin{lemma}\label{lem:chaining2}
-There exists a positive constant $K_1$ such that
-\[
-\expect_\Phi \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ Z_N(\psibf) } \leq K_1 \sqrt{C_\epsilon}.
-\]
-\end{lemma}
-\begin{proof}  Let $B_\epsilon = \{\xbf \in \reals \mid \|\xbf\|_\infty < \epsilon \}$.
-%We are now required to prove that
-%\[
-%\expect_\Phi \sup_{\psibf \in B_\epsilon} \abs{ Z_N(\psibf) } \leq K_1 \sqrt{C_\epsilon}.
-%\]
-For each non negative integer $k$, let $T_\epsilon(k)$ be a discrete subset of $\reals^{m+1}$ with the property that for every $\psibf \in B_\epsilon$ there exists some $\psibf^* \in T_\epsilon(k)$ such that the pseudometric
-\[
-d(\psibf, \psibf^*) = \sum_{n = 1}^N \big( f_{nN}(\psibf) - f_{nN}(\psibf^*) \big)^2 \leq 2^{-k} C_\epsilon N.
-\]
-We define $T_\epsilon(0)$ to contain a single point, the origin $\zerobf$.
-Defined this way $T_\epsilon(0)$ satisfies the inequality above because $d(\psibf, \zerobf) = \sum_{n = 1}^N f_{nN}(\psibf)^2  \leq C_\epsilon N$ for all $\psibf \in B_\epsilon$, as a result of Lemma~\ref{lem:epslmlemma}.
-
-The existence of $T_\epsilon(k)$ for each positive integer $k$ will be proved in Lemma~\ref{lem:metricentropy}.
-It is worth giving some intuition regarding $T_\epsilon(k)$.
-If we place a `ball' of radius $2^{-k} C_\epsilon N$  with respect to the pseudometric $d(\cdot, \cdot)$ around each point in $T_\epsilon(k)$, then, by definition, the union of these balls is a superset of $B_\epsilon$.
-The balls are said to \emph{cover} $B_\epsilon$ and $T_\epsilon(k)$ is said to form a \emph{covering} of $B_\epsilon$~\cite[Section 1.2]{Dudley_unif_central_lim_th_1999}.
-The minimum number of such balls required to cover $B_\epsilon$ is called a \emph{covering number} of $B_\epsilon$.
-In Lemma~\ref{lem:metricentropy} we show that no more than $K_3 2^{(m+1)k}$ balls of radius $2^{-k} C_\epsilon N$ are required to cover $B_\epsilon$, that is $\abs{T_\epsilon(k)} \leq K_3 2^{(m+1)k}$, where $K_3$ is a constant, independent of $N$ and $\epsilon$.
-%We make use of this upper bound now.
-%The lemma also upper bounds the number of elements by $\abs{T_\epsilon(k)} < K_3 2^{(m+1)k}$ where $K_3$ is a constant independent of $N$ and $\epsilon$.
-
-Since $f_{nN}(\psibf) = \left( \tfrac{n}{N}\right)^\ell \round{\Phi_n + p_{nN}(\psibf)}$ is a multiple of $N^{-\ell}$ for all $\psibf \in \reals^{m+1}$, it follows that $d(\psibf, \psibf^*)$ is a multiple of $N^{-2\ell}$.
-When $2^k > C_\epsilon N^{1+2\ell}$ we have $0 \leq d(\psibf, \psibf^*) < 2^{-k} C_\epsilon N < N^{-2\ell}$, and so $d(\psibf, \psibf^*) = 0$, and consequently $f_{nN}(\psibf) = f_{nN}(\psibf^*)$  for every $n = 1, \dots, N$ and $Z_N(\psibf) = Z_N(\psibf^*)$.
-Thus,
-\[
-\sup_{\|\psibf\|_{\infty} < \epsilon } \abs{ Z_N(\psibf) } = \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) }
-\]
-for all $k$ large enough that $2^{k} > C_\epsilon N^{1+2\ell}$.
-So, to analyse the supremum of $Z_N(\psibf)$ over the continuous interval $B_\epsilon$ it is enough to analyse the supremum over the discrete set $T_\epsilon(k)$ for large $k$.
-% The chaining argument~\cite[Sec.~3]{Pollard_asymp_empi_proc_1989} can  be used to show that
-% \[
-% \expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) } < \expect_{\Phi} \sup_{\psibf \in T_\epsilon(k-1)} \sabs{ Z_N(\psibf) }  + \sqrt{C_\epsilon}\frac{6 \sqrt{\log\sabs{T_\epsilon(k)}}}{2^{k/2}},
-% \]
-% which involves a recursion in $k$, where $\log(\cdot)$ is the natural logarithm.
-Since $T_\epsilon(0)$ contains only the origin, $\expect_{\Phi}\sup_{\psibf \in T_\epsilon(0)} \abs{ Z_N(\psibf) } = \expect_{\Phi}\abs{ Z_N(0) } = 0$ and, by unravelling the recursion, we obtain
-% \[
-% \expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) } < \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{6\sqrt{\log\abs{T_\epsilon(i)}}}{2^{i/2}}.
-% \]
-% Now $\log\abs{T_\epsilon(i)} \leq \log(K_3) + i(m+1)\log(2)$ by Lemma~\ref{lem:metricentropy}, and so the sum above converges as $k\rightarrow\infty$.
-The lemma holds with $K_1 =  \sum_{i=1}^{\infty}\frac{6\sqrt{\log\abs{T_\epsilon(i)}}}{2^{i/2}}$.
-% as required.
-Lemma~\ref{lem:chaining3} shows that 
-\[
-\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) } \leq \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{\sqrt{ i A_1 + A_2}}{2^{i/2}} < \infty
-\]
-for every positive integer $k$, where $A_1 = 18(m+1)\log 2$ and $A_2 = 18\log K_3$ are constants and $\log(\cdot)$ is the natural logarithm.
-The lemma holds with $K_1 =  \sum_{i=1}^{\infty}2^{-i/2}\sqrt{ i A_1 + A_2 }$.
-\end{proof}
-
-\begin{lemma}\label{lem:epslmlemma}
-For $\epsilon < \frac{1}{m+1}$ and all $\psibf \in B_\epsilon$ and $n = 1, \dots, N$,
-\[
-f_{nN}(\psibf)^2 \leq \abs{f_{nN}(\psibf)} \leq I_\epsilon(\abs{\Phi_n})
-\]
-and consequently,
-\begin{equation}\label{eq:sumf2absfCe}
-\sum_{n=1}^{N}f_{nN}(\psibf)^2 \leq \sum_{n=1}^{N}\abs{f_{nN}(\psibf)} \leq N C_\epsilon.
-\end{equation}
-\end{lemma}
-\begin{proof}
-Recall that $f_{nN}(\psibf) = \left( \tfrac{n}{N}\right)^\ell \round{\Phi_n + p_{nN}(\psibf)}$.  Because $\abs{\psi_i} < \epsilon$ for all $i = 0, \dots, m$, 
-\begin{equation}\label{eq:pnNsmall}
-\abs{p_{nN}(\psibf)} = \abs{\sum_{i=0}^{m}\big(\tfrac{n}{N}\big)^i \psi_i} \leq (m+1)\epsilon < 1.
-\end{equation}
-%for all $n = 1, \dots, N$.  
-Since $\Phi_n \in [-\nicefrac{1}{2}, \nicefrac{1}{2})$, it follows that $f_{nN}(\psibf)$ equals either $-(\tfrac{n}{N})^\ell$, $(\tfrac{n}{N})^\ell$ or $0$ and so
-\begin{equation}\label{eq:fnNleqfnNs}
-   f_{nN}(\psibf)^2 \leq \abs{f_{nN}(\psibf)} \leq 1.
-\end{equation}
-Whenever $f_{nN}(\psibf) \neq 0$ we must have 
-\[
-\abs{\Phi_n} \geq \nicefrac{1}{2} - \abs{p_{nN}(\psibf)} \geq \nicefrac{1}{2} - (m+1)\epsilon
-\]
-and $I_\epsilon(\abs{\Phi_n}) = 1$.  Thus, for all $n = 1, \dots, N$,
-\[
-f_{nN}(\psibf)^2 \leq \abs{f_{nN}(\psibf)} \leq I_\epsilon(\abs{\Phi_n}). %\qquad \text{for all $n = 1, \dots, N$.}
-\]
-Summing the terms in this inequality over $n = 1, \dots, N$ and using~\eqref{eq:Cedefn} gives~\eqref{eq:sumf2absfCe}.
-\end{proof}
-
-\begin{lemma}\label{lem:chaining3}(Chaining)
-For all positive integers $k$,
-\[
-\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) } \leq \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{\sqrt{ i A_1 + A_2 }}{2^{i/2}},
-\]
-where the constants $A_1 = 18(m+1)\log 2$ and $A_2 = 18\log K_3$.
-\end{lemma}
-\begin{proof}
-Let $b_k$ be a function that maps each $\psibf \in T_\epsilon(k)$ to $b_k(\psibf) \in T_\epsilon(k-1)$ such that  $d(\psibf,b_k(\psibf)) \leq 2^{1-k} C_\epsilon N$.
-The existence of the function $b_k$ is guaranteed by the definition of $T_\epsilon(k)$.
-By the triangle inequality,
-\[
-\abs{ Z_N(\psibf) } \leq \abs{ Z_N(b_k(\psibf)) } + \abs{ Z_N(\psibf) - Z_N(b_k(\psibf))  },
-\]
-and by taking supremums on both sides,
-\begin{align}
-\sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) } &\leq \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(b_k(\psibf)) } + \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) - Z_N(b_k(\psibf))  } \nonumber \\
-&\leq \sup_{\psibf \in T_\epsilon(k-1)} \sabs{ Z_N(\psibf) } + \sup_{\psibf \in T_\epsilon(k)} \sabs{ Z_N(\psibf) - Z_N(b_k(\psibf)) }, \label{eq:supZchain}
-\end{align}
-the last line following since $b_k(\psibf) \in T_\epsilon(k-1)$ and so 
-\[
-\sup_{\psibf \in T_\epsilon(k)} \abs{ Z_N(b_k(\psibf)) } \leq \sup_{\psibf \in T_\epsilon(k-1)} \abs{ Z_N(\psibf) }.
-\]
-Conditional on $\Phi_1, \Phi_2, \dots$ the random variable $X(\psibf) = Z_N(\psibf) - Z_N(b_k(\psibf))$ has zero mean, and is normally distributed with variance
-\begin{align*}
-\sigma_X^2 &= \expect_{\Phi} \frac{1}{N} \sum_{n=1}^N g_n^2 \big( f_{nN}(\psibf) - f_{nN}(b_k(\psibf)) \big)^2 \\
-&= \frac{1}{N} \sum_{n=1}^N \big( f_{nN}(\psibf) - f_{nN}(b_k(\psibf)) \big)^2 = d(\psibf, b_k(\psibf)) \leq 2^{1-k} C_\epsilon,
-\end{align*}
-because $\expect_{\Phi} g_n^2 = 1$.  Using Lemma~\ref{lem:maxineq}, 
-\begin{align*}
-\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k)} \sabs{X(\psibf)} \leq 3\sqrt{2^{1-k} C_\epsilon \log\sabs{T_\epsilon(k)}} \leq \sqrt{C_\epsilon}\frac{\sqrt{  k A_1 + A_2 }}{2^{k/2}}
-\end{align*}
-because $\log\abs{T_\epsilon(k)} \leq k(m+1)\log 2 +  \log K_3$.  Taking expectations on both sides of~\eqref{eq:supZchain} gives
-\begin{align*}
-\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) } \leq \expect_{\Phi} \sup_{\psibf \in T_\epsilon(k-1)} \sabs{ Z_N(\psibf) }  + \sqrt{C_\epsilon}\frac{\sqrt{  k A_1 + A_2 }}{2^{k/2}},
-\end{align*}
-which involves a recursion in $k$.  By unravelling the recursion, and using the fact $T_\epsilon(0)$ contains only the origin, and therefore 
-\[
-\expect_{\Phi}\sup_{\psibf \in T_\epsilon(0)} \abs{ Z_N(\psibf) } = \expect_{\Phi}\abs{ Z_N(0) } = 0,
-\]
-we obtain $\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) } < \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{\sqrt{  k A_1 + A_2 }}{2^{k/2}}$
-as required.
-\end{proof}
-
-\begin{lemma}\label{lem:maxineq}(Maximal inequality)
-Suppose $X_1, \dots, X_N$ are zero mean Gaussian random variables each with variance less than some positive constant $K$, then $\expect \sup_{n = 1, \dots, N} \abs{X_n} \leq 3  \sqrt{K \log N}$ where $\log N$ is the natural logarithm of $N$.
-\end{lemma}
-\begin{proof}
-This result is well known, see for example~\cite[Section 3]{Pollard_asymp_empi_proc_1989}  
-\end{proof}
-
-\begin{lemma}\label{lem:metricentropy} (Covering numbers)
-For $k \in \ints$ there exists a discrete set $T_\epsilon(k) \subset \reals^{m+1}$ with the property that, for every $\psibf \in B_\epsilon$, there is a $\psibf^* \in T_\epsilon(k)$ such that,
-\[
-d(\psibf,\psibf^*) = \sum_{n = 1}^N \big( f_{nN}(\psibf) - f_{nN}(\psibf^*) \big)^2 \leq \frac{C_\epsilon N}{2^k}.
-\]
-The number of elements in $T_\epsilon(k)$ is no more than $K_3 2^{(m+1)k}$ where $K_3$ is a positive constant, independent of $N$, $\epsilon$ and $k$.
-\end{lemma}
-
-Before we give the proof of this lemma we need some results from the literature on hyperplane arrangements and what are called \emph{$\epsilon$-cuttings}~\cite{Chazelle_discrepency_method_2000,Matousek_lect_disc_geom_2002}.
-Let $H$ be a set of $m$-dimensional affine hyperplanes lying in $\reals^{m+1}$.
-By \emph{affine} it is meant that  the hyperplanes need not pass through the origin.
-For each hyperplane $h \in H$ let $D(h)$ and its complement $\bar{D}(h)$ be the corresponding half spaces of $\mathbb{R}^{m+1}$.
-For a point $\xbf \in \mathbb{R}^{m+1}$, let
-\begin{equation}\label{eq:bhsi}
-b(h,\xbf) = \begin{cases} 
-1 & \xbf \in D(h) \\
-0 & \xbf \in \bar{D}(h).
-\end{cases}
-\end{equation}
-Note that $b(h,\xbf)$ is piecewise constant in $\xbf$.  Two points $\xbf$ and $\ybf$ from $\reals^{m+1}$ are in the same halfspace of $h$ if and only if $b(h, \xbf) = b(h, \ybf)$.  So the pseudometric
-\[
-\sigma(\xbf,\ybf) = \sum_{h \in H} |b(h,\xbf) - b(h,\ybf)|
-\]
-is the number of hyperplanes in $H$ that pass between the points $\xbf$ and $\ybf$.  %Observe that $\sigma(\xbf, \ybf)$ is peicewise constant in both $\xbf$ and $\ybf$.
-
-The next theorem considers the partitioning of $\reals^{m+1}$ into subsets so that not too many hyperplanes intersect with any subset.
-Proofs can be found in Theorem 5.1 on page 206 of~\cite{Chazelle_discrepency_method_2000} and also Theorem 6.5.3 on page 144 of~\cite{Matousek_lect_disc_geom_2002}.
-
-\begin{theorem}\label{thm:epscutting}
-There exists a constant $K$, independent of the set of hyperplanes $H$, such that for any positive real number $r$, we can partition $\reals^{m+1}$ into $K r^{m+1}$ generalised $(m+1)$-dimensional simplices with the property that no more than $\abs{H}/r$ hyperplanes from $H$ pass through the interior of any simplex.
-\end{theorem}
-
-By the phrase `There exists a constant $K$, independent of the set of hyperplanes $H$', it is meant that the constant $K$ is valid for every possible set of hyperplanes in $\reals^{m+1}$, regardless of the number of hyperplanes or their position and orientation.
-A \emph{generalised} $(m+1)$-dimensional simplex is the region defined by the intersection of $m+2$ half spaces in $\reals^{m+1}$.
-Note that a generalised simplex (unlike an ordinary simplex) can be unbounded.
-For our purposes Theorem~\ref{thm:epscutting} is important because of the following corollary.
-
-\begin{corollary}\label{cor:epscutting}
-There exists a constant $K$, independent of the set of hyperplanes $H$, such that for every positive real number $r$ there is a discrete subset $T \subset \reals^{m+1}$ containing no more than $K r^{m+1}$ elements with the property that for every $\xbf \in \reals^{m+1}$ there exists $\ybf \in T$ with $\sigma(\xbf, \ybf) \leq \abs{H}/r$.
-\end{corollary}
-\begin{proof}
-Let $C$ be the set of generalised simplices constructed according to Theorem~\ref{thm:epscutting}.
-Define $T$ as a set containing precisely one point from the interior of each simplex in $C$.
-Let $\xbf \in \reals^{m+1}$.
-Since $b(h, \xbf)$ is piecewise constant for each $h \in H$, and since $C$ partitions $\reals^{m+1}$, there must exist a simplex $c \in C$ with a point $\zbf$ in its interior such that $b(h,\zbf) = b(h, \xbf)$ for all $h \in H$, and correspondingly $\sigma(\zbf, \xbf) = 0$.
-Let $\ybf$ be the element from $T$ that is in the interior of $c$.
-Since at most $\abs{H}/r$ hyperplanes cross the interior of $c$ there can be at most $\abs{H}/r$ hyperplanes between $\zbf$ and $\ybf$, and so $\sigma(\zbf, \ybf) \leq \abs{H}/r$.
-Now $\sigma(\xbf, \ybf) < \sigma(\zbf, \xbf) + \sigma(\zbf, \ybf) \leq \frac{\abs{H}}{r}$ follows from the triangle inequality.
-\end{proof}
-
-The previous corollary ensures that we can \emph{cover} $\reals^{m+1}$ using $K r^{m+1}$ `balls' of radius $\abs{H}/r$ with respect to the pseudometric $\sigma(\xbf, \ybf)$.
-The balls are placed at the positions defined by points in the set $T$, and these points could be anywhere in $\reals^{m+1}$.
-The next corollary asserts that we can cover a subset of $\reals^{m+1}$, by placing the balls at points only within this subset.
-
-\begin{corollary}\label{cor:epscuttingsubset}
-Let $B$ be a subset of $\reals^{m+1}$.
-There exists a constant $K$, independent of the set of hyperplanes $H$, such that for every positive real number $r$ there is a discrete subset $T_B \subset B$ containing no more than $K r^{m+1}$ elements with the property that for every $\xbf \in B$ there exists $\ybf \in T_B$ with $\sigma(\xbf, \ybf) \leq \abs{H}/r$.
-\end{corollary}
-\begin{proof}
-Let $C$ be the set of generalised simplices constructed according to Theorem~\ref{thm:epscutting} and let $C_B$ be the subset of those indices that intersect $B$.
-Let $T_B$ contain a point from $c \cap B$ for each simplex $c \in C_B$.
-The proof now follows similarly to Corollary~\ref{cor:epscutting}.
-\end{proof}
-
-We are now ready to prove Lemma~\ref{lem:metricentropy}.
-
-\begin{proof} (Lemma~\ref{lem:metricentropy})
-Put $g_{nN}(\psibf) = (\tfrac{N}{n})^\ell f_{nN}(\psibf) = \round{ \Phi_n + p_{nN}(\psibf)}$, and let
-\[
-d_g(\psibf, \psibf^*) = \sum_{n=1}^{N} \big( g_{nN}(\psibf) - g_{nN}(\psibf^*) \big)^2.
-\]
-We have $d(\psibf, \psibf^*) \leq d_g(\psibf, \psibf^*)$, and so it suffices to prove the lemma with $d$ replaced by $d_g$.  
-%It will be more convenient to work with $d_g$ rather than $d$.
-From~\eqref{eq:pnNsmall} it follows that $\abs{p_{nN}(\psibf)} \leq (m+1)\epsilon < 1$.  Since $\Phi_n \in [-\nicefrac{1}{2}, \nicefrac{1}{2})$, when $\Phi_n \geq 0$,
-\[
-g_{nN}(\psibf) = \begin{cases}
-1 & p_{nN}(\psibf) \geq \nicefrac{1}{2} - \Phi_n \\
-0 & \text{otherwise},
-\end{cases}
-\]
-and when $\Phi_n < 0$,
-\[
-g_{nN}(\psibf) = \begin{cases}
--1 & p_{nN}(\psibf) < -\nicefrac{1}{2} - \Phi_n \\
-0 & \text{otherwise}.
-\end{cases}
-\]
-Thus, $( g_{nN}(\psibf) - g_{nN}(\psibf^*) )^2$ is either equal to one when $g_{nN}(\psibf) \neq g_{nN}(\psibf^*)$ or zero when $g_{nN}(\psibf) = g_{nN}(\psibf^*)$.
-Now $g_{nN}(\psibf) \neq 0$ only if
-\[
-\abs{\Phi_n} \geq \nicefrac{1}{2} - \abs{p_{nN}(\psibf)} \geq \nicefrac{1}{2} - (m+1)\epsilon,
-\]
-that is, only if $I_\epsilon(\Phi_n) = 1$.
-Let $A = \{ n \in \{1, \dots, N\} \mid I_\epsilon(\Phi_n) = 1 \}$ be the subset of the indices where $I_\epsilon(\Phi_n) = 1$.
-By definition the number of elements in $A$ is $C_\epsilon N$ (see~\eqref{eq:Cedefn}).
-If both $\psibf$ and $\psibf^*$ are in $B_\epsilon$, then
-\[
-(g_{nN}(\psibf) - g_{nN}(\psibf^*) )^2 \neq 0 
-\]
-only if $n \notin A$.  Thus,
-\begin{align*}
-d_g(\psibf, \psibf^*) = \sum_{n=1}^{N} \big( g_{nN}(\psibf) - g_{nN}(\psibf^*) \big)^2 = \sum_{n \in A} \big( g_{nN}(\psibf) - g_{nN}(\psibf^*) \big)^2.
-\end{align*}
-
-We now use Corollary~\ref{cor:epscuttingsubset}.  Let $h_n$ be the $m$ dimensional hyperplane in $\reals^{m+1}$ satisfying
-\[
-p_{nN}(\psibf) =  \sum_{i=0}^{m} \big(\tfrac{n}{N}\big)^i \psi_i = \tfrac{1}{2} \sign{\Phi_n} - \Phi_n
-\]
-where $\sign{\Phi_n}$ is equal to $1$ when $\Phi_n \geq 0$ and $-1$ otherwise.
-The hyperplane $h_n$ divides $\reals^{m+1}$ into two halfspaces, $D(h_n)$ and its complement $\bar{D}(h_n)$.
-If $\psibf$ and $\psibf^*$ are in the same halfspace, then $\abs{ b(h_n,\psibf) - b(h_n,\psibf^*) } = 0$ and $g_{nN}(\psibf) = g_{nN}(\psibf^*)$, and therefore $( g_{nN}(\psibf) - g_{nN}(\psibf^*) )^2 = 0$.
-Otherwise, if $\psibf$ and $\psibf^*$ are in different halfspaces, then $\abs{ b(h_n,\psibf) - b(h_n,\psibf^*) } = 1$ and $g_{nN}(\psibf) \neq g_{nN}(\psibf^*)$, and therefore $( g_{nN}(\psibf) - g_{nN}(\psibf^*) )^2 = 1$.
-Thus,
-\[
-(g_{nN}(\psibf) - g_{nN}(\psibf^*) )^2 = \abs{ b(h_n,\psibf) - b(h_n,\psibf^*) }
-\]
-for $n = 1, \dots,  N$.
-Let $H$ be the finite set of hyperplanes $\{ h_n, n \in A\}$ and observe that the number of hyperplanes is $\abs{H} = \abs{A} = C_\epsilon N$.
-When both $\psibf$ and $\psibf^*$ are inside $B_\epsilon$, $d_g$ can be written as
-\begin{align*}
-d_g(\psibf, \psibf^*) &= \sum_{n\in A} \abs{ b(h_n,\psibf) - b(h_n,\psibf^*) } \\
-&=  \sum_{h \in H} \abs{ b(h,\psibf) - b(h,\psibf^*) } = \sigma(\psibf, \psibf^*).
-\end{align*}
-That is, when both $\psibf, \psibf^* \in B_\epsilon$, $d_g(\psibf, \psibf^*)$ is the number of hyperplanes from $H$ that pass between the points $\psibf$ and $\psibf^*$.
-
-It follows from Corollary~\ref{cor:epscuttingsubset} that for any positive $r$ there exists a finite subset $T_B$ of $B_\epsilon$ containing at most $K_3 r^{m+1}$ elements, such that for every $\psibf \in B_\epsilon$ there is a $\psibf^* \in T_B$ with
-\[
-d_g(\psibf, \psibf^*) = \sigma(\psibf, \psibf^*) \leq \frac{\abs{H}}{r} = \frac{\abs{A}}{r} = \frac{C_\epsilon N}{r}.
-\]
-Putting $r = 2^k$ and choosing $T_\epsilon(k) = T_B$ completes the proof.
-\end{proof}
-
-\bibliographystyle{acmtrans-ims}
-\bibliography{bib}
-
 \end{document}

### 2012-10-26 paper2.tex
added 2nd part.

%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
\documentclass[journal]{IEEEtran}

\usepackage{mathbf-abbrevs}

%\newcommand{\dealias}{\operatorname{dealias}}

\input{defs}

%\newcommand{\term}{\emph}
%\renewcommand{\index}[1]{}

\title{Polynomial phase estimation by phase unwrapping 2: Asymptotic normality and simulations}
%\title{Polynomial phase estimation by phase unwrapping}

\author{Robby~G.~McKilliam, Barry~G.~Quinn, I.~Vaughan~L.~Clarkson, Bill~Moran and Badri~N.~Vellambi%
    \thanks{%A preliminary version of some of this material is contained in Part 3 of Robby McKilliam's PhD thesis \cite{McKilliam2010thesis}. 
Robby~McKilliam and Badri Vellambi are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.  Barry~Quinn is with the Department of Statistics, Macquarie University, Sydney, NSW, 2109, Australia.   Vaughan~Clarkson is with the School of Information Technology \& Electrical Engineering, The University of Queensland, QLD., 4072, Australia.  B. Moran is with the Department of Electrical Engineering and Computer
Science, Melbourne Systems Lab, Dept of Elec \& Electronic Eng, Uni of Melbourne, Vic. 3010, Australia.}}
% The paper headers 
\markboth{Polynomial phase estimation by phase unwrapping}{DRAFT \today}

% make the title area

\begin{document} 
\maketitle

\begin{abstract}
Estimating the coefficients of a noisy polynomial phase signal is important in fields including radar, biology and radio communications. One approach attempts to perform polynomial regression on the phase of the signal.  This is complicated by the fact that the phase is \emph{wrapped} modulo $2\pi$ and must be \emph{unwrapped} before regression can be performed. %A recent approach suggested by the authors is to perform the unwrapping in a least squares manner.  %It was shown how the resulting estimator could be posed as a \emph{nearest lattice point problem} in a specific lattice and could be solved using existing algorithms. 
%It was shown by Monte Carlo simulation that this produces a remarkably accurate estimator.  
In this paper we consider an estimator that performs phase unwrapping in a least squares manner.  We describe the asymptotic properties of this estimator, showing that it is strongly consistent and asymptotically normally distributed. %We hypothesise that the estimator produces very near maximum likelihood performance.
\end{abstract}

\begin{keywords}
Polynomial phase signals, phase unwrapping, asymptotic properties, nearest lattice point problem
\end{keywords}

%\bibliographystyle{unsrt}

\section{Introduction} \label{intro}

Polynomial phase signals arise in fields including radar, sonar, geophysics, biology, and radio communication \cite{Angeby_estimating_2000}. In radar and sonar applications polynomial phase signals arise when acquiring radial velocity and acceleration (and higher order motion descriptors) of a target from a reflected signal, and also in continuous wave radar and low probability of intercept radar~\cite{Levanon_Radar_signals_2004}.  In biology, polynomial phase signals are used to describe the sounds emitted by bats and dolphins for echo location \citep{Suga_1975_bats_echolocation, Moss_2005echolocation}.

A polynomial phase signal of order $m$ is a function of the form
\[
s(t) = e^{2\pi j y(t)},
\]
where $j = \sqrt{-1}$, and $t$ is a real number, often representing time, and 
\[
y(t) = \tilde{\mu}_0 +\tilde{\mu}_1 t + \tilde{\mu}_2 t^2 + \dots \tilde{\mu}_m t^m
\]
is a polynomial of order $m$.  In practice the signal is typically sampled at discrete points in `time', $t$. In this paper we only consider uniform sampling, where the gap between consecutive samples is constant. In this case we can always consider the samples to be taken at some set of consecutive integers and our sampled polynomial phase signal looks like
\[
s_n = s(n) = e^{2\pi j y(n)},
\] 
where $n$ is an integer.  Of practical importance is the estimation of the coefficients $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ from a number, say $N$, of observations of the noisy sampled signal
\begin{equation}\label{eq:Y_nsamplednoisey}
Y_n = \rho s_n + X_n,
\end{equation}
where $\rho$ is a real number greater than zero representing the (usually unknown) signal amplitude and $\{X_n, n \in \ints\}$ is a sequence complex noise variables. In order to ensure identifiability it is necessary to restrict the $m+1$ coefficients to a region of $m+1$ dimensional Euclidean space $\reals^{m+1}$ called an \emph{identifiable region}.  It was shown in \cite{McKilliam2009IndentifiabliltyAliasingPolyphase} that an identifiable region tessellates a particular $m+1$ dimensional lattice.  We discuss this in Section \ref{sec:ident_aliasing}.

%A popular estimator is based on the \emph{discrete polynomial phase transform} (DPT) first introduced by Peleg and Porat \cite{Peleg_DPT_1995}.  The transform enables each parameter to be be estimated iteratively using the Fast Fourier transform in a way analogous to the frequency estimator of Rife and Boorstyn \cite{Rife1974, Quinn2001}.  The DPT estimator is characterized by good statistical performance and computational efficiency.  Variants of the DPT have been suggested by numerous authors including Golden and Friedlander \cite{Golden_mod_dpt_1998} and O'Shea \cite{Oshea_iterative_1996}. The recent \emph{high-order phase function} estimators are based on similar ideas \cite{Farquharson_another_poly_est_2005}.  A substantial drawback of all of these estimators is that they only work correctly over a small set of parameters, generally much smaller than the identifiable region.  This property appears to have been first noticed in \cite{McKilliam2009IndentifiabliltyAliasingPolyphase, McKilliam2009asilomar_polyest_lattice} and has been analysed further in \cite[Ch. 10]{McKilliam2010thesis}.
%
%Another estimator, described by Kitchen \cite{Kitchen_polyphase_unwrapping_1994}, is based on the phase unwrapping approach to frequency estimation suggested by Kay \cite{Kay1989}.  The primary advantage of Kitchen's estimator is that it requires only $O(N)$ arithmetic operations to compute. The primary disadvantage is that it only works correctly when the signal to noise ratio (SNR) is high. It was also observed in \cite{McKilliam2009asilomar_polyest_lattice} that Kitchen's estimator does not work correctly for all coefficients in the identifiable region, but, the problem is far less severe than for estimators based on the DPT \cite[Sec. 10.3]{McKilliam2010thesis}.

An obvious estimator of the unknown coefficients is the least squares estimator (LSE).  This is also the maximum likelihood estimator (MLE) when the noise sequence $\{X_n\}$ is white and Gaussian.  When $m=0$ (phase estimation) or $m=1$ (frequency estimation) the LSE is an effective approach, being both computationally efficient and statistically accurate \cite{Quinn2009_dasp_phase_only_information_loss,Hannan1973,Quinn2001,McKilliam_mean_dir_est_sq_arc_length2010}\cite[Sec.~6.4~and~9.1]{McKilliam2010thesis}. When $m \geq 2$ the computational complexity of the LSE is large~\cite[Sec.~10.1]{McKilliam2010thesis}\cite{Abatzoglou_ml_chirp_1986}. For this reason many authors have considered alternative approaches to polynomial phase estimation. These can loosely be grouped into two classes, estimators based on polynomial phase transforms, such as the discrete polynomial phase transform (DPT)~\cite{Peleg_DPT_1995,Peleg1991_est_class_PPS_1991,Porat_asympt_HAF_DPT_1996} and the high order phase function~\cite{Farquharson_another_poly_est_2005}, and estimators based on phase unwrapping, such as Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, the estimator of Djuric and Kay~\cite{Djuric_phase_unwrap_chirp_1990}, and Morelande's Bayesian unwrapping estimator~\cite{Morelande_bayes_unwrapping_2009_tsp}.

In this paper we consider the estimator that results from unwrapping the phase in a least squares manner.  We call this the \emph{least squares unwrapping estimator} (LSU)~\cite{McKilliam2009asilomar_polyest_lattice, McKilliamFrequencyEstimationByPhaseUnwrapping2009}\cite[Chap. 8]{McKilliam2010thesis}.  It was shown in \cite{McKilliam2009asilomar_polyest_lattice, McKilliamFrequencyEstimationByPhaseUnwrapping2009} that the LSU estimator can be represented as a \emph{nearest lattice point problem}, and Monte-Carlo simulations were used to show the LSU estimator's favourable statistical performance. %Moreover, it was noticed that the LSU appears to work correctly for coefficients anywhere in the identifiable region.  A drawback of the LSU estimator is that computing a nearest lattice point is, in general, computationally difficult.  In \cite{McKilliam2009asilomar_polyest_lattice}, two standard techniques were considered, the \emph{sphere decoder} \cite{Pohst_sphere_decoder_1981}, and \emph{Babai's nearest plane algorithm} \cite{Babai1986}. The sphere decoder was observed to have excellent statistical performance but can only be computed efficiently for $N<50$ \cite{Jalden2005_sphere_decoding_complexity}.  Computing the Babai point requires only $O(N^2)$ operations, but its statistical performance is comparatively poor at low signal to noise ratio (SNR). A major point of interest is that the lattices considered are not \emph{random}, and therefore may admit fast nearest point algorithms. 
%This has been studied in~\cite[Sec.~4.2]{McKilliam2010thesis} where polynomial time algorithms where found that compute the nearest point exactly.  Unfortantely, although polynomial time, the algorithm are still computationally very slow in practice. 
% Two main questions were raised in \cite{McKilliam2009asilomar_polyest_lattice}:
% \begin{enumerate}
% \vspace{-0.1cm}
% \item What are the asymptotic statistical properties of the LSU estimator?
% \vspace{-0.18cm}
% \item Can the structure of the lattice be exploited to obtain fast (exact or approximate) nearest point algorithms and thereby a computationally efficient estimator?
% \vspace{-0.1cm}
% \end{enumerate}
% The purpose of this paper is to answer the first question. We prove that the LSU estimator is strongly consistent and asymptotically normally distributed. 
%We also apply another known approximate nearest point algorithm, called the $K$-best method \cite{Zhan2006_K_best_sphere_decoder}, that we show provides near sphere decoder performance, but can be computed in a reasonable amount of time if $N$ is less than about 300.
In this paper we derive the asymptotic properties of the LSU estimator, showing that it is strongly consistent and asymptotically normally distributed.  Similar results were stated without a complete proof in~\cite{McKilliam_polyphase_est_icassp_2011}.  Here, we give a proof.  The results here are also more general than in~\cite{McKilliam_polyphase_est_icassp_2011}, allowing for a wider class of noise distributions.

An interesting property is that the estimator of the $k$th polynomial phase coefficient converges almost surely to $\tilde{\mu}_k$ at rate $o(N^{-k})$.  This is perhaps not surprising, since it is the same rate observed in polynomial regression.  However, asserting that convergence at this rate occurs in the polynomial phase setting is not trivial.  For this purpose we make use of an elementary result about the number of arithmetic progressions contained inside subsets of $\{1,2,\dots,N\}$~\cite{Erdos_on_some_sequence_of_integers1936,Szemeredi_setint_no_k_arth1975,Gowers_new_proof2001}.  %This proof technique appears to be novel, and may be useful in other problems.  
The proof of asymptotic normality is complicated by the fact that the objective function corresponding with the LSU estimator is not differentiable everywhere.  Empirical process techniques~\cite{Pollard_new_ways_clts_1986,Pollard_asymp_empi_proc_1989,van2009empirical,Dudley_unif_central_lim_th_1999} and results from the literature on hyperplane arrangements~\cite{Chazelle_discrepency_method_2000,Matousek_lect_disc_geom_2002} become useful here.  We are hopeful that the proof techniques developed here will be useful for purposes other than polynomial phase estimation, and in particular other applications involving data that is `wrapped' in some sense.  Potential candidates are the phase wrapped images observed in modern radar and medical imaging devices such as synthetic aperture radar and magnetic resonance imaging~\cite{Nico_phaseunwrappingSAR_2000,Friedlander_PD_phaseunwrapping_1996}.

The paper is organised in the following way. Section~\ref{sec:lattice-theory} describes the required concepts from lattice theory.  In Section~\ref{sec:ident_aliasing} we briefly describe the identifiable region that was also derived in~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}.  These identifiability results are required in order to properly understand the statistical properties of polynomial phase estimators. In Section \ref{sec:least-squar-unwr} we describe the LSU estimator and derive its asymptotic statistical properties.
%and also consider techniques for computing the LSU as a nearest lattice point problem. 
Section \ref{sec:simulations} describes the results of Monte Carlo simulations that compare the performance of the LSU estimator with existing estimators.  These simulations agree with the derived asymptotic properties.

%\subsection{Notation}
%We write random variables using capital letters, such as $X$ and $Y$ and circular random variables using the capital Greek letters $\Theta$ and $\Phi$.  We use $\round{x}$ to denote the nearest integer to $x$ with half integers rounded up and use $\fracpart{x} = x - \round{x}$ to denote the \emph{centred} fractional part.  Both $\round{\cdot}$ and $\fractpart{\cdot}$ operate elementwise on vectors.

\section{The least squares unwrapping estimator}\label{sec:least-squar-unwr}

We now describe the least squares unwrapping (LSU) estimator of the polynomial coefficients. Recall that we desire to estimate the coefficients $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ from the noisy samples $Y_1, \dots, Y_N$ given in~\eqref{eq:Y_nsamplednoisey}.  We take the complex argument of the $Y_n$ and divide by $2\pi$ to obtain
\begin{equation}\label{eq:noise_circ_poly}
\Theta_n = \frac{\angle{Y_n}}{2\pi} = \fracpart{ \Phi_n + y(n) }
\end{equation}
where $\angle{z}$ denotes the complex argument of the complex number $z$, and 
\[
\Phi_n = \frac{1}{2\pi}\angle(1 + \rho^{-1}s_n^{-1}X_n)
\] 
are random variables representing the phase noise induced by the $X_n$~\cite{Tretter1985,Quinn2009_dasp_phase_only_information_loss}.  If the distribution of $X_n$ is circularly symmetric (i.e., the angle $\angle X_n$ is uniformly distributed on $[-\pi, \pi)$ and is independent of the magnitude $\abs{X_n}$) then the distribution of $\Phi_n$ is the same as the distribution of $\tfrac{1}{2\pi}\angle(1 + \rho^{-1}X_n)$.  If the $X_1, \dots, X_N$ are circularly symmetric and identically distributed, then $\Phi_1, \dots, \Phi_n$ are also identically distributed.

Let $\mubf$ be the vector $[\mu_0, \mu_1, \dots, \mu_m]$ and put,
\begin{equation} \label{eq:sumofsquaresfunction}
SS(\mubf) = \sum_{n=1}^{N}\fracpart{  \Theta_{n} - \sum_{k = 0}^{m}{\mu_k n^k} }^{2}.
\end{equation}
The least squares unwrapping estimator is defined as those coefficients $\widehat{\mu}_0, \dots, \widehat{\mu}_m$ that minimise $SS$ over the identifiable region $B$. That is, the LSU estimator is,
\begin{equation}\label{eq:hatmubfLSUdefn}
\widehat{\mubf} = \arg \min_{\mubf \in B} SS(\mubf). 
\end{equation}

It is shown in~\cite[Sec~8.1]{McKilliam2010thesis}\cite{McKilliam2009asilomar_polyest_lattice} how this minimisation problem can be posed as that of computing a nearest lattice point in a particular lattice. Polynomial time algorithms that compute the nearest point are described in~\cite[Sec.~4.3]{McKilliam2010thesis}.  Although polynomial in complexity, these algorithms are not fast in practice.  The existence of practically fast nearest point algorithms for these lattices is an interesting open problem.  In this paper we focus on the asymptotic statistical properties of the LSU estimator, rather than computational aspects.

The next theorem describes the asymptotic properties of this estimator.  Before we give the proof it is necessary to understand some of the properties of the phase noise $\Phi_1,\dots,\Phi_N$, which are \emph{circular} random variables with support on $[-\nicefrac{1}{2}, \nicefrac{1}{2})$~\cite{McKilliam2010thesis,McKilliam_mean_dir_est_sq_arc_length2010,Mardia_directional_statistics,Fisher1993}.  Circular random variables are often considered modulo $2\pi$ and therefore have support $[-\pi, \pi)$ with $-\pi$ and $\pi$ being identified as equivalent.  Here we instead consider circular random variables modulo 1 with support $[-\nicefrac{1}{2}, \nicefrac{1}{2})$ and with $-\nicefrac{1}{2}$ and $\nicefrac{1}{2}$ being equivalent.  This is nonstandard but it allows us to use notation such as $\round{\cdot}$ for rounding and $\fracpart{\cdot}$ for the centered fractional part in a convenient way.   %The random variable $2\pi \Phi_n$ has support on $[-\pi, \pi)$ and can be identified as a \emph{circular random variable}~\cite{Mardia_directional_statistics,Fisher1993}.  %Let $f$ be the probability density function (pdf) of $\Phi_n$.  Then $f$ takes nonzero values only on $[-\nicefrac{1}{2}, \nicefrac{1}{2})$.  It is instructive to think of $f$ as describing a describing a distribution on a circle.  For example, when the noise term $X_n$ is complex Gaussian with independent and identically distributed real and imaginary parts, the distribution of $\Phi_n$ is the projected Guassian~\cite[Section 5.6.1]{McKilliam2010thesis}\cite[page 46]{Mardia_directional_statistics}.  It is instrutive the consider a periodic version of $f$, that we write as $f(\fracpart{x})$, which has period equal to one, because the fractional part function $\fracpart{x}$ has period one.

The \emph{intrinsic mean} or \emph{Fr\'{e}chet mean} of $\Phi_n$ is defined as~\cite{McKilliam_mean_dir_est_sq_arc_length2010,bwhk07a,Bhattacharya_int_ext_means_2003,Bhattacharya_int_ext_means_2005},
\begin{equation}\label{eq:intrmeandefn}
 \mu_{\text{intr}}  = \arg \min_{\mu \in [-\nicefrac{1}{2}, \nicefrac{1}{2})} \expect \fracpart{\Phi_n - \mu}^2, 
% %&= \arg \min_{\mu \in [-\nicefrac{1}{2}, \nicefrac{1}{2})}\int_{-\nicefrac{1}{2}}^{\nicefrac{1}{2}}\fracpart{\phi - \mu}^2f(\theta) d\theta. \label{eq:minunrappedmeandef}
\end{equation}
and the \emph{intrinsic variance} is
\[
\sigma_{\text{intr}}^2 = E\fracpart{\Theta - \mu_{\text{intr}}}^2 = \min_{\mu \in [-\nicefrac{1}{2}, \nicefrac{1}{2})} \expect \fracpart{\Phi_n - \mu}^2,
\]
where $\expect$ denotes the expected value.  Depending on the distribution of $\Phi_n$ the argument that minimises~\eqref{eq:intrmeandefn} may not be unique.  The set of minima is often called the~\emph{Fr\'{e}chet mean set}~\cite{Bhattacharya_int_ext_means_2003,Bhattacharya_int_ext_means_2005}.  If the minimiser is not unique we say that $\Phi_n$ has no intrinsic mean.  We are now equipped to state the asymptotic properties of the LSU estimator.

%\begin{figure}[tp]
% 	\centering 
% 		\includegraphics[width=\linewidth]{plots/distplots.class.distributions.circular.ProjectedNormalDistribution.var3.0-1.mps}
% 		\caption{The projected normal distribution.}
% 		\label{fig:circularuniformdist}
%\end{figure}

%BLERG NOTE: theorem should be rewritten to instead assume that the X_1, \dots X_N are circularly symmetric.  Then the continuity stuff for the f follows immediately.  You can then talk about how it could apply more generally, but gets complicated.

\begin{theorem} \label{thm:asymp_proof} 
Let $\widehat{\mubf}$ be defined by~\eqref{eq:hatmubfLSUdefn} and put $\widehat{\lambdabf}_N = \dealias(\tilde{\mubf} - \widehat{\mubf})$.  Denote the elements of $\widehat{\lambdabf}_N$ by $\widehat{\lambda}_{0,N}, \dots, \widehat{\lambda}_{m,N}$.  Suppose $\Phi_1, \dots, \Phi_N$ are independent and identically distributed with zero intrinsic mean, intrinsic variance $\sigma^2$, and pdf $f$, then: 
\begin{enumerate}
\item (Strong consistency) $N^k \widehat{\lambda}_{k,N}$ converges almost surely to $0$ as $N\rightarrow\infty$ for all $k = 0, 1, \dots, m$.
\item (Asymptotic normality) If $f(\fracpart{x})$ is continuous at $x = -\nicefrac{1}{2}$ and $f(-\nicefrac{1}{2}) < 1$ then the distribution of
\[
\left[
\begin{array}
[c]{cccc}%
\sqrt{N} \widehat{\lambda}_{0,N} & N \sqrt{N}\widehat{\lambda}_{1,N}  & \dots & N^m\sqrt{N} \widehat{\lambda}_{m,N}
\end{array}
\right]^\prime
\]
converges to the normal with zero mean and covariance
\[
\frac{\sigma^2}{\left(1-f( -\nicefrac{1}{2}) \right)^{2}} \Cbf^{-1},
\]
where $\Cbf$ is the $(m+1)\times (m+1)$ Hilbert matrix with elements $C_{ik} = 1/(i + k + 1)$ for $i,k = \{0, 1, \dots, m\}$.
\end{enumerate}
\end{theorem}
The proof of Theorem~\ref{thm:asymp_proof} is contained within the next two sections. Section~\ref{sec:strongconstproof} proves strong consistency and Section~\ref{sec:centlimitproof} proves asymptotic normality.  The proofs for the case of $m=0$ were given in~\cite{McKilliam_mean_dir_est_sq_arc_length2010} and for the case when $m=1$ were given in~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.  The proofs here take a similar approach, but require new techniques.  %Before giving the proof we make the following remarks.
The theorem gives conditions on the \emph{dealiased} difference $\dealias(\tilde{\mubf} - \widehat{\mubf})$ between the true coefficients $\tilde{\mubf}$ and the estimated coefficients $\widehat{\mubf}$ rather than directly on the difference $\tilde{\mubf} - \widehat{\mubf}$.   To see why this makes sense, consider the case when $m=0$, $\tilde{\mu}_0 = -0.5$ and $\widehat{\mu}_0 = 0.49$, so that $\tilde{\mu}_0 - \widehat{\mu}_0 = -0.99$.  However, the two phases are obviously close, since the phase $\pm 0.5$ are actually the same.  In this case 
\[
\dealias(\tilde{\mu}_0 - \widehat{\mu}_0) = \fracpart{\tilde{\mu}_0 - \widehat{\mu}_0} = 0.01.
\] 
The same reasoning holds for $m > 0$.

The requirement that $\Phi_1, \dots, \Phi_N$ be identically distributed will typically hold only when the complex random variables $X_1, \dots, X_N$ are identically distributed and circularly symmetric.  It would be possible to drop the assumption that $\Phi_1, \dots, \Phi_N$ be identically distributed, but this complicates the theorem statement and the proof.  In the interest of simplicity we only consider the case when $\Phi_1, \dots, \Phi_N$ are identically distributed here.  If $X_n$ is circularly symmetric with density function nonincreasing with magnitude $\abs{X_n}$, then, the corresponding $\Phi_n$ necessarily has zero intrinsic mean~\cite[Theorem 5.2, page 78]{McKilliam2010thesis}.  Thus, our theorem covers commonly used distributions for $X_1, \dots, X_N$, such as the normal distribution.

The proof of asymptotic normality places requirements on the pdf $f$ of the phase noise.  The requirement that $\Phi_1, \dots, \Phi_N$ have zero intrinsic mean implies that $f(-\nicefrac{1}{2}) \leq 1$~\cite[Lemma~1]{McKilliam_mean_dir_est_sq_arc_length2010}, so the only case not handled is when equality holds, i.e., when $f(-\nicefrac{1}{2}) = 1$ or when $f(\fracpart{x})$ is discontinuous at $x = -\nicefrac{1}{2}$. In this exceptional case other expressions for the asymptotic variance can be found (similar to \cite[Theorem 3.1]{Hotz_circle_means_2011}), but this comes at a substantial increase in complexity and for this reason we have omitted them. %Distributions that do not satisfy these requirements are unlikely to be needed in practice.

\section{Proof of asymptotic normality}\label{sec:centlimitproof}

Let $\psibf$ be the vector with $k$th component $\psi_k = N^k \lambda_k$, $k=0, \dots, m$ and let 
\[
T_{N}(\psibf) = S_N(\lambdabf) = \frac{1}{N} \sum_{n=1}^{N} \fracpart{ \Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \psi_k }^2.
\]
Let $\widehat{\psibf}_N$ be the vector with elements $\widehat{\psi}_{k,N} = N^k \widehat{\lambda}_{k,N}$ so that $\widehat{\psibf}_N$ is the minimiser of $T_N$.  Because each of $N^k \widehat{\lambda}_{k,N}$ converges almost surely to zero as $N \rightarrow \infty$, then $\widehat{\psibf}_{N}$ converges almost surely to $\zerobf$ as $N \rightarrow \infty$.  We want to find the asymptotic distribution of
\[
\sqrt{N}\widehat{\psibf}_{N} = 
\left[
\begin{array}
[c]{c}%
\sqrt{N} \widehat{\psi}_{0,N} \\ \sqrt{N}\widehat{\psi}_{1,N}  \\ \vdots \\ \sqrt{N} \widehat{\psi}_{m,N}
\end{array}
\right]
=
\left[
\begin{array}
[c]{c}%
\sqrt{N} \widehat{\lambda}_{0,N} \\ N\sqrt{N}\widehat{\lambda}_{1,N} \\ \vdots \\ N^m\sqrt{N} \widehat{\lambda}_{m,N}
\end{array}
\right].
\]
The proof is complicated by the fact that $T_N$ is not differentiable everywhere as $\fracpart{x}^2$ is not differentiable when $\fracpart{x} = \tfrac{1}{2}$.  This precludes the use of ``standard approaches'' to proving asymptotic normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.  However, we show in Lemma~\ref{lem:diffathatpsi} that all the partial derivatives $\frac{\partial T_N}{\partial \psi_\ell}$ for $\ell = 0, \dots, m$ exist, and are equal to zero, at the minimiser $\widehat{\psibf}_N$.  Thus, putting
\begin{equation}\label{eq:Wn}
W_{n} = \round{\Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \widehat{\psi}_{k,N}},
\end{equation}
we have, for each $\ell = 0, \dots, m$,
\begin{align*}
0 & = \frac{\partial T_N}{\partial \psi_\ell}(\widehat{\psibf}_N)\\
&= \frac{2}{N}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\left( \Phi_n - W_{n}+ \sum_{k=0}^{m}(\tfrac{n}{N})^k \widehat{\psi}_{k,N}  \right),
\end{align*}
so that
\[
D_{\ell,N} = K_{\ell,N},
\]
where
\[
D_{\ell,N} = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}(\tfrac{n}{N})^\ell \Phi_n,
\]
and
\begin{equation}\label{eq:KellN}
K_{\ell,N} = \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\left(W_{n}- \sum_{k=0}^{m}(\tfrac{n}{N})^k \widehat{\psi}_{k,N}  \right).
\end{equation}
Lemma~\ref{lem:Kconvfhalf} shows that, for all $\ell = 0, \dots m$,
\begin{equation}\label{eq:KellNconv}
K_{\ell,N} =  (h - 1) \sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big( C_{\ell k} + o_P(1) \big) + o_P(1)
\end{equation}
where $C_{\ell k} =  \tfrac{1}{\ell + k + 1}$, and $h = f(-\nicefrac{1}{2})$, and $o_P(1)$ denotes a random variable converging in probability to zero as $N\rightarrow\infty$.

It is now convenient to write in vector form.  Let 
\begin{equation}\label{eq:dnandkn}
\kbf_N = 
\left[
\begin{array}
[c]{c}%
K_{0,N}  \\ \vdots \\ K_{m,N}
\end{array}
\right]
\qquad \text{and} \qquad 
\dbf_N = \left[
\begin{array}
[c]{c}%
D_{0,N}  \\ \vdots \\ D_{m,N}
\end{array}
\right].
\end{equation}
From~\eqref{eq:KellNconv},
\[
\dbf_N = \kbf_N = \sqrt{N} (h - 1)(\Cbf + o_P(1)) \widehat{\psibf}_N + o_P(1)
\]
where $o_P(1)$ here means a vector or matrix of the appropriate dimension with every element converging in probability to zero as $N\rightarrow\infty$.  Thus $\sqrt{N}\widehat{\psibf}_N$ has the same asymptotic distribution as $(h - 1)^{-1}\Cbf^{-1} \dbf_N$.  Lemma~\ref{eq:convdn} shows that $\dbf_N$ is asymptotically normally distributed with zero mean and covariance matrix $\sigma^2\Cbf$.  Hence $\sqrt{N}\widehat{\psibf}_N$ is asymptotically normal with zero mean and covariance matrix
\[
\frac{\sigma^2\Cbf^{-1}\Cbf(\Cbf^{-1})^\prime}{(1 - h)^2} = 
\frac{\sigma^2\Cbf^{-1}}{(1 - h)^2}.
\] 
It remains to prove Lemmas~\ref{lem:diffathatpsi},~\ref{lem:Kconvfhalf} and~\ref{eq:convdn}.

\begin{lemma}\label{lem:diffathatpsi}
For all $\ell = 0, \dots, m$ the partial derivatives $\frac{\partial T_N}{\partial \psi_\ell}$ exist, and are equal to zero, at the minimiser $\widehat{\psibf}_N$.  That is,
\[
\frac{\partial T_N}{\partial \psi_\ell}(\widehat{\psibf}_N) = 0 \qquad \text{for each $\ell = 0, \dots, m$.}
\]
\end{lemma}
\begin{IEEEproof}
The function $\fracpart{x}^2$ is differentiable everywhere except if $\fracpart{x} \neq -\nicefrac{1}{2}$.  Recalling that
\[
T_{N}(\psibf) = \frac{1}{N} \sum_{n=1}^{N} \fracpart{ \Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \psi_k }^2
\]
we see that $T_N$ will be differentiable with respect to $\psibf$ at $\widehat{\psibf}_N$ if 
\[
\fracpart{\Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \widehat{\psi}_{k,N}} \neq -\nicefrac{1}{2} \qquad \text{for all $n = 1, \dots, N$.}
\] 
This is proved in Lemma~\ref{lem:boundonhatlambda}.  So the partial derivatives $\frac{\partial T_N}{\partial \psi_\ell}$ exist for all $\ell = 0, \dots, m$ at $\widehat{\psibf}_N$.  That each of the partial derivatives is equal to zero at $\widehat{\psibf}_N$ follows immediately from the fact that $\widehat{\psibf}_N$ is a minimiser of $T_N$.
\end{IEEEproof}

\begin{lemma}\label{lem:boundonhatlambda} $\sabs{\sfracpart{\Phi_n + \sum_{k=0}^m (\nicefrac{n}{N})^k \widehat{\psi}_{k,N}}} \leq \frac{1}{2} - \frac{1}{2N}$ for all $n = 1, \dots, N$.
\end{lemma}
\begin{IEEEproof}
To simplify our notation let 
\[
B_n = \Phi_n + \sum_{k=1}^m (\nicefrac{n}{N})^k \widehat{\psi}_{k,N}
\]
so that we now require to prove $\abs{\fracpart{B_n + \widehat{\psi}_{0,N}}} \leq \frac{1}{2} - \frac{1}{2N}$
for all $n = 1, \dots, N$.  From~\eqref{eq:Wn}, $W_n = \round{B_n + \widehat{\psi}_{0,N}}$, and 
\begin{align*}
T_N(\widehat{\psibf}_N) &= \frac{1}{N}\sum_{n=1}^N \fracpart{B_n + \widehat{\psi}_{0,N}}^2
\\ &= \frac{1}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N} - W_n)^2.
\end{align*}
Since $\widehat{\psi}_{0,N}$ is the minimiser of the quadratic above,
\begin{equation}\label{eq:lamsum}
\widehat{\psi}_{0,N} = -\frac{1}{N}\sum_{n=1}^N(B_n - W_n).
\end{equation}
The proof now proceeds by contradiction.  Assume that for some $k$,
\begin{equation}\label{eq:philamfraccontra}
\fracpart{B_k + \widehat{\psi}_{0,N}} > \frac{1}{2} - \frac{1}{2N}.
\end{equation}
Let $F_n = W_n$ for all $n \neq k$ and $F_k = W_k + 1$, and let
\[
\phi = -\frac{1}{N}\sum_{n=1}^N(B_n - F_n) = \widehat{\psi}_{0,N} + \frac{1}{N}.
\]
Then,
\begin{align}
(B_k + &\phi - F_k)^2 = (B_k + \phi - W_k - 1)^2 \nonumber \\
&= (B_k + \phi - W_k)^2 - 2(B_k + \phi - W_k) + 1 \nonumber \\
&= (B_k + \phi - W_k)^2 - 2(B_k + \widehat{\psi}_{0,N} - W_k) + 1 - \frac{2}{N} \nonumber \\
&= (B_k + \phi - W_k)^2 - 2\fracpart{B_k + \widehat{\psi}_{0,N}} + 1 - \frac{2}{N} \nonumber \\
&< (B_k + \phi - W_k)^2 - \frac{1}{N}, \label{eq:Bkineq}
\end{align}
where the inequality in the last line follows from~\eqref{eq:philamfraccontra}. Let
\[
\bbf = \left[
\begin{array}
[c]{cccc}%
\phi & \widehat{\psi}_{1,N} & \cdots & \widehat{\psi}_{m,N}
\end{array}
\right]^\prime
\]
be the vector of length $m+1$ with components $b_0 = \phi$ and $b_\ell = \widehat{\psi}_{\ell,N}$ for $\ell = 1 , \dots m$.  Now,
\[
N T_N(\bbf) = \sum_{n=1}^N \fracpart{B_n + \phi }^2 \leq  \sum_{n=1}^N (B_n + \phi  - F_n)^2, 
\]
and using the inequality from~\eqref{eq:Bkineq},
\begin{align*}
N &T_N(\bbf) < - \frac{1}{N} + \sum_{n=1}^N (B_n + \phi  - W_n)^2 \\
&= - \frac{1}{N} + \sum_{n=1}^N (B_n + \widehat{\psi}_{0,N} + \frac{1}{N}  - W_n)^2 \\
&= \sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n)^2 +  \frac{2}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n)\\
%&= N T_N(\widehat{\psibf}_N) + \frac{2}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n) \\
&= N T_N(\widehat{\psibf}_N),
\end{align*}
because $\frac{2}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n) = 0$ as a result of~\eqref{eq:lamsum}.  But, now $T_N(\bbf) < T_N(\widehat{\psibf}_N)$ violating the fact that $\widehat{\psibf}_N$ is a minimiser of $T_N$.  So~\eqref{eq:philamfraccontra} is false by contradiction.

If $\fracpart{B_k + \widehat{\psi}_{0,N}} < -\frac{1}{2} + \frac{1}{2N}$ for some $k$, we set $F_k = W_k - 1$ and using the same procedure as before obtain $T_N(\bbf) < T_N(\widehat{\psibf}_N)$ again.  The proof follows.
\end{IEEEproof}

\begin{lemma}\label{lem:Kconvfhalf}
With $K_{\ell,N}$ defined as in~\eqref{eq:KellN} and $h = f(-\nicefrac{1}{2})$,
\[
K_{\ell,N} = (h - 1) \sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big( C_{\ell k} + o_P(1) \big) + o_P(1) 
\] 
for all $\ell = 0, \dots, m$, where $C_{\ell k} =  \frac{1}{\ell + k + 1}$.% is an element of the $m+1$ by $m+1$ Hilbert matrix, and $o_P(1)$ denotes a random variable converging in probability to zero as $N\rightarrow\infty$.
\end{lemma}
\begin{IEEEproof}
Care must be taken since $\widehat{\psibf}_N$ depends on the sequence $\{ \Phi_n \}$.  For $n = 1, \dots, N$ and positive $N$, let
\begin{equation}
p_{nN}(\psibf) = \sum_{k=0}^{m}\left( \tfrac{n}{N}\right)^k \psi_k,
\end{equation}
and put
\[
q_{n}(x) = \round{\Phi_n + x}, \qquad Q(x) = \expect q_{n}(x) =  \expect q_{1}(x).
\]
Let
\begin{equation}\label{eq:GNdef}
G_N(\psibf) = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}\left( \tfrac{n}{N}\right)^\ell  \big(q_n(p_{nN}(\psibf)) - Q(p_{nN}(\psibf)) \big).
\end{equation}
and put
\begin{equation}\label{eq:hatpnN}
\widehat{p}_{nN} = p_{nN}(\widehat{\psibf}_N) = \sum_{k=0}^m \left(\tfrac{n}{N}\right)^k \widehat{\psi}_{k,N}.
\end{equation}
Observe that $G_N$ depends on $\ell$ and we could write $G_{\ell,N}$ but have suppressed the subscript $\ell$ for notational simplicity.  Now $W_n$ from~\eqref{eq:Wn} can be written as
\[
W_n = \round{\Phi_n + \widehat{p}_{nN}} = q_{n}(\widehat{p}_{nN})
\]
and $K_{\ell,N}$ from~\eqref{eq:KellN} can be written as
\begin{align*}
K_{\ell,N} &= \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\big( q_{n}(\widehat{p}_{nN}) - \widehat{p}_{nN}  \big) \\
&= \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\big(q_{n}(\widehat{p}_{nN}) - \widehat{p}_{nN} + Q(\widehat{p}_{nN}) - Q(\widehat{p}_{nN}) \big) \\
&= G_N(\widehat{\psibf}_N) + H_{\ell,N},
\end{align*}
where 
\begin{equation}\label{eq:HellNdef}
H_{\ell,N} = \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell \big( Q(\widehat{p}_{nN}) - \widehat{p}_{nN} \big).
\end{equation}
Lemma~\ref{lem:unifprobG} in the Appendix shows that for any $\delta >0$ and $\nu > 0$ there exists an $\epsilon > 0$ such that
\[
\prob\left\{ \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf) } > \delta   \right\} < \nu
\]
for all positive integers $N$, where $\|\psibf\|_\infty = \sup_{k} \abs{\psi_k}$.  Since $\widehat{\psibf}_N$ converges almost surely to zero, it follows that for any $\epsilon > 0$,
\[
\lim_{N\rightarrow\infty}\prob\left\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} = 0
\] 
and therefore $\prob\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \} < \nu$ for all sufficiently large $N$.  Now
\begin{align*}
  \prob&\left\{\abs{ G_N(\widehat{\psibf}_N) } > \delta \right\} \\
&= \prob\left\{ \abs{G_N(\widehat{\psibf}_N)} > \delta \;, \; \|\widehat{\psibf}_N\|_{\infty} < \epsilon \right\} \\
&\hspace{0.7cm} + \prob\left\{ \abs{G_N(\widehat{\psibf}_N)} > \delta  \;, \; \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} \\
&\leq \prob\left\{  \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf) } > \delta \right\} + \prob\left\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} \\
&\leq 2\nu
\end{align*}
for all sufficiently large $N$.  Since $\nu$ and $\delta$ can be chosen arbitrarily small, it follows that $G_N(\widehat{\psibf}_N)$ converges in probability to zero as $N\rightarrow\infty$, and therefore $K_{\ell,N} = H_{\ell,N} + o_P(1)$.  Lemma~\ref{lem:sumpNhilb} shows that
\begin{align*}
H_{\ell,N} =  (h-1)\sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big(C_{\ell k} + o_P(1)\big).
\end{align*}
The proof follows.
\end{IEEEproof}

\begin{lemma}\label{lem:sumpNhilb}
With $H_{\ell,N}$ defined in~\eqref{eq:HellNdef} and $\widehat{p}_{nN}$ defined in~\eqref{eq:hatpnN}, and with $h = f(-\nicefrac{1}{2})$,
\[
H_{\ell,N} = (h-1)\sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big(C_{\ell k} + o_P(1)\big),
\]
where $C_{\ell k} = \tfrac{1}{\ell + k + 1}$.
\end{lemma}
\begin{IEEEproof}
If $\abs{x} < 1$, then
\[
q_n(x)  = \round{\Phi_n + x} = \begin{cases}
1, & \Phi_n + x  \geq \nicefrac{1}{2} \\
-1, & \Phi_n + x  < -\nicefrac{1}{2} \\
0, & \text{otherwise},
\end{cases}
\]
and,
\begin{align*}
Q(x) = Eq_1(x) &=  \begin{cases}
\int_{\nicefrac{1}{2} -x }^{\nicefrac{1}{2}}f(t)\,dt, &   x \geq 0 \\
-\int_{-\nicefrac{1}{2}}^{-\nicefrac{1}{2} - x }f(t)\,dt, &  x  < 0.
\end{cases} 
% \\
% &= \begin{cases}
% 1 - F(\nicefrac{1}{2} - x ), & x  \geq 0 \\
% -F(-\nicefrac{1}{2} - x ), &  x < 0,
% \end{cases} 
\end{align*}
Because $f(\fracpart{x})$ is continuous at $-\nicefrac{1}{2}$ it follows from the mean value theorem that
\[
Q(x) = x \big( h + \zeta(x) \big),
\] 
where $\zeta(x)$ is a function that converges to zero as $x$ converges to zero.  
%The equation above is still valid when $\abs{x} \geq 1$ by an appropriate choice of the function $\zeta(x)$.  
Observe that $\abs{\widehat{p}_{nN}} \leq \sum_{k=0}^m\sabs{ \widehat{\psi}_{k,N}}$ and, since each of the $\widehat{\psi}_{k,N} \rightarrow 0$ almost surely as $N\rightarrow\infty$, it follows that $\widehat{p}_{nN} \rightarrow 0$ almost surely uniformly in $n = 1, \dots, N$ as $N\rightarrow\infty$.  Thus, $\zeta(\widehat{p}_{nN}) \rightarrow 0$ almost surely (and therefore also in probability) uniformly in $n = 1, \dots, N$ as $N\rightarrow\infty$.  Now,
\begin{align*}
Q(\widehat{p}_{nN}) - \widehat{p}_{nN} &= \widehat{p}_{nN}\big( h - 1 + \zeta(\widehat{p}_{nN}) \big) \\
&= \widehat{p}_{nN}\big( h - 1 + o_P(1) \big), 
\end{align*}
and so,
\begin{align*}
H_{\ell,N} &= \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell \widehat{p}_{nN}\big( h - 1 + o_P(1) \big)  \\
&=  \frac{1}{\sqrt{N}} \sum_{n=1}^{N}(\tfrac{n}{N})^\ell \sum_{k=0}^m (\tfrac{n}{N})^k \widehat{\psi}_{k,N} \big( h - 1 + o_P(1) \big)   \\
&= \sqrt{N}\sum_{k=0}^m \widehat{\psi}_{k,N} \frac{1}{N} \sum_{n=1}^{N} \frac{n^{\ell + k}}{N^{\ell + k + 1}}\big( h - 1 + o_P(1)  \big).
\end{align*}
The Riemann sum
\[
\frac{1}{N} \sum_{n=1}^{N} \frac{n^{\ell + k}}{N^{\ell + k + 1}} = \int_{0}^{1}x^{k+\ell+1}dx + o_P(1),
\]
and since the integral above evaluates to $\frac{1}{k+\ell+1}$, we have
\[
H_{\ell,N} = (h - 1)\sqrt{N}\sum_{k=0}^m \widehat{\psi}_{k,N}\left( \frac{1}{k+\ell+1} + o_P(1)  \right)
\]
as required.
% Put
% \[
% C_{\ell + k,N} = \sum_{n=1}^{N}\frac{n^{\ell + k}}{N^{\ell + k+1}}.
% \]
% It is well known (see for example~\cite[eq. 21]{Peleg1991_CRB_PPS_1991}) that,
% \begin{equation}\label{eq:ClkNHilbert}
% C_{\ell + k,N} = C_{\ell k} + O(N^{-1}).
% \end{equation}
% Since $\zeta(\widehat{p}_{nN}) = o_P(1)$ for all $n = 1, \dots, N$ then
% \[
% \sum_{n=1}^{N}\frac{n^{\ell + k}}{N^{\ell + k + 1}}\zeta(\widehat{p}_{nN}) 
% \]
% converges in probability to zero as $N\rightarrow\infty$.  So,
% \begin{align*}
% \sum_{n=1}^{N}\frac{n^{\ell + k}}{N^{\ell + k + 1}}\big( h - 1 +  \zeta(\widehat{p}_{nN}) \big) = (h-1)\big( C_{\ell k} + o_P(1) \big).
% \end{align*}
% The proof follows by substituting this into~\eqref{eq:HellNre}.
\end{IEEEproof}

\begin{lemma}\label{eq:convdn}
The distribution of the vector $\dbf_N$, defined in~\eqref{eq:dnandkn}, converges to the multivariate normal with zero mean and covariance matrix $\sigma^2\Cbf$.
\end{lemma}
\begin{IEEEproof}
For any constant vector $\alphabf$, let
\[
z_N = \alphabf^\prime \dbf_N = \frac{1}{\sqrt{N}} \sum_{n=1}^{N} \Phi_n \sum_{\ell=0}^{m}\alpha_\ell\left( \frac{n}{N} \right)^\ell.
\]
By Lypanov's central limit theorem $\zbf_N$ is asymptotically normally distributed with zero mean and variance
\[
\lim_{N\rightarrow\infty} \sigma^2 \frac{1}{N} \sum_{n=1}^{N}\left( \sum_{\ell=0}^{m}\alpha_\ell\left( \frac{n}{N} \right)^\ell \right)^2 = \sigma^2 \alphabf^\prime \Cbf \alphabf.
\]
By the Cramer-Wold theorem it follows that $\dbf_N$ is asymptotically normally distributed with zero mean and covariance $\sigma^2 \Cbf$. 
% Recall that the $\ell$th element of $\dbf_N$ is
% \[
% D_{\ell,N} = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}(\tfrac{n}{N})^\ell \Phi_n.
% \]
% Since the $\Phi_1, \dots, \Phi_N$ are independent, each with variance $\sigma^2$, the distribution of $D_{\ell,N}$ converges to the normal with zero mean and variance
% \[
% \lim_{N\rightarrow\infty}\var D_{\ell,N} = \sigma^2 \lim_{N\rightarrow\infty} C_{2\ell,N} = \sigma^2 C_{\ell\ell}
% \]
% by Lypanov's central limit theorem and by using~\eqref{eq:ClkNHilbert}.  The covariance between $D_{\ell,N}$ and $D_{k,N}$ is
% \[
% \covar(D_{\ell,N}, D_{k,N}) = \sigma^2 C_{\ell+k,N} \rightarrow \sigma^2 C_{\ell k}
% \]
% as $N \rightarrow \infty$, again using~\eqref{eq:ClkNHilbert}.  The proof follows.
\end{IEEEproof}

% \begin{figure*}[tb] 
%    	\centering 
%   		\includegraphics{code/lseplot-1.mps} 
%    		\caption{Performance of the LSE and LSU estimator for a polynomial phase signal of order two, i.e. $m=2$.  (Left) Mean square error (MSE) in the second order parameter $\mu_2$ for $N=10,50$ and $200$ versus the variance $\sigma_c^2$ of the $X_1, \dots, X_N$. The $X_1, \dots, X_N$ are independent and identically distributed (i.i.d.) complex Gaussian random variables.  (Right) MSE in second order parameter $\mu_2$ versus the variance $\sigma^2$ of the $\Phi_1, \dots, \Phi_N$. The $\Phi_1, \dots, \Phi_N$ are i.i.d. uniform random variables.} 
%    		\label{plot:lsetest} 
%   \end{figure*}

\section{Simulations}\label{sec:simulations}

% This section describes the results of Monte-Carlo simulations involving the least squares unwrapping estimator (LSU), the least squares estimator (LSE), Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, and the discrete polynomial phase transform (DPT)~\cite{Peleg_DPT_1995}.  In all simulations the sample sizes considered are $N = 10, 50$ and $200$ and the unknown amplitude is $\rho = 1$.  The number of replications of each experiment is 2000 and the $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ are distributed uniformly randomly over the identifiable region.

% Figure~\ref{plot:lsetest} shows the performance of the LSU estimator and LSE when $m=2$ (linear chirp). Only results for the highest order parameter $\widehat{\mu}_2$ are displayed, the results for the lower order parameters leading to similar conclusions.  The LSE is computed using the approach described in~\cite[Sec.~10.1]{McKilliam2010thesis} which is based on that of Abatzoglou~\cite{Abatzoglou_ml_chirp_1986}. The LSE is computationally intensive.  We were able to compute it within a reasonable amount of time only with $N$ less than $50$ when $m=2$. Results for the exact LSU estimator computed using the sphere decoder and the approximate LSU estimator computed using Babai's algorithm are displayed. The sphere decoder is computationally intractable when $N = 200$. It is evident that a significant performance penalty is suffered by using Babai's algorithm when the noise variance is large.  We have also displayed results using another approximate nearest point algorithm called the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder}. Our implementation requires $O(N^3\log{N})$ operations. This is more computationally expensive than Babai's algorithm but, as can be seen, the $K$-best method is more statistically accurate.

% For the simulations displayed in the left hand plot of Figure~\ref{plot:lsetest} the signal model used is that from~\eqref{eq:Y_nsamplednoisey}.  The noise terms $X_1, \dots, X_N$ are independent and identically distributed zero mean complex Gaussian random variables with independent real and imaginary parts having variance $\sigma_c^2$.  In this case the LSE is also the maximum likelihood estimator.  The \term{Cram\`{e}r-Rao lower bound} (CRB) for the variance of unbiased estimators of $\tilde{\mu}_j$ is given by Peleg and Porat~\cite{Peleg1991_CRB_PPS_1991} to be
% \[ 
% \frac{\sigma_c^2}{4\pi^2 N^{2j + 1}} C^{-1}_{jj} + o(N^{-2j-1})
% \]
% where $C^{-1}_{jj}$ is the $j$th diagonal element of the inverse of the $m+1$ by $m+1$ Hilbert matrix $\Cbf$.  Observe that the performance of the exact LSU estimator (computed using the sphere decoder) is close to the maximum likelihood LSE.  The asymptotic Cramer-Rao bound (CRB) (solid line) and the asymptotic variance predicted in Theorem~\ref{thm:asymp_proof} (dashed line) are also shown. Provided the noise variance is small enough (so that the `threshold' is avoided) the LSE variance is close to the CRB (as is expected) and the LSU variance is close to that predicted by Theorem~\ref{thm:asymp_proof}. 
% %Note the gap between the CRB and the variance predicted by Theorem~\ref{thm:asymp_proof}.  In practice this gap can be overcome using a numerical technique, such as Newton's method, with the LSU estimate as a starting point.  We have not done this here in order to show the accuracy of Theorem~\ref{thm:asymp_proof}.  The other noticeable gap is that between the threshold behaviour of the LSU estimator and the LSE.  When $N=50$ the threshold \emph{kicks in} a little earlier for the LSE than it does for the LSU. We suspect that this gap will actually \emph{close} as $N$ increases. This hypothesis is supported by the fact that this behaviour is observed in the case of frequency estimation (when $m=1$)~\cite[Sec. 9.5]{McKilliam2010thesis}\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}. Unfortunately it is not feasible to test this hypothesis because of the computational complexity of the LSE. 
% For the simulations displayed in the right hand plot of Figure~\ref{plot:lsetest} the signal model used is that from~\eqref{eq:noise_circ_poly}.  The phase noise terms $\Phi_1, \dots, \Phi_N$ are uniform random variables with variance $\sigma^2$.  In this case, the LSU estimator performs slightly better than the LSE.  This result agrees with similar observations made in circular statistics~\cite{McKilliam_mean_dir_est_sq_arc_length2010}.  The asymptotic variance predicted by Theorem~\ref{thm:asymp_proof} again accurately predicts the performance of the LSU estimator.

% Figure~\ref{plot:polyest} displays the performance of the LSU estimator, the DPT and Kitchen's estimator when $m=3$. Both the DPT and Kitchen's estimator perform poorly because they do not work correctly for some parameters in the identifiable region $B$. This behaviour has been observed previously~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase,McKilliam_polyphase_est_icassp_2011}. Again the performance of the LSU is well predicted by Theorem~\ref{thm:asymp_proof} provided that the noise variance is small enough to avoid the threshold effect and provided that $N$ is sufficiently large.  The asymptotic CRB is also plotted.

This section describes the results of Monte-Carlo simulations with the least squares unwrapping (LSU) estimator and with the discrete poylnomial phase transform (DPT) estimator of Peleg and Porat~\cite{Peleg_DPT_1995}.  The sample sizes considered are $N = 10, 50, 200$ and the unknown amplitude is $\rho = 1$.  The $X_1, \dots, X_N$ are pseudorandomly generated independent and identically distributed circularly symmetric complex Gaussian random variables with variance $\sigma_c^2$.  The coefficients $\tilde{\mubf} = [\tilde{\mu}_0, \dots, \tilde{\mu}_m]$ are distributed uniformly randomly in the identifiable region $B$.  The number of replications of each experiment is $T = 2000$ to obtain estimates $\widehat{\mubf}_1, \dots, \widehat{\mubf}_T$ and the corresponding dealiased errors $\widehat{\lambdabf}_t = \dealias(\widehat{\mubf}_T - \tilde{\mubf})$ are computed.  The sample mean square error (MSE) of the $k$th coefficient is computed according to $\tfrac{1}{T}\sum_{t=1}^T \widehat{\lambda}_{k,t}^2$ where $\widehat{\lambda}_{k,t}$ is the $k$th element of $\widehat{\lambdabf}_t$.

Figure~\ref{plot:polyest1} shows the sample MSEs obtained for a polynomial phase signal of order $m=3$.  %Results are displayed results for the zeroth and third order coefficients $\widehat{\mu}_1$ and $\widehat{\mu}_3$.  The results for $\widehat{\mu}_1$ and $\widehat{\mu}_2$ lead to similar conclusions. 
When $N = 10$ and $50$ the LSU estimator can be computed exactly using a general purpose algorithm for finding nearest lattice points called the sphere~decoder~\cite{Pohst_sphere_decoder_1981,Agrell2002,Viterbo_sphere_decoder_1999}.  This is displayed by the circles in the figures.  When $N=200$ the sphere decoder is computationally intractable and we instead use an approximate nearest point algorithm called the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder}.  This is displayed by the dots.  For the purpose of comparison we have also plotted the results for the $K$-best method when $N = 10$ and $50$.  The asymptotic variance predicted in Theorem~\ref{thm:asymp_proof} is displayed by the dashed line.  Provided the noise variance is small enough (so that the `threshold' is avoided) the sample MSE of the LSU estimator is close to that predicted by Theorem~\ref{thm:asymp_proof}.  The Cram\`{e}r-Rao lower bound for the variance of unbiased polynomial phase estimators in Gaussian noise is also plotted using the solid line~\cite{Peleg1991_CRB_PPS_1991}.  When the noise variance is small the asymptotic variance of the LSU estimator is close to the Cram\`{e}r-Rao lower bound.

\begin{figure*}[p] 
   	\centering 
  		\includegraphics{code/gaussianplot4-1.mps} 
  		\caption{Mean square error (MSE) of polynomial phase estimators for $N=10,50$ and $200$ versus the variance $\sigma_c^2$ of the complex noise terms $X_1, \dots X_N$. The polynomial phase signal has order 3, and the $X_1, \dots, X_N$ are independent and identically distributed complex Gaussian random variables. (Top left) MSE in the phase coefficient $\mu_0$.  (Top right) MSE in the frequency coefficient $\mu_1$.  (Bottom left) MSE in the quadratic coefficient $\mu_2$.  (Bottom right) MSE in the cubic coefficient $\mu_3$.} 
  		\label{plot:polyest} 
 \end{figure*}

% \section{Computational considerations}

% When $m=0$, polynomial phase estimation is equivalent to the problem of \emph{direction estimation} from circular statistics, and the least squares estimator corresponds with the \emph{extrinsic mean} (or \emph{sample circular mean})~\cite{McKilliam_mean_dir_est_sq_arc_length2010,Fisher1993,Mardia_directional_statistics}.  In this case the LSE can be computed efficiently in $O(N)$ by averaging $N$ complex numbers.  When $m=1$, polynomial phase estimation is equivalent to single frequency estimation, and the least squares estimator can be computed efficiently in $O(N\log N)$ operations by maximising the \emph{periodogram} using the fast Fourier transform followed by an appropriate numerical optimisation technique~\cite{Quinn2008maximizing_the_periodogram,Rife1974,Quinn2001}.  When $m=2$, the least squares estimator can be computed in a reasonable amount of time when $N$ is approximately less than $50$.  The LSE is computationally intensive for any $N$ when $m > 2$. By comparison the exact LSU estimator implemented using the sphere decoder can be computed quickly for any $m$ with $N$ approximately less than $50$.  In situations where $N$ is small, but high statistical accuracy is required, the sphere decoder is computationally a better choice than the LSE for polynomial phase signals of order greater than or equal to $2$.

% Babai's nearest plane algorithm requires $O(N^2)$ operations and can be run in a reasonable amount of time for any $m$ and reasonably large $N$. It should be noted that Babai's nearest plane algorithm requires a Lov\`as reduced lattice basis~\cite{Lenstra1982}.  This requires $O(N^4)$ operations to compute, but, the Lov\`as reduced basis can be computed once offline. It does not need to be computed each time the estimator is run. The $K$-best algorithm runs in a reasonable amount of time for any $m$ and when $N$ is less than about $300$. If the noise variance is small, then both the sphere decoder and the $K$-best algorithm are actually fast.  This is because both of these estimators begin with the lattice point found by Babai's nearest plane algorithm.  If this point is close to (or \emph{is}) the nearest point then both the sphere decoder and the $K$-best algorithm terminate quickly.  This is in contrast to the least squares estimator that is slow regardless of the noise variance.

% Both Kitchen's estimator and the DPT are fast to compute.  Kitchen's estimator requires only $O(N)$ arithmetic operations and the DPT requires $O(N \log N)$ operations.  However, these estimators are not as accurate as the LSE or the LSU.  They also fail to operate correctly over the entire identifiable region of polynomial phase coefficients~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}~\cite[Ch.~10]{McKilliam2010thesis}.

%\section{Discussion}

\section{Conclusion} \label{sec:conclusion}

This paper has considered the estimation of the coefficients of a noisy polynomial phase signal by least squares phase unwrapping (LSU). It has been shown that the LSU estimator is strongly consistent and asymptotically normally distributed. %Simulations are used to show that the LSU has near maximum likelihood performance in the case that the noise is additive white and Gaussian.  
The LSU estimator can be computed by finding a nearest lattice point in a lattice~\cite{McKilliam2009asilomar_polyest_lattice}.  %Here, we consider using the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder} to achieve this in reasonable time when the number of observations, $N$, is less than about 300.  
Polynomial time nearest point algorithms for these lattices exist~\cite[Sec 4.3]{McKilliam2010thesis}, but these algorithms are not fast in practice.  Instead, we considered some general purpose algorithms, the \term{sphere decoder}, \term{Babai's nearest plane algorithm} and the \term{$K$-best algorithm}. The sphere decoder and $K$-best algorithm result in accurate estimators.

The major outstanding question is whether faster nearest point algorithms exist for these specific lattices.  Considering the excellent statistical performance (both theoretically and practically) of the LSU estimator, even fast \emph{approximate} nearest point algorithms are likely to prove useful for the estimation of polynomial phase signals.

%\begin{figure}[p] 
%   	\centering 
%  		\includegraphics[width=\linewidth]{polyfig/polyphase4plot.4} 
%  		\caption{MSE in the third order coefficient $\mu_3$ versus $\var{X_n} = \sigma^2$. The true coefficients are uniformly spread in the identifiable region.} 
%  		\label{plot:polyest_projnorm_mu3} 
% \end{figure}

% \begin{figure}[p] 
%  	\centering 
% 		\includegraphics[width=\linewidth]{polyfig/polyphase4plot_unfair.4} 
% 		\caption{MSE in the third order coefficient $\mu_3$ versus $\var{X_n} = \sigma_c^2$. The coefficient have been restricted for Kitchen's estimator and the DPT.} 
% 		\label{plot:polyest_projnorm_mu3_unfair} 
%\end{figure} 
% 
% \begin{figure}[p] 
%  	\centering 
% 		\includegraphics[width=\linewidth]{polyfig/polyphase4plot_fair.4} 
% 		\caption{MSE in the third order coefficient $\mu_3$ versus $\var{X_n} = \sigma_c^2$. The DPT estimator runs at the higher sampling rate $\delta$ so that the volumes $V_{DPT}(\delta) = V_m$.} 
% 		\label{plot:polyest_projnorm_mu3_fair} 
%\end{figure}

%\addtolength{\parskip}{-0.2cm} 
%\bibliographystyle{../../bib/IEEEbib} 
\small 
\bibliography{bib}

\normalsize
\appendix

%\section{Tricks with fractional parts}
%Throughout the paper, and particularly in  have made use of a number of results involving the fractional part function

\subsection{A tightness result}

During the proof of asymptotic normality in Lemma~\ref{lem:Kconvfhalf} we made use of the following result regarding the function,
\[
G_N(\psibf) = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}\left( \tfrac{n}{N}\right)^\ell  \big(q_n(p_{nN}(\psibf)) - Q(p_{nN}(\psibf)) \big),
\]
where the functions $q_n$, $Q$ and $p_{nN}$ are defined above~\eqref{eq:GNdef} and $\ell \in \{0, 1, \dots, m\}$.

\begin{lemma}\label{lem:unifprobG}
For any $\delta >0$ and $\nu > 0$ there exists $\epsilon > 0$ such that
\[
\prob\left\{ \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf) } > \delta   \right\} < \nu %\qquad \text{for all $N > N_0$}
\]
for all positive integers $N$.
\end{lemma}

This result is related to what is called \emph{tightness} or \emph{asymptotic continuity} in the literature on empirical processes and weak convergence on metric spaces~\cite{Billingsley1999_convergence_of_probability_measures,Dudley_unif_central_lim_th_1999,Shorak_emp_proc_stat_2009,van2009empirical}.  The lemma is different from what is usually proved in the literature because the function 
\[
p_{nN}(\psibf) = \sum_{k=0}^m\left( \tfrac{n}{N}\right)^k\psi_k
\]
depends on $n$.  Nevertheless, the methods of proof from the literature can be used if we include a known result about hyperplane arrangements~\cite[Ch. 5]{Chazelle_discrepency_method_2000}\cite[Ch. 6]{Matousek_lect_disc_geom_2002}. Our proof is based on a technique called \emph{symmetrisation} and another technique called \emph{chaining} (also known as \emph{bracketing})~\cite{Pollard_asymp_empi_proc_1989,van2009empirical}.

\begin{proof}
Define the function 
\begin{align*}
f_{nN}(\psibf, \Phi_n) &= \left( \tfrac{n}{N}\right)^\ell q_n(p_{nN}(\psibf)) \\
%&= \left( \tfrac{n}{N}\right)^\ell \round{\Phi_n + p_{nN}(\psibf)} \\
&= \left( \tfrac{n}{N}\right)^\ell \round{\Phi_n + \sum_{k=0}^m\left( \tfrac{n}{N}\right)^k\psi_k}
\end{align*}
%and observe that 
%\[
%\expect f_{nN}(\psibf, \Phi_n) = \expect f_{nN}(\psibf, \Phi_1) = \left( \tfrac{n}{N}\right)^\ell Q(p_{nN}(\psibf))
%\]
so that $G_N$ can be written as
\[
G_N(\psibf) = \frac{1}{\sqrt{N}} \sum_{n=1}^{N} \big( f_{nN}(\psibf, \Phi_n) - \expect f_{nN}(\psibf, \Phi_n) \big).
\]
Let $\{g_n\}$ be a sequence of independent standard normal random variables, independent of the phase noise sequence $\{\Phi_n\}$.  Lemma~\ref{lem:symmetrisation} shows that
\[
\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} \leq \sqrt{2\pi} \; \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ Z_N(\psibf) },
\]
where 
\begin{equation}\label{eq:ZpsiCondGaussProc}
Z_N(\psibf) = \frac{1}{\sqrt{N}} \sum_{n=1}^{N} g_n f_{nN}(\psibf, \Phi_n),
\end{equation}
and where $\expect$ runs over both $\{g_n\}$ and $\{\Phi_n\}$.  Conditionally on $\{\Phi_n\}$, $\{Z_N(\psibf), \psibf \in \reals^{m+1} \}$ is a \emph{Gaussian process}, and numerous techniques exist for its analysis.  Lemma~\ref{lem:chaining} shows that for any $\kappa > 0$ there exists an $\epsilon > 0$ such that
\[
\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ Z_N(\psibf) } < \kappa.
\]
It follows that,
\[
\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)}  <  \sqrt{2\pi} \; \kappa,
\]
and by Markov's inequality,
\[
\prob \cubr{  \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} > \delta } \leq  \sqrt{2\pi} \; \frac{\kappa}{\delta},
\]
for any $\delta > 0$.  The proof follows with $\nu =  \sqrt{2\pi} \kappa/\delta$.  It remains to prove Lemmas~\ref{lem:symmetrisation}~and~\ref{lem:chaining}.

\end{proof}

The proof of the next lemma is based on a technique called \emph{symmetrisation}~\cite{Gine_Zinn_symmetrisation_1984,van2009empirical} and the proof we give closely follows that of Pollard~\cite[Section 4]{Pollard_asymp_empi_proc_1989}.

\begin{lemma} \label{lem:symmetrisation}(Symmetrisation)
\[
\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} \leq \sqrt{2\pi} \; \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ Z_N(\psibf) }.
\]
\end{lemma}
\begin{IEEEproof}
Let $\{\Phi_n^\prime\}$ be a sequence of random variables distributed identically to $\{\Phi_n\}$ and independent of both $\{\Phi_n\}$ and $\{g_n\}$.  Write $\expect_\Phi$ to denote expectation conditional on $\{\Phi_n\}$.  %That is, expectation taken over $\{g_n\}$ and $\{\Phi_n^\prime\}$ treating $\{\Phi_n\}$ as fixed.  
Since $\Phi_n$ and $\Phi_n^\prime$ are identically distributed,
\[
\expect f_{nN}(\psibf, \Phi_n) = \expect_\Phi f_{nN}(\psibf, \Phi_n^\prime),
\]
so that,
\begin{align*}
G_N(\psibf) &= \frac{1}{\sqrt{N}} \sum_{n=1}^{N} \big( f_{nN}(\psibf, \Phi_n) - \expect_\Phi f_{nN}(\psibf, \Phi_n^\prime) \big)  \\
&= \expect_\Phi \frac{1}{\sqrt{N}} \sum_{n=1}^{N} ( f_{nN} - f_{nN}^\prime ),
\end{align*}
where, for notational convenience, we put 
\[
f_{nN} = f_{nN}(\psibf, \Phi_n) \qquad \text{and} \qquad  f_{nN}^\prime = f_{nN}(\psibf, \Phi_n^\prime).
\] 
Taking absolute values followed by supremums,
\begin{align*}
 \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} &= \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{\frac{1}{\sqrt{N}} \expect_\Phi \sum_{n=1}^{N} (f_{nN} - f_{nN}^\prime )} \\
&\leq \sup_{\|\psibf\|_{\infty} < \epsilon} \expect_\Phi  \abs{ \frac{1}{\sqrt{N}} \sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) },
\end{align*}
the upper bound following from Jensen's inequality.  Since $\sup \expect \abs{\dots} \leq \expect \sup \abs{\dots}$, %the bound can be increased by moving $\expect_{\Phi}$ outside the supremum,
\begin{equation}\label{eq:Eoutsidesup}
 \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} \leq \expect_\Phi \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} ( f_{nN} - f_{nN}^\prime ) }.
\end{equation}
Applying $\expect$ to both sides gives the inequality
\begin{equation}\label{eq:Gninequfin}
 \expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} \leq  \expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) },
\end{equation}
since $\expect \expect_\Phi$ is equivalent to $\expect$.   Let
\[
\sigma_n = \frac{g_n}{\abs{g_n}},
\]
and put $\sigma_n = 1$ if $g_n = 0$.  The $\sigma_n$ are thus independent with 
\[
\prob\{ \sigma_n = -1 \} = \tfrac{1}{2} \qquad \text{and} \qquad  \prob\{ \sigma_n = 1 \} = \tfrac{1}{2}.
\]  
As $\Phi_n$ and $\Phi_n^\prime$ are independent and identically distributed, the random variable $f_{nN} - f_{nN}^\prime$ is symmetrically distributed about zero, and is therefore distributed identically to $\sigma_n( f_{nN} - f_{nN}^\prime)$.  Thus,
%\[
%\sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) \qquad \text{and} \qquad \sum_{n=1}^{N}  \sigma_n(f_{nN} - f_{nN}^\prime)
%\]
%are identically distributed, and therefore
\begin{align*}
 \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  &\abs{  \frac{1}{\sqrt{N}}\sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) } \\
&= \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{  \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n ( f_{nN} - f_{nN}^\prime) }.
\end{align*}
By the triangle inequality
\[
\abs{ \sum_{n=1}^{N} \sigma_n (f_{nN} - f_{nN}^\prime)} \leq \abs{ \sum_{n=1}^{N} \sigma_n f_{nN} } + \abs{ \sum_{n=1}^{N}  \sigma_n f_{nN}^\prime },
\]
and it follows from~\eqref{eq:Gninequfin}, that
\begin{align*}
 \expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} &\leq \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} } \\
&\hspace{1cm} + \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{\frac{1}{\sqrt{N}} \sum_{n=1}^{N}  \sigma_n  f_{nN}^\prime } \\
&= 2 \; \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} }.
\end{align*}
Since $g_n$ is a standard normal random variable, $\expect \abs{g_n} = \expect_\Phi \abs{g_n} = \sqrt{2/\pi}$,
and
\begin{align*}
\expect \sup_{\|\psibf\|_{\infty} < \epsilon}  &\abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} } \\
&= \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} \sqrt{\frac{\pi}{2}}\expect_\Phi \abs{g_n}} \\
&\leq \sqrt{\frac{\pi}{2}} \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{ \sqrt{N}}\sum_{n=1}^{N} \sigma_n \abs{g_n}  f_{nN}},
\end{align*}
the last line following from Jensen's inequality and by moving $\expect_{\Phi}$ outside the supremum similarly to~\eqref{eq:Eoutsidesup}. As $g_n = \sigma_n \abs{g_n}$, it follows that
\begin{align*}
\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{G_N(\psibf)}  &\leq 2\sqrt{\frac{\pi}{2}} \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} g_n  f_{nN}} \\
&= \sqrt{2\pi} \; \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ Z_N(\psibf) } 
\end{align*}
as required.
\end{IEEEproof}

\begin{lemma} \label{lem:chaining}
For any $\kappa > 0$ there exists $\epsilon > 0$ such that
\[
\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ Z_N(\psibf) } < \kappa,
\]
where $Z_N(\psibf)$ is defined by~\eqref{eq:ZpsiCondGaussProc}.
\end{lemma}
\begin{IEEEproof}
Without loss of generality, assume that $\epsilon < \frac{1}{m+1}$.  Lemma~\ref{lem:chaining2} shows that
\begin{equation}\label{eq:supZCK1} 
\expect_\Phi \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ Z_N(\psibf) } \leq K_1 \sqrt{C_\epsilon(\{\Phi_n\})}
\end{equation}
where $K_1$ is a finite, positive constant, and $C_\epsilon(\{\Phi_n\})$ is the average number of times $\abs{\Phi_1}, \dots, \abs{\Phi_N}$ is greater than or equal to $\nicefrac{1}{2} - (m+1)\epsilon$.  That is,
\begin{equation}\label{eq:Cedefn}
C_\epsilon(\{\Phi_n\}) = \frac{1}{N} \sum_{n=1}^{N} I_\epsilon(\abs{\Phi_n})
\end{equation}
where $I_\epsilon(\abs{\Phi_n})$ is 1 when $\abs{\Phi_n} \geq \nicefrac{1}{2} - (m+1)\epsilon$ and zero otherwise.  Recall that $f$ is the probability density function of $\Phi_n$, and (by assumption in Theorem~\ref{thm:asymp_proof}) that $f(\fracpart{x})$ is continuous at $x = -\nicefrac{1}{2}$.  Because of this, the expected value of $C_\epsilon(\{\Phi_n\})$ is small when $\epsilon$ is small, since
\begin{align*}
\expect C_\epsilon(\{\Phi\}) &= \frac{1}{N} \sum_{n=1}^{N} \expect I_\epsilon(\abs{\Phi_n}) \\
&= \prob\cubr{\abs{\Phi_1} \geq \nicefrac{1}{2} - (m+1)\epsilon} \\
&= \int_{-1/2}^{-1/2 + (m+1)\epsilon} f(\phi) d\phi + \int_{1/2 - (m+1)\epsilon}^{1/2} f(\phi) d\phi \\
&= \int_{-1/2 + (m+1)\epsilon}^{1/2 - (m+1)\epsilon} f(\fracpart{\phi}) d\phi \\
&= 2(m+1)\epsilon(f(-\nicefrac{1}{2}) + o(1))
\end{align*}
where $o(1)$ goes to zero as $\epsilon$ goes to zero, by the mean value theorem.  Since $\sqrt{\cdot}$ is a concave function on the positive real line, and since $C_{\epsilon}(\{\Phi_n\})$ is always positive,
\[
\expect \sqrt{C_\epsilon(\{\Phi_n\})} \leq  \sqrt{\expect  C_\epsilon(\{\Phi_n\} )} < \sqrt{K_2 \epsilon }
\]
by Jensen's inequality, for some constant $K_2$.  Applying $\expect$ to both sides of~\eqref{eq:supZCK1},
\[
\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ Z_N(\psibf) } \leq K_1 \sqrt{\expect C_\epsilon(\{\Phi_n\})} < K_1 \sqrt{K_2 \epsilon}.
\]
Choosing $\epsilon = \kappa^2/(K_1^2 K_2)$ completes the proof.  It remains to prove Lemma~\ref{lem:chaining2}.
\end{IEEEproof}

The proofs of Lemmas~\ref{lem:chaining2}~and~\ref{lem:chaining3} are based on a technique called \emph{chaining} (or \emph{bracketing})~\cite{Dudley_unif_central_lim_th_1999,Ossiander_clt_bracketing_1984,Pollard_asymp_empi_proc_1989,Pollard_new_ways_clts_1986}.  The proofs here follow those of Pollard~\cite[Section 3]{Pollard_asymp_empi_proc_1989}.  Within the remaining proofs, we are interested in expectations conditional on the phase noise sequence $\{\Phi_n\}$, so for the purpose of these lemmas, $\{\Phi_n\}$ is treated as a fixed realisation.  For notational convenience we use the abbreviations 
\[
C_\epsilon = C_\epsilon(\{\Phi_n\}) \qquad \text{and} \qquad f_{nN}(\psibf) = f_{nN}(\psibf, \Phi_n)
\]
in what follows.  As in Lemma~\ref{lem:chaining} we assume, without loss of generality, that $\epsilon < \tfrac{1}{m+1}$.  %This comes at no loss since Lemma~\ref{lem:chaining} only asks for the existence of an $\epsilon > 0$.

\begin{lemma}\label{lem:chaining2}
There exists a positive constant $K_1$ such that
\[
\expect_\Phi \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ Z_N(\psibf) } \leq K_1 \sqrt{C_\epsilon}.
\]
\end{lemma}
\begin{IEEEproof}
Let
\[
B_\epsilon = \{\xbf \in \reals \mid \|\xbf\|_\infty < \epsilon \}.
\]
%We are now required to prove that
%\[
%\expect_\Phi \sup_{\psibf \in B_\epsilon} \abs{ Z_N(\psibf) } \leq K_1 \sqrt{C_\epsilon}.
%\]
For each non negative integer $k$, let $T_\epsilon(k)$ be a discrete subset of $\reals^{m+1}$ with the property that for every $\psibf \in B_\epsilon$ there exists some $\psibf^* \in T_\epsilon(k)$ such that the pseudometric
\[
d(\psibf, \psibf^*) = \sum_{n = 1}^N \big( f_{nN}(\psibf) - f_{nN}(\psibf^*) \big)^2 \leq 2^{-k} C_\epsilon N.
\]
We specifically define $T_\epsilon(0)$ to contain a single point, that being the origin $\zerobf$.  Defined this way, $T_\epsilon(0)$ satisfies the inequality above because
\[
d(\psibf, \zerobf) = \sum_{n = 1}^N f_{nN}(\psibf)^2  \leq C_\epsilon N ,
\]
for all $\psibf \in B_\epsilon$, as a result of Lemma~\ref{lem:epslmlemma}.

The existence of $T_\epsilon(k)$ for each positive integer $k$ will be proved in Lemma~\ref{lem:metricentropy}.  It is worth giving some intuition regarding $T_\epsilon(k)$.  If we place a `ball' of radius $2^{-k} C_\epsilon N$  with respect to the pseudometric $d(\cdot, \cdot)$ around each point in $T_\epsilon(k)$, then, by definition, the union of these balls is a superset of $B_\epsilon$.  The balls are said to \emph{cover} $B_\epsilon$ and $T_\epsilon(k)$ is said to form a \emph{covering} of $B_\epsilon$~\cite[Section 1.2]{Dudley_unif_central_lim_th_1999}.  The minimum number of such balls required to cover $B_\epsilon$ is called a \emph{covering number} of $B_\epsilon$.  In Lemma~\ref{lem:metricentropy} we show that no more than $K_3 2^{(m+1)k}$ balls of radius $2^{-k} C_\epsilon N$ are required to cover $B_\epsilon$, where $K_3$ is a constant, independent of $N$ and $\epsilon$.  We make use of this upper bound in Lemma~\ref{lem:chaining3}.

We now continue the proof.  For all $\psibf \in \reals^{m+1}$, 
\[
f_{nN}(\psibf) = \left( \tfrac{n}{N}\right)^\ell \round{\Phi_n + p_{nN}(\psibf)}
\]
is a multiple of $N^{-\ell}$ and so $d(\psibf, \psibf^*)$ is a multiple of $N^{-2\ell}$.  When $2^k > C_\epsilon N^{1+2\ell}$ we have 
\[
0 \leq d(\psibf, \psibf^*) < 2^{-k} C_\epsilon N < N^{-2\ell},
\]
and so we must have $d(\psibf, \psibf^*) = 0$.  Consequently $f_{nN}(\psibf) = f_{nN}(\psibf^*)$  for every $n = 1, \dots, N$, and therefore $Z_N(\psibf) - Z_N(\psibf^*) = 0$.  Thus,
\[
\sup_{\|\psibf\|_{\infty} < \epsilon } \abs{ Z_N(\psibf) } = \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) }
\]
for all $k$ large enough that $2^{k} > C_\epsilon N^{1+2\ell}$.  So, to analyse the supremum of $Z_N(\psibf)$ over the continuous interval $B_\epsilon$ it is enough to analyse the supremum over the discrete set $T_\epsilon(k)$ for large $k$.  Lemma~\ref{lem:chaining3} shows that 
\[
\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) } \leq \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{\sqrt{ i A_1 + A_2}}{2^{i/2}} < \infty
\]
for every positive integer $k$, where $A_1$ and $A_2$ are the constants,
\begin{equation}\label{eq:A1andA2}
A_1 = 18(m+1)\log 2\ \qquad \text{and} \qquad A_2 = 18\log K_3
\end{equation}
and $\log(\cdot)$ is the natural logarithm.  The lemma holds with $K_1 =  \sum_{i=1}^{\infty}2^{-i/2}\sqrt{ i A_1 + A_2 }$.
\end{IEEEproof}

\begin{lemma}\label{lem:epslmlemma}
For $\epsilon < \frac{1}{m+1}$ and all $\psibf \in B_\epsilon$ and $n = 1, \dots, N$,
\[
f_{nN}(\psibf)^2 \leq \abs{f_{nN}(\psibf)} \leq I_\epsilon(\abs{\Phi_n})
\]
and consequently,
\begin{equation}\label{eq:sumf2absfCe}
\sum_{n=1}^{N}f_{nN}(\psibf)^2 \leq \sum_{n=1}^{N}\abs{f_{nN}(\psibf)} \leq N C_\epsilon.
\end{equation}
\end{lemma}
\begin{IEEEproof}
Recall that
\[
f_{nN}(\psibf) = \left( \tfrac{n}{N}\right)^\ell \round{\Phi_n + p_{nN}(\psibf)}.
\]
Because $\abs{\psi_i} < \epsilon$ for all $i = 0, \dots, m$, 
\begin{equation}\label{eq:pnNsmall}
\abs{p_{nN}(\psibf)} = \abs{\sum_{i=0}^{m}\big(\tfrac{n}{N}\big)^i \psi_i} \leq (m+1)\epsilon < 1.
\end{equation}
%for all $n = 1, \dots, N$.  
Since $\Phi_n \in [-\nicefrac{1}{2}, \nicefrac{1}{2})$, it follows that $f_{nN}(\psibf)$ equals either $-(\tfrac{n}{N})^\ell$, $(\tfrac{n}{N})^\ell$ or $0$ and so
\begin{equation}\label{eq:fnNleqfnNs}
   f_{nN}(\psibf)^2 \leq \abs{f_{nN}(\psibf)} \leq 1.
\end{equation}
Whenever $f_{nN}(\psibf) \neq 0$ we must have 
\[
\abs{\Phi_n} \geq \nicefrac{1}{2} - \abs{p_{nN}(\psibf)} \geq \nicefrac{1}{2} - (m+1)\epsilon
\]
and $I_\epsilon(\abs{\Phi_n}) = 1$.  Thus, for all $n = 1, \dots, N$,
\[
f_{nN}(\psibf)^2 \leq \abs{f_{nN}(\psibf)} \leq I_\epsilon(\abs{\Phi_n}). %\qquad \text{for all $n = 1, \dots, N$.}
\]
Summing the terms in this inequality over $n = 1, \dots, N$ and using~\eqref{eq:Cedefn} gives~\eqref{eq:sumf2absfCe}.
\end{IEEEproof}

\begin{lemma}\label{lem:chaining3}(Chaining)
For all positive integers $k$,
\[
\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) } \leq \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{\sqrt{ i A_1 + A_2 }}{2^{i/2}},
\]
where the constants $A_1$ and $A_2$ are defined in~\eqref{eq:A1andA2}.
\end{lemma}
\begin{IEEEproof}
Let $b_k$ be a function that maps each $\psibf \in T_\epsilon(k)$ to $b_k(\psibf) \in T_\epsilon(k-1)$ such that 
\[
d(\psibf,b_k(\psibf)) \leq 2^{1-k} C_\epsilon N.
\]
The existence of the function $b_k$ is guaranteed by the definition of $T_\epsilon(k)$.  By the triangle inequality,
\[
\abs{ Z_N(\psibf) } \leq \abs{ Z_N(b_k(\psibf)) } + \abs{ Z_N(\psibf) - Z_N(b_k(\psibf))  },
\]
and by taking supremums on both sides,
\begin{align}
&\sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) } \nonumber \\ %\leq \sup_{\psibf \in T_\epsilon(k) } \big( \abs{ Z_N(\psibf^*) } + \abs{ Z_N(\psibf) - Z_N(\psibf^*)  } \big) \nonumber \\
&\leq \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(b_k(\psibf)) } + \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) - Z_N(b_k(\psibf))  } \nonumber \\
&\leq \sup_{\psibf \in T_\epsilon(k-1)} \sabs{ Z_N(\psibf) } + \sup_{\psibf \in T_\epsilon(k)} \sabs{ Z_N(\psibf) - Z_N(b_k(\psibf)) }, \label{eq:supZchain}
\end{align}
the last line following since $b_k(\psibf) \in T_\epsilon(k-1)$ and so 
\[
\sup_{\psibf \in T_\epsilon(k)} \abs{ Z_N(b_k(\psibf)) } \leq \sup_{\psibf \in T_\epsilon(k-1)} \abs{ Z_N(\psibf) }.
\]
Conditional on $\Phi_1, \Phi_2, \dots$ the random variable
\[
X(\psibf) = Z_N(\psibf) - Z_N(b_k(\psibf))
\]
has zero mean, and is normally distributed with variance
\begin{align*}
\sigma_X^2 &= \expect_{\Phi} \frac{1}{N} \sum_{n=1}^N g_n^2 ( f_{nN}(\psibf) - f_{nN}(b_k(\psibf)) )^2 \\
&= \frac{1}{N} \sum_{n=1}^N ( f_{nN}(\psibf) - f_{nN}(b_k(\psibf)) )^2 \\
&= d(\psibf, b_k(\psibf)) \leq 2^{1-k} C_\epsilon,
\end{align*}
since $\expect_{\Phi} g_n^2 = 1$.  Using Lemma~\ref{lem:maxineq}, 
\begin{align*}
\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k)} \sabs{X(\psibf)} &\leq 3\sqrt{2^{1-k} C_\epsilon \log\sabs{T_\epsilon(k)}} \\
&\leq \sqrt{C_\epsilon}\frac{\sqrt{  k A_1 + A_2 }}{2^{k/2}}
\end{align*}
because $\log\abs{T_\epsilon(k)} \leq k(m+1)\log 2 +  \log K_3$.  Taking expectations on both side of~\eqref{eq:supZchain},
\begin{align*}
\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) } &\leq \expect_{\Phi} \sup_{\psibf \in T_\epsilon(k-1)} \sabs{ Z_N(\psibf) } \\
&\hspace{1cm} + \sqrt{C_\epsilon}\frac{\sqrt{  k A_1 + A_2 }}{2^{k/2}},
\end{align*}
which involves a recursion in $k$.  By unravelling the recursion, and using the fact $T_\epsilon(0)$ contains only the origin, and therefore 
\[
\expect_{\Phi}\sup_{\psibf \in T_\epsilon(0)} \abs{ Z_N(\psibf) } = \expect_{\Phi}\abs{ Z_N(0) } = 0,
\]
we obtain
\begin{align*}
\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) } < \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{\sqrt{  k A_1 + A_2 }}{2^{k/2}}
\end{align*}
as required.
\end{IEEEproof}

\begin{lemma}\label{lem:maxineq}(Maximal inequalities)
Suppose $X_1, \dots, X_N$ are zero mean Gaussian random variables each with variance less than some positive constant $K$, then
\[
\expect \sup_{n = 1, \dots, N} \abs{X_n} \leq 3  \sqrt{K \log N}
\]
where $\log N$ is the natural logarithm of $N$.
\end{lemma}
\begin{IEEEproof}
This result is well known, see for example~\cite[Section 3]{Pollard_asymp_empi_proc_1989}  
\end{IEEEproof}

\begin{lemma}\label{lem:metricentropy}(Covering numbers)
For $k \in \ints$ there exists a discrete set $T_\epsilon(k) \subset \reals^{m+1}$ with the property that, for every $\psibf \in B_\epsilon$, there is a $\psibf^* \in T_\epsilon(k)$ such that,
\[
d(\psibf,\psibf^*) = \sum_{n = 1}^N \big( f_{nN}(\psibf) - f_{nN}(\psibf^*) \big)^2 \leq \frac{C_\epsilon N}{2^k}.
\]
The number of elements in $T_\epsilon(k)$ is no more than $K_3 2^{(m+1)k}$ where $K_3$ is a positive constant, independent of $N$, $\epsilon$ and $k$.
\end{lemma}

Before we give the proof of this lemma we need some results from the literature on hyperplane arrangements and what are called \emph{$\epsilon$-cuttings}~\cite{Chazelle_discrepency_method_2000,Matousek_lect_disc_geom_2002}.  Let $H$ be a set of $m$-dimensional affine hyperplanes lying in $\reals^{m+1}$.  By \emph{affine} it is meant that  the hyperplanes need not pass through the origin.  For each hyperplane $h \in H$ let $D(h)$ and its complement $\bar{D}(h)$ be the corresponding half spaces of $\mathbb{R}^{m+1}$.  For a point $\xbf \in \mathbb{R}^{m+1}$, let
\begin{equation}\label{eq:bhsi}
b(h,\xbf) = \begin{cases} 
1 & \xbf \in D(h) \\
0 & \xbf \in \bar{D}(h).
\end{cases}
\end{equation}
Note that $b(h,\xbf)$ is piecewise constant in $\xbf$.  Two points $\xbf$ and $\ybf$ from $\reals^{m+1}$ are in the same halfspace of $h$ if and only if $b(h, \xbf) = b(h, \ybf)$.  Then the pseudometric
\[
\sigma(\xbf,\ybf) = \sum_{h \in H} |b(h,\xbf) - b(h,\ybf)|
\]
is the number of hyperplanes in $H$ that pass between the points $\xbf$ and $\ybf$.  %Observe that $\sigma(\xbf, \ybf)$ is peicewise constant in both $\xbf$ and $\ybf$.

The next theorem considers the partitioning of $\reals^{m+1}$ into subsets so that not too many hyperplanes intersect with any subset.  Proofs can be found in Theorem 5.1 on page 206 of~\cite{Chazelle_discrepency_method_2000} and also Theorem 6.5.3 on page 144 of~\cite{Matousek_lect_disc_geom_2002}.

\begin{theorem}\label{thm:epscutting}
There exists a constant $K$, independent of the set of hyperplanes $H$, such that for any positive real number $r$, we can partition $\reals^{m+1}$ into $K r^{m+1}$ generalised $(m+1)$-dimensional simplices with the property that no more than $\abs{H}/r$ hyperplanes from $H$ pass through the interior of any simplex.
\end{theorem}

By the phrase `There exists a constant $K$, independent of the set of hyperplanes $H$', it is meant that the constant $K$ is valid for every possible set of hyperplanes in $\reals^{m+1}$, regardless of the number of hyperplanes or their position and orientation.  A \emph{generalised} $(m+1)$-dimensional simplex is the region defined by the intersection of $m+2$ half spaces in $\reals^{m+1}$.  Note that a generalised simplex (unlike an ordinary simplex) can be unbounded.  For our purposes Theorem~\ref{thm:epscutting} is important because of the following corollary.

\begin{corollary}\label{cor:epscutting}
There exists a constant $K$, independent of the set of hyperplanes $H$, such that for every positive real number $r$ there is a discrete subset $T \subset \reals^{m+1}$ containing no more than $K r^{m+1}$ elements with the property that for every $\xbf \in \reals^{m+1}$ there exists $\ybf \in T$ with $\sigma(\xbf, \ybf) \leq \abs{H}/r$.
\end{corollary}
\begin{IEEEproof}
Let $C$ be the set of generalised simplices constructed according to Theorem~\ref{thm:epscutting}.  Define $T$ as a set containing precisely one point from the interior of each simplex in $C$.  Let $\xbf \in \reals^{m+1}$.  Since $b(h, \xbf)$ is piecewise constant for each $h \in H$, and since $C$ partitions $\reals^{m+1}$, there must exist a simplex $c \in C$ with a point $\zbf$ in its interior such that $b(h,\zbf) = b(h, \xbf)$ for all $h \in H$, and correspondingly $\sigma(\zbf, \xbf) = 0$.  Let $\ybf$ be the element from $T$ that is in the interior of $c$.  Since at most $\abs{H}/r$ hyperplanes cross the interior of $c$ there can be at most $\abs{H}/r$ hyperplanes between $\zbf$ and $\ybf$, and so $\sigma(\zbf, \ybf) \leq \abs{H}/r$.  Now, 
\[
\sigma(\xbf, \ybf) < \sigma(\zbf, \xbf) + \sigma(\zbf, \ybf) \leq \frac{\abs{H}}{r}
\]
follows from the triangle inequality.
\end{IEEEproof}

The previous corollary ensures that we can \emph{cover} $\reals^{m+1}$ using $K r^{m+1}$ `balls' of radius $\abs{H}/r$ with respect to the pseudometric $\sigma(\xbf, \ybf)$.  The balls are placed at the positions defined by points in the set $T$, and these points could be anywhere in $\reals^{m+1}$.  The next corollary asserts that we can cover a subset of $\reals^{m+1}$, by placing the balls at points only within this subset.

\begin{corollary}\label{cor:epscuttingsubset}
Let $B$ be a subset of $\reals^{m+1}$.  There exists a constant $K$, independent of the set of hyperplanes $H$, such that for every positive real number $r$ there is a discrete subset $T_B \subset B$ containing no more than $K r^{m+1}$ elements with the property that for every $\xbf \in B$ there is a $\ybf \in T_B$ with $\sigma(\xbf, \ybf) \leq \abs{H}/r$.
\end{corollary}
\begin{IEEEproof}
Let $C$ be the set of generalised simplices constructed according to Theorem~\ref{thm:epscutting} and let $C_B$ be the subset of those indices that intersect $B$.  Let $T_B$ contain a point from $c \cap B$ for each simplex $c \in C_B$.  The proof now follows similarly to Corollary~\ref{cor:epscutting}.
\end{IEEEproof}

We are now ready to prove Lemma~\ref{lem:metricentropy}.

\begin{IEEEproof} (Lemma~\ref{lem:metricentropy})
Put 
\[
g_{nN}(\psibf) = (\tfrac{N}{n})^\ell f_{nN}(\psibf) = \round{ \Phi_n + p_{nN}(\psibf)},
\]
and let
\[
d_g(\psibf, \psibf^*) = \sum_{n=1}^{N} \big( g_{nN}(\psibf) - g_{nN}(\psibf^*) \big)^2.
\]
We have $d(\psibf, \psibf^*) \leq d_g(\psibf, \psibf^*)$, and so it suffices to prove the Lemma with $d$ replaced by $d_g$.  
%It will be more convenient to work with $d_g$ rather than $d$.

From~\eqref{eq:pnNsmall}, 
\[
\abs{p_{nN}(\psibf)} \leq (m+1)\epsilon < 1.
\]
Since $\Phi_n \in [-\nicefrac{1}{2}, \nicefrac{1}{2})$, when $\Phi_n \geq 0$,
\[
g_{nN}(\psibf) = \begin{cases}
1 & p_{nN}(\psibf) \geq \nicefrac{1}{2} - \Phi_n \\
0 & \text{otherwise},
\end{cases}
\]
and when $\Phi_n < 0$,
\[
g_{nN}(\psibf) = \begin{cases}
-1 & p_{nN}(\psibf) < -\nicefrac{1}{2} - \Phi_n \\
0 & \text{otherwise}.
\end{cases}
\]
Thus, $( g_{nN}(\psibf) - g_{nN}(\psibf^*) )^2$ is either equal to one when $g_{nN}(\psibf) \neq g_{nN}(\psibf^*)$ or zero when $g_{nN}(\psibf) = g_{nN}(\psibf^*)$.  Now $g_{nN}(\psibf) \neq 0$ only if 
\[
\abs{\Phi_n} \geq \nicefrac{1}{2} - \abs{p_{nN}(\psibf)} \geq \nicefrac{1}{2} - (m+1)\epsilon,
\]
that is, only if $I_\epsilon(\Phi_n) = 1$.  Let
\[
A = \{ n \in \{1, \dots, N\} \mid I_\epsilon(\Phi_n) = 1 \}
\]
be the subset of the indices where $I_\epsilon(\Phi_n) = 1$.  By definition the number of elements in $A$ is $C_\epsilon N$ (see~\eqref{eq:Cedefn}).  If both $\psibf$ and $\psibf^*$ are in $B_\epsilon$, then 
\[
( g_{nN}(\psibf) - g_{nN}(\psibf^*) )^2 \neq 0 
\]
only if $n \notin A$.  Thus,
\begin{align*}
d_g(\psibf, \psibf^*) &= \sum_{n=1}^{N} \big( g_{nN}(\psibf) - g_{nN}(\psibf^*) \big)^2 \\
&= \sum_{n \in A} \big( g_{nN}(\psibf) - g_{nN}(\psibf^*) \big)^2.
\end{align*}

We now use Corollary~\ref{cor:epscuttingsubset}.  Let $h_n$ be the $m$ dimensional hyperplane in $\reals^{m+1}$ satisfying
\[
p_{nN}(\psibf) =  \sum_{i=0}^{m} \big(\tfrac{n}{N}\big)^i \psi_i = \tfrac{1}{2} \sign{\Phi_n} - \Phi_n
\]
where $\sign{\Phi_n}$ is equal to $1$ when $\Phi_n \geq 0$ and $-1$ otherwise.  The hyperplane $h_n$ divides $\reals^{m+1}$ into two halfspaces, $D(h_n)$ and its complement $\bar{D}(h_n)$.  If $\psibf$ and $\psibf^*$ are in the same halfspace, then $\abs{ b(h_n,\psibf) - b(h_n,\psibf^*) } = 0$ and $g_{nN}(\psibf) = g_{nN}(\psibf^*)$, and therefore $( g_{nN}(\psibf) - g_{nN}(\psibf^*) )^2 = 0$.  Otherwise, if $\psibf$ and $\psibf^*$ are in different halfspaces, then $\abs{ b(h_n,\psibf) - b(h_n,\psibf^*) } = 1$ and $g_{nN}(\psibf) \neq g_{nN}(\psibf^*)$, and therefore $( g_{nN}(\psibf) - g_{nN}(\psibf^*) )^2 = 1$.  Thus, 
\[
(g_{nN}(\psibf) - g_{nN}(\psibf^*) )^2 = \abs{ b(h_n,\psibf) - b(h_n,\psibf^*) }
\]
for $n = 1, \dots,  N$.  Let $H$ be the finite set of hyperplanes $\{ h_n, n \in A\}$ and observe that the number of hyperplanes is $\abs{H} = \abs{A} = C_\epsilon N$.  When both $\psibf$ and $\psibf^*$ are inside $B_\epsilon$, $d_g$ can be written as
\begin{align*}
d_g(\psibf, \psibf^*) &= \sum_{n\in A} \abs{ b(h_n,\psibf) - b(h_n,\psibf^*) } \\
&=  \sum_{h \in H} \abs{ b(h,\psibf) - b(h,\psibf^*) } = \sigma(\psibf, \psibf^*).
\end{align*}
That is, when both $\psibf, \psibf^* \in B_\epsilon$, $d_g(\psibf, \psibf^*)$ is the number of hyperplanes from $H$ that pass between the points $\psibf$ and $\psibf^*$.

It follows from Corollary~\ref{cor:epscuttingsubset} that for any positive $r$ there exists a finite subset $T_B$ of $B_\epsilon$ containing at most $K_3 r^{m+1}$ elements, such that for every $\psibf \in B_\epsilon$ there is a $\psibf^* \in T_B$ with 
\[
d_g(\psibf, \psibf^*) = \sigma(\psibf, \psibf^*) \leq \frac{\abs{H}}{r} = \frac{\abs{A}}{r} = \frac{C_\epsilon N}{r}.
\]
Putting $r = 2^k$ and choosing $T_\epsilon(k) = T_B$ completes the proof.
\end{IEEEproof}

%\begin{figure}[p] 
%   	\centering 
%  		\includegraphics[width=\linewidth]{polyfig/polyphase4plot.4} 
%  		\caption{MSE in the third order coefficient $\mu_3$ versus $\var{X_n} = \sigma^2$. The true coefficients are uniformly spread in the identifiable region.} 
%  		\label{plot:polyest_projnorm_mu3} 
% \end{figure}

% \begin{figure}[p] 
%  	\centering 
% 		\includegraphics[width=\linewidth]{polyfig/polyphase4plot_unfair.4} 
% 		\caption{MSE in the third order coefficient $\mu_3$ versus $\var{X_n} = \sigma_c^2$. The coefficient have been restricted for Kitchen's estimator and the DPT.} 
% 		\label{plot:polyest_projnorm_mu3_unfair} 
%\end{figure} 
% 
% \begin{figure}[p] 
%  	\centering 
% 		\includegraphics[width=\linewidth]{polyfig/polyphase4plot_fair.4} 
% 		\caption{MSE in the third order coefficient $\mu_3$ versus $\var{X_n} = \sigma_c^2$. The DPT estimator runs at the higher sampling rate $\delta$ so that the volumes $V_{DPT}(\delta) = V_m$.} 
% 		\label{plot:polyest_projnorm_mu3_fair} 
%\end{figure}

\end{document}

### 2012-10-29 paper.tex
added to two papers

@@
 %It was shown by Monte Carlo simulation that this produces a remarkably accurate estimator.  
-In this paper we consider an estimator that performs phase unwrapping in a least squares manner.
-We describe the asymptotic properties of this estimator, showing that it is strongly consistent and asymptotically normally distributed.
+In this two part series of papers we consider an estimator that performs phase unwrapping in a least squares manner.
+In this first part we describe conditions for the identifiability of polynomial phase signals and we prove the strong consistency of our unwrapping estimator.
+%We describe the asymptotic properties of this estimator, showing that it is strongly consistent and asymptotically normally distributed.
 %We hypothesise that the estimator produces very near maximum likelihood performance.
@@
 
-An obvious estimator of the unknown coefficients is the least squares estimator (LSE).
+%An obvious estimator of the unknown coefficients is the least squares estimator (LSE).
 This is also the maximum likelihood estimator (MLE) when the noise sequence $\{X_n\}$ is white and Gaussian.
@@
 
-In this paper we consider the estimator that results from unwrapping the phase in a least squares manner.
+One estimation approach attempts to perform polynomial regression on the phase of the signal~\cite{Slocumb_polynomial_1994,Tretter1985,Kitchen_polyphase_unwrapping_1994,Djuric_phase_unwrap_chirp_1990,Morelande_bayes_unwrapping_2009_tsp}.
+This is complicated by the fact that the phase is \emph{wrapped} modulo $2\pi$ and must be \emph{unwrapped} before regression can be performed.
+In this two part series of papers we consider the estimator that results from unwrapping the phase in a least squares manner.
 We call this the \emph{least squares unwrapping estimator} (LSU)~\cite{McKilliam2009asilomar_polyest_lattice, McKilliamFrequencyEstimationByPhaseUnwrapping2009}\cite[Chap.
@@
 %We also apply another known approximate nearest point algorithm, called the $K$-best method \cite{Zhan2006_K_best_sphere_decoder}, that we show provides near sphere decoder performance, but can be computed in a reasonable amount of time if $N$ is less than about 300.
-In this paper we derive the asymptotic properties of the LSU estimator, showing that it is strongly consistent and asymptotically normally distributed.
+In this two part series of papers we derive the asymptotic statistical properties of the LSU estimator, showing that it is strongly consistent and asymptotically normally distributed.
+Strong consistency is proved in this paper, while asymptotic normality is proved in the second part~\cite{McKilliam_pps2_2012}.
 Similar results were stated without a complete proof in~\cite{McKilliam_polyphase_est_icassp_2011}.
@@
 
-An interesting property is that the estimator of the $k$th polynomial phase coefficient converges almost surely to $\tilde{\mu}_k$ at rate $o(N^{-k})$.
+An interesting property is that the estimator of the $k$th polynomial phase coefficient converges to $\tilde{\mu}_k$ at rate $o(N^{-k})$.
 This is perhaps not surprising, since it is the same rate observed in polynomial regression.
@@
 %This proof technique appears to be novel, and may be useful in other problems.
-The proof of asymptotic normality is complicated by the fact that the objective function corresponding with the LSU estimator is not differentiable everywhere.
+%The proof of asymptotic normality is complicated by the fact that the objective function corresponding with the LSU estimator is not differentiable everywhere.
 Empirical process techniques~\cite{Pollard_new_ways_clts_1986,Pollard_asymp_empi_proc_1989,van2009empirical,Dudley_unif_central_lim_th_1999} and results from the literature on hyperplane arrangements~\cite{Chazelle_discrepency_method_2000,Matousek_lect_disc_geom_2002} become useful here.
@@
 The paper is organised in the following way.
-Section~\ref{sec:lattice-theory} describes the required concepts from lattice theory.
-In Section~\ref{sec:ident_aliasing} we briefly describe the identifiable region that was also derived in~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}.
+Section~\ref{sec:lattice-theory} describes some required concepts from lattice theory.
+In Section~\ref{sec:ident_aliasing} we describe the identifiable region that was also derived in~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}.
 These identifiability results are required in order to properly understand the statistical properties of polynomial phase estimators.
-In Section \ref{sec:least-squar-unwr} we describe the LSU estimator and derive its asymptotic statistical properties.
-%and also consider techniques for computing the LSU as a nearest lattice point problem. 
-Section \ref{sec:simulations} describes the results of Monte Carlo simulations that compare the performance of the LSU estimator with existing estimators.
-These simulations agree with the derived asymptotic properties.
+Section~\ref{sec:least-squar-unwr} describes the LSU estimator and states a theorem asserting the estimator to be strongly consistent.
+The theorem is proved in Section~\ref{sec:strongconstproof}.
+In the second paper~\cite{McKilliam_pps2_2012} we prove the asymptotic normality of the LSU estimator and describe the results of Monte Carlo simulations.
+These simulations agree with the derived asymptotic statistical properties.
 
@@
 One choice for a set of coset representatives is a fundamental parallelepiped; the parallelepiped generated by the columns of a generator matrix.
-Another choice is based on the Voronoi cell those points from $\reals^n$ nearest (with respect to the Euclidean norm here) to the lattice point at the origin.
+Another choice is based on the Voronoi cell; those points from $\reals^n$ nearest (with respect to the Euclidean norm in this paper) to the lattice point at the origin.
 It is always possible to construct a rectangular set of representatives, as the next proposition will show.
 We will use these rectangular regions for describing the aliasing properties of polynomial phase signals in Section~\ref{sec:ident_aliasing}.
-These rectangular regions will be important for the derivation of the asymptotic properties of the LSU estimator in Section~\ref{sec:least-squar-unwr}.
+These rectangular regions will be important for the derivation of the asymptotic properties of the LSU estimator.
+% in Section~\ref{sec:least-squar-unwr}.
 
@@
 These aliasing results are also given in~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}, but the presentation here is different, and is better suited to studying the asymptotic properties of the LSU estimator.
-A description of the aliasing properties of polynomial phase signals of order $2$ has been independently discovered in~\cite{Abatzoglou_ml_chirp_1986}~and~\cite{Angeby_PPS_aliasing_2000}.
+%A description of the aliasing properties of polynomial phase signals of order $2$ has been independently discovered in~\cite{Abatzoglou_ml_chirp_1986}~and~\cite{Angeby_PPS_aliasing_2000}.
 The results here are for polynomial phase signals of any order.
@@
 It may be helpful to observe Figures~\ref{fig:circstatplot_zero} to~\ref{fig:circstatplot_cube}.
-In these, the phase (divided by $2\pi$) of two distinct polynomial phase signals is plotted on the left, and on the right the principal component of the phase is plotted.
+In these, the phase (divided by $2\pi$) of two distinct polynomial phase signals is plotted on the left, and on the right the principal component of the phase (also divided by $2\pi$) is plotted.
 The circles display the samples at the integers.
@@
 \[
-\coef(g) = \left[ \begin{array}{ccccc} a_0 & a_1 & a_2 & \dots & a_m, \end{array} \right]^\prime,
+\coef(g) = \left[ \begin{array}{ccccc} a_0 & a_1 & a_2 & \dots & a_m \end{array} \right]^\prime,
 \]
@@
 If the minimiser is not unique we say that $\Phi_n$ has no intrinsic mean.
-We are now equipped to state the asymptotic properties of the LSU estimator.
+%We are now equipped to state the asymptotic properties of the LSU estimator.
+We are now equipped to state the main result of this paper.
 
@@
 
-\begin{theorem} \label{thm:asymp_proof} 
+\begin{theorem} \label{thm:asymp_proof}  (Strong consistency) 
 Let $\widehat{\mubf}$ be defined by~\eqref{eq:hatmubfLSUdefn} and put $\widehat{\lambdabf}_N = \dealias(\tilde{\mubf} - \widehat{\mubf})$.
 Denote the elements of $\widehat{\lambdabf}_N$ by $\widehat{\lambda}_{0,N}, \dots, \widehat{\lambda}_{m,N}$.
-Suppose $\Phi_1, \dots, \Phi_N$ are independent and identically distributed with zero intrinsic mean, intrinsic variance $\sigma^2$, and pdf $f$, then:
-\begin{enumerate}
-\item (Strong consistency) $N^k \widehat{\lambda}_{k,N}$ converges almost surely to $0$ as $N\rightarrow\infty$ for all $k = 0, 1, \dots, m$.
-\item (Asymptotic normality) If $f(\fracpart{x})$ is continuous at $x = -\nicefrac{1}{2}$ and $f(-\nicefrac{1}{2}) < 1$ then the distribution of
-\[
-\left[
-\begin{array}
-[c]{cccc}%
-\sqrt{N} \widehat{\lambda}_{0,N} & N \sqrt{N}\widehat{\lambda}_{1,N}  & \dots & N^m\sqrt{N} \widehat{\lambda}_{m,N}
-\end{array}
-\right]^\prime
-\]
-converges to the normal with zero mean and covariance
-\[
-\frac{\sigma^2}{\left(1-f( -\nicefrac{1}{2}) \right)^{2}} \Cbf^{-1},
-\]
-where $\Cbf$ is the $(m+1)\times (m+1)$ Hilbert matrix with elements $C_{ik} = 1/(i + k + 1)$ for $i,k = \{0, 1, \dots, m\}$.
-\end{enumerate}
+Suppose $\Phi_1, \dots, \Phi_N$ are independent and identically distributed with zero intrinsic mean and intrinsic variance $\sigma^2$, then $N^k \widehat{\lambda}_{k,N}$ converges almost surely to $0$ as $N\rightarrow\infty$ for all $k = 0, 1, \dots, m$.
 \end{theorem}
-The proof of Theorem~\ref{thm:asymp_proof} is contained within the next two sections.
-Section~\ref{sec:strongconstproof} proves strong consistency and Section~\ref{sec:centlimitproof} proves asymptotic normality.
-The proofs for the case of $m=0$ were given in~\cite{McKilliam_mean_dir_est_sq_arc_length2010} and for the case when $m=1$ were given in~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.
-The proofs here take a similar approach, but require new techniques.
+
+A proof of this theorem is given in Section~\ref{sec:strongconstproof}.
+A proofs for the case when $m=0$ was given in~\cite{McKilliam_mean_dir_est_sq_arc_length2010} and for the case when $m=1$ was given in~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.
+The proof here takes a similar approach, but requires new techniques.
 %Before giving the proof we make the following remarks.
@@
 
-The proof of asymptotic normality places requirements on the pdf $f$ of the phase noise.
+%The proof of asymptotic normality places requirements on the pdf $f$ of the phase noise.
 The requirement that $\Phi_1, \dots, \Phi_N$ have zero intrinsic mean implies that $f(-\nicefrac{1}{2}) \leq 1$~\cite[Lemma~1]{McKilliam_mean_dir_est_sq_arc_length2010}, so the only case not handled is when equality holds, i.e., when $f(-\nicefrac{1}{2}) = 1$ or when $f(\fracpart{x})$ is discontinuous at $x = -\nicefrac{1}{2}$.
@@
  \end{equation} 
-Now $\widehat{\lambdabf}_N = \dealias(\tilde{\mubf} - \widehat{\mubf})$ is the minimiser of $S_{N}$ in $B$.
-We shall show that $N^k\widehat{\lambda}_{k,N}\rightarrow0$ almost surely as $N\rightarrow\infty$ for all $k = 0,1, \dots, m$ and from this the proof of strong consistency follows.
+Now $\widehat{\lambdabf}_N = \dealias(\tilde{\mubf} - \widehat{\mubf})$ is the minimiser of $S_{N}$ in $B$.  %We now analyse the minimiser of $S_N$.  
 Let
@@
  \begin{equation}\label{eq:SNVNunifmlln}
-\sup_{\lambdabf \in B}\sabs{S_N(\lambdabf) - V_N(\lambdabf)} \rightarrow 0 \qquad \text{almost surely as $N\rightarrow\infty$}.  
- \end{equation}
+\sup_{\lambdabf \in B}\sabs{S_N(\lambdabf) - V_N(\lambdabf)} \rightarrow 0  
+ \end{equation}
+almost surely as $N\rightarrow\infty$.
 This type of result has been called a \emph{uniform law of large numbers} and follows from standard techniques~\cite{Pollard_conv_stat_proc_1984}.
@@
 \begin{proof}
-By definition $\widehat{\lambdabf}_N =\arg\min_{\lambdabf\in B} S_N(\lambdabf)$ so $0 \leq S_{N}(\zerobf) - S_N(\widehat{\lambdabf}_N)$.
-Also, because $V_N$ is minimised at $\zerobf$, it follows that $0 \leq V_N(\widehat{\lambdabf}_N) - V_N(\zerobf)$.
-Thus,
+By definition $\widehat{\lambdabf}_N =\arg\min_{\lambdabf\in B} S_N(\lambdabf)$ so 
+\[
+0 \leq S_{N}(\zerobf) - S_N(\widehat{\lambdabf}_N).
+\]  
+Also, because $V_N$ is minimised at $\zerobf$, it follows that 
+%\[
+%0 \leq V_N(\widehat{\lambdabf}_N) - V_N(\zerobf).
+%\]  
+%Combining the two inequalities above,
 \begin{align*}
@@
 However, the next two lemmas and the following corollary will show that if $K$ is sufficiently large then it always contains at least one arithmetic progression (for all sufficiently small $h$) and therefore $A(h,K)$ is not empty.
-We do not wish to claim any novelty here, the study of arithmetic progressions within subsets of $W$ has a considerable history~\cite{Erdos_on_some_sequence_of_integers1936,Szemeredi_setint_no_k_arth1975,Gowers_new_proof2001}.
+We do not wish to claim any novelty here.
+The study of arithmetic progressions within subsets of $W$ has a considerable history~\cite{Erdos_on_some_sequence_of_integers1936,Szemeredi_setint_no_k_arth1975,Gowers_new_proof2001}.
 In particular, Gower's~\cite[Theorem 1.3]{Gowers_new_proof2001} gives a result far stronger than we require here.
@@
 \[
-\Delta_h d(n) = d(n+h) - d(n)
+\Delta_h^1 d(n) = d(n+h) - d(n)
 \] 
@@
 
-The operator $\Delta_h$ has special properties when applied to polynomials. If $d(n) = a_r n^r + \dots + a_0$ is a polynomial of order $r$ then
+The operator $\Delta_h^r$ has special properties when applied to polynomials. If $d(n) = a_r n^r + \dots + a_0$ is a polynomial of order $r$ then
  \begin{equation} \label{eq:mfinitediffpoly}
@@
 \[
-z_N(n) = \lambda_{0,N} + \lambda_{1,N} n + \lambda_{2,N} n^2 + \dots + \lambda_{m,N} n^m
-\]
-so that $V_N(\lambdabf_N) - \sigma^2 = \frac{1}{N}\sum_{n=1}^{N} g\left( \fracpart{z_N(n)} \right) \rightarrow 0$ as $N \rightarrow \infty$.
+z_N(n) = \lambda_{0,N} + \lambda_{1,N} n + \dots + \lambda_{m,N} n^m
+\]
+so that 
+\[
+V_N(\lambdabf_N) - \sigma^2 = \frac{1}{N}\sum_{n=1}^{N} g\left( \fracpart{z_N(n)} \right) \rightarrow 0
+\] 
+as $N \rightarrow \infty$.
 %At this stage we can \emph{see} the proof.
@@
 \[
-z_{N,r}(n) = \lambda_{0,N} + \lambda_{1,N} n + \lambda_{2,N} n^2 + \dots + \lambda_{r,N} n^r.
-\]
-Because the $m-r$ highest order coefficients converge we can write $z_N(n) = z_{N,r}(n) + \gamma_N(n)$ where $\sup_{n\in\{1,\dots,N\}}\abs{\gamma_N(n)} \rightarrow 0$ as $N\rightarrow\infty$.
+z_{N,r}(n) = \lambda_{0,N} + \lambda_{1,N} n + \dots + \lambda_{r,N} n^r.
+\]
+Because the $m-r$ highest order coefficients converge we can write $z_N(n) = z_{N,r}(n) + \gamma_N(n)$ where 
+\[
+\sup_{n\in\{1,\dots,N\}}\abs{\gamma_N(n)} \rightarrow 0 \qquad \text{as $N\rightarrow\infty$}. 
+\]
 Now the bound from \eqref{eq:Deltazfracbound}, but applied using the $r$th difference, gives
@@
 for all $h$ such that $1 \leq h \leq \tfrac{N}{2m}$.
-Choosing $2^r\delta + |\epsilon| < \frac{1}{2^{2r+1}}$ and using the same iterative process as for the highest order coefficient $\lambda_{m,N}$  (see~\eqref{eq:startiterativearg}~to~\eqref{eq:enditerativearg}) we find that $N^r \lambda_{r,N} \rightarrow 0$ as $N\rightarrow\infty$.
+Choosing $2^r\delta + |\epsilon| < 2^{-2r-1}$ and using the same iterative process as for the highest order coefficient $\lambda_{m,N}$  (see~\eqref{eq:startiterativearg}~to~\eqref{eq:enditerativearg}) we find that $N^r \lambda_{r,N} \rightarrow 0$ as $N\rightarrow\infty$.
 The proof now follows by induction.
@@
 We are now in a position to complete the proof of strong consistency.
-Let $A$ be the subset of the sample space on which  $V_N(\widehat{\lambdabf}_N) - \sigma^2 \rightarrow 0$ as $N\rightarrow\infty$.
-From Lemma~\ref{lem:ESNconv}, the $\prob\{A\} =1$.
-Let $A'$ be the subset of the sample space on which $N^k\widehat{\lambda}_{k,N} \rightarrow 0$ for $k=0,\dots,m$ as $N\rightarrow\infty$.
+As is customary, let $(\Omega, \mathcal F, \prob)$ be the probability space over which the random variables $\{X_i\}$ and $\{\Phi_i\}$ are defined.
+Let $A$ be the subset of the sample space $\Omega$ on which  $V_N(\widehat{\lambdabf}_N) - \sigma^2 \rightarrow 0$ as $N\rightarrow\infty$.
+Now $\prob\{A\} =1$ as a result of Lemma~\ref{lem:ESNconv}.
+Let $A'$ be the subset of $\Omega$ on which $N^k\widehat{\lambda}_{k,N} \rightarrow 0$ for $k=0,\dots,m$ as $N\rightarrow\infty$.
 As a result of Lemma~\ref{lem:moran2}, $A \subseteq A'$, and so $\prob\{A'\} \geq \prob\{A\} = 1$.
@@
 This paper has considered the estimation of the coefficients of a noisy polynomial phase signal by least squares phase unwrapping (LSU).
-It has been shown that the LSU estimator is strongly consistent and asymptotically normally distributed.
+We have derived conditions under which the polynomial phase estimation problem is identifiable.
+Under these conditions, and under some assumptions on the distribution of the noise, the LSU estimator is shown to be strongly consistent.
 %Simulations are used to show that the LSU has near maximum likelihood performance in the case that the noise is additive white and Gaussian.
-The LSU estimator can be computed by finding a nearest lattice point in a lattice~\cite{McKilliam2009asilomar_polyest_lattice}.
-%Here, we consider using the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder} to achieve this in reasonable time when the number of observations, $N$, is less than about 300.
-Polynomial time nearest point algorithms for these lattices exist~\cite[Sec 4.3]{McKilliam2010thesis}, but these algorithms are not fast in practice.
-Instead, we considered some general purpose algorithms, the \term{sphere decoder}, \term{Babai's nearest plane algorithm} and the \term{$K$-best algorithm}.
-The sphere decoder and $K$-best algorithm result in accurate estimators.
-
-The major outstanding question is whether faster nearest point algorithms exist for these specific lattices.
-Considering the excellent statistical performance (both theoretically and practically) of the LSU estimator, even fast \emph{approximate} nearest point algorithms are likely to prove useful for the estimation of polynomial phase signals.
+In the second paper in this series~\cite{McKilliam_pps2_2012} we show the LSU estimator to be asymptotically normal and present the results of Monte Carlo simulations that support our asymptotic results.
 

### 2012-10-29 paper2.tex
added to two papers

@@
 %It was shown by Monte Carlo simulation that this produces a remarkably accurate estimator.  
-In this paper we consider an estimator that performs phase unwrapping in a least squares manner.
-We describe the asymptotic properties of this estimator, showing that it is strongly consistent and asymptotically normally distributed.
-%We hypothesise that the estimator produces very near maximum likelihood performance.
+In this two part series of papers we consider an estimator that performs phase unwrapping in a least squares manner.
+In this second part we prove that the estimator is asymptotically normally distributed.
+The results of Monte Carlo simulations are presented and these support our asymptotic theory.
 \end{abstract}
@@
 
-A polynomial phase signal of order $m$ is a function of the form
-\[
-s(t) = e^{2\pi j y(t)},
-\]
-where $j = \sqrt{-1}$, and $t$ is a real number, often representing time, and 
-\[
-y(t) = \tilde{\mu}_0 +\tilde{\mu}_1 t + \tilde{\mu}_2 t^2 + \dots \tilde{\mu}_m t^m
-\]
-is a polynomial of order $m$.
+% A polynomial phase signal of order $m$ is a function of the form
+% \[
+% s(t) = e^{2\pi j y(t)},
+% \]
+% where $j = \sqrt{-1}$, and $t$ is a real number, often representing time, and 
+% \[
+% y(t) = \tilde{\mu}_0 +\tilde{\mu}_1 t + \tilde{\mu}_2 t^2 + \dots \tilde{\mu}_m t^m
+% \]
+% is a polynomial of order $m$.
 In practice the signal is typically sampled at discrete points in `time', $t$.
@@
 In this case we can always consider the samples to be taken at some set of consecutive integers and our sampled polynomial phase signal looks like
-\[
-s_n = s(n) = e^{2\pi j y(n)},
-\] 
-where $n$ is an integer.  Of practical importance is the estimation of the coefficients $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ from a number, say $N$, of observations of the noisy sampled signal
+A uniformly sampled polynomial-phase signal of order $m$ is a signal of the form
+\[
+s_n = e^{2\pi j y(n)},
+\] 
+where $n$ is an integer and 
+\[
+y(n) = \tilde{\mu}_0 +\tilde{\mu}_1 n + \tilde{\mu}_2 n^2 + \dots \tilde{\mu}_m n^m
+\]
+is a polynomial of order $m$.  Of practical importance is the estimation of the coefficients $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ from a number, say $N$, of observations of the noisy sampled signal
 \begin{equation}\label{eq:Y_nsamplednoisey}
@@
 where $\rho$ is a real number greater than zero representing the (usually unknown) signal amplitude and $\{X_n, n \in \ints\}$ is a sequence complex noise variables.
-In order to ensure identifiability it is necessary to restrict the $m+1$ coefficients to a region of $m+1$ dimensional Euclidean space $\reals^{m+1}$ called an \emph{identifiable region}.
+%In order to ensure identifiability it is necessary to restrict the $m+1$ coefficients to a region of $m+1$ dimensional Euclidean space $\reals^{m+1}$ called an \emph{identifiable region}.
 It was shown in \cite{McKilliam2009IndentifiabliltyAliasingPolyphase} that an identifiable region tessellates a particular $m+1$ dimensional lattice.
@@
 %We also apply another known approximate nearest point algorithm, called the $K$-best method \cite{Zhan2006_K_best_sphere_decoder}, that we show provides near sphere decoder performance, but can be computed in a reasonable amount of time if $N$ is less than about 300.
-In this paper we derive the asymptotic properties of the LSU estimator, showing that it is strongly consistent and asymptotically normally distributed.
-Similar results were stated without a complete proof in~\cite{McKilliam_polyphase_est_icassp_2011}.
-Here, we give a proof.
-The results here are also more general than in~\cite{McKilliam_polyphase_est_icassp_2011}, allowing for a wider class of noise distributions.
-
-An interesting property is that the estimator of the $k$th polynomial phase coefficient converges almost surely to $\tilde{\mu}_k$ at rate $o(N^{-k})$.
-This is perhaps not surprising, since it is the same rate observed in polynomial regression.
-However, asserting that convergence at this rate occurs in the polynomial phase setting is not trivial.
-For this purpose we make use of an elementary result about the number of arithmetic progressions contained inside subsets of $\{1,2,\dots,N\}$~\cite{Erdos_on_some_sequence_of_integers1936,Szemeredi_setint_no_k_arth1975,Gowers_new_proof2001}.
-%This proof technique appears to be novel, and may be useful in other problems.
+In this two part series of papers we derive the asymptotic statistical properties of the LSU estimator.
+The first part in the series~\cite{McKilliam_pps1_2012} asserts the strong consistency of the LSU estimator, under some assumption on the distribution on the noise $\{X_n\}$.
+In this second part we prove, under similar conditions on $\{X_n\}$, the asymptotic normality of the LSU estimator.
 The proof of asymptotic normality is complicated by the fact that the objective function corresponding with the LSU estimator is not differentiable everywhere.
@@
 The paper is organised in the following way.
-Section~\ref{sec:lattice-theory} describes the required concepts from lattice theory.
-In Section~\ref{sec:ident_aliasing} we briefly describe the identifiable region that was also derived in~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}.
-These identifiability results are required in order to properly understand the statistical properties of polynomial phase estimators.
-In Section \ref{sec:least-squar-unwr} we describe the LSU estimator and derive its asymptotic statistical properties.
-%and also consider techniques for computing the LSU as a nearest lattice point problem. 
+Section \ref{sec:least-squar-unwr} describes the LSU estimator and states a theorem asserting the estimator to be asymptotically normally distributed.
+The theorem is proved in Section~\ref{sec:centlimitproof}.
 Section \ref{sec:simulations} describes the results of Monte Carlo simulations that compare the performance of the LSU estimator with existing estimators.
 These simulations agree with the derived asymptotic properties.
+This paper is self contained and can be read independently of the first part~\cite{McKilliam_pps1_2012}.
+Those results required from~\cite{McKilliam_pps1_2012} are referenced here.
+However, it is intended (and recommended) that the papers be read in order.
 
@@
 
-\begin{theorem} \label{thm:asymp_proof} 
+\begin{theorem} \label{thm:asymp_proof} (Asymptotic normality)
 Let $\widehat{\mubf}$ be defined by~\eqref{eq:hatmubfLSUdefn} and put $\widehat{\lambdabf}_N = \dealias(\tilde{\mubf} - \widehat{\mubf})$.
 Denote the elements of $\widehat{\lambdabf}_N$ by $\widehat{\lambda}_{0,N}, \dots, \widehat{\lambda}_{m,N}$.
-Suppose $\Phi_1, \dots, \Phi_N$ are independent and identically distributed with zero intrinsic mean, intrinsic variance $\sigma^2$, and pdf $f$, then:
-\begin{enumerate}
-\item (Strong consistency) $N^k \widehat{\lambda}_{k,N}$ converges almost surely to $0$ as $N\rightarrow\infty$ for all $k = 0, 1, \dots, m$.
-\item (Asymptotic normality) If $f(\fracpart{x})$ is continuous at $x = -\nicefrac{1}{2}$ and $f(-\nicefrac{1}{2}) < 1$ then the distribution of
+Suppose $\Phi_1, \dots, \Phi_N$ are independent and identically distributed with zero intrinsic mean, intrinsic variance $\sigma^2$, and pdf $f$.
+If $f(\fracpart{x})$ is continuous at $x = -\nicefrac{1}{2}$ and $f(-\nicefrac{1}{2}) < 1$ then the distribution of
 \[
@@
 where $\Cbf$ is the $(m+1)\times (m+1)$ Hilbert matrix with elements $C_{ik} = 1/(i + k + 1)$ for $i,k = \{0, 1, \dots, m\}$.
-\end{enumerate}
 \end{theorem}

### 2012-10-29 paper.tex
minor edits

@@
 In the interest of simplicity we only consider the case when $\Phi_1, \dots, \Phi_N$ are identically distributed here.
-If $X_n$ is circularly symmetric with density function nonincreasing with magnitude $\abs{X_n}$, then, the corresponding $\Phi_n$ necessarily has zero intrinsic mean~\cite[Theorem 5.2, page 78]{McKilliam2010thesis}.
+If $X_n$ is circularly symmetric with density function nonincreasing with magnitude $\abs{X_n}$, then, the corresponding $\Phi_n$ necessarily has zero intrinsic mean~\cite[Theorem~5.2,~page~78]{McKilliam2010thesis}.
 Thus, our theorem covers commonly used distributions for $X_1, \dots, X_N$, such as the normal distribution.
+
+Although we will not prove it here the assumption that $\Phi_1,\dots,\Phi_N$ have zero intrinsic mean is not only sufficient, but also necessary, for if $\Phi_1,\dots,\Phi_N$ have intrinsic mean $x \in [-\nicefrac{1}{2},\nicefrac{1}{2})$ with $x \neq 0$ then $\fracpart{\widehat{\lambda}_{0,N} - x}\rightarrow 0$ almost surely as $N\rightarrow\infty$, and so $\widehat{\lambda}_{0,N}$ does not converge to zero.
+On the other hand if $\Phi_1,\dots,\Phi_N$ do not have an intrinsic mean then $\widehat{\lambda}_{0,N}$ will not converge.
 

### 2012-10-29 paper2.tex
minor edits

@@
 
-We now describe the least squares unwrapping (LSU) estimator of the polynomial coefficients.
-Recall that we desire to estimate the coefficients $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ from the noisy samples $Y_1, \dots, Y_N$ given in~\eqref{eq:Y_nsamplednoisey}.
-We take the complex argument of the $Y_n$ and divide by $2\pi$ to obtain
+As in~\cite{McKilliam_pps1_2012} the least squares unwrapping (LSU) estimator of the polynomial coefficients is the minimiser
+\begin{equation}\label{eq:hatmubfLSUdefn}
+\widehat{\mubf} = \arg \min_{\mubf \in B} SS(\mubf)
+\end{equation}
+where the objective function
+\begin{equation} \label{eq:sumofsquaresfunction}
+SS(\mubf) = \sum_{n=1}^{N}\fracpart{  \Theta_{n} - \sum_{k = 0}^{m}{\mu_k n^k} }^{2},
+\end{equation}
+and
 \begin{equation}\label{eq:noise_circ_poly}
-\Theta_n = \frac{\angle{Y_n}}{2\pi} = \fracpart{ \Phi_n + y(n) }
-\end{equation}
-where $\angle{z}$ denotes the complex argument of the complex number $z$, and 
-\[
-\Phi_n = \frac{1}{2\pi}\angle(1 + \rho^{-1}s_n^{-1}X_n)
-\] 
-are random variables representing the phase noise induced by the $X_n$~\cite{Tretter1985,Quinn2009_dasp_phase_only_information_loss}.
-If the distribution of $X_n$ is circularly symmetric (i.e., the angle $\angle X_n$ is uniformly distributed on $[-\pi, \pi)$ and is independent of the magnitude $\abs{X_n}$) then the distribution of $\Phi_n$ is the same as the distribution of $\tfrac{1}{2\pi}\angle(1 + \rho^{-1}X_n)$.
-If the $X_1, \dots, X_N$ are circularly symmetric and identically distributed, then $\Phi_1, \dots, \Phi_n$ are also identically distributed.
-
-Let $\mubf$ be the vector $[\mu_0, \mu_1, \dots, \mu_m]$ and put,
-\begin{equation} \label{eq:sumofsquaresfunction}
-SS(\mubf) = \sum_{n=1}^{N}\fracpart{  \Theta_{n} - \sum_{k = 0}^{m}{\mu_k n^k} }^{2}.
-\end{equation}
-The least squares unwrapping estimator is defined as those coefficients $\widehat{\mu}_0, \dots, \widehat{\mu}_m$ that minimise $SS$ over the identifiable region $B$. That is, the LSU estimator is,
-\begin{equation}\label{eq:hatmubfLSUdefn}
-\widehat{\mubf} = \arg \min_{\mubf \in B} SS(\mubf). 
-\end{equation}
-
-It is shown in~\cite[Sec~8.1]{McKilliam2010thesis}\cite{McKilliam2009asilomar_polyest_lattice} how this minimisation problem can be posed as that of computing a nearest lattice point in a particular lattice.
+\Theta_n = \frac{\angle{Y_n}}{2\pi} = \fracpart{ \Phi_n + y(n) },
+\end{equation}
+is the phase (divided by $2\pi$) of the observation $Y_1,\dots,Y_N$, and
+\[
+\Phi_n = \frac{1}{2\pi}\angle(1 + \rho^{-1}s_n^{-1}X_n).
+\] 
+are random variables representing the \emph{phase noise} observed at the reciever~\cite{Tretter1985}.
+We denote by $\angle{z}$ the complex argument of the complex number $z$, and by $\fracpart{x} = x - \round{x}$ the (centered) fractional part of the real number $x$, and by $\round{x}$ the nearest integer to $x$.
+We have chosen to round half integers up here.
+The next theorem describes the asymptotic distribution of the LSU estimator.
+
+%It is shown in~\cite[Sec~8.1]{McKilliam2010thesis}\cite{McKilliam2009asilomar_polyest_lattice} how the minimisation problem in~\eqref{eq:hatmubfLSUdefn} can be posed as that of computing a nearest lattice point in a particular lattice.
 Polynomial time algorithms that compute the nearest point are described in~\cite[Sec.~4.3]{McKilliam2010thesis}.
@@
 
-The next theorem describes the asymptotic properties of this estimator.
+%The next theorem describes the asymptotic properties of this estimator.
 Before we give the proof it is necessary to understand some of the properties of the phase noise $\Phi_1,\dots,\Phi_N$, which are \emph{circular} random variables with support on $[-\nicefrac{1}{2}, \nicefrac{1}{2})$~\cite{McKilliam2010thesis,McKilliam_mean_dir_est_sq_arc_length2010,Mardia_directional_statistics,Fisher1993}.
@@
 
-The \emph{intrinsic mean} or \emph{Fr\'{e}chet mean} of $\Phi_n$ is defined as~\cite{McKilliam_mean_dir_est_sq_arc_length2010,bwhk07a,Bhattacharya_int_ext_means_2003,Bhattacharya_int_ext_means_2005},
-\begin{equation}\label{eq:intrmeandefn}
- \mu_{\text{intr}}  = \arg \min_{\mu \in [-\nicefrac{1}{2}, \nicefrac{1}{2})} \expect \fracpart{\Phi_n - \mu}^2, 
-% %&= \arg \min_{\mu \in [-\nicefrac{1}{2}, \nicefrac{1}{2})}\int_{-\nicefrac{1}{2}}^{\nicefrac{1}{2}}\fracpart{\phi - \mu}^2f(\theta) d\theta. \label{eq:minunrappedmeandef}
-\end{equation}
-and the \emph{intrinsic variance} is
-\[
-\sigma_{\text{intr}}^2 = E\fracpart{\Theta - \mu_{\text{intr}}}^2 = \min_{\mu \in [-\nicefrac{1}{2}, \nicefrac{1}{2})} \expect \fracpart{\Phi_n - \mu}^2,
-\]
-where $\expect$ denotes the expected value.
+% The \emph{intrinsic mean} or \emph{Fr\'{e}chet mean} of $\Phi_n$ is defined as~\cite{McKilliam_mean_dir_est_sq_arc_length2010,bwhk07a,Bhattacharya_int_ext_means_2003,Bhattacharya_int_ext_means_2005},
+% \begin{equation}\label{eq:intrmeandefn}
+%  \mu_{\text{intr}}  = \arg \min_{\mu \in [-\nicefrac{1}{2}, \nicefrac{1}{2})} \expect \fracpart{\Phi_n - \mu}^2, 
+% % %&= \arg \min_{\mu \in [-\nicefrac{1}{2}, \nicefrac{1}{2})}\int_{-\nicefrac{1}{2}}^{\nicefrac{1}{2}}\fracpart{\phi - \mu}^2f(\theta) d\theta. \label{eq:minunrappedmeandef}
+% \end{equation}
+% and the \emph{intrinsic variance} is
+% \[
+% \sigma_{\text{intr}}^2 = E\fracpart{\Theta - \mu_{\text{intr}}}^2 = \min_{\mu \in [-\nicefrac{1}{2}, \nicefrac{1}{2})} \expect \fracpart{\Phi_n - \mu}^2,
+% \]
+% where $\expect$ denotes the expected value.
 Depending on the distribution of $\Phi_n$ the argument that minimises~\eqref{eq:intrmeandefn} may not be unique.
@@
 Denote the elements of $\widehat{\lambdabf}_N$ by $\widehat{\lambda}_{0,N}, \dots, \widehat{\lambda}_{m,N}$.
-Suppose $\Phi_1, \dots, \Phi_N$ are independent and identically distributed with zero intrinsic mean, intrinsic variance $\sigma^2$, and pdf $f$.
-If $f(\fracpart{x})$ is continuous at $x = -\nicefrac{1}{2}$ and $f(-\nicefrac{1}{2}) < 1$ then the distribution of
+Suppose $\Phi_1, \dots, \Phi_N$ are independent and identically distributed with zero mean, variance $\sigma^2$, and pdf $f$ such that $f(\fracpart{x})$ is continuous at $x = -\nicefrac{1}{2}$ and $f(-\nicefrac{1}{2}) < 1$.
+Suppose also that  $N^k \widehat{\lambda}_{k,N} \rightarrow 0$ in probability as $N\rightarrow\infty$ for all $k = 0, 1, \dots, m$.
+Then the distribution of
 \[

### 2012-10-29 paper2.tex
minor

@@
 \end{equation}
-where the objective function
+where $B = \prod_{k=0}^{m}\left[ -\frac{0.5}{k!}, \frac{0.5}{k!}  \right)$ is a subset of $\reals^{m+1}$ defined in~\cite[Sec.~3]{McKilliam_pps1_2012} called the identifiable region, and where the objective function
 \begin{equation} \label{eq:sumofsquaresfunction}
@@
 \end{equation}
-and
+where
 \begin{equation}\label{eq:noise_circ_poly}
@@
 \end{equation}
-is the phase (divided by $2\pi$) of the observation $Y_1,\dots,Y_N$, and
-\[
-\Phi_n = \frac{1}{2\pi}\angle(1 + \rho^{-1}s_n^{-1}X_n).
+is the phase (divided by $2\pi$) of the observations $Y_1,\dots,Y_N$, and
+\[
+\Phi_n = \frac{1}{2\pi}\angle(1 + \rho^{-1}s_n^{-1}X_n)
 \] 
@@
 We have chosen to round half integers up here.
+
+As in~\cite[Sec.~4]{McKilliam_pps1_2012} the phase noise $\Phi_1,\dots,\Phi_N$ are \emph{circular} random variables with support on $[-\nicefrac{1}{2}, \nicefrac{1}{2})$~\cite{McKilliam2010thesis,McKilliam_mean_dir_est_sq_arc_length2010,Mardia_directional_statistics,Fisher1993}.
+The \emph{intrinsic mean} or \emph{Fr\'{e}chet mean} of $\Phi_n$ is defined as~\cite{McKilliam_mean_dir_est_sq_arc_length2010,bwhk07a,Bhattacharya_int_ext_means_2003,Bhattacharya_int_ext_means_2005},
+\begin{equation}\label{eq:intrmeandefn}
+ \mu_{\text{intr}}  = \arg \min_{\mu \in [-\nicefrac{1}{2}, \nicefrac{1}{2})} \expect \fracpart{\Phi_n - \mu}^2, 
+% %&= \arg \min_{\mu \in [-\nicefrac{1}{2}, \nicefrac{1}{2})}\int_{-\nicefrac{1}{2}}^{\nicefrac{1}{2}}\fracpart{\phi - \mu}^2f(\theta) d\theta. \label{eq:minunrappedmeandef}
+\end{equation}
+and the \emph{intrinsic variance} is
+\[
+\sigma_{\text{intr}}^2 = E\fracpart{\Phi_n - \mu_{\text{intr}}}^2 = \min_{\mu \in [-\nicefrac{1}{2}, \nicefrac{1}{2})} \expect \fracpart{\Phi_n - \mu}^2,
+\]
+where $\expect$ denotes the expected value.
+Observe that if $\Phi_n$ has zero intrinsic mean then it also has zero mean and the variance and the intrinsic variance of $\Phi_n$ are equal.
 The next theorem describes the asymptotic distribution of the LSU estimator.
@@
 Denote the elements of $\widehat{\lambdabf}_N$ by $\widehat{\lambda}_{0,N}, \dots, \widehat{\lambda}_{m,N}$.
-Suppose $\Phi_1, \dots, \Phi_N$ are independent and identically distributed with zero mean, variance $\sigma^2$, and pdf $f$ such that $f(\fracpart{x})$ is continuous at $x = -\nicefrac{1}{2}$ and $f(-\nicefrac{1}{2}) < 1$.
-Suppose also that  $N^k \widehat{\lambda}_{k,N} \rightarrow 0$ in probability as $N\rightarrow\infty$ for all $k = 0, 1, \dots, m$.
+Suppose $\Phi_1, \dots, \Phi_N$ are independent and identically distributed with zero intrinsic mean, intrinsic variance $\sigma^2$, and pdf $f$ such that $f(\fracpart{x})$ is continuous at $x = -\nicefrac{1}{2}$ and $f(-\nicefrac{1}{2}) < 1$.
 Then the distribution of
@@
 \end{theorem}
-The proof of Theorem~\ref{thm:asymp_proof} is contained within the next two sections.
-Section~\ref{sec:strongconstproof} proves strong consistency and Section~\ref{sec:centlimitproof} proves asymptotic normality.
-The proofs for the case of $m=0$ were given in~\cite{McKilliam_mean_dir_est_sq_arc_length2010} and for the case when $m=1$ were given in~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.
+
+A proof of this theorem is given in the next section.
+%The proofs for the case of $m=0$ were given in~\cite{McKilliam_mean_dir_est_sq_arc_length2010} and for the case when $m=1$ were given in~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.
 The proofs here take a similar approach, but require new techniques.
 %Before giving the proof we make the following remarks.
-The theorem gives conditions on the \emph{dealiased} difference $\dealias(\tilde{\mubf} - \widehat{\mubf})$ between the true coefficients $\tilde{\mubf}$ and the estimated coefficients $\widehat{\mubf}$ rather than directly on the difference $\tilde{\mubf} - \widehat{\mubf}$.
-To see why this makes sense, consider the case when $m=0$, $\tilde{\mu}_0 = -0.5$ and $\widehat{\mu}_0 = 0.49$, so that $\tilde{\mu}_0 - \widehat{\mu}_0 = -0.99$.
+As in~\cite{McKilliam_pps1_2012} the theorem gives conditions on the \emph{dealiased} difference $\dealias(\tilde{\mubf} - \widehat{\mubf})$ between the true coefficients $\tilde{\mubf}$ and the estimated coefficients $\widehat{\mubf}$ rather than directly on the difference $\tilde{\mubf} - \widehat{\mubf}$.
+This makes sense in view of the identifiability conditions described in~\cite[Sec.~3]{McKilliam_pps1_2012}  %To see why this makes sense, consider the case when $m=0$, $\tilde{\mu}_0 = -0.5$ and $\widehat{\mu}_0 = 0.49$, so that $\tilde{\mu}_0 - \widehat{\mu}_0 = -0.99$.
 However, the two phases are obviously close, since the phase $\pm 0.5$ are actually the same.
 In this case
-\[
-\dealias(\tilde{\mu}_0 - \widehat{\mu}_0) = \fracpart{\tilde{\mu}_0 - \widehat{\mu}_0} = 0.01.
-\] 
-The same reasoning holds for $m > 0$.
-
-The requirement that $\Phi_1, \dots, \Phi_N$ be identically distributed will typically hold only when the complex random variables $X_1, \dots, X_N$ are identically distributed and circularly symmetric.
+%\[
+%\dealias(\tilde{\mu}_0 - \widehat{\mu}_0) = \fracpart{\tilde{\mu}_0 - \widehat{\mu}_0} = 0.01.
+%\] 
+%The same reasoning holds for $m > 0$.
+
+%The requirement that $\Phi_1, \dots, \Phi_N$ be identically distributed will typically hold only when the complex random variables $X_1, \dots, X_N$ are identically distributed and circularly symmetric.
 It would be possible to drop the assumption that $\Phi_1, \dots, \Phi_N$ be identically distributed, but this complicates the theorem statement and the proof.
@@
 The requirement that $\Phi_1, \dots, \Phi_N$ have zero intrinsic mean implies that $f(-\nicefrac{1}{2}) \leq 1$~\cite[Lemma~1]{McKilliam_mean_dir_est_sq_arc_length2010}, so the only case not handled is when equality holds, i.e., when $f(-\nicefrac{1}{2}) = 1$ or when $f(\fracpart{x})$ is discontinuous at $x = -\nicefrac{1}{2}$.
-In this exceptional case other expressions for the asymptotic variance can be found (similar to \cite[Theorem 3.1]{Hotz_circle_means_2011}), but this comes at a substantial increase in complexity and for this reason we have omitted them.
+In these exceptional cases other expressions for the asymptotic variance can be found (similar to \cite[Theorem 3.1]{Hotz_circle_means_2011}), but this comes at a substantial increase in complexity and for this reason we have omitted them.
 %Distributions that do not satisfy these requirements are unlikely to be needed in practice.

### 2012-10-30 paper.tex
minor
added some newplot data

@@
 \end{equation}
-where $\rho$ is a real number greater than zero representing the (usually unknown) signal amplitude and $\{X_n, n \in \ints\}$ is a sequence complex noise variables.
+where $\rho$ is a positive real number representing the (usually unknown) signal amplitude and $\{X_n, n \in \ints\}$ is a sequence of complex noise variables.
 In order to ensure identifiability it is necessary to restrict the $m+1$ coefficients to a region of $m+1$ dimensional Euclidean space $\reals^{m+1}$ called an \emph{identifiable region}.
@@
 %We also apply another known approximate nearest point algorithm, called the $K$-best method \cite{Zhan2006_K_best_sphere_decoder}, that we show provides near sphere decoder performance, but can be computed in a reasonable amount of time if $N$ is less than about 300.
-In this two part series of papers we derive the asymptotic statistical properties of the LSU estimator, showing that it is strongly consistent and asymptotically normally distributed.
+In this two part series of we derive the asymptotic statistical properties of the LSU estimator showing, under some assumptions on the noise $X_1,\dots,X_N$, that it is strongly consistent and asymptotically normally distributed.
 Strong consistency is proved in this paper, while asymptotic normality is proved in the second part~\cite{McKilliam_pps2_2012}.
@@
 These identifiability results are required in order to properly understand the statistical properties of polynomial phase estimators.
-Section~\ref{sec:least-squar-unwr} describes the LSU estimator and states a theorem asserting the estimator to be strongly consistent.
+Section~\ref{sec:least-squar-unwr} describes the LSU estimator and states a theorem asserting the estimator to be strongly consistent under some assumptions on the noise $X_1,\dots,X_N$.
 The theorem is proved in Section~\ref{sec:strongconstproof}.
 In the second paper~\cite{McKilliam_pps2_2012} we prove the asymptotic normality of the LSU estimator and describe the results of Monte Carlo simulations.
-These simulations agree with the derived asymptotic statistical properties.
+These simulations agree with our derived asymptotic statistical properties.
 
@@
 \[
-\sigma_{\text{intr}}^2 = E\fracpart{\Theta - \mu_{\text{intr}}}^2 = \min_{\mu \in [-\nicefrac{1}{2}, \nicefrac{1}{2})} \expect \fracpart{\Phi_n - \mu}^2,
+\sigma_{\text{intr}}^2 = \expect\fracpart{\Phi_n - \mu_{\text{intr}}}^2 = \min_{\mu \in [-\nicefrac{1}{2}, \nicefrac{1}{2})} \expect \fracpart{\Phi_n - \mu}^2,
 \]
@@
 
-A proof of this theorem is given in Section~\ref{sec:strongconstproof}.
-A proofs for the case when $m=0$ was given in~\cite{McKilliam_mean_dir_est_sq_arc_length2010} and for the case when $m=1$ was given in~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.
+A proof of this theorem is given in the next section.
+Proofs for the case when $m=0$ was given in~\cite{McKilliam_mean_dir_est_sq_arc_length2010} and for the case when $m=1$ was given in~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.
 The proof here takes a similar approach, but requires new techniques.

### 2012-10-30 paper2.tex
added some newplot data

@@
 \end{equation}
-where $\rho$ is a real number greater than zero representing the (usually unknown) signal amplitude and $\{X_n, n \in \ints\}$ is a sequence complex noise variables.
+where $\rho$ is a positive real number representing the (usually unknown) signal amplitude and $\{X_n, n \in \ints\}$ is a sequence of complex noise variables.
 %In order to ensure identifiability it is necessary to restrict the $m+1$ coefficients to a region of $m+1$ dimensional Euclidean space $\reals^{m+1}$ called an \emph{identifiable region}.
@@
 In this two part series of papers we derive the asymptotic statistical properties of the LSU estimator.
-The first part in the series~\cite{McKilliam_pps1_2012} asserts the strong consistency of the LSU estimator, under some assumption on the distribution on the noise $\{X_n\}$.
-In this second part we prove, under similar conditions on $\{X_n\}$, the asymptotic normality of the LSU estimator.
+The first part in the series~\cite{McKilliam_pps1_2012} asserts the strong consistency of the LSU estimator under some conditions on the distribution on the noise $\{X_n\}$.
+In this second part we prove, under similar conditions on $\{X_n\}$, that the LSU estimator is asymptotically normallly distributed.
+
 The proof of asymptotic normality is complicated by the fact that the objective function corresponding with the LSU estimator is not differentiable everywhere.
@@
 The theorem is proved in Section~\ref{sec:centlimitproof}.
-Section \ref{sec:simulations} describes the results of Monte Carlo simulations that compare the performance of the LSU estimator with existing estimators.
+Section \ref{sec:simulations} describes the results of Monte Carlo simulations that compare the performance of the LSU estimator with some existing estimators.
 These simulations agree with the derived asymptotic properties.
@@
 \end{equation}
-where $B = \prod_{k=0}^{m}\left[ -\frac{0.5}{k!}, \frac{0.5}{k!}  \right)$ is a subset of $\reals^{m+1}$ defined in~\cite[Sec.~3]{McKilliam_pps1_2012} called the identifiable region, and where the objective function
+where $B = \prod_{k=0}^{m}\left[ -\frac{0.5}{k!}, \frac{0.5}{k!}  \right)$ is a subset of $\reals^{m+1}$ defined in~\cite[Sec.~3]{McKilliam_pps1_2012} called the \emph{identifiable region}, and where the objective function
 \begin{equation} \label{eq:sumofsquaresfunction}
@@
 We denote by $\angle{z}$ the complex argument of the complex number $z$, and by $\fracpart{x} = x - \round{x}$ the (centered) fractional part of the real number $x$, and by $\round{x}$ the nearest integer to $x$.
-We have chosen to round half integers up here.
+We have chosen to round half integers up.
 
 As in~\cite[Sec.~4]{McKilliam_pps1_2012} the phase noise $\Phi_1,\dots,\Phi_N$ are \emph{circular} random variables with support on $[-\nicefrac{1}{2}, \nicefrac{1}{2})$~\cite{McKilliam2010thesis,McKilliam_mean_dir_est_sq_arc_length2010,Mardia_directional_statistics,Fisher1993}.
-The \emph{intrinsic mean} or \emph{Fr\'{e}chet mean} of $\Phi_n$ is defined as~\cite{McKilliam_mean_dir_est_sq_arc_length2010,bwhk07a,Bhattacharya_int_ext_means_2003,Bhattacharya_int_ext_means_2005},
+The \emph{intrinsic mean} or \emph{Fr\'{e}chet mean} of $\Phi_n$ is defined as~\cite{McKilliam_mean_dir_est_sq_arc_length2010,Bhattacharya_int_ext_means_2003,Bhattacharya_int_ext_means_2005},
 \begin{equation}\label{eq:intrmeandefn}
@@
 \[
-\sigma_{\text{intr}}^2 = E\fracpart{\Phi_n - \mu_{\text{intr}}}^2 = \min_{\mu \in [-\nicefrac{1}{2}, \nicefrac{1}{2})} \expect \fracpart{\Phi_n - \mu}^2,
+\sigma_{\text{intr}}^2 = \expect\fracpart{\Phi_n - \mu_{\text{intr}}}^2 = \min_{\mu \in [-\nicefrac{1}{2}, \nicefrac{1}{2})} \expect \fracpart{\Phi_n - \mu}^2,
 \]
 where $\expect$ denotes the expected value.
-Observe that if $\Phi_n$ has zero intrinsic mean then it also has zero mean and the variance and the intrinsic variance of $\Phi_n$ are equal.
+Observe that if $\Phi_n$ has zero intrinsic mean then it also has zero mean and the variance and the intrinsic variance of $\Phi_n$ coincide.
 The next theorem describes the asymptotic distribution of the LSU estimator.
+The theorem statement makes use of the function $\dealias(\cdot)$ that is defined in~\cite[Sec.~3]{McKilliam_pps1_2012}.
 
@@
 Let $\widehat{\psibf}_N$ be the vector with elements $\widehat{\psi}_{k,N} = N^k \widehat{\lambda}_{k,N}$ so that $\widehat{\psibf}_N$ is the minimiser of $T_N$.
-Because each of $N^k \widehat{\lambda}_{k,N}$ converges almost surely to zero as $N \rightarrow \infty$, then $\widehat{\psibf}_{N}$ converges almost surely to $\zerobf$ as $N \rightarrow \infty$.
+Because each of $N^k \widehat{\lambda}_{k,N}$ converges almost surely to zero as $N \rightarrow \infty$~\cite[Theorem~2]{McKilliam_pps1_2012}, it follows that $\widehat{\psibf}_{N}$ converges almost surely to $\zerobf$ as $N \rightarrow \infty$.
 We want to find the asymptotic distribution of
@@
 \]
-so that we now require to prove $\abs{\fracpart{B_n + \widehat{\psi}_{0,N}}} \leq \frac{1}{2} - \frac{1}{2N}$
-for all $n = 1, \dots, N$.  From~\eqref{eq:Wn}, $W_n = \round{B_n + \widehat{\psi}_{0,N}}$, and 
-\begin{align*}
-T_N(\widehat{\psibf}_N) &= \frac{1}{N}\sum_{n=1}^N \fracpart{B_n + \widehat{\psi}_{0,N}}^2
+so that we now require to prove $\sabs{\sfracpart{B_n + \widehat{\psi}_{0,N}}} \leq \frac{1}{2} - \frac{1}{2N}$
+for all $n = 1, \dots, N$.  From~\eqref{eq:Wn}, $W_n = \sround{B_n + \widehat{\psi}_{0,N}}$, and 
+\begin{align*}
+T_N(\widehat{\psibf}_N) &= \frac{1}{N}\sum_{n=1}^N \sfracpart{B_n + \widehat{\psi}_{0,N}}^2
 \\ &= \frac{1}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N} - W_n)^2.
@@
 \begin{equation}\label{eq:philamfraccontra}
-\fracpart{B_k + \widehat{\psi}_{0,N}} > \frac{1}{2} - \frac{1}{2N}.
+\sfracpart{B_k + \widehat{\psi}_{0,N}} > \frac{1}{2} - \frac{1}{2N}.
 \end{equation}
@@
 &= (B_k + \phi - W_k)^2 - 2(B_k + \widehat{\psi}_{0,N} - W_k) + 1 - \frac{2}{N} \nonumber \\
-&= (B_k + \phi - W_k)^2 - 2\fracpart{B_k + \widehat{\psi}_{0,N}} + 1 - \frac{2}{N} \nonumber \\
+&= (B_k + \phi - W_k)^2 - 2\sfracpart{B_k + \widehat{\psi}_{0,N}} + 1 - \frac{2}{N} \nonumber \\
 &< (B_k + \phi - W_k)^2 - \frac{1}{N}, \label{eq:Bkineq}
@@
 
-If $\fracpart{B_k + \widehat{\psi}_{0,N}} < -\frac{1}{2} + \frac{1}{2N}$ for some $k$, we set $F_k = W_k - 1$ and using the same procedure as before obtain $T_N(\bbf) < T_N(\widehat{\psibf}_N)$ again.
+If $\sfracpart{B_k + \widehat{\psi}_{0,N}} < -\frac{1}{2} + \frac{1}{2N}$ for some $k$, we set $F_k = W_k - 1$ and using the same procedure as before obtain $T_N(\bbf) < T_N(\widehat{\psibf}_N)$ again.
 The proof follows.
@@
 \end{equation}
-Observe that $G_N$ depends on $\ell$ and we could write $G_{\ell,N}$ but have suppressed the subscript $\ell$ for notational simplicity.  Now $W_n$ from~\eqref{eq:Wn} can be written as
-\[
-W_n = \round{\Phi_n + \widehat{p}_{nN}} = q_{n}(\widehat{p}_{nN})
-\]
+Observe that $G_N$ depends on $\ell$ and we could write $G_{\ell,N}$ but have suppressed the subscript $\ell$ for notational simplicity.
+Now $W_n$ from~\eqref{eq:Wn} can be written as $W_n = \round{\Phi_n + \widehat{p}_{nN}} = q_{n}(\widehat{p}_{nN})$
 and $K_{\ell,N}$ from~\eqref{eq:KellN} can be written as
@@
 \end{align*}
-Because $f(\fracpart{x})$ is continuous at $-\nicefrac{1}{2}$ it follows from the mean value theorem that
+Because $f(\fracpart{x})$ is continuous at $-\nicefrac{1}{2}$ it follows that 
 \[
@@
 Observe that $\abs{\widehat{p}_{nN}} \leq \sum_{k=0}^m\sabs{ \widehat{\psi}_{k,N}}$ and, since each of the $\widehat{\psi}_{k,N} \rightarrow 0$ almost surely as $N\rightarrow\infty$, it follows that $\widehat{p}_{nN} \rightarrow 0$ almost surely uniformly in $n = 1, \dots, N$ as $N\rightarrow\infty$.
-Thus, $\zeta(\widehat{p}_{nN}) \rightarrow 0$ almost surely (and therefore also in probability) uniformly in $n = 1, \dots, N$ as $N\rightarrow\infty$.
+Thus $\zeta(\widehat{p}_{nN}) \rightarrow 0$ almost surely (and therefore also in probability) uniformly in $n = 1, \dots, N$ as $N\rightarrow\infty$.
 Now,
@@
 
-This section describes the results of Monte-Carlo simulations with the least squares unwrapping (LSU) estimator and with the discrete poylnomial phase transform (DPT) estimator of Peleg and Porat~\cite{Peleg_DPT_1995}.
+This section describes the results of Monte-Carlo simulations with the least squares unwrapping (LSU) estimator, Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, and the discrete poylnomial phase transform (DPT) estimator of Peleg and Porat~\cite{Peleg_DPT_1995}.
 The sample sizes considered are $N = 10, 50, 200$ and the unknown amplitude is $\rho = 1$.
 The $X_1, \dots, X_N$ are pseudorandomly generated independent and identically distributed circularly symmetric complex Gaussian random variables with variance $\sigma_c^2$.
-The coefficients $\tilde{\mubf} = [\tilde{\mu}_0, \dots, \tilde{\mu}_m]$ are distributed uniformly randomly in the identifiable region $B$.
+The coefficients $\tilde{\mubf} = [\tilde{\mu}_0, \dots, \tilde{\mu}_m]$ are distributed uniformly randomly in the identifiable region $B$~\cite[Sec.~3]{McKilliam_pps1_2012}.
 The number of replications of each experiment is $T = 2000$ to obtain estimates $\widehat{\mubf}_1, \dots, \widehat{\mubf}_T$ and the corresponding dealiased errors $\widehat{\lambdabf}_t = \dealias(\widehat{\mubf}_T - \tilde{\mubf})$ are computed.
@@
 
-Figure~\ref{plot:polyest1} shows the sample MSEs obtained for a polynomial phase signal of order $m=3$.
+Figure~\ref{plot:polyest} shows the sample MSEs obtained for a polynomial phase signal of order $m=3$.
 %Results are displayed results for the zeroth and third order coefficients $\widehat{\mu}_1$ and $\widehat{\mu}_3$.
 The results for $\widehat{\mu}_1$ and $\widehat{\mu}_2$ lead to similar conclusions.
+The LSU estimator can be computed by finding a nearest lattice point in a particular lattice~\cite{McKilliam2010thesis,McKilliam2009asilomar_polyest_lattice}.
 When $N = 10$ and $50$ the LSU estimator can be computed exactly using a general purpose algorithm for finding nearest lattice points called the sphere~decoder~\cite{Pohst_sphere_decoder_1981,Agrell2002,Viterbo_sphere_decoder_1999}.
@@
 
+Kitchen's unwrapping estimator and the DPT estimator perform very poorly in Figure~\ref{plot:polyest}.
+The reason is that both estimators do not work correctly for all parameters in the identifiable region $B$.
+For this reaon Figure~\ref{plot:polyest2} plots the sample MSE's for the estimators with the true coefficients fixed to $\tilde{\mubf} = [0.1,0.2,10^{-3},10^{-5}]$.
+This is within the range suitable for the DPT estimator~\cite{Peleg_DPT_1995}.
+Both Kitchen's unwrapping estimator and the DPT estimator perform better in this case.
+Only results for the zeroth and third order parameters $\tilde{\mu}_0$ and $\tilde{\mu}_3$ are displayed.
+The results for the other parameters are similar.
+
 \begin{figure*}[p] 
@@
   		\includegraphics{code/gaussianplot4-1.mps} 
-  		\caption{Mean square error (MSE) of polynomial phase estimators for $N=10,50$ and $200$ versus the variance $\sigma_c^2$ of the complex noise terms $X_1, \dots X_N$.
-The polynomial phase signal has order 3, and the $X_1, \dots, X_N$ are independent and identically distributed complex Gaussian random variables.
+  		\caption{Mean square error (MSE) of polynomial phase estimators.
+The true coefficients $\tilde{\mubf}$ are uniformly distributed in the identifiable region $B$.
 (Top left) MSE in the phase coefficient $\mu_0$.
@@
  \end{figure*}
+
+\begin{figure}[p] 
+   	\centering 
+  		\includegraphics{code/gaussianplot4-2.mps} 
+  		\caption{Mean square error (MSE) of polynomial phase estimators.
+The true coefficients are fixed to $\tilde{\mubf} = [0.1,0.2,10^{-3},10^{-5}]$.
+(Top) MSE in the phase coefficient $\widehat{\mu}_0$.
+(Bottom right) MSE in the cubic coefficient $\widehat{\mu}_3$.}
+  		\label{plot:polyest2} 
+ \end{figure}
 
@@
 
-This paper has considered the estimation of the coefficients of a noisy polynomial phase signal by least squares phase unwrapping (LSU).
-It has been shown that the LSU estimator is strongly consistent and asymptotically normally distributed.
+This series of papers has considered the estimation of the coefficients of a noisy polynomial phase signal by least squares phase unwrapping (LSU).
+Under some assumptions on the distribution of the noise, it has been shown that the LSU estimator is strongly consistent and asymptotically normally distributed.
 %Simulations are used to show that the LSU has near maximum likelihood performance in the case that the noise is additive white and Gaussian.
-The LSU estimator can be computed by finding a nearest lattice point in a lattice~\cite{McKilliam2009asilomar_polyest_lattice}.
+The results of Monte Carlo experiment were described and these support the asymptotic analysis.
+
+It is shown in~\cite{McKilliam2010thesis,McKilliam2009asilomar_polyest_lattice} how the LSU estimator can be computed by finding a nearest lattice point in a particular lattice.
 %Here, we consider using the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder} to achieve this in reasonable time when the number of observations, $N$, is less than about 300.
 Polynomial time nearest point algorithms for these lattices exist~\cite[Sec 4.3]{McKilliam2010thesis}, but these algorithms are not fast in practice.
-Instead, we considered some general purpose algorithms, the \term{sphere decoder}, \term{Babai's nearest plane algorithm} and the \term{$K$-best algorithm}.
+%Instead, we considered some general purpose algorithms, the \term{sphere decoder}, \term{Babai's nearest plane algorithm} and the \term{$K$-best algorithm}.
 The sphere decoder and $K$-best algorithm result in accurate estimators.
-
 The major outstanding question is whether faster nearest point algorithms exist for these specific lattices.
@@
 \end{equation}
-since $\expect \expect_\Phi$ is equivalent to $\expect$.   Let
-\[
-\sigma_n = \frac{g_n}{\abs{g_n}},
-\]
-and put $\sigma_n = 1$ if $g_n = 0$.  The $\sigma_n$ are thus independent with 
-\[
-\prob\{ \sigma_n = -1 \} = \tfrac{1}{2} \qquad \text{and} \qquad  \prob\{ \sigma_n = 1 \} = \tfrac{1}{2}.
-\]  
+since $\expect \expect_\Phi$ is equivalent to $\expect$.
+Let $\sigma_n = g_n/\abs{g_n}$, and put $\sigma_n = 1$ if $g_n = 0$.
+The $\sigma_n$ are thus independent with $\prob\{ \sigma_n = -1 \} = \tfrac{1}{2}$ and $\prob\{ \sigma_n = 1 \} = \tfrac{1}{2}$.
+The symmetry of the distribution of $g_n$ implies that $g_n$ and $\sigma_n$ are independent.
 As $\Phi_n$ and $\Phi_n^\prime$ are independent and identically distributed, the random variable $f_{nN} - f_{nN}^\prime$ is symmetrically distributed about zero, and is therefore distributed identically to $\sigma_n( f_{nN} - f_{nN}^\prime)$.
@@
 \end{align*}
-Since $g_n$ is a standard normal random variable, $\expect \abs{g_n} = \expect_\Phi \abs{g_n} = \sqrt{2/\pi}$,
+Write $\expect_{\Phi\sigma}$ to denote expectation conditional on $\{\Phi_n\}$ and $\{\sigma_n\}$.
+Since $g_n$ is a standard normal random variable, $\expect \abs{g_n} = \expect_{\Phi\sigma} \abs{g_n} = \sqrt{2/\pi}$,
 and
@@
 \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  &\abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} } \\
-&= \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} \sqrt{\frac{\pi}{2}}\expect_\Phi \abs{g_n}} \\
+&= \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} \sqrt{\frac{\pi}{2}}\expect_{\Phi\sigma} \abs{g_n}} \\
 &\leq \sqrt{\frac{\pi}{2}} \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{ \sqrt{N}}\sum_{n=1}^{N} \sigma_n \abs{g_n}  f_{nN}},
 \end{align*}
-the last line following from Jensen's inequality and by moving $\expect_{\Phi}$ outside the supremum similarly to~\eqref{eq:Eoutsidesup}. As $g_n = \sigma_n \abs{g_n}$, it follows that
+the last line following from Jensen's inequality and by moving $\expect_{\Phi\sigma}$ outside the supremum similarly to~\eqref{eq:Eoutsidesup}. As $g_n = \sigma_n \abs{g_n}$, it follows that
 \begin{align*}
@@
 \end{align*}
-where $o(1)$ goes to zero as $\epsilon$ goes to zero, by the mean value theorem.
-Since $\sqrt{\cdot}$ is a concave function on the positive real line, and since $C_{\epsilon}(\{\Phi_n\})$ is always positive,
+where $o(1)$ goes to zero as $\epsilon$ goes to zero.  Since $\sqrt{\cdot}$ is a concave function on the positive real line, and since $C_{\epsilon}(\{\Phi_n\})$ is always positive,
 \[
@@
 
-The proofs of Lemmas~\ref{lem:chaining2}~and~\ref{lem:chaining3} are based on a technique called \emph{chaining} (or \emph{bracketing})~\cite{Dudley_unif_central_lim_th_1999,Ossiander_clt_bracketing_1984,Pollard_asymp_empi_proc_1989,Pollard_new_ways_clts_1986}.
-The proofs here follow those of Pollard~\cite[Section 3]{Pollard_asymp_empi_proc_1989}.
-Within the remaining proofs, we are interested in expectations conditional on the phase noise sequence $\{\Phi_n\}$, so for the purpose of these lemmas, $\{\Phi_n\}$ is treated as a fixed realisation.
-For notational convenience we use the abbreviations
+The proofs of Lemmas~\ref{lem:chaining2}~and~\ref{lem:chaining3} are based on a technique called \emph{chaining} (or \emph{bracketing})~\cite{Dudley_unif_central_lim_th_1999,Ossiander_clt_bracketing_1984,Pollard_asymp_empi_proc_1989,Pollard_new_ways_clts_1986,van2009empirical}.
+The proofs here follow those of Pollard~\cite{Pollard_asymp_empi_proc_1989}.
+In the remaining lemmas we consider expectation conditional on $\{\Phi_n\}$ and consequently treat $\{\Phi_n\}$ as a fixed realisation.
+We also use the abbreviations
 \[
@@
 \]
-in what follows.
 As in Lemma~\ref{lem:chaining} we assume, without loss of generality, that $\epsilon < \tfrac{1}{m+1}$.
@@
 \]
-and so we must have $d(\psibf, \psibf^*) = 0$.  Consequently $f_{nN}(\psibf) = f_{nN}(\psibf^*)$  for every $n = 1, \dots, N$, and therefore $Z_N(\psibf) - Z_N(\psibf^*) = 0$.  Thus,
+and so we must have $d(\psibf, \psibf^*) = 0$.  Consequently $f_{nN}(\psibf) = f_{nN}(\psibf^*)$  for every $n = 1, \dots, N$, and $Z_N(\psibf) = Z_N(\psibf^*)$.  Thus,
 \[
@@
 \begin{IEEEproof}
-Recall that
-\[
-f_{nN}(\psibf) = \left( \tfrac{n}{N}\right)^\ell \round{\Phi_n + p_{nN}(\psibf)}.
-\]
+%Recall that
+%\[
+%f_{nN}(\psibf) = \left( \tfrac{n}{N}\right)^\ell \round{\Phi_n + p_{nN}(\psibf)}.
+%\]
 Because $\abs{\psi_i} < \epsilon$ for all $i = 0, \dots, m$, 
@@
 \begin{IEEEproof}
-Let $b_k$ be a function that maps each $\psibf \in T_\epsilon(k)$ to $b_k(\psibf) \in T_\epsilon(k-1)$ such that 
-\[
-d(\psibf,b_k(\psibf)) \leq 2^{1-k} C_\epsilon N.
-\]
+Let $b_k$ be a function that maps each $\psibf \in T_\epsilon(k)$ to $b_k(\psibf) \in T_\epsilon(k-1)$ such that $d(\psibf,b_k(\psibf)) \leq 2^{1-k} C_\epsilon N$.
 The existence of the function $b_k$ is guaranteed by the definition of $T_\epsilon(k)$.  By the triangle inequality,
@@
 \end{align*}
-because $\log\abs{T_\epsilon(k)} \leq k(m+1)\log 2 +  \log K_3$.  Taking expectations on both side of~\eqref{eq:supZchain},
+because $\log\abs{T_\epsilon(k)} \leq k(m+1)\log 2 +  \log K_3$.  Taking expectations on both sides of~\eqref{eq:supZchain},
 \begin{align*}
@@
 \]
-Thus, $( g_{nN}(\psibf) - g_{nN}(\psibf^*) )^2$ is either equal to one when $g_{nN}(\psibf) \neq g_{nN}(\psibf^*)$ or zero when $g_{nN}(\psibf) = g_{nN}(\psibf^*)$.
+Thus $( g_{nN}(\psibf) - g_{nN}(\psibf^*) )^2$ is either equal to one when $g_{nN}(\psibf) \neq g_{nN}(\psibf^*)$ or zero when $g_{nN}(\psibf) = g_{nN}(\psibf^*)$.
 Now $g_{nN}(\psibf) \neq 0$ only if

### 2012-10-30 build.sh

@@ -19,3 +19,8 @@ pdflatex paper.tex
 bibtex paper.aux 
 pdflatex paper.tex 
 pdflatex paper.tex 
+
+pdflatex paper2.tex 
+bibtex paper2.aux 
+pdflatex paper2.tex 
+pdflatex paper2.tex 

### 2012-10-30 code/runsim.sh

@@ -1,4 +1,5 @@
 #export JAVA_OPTS="-Xprof -server -Xms1g -Xmx1g"
 export JAVA_OPTS="-d64 -server -Xms1g -Xmx1g"
-scala -nocompdaemon -cp PubSim.jar:Jama-1.0.2.jar:flanagan.jar:colt.jar:RngPack.jar simgaussian.scala
-scala -nocompdaemon -cp PubSim.jar:Jama-1.0.2.jar:flanagan.jar:colt.jar:RngPack.jar simwrappedu.scala
+#scala -nocompdaemon -cp PubSim.jar:Jama-1.0.2.jar:flanagan.jar:colt.jar:RngPack.jar simgaussian.scala
+scala -nocompdaemon -cp PubSim.jar:Jama-1.0.2.jar:flanagan.jar:colt.jar:RngPack.jar simgaussianfixedmu.scala
+#scala -nocompdaemon -cp PubSim.jar:Jama-1.0.2.jar:flanagan.jar:colt.jar:RngPack.jar simwrappedu.scala

### 2012-10-30 code/simgaussian.scala

@@ -19,7 +19,7 @@ import pubsim.distributions.circular.ProjectedNormalDistribution
 
 val iters = 2000 //number of Monte-Carlo trials.
 val Ns = List(10,50,200) //values of N we will generate curves for
-val ms = List(2,3) //order of our polynomial phase signals
+val ms = List(3) //order of our polynomial phase signals
 
 def gp( a : Double, r : Double, n : Int ) : Double = a * scala.math.pow(r,n)
 
@@ -31,12 +31,12 @@ def estfactory(m : Int, N : Int) : List[() => PolynomialPhaseEstimator] = {
   var ret = List( 
     () => new KitchenEstimator(m,N),
     () => new DPTEstimator(m,N),
-    () => new BabaiEstimator(m,N),
+    //() => new BabaiEstimator(m,N),
     () => new MbestEstimator(m,N,4*N) 
   )
   //add the sphere decoder and Least squares estimators if N and m are small
   if( N < 60 ) ret = ret :+ ( () => new SphereDecoderEstimator(m,N) )
-  if( N < 60 && m <= 2 ) ret = ret :+ ( () => new MaximumLikelihood(m,N) )
+  //if( N < 60 && m <= 2 ) ret = ret :+ ( () => new MaximumLikelihood(m,N) )
   return ret
 }
 
@@ -47,7 +47,7 @@ for( N <-  Ns; m <- ms ) {
 
   for(estf <- estfactory(m,N) ){
     
-    val estname = estf().getClass.getSimpleName + "N" + N.toString + "m" + m + "WU"
+    val estname = estf().getClass.getSimpleName + "N" + N.toString + "m" + m + "Gaussian"
     print("Running " + estname)
     val eststarttime = (new java.util.Date).getTime
     

### 2012-10-30 paper2.tex
Added to paper 2.  Hopefully nothing else needs to be done with this.

@@
 The first part in the series~\cite{McKilliam_pps1_2012} asserts the strong consistency of the LSU estimator under some conditions on the distribution on the noise $\{X_n\}$.
-In this second part we prove, under similar conditions on $\{X_n\}$, that the LSU estimator is asymptotically normallly distributed.
+In this second part we prove, under similar conditions on $\{X_n\}$, that the LSU estimator is asymptotically normally distributed.
 
@@
 \] 
-are random variables representing the \emph{phase noise} observed at the reciever~\cite{Tretter1985}.
+are random variables representing the \emph{phase noise} observed at the receiver~\cite{Tretter1985}.
 We denote by $\angle{z}$ the complex argument of the complex number $z$, and by $\fracpart{x} = x - \round{x}$ the (centered) fractional part of the real number $x$, and by $\round{x}$ the nearest integer to $x$.
@@
 where $\expect$ denotes the expected value.  
-Observe that if $\Phi_n$ has zero intrinsic mean then it also has zero mean and the variance and the intrinsic variance of $\Phi_n$ coincide.
-The next theorem describes the asymptotic distribution of the LSU estimator.
-The theorem statement makes use of the function $\dealias(\cdot)$ that is defined in~\cite[Sec.~3]{McKilliam_pps1_2012}.
+Observe the following property of circular random variables with zero intrinsic mean.
+\begin{proposition}\label{prop:zerointmeanzeromean}
+Let $\Phi$ be a circular random variable with intrinsic mean $\mu_{\text{intr}} = 0$ and intrinsic variance $\sigma^2$.
+Then $\Phi$ has zero mean and variance $\sigma^2$, that is, $\expect \Phi = 0$ and $\expect (\Phi - \expect \Phi)^2  = \sigma^2$.
+\end{proposition}
+\begin{proof}
+Assume the proposition is false and that $\mu = \expect \Phi \neq 0$.  But, then 
+\begin{align*}
+\sigma^2 &= \expect\fracpart{\Phi -  \mu_{\text{intr}}}^2 = \expect\fracpart{\Phi}^2 \\
+&= \expect \Phi^2 >  \expect (\Phi - \mu)^2 \geq  \expect \fracpart{\Phi - \mu}^2,
+\end{align*}
+violating the fact that $\mu_{\text{intr}} = 0$ is the minimiser of~\eqref{eq:intrmeandefn}.
+\end{proof}
+The next theorem describes the asymptotic distribution of the LSU estimator.  The theorem statement makes use of the function $\dealias(\cdot)$ that is defined in~\cite[Sec.~3]{McKilliam_pps1_2012}.
 
@@
 \]
+Observe that $\Phi_n$ has zero mean and variance $\sigma^2$ as a result of Proposition~\ref{prop:zerointmeanzeromean}.
 By Lypanov's central limit theorem $\zbf_N$ is asymptotically normally distributed with zero mean and variance
@@
 \]
-By the Cramer-Wold theorem it follows that $\dbf_N$ is asymptotically normally distributed with zero mean and covariance $\sigma^2 \Cbf$. 
+By the Cram\'{e}r-Wold theorem it follows that $\dbf_N$ is asymptotically normally distributed with zero mean and covariance $\sigma^2 \Cbf$. 
 % Recall that the $\ell$th element of $\dbf_N$ is
@@
 
-This section describes the results of Monte-Carlo simulations with the least squares unwrapping (LSU) estimator, Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, and the discrete poylnomial phase transform (DPT) estimator of Peleg and Porat~\cite{Peleg_DPT_1995}.
+This section describes the results of Monte-Carlo simulations with the least squares unwrapping (LSU) estimator, Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, and the discrete polynomial phase transform (DPT) estimator of Peleg and Porat~\cite{Peleg_DPT_1995}.
 The sample sizes considered are $N = 10, 50, 200$ and the unknown amplitude is $\rho = 1$.
@@
 Provided the noise variance is small enough (so that the `threshold' is avoided) the sample MSE of the LSU estimator is close to that predicted by Theorem~\ref{thm:asymp_proof}.
-The Cram\`{e}r-Rao lower bound for the variance of unbiased polynomial phase estimators in Gaussian noise is also plotted using the solid line~\cite{Peleg1991_CRB_PPS_1991}.
-When the noise variance is small the asymptotic variance of the LSU estimator is close to the Cram\`{e}r-Rao lower bound.
+The Cram\'{e}r-Rao lower bound for the variance of unbiased polynomial phase estimators in Gaussian noise is also plotted using the solid line~\cite{Peleg1991_CRB_PPS_1991}.
+When the noise variance is small the asymptotic variance of the LSU estimator is close to the Cram\'{e}r-Rao lower bound.
 
@@
 The reason is that both estimators do not work correctly for all parameters in the identifiable region $B$.
-For this reaon Figure~\ref{plot:polyest2} plots the sample MSE's for the estimators with the true coefficients fixed to $\tilde{\mubf} = [0.1,0.2,10^{-3},10^{-5}]$.
+For this reason Figure~\ref{plot:polyest2} plots the sample MSEs for the estimators with the true coefficients fixed to $\tilde{\mubf} = [0.1,0.2,10^{-3},10^{-5}]$.
 This is within the range suitable for the DPT estimator~\cite{Peleg_DPT_1995}.

### 2012-10-31 paper.tex
minor

@@
 %Simulations are used to show that the LSU has near maximum likelihood performance in the case that the noise is additive white and Gaussian.
-In the second paper in this series~\cite{McKilliam_pps2_2012} we show the LSU estimator to be asymptotically normal and present the results of Monte Carlo simulations that support our asymptotic results.
+In the second paper in this series~\cite{McKilliam_pps2_2012} we show the LSU estimator to be asymptotically normally distributed and present the results of Monte Carlo simulations that support our asymptotic results.
 

### 2012-11-02 paper.tex
minor

@@
 
-Although we will not prove it here the assumption that $\Phi_1,\dots,\Phi_N$ have zero intrinsic mean is not only sufficient, but also necessary, for if $\Phi_1,\dots,\Phi_N$ have intrinsic mean $x \in [-\nicefrac{1}{2},\nicefrac{1}{2})$ with $x \neq 0$ then $\fracpart{\widehat{\lambda}_{0,N} - x}\rightarrow 0$ almost surely as $N\rightarrow\infty$, and so $\widehat{\lambda}_{0,N}$ does not converge to zero.
+Although we will not prove it here the assumption that $\Phi_1,\dots,\Phi_N$ have zero intrinsic mean is not only sufficient, but also necessary, for if $\Phi_1,\dots,\Phi_N$ have intrinsic mean $x \in [-\nicefrac{1}{2},\nicefrac{1}{2})$ with $x \neq 0$ then $\sfracpart{\widehat{\lambda}_{0,N} - x}\rightarrow 0$ almost surely as $N\rightarrow\infty$, and so $\widehat{\lambda}_{0,N}$ does not converge to zero.
 On the other hand if $\Phi_1,\dots,\Phi_N$ do not have an intrinsic mean then $\widehat{\lambda}_{0,N}$ will not converge.

### 2012-11-02 paper.tex
added simple proposition

--- 2012-10-30 paper.tex
@@
-%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
-\documentclass[journal]{IEEEtran}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGIN AOAP START %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Template for the submission to:
+%   The Annals of Probability           [aop]
+%   The Annals of Applied Probability   [aap]
+%   The Annals of Statistics            [aos] 
+%   The Annals of Applied Statistics    [aoas]
+%   Stochastic Systems                  [ssy]
+%
+%Author: In this template, the places where you need to add information
+%        (or delete line) are indicated by {???}.  Mostly the information
+%        required is obvious, but some explanations are given in lines starting
+%Author:
+%All other lines should be ignored.  After editing, there should be
+%no instances of ??? after this line.
+
+% use option [preprint] to remove info line at bottom
+% journal options: aop,aap,aos,aoas,ssy
+% natbib option: authoryear
+\documentclass[aap]{imsart}
+
+\usepackage{amsthm,amsmath,amsfonts,amssymb}
+\usepackage[square,comma,numbers,sort&compress]{natbib}
+%\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
+\usepackage{units}
+\usepackage{ifpdf}
+\ifpdf
+  \usepackage[pdftex]{graphicx}
+\else
+  \usepackage{graphicx}% standard graphics package for inclusion images and eps files into LaTeX document
+\fi
+
+% provide arXiv number if available:
+%\arxiv{arXiv:0000.0000}
+
+% put your definitions there:
+\startlocaldefs
+\newcommand{\reals}{{\mathbb R}}
+\newcommand{\ints}{{\mathbb Z}}
+\newcommand{\term}{\emph}
+\newcommand{\var}{\operatorname{var}}
+\newcommand{\covar}{\operatorname{covar}}
+%\newcommand{\prob}{{\mathbb P}}
+\newcommand{\prob}{\operatorname{Pr}}
+\newcommand{\expect}{{\mathbb E}}
+\newcommand{\dealias}{\operatorname{dealias}}
+\renewcommand{\mid}{\; ; \;}
+\newcommand{\fracpart}[1]{\left\langle #1 \right\rangle}
+\newcommand{\sfracpart}[1]{\langle #1 \rangle}
+\newcommand{\abs}[1]{\left\vert #1 \right\vert}
+\newcommand{\sabs}[1]{\vert #1 \vert}
+\newcommand{\floor}[1]{{\left\lfloor #1 \right\rfloor}}
+\newcommand{\ceiling}[1]{{\left\lceil #1 \right\rceil}}
+\newcommand{\ceil}[1]{\lceil #1 \rceil}
+\newcommand{\round}[1]{{\left\lceil #1 \right\rfloor}}
 \usepackage{mathbf-abbrevs}
-
-%\newcommand{\dealias}{\operatorname{dealias}}
-
-\input{defs}
-
-%\newcommand{\term}{\emph}
-%\renewcommand{\index}[1]{}
-
-\title{Polynomial phase estimation by phase unwrapping 1: Identifiability and strong consistency}
-%\title{Polynomial phase estimation by phase unwrapping}
-
-\author{Robby~G.~McKilliam, Barry~G.~Quinn, I.~Vaughan~L.~Clarkson, Bill~Moran and Badri~N.~Vellambi%
-    \thanks{%A preliminary version of some of this material is contained in Part 3 of Robby McKilliam's PhD thesis \cite{McKilliam2010thesis}. 
-Robby~McKilliam and Badri Vellambi are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.
-Barry~Quinn is with the Department of Statistics, Macquarie University, Sydney, NSW, 2109, Australia.
-Vaughan~Clarkson is with the School of Information Technology \& Electrical Engineering, The University of Queensland, QLD., 4072, Australia.
-B.
-Moran is with the Department of Electrical Engineering and Computer
-Science, Melbourne Systems Lab, Dept of Elec \& Electronic Eng, Uni of Melbourne, Vic. 3010, Australia.}}
-% The paper headers 
-\markboth{Polynomial phase estimation by phase unwrapping}{DRAFT \today}
-
-% make the title area
+\newtheorem{proposition}{Proposition}
+\newtheorem{corollary}{Corollary}
+\newtheorem{theorem}{Theorem}
+\newtheorem{lemma}{Lemma}
+\newtheorem{definition}{Definition}
+\newcommand{\coef}{\operatorname{coef}}
+\newcommand{\cubr}[1]{{\left\{ #1 \right\}}}
+\newcommand{\scubr}[1]{{\{ #1 \}}}
+\newcommand{\sign}[1]{\operatorname{sgn}\left( #1 \right)}
+\newcommand{\bfracpart}[1]{\big\langle #1 \big\rangle}
+\newcommand{\babs}[1]{\big\vert #1 \big\vert}
+\endlocaldefs
 
 \begin{document}
-\maketitle
+
+\begin{frontmatter}
+
+% "Title of the paper"
+\title{Polynomial phase estimation by phase unwrapping}
+\runtitle{Polynomial phase estimation by phase unwrapping}
+
+% indicate corresponding author with \corref{}
+% \author{\fnms{John} \snm{Smith}\corref{}\ead[label=e1]{smith@foo.com}\thanksref{t1}}
+% \thankstext{t1}{Thanks to somebody} 
+% \address{line 1\\ line 2\\ printead{e1}}
+% \affiliation{Some University}
+
+\begin{aug}
+\author{\fnms{Robby G.} \snm{McKilliam}\thanksref{m1}\ead[label=e1]{robby.mckilliam@unisa.edu.au}},
+\author{\fnms{Barry G.} \snm{Quinn}\thanksref{m2}\ead[label=e2]{barry.quinn@mq.edu.au}},
+\author{\fnms{I. Vaughan L.} \snm{Clarkson}\thanksref{m3}\ead[label=e3]{v.clarkson@uq.edu.au}},
+\author{\fnms{Bill} \snm{Moran}\thanksref{m4}\ead[label=e4]{wmoran@unimelb.edu.au}},
+\and
+\author{\fnms{Badri N.} \snm{Vellambi}\thanksref{m1}\ead[label=e5]{badri.vellambi@unisa.edu.au}}
+\runauthor{R. McKilliam et. al.}
+
+\affiliation{University of South Australia\thanksmark{m1}, Macquarie University\thanksmark{m2}, University of Queensland\thanksmark{m3} and Melbourne University\thanksmark{m4}}
+
+\address{R. G. McKilliam and B. N. Vellambi\\
+Institute for Telecommunications Research\\
+University of South Australia, Adelaide\\
+SA, 5095, Australia\\
+\printead{e1}\\
+\phantom{E-mail:\ }\printead*{e5}}
+
+\address{B. G. Quinn\\
+Department of Statistics\\
+Macquarie University, Sydney,\\
+NSW, 2109, Australia\\
+\printead{e3}}
+
+\address{I. V. L. Clarkson\\
+School of Information Technology \& Electrical\\
+Engineering, University of Queensland,\\
+Brisbane, QLD, 4072, Australia\\
+\printead{e3}}
+
+\address{B. Moran\\
+Department of Electrical \& Electronic\\
+Engineering, Melbourne University,\\
+Melbourne, VIC, 3010, Australia\\
+\printead{e4}}
+
+\end{aug}
 
@@
 %It was shown by Monte Carlo simulation that this produces a remarkably accurate estimator.  
-In this two part series of papers we consider an estimator that performs phase unwrapping in a least squares manner.
-In this first part we describe conditions for the identifiability of polynomial phase signals and we prove the strong consistency of our unwrapping estimator.
-%We describe the asymptotic properties of this estimator, showing that it is strongly consistent and asymptotically normally distributed.
+In this paper we consider an estimator that performs phase unwrapping in a least squares manner.
+We describe the asymptotic properties of this estimator, showing that it is strongly consistent and asymptotically normally distributed.
 %We hypothesise that the estimator produces very near maximum likelihood performance.
@@
 
-\begin{keywords}
-Polynomial phase signals, phase unwrapping, asymptotic properties, nearest lattice point problem
-\end{keywords}
-
-%\bibliographystyle{unsrt}
+%\begin{keyword}[class=AMS]
+%\kwd[Primary ]{}
+%\kwd{}
+%\kwd[; secondary ]{}
+%\end{keyword}
+
+\begin{keyword}
+\kwd{Polynomial phase signals, phase unwrapping, asymptotic statistics}
+\end{keyword}
+
+\end{frontmatter}
+
+% AOS,AOAS: If there are supplements please fill:
+% \begin{supplement}[id=suppA]
+%  \sname{Supplement A}
+%  \stitle{Title}
+%  \slink[url]{http://lib.stat.cmu.edu/aoas/???/???}
+%  \sdescription{Some text}
+% \end{supplement}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%% END AOAP START %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGIN IEEE START %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% %\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
+% \documentclass[journal]{IEEEtran}
+
+% \usepackage{mathbf-abbrevs}
+
+% %\newcommand{\dealias}{\operatorname{dealias}}
+
+% \input{defs}
+
+% %\newcommand{\term}{\emph}
+% %\renewcommand{\index}[1]{}
+
+% %\title{Polynomial phase estimation by phase unwrapping 1: Identifiability and strong consistency}
+% \title{Polynomial phase estimation by phase unwrapping}
+
+% \author{Robby~G.~McKilliam, Barry~G.~Quinn, I.~Vaughan~L.~Clarkson, Bill~Moran and Badri~N.~Vellambi%
+%     \thanks{%A preliminary version of some of this material is contained in Part 3 of Robby McKilliam's PhD thesis \cite{McKilliam2010thesis}. 
+% Robby~McKilliam and Badri Vellambi are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.
+Barry~Quinn is with the Department of Statistics, Macquarie University, Sydney, NSW, 2109, Australia.
+Vaughan~Clarkson is with the School of Information Technology \& Electrical Engineering, The University of Queensland, QLD., 4072, Australia.
+B.
+Moran is with the Department of Electrical Engineering and Computer
+% Science, Melbourne Systems Lab, Dept of Elec \& Electronic Eng, Uni of Melbourne, Vic. 3010, Australia.}}
+% % The paper headers 
+% \markboth{Polynomial phase estimation by phase unwrapping}{DRAFT \today}
+
+% % make the title area
+
+% \begin{document} 
+% \maketitle
+
+% \begin{abstract}
+% Estimating the coefficients of a noisy polynomial phase signal is important in many fields including radar, biology and radio communications.
+One approach attempts to perform polynomial regression on the phase of the signal.
+This is complicated by the fact that the phase is \emph{wrapped} modulo $2\pi$ and must be \emph{unwrapped} before regression can be performed.
+%A recent approach suggested by the authors is to perform the unwrapping in a least squares manner.
+%It was shown how the resulting estimator could be posed as a \emph{nearest lattice point problem} in a specific lattice and could be solved using existing algorithms.
+% %It was shown by Monte Carlo simulation that this produces a remarkably accurate estimator.  
+% In this paper we consider an estimator that performs phase unwrapping in a least squares manner.
+We describe the asymptotic properties of this estimator, showing that it is strongly consistent and asymptotically normally distributed.
+%We hypothesise that the estimator produces very near maximum likelihood performance.
+% \end{abstract}
+
+% \begin{keywords}
+% Polynomial phase signals, phase unwrapping, asymptotic properties, nearest lattice point problem
+% \end{keywords}
+
+% %\bibliographystyle{unsrt}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%% END IEEE START %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-Polynomial phase signals arise in fields including radar, sonar, geophysics, biology, and radio communication \cite{Angeby_estimating_2000}.
-In radar and sonar applications polynomial phase signals arise when acquiring radial velocity and acceleration (and higher order motion descriptors) of a target from a reflected signal, and also in continuous wave radar and low probability of intercept radar~\cite{Levanon_Radar_signals_2004}.
-In biology, polynomial phase signals are used to describe the sounds emitted by bats and dolphins for echo location \citep{Suga_1975_bats_echolocation, Moss_2005echolocation}.
+Polynomial phase signals arise in fields including radar, sonar, geophysics, speech analysis, biology, and radio communication~\cite{Hlawatsch_lin_quad_time_freq_spmag_1992,Ridleyspeechpolyphase1989, Suga_1975_bats_echolocation, Moss_2005echolocation}.
+In radar and sonar applications polynomial phase signals arise when acquiring radial velocity and acceleration (and higher order motion descriptors) of a target from a reflected signal, and also in continuous wave radar and low probability of intercept radar.
+In biology, polynomial phase signals are used to describe the sounds emitted by bats and dolphins for echo location.
 
@@
 In this paper we only consider uniform sampling, where the gap between consecutive samples is constant.
-In this case we can always consider the samples to be taken at some set of consecutive integers and our sampled polynomial phase signal looks like
-\[
-s_n = s(n) = e^{2\pi j y(n)},
-\] 
-where $n$ is an integer.  Of practical importance is the estimation of the coefficients $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ from a number, say $N$, of observations of the noisy sampled signal
+In this case we can always consider the samples to be taken at some set of consecutive integers and our sampled polynomial phase signal is $s_n = s(n) = e^{2\pi j y(n)}$, where $n$ is an integer.
+Of practical importance is the estimation of the coefficients $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ from a number, say $N$, of consecutive observations of the noisy sampled signal
 \begin{equation}\label{eq:Y_nsamplednoisey}
@@
 \end{equation}
-where $\rho$ is a positive real number representing the (usually unknown) signal amplitude and $\{X_n, n \in \ints\}$ is a sequence of complex noise variables.
+where $\rho$ is a real number greater than zero representing the (usually unknown) signal amplitude and $\{X_n, n \in \ints\}$ is a sequence of complex noise variables.
 In order to ensure identifiability it is necessary to restrict the $m+1$ coefficients to a region of $m+1$ dimensional Euclidean space $\reals^{m+1}$ called an \emph{identifiable region}.
-It was shown in \cite{McKilliam2009IndentifiabliltyAliasingPolyphase} that an identifiable region tessellates a particular $m+1$ dimensional lattice.
+It was shown by some of the authors \cite{McKilliam2009IndentifiabliltyAliasingPolyphase} that an identifiable region tessellates a particular $m+1$ dimensional lattice.
 We discuss this in Section \ref{sec:ident_aliasing}.
@@
 
-%An obvious estimator of the unknown coefficients is the least squares estimator (LSE).
-This is also the maximum likelihood estimator (MLE) when the noise sequence $\{X_n\}$ is white and Gaussian.
-When $m=0$ (phase estimation) or $m=1$ (frequency estimation) the LSE is an effective approach, being both computationally efficient and statistically accurate \cite{Quinn2009_dasp_phase_only_information_loss,Hannan1973,Quinn2001,McKilliam_mean_dir_est_sq_arc_length2010}\cite[Sec.~6.4~and~9.1]{McKilliam2010thesis}.
-When $m \geq 2$ the computational complexity of the LSE is large~\cite[Sec.~10.1]{McKilliam2010thesis}\cite{Abatzoglou_ml_chirp_1986}.
-For this reason many authors have considered alternative approaches to polynomial phase estimation.
-These can loosely be grouped into two classes, estimators based on polynomial phase transforms, such as the discrete polynomial phase transform (DPT)~\cite{Peleg_DPT_1995,Peleg1991_est_class_PPS_1991,Porat_asympt_HAF_DPT_1996} and the high order phase function~\cite{Farquharson_another_poly_est_2005}, and estimators based on phase unwrapping, such as Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, the estimator of Djuric and Kay~\cite{Djuric_phase_unwrap_chirp_1990}, and Morelande's Bayesian unwrapping estimator~\cite{Morelande_bayes_unwrapping_2009_tsp}.
-
-One estimation approach attempts to perform polynomial regression on the phase of the signal~\cite{Slocumb_polynomial_1994,Tretter1985,Kitchen_polyphase_unwrapping_1994,Djuric_phase_unwrap_chirp_1990,Morelande_bayes_unwrapping_2009_tsp}.
-This is complicated by the fact that the phase is \emph{wrapped} modulo $2\pi$ and must be \emph{unwrapped} before regression can be performed.
-In this two part series of papers we consider the estimator that results from unwrapping the phase in a least squares manner.
-We call this the \emph{least squares unwrapping estimator} (LSU)~\cite{McKilliam2009asilomar_polyest_lattice, McKilliamFrequencyEstimationByPhaseUnwrapping2009}\cite[Chap.
-8]{McKilliam2010thesis}.
-It was shown in \cite{McKilliam2009asilomar_polyest_lattice, McKilliamFrequencyEstimationByPhaseUnwrapping2009} that the LSU estimator can be represented as a \emph{nearest lattice point problem}, and Monte-Carlo simulations were used to show the LSU estimator's favourable statistical performance.
+An obvious estimator of the unknown coefficients is the least squares estimator.  
+%This is also the maximum likelihood estimator (MLE) when the noise sequence $\{X_n\}$ is white and Gaussian.  
+When $m=0$ (phase estimation) or $m=1$ (frequency estimation) the least squares estimator is an effective approach, being both computationally efficient and statistically accurate~\cite{Hannan1973,Quinn2001,McKilliam_mean_dir_est_sq_arc_length2010}.
+When $m \geq 2$ the computational complexity of the least squares estimator is large, and alternative estimators have been considered for this reason.
+These can loosely be grouped into two classes; estimators based on polynomial phase transforms, such as the discrete polynomial phase transform~\cite{Peleg_DPT_1995} and the high order phase function~\cite{Farquharson_another_poly_est_2005,Porat_asympt_HAF_DPT_1996}; and estimators based on phase unwrapping, such as Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, and Morelande's Bayesian unwrapping estimator~\cite{Morelande_bayes_unwrapping_2009_tsp}.
+
+In this paper we consider the estimator that results from unwrapping the phase in a least squares manner.
+We call this the \emph{least squares unwrapping} (LSU) estimator.
+It was shown by some of the authors \cite[Sec.
+8.1]{McKilliam2010thesis}\cite{McKilliam2009asilomar_polyest_lattice} that the LSU estimator can be computed by finding a nearest point in a lattice~\cite{Agrell2002}, and Monte-Carlo simulations were used to show the LSU estimator's favourable statistical performance.
 %Moreover, it was noticed that the LSU appears to work correctly for coefficients anywhere in the identifiable region.
@@
 %We also apply another known approximate nearest point algorithm, called the $K$-best method \cite{Zhan2006_K_best_sphere_decoder}, that we show provides near sphere decoder performance, but can be computed in a reasonable amount of time if $N$ is less than about 300.
-In this two part series of we derive the asymptotic statistical properties of the LSU estimator showing, under some assumptions on the noise $X_1,\dots,X_N$, that it is strongly consistent and asymptotically normally distributed.
-Strong consistency is proved in this paper, while asymptotic normality is proved in the second part~\cite{McKilliam_pps2_2012}.
-Similar results were stated without a complete proof in~\cite{McKilliam_polyphase_est_icassp_2011}.
+In this paper we derive the asymptotic properties of the LSU estimator.
+Under some assumptions about the distribution of the noise $X_1, \dots, X_N$, we show the estimator to be strongly consistent and asymptotically normally distributed.
+Similar results were stated without proof in~\cite{McKilliam_polyphase_est_icassp_2011}.
 Here, we give a proof.
@@
 
-An interesting property is that the estimator of the $k$th polynomial phase coefficient converges to $\tilde{\mu}_k$ at rate $o(N^{-k})$.
+An interesting property is that the estimator of the $k$th polynomial phase coefficient converges almost surely to $\tilde{\mu}_k$ at rate $o(N^{-k})$.
 This is perhaps not surprising, since it is the same rate observed in polynomial regression.
@@
 %This proof technique appears to be novel, and may be useful in other problems.
-%The proof of asymptotic normality is complicated by the fact that the objective function corresponding with the LSU estimator is not differentiable everywhere.
+The proof of asymptotic normality is complicated by the fact that the objective function corresponding with the LSU estimator is not differentiable everywhere.
 Empirical process techniques~\cite{Pollard_new_ways_clts_1986,Pollard_asymp_empi_proc_1989,van2009empirical,Dudley_unif_central_lim_th_1999} and results from the literature on hyperplane arrangements~\cite{Chazelle_discrepency_method_2000,Matousek_lect_disc_geom_2002} become useful here.
@@
 The paper is organised in the following way.
-Section~\ref{sec:lattice-theory} describes some required concepts from lattice theory.
-In Section~\ref{sec:ident_aliasing} we describe the identifiable region that was also derived in~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}.
+Section~\ref{sec:lattice-theory} describes some preliminary concepts from lattice theory, and in Section~\ref{sec:ident_aliasing} we use these results to describe an identifiable region for the set of polynomial phase coefficients.
 These identifiability results are required in order to properly understand the statistical properties of polynomial phase estimators.
-Section~\ref{sec:least-squar-unwr} describes the LSU estimator and states a theorem asserting the estimator to be strongly consistent under some assumptions on the noise $X_1,\dots,X_N$.
-The theorem is proved in Section~\ref{sec:strongconstproof}.
-In the second paper~\cite{McKilliam_pps2_2012} we prove the asymptotic normality of the LSU estimator and describe the results of Monte Carlo simulations.
-These simulations agree with our derived asymptotic statistical properties.
-
-%\subsection{Notation}
-%We write random variables using capital letters, such as $X$ and $Y$ and circular random variables using the capital Greek letters $\Theta$ and $\Phi$.
-We use $\round{x}$ to denote the nearest integer to $x$ with half integers rounded up and use $\fracpart{x} = x - \round{x}$ to denote the \emph{centred} fractional part.
-Both $\round{\cdot}$ and $\fractpart{\cdot}$ operate elementwise on vectors.
-
-\section{Lattice Theory}\label{sec:lattice-theory}
+In Section \ref{sec:least-squar-unwr} we describe the LSU estimator and state its asymptotic statistical properties.
+Section~\ref{sec:strongconstproof} gives the proof of strong consistency and Section~\ref{sec:centlimitproof} gives the proof of asymptotic normality.
+%and also consider techniques for computing the LSU as a nearest lattice point problem. 
+Section~\ref{sec:simulations} describes the results of Monte Carlo simulations with the LSU estimator.  These simulations agree with the derived asymptotic properties.
+
+\section{Lattices}\label{sec:lattice-theory}
 
@@
 One choice for a set of coset representatives is a fundamental parallelepiped; the parallelepiped generated by the columns of a generator matrix.
-Another choice is based on the Voronoi cell; those points from $\reals^n$ nearest (with respect to the Euclidean norm in this paper) to the lattice point at the origin.
+Another choice is based on the Voronoi cell those points from $\reals^n$ nearest (with respect to the Euclidean norm here) to the lattice point at the origin.
 It is always possible to construct a rectangular set of representatives, as the next proposition will show.
 We will use these rectangular regions for describing the aliasing properties of polynomial phase signals in Section~\ref{sec:ident_aliasing}.
-These rectangular regions will be important for the derivation of the asymptotic properties of the LSU estimator.
-% in Section~\ref{sec:least-squar-unwr}.
+These rectangular regions will be important for the derivation of the asymptotic properties of the LSU estimator in Section~\ref{sec:least-squar-unwr}.
 
@@
 Let $\Bbf = \Qbf\Rbf$ where $\Qbf$ is orthonormal and $\Rbf$ is upper triangular with elements $r_{ij}$.
-Then the rectangular prism $\Qbf P$ where
-\[
-P = \prod_{k=1}^{n}{\left[-\frac{r_{kk}}{2}, \frac{r_{kk}}{2}\right)}
-\]
-is a set of coset representatives for $\reals^n / \Lambda$.
+Then the rectangular prism $\Qbf P$ where $P = \prod_{k=1}^{n}{[-\tfrac{r_{kk}}{2}, \frac{r_{kk}}{2})}$ is a set of coset representatives for $\reals^n / \Lambda$.
 \end{proposition}
-\begin{IEEEproof}
+\begin{proof}
 This result is well known~\cite[Chapter IX, Theorem IV]{Cassels_geom_numbers_1997}~\cite[Proposition 2.1]{McKilliam2010thesis}.
@@
 A result in the general case can be obtained similarly, but is not required here.
-\end{IEEEproof}
-
-\begin{figure}[tp]
+\end{proof}
+
+\begin{figure}[t]
 	\centering
-		\includegraphics[width=\linewidth]{plots/tesselationfigures-2.mps}
+		\includegraphics[width=0.8\linewidth]{plots/tesselationfigures-2.mps}
 		\caption{Rectangular tessellation constructed according to Proposition~\ref{prop:lattice-theory-constructing_a_rectangular_tesselating_region} where $\Lambda$ is a 2 dimensional lattice with generator matrix having columns $[1, 0.2]^\prime$ and $[0.2, 1]^\prime$.
@@
 \end{figure}
+
+% Given a point $\ybf\in\reals^n$ and a lattice $\Lambda$ contained in~$\reals^n$, the \emph{nearest lattice point problem} is to find the point $\xbf \in \Lambda$ such that the distance (with respect to a given norm) between $\ybf$ and $\xbf$ is minimised.
+Here, the Euclidean norm is assumed, so that the nearest point is
+% \[ 
+% \arg\min_{\xbf \in \Lambda} \| \xbf - \ybf \|^2.
+% \]
+% The nearest lattice point problem is known to be NP-hard under certain conditions when the lattice itself, or rather a basis thereof, is considered as an additional input parameter~\cite{micciancio_hardness_2001, Jalden2005_sphere_decoding_complexity}.
+Nevertheless, algorithms exist that can compute a nearest lattice point in reasonable time if the dimension is small.
+One such algorithm, introduced by Pohst~\cite{Pohst_sphere_decoder_1981}, was popularised in the signal processing and communications fields by Viterbo and Boutros~\cite{Viterbo_sphere_decoder_1999} and has since been called the \emph{sphere decoder}.
+
+% Approximate algorithms for computing the nearest point have also been studied.
+A classic example is \emph{Babai's nearest plane algorithm}~\cite{Babai1986}, which requires $O(n^4)$ arithmetic operations in the worst case where $n$ is the dimension of the lattice and only $O(n^2)$ if the lattice basis is \emph{Lov\'asz reduced}~\cite{Lenstra1982}\index{Lov\'asz reduced}.
+Recently, various approximate techniques for solving the nearest lattice point problem have been motivated by applications to multiple-input multiple-output communications.
+An example is the \emph{$K$-best algorithm}~\cite{Zhan2006_K_best_sphere_decoder} that works similarly to the sequential $M$-algorithm~\cite{Anderson1984_seq_coding_alg} used in coding theory.
+Yet another example is the \emph{fixed sphere decoder}~\cite{Jalden2009_error_prob_fixed_sphere_decoder,Barbero2008_fixed_sphere_decoder}.
+
+% In~\cite{McKilliam2009asilomar_polyest_lattice}, it was shown how the least squares phase unwrapping estimator of a polynomial phase signal can be computed by finding a nearest point in a lattice.
+We will briefly describe this result in Section~\ref{sec:least-squar-unwr}.
 
@@
 That is, two or more distinct polynomial-phase signals can take exactly the same values at the sample points.
-These aliasing results are also given in~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}, but the presentation here is different, and is better suited to studying the asymptotic properties of the LSU estimator.
+Understanding how aliasing occurs is crucial to understanding the behaviour of polynomial phase estimators.
+The aliasing properties are described in~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}, but, here we present the properties in a way that is better suited to studying the LSU estimator.
 %A description of the aliasing properties of polynomial phase signals of order $2$ has been independently discovered in~\cite{Abatzoglou_ml_chirp_1986}~and~\cite{Angeby_PPS_aliasing_2000}.
@@
 That is, $\mathcal{Z}$ contains all polynomials $p$ such that $p(n)$ is an integer whenever $n$ is an integer.
-Let $y$ and $z$ be two \emph{distinct} polynomials such that $z = y + p$ for some polynomial $p$ in $\mathcal{Z}$. The two polynomial phase signals
-\[
-s(t) = e^{2\pi j y(t)} \qquad \text{and} \qquad r(t) = e^{2\pi j z(t)}
-\]
-are distinct because $y$ and $z$ are distinct, but if we sample $s$ and $r$ at the integers  
-\begin{align*}
-s(n) &= e^{2\pi j y(n)} =  e^{2\pi j y(n)} e^{2\pi j p(n)} \\
-&= e^{2\pi j (y(n) + p(n))} = e^{2\pi j z(n)} = r(n)
+Let $y$ and $z$ be two distinct polynomials such that $z = y + p$ for some polynomial $p$ in $\mathcal{Z}$.
+The two polynomial phase signals $s(t) = e^{2\pi j y(t)}$ and $r(t) = e^{2\pi j z(t)}$ are distinct because $y$ and $z$ are distinct, but if we sample $s$ and $r$ at the integers
+\begin{align*}
+s(n) = e^{2\pi j y(n)} =  e^{2\pi j y(n)} e^{2\pi j p(n)} = e^{2\pi j (y(n) + p(n))} = e^{2\pi j z(n)} = r(n)
 \end{align*}
@@
 
-It may be helpful to observe Figures~\ref{fig:circstatplot_zero} to~\ref{fig:circstatplot_cube}.
+It may be helpful to observe Figures~\ref{fig:circstatplot_line},~\ref{fig:circstatplot_quad}~and~\ref{fig:circstatplot_cube}.
 In these, the phase (divided by $2\pi$) of two distinct polynomial phase signals is plotted on the left, and on the right the principal component of the phase (also divided by $2\pi$) is plotted.
@@
 \begin{equation} \label{eq:lem_polynomial}
-c_0 p_0 + c_1 p_1 + \dots + c_m p_m,
-\end{equation}
-where the $c_i \in \ints$.
-\end{lemma}
-\begin{IEEEproof}
+c_0 p_0 + c_1 p_1 + \dots + c_m p_m, \qquad c_0,c_1,\dots,c_m \in \ints.
+\end{equation}
+\end{lemma}
+\begin{proof}
 See~\citep[p. 2]{cahen_integer-valued_1997} or~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}. 
-\end{IEEEproof}
-% \begin{IEEEproof}
+\end{proof}
+% \begin{proof}
 % Note that $x(x-1)(x-2)\dots(x-k+1)$ is divisible by all integers $1,2,\dots,k$ and so $p_k$ takes integer values for all $x\in\ints$.
@@
 \[
-\coef(g) = \left[ \begin{array}{ccccc} a_0 & a_1 & a_2 & \dots & a_m \end{array} \right]^\prime,
-\]
-where superscript $^\prime$ indicates the transpose, denote the column vector of length $m+1$ containing the coefficients of $g$.
-If $y$ and $z$ differ by a polynomial from $\mathcal{Z}$ then we can write $y = z + p$ where $p \in \mathcal{Z}$ and then also $\coef(y) = \coef(z) + \coef(p)$.
+\coef(g) = \left[ \begin{array}{ccccc} a_0 & a_1 & a_2 & \dots & a_m \end{array} \right]^\prime
+\]
+denote the column vector of length $m+1$ containing the coefficients of $g$.
+We use superscript $^\prime$ to indicate the vector or matrix transpose.
+If $y$ and $z$ differ by a polynomial from $\mathcal{Z}$ then $y = z + p$ where $p \in \mathcal{Z}$ and also $\coef(y) = \coef(z) + \coef(p)$.
 %\footnote{In group theory terminology $\coef(\cdot)$ coupled with vector addition is called a \term{group homomorphism}.\index{group homomorphism}}. 
@@
 
+%\begin{figure}[tp]
+%	\centering
+%		\includegraphics[width=\linewidth]{plots/circstatfigzero-1.mps}
+%		\caption{The zeroth order polynomials $\tfrac{7}{10}$ (solid line) and $\tfrac{17}{10}$ (dashed line).}
+%		\label{fig:circstatplot_zero}
+%\end{figure}
+
 \begin{figure}[p]
 	\centering
-		\includegraphics[width=\linewidth]{plots/circstatfigzero-1.mps}
-		\caption{The zeroth order polynomials $\tfrac{7}{10}$ (solid line) and $\tfrac{17}{10}$ (dashed line).}		
-\label{fig:circstatplot_zero}
-\end{figure}
-
-\begin{figure}[p]
-	\centering
-		\includegraphics[width=\linewidth]{plots/circstatfig-1.mps}
+		\includegraphics[width=0.85\linewidth]{plots/circstatfig-1.mps}
 		\caption{The first order polynomials $\tfrac{1}{10}(3 + 8t)$ (solid) and $\tfrac{1}{10}(33 - 2t)$ (dashed line).}
@@
 	\centering
-		\includegraphics[width=\linewidth]{plots/circstatfigquad-1.mps}
+		\includegraphics[width=0.85\linewidth]{plots/circstatfigquad-1.mps}
 		\caption{The quadratic polynomials $\tfrac{1}{10} (15 - 15 t + 4 t^2)$ (solid line) and $\tfrac{1}{10}(25 -  t^2)$  (dashed line).}
@@
 	\centering
-		\includegraphics[width=\linewidth]{plots/circstatfigcube-1.mps}
+		\includegraphics[width=0.85\linewidth]{plots/circstatfigcube-1.mps}
 		\caption{The cubic polynomials $\tfrac{1}{160} (174 + 85 t - 118 t^2 + 40 t^3)$ (solid line) and $\tfrac{1}{48} (84 + 19 t + 12 t^2 - 4 t^3)$  (dashed line).}
@@
 \end{equation}
-We will make use of this region when deriving the statistical properties of the LSU estimator in the next section.
-
-Given vectors $\xbf$ and $\ybf$ in $\reals^{m+1}$ we say that $\xbf \equiv \ybf \bmod L_{m+1}$ if $\xbf$ and $\ybf$ differ by a lattice point in $L_{m+1}$.
-We define the function $\dealias(\xbf)$ to take $\xbf$ to its coset representative inside $B$.
+We will make use of this set when deriving the statistical properties of the LSU estimator in the next section.
+
+We define the function $\dealias(\xbf)$ to take $\xbf\in\reals^{m+1}$ to its coset representative inside $B$.
 That is, $\dealias(\xbf) = \zbf \in B$ where $\xbf - \zbf \in L_{m+1}$.
 %The effect of $\dealias(\xbf)$ is to \emph{dealias} the polynomial phase coefficients.
-When $m = 0$ or $1$ then $\dealias(\xbf) = \fracpart{\xbf}$ where $\fracpart{\xbf} = \xbf - \round{\xbf}$ denotes the (centered) fractional part and $\round{\xbf}$ denotes the nearest integer to $\xbf$ with half integers rounded upwards and both $\fracpart{\cdot}$ and $\round{\cdot}$ operate on vectors elementwise.
+When $m = 0$ or $1$ $\dealias(\xbf) = \fracpart{\xbf}$ where $\fracpart{\xbf} = \xbf - \round{\xbf}$ denotes the (centered) fractional part and $\round{\xbf}$ denotes the nearest integer to $\xbf$ with half integers rounded upwards and both $\fracpart{\cdot}$ and $\round{\cdot}$ operate on vectors elementwise.
 For $m \geq 2$ the function $\dealias(\xbf)$ can be computed by a simple sequential algorithm \cite[Sec.
@@
 \] 
-are random variables representing the phase noise induced by the $X_n$~\cite{Tretter1985,Quinn2009_dasp_phase_only_information_loss}.
-If the distribution of $X_n$ is circularly symmetric (i.e., the angle $\angle X_n$ is uniformly distributed on $[-\pi, \pi)$ and is independent of the magnitude $\abs{X_n}$) then the distribution of $\Phi_n$ is the same as the distribution of $\tfrac{1}{2\pi}\angle(1 + \rho^{-1}X_n)$.
-If the $X_1, \dots, X_N$ are circularly symmetric and identically distributed, then $\Phi_1, \dots, \Phi_n$ are also identically distributed.
+is a random variable representing the `phase noise' induced by $X_n$.
+If the distribution of $X_n$ is circularly symmetric (the phase $\angle X_n$ is uniformly distributed on $[-\pi, \pi)$ and is independent of the magnitude $\abs{X_n}$) then the distribution of $\Phi_n$ is the same as the distribution of $\tfrac{1}{2\pi}\angle(1 + \rho^{-1}X_n)$.
+If $X_1, \dots, X_N$ are circularly symmetric and identically distributed, then $\Phi_1, \dots, \Phi_n$ are also identically distributed.
 
@@
 
-The next theorem describes the asymptotic properties of this estimator.
-Before we give the proof it is necessary to understand some of the properties of the phase noise $\Phi_1,\dots,\Phi_N$, which are \emph{circular} random variables with support on $[-\nicefrac{1}{2}, \nicefrac{1}{2})$~\cite{McKilliam2010thesis,McKilliam_mean_dir_est_sq_arc_length2010,Mardia_directional_statistics,Fisher1993}.
+The next theorem describes the asymptotic properties of the LSU estimator.
+Before we state the theorem it is necessary to understand some of the properties of the phase noise $\Phi_1,\dots,\Phi_N$, which are \emph{circular} random variables with support on $[-\nicefrac{1}{2}, \nicefrac{1}{2})$~\cite{McKilliam2010thesis,McKilliam_mean_dir_est_sq_arc_length2010,Mardia_directional_statistics,Fisher1993}.
 Circular random variables are often considered modulo $2\pi$ and therefore have support $[-\pi, \pi)$ with $-\pi$ and $\pi$ being identified as equivalent.
@@
 \[
-\sigma_{\text{intr}}^2 = \expect\fracpart{\Phi_n - \mu_{\text{intr}}}^2 = \min_{\mu \in [-\nicefrac{1}{2}, \nicefrac{1}{2})} \expect \fracpart{\Phi_n - \mu}^2,
+\sigma_{\text{intr}}^2 = \expect\fracpart{\Theta - \mu_{\text{intr}}}^2 = \min_{\mu \in [-\nicefrac{1}{2}, \nicefrac{1}{2})} \expect \fracpart{\Phi_n - \mu}^2,
 \]
@@
 If the minimiser is not unique we say that $\Phi_n$ has no intrinsic mean.
-%We are now equipped to state the asymptotic properties of the LSU estimator.
-We are now equipped to state the main result of this paper.
+Observe the following property of circular random variables with zero intrinsic mean.
+\begin{proposition}\label{prop:zerointmeanzeromean}
+Let $\Phi$ be a circular random variable with intrinsic mean $\mu_{\text{intr}} = 0$ and intrinsic variance $\sigma^2$.
+Then $\Phi$ has zero mean and variance $\sigma^2$, that is, $\expect \Phi = 0$ and $\expect (\Phi - \expect \Phi)^2  = \sigma^2$.
+\end{proposition}
+\begin{proof}
+Assume the proposition is false and that $\mu = \expect \Phi \neq 0$.  But, then 
+\begin{align*}
+\sigma^2 = \expect\fracpart{\Phi -  \mu_{\text{intr}}}^2 = \expect\fracpart{\Phi}^2 = \expect \Phi^2 >  \expect (\Phi - \mu)^2 \geq  \expect \fracpart{\Phi - \mu}^2,
+\end{align*}
+violating the fact that $\mu_{\text{intr}} = 0$ is the minimiser of~\eqref{eq:intrmeandefn}.
+\end{proof}
+We are now equipped to state the asymptotic properties of the LSU estimator.
 
@@
 
-\begin{theorem} \label{thm:asymp_proof}  (Strong consistency) 
+\begin{theorem} \label{thm:asymp_proof} 
 Let $\widehat{\mubf}$ be defined by~\eqref{eq:hatmubfLSUdefn} and put $\widehat{\lambdabf}_N = \dealias(\tilde{\mubf} - \widehat{\mubf})$.
 Denote the elements of $\widehat{\lambdabf}_N$ by $\widehat{\lambda}_{0,N}, \dots, \widehat{\lambda}_{m,N}$.
-Suppose $\Phi_1, \dots, \Phi_N$ are independent and identically distributed with zero intrinsic mean and intrinsic variance $\sigma^2$, then $N^k \widehat{\lambda}_{k,N}$ converges almost surely to $0$ as $N\rightarrow\infty$ for all $k = 0, 1, \dots, m$.
+Suppose $\Phi_1, \dots, \Phi_N$ are independent and identically distributed with zero intrinsic mean, intrinsic variance $\sigma^2$, and probability density function $f$, then:
+\begin{enumerate}
+\item (Strong consistency) $N^k \widehat{\lambda}_{k,N}$ converges almost surely to $0$ as $N\rightarrow\infty$ for all $k = 0, 1, \dots, m$.
+\item (Asymptotic normality) If $f(\fracpart{x})$ is continuous at $x = -\nicefrac{1}{2}$ and if $f(-\nicefrac{1}{2}) < 1$ then the distribution of the vector
+\[
+\left[
+\begin{array}
+[c]{cccc}%
+\sqrt{N} \widehat{\lambda}_{0,N} & N \sqrt{N}\widehat{\lambda}_{1,N}  & \dots & N^m\sqrt{N} \widehat{\lambda}_{m,N}
+\end{array}
+\right]^\prime
+\]
+converges to the normal with zero mean and covariance
+\[
+\frac{\sigma^2}{\left(1-f( -\nicefrac{1}{2}) \right)^{2}} \Cbf^{-1},
+\]
+where $\Cbf$ is the $m+1$ by $m+1$ Hilbert matrix with elements $C_{ik} = 1/(i + k + 1)$ for $i,k \in \{0, 1, \dots, m\}$.
+\end{enumerate}
 \end{theorem}
-
-A proof of this theorem is given in the next section.
-Proofs for the case when $m=0$ was given in~\cite{McKilliam_mean_dir_est_sq_arc_length2010} and for the case when $m=1$ was given in~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.
-The proof here takes a similar approach, but requires new techniques.
-%Before giving the proof we make the following remarks.
-The theorem gives conditions on the \emph{dealiased} difference $\dealias(\tilde{\mubf} - \widehat{\mubf})$ between the true coefficients $\tilde{\mubf}$ and the estimated coefficients $\widehat{\mubf}$ rather than directly on the difference $\tilde{\mubf} - \widehat{\mubf}$.
+The proof of Theorem~\ref{thm:asymp_proof} is contained within the next two sections.
+Section~\ref{sec:strongconstproof} proves strong consistency and Section~\ref{sec:centlimitproof} proves asymptotic normality.
+Proofs for the case when $m=0$ were given in~\cite{McKilliam_mean_dir_est_sq_arc_length2010} and for the case when $m=1$ were given in~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.
+The proofs here take a similar approach, but require new techniques.
+%The proof of asymptotic normality in~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009} contains a subtle error that is amended by the proof given in this paper.
+%Before begining the proof we first discuss the assumptions made in the theorem statement.
+The theorem gives conditions on $\dealias(\tilde{\mubf} - \widehat{\mubf})$ rather than directly on the difference $\tilde{\mubf} - \widehat{\mubf}$.
 To see why this makes sense, consider the case when $m=0$, $\tilde{\mu}_0 = -0.5$ and $\widehat{\mu}_0 = 0.49$, so that $\tilde{\mu}_0 - \widehat{\mu}_0 = -0.99$.
-However, the two phases are obviously close, since the phase $\pm 0.5$ are actually the same.
-In this case
-\[
-\dealias(\tilde{\mu}_0 - \widehat{\mu}_0) = \fracpart{\tilde{\mu}_0 - \widehat{\mu}_0} = 0.01.
-\] 
+However, the two phases are obviously close, since the phases $\pm 0.5$ are actually the same.
+In this case $\dealias(\tilde{\mu}_0 - \widehat{\mu}_0) = \fracpart{\tilde{\mu}_0 - \widehat{\mu}_0} = 0.01$ as expected.
 The same reasoning holds for $m > 0$.
@@
 In the interest of simplicity we only consider the case when $\Phi_1, \dots, \Phi_N$ are identically distributed here.
-If $X_n$ is circularly symmetric with density function nonincreasing with magnitude $\abs{X_n}$, then, the corresponding $\Phi_n$ necessarily has zero intrinsic mean~\cite[Theorem~5.2,~page~78]{McKilliam2010thesis}.
+If $X_n$ is circularly symmetric with density function nonincreasing with magnitude $\abs{X_n}$, then the corresponding $\Phi_n$ necessarily has zero intrinsic mean~\cite[Theorem 5.2]{McKilliam2010thesis}.
 Thus, our theorem covers commonly used distributions for $X_1, \dots, X_N$, such as the normal distribution.
 
-Although we will not prove it here the assumption that $\Phi_1,\dots,\Phi_N$ have zero intrinsic mean is not only sufficient, but also necessary, for if $\Phi_1,\dots,\Phi_N$ have intrinsic mean $x \in [-\nicefrac{1}{2},\nicefrac{1}{2})$ with $x \neq 0$ then $\fracpart{\widehat{\lambda}_{0,N} - x}\rightarrow 0$ almost surely as $N\rightarrow\infty$, and so $\widehat{\lambda}_{0,N}$ does not converge to zero.
-On the other hand if $\Phi_1,\dots,\Phi_N$ do not have an intrinsic mean then $\widehat{\lambda}_{0,N}$ will not converge.
-
-%The proof of asymptotic normality places requirements on the pdf $f$ of the phase noise.
-The requirement that $\Phi_1, \dots, \Phi_N$ have zero intrinsic mean implies that $f(-\nicefrac{1}{2}) \leq 1$~\cite[Lemma~1]{McKilliam_mean_dir_est_sq_arc_length2010}, so the only case not handled is when equality holds, i.e., when $f(-\nicefrac{1}{2}) = 1$ or when $f(\fracpart{x})$ is discontinuous at $x = -\nicefrac{1}{2}$.
-In this exceptional case other expressions for the asymptotic variance can be found (similar to \cite[Theorem 3.1]{Hotz_circle_means_2011}), but this comes at a substantial increase in complexity and for this reason we have omitted them.
+The proof of asymptotic normality places requirements on the probability density function $f$ of the phase noise $\Phi_1, \dots, \Phi_N$.
+The requirement that $\Phi_1, \dots, \Phi_N$ have zero intrinsic mean implies $f(-\nicefrac{1}{2}) \leq 1$~\cite[Lemma~1]{McKilliam_mean_dir_est_sq_arc_length2010}, so the only case not handled is when $f(-\nicefrac{1}{2}) = 1$ or when $f(\fracpart{x})$ is discontinuous at $x = -\nicefrac{1}{2}$.
+In this exceptional case other expressions for the asymptotic variance can be found (similar to \cite[Theorem 3.1]{Hotz_circle_means_2011}), but this comes at a substantial increase in complexity and we have omitted them for this reason.
 %Distributions that do not satisfy these requirements are unlikely to be needed in practice.
-
-%Finally note that $\Cbf$ is a \emph{Hilbert matrix}\index{Hilbert matrix} and the elements of the inverse $\Cbf^{-1}$ are given by
- %\[
- %C^{-1}_{ik} =\frac{i+k+1}{(-1)^{i+k}}{m+i+1 \choose m-k}{m+k+1 \choose m-i}{i+k \choose i}^2.
- %\]
-%Hilbert matrices are particularly ill conditioned and difficult to numerically invert.  This problem can be avoided using the above formula. 
-%I MIGHT HAVE AN ERROR IN THIS FORMULA!
 
@@
  \end{equation} 
-Now $\widehat{\lambdabf}_N = \dealias(\tilde{\mubf} - \widehat{\mubf})$ is the minimiser of $S_{N}$ in $B$.  %We now analyse the minimiser of $S_N$.  
+Now $\widehat{\lambdabf}_N = \dealias(\tilde{\mubf} - \widehat{\mubf})$ is the minimiser of $S_{N}$ in $B$.
+We shall show that $N^k\widehat{\lambda}_{k,N}\rightarrow0$ almost surely as $N\rightarrow\infty$ for all $k = 0,1, \dots, m$ and from this the proof of strong consistency follows.
 Let
@@
  \begin{equation}\label{eq:SNVNunifmlln}
-\sup_{\lambdabf \in B}\sabs{S_N(\lambdabf) - V_N(\lambdabf)} \rightarrow 0  
- \end{equation}
-almost surely as $N\rightarrow\infty$.
+\sup_{\lambdabf \in B}\sabs{S_N(\lambdabf) - V_N(\lambdabf)} \rightarrow 0 \qquad \text{almost surely as $N\rightarrow\infty$}.  
+ \end{equation}
 This type of result has been called a \emph{uniform law of large numbers} and follows from standard techniques~\cite{Pollard_conv_stat_proc_1984}.
@@
 \begin{proof}
-By definition $\widehat{\lambdabf}_N =\arg\min_{\lambdabf\in B} S_N(\lambdabf)$ so 
-\[
-0 \leq S_{N}(\zerobf) - S_N(\widehat{\lambdabf}_N).
-\]  
-Also, because $V_N$ is minimised at $\zerobf$, it follows that 
-%\[
-%0 \leq V_N(\widehat{\lambdabf}_N) - V_N(\zerobf).
-%\]  
-%Combining the two inequalities above,
+By definition $\widehat{\lambdabf}_N =\arg\min_{\lambdabf\in B} S_N(\lambdabf)$ so $0 \leq S_{N}(\zerobf) - S_N(\widehat{\lambdabf}_N)$.
+Also, because $V_N$ is minimised at $\zerobf$, it follows that $0 \leq V_N(\widehat{\lambdabf}_N) - V_N(\zerobf)$.
+Thus,
 \begin{align*}
@@
 However, the next two lemmas and the following corollary will show that if $K$ is sufficiently large then it always contains at least one arithmetic progression (for all sufficiently small $h$) and therefore $A(h,K)$ is not empty.
-We do not wish to claim any novelty here.
-The study of arithmetic progressions within subsets of $W$ has a considerable history~\cite{Erdos_on_some_sequence_of_integers1936,Szemeredi_setint_no_k_arth1975,Gowers_new_proof2001}.
+We do not wish to claim any novelty here, the study of arithmetic progressions within subsets of $W$ has a considerable history~\cite{Erdos_on_some_sequence_of_integers1936,Szemeredi_setint_no_k_arth1975,Gowers_new_proof2001}.
 In particular, Gower's~\cite[Theorem 1.3]{Gowers_new_proof2001} gives a result far stronger than we require here.
@@
 \[
-\Delta_h^1 d(n) = d(n+h) - d(n)
+\Delta_h d(n) = d(n+h) - d(n)
 \] 
@@
 \begin{equation}\label{eq:mthdiffformula}
-\begin{split}
-\Delta_h^r d(n) &= \Delta_h^{r-1} d(n+h) - \Delta_h^{r-1} d(n) \\
-&= \sum_{k=0}^{r}\binom{r}{k}(-1)^{r-k}d(n+kh)
-\end{split}
+\Delta_h^r d(n) = \Delta_h^{r-1} d(n+h) - \Delta_h^{r-1} d(n) = \sum_{k=0}^{r}\binom{r}{k}(-1)^{r-k}d(n+kh)
 \end{equation}
@@
 
-The operator $\Delta_h^r$ has special properties when applied to polynomials. If $d(n) = a_r n^r + \dots + a_0$ is a polynomial of order $r$ then
+The operator $\Delta_h$ has special properties when applied to polynomials. If $d(n) = a_r n^r + \dots + a_0$ is a polynomial of order $r$ then
  \begin{equation} \label{eq:mfinitediffpoly}
@@
 \[
-z_N(n) = \lambda_{0,N} + \lambda_{1,N} n + \dots + \lambda_{m,N} n^m
-\]
-so that 
-\[
-V_N(\lambdabf_N) - \sigma^2 = \frac{1}{N}\sum_{n=1}^{N} g\left( \fracpart{z_N(n)} \right) \rightarrow 0
-\] 
-as $N \rightarrow \infty$.
+z_N(n) = \lambda_{0,N} + \lambda_{1,N} n + \lambda_{2,N} n^2 + \dots + \lambda_{m,N} n^m
+\]
+so that $V_N(\lambdabf_N) - \sigma^2 = \frac{1}{N}\sum_{n=1}^{N} g\left( \fracpart{z_N(n)} \right) \rightarrow 0$ as $N \rightarrow \infty$.
 %At this stage we can \emph{see} the proof.
@@
 \[
-z_{N,r}(n) = \lambda_{0,N} + \lambda_{1,N} n + \dots + \lambda_{r,N} n^r.
-\]
-Because the $m-r$ highest order coefficients converge we can write $z_N(n) = z_{N,r}(n) + \gamma_N(n)$ where 
-\[
-\sup_{n\in\{1,\dots,N\}}\abs{\gamma_N(n)} \rightarrow 0 \qquad \text{as $N\rightarrow\infty$}. 
-\]
+z_{N,r}(n) = \lambda_{0,N} + \lambda_{1,N} n + \lambda_{2,N} n^2 + \dots + \lambda_{r,N} n^r.
+\]
+Because the $m-r$ highest order coefficients converge we can write $z_N(n) = z_{N,r}(n) + \gamma_N(n)$ where $\sup_{n\in\{1,\dots,N\}}\abs{\gamma_N(n)} \rightarrow 0$ as $N\rightarrow\infty$.
 Now the bound from \eqref{eq:Deltazfracbound}, but applied using the $r$th difference, gives
  \begin{equation}\label{eq:zrbound}
-\begin{split}
-\left|\fracpart{  \Delta_h^r z_N(n')}\right| &= \left|\fracpart{  \Delta_h^r\gamma_N(n') + \Delta_h^r z_r(n') }\right| 
-\\ &= |\fracpart{ \epsilon + h^r r! \lambda_{r,N} }| < 2^r\delta,
-\end{split}
+\left|\fracpart{  \Delta_h^r z_N(n')}\right| = \left|\fracpart{  \Delta_h^r\gamma_N(n') + \Delta_h^r z_r(n') }\right| = |\fracpart{ \epsilon + h^r r! \lambda_{r,N} }| < 2^r\delta,
  \end{equation}
@@
 for all $h$ such that $1 \leq h \leq \tfrac{N}{2m}$.
-Choosing $2^r\delta + |\epsilon| < 2^{-2r-1}$ and using the same iterative process as for the highest order coefficient $\lambda_{m,N}$  (see~\eqref{eq:startiterativearg}~to~\eqref{eq:enditerativearg}) we find that $N^r \lambda_{r,N} \rightarrow 0$ as $N\rightarrow\infty$.
+Choosing $2^r\delta + |\epsilon| < \frac{1}{2^{2r+1}}$ and using the same iterative process as for the highest order coefficient $\lambda_{m,N}$  (see~\eqref{eq:startiterativearg}~to~\eqref{eq:enditerativearg}) we find that $N^r \lambda_{r,N} \rightarrow 0$ as $N\rightarrow\infty$.
 The proof now follows by induction.
@@
 We are now in a position to complete the proof of strong consistency.
-As is customary, let $(\Omega, \mathcal F, \prob)$ be the probability space over which the random variables $\{X_i\}$ and $\{\Phi_i\}$ are defined.
-Let $A$ be the subset of the sample space $\Omega$ on which  $V_N(\widehat{\lambdabf}_N) - \sigma^2 \rightarrow 0$ as $N\rightarrow\infty$.
-Now $\prob\{A\} =1$ as a result of Lemma~\ref{lem:ESNconv}.
-Let $A'$ be the subset of $\Omega$ on which $N^k\widehat{\lambda}_{k,N} \rightarrow 0$ for $k=0,\dots,m$ as $N\rightarrow\infty$.
+Let $A$ be the subset of the sample space on which  $V_N(\widehat{\lambdabf}_N) - \sigma^2 \rightarrow 0$ as $N\rightarrow\infty$.
+From Lemma~\ref{lem:ESNconv}, the $\prob\{A\} =1$.
+Let $A'$ be the subset of the sample space on which $N^k\widehat{\lambda}_{k,N} \rightarrow 0$ for $k=0,\dots,m$ as $N\rightarrow\infty$.
 As a result of Lemma~\ref{lem:moran2}, $A \subseteq A'$, and so $\prob\{A'\} \geq \prob\{A\} = 1$.
@@
 
-%\section{Discussion}
-
-\section{Conclusion} \label{sec:conclusion}
-
-This paper has considered the estimation of the coefficients of a noisy polynomial phase signal by least squares phase unwrapping (LSU).
-We have derived conditions under which the polynomial phase estimation problem is identifiable.
-Under these conditions, and under some assumptions on the distribution of the noise, the LSU estimator is shown to be strongly consistent.
-%Simulations are used to show that the LSU has near maximum likelihood performance in the case that the noise is additive white and Gaussian.
-In the second paper in this series~\cite{McKilliam_pps2_2012} we show the LSU estimator to be asymptotically normal and present the results of Monte Carlo simulations that support our asymptotic results.
+\section{Proof of asymptotic normality}\label{sec:centlimitproof}
+
+Let $\psibf$ be the vector with $k$th component $\psi_k = N^k \lambda_k$, $k=0, \dots, m$ and let 
+\[
+T_{N}(\psibf) = S_N(\lambdabf) = \frac{1}{N} \sum_{n=1}^{N} \fracpart{ \Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \psi_k }^2.
+\]
+Let $\widehat{\psibf}_N$ be the vector with elements $\widehat{\psi}_{k,N} = N^k \widehat{\lambda}_{k,N}$ so that $\widehat{\psibf}_N$ is the minimiser of $T_N$.
+Because each of $N^k \widehat{\lambda}_{k,N}$ converges almost surely to zero as $N \rightarrow \infty$, then $\widehat{\psibf}_{N}$ converges almost surely to $\zerobf$ as $N \rightarrow \infty$.
+We want to find the asymptotic distribution of
+\[
+\sqrt{N}\widehat{\psibf}_{N} = 
+\left[
+\begin{array}
+[c]{c}%
+\sqrt{N} \widehat{\psi}_{0,N} \\ \sqrt{N}\widehat{\psi}_{1,N}  \\ \vdots \\ \sqrt{N} \widehat{\psi}_{m,N}
+\end{array}
+\right]
+=
+\left[
+\begin{array}
+[c]{c}%
+\sqrt{N} \widehat{\lambda}_{0,N} \\ N\sqrt{N}\widehat{\lambda}_{1,N} \\ \vdots \\ N^m\sqrt{N} \widehat{\lambda}_{m,N}
+\end{array}
+\right].
+\]
+The proof is complicated by the fact that $T_N$ is not differentiable everywhere because $\fracpart{x}^2$ is not differentiable when $\fracpart{x} = \tfrac{1}{2}$.
+This precludes the use of `standard approaches' to proving asymptotic normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.
+However, we show in Lemma~\ref{lem:diffathatpsi} that all the partial derivatives $\frac{\partial T_N}{\partial \psi_\ell}$ for $\ell = 0, \dots, m$ exist, and are equal to zero, at the minimiser $\widehat{\psibf}_N$.
+Thus, putting
+\begin{equation}\label{eq:Wn}
+W_{n} = \round{\Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \widehat{\psi}_{k,N}},
+\end{equation}
+so that 
+\[
+T_{N}(\psibf) = \frac{1}{N} \sum_{n=1}^{N} \big( \Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \psi_k  - W_n \big)^2,
+\] 
+we have,
+\begin{align*}
+0  = \frac{\partial T_N}{\partial \psi_\ell}(\widehat{\psibf}_N) = \frac{2}{N}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\left( \Phi_n - W_{n}+ \sum_{k=0}^{m}(\tfrac{n}{N})^k \widehat{\psi}_{k,N}  \right),
+\end{align*}
+for each $\ell = 0, \dots, m$.  Now $D_{\ell,N} = K_{\ell,N}$, where $D_{\ell,N} = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}(\tfrac{n}{N})^\ell \Phi_n$, and
+\begin{equation}\label{eq:KellN}
+K_{\ell,N} = \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\left(W_{n}- \sum_{k=0}^{m}(\tfrac{n}{N})^k \widehat{\psi}_{k,N}  \right).
+\end{equation}
+Lemma~\ref{lem:Kconvfhalf} shows that,
+\begin{equation}\label{eq:KellNconv}
+K_{\ell,N} =  (h - 1) \sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big( C_{\ell k} + o_P(1) \big) + o_P(1),
+\end{equation}
+for all $\ell = 0, \dots m$, where $C_{\ell k} =  \tfrac{1}{\ell + k + 1}$, and $h = f(-\nicefrac{1}{2})$, and $o_P(1)$ denotes a random variable converging in probability to zero as $N\rightarrow\infty$.
+
+It is now convenient to write in vector form.  Let 
+\begin{equation}\label{eq:dnandkn}
+\kbf_N = 
+\left[
+\begin{array}{ccc}%
+K_{0,N}  & \cdots & K_{m,N}
+\end{array}
+\right]^\prime
+%\qquad \text{and} \qquad 
+=
+\dbf_N = \left[
+\begin{array}{ccc}%
+D_{0,N}  & \cdots & D_{m,N}
+\end{array}
+\right]^\prime.
+\end{equation}
+From~\eqref{eq:KellNconv},
+\[
+\dbf_N = \kbf_N = \sqrt{N} (h - 1)(\Cbf + o_P(1)) \widehat{\psibf}_N + o_P(1)
+\]
+where $o_P(1)$ here means a vector or matrix of the appropriate dimension with every element converging in probability to zero as $N\rightarrow\infty$.
+Thus $\sqrt{N}\widehat{\psibf}_N$ has the same asymptotic distribution as $(h - 1)^{-1}\Cbf^{-1} \dbf_N$.
+Lemma~\ref{eq:convdn} shows that $\dbf_N$ is asymptotically normally distributed with zero mean and covariance matrix $\sigma^2\Cbf$.
+Thus $\sqrt{N}\widehat{\psibf}_N$ is asymptotically normal with zero mean and covariance matrix
+\[
+\frac{\sigma^2\Cbf^{-1}\Cbf(\Cbf^{-1})^\prime}{(1 - h)^2} = 
+\frac{\sigma^2\Cbf^{-1}}{(1 - h)^2}.
+\] 
+It remains to prove Lemmas~\ref{lem:diffathatpsi},~\ref{lem:Kconvfhalf} and~\ref{eq:convdn}.
+
+\begin{lemma}\label{lem:diffathatpsi}
+For all $\ell = 0, \dots, m$ the partial derivatives $\frac{\partial T_N}{\partial \psi_\ell}$ exist, and are equal to zero, at the minimiser $\widehat{\psibf}_N$.
+That is $\frac{\partial T_N}{\partial \psi_\ell}(\widehat{\psibf}_N) = 0$ for each $\ell = 0, \dots, m$.
+\end{lemma}
+\begin{proof}
+The function $\fracpart{x}^2$ is differentiable everywhere except if $\fracpart{x} \neq -\frac{1}{2}$, and so $T_N$ is differentiable with respect to $\psibf$ at $\widehat{\psibf}_N$ if $\sfracpart{\Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \widehat{\psi}_{k,N}} \neq -\tfrac{1}{2}$ for all $n = 1, \dots, N$.
+This is proved in Lemma~\ref{lem:boundonhatlambda}.
+So the partial derivatives $\frac{\partial T_N}{\partial \psi_\ell}$ exist for all $\ell = 0, \dots, m$ at $\widehat{\psibf}_N$.
+That each of the partial derivatives is equal to zero at $\widehat{\psibf}_N$ follows since $\widehat{\psibf}_N$ is a minimiser of $T_N$.
+\end{proof}
+
+\begin{lemma}\label{lem:boundonhatlambda} $\sabs{\sfracpart{\Phi_n + \sum_{k=0}^m (\nicefrac{n}{N})^k \widehat{\psi}_{k,N}}} \leq \frac{1}{2} - \frac{1}{2N}$ for all $n = 1, \dots, N$.
+\end{lemma}
+\begin{proof}
+To simplify our notation let $B_n = \Phi_n + \sum_{k=1}^m (\nicefrac{n}{N})^k \widehat{\psi}_{k,N}$ so that we now require to prove $\abs{\fracpart{B_n + \widehat{\psi}_{0,N}}} \leq \frac{1}{2} - \frac{1}{2N}$ for all $n = 1, \dots, N$.
+From~\eqref{eq:Wn}, $W_n = \round{B_n + \widehat{\psi}_{0,N}}$, and
+\begin{align*}
+T_N(\widehat{\psibf}_N) = \frac{1}{N}\sum_{n=1}^N \fracpart{B_n + \widehat{\psi}_{0,N}}^2= \frac{1}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N} - W_n)^2.
+\end{align*}
+Since $\widehat{\psi}_{0,N}$ is the minimiser of the quadratic above,
+\begin{equation}\label{eq:lamsum}
+\widehat{\psi}_{0,N} = -\frac{1}{N}\sum_{n=1}^N(B_n - W_n).
+\end{equation}
+The proof now proceeds by contradiction.  Assume that for some $k$,
+\begin{equation}\label{eq:philamfraccontra}
+\fracpart{B_k + \widehat{\psi}_{0,N}} > \frac{1}{2} - \frac{1}{2N}.
+\end{equation}
+Let $F_n = W_n$ for all $n \neq k$ and $F_k = W_k + 1$, and let
+\[
+\phi = -\frac{1}{N}\sum_{n=1}^N(B_n - F_n) = \widehat{\psi}_{0,N} + \frac{1}{N}.
+\]
+Now,
+\begin{align}
+(B_k + \phi - F_k)^2 &= (B_k + \phi - W_k - 1)^2 \nonumber \\
+&= (B_k + \phi - W_k)^2 - 2(B_k + \phi - W_k) + 1 \nonumber \\
+&= (B_k + \phi - W_k)^2 - 2(B_k + \widehat{\psi}_{0,N} - W_k) + 1 - \frac{2}{N} \nonumber \\
+&= (B_k + \phi - W_k)^2 - 2\fracpart{B_k + \widehat{\psi}_{0,N}} + 1 - \frac{2}{N} \nonumber \\
+&< (B_k + \phi - W_k)^2 - \frac{1}{N}, \label{eq:Bkineq}
+\end{align}
+where the inequality in the last line follows from~\eqref{eq:philamfraccontra}. Let $\bbf = [ \phi,  \widehat{\psi}_{1,N} , \dots , \widehat{\psi}_{m,N}]$
+% \[
+% \bbf = \left[
+% \begin{array}
+% [c]{cccc}%
+% \phi & \widehat{\psi}_{1,N} & \cdots & \widehat{\psi}_{m,N}
+% \end{array}
+% \right]^\prime
+% \]
+be the vector of length $m+1$ with components $b_0 = \phi$ and $b_\ell = \widehat{\psi}_{\ell,N}$ for $\ell = 1 , \dots m$.  Now,
+\[
+N T_N(\bbf) = \sum_{n=1}^N \fracpart{B_n + \phi }^2 \leq  \sum_{n=1}^N (B_n + \phi  - F_n)^2, 
+\]
+and using the inequality from~\eqref{eq:Bkineq},
+\begin{align*}
+N T_N(\bbf) &< - \frac{1}{N} + \sum_{n=1}^N (B_n + \phi  - W_n)^2 \\
+&= - \frac{1}{N} + \sum_{n=1}^N (B_n + \widehat{\psi}_{0,N} + \frac{1}{N}  - W_n)^2 \\
+&= \sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n)^2 +  \frac{2}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n)\\
+%&= N T_N(\widehat{\psibf}_N) + \frac{2}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n) \\
+&= N T_N(\widehat{\psibf}_N),
+\end{align*}
+because $\frac{2}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n) = 0$ as a result of~\eqref{eq:lamsum}.
+But, now $T_N(\bbf) < T_N(\widehat{\psibf}_N)$ violating the fact that $\widehat{\psibf}_N$ is a minimiser of $T_N$.
+So~\eqref{eq:philamfraccontra} is false by contradiction.
+
+If $\fracpart{B_k + \widehat{\psi}_{0,N}} < -\frac{1}{2} + \frac{1}{2N}$ for some $k$, we set $F_k = W_k - 1$ and using the same procedure as before obtain $T_N(\bbf) < T_N(\widehat{\psibf}_N)$ again.
+The proof follows.
+\end{proof}
+
+\begin{lemma}\label{lem:Kconvfhalf}
+With $K_{\ell,N}$ defined in~\eqref{eq:KellN}, $h = f(-\nicefrac{1}{2})$, and $C_{\ell k} =  \frac{1}{\ell + k + 1}$ we have $K_{\ell,N} = (h - 1) \sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big( C_{\ell k} + o_P(1) \big) + o_P(1)$ for all $\ell = 0, \dots, m$.% is an element of the $m+1$ by $m+1$ Hilbert matrix, and $o_P(1)$ denotes a random variable converging in probability to zero as $N\rightarrow\infty$.
+\end{lemma}
+\begin{proof}
+Care must be taken since $\widehat{\psibf}_N$ depends on the sequence $\{ \Phi_n \}$.  For $n = 1, \dots, N$ and positive $N$, let
+\begin{equation}
+p_{nN}(\psibf) = \sum_{k=0}^{m}\left( \tfrac{n}{N}\right)^k \psi_k,
+\end{equation}
+and put $q_{n}(x) = \round{\Phi_n + x}$ and $Q(x) = \expect q_{n}(x) =  \expect q_{1}(x)$.  Let
+\begin{equation}\label{eq:GNdef}
+G_{\ell,N}(\psibf) = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}\left( \tfrac{n}{N}\right)^\ell  \big(q_n(p_{nN}(\psibf)) - Q(p_{nN}(\psibf)) \big),
+\end{equation}
+and put
+\begin{equation}\label{eq:hatpnN}
+\widehat{p}_{nN} = p_{nN}(\widehat{\psibf}_N) = \sum_{k=0}^m \left(\tfrac{n}{N}\right)^k \widehat{\psi}_{k,N}.
+\end{equation}
+Now $W_n$ from~\eqref{eq:Wn} can be written as $W_n = \round{\Phi_n + \widehat{p}_{nN}} = q_{n}(\widehat{p}_{nN})$ and $K_{\ell,N}$ from~\eqref{eq:KellN} can be written as
+\begin{align*}
+K_{\ell,N} &= \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\big( q_{n}(\widehat{p}_{nN}) - \widehat{p}_{nN}  \big) \\
+&= \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\big(q_{n}(\widehat{p}_{nN}) - \widehat{p}_{nN} + Q(\widehat{p}_{nN}) - Q(\widehat{p}_{nN}) \big) \\
+&= G_{\ell,N}(\widehat{\psibf}_N) + H_{\ell,N},
+\end{align*}
+where 
+\begin{equation}\label{eq:HellNdef}
+H_{\ell,N} = \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell \big( Q(\widehat{p}_{nN}) - \widehat{p}_{nN} \big).
+\end{equation}
+Lemma~\ref{lem:unifprobG} in the Appendix shows that for any $\delta >0$ and $\nu > 0$ there exists an $\epsilon > 0$ such that
+\[
+\prob\left\{ \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_{\ell,N}(\psibf) } > \delta   \right\} < \nu
+\]
+for all positive integers $N$ and all $\ell = 0,\dots,m$, where $\|\psibf\|_\infty = \sup_{k} \abs{\psi_k}$.  Since $\widehat{\psibf}_N$ converges almost surely to zero, it follows that 
+\[
+\lim_{N\rightarrow\infty}\prob\left\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} = 0
+\]
+for any $\epsilon > 0$, and therefore $\prob\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \} < \nu$ for all sufficiently large $N$.  Now
+\begin{align*}
+  \prob\left\{\abs{ G_{\ell,N}(\widehat{\psibf}_N) } > \delta \right\} &= \prob\left\{ \abs{G_{\ell,N}(\widehat{\psibf}_N)} > \delta \;, \; \|\widehat{\psibf}_N\|_{\infty} < \epsilon \right\} \\
+&\hspace{1cm} + \prob\left\{ \abs{G_{\ell,N}(\widehat{\psibf}_N)} > \delta  \;, \; \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} \\
+&\leq \prob\left\{  \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_{\ell,N}(\psibf) } > \delta \right\} + \prob\left\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} \\
+&\leq 2\nu
+\end{align*}
+for all sufficiently large $N$.
+Since $\nu$ and $\delta$ can be chosen arbitrarily small, it follows that $G_{\ell,N}(\widehat{\psibf}_N)$ converges in probability to zero as $N\rightarrow\infty$, and therefore $K_{\ell,N} = H_{\ell,N} + o_P(1)$.
+Lemma~\ref{lem:sumpNhilb} shows that
+\begin{align*}
+H_{\ell,N} =  (h-1)\sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big(C_{\ell k} + o_P(1)\big).
+\end{align*}
+%The proof follows.
+\end{proof}
+
+\begin{lemma}\label{lem:sumpNhilb}
+With $H_{\ell,N}$ defined in~\eqref{eq:HellNdef}, $h = f(-\nicefrac{1}{2})$, and  $C_{\ell k} = \tfrac{1}{\ell + k + 1}$ we have $H_{\ell,N} = (h-1)\sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big(C_{\ell k} + o_P(1)\big)$.
+\end{lemma}
+\begin{proof}
+If $\abs{x} < 1$, then
+\[
+q_n(x)  = \round{\Phi_n + x} = \begin{cases}
+1, & \Phi_n + x  \geq \nicefrac{1}{2} \\
+-1, & \Phi_n + x  < -\nicefrac{1}{2} \\
+0, & \text{otherwise},
+\end{cases}
+\]
+and,
+\begin{align*}
+Q(x) = Eq_1(x) &=  \begin{cases}
+\int_{\nicefrac{1}{2} -x }^{\nicefrac{1}{2}}f(t)\,dt, &   x \geq 0 \\
+-\int_{-\nicefrac{1}{2}}^{-\nicefrac{1}{2} - x }f(t)\,dt, &  x  < 0.
+\end{cases} 
+% \\
+% &= \begin{cases}
+% 1 - F(\nicefrac{1}{2} - x ), & x  \geq 0 \\
+% -F(-\nicefrac{1}{2} - x ), &  x < 0,
+% \end{cases} 
+\end{align*}
+Because $f(\fracpart{x})$ is continuous at $-\nicefrac{1}{2}$ it follows that $Q(x) = x \big( h + \zeta(x) \big)$
+where $\zeta(x)$ is a function that converges to zero as $x$ converges to zero.  
+%The equation above is still valid when $\abs{x} \geq 1$ by an appropriate choice of the function $\zeta(x)$.  
+Observe that $\abs{\widehat{p}_{nN}} \leq \sum_{k=0}^m\sabs{ \widehat{\psi}_{k,N}}$ and, since each of the $\widehat{\psi}_{k,N} \rightarrow 0$ almost surely as $N\rightarrow\infty$, it follows that $\widehat{p}_{nN} \rightarrow 0$ almost surely uniformly in $n = 1, \dots, N$ as $N\rightarrow\infty$.
+Thus, $\zeta(\widehat{p}_{nN}) \rightarrow 0$ almost surely (and therefore also in probability) uniformly in $n = 1, \dots, N$ as $N\rightarrow\infty$.
+Now,
+\begin{align*}
+Q(\widehat{p}_{nN}) - \widehat{p}_{nN} = \widehat{p}_{nN}\big( h - 1 + \zeta(\widehat{p}_{nN}) \big) = \widehat{p}_{nN}\big( h - 1 + o_P(1) \big), 
+\end{align*}
+and, using~\eqref{eq:HellNdef},
+\begin{align*}
+H_{\ell,N} &= \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell \widehat{p}_{nN}\big( h - 1 + o_P(1) \big)  \\
+&=  \frac{1}{\sqrt{N}} \sum_{n=1}^{N}(\tfrac{n}{N})^\ell \sum_{k=0}^m (\tfrac{n}{N})^k \widehat{\psi}_{k,N} \big( h - 1 + o_P(1) \big)   \\
+&= \sqrt{N}\sum_{k=0}^m \widehat{\psi}_{k,N} \frac{1}{N} \sum_{n=1}^{N} \frac{n^{\ell + k}}{N^{\ell + k + 1}}\big( h - 1 + o_P(1)  \big).
+\end{align*}
+The Riemann sum
+\[
+\frac{1}{N} \sum_{n=1}^{N} \frac{n^{\ell + k}}{N^{\ell + k + 1}} = \int_{0}^{1}x^{k+\ell+1}dx + o_P(1),
+\]
+and since the integral above evaluates to $C_{\ell k} = \frac{1}{k+\ell+1}$, we have
+\[
+H_{\ell,N} = (h - 1)\sqrt{N}\sum_{k=0}^m \widehat{\psi}_{k,N}\big( C_{\ell k} + o_P(1)  \big).C_{\ell k}
+\]
+% Put
+% \[
+% C_{\ell + k,N} = \sum_{n=1}^{N}\frac{n^{\ell + k}}{N^{\ell + k+1}}.
+% \]
+% It is well known (see for example~\cite[eq. 21]{Peleg1991_CRB_PPS_1991}) that,
+% \begin{equation}\label{eq:ClkNHilbert}
+% C_{\ell + k,N} = C_{\ell k} + O(N^{-1}).
+% \end{equation}
+% Since $\zeta(\widehat{p}_{nN}) = o_P(1)$ for all $n = 1, \dots, N$ then
+% \[
+% \sum_{n=1}^{N}\frac{n^{\ell + k}}{N^{\ell + k + 1}}\zeta(\widehat{p}_{nN}) 
+% \]
+% converges in probability to zero as $N\rightarrow\infty$.  So,
+% \begin{align*}
+% \sum_{n=1}^{N}\frac{n^{\ell + k}}{N^{\ell + k + 1}}\big( h - 1 +  \zeta(\widehat{p}_{nN}) \big) = (h-1)\big( C_{\ell k} + o_P(1) \big).
+% \end{align*}
+% The proof follows by substituting this into~\eqref{eq:HellNre}.
+\end{proof}
+
+\begin{lemma}\label{eq:convdn}
+The distribution of the vector $\dbf_N$, defined in~\eqref{eq:dnandkn}, converges to the multivariate normal with zero mean and covariance matrix $\sigma^2\Cbf$.
+\end{lemma}
+\begin{proof}
+For any constant vector $\alphabf$, let
+\[
+z_N = \alphabf^\prime \dbf_N = \frac{1}{\sqrt{N}} \sum_{n=1}^{N} \Phi_n \sum_{\ell=0}^{m}\alpha_\ell\left( \frac{n}{N} \right)^\ell.
+\]
+By Lypanov's central limit theorem $z_N$ is asymptotically normally distributed with zero mean and variance
+\[
+\lim_{N\rightarrow\infty} \sigma^2 \frac{1}{N} \sum_{n=1}^{N}\left( \sum_{\ell=0}^{m}\alpha_\ell\left( \frac{n}{N} \right)^\ell \right)^2 = \sigma^2 \alphabf^\prime \Cbf \alphabf.
+\]
+By the Cram\`{e}r-Wold theorem it follows that $\dbf_N$ is asymptotically normally distributed with zero mean and covariance $\sigma^2 \Cbf$. 
+% Recall that the $\ell$th element of $\dbf_N$ is
+% \[
+% D_{\ell,N} = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}(\tfrac{n}{N})^\ell \Phi_n.
+% \]
+% Since the $\Phi_1, \dots, \Phi_N$ are independent, each with variance $\sigma^2$, the distribution of $D_{\ell,N}$ converges to the normal with zero mean and variance
+% \[
+% \lim_{N\rightarrow\infty}\var D_{\ell,N} = \sigma^2 \lim_{N\rightarrow\infty} C_{2\ell,N} = \sigma^2 C_{\ell\ell}
+% \]
+% by Lypanov's central limit theorem and by using~\eqref{eq:ClkNHilbert}.  The covariance between $D_{\ell,N}$ and $D_{k,N}$ is
+% \[
+% \covar(D_{\ell,N}, D_{k,N}) = \sigma^2 C_{\ell+k,N} \rightarrow \sigma^2 C_{\ell k}
+% \]
+% as $N \rightarrow \infty$, again using~\eqref{eq:ClkNHilbert}.  The proof follows.
+\end{proof}
+
+\section{Simulations}\label{sec:simulations}
+
+This section describes the results of Monte-Carlo simulations with the least squares unwrapping (LSU) estimator.
+The sample sizes considered are $N = 10, 50, 200$ and the unknown amplitude is $\rho = 1$.
+The $X_1, \dots, X_N$ are pseudorandomly generated independent and identically distributed circularly symmetric complex Gaussian random variables with variance $\sigma_c^2$.
+The coefficients $\tilde{\mubf} = [\tilde{\mu}_0, \dots, \tilde{\mu}_m]$ are distributed uniformly randomly in the identifiable region $B$.
+The number of replications of each experiment is $T = 2000$ to obtain estimates $\widehat{\mubf}_1, \dots, \widehat{\mubf}_T$ and the corresponding dealiased errors $\widehat{\lambdabf}_t = \dealias(\widehat{\mubf}_T - \tilde{\mubf})$ are computed.
+The sample mean square error (MSE) of the $k$th coefficient is computed according to $\tfrac{1}{T}\sum_{t=1}^T \widehat{\lambda}_{k,t}^2$ where $\widehat{\lambda}_{k,t}$ is the $k$th element of $\widehat{\lambdabf}_t$.
+
+Figure~\ref{plot:polyest1} shows the sample MSEs obtained for a polynomial phase signal of order $m=3$.
+Results are displayed results for the zeroth and third order coefficients $\widehat{\mu}_1$ and $\widehat{\mu}_3$.
+The results for $\widehat{\mu}_1$ and $\widehat{\mu}_2$ lead to similar conclusions.
+When $N = 10$ and $50$ the LSU estimator can be computed exactly using a general purpose algorithm for finding nearest lattice points called the sphere~decoder~\cite{Pohst_sphere_decoder_1981,Agrell2002,Viterbo_sphere_decoder_1999}.
+This is displayed by the circles in the figures.
+When $N=200$ the sphere decoder is computationally intractable and we instead use an approximate nearest point algorithm called the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder}.
+This is displayed by the dots.
+For the purpose of comparison we have also plotted the results for the $K$-best method when $N = 10$ and $50$.
+The asymptotic variance predicted in Theorem~\ref{thm:asymp_proof} is displayed by the dashed line.
+Provided the noise variance is small enough (so that the `threshold' is avoided) the sample MSE of the LSU estimator is close to that predicted by Theorem~\ref{thm:asymp_proof}.
+The Cram\`{e}r-Rao lower bound for the variance of unbiased polynomial phase estimators in Gaussian noise is also plotted using the solid line~\cite{Peleg1991_CRB_PPS_1991}.
+When the noise variance is small the asymptotic variance of the LSU estimator is close to the Cram\`{e}r-Rao lower bound.
+
+%Note the gap between the CRB and the variance predicted by Theorem~\ref{thm:asymp_proof}.
+In practice this gap can be overcome using a numerical technique, such as Newton's method, with the LSU estimate as a starting point.
+We have not done this here in order to show the accuracy of Theorem~\ref{thm:asymp_proof}.
+The other noticeable gap is that between the threshold behaviour of the LSU estimator and the LSE.
+When $N=50$ the threshold \emph{kicks in} a little earlier for the LSE than it does for the LSU.
+We suspect that this gap will actually \emph{close} as $N$ increases.
+This hypothesis is supported by the fact that this behaviour is observed in the case of frequency estimation (when $m=1$)~\cite[Sec.
+9.5]{McKilliam2010thesis}\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.
+Unfortunately it is not feasible to test this hypothesis because of the computational complexity of the LSE.
+
+% \begin{figure}[p] 
+ %   	\centering 
+ %  		\includegraphics[width=\linewidth]{code/lseplot-1.mps} 
+ %               %\includegraphics{code/lseplot-1.mps}
+ %   		\caption{Performance of the LSE and LSU estimator for a polynomial phase signal of order two, i.e.
+$m=2$.
+(Top) Mean square error (MSE) in the second order parameter $\mu_2$ for $N=10,50$ and $200$ versus the variance $\sigma_c^2$ of the $X_1, \dots, X_N$.
+The $X_1, \dots, X_N$ are independent and identically distributed (i.i.d.) complex Gaussian random variables.
+(Bottom) MSE in second order parameter $\mu_2$ versus the variance $\sigma^2$ of the $\Phi_1, \dots, \Phi_N$.
+The $\Phi_1, \dots, \Phi_N$ are i.i.d.
+uniform random variables.}
+ %   		\label{plot:lsetest} 
+ %  \end{figure}
+
+\begin{figure}[p] 
+   	\centering 
+  		\includegraphics[width=\linewidth]{code/gaussianplot4-3.mps}
+                %\includegraphics{code/gaussianplot4-1.mps}
+  		\caption{Sample mean square error (MSE) of the least squares unwrapping estimator for $N=10,50$ and $200$ for a polynomial phase signal of order $m=3$.
+(Top) MSE of the frequency coefficient $\mu_0$.
+(Bottom) MSE of the cubic coefficient $\mu_3$.}
+  		\label{plot:polyest1} 
+ \end{figure}
 
@@
 %   	\centering 
-%  		\includegraphics[width=\linewidth]{polyfig/polyphase4plot.4} 
-%  		\caption{MSE in the third order coefficient $\mu_3$ versus $\var{X_n} = \sigma^2$. The true coefficients are uniformly spread in the identifiable region.} 
-%  		\label{plot:polyest_projnorm_mu3} 
+%  		\includegraphics[width=\linewidth]{code/gaussianplot4-2.mps}
+%                %\includegraphics{code/gaussianplot4-1.mps}
+%  		\caption{Sample mean square error (MSE) of the least squares unwrapping estimator for $N=10,50$ and $200$ for a polynomial phase signal of order $m=3$.
+(Top) MSE of the quadratic coefficient $\mu_2$.
+(Bottom) MSE of the cubic coefficient $\mu_3$.}
+%  		\label{plot:polyest2} 
 % \end{figure}
 
-% \begin{figure}[p] 
-%  	\centering 
-% 		\includegraphics[width=\linewidth]{polyfig/polyphase4plot_unfair.4} 
-% 		\caption{MSE in the third order coefficient $\mu_3$ versus $\var{X_n} = \sigma_c^2$. The coefficient have been restricted for Kitchen's estimator and the DPT.} 
-% 		\label{plot:polyest_projnorm_mu3_unfair} 
-%\end{figure} 
-% 
-% \begin{figure}[p] 
-%  	\centering 
-% 		\includegraphics[width=\linewidth]{polyfig/polyphase4plot_fair.4} 
-% 		\caption{MSE in the third order coefficient $\mu_3$ versus $\var{X_n} = \sigma_c^2$. The DPT estimator runs at the higher sampling rate $\delta$ so that the volumes $V_{DPT}(\delta) = V_m$.} 
-% 		\label{plot:polyest_projnorm_mu3_fair} 
-%\end{figure}
-
-%\addtolength{\parskip}{-0.2cm} 
-%\bibliographystyle{../../bib/IEEEbib} 
-\small 
-\bibliography{bib}
-
-\normalsize
+% \section{Computational considerations}
+
+% When $m=0$, polynomial phase estimation is equivalent to the problem of \emph{direction estimation} from circular statistics, and the least squares estimator corresponds with the \emph{extrinsic mean} (or \emph{sample circular mean})~\cite{McKilliam_mean_dir_est_sq_arc_length2010,Fisher1993,Mardia_directional_statistics}.
+In this case the LSE can be computed efficiently in $O(N)$ by averaging $N$ complex numbers.
+When $m=1$, polynomial phase estimation is equivalent to single frequency estimation, and the least squares estimator can be computed efficiently in $O(N\log N)$ operations by maximising the \emph{periodogram} using the fast Fourier transform followed by an appropriate numerical optimisation technique~\cite{Quinn2008maximizing_the_periodogram,Rife1974,Quinn2001}.
+When $m=2$, the least squares estimator can be computed in a reasonable amount of time when $N$ is approximately less than $50$.
+The LSE is computationally intensive for any $N$ when $m > 2$.
+By comparison the exact LSU estimator implemented using the sphere decoder can be computed quickly for any $m$ with $N$ approximately less than $50$.
+In situations where $N$ is small, but high statistical accuracy is required, the sphere decoder is computationally a better choice than the LSE for polynomial phase signals of order greater than or equal to $2$.
+
+% Babai's nearest plane algorithm requires $O(N^2)$ operations and can be run in a reasonable amount of time for any $m$ and reasonably large $N$.
+It should be noted that Babai's nearest plane algorithm requires a Lov\`as reduced lattice basis~\cite{Lenstra1982}.
+This requires $O(N^4)$ operations to compute, but, the Lov\`as reduced basis can be computed once offline.
+It does not need to be computed each time the estimator is run.
+The $K$-best algorithm runs in a reasonable amount of time for any $m$ and when $N$ is less than about $300$.
+If the noise variance is small, then both the sphere decoder and the $K$-best algorithm are actually fast.
+This is because both of these estimators begin with the lattice point found by Babai's nearest plane algorithm.
+If this point is close to (or \emph{is}) the nearest point then both the sphere decoder and the $K$-best algorithm terminate quickly.
+This is in contrast to the least squares estimator that is slow regardless of the noise variance.
+
+% Both Kitchen's estimator and the DPT are fast to compute.
+Kitchen's estimator requires only $O(N)$ arithmetic operations and the DPT requires $O(N \log N)$ operations.
+However, these estimators are not as accurate as the LSE or the LSU.
+They also fail to operate correctly over the entire identifiable region of polynomial phase coefficients~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}~\cite[Ch.~10]{McKilliam2010thesis}.
+
+\section{Conclusion} \label{sec:conclusion}
+
+This paper has considered the estimation of the coefficients of a noisy polynomial phase signal by least squares phase unwrapping (LSU).
+It has been shown that the LSU estimator is strongly consistent and asymptotically normally distributed.
+%Simulations are used to show that the LSU has near maximum likelihood performance in the case that the noise is additive white and Gaussian.
+%It is shown in~\cite[Sec~8.1]{McKilliam2010thesis}\cite{McKilliam2009asilomar_polyest_lattice} how the LSU estimator can be computed by finding a nearest lattice point in a particular lattice.  
+Polynomial time algorithms that compute the LSU estimator are described in~\cite{McKilliam2010thesis}, but these are slow algorithms in practice.
+A significant outstanding question is whether practically fast algorithms exist.
+Considering the excellent statistical performance (both theoretically and practically) of the LSU estimator, even fast \emph{approximate} algorithms are likely to prove useful for the estimation of polynomial phase signals.
+
 \appendix
@@
 
-\subsection{A uniform law of large numbers} \label{app:uniform-law-large}
+\section{A uniform law of large numbers} \label{app:uniform-law-large}
 
@@
 \end{equation}
-almost surely as $N\rightarrow\infty$, where $V_N(\lambdabf) = \expect S_N(\lambdabf)$.  We prove this result here.  Put
-\[
-D_N(\lambdabf) = S_N(\lambdabf) - V_N(\lambdabf).
-\]  
-Now, for any $\epsilon > 0$,
-\[
-\sum_{N=1}^\infty \prob \left\{ \sup_{\lambdabf \in B}\abs{ D_N(\lambdabf) } > \epsilon \right\} < \infty
-\]
-by Lemma~\ref{lem:vn}.
-So~\eqref{eq:SNVNunifmlln2} follows from the Borel-Cantelli lemma.
+almost surely as $N\rightarrow\infty$, where $V_N(\lambdabf) = \expect S_N(\lambdabf)$.
+We prove this result here.
+Put $D_N(\lambdabf) = S_N(\lambdabf) - V_N(\lambdabf)$.
+Now $\sum_{N=1}^\infty \prob \left\{ \sup_{\lambdabf \in B}\abs{ D_N(\lambdabf) } > \epsilon \right\} < \infty$ for any $\epsilon > 0$ by Lemma~\ref{lem:vn}, and~\eqref{eq:SNVNunifmlln2} follows from the Borel-Cantelli lemma.
 In what follows we use order notation in the standard way, that is, for functions $h$ and $g$, we write $h(N) = O(g(N))$ to mean that there exists a constant $K > 0$ and a finite $N_0$ such that $h(N) \leq K g(N)$ for all $N > N_0$.
 
-\begin{lemma} \label{lem:vn} 
-For any $\epsilon > 0$ and $c < 2$,
-\[
-\prob \left\{ \sup_{\lambdabf \in B}\abs{ D_N(\lambdabf) } > \epsilon \right\} = O(e^{-c \epsilon^2 N}).
-\]
+\begin{lemma} \label{lem:vn} $\prob \left\{ \sup_{\lambdabf \in B}\abs{ D_N(\lambdabf) } > \epsilon \right\} = O(e^{-c \epsilon^2 N})$ for any $\epsilon > 0$ and $c < 2$.
  \end{lemma}
@@
 \begin{align}
- &\sup_{\lambdabf \in B}\sabs{ D_N(\lambdabf) }  \nonumber \\
-&= \sup_{\rbf \in G}\sup_{ \lambdabf \in B[\rbf]}\sabs{ D_N(\lambdabf[\rbf]) + D_N(\lambdabf) - D_N(\lambdabf[\rbf])  } \nonumber \\
+ \sup_{\lambdabf \in B}\sabs{ D_N(\lambdabf) }  &= \sup_{\rbf \in G}\sup_{ \lambdabf \in B[\rbf]}\sabs{ D_N(\lambdabf[\rbf]) + D_N(\lambdabf) - D_N(\lambdabf[\rbf])  } \nonumber \\
  &\leq \sup_{\rbf \in G}\sabs{ D_N(\lambdabf[\rbf])} + \sup_{\rbf \in G}\sup_{\lambdabf\in B[\rbf]}\sabs{ D_N(\lambdabf) - D_N(\lambdabf[\rbf])  }. \label{eq:gridanddensespace}
  \end{align}
-From Lemma~\ref{lem:supVjk} it will follow that 
-\[
-\prob\left\{   \sup_{\rbf \in G }\abs{ D_N(\lambdabf[\rbf])  } > \frac{\epsilon}{2} \right\} = O(e^{-c \epsilon^2 N})
-\]
+From Lemma~\ref{lem:supVjk} it will follow that $\prob\left\{   \sup_{\rbf \in G }\abs{ D_N(\lambdabf[\rbf])  } > \frac{\epsilon}{2} \right\} = O(e^{-c \epsilon^2 N})$
 for any $\epsilon > 0$ and $c < 2$.  In Lemma~\ref{lem:supBVn} we show that
@@
 \begin{lemma}\label{lem:supVjk}
-For any $\epsilon > 0$ and $c < 8$,
-\[
-\prob\cubr{   \sup_{\rbf \in G}\abs{ D_N(\lambdabf[\rbf])  } > \epsilon } = O(e^{-c \epsilon^2 N}).
-\]
-\end{lemma}
- \begin{proof}
-Fix $\lambdabf$ and write 
-\[
-D_N(\lambdabf) = \bar{Z} = \frac{1}{N}  \sum_{n=1}^{N}Z_{n},
-\] 
-where
+$\prob\cubr{   \sup_{\rbf \in G}\abs{ D_N(\lambdabf[\rbf])  } > \epsilon } = O(e^{-c \epsilon^2 N})$ for any $\epsilon > 0$ and $c < 8$.
+\end{lemma}
+ \begin{proof}
+Fix $\lambdabf$ and write $D_N(\lambdabf) = \bar{Z} = \frac{1}{N}  \sum_{n=1}^{N}Z_{n}$, where
  \[
@@
  \]
- are independent with zero mean and $\abs{Z_n} \leq \tfrac{1}{4}$. It follows from Hoeffding's inequality~\cite{Hoeffding_inequality_1963} that,  
-\[
-\prob\{ \abs{D_N(\lambdabf)} > \epsilon \} \leq 2e^{ - 8 \epsilon^2 N},
-\] 
-and so,
+ are independent with zero mean and $\abs{Z_n} \leq \tfrac{1}{4}$.
+It follows from Hoeffding's inequality~\cite{Hoeffding_inequality_1963} that,  $\prob\{ \abs{D_N(\lambdabf)} > \epsilon \} \leq 2e^{ - 8 \epsilon^2 N}$, and so,
  \begin{align*}
@@
 
-\begin{lemma}\label{lem:boundedsquarefracparts} For real numbers $x$ and $delta$,
-\[
-\fracpart{x}^2 - |\delta| \leq \fracpart{x + \delta}^2 \leq \fracpart{x}^2 + |\delta|.
-\]
+\begin{lemma}\label{lem:boundedsquarefracparts}
+$\fracpart{x}^2 - |\delta| \leq \fracpart{x + \delta}^2 \leq \fracpart{x}^2 + |\delta|$ for all $x, \delta \in \reals$.
 \end{lemma}
@@
 
- \begin{lemma}\label{lem:supBVn} 
-For all positive integers $N$,
-\[
-\sup_{\rbf\in G}\sup_{ \lambdabf  \in B[\rbf] }\left\vert D_N\left(  \lambdabf \right) - D_N\left(  \lambdabf[\rbf] \right)\right\vert < 2\frac{m+1}{N^b}.
-\]
+ \begin{lemma}\label{lem:supBVn} $\sup_{\rbf\in G}\sup_{ \lambdabf  \in B[\rbf] }\left\vert D_N\left(  \lambdabf \right) - D_N\left(  \lambdabf[\rbf] \right)\right\vert < 2\frac{m+1}{N^b}$ for all $N$.
  \end{lemma}
@@
 %The proof we give is not just almost surely but \emph{surely} as the convergence will be shown to occur irrespective of the values of the $\Phi_n$. 
-Put 
-\[
-b_n = \Phi_n + \sum_{k = 0}^{m}{\lambda_k n^k} \;\;\; \text{and}  \;\;\; a_n = \Phi_n + \sum_{k = 0}^{m}{\lambda[\rbf]_k n^k},
-\]
-where $\lambda[\rbf]_k$ denotes the $k$th element of the grid point $\lambdabf[\rbf]$. For $\lambdabf \in B[\rbf]$ we have $b_n = a_n + \delta_n$,
-where 
-\[
-\sabs{\delta_n} \leq \sum_{k=0}^m\frac{n^k}{k! N^{b+k}} \leq \frac{m+1}{N^{b}}.
-\]
-From Lemma~\ref{lem:boundedsquarefracparts} it follows that 
-\[
--|\delta_n| \leq \fracpart{x + b_n}^2 -  \fracpart{x + a_n}^2  \leq |\delta_n|,
-\]
-and consequently 
-\[
-|\fracpart{x + b_n}^2 -  \fracpart{x + a_n}^2| \leq \tfrac{m+1}{N^b}
-\] 
-for all $x \in \reals$. Now
+Put $b_n = \Phi_n + \sum_{k = 0}^{m}{\lambda_k n^k}$ and $a_n = \Phi_n + \sum_{k = 0}^{m}{\lambda[\rbf]_k n^k}$, where $\lambda[\rbf]_k$ denotes the $k$th element of the grid point $\lambdabf[\rbf]$.
+For $\lambdabf \in B[\rbf]$ we have $b_n = a_n + \delta_n$,
+where $\sabs{\delta_n} \leq \sum_{k=0}^m\frac{n^k}{k!
+N^{b+k}} \leq \frac{m+1}{N^{b}}$.
+From Lemma~\ref{lem:boundedsquarefracparts} it follows that $-|\delta_n| \leq \fracpart{x + b_n}^2 -  \fracpart{x + a_n}^2  \leq |\delta_n|$,
+and consequently $|\fracpart{x + b_n}^2 -  \fracpart{x + a_n}^2| \leq \tfrac{m+1}{N^b}$ for all $x \in \reals$. Now
 \[
@@
 \]
-and therefore 
-\[
-\sabs{S_N(\lambdabf) - S_N(\lambdabf[\rbf])} \leq \frac{m+1}{N^b}
-\] 
-for all $\lambdabf \in B[\rbf]$.  As this bound is independent of $\Phi_1 \dots \Phi_N$, we have
+and therefore $\sabs{S_N(\lambdabf) - S_N(\lambdabf[\rbf])} \leq \frac{m+1}{N^b}$ for all $\lambdabf \in B[\rbf]$.  As this bound is independent of $\Phi_1 \dots \Phi_N$, we have
 \[
@@
 \]
-by Jensen's inequality.  Therefore,
- \begin{align*}
- &\sabs{ D_N(\lambdabf) - D_N(\lambdabf[\rbf]) } \\
-&= \sabs{ S_N(\lambdabf) - S_N(\lambdabf[\rbf]) + V_N(\lambdabf) - V_N(\lambdabf[\rbf]) } \\
+by Jensen's inequality.  Therefore, for all $\lambdabf \in B[\rbf]$,
+ \begin{align*}
+ \sabs{ D_N(\lambdabf) - D_N(\lambdabf[\rbf]) } &= \sabs{ S_N(\lambdabf) - S_N(\lambdabf[\rbf]) + V_N(\lambdabf) - V_N(\lambdabf[\rbf]) } \\
 &\leq \sabs{ S_N(\lambdabf) - S_N(\lambdabf[\rbf]) \vert + \vert V_N(\lambdabf) - V_N(\lambdabf[\rbf]) } \\
@@
 \end{align*}
-for all $\lambdabf \in B[\rbf]$.  The lemma follows because this bound is independent of $\rbf$.
+and the lemma follows because this bound is independent of $\rbf$.
 %\[
@@
 
+\section{A tightness result}\label{app:tightness-result}
+
+During the proof of asymptotic normality in Lemma~\ref{lem:Kconvfhalf} we made use of the following result regarding the function,
+\[
+G_{\ell,N}(\psibf) = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}\left( \tfrac{n}{N}\right)^\ell  \big(q_n(p_{nN}(\psibf)) - Q(p_{nN}(\psibf)) \big),
+\]
+where the functions $q_n$, $Q$ and $p_{nN}$ are defined above~\eqref{eq:GNdef} and $\ell \in \{0, 1, \dots, m\}$.
+To simplify notation we drop the subscript $\ell$ and write $G_{\ell,N}$ as $G_N$ in what follows.
+The proof we will give holds for any nonnegative integers $\ell$.
+
+\begin{lemma}\label{lem:unifprobG}
+For any $\delta >0$ and $\nu > 0$ there exists $\epsilon > 0$ such that
+\[
+\prob\left\{ \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf) } > \delta   \right\} < \nu %\qquad \text{for all $N > N_0$}
+\]
+for all positive integers $N$.
+\end{lemma}
+
+This result is related to what is called \emph{tightness} or \emph{asymptotic continuity} in the literature on empirical processes and weak convergence on metric spaces~\cite{Billingsley1999_convergence_of_probability_measures,Dudley_unif_central_lim_th_1999,Shorak_emp_proc_stat_2009,van2009empirical}.
+The lemma is different from what is usually proved in the literature because the function $p_{nN}(\psibf) = \sum_{k=0}^m\left( \tfrac{n}{N}\right)^k\psi_k$
+depends on $n$.
+Nevertheless, the methods of proof from the literature can be used if we include a known result about hyperplane arrangements~\cite[Ch.
+5]{Chazelle_discrepency_method_2000}\cite[Ch.
+6]{Matousek_lect_disc_geom_2002}.
+Our proof is based on a technique called \emph{symmetrisation} and another technique called \emph{chaining} (also known as \emph{bracketing})~\cite{Pollard_asymp_empi_proc_1989,van2009empirical}.
+
+\begin{proof}
+Define the function 
+\begin{align*}
+f_{nN}(\psibf, \Phi_n) = \left( \tfrac{n}{N}\right)^\ell q_n(p_{nN}(\psibf)) = \left( \tfrac{n}{N}\right)^\ell \round{\Phi_n + \sum_{k=0}^m\left( \tfrac{n}{N}\right)^k\psi_k}
+\end{align*}
+%and observe that 
+%\[
+%\expect f_{nN}(\psibf, \Phi_n) = \expect f_{nN}(\psibf, \Phi_1) = \left( \tfrac{n}{N}\right)^\ell Q(p_{nN}(\psibf))
+%\]
+so that $G_N$ can be written as
+\[
+G_N(\psibf) = \frac{1}{\sqrt{N}} \sum_{n=1}^{N} \big( f_{nN}(\psibf, \Phi_n) - \expect f_{nN}(\psibf, \Phi_n) \big).
+\]
+Let $\{g_n\}$ be a sequence of independent standard normal random variables, independent of the phase noise sequence $\{\Phi_n\}$.
+The symmetrisation argument~\cite[Sec.~4]{Pollard_asymp_empi_proc_1989}\cite{Gine_Zinn_symmetrisation_1984,van2009empirical} can be used to show that
+%Lemma~\ref{lem:symmetrisation} shows that
+\[
+\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} \leq \sqrt{2\pi} \; \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ Z_N(\psibf) },
+\]
+where 
+\begin{equation}\label{eq:ZpsiCondGaussProc}
+Z_N(\psibf) = \frac{1}{\sqrt{N}} \sum_{n=1}^{N} g_n f_{nN}(\psibf, \Phi_n),
+\end{equation}
+and where $\expect$ runs over both $\{g_n\}$ and $\{\Phi_n\}$.
+Conditionally on $\{\Phi_n\}$, the process $\{Z_N(\psibf), \psibf \in \reals^{m+1} \}$ is a \emph{Gaussian process}, and numerous techniques exist for its analysis.
+Lemma~\ref{lem:chaining} shows that for any $\kappa > 0$ there exists an $\epsilon > 0$ such that $\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ Z_N(\psibf) } < \kappa.$  Thus $\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)}  <  \sqrt{2\pi} \; \kappa$, and by Markov's inequality,
+\[
+\prob \cubr{  \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} > \delta } \leq  \sqrt{2\pi} \; \frac{\kappa}{\delta},
+\]
+for any $\delta > 0$.  The proof follows with $\nu =  \sqrt{2\pi} \kappa/\delta$.  It remains to prove Lemma~\ref{lem:chaining}.
+%Lemmas~\ref{lem:symmetrisation}~and~\ref{lem:chaining}.
+
+\end{proof}
+
+% The proof of the next lemma is based on a technique called \emph{symmetrisation}~\cite{Gine_Zinn_symmetrisation_1984} and the proof we give closely follows Pollard~\cite[Sec.~4]{Pollard_asymp_empi_proc_1989}.
+
+% \begin{lemma} \label{lem:symmetrisation}(Symmetrisation)
+% \[
+% \expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} \leq \sqrt{2\pi} \; \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ Z_N(\psibf) }.
+% \]
+% \end{lemma}
+% \begin{proof}
+% Let $\{\Phi_n^\prime\}$ be a sequence of random variables distributed identically to $\{\Phi_n\}$ and independent of both $\{\Phi_n\}$ and $\{g_n\}$.
+Write $\expect_\Phi$ to denote expectation conditional on $\{\Phi_n\}$.
+%That is, expectation taken over $\{g_n\}$ and $\{\Phi_n^\prime\}$ treating $\{\Phi_n\}$ as fixed.
+% Since $\Phi_n$ and $\Phi_n^\prime$ are identically distributed $\expect f_{nN}(\psibf, \Phi_n) = \expect_\Phi f_{nN}(\psibf, \Phi_n^\prime)$ so that,
+% \begin{align*}
+% G_N(\psibf) &= \frac{1}{\sqrt{N}} \sum_{n=1}^{N} \big( f_{nN}(\psibf, \Phi_n) - \expect_\Phi f_{nN}(\psibf, \Phi_n^\prime) \big)  \\
+% &= \expect_\Phi \frac{1}{\sqrt{N}} \sum_{n=1}^{N} ( f_{nN} - f_{nN}^\prime ),
+% \end{align*}
+% where we put $f_{nN} = f_{nN}(\psibf, \Phi_n)$ and  $f_{nN}^\prime = f_{nN}(\psibf, \Phi_n^\prime)$  for notational convenience.
+% Taking absolute values followed by supremums,
+% \begin{align}
+%  \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} &= \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{\frac{1}{\sqrt{N}} \expect_\Phi \sum_{n=1}^{N} (f_{nN} - f_{nN}^\prime )} \nonumber \\
+% &\leq \sup_{\|\psibf\|_{\infty} < \epsilon} \expect_\Phi  \abs{ \frac{1}{\sqrt{N}} \sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) } \nonumber  \\
+% &\leq \expect_\Phi \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} ( f_{nN} - f_{nN}^\prime ) }, \label{eq:Eoutsidesup}
+% \end{align}
+% the second line following from Jensen's inequality and the third line since $\sup \expect \abs{\dots} \leq \expect \sup \abs{\dots}$.  Applying $\expect$ to both sides gives the inequality
+% \begin{equation}\label{eq:Gninequfin}
+%  \expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} \leq  \expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) },
+% \end{equation}
+% because $\expect \expect_\Phi$ is equivalent to $\expect$.
+Let $\sigma_n = \frac{g_n}{\abs{g_n}}$ and put $\sigma_n = 1$ if $g_n = 0$.
+The $\sigma_n$ are independent with $\prob\{ \sigma_n = -1 \} = \tfrac{1}{2}$ and $\prob\{ \sigma_n = 1 \} = \tfrac{1}{2}$.
+The symmetry of the distribution of $g_n$ implies that $g_n$ and $\sigma_n$ are independent.
+As $\Phi_n$ and $\Phi_n^\prime$ are independent and identically distributed, the random variable $f_{nN} - f_{nN}^\prime$ is symmetrically distributed about zero, and is therefore distributed identically to $\sigma_n( f_{nN} - f_{nN}^\prime)$.
+Thus,
+% %\[
+% %\sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) \qquad \text{and} \qquad \sum_{n=1}^{N}  \sigma_n(f_{nN} - f_{nN}^\prime)
+% %\]
+% %are identically distributed, and therefore
+% \begin{align*}
+%  \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{  \frac{1}{\sqrt{N}}\sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) } = \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{  \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n ( f_{nN} - f_{nN}^\prime) }.
+% \end{align*}
+% By the triangle inequality
+% \[
+% \abs{ \sum_{n=1}^{N} \sigma_n (f_{nN} - f_{nN}^\prime)} \leq \abs{ \sum_{n=1}^{N} \sigma_n f_{nN} } + \abs{ \sum_{n=1}^{N}  \sigma_n f_{nN}^\prime },
+% \]
+% and it follows from~\eqref{eq:Gninequfin}, that 
+% \begin{align*}
+%  \expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf)} &\leq \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} } + \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{\frac{1}{\sqrt{N}} \sum_{n=1}^{N}  \sigma_n  f_{nN}^\prime } \\
+% &= 2 \; \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} }.
+% \end{align*}
+% Write $\expect_{\Phi\sigma}$ to denote expectation conditional on $\{\Phi_n\}$ and $\{\sigma_n\}$.
+Since $g_n$ is a standard normal random variable, $\expect \abs{g_n} = \expect_{\Phi\sigma} \abs{g_n} = \sqrt{2/\pi}$, and
+% \begin{align*}
+% \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} } &= \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} \sigma_n f_{nN} \sqrt{\frac{\pi}{2}}\expect_{\Phi\sigma} \abs{g_n}} \\
+% &\leq \sqrt{\frac{\pi}{2}} \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{ \sqrt{N}}\sum_{n=1}^{N} \sigma_n \abs{g_n}  f_{nN}},
+% \end{align*}
+% the last line following from Jensen's inequality and by moving $\expect_{\Phi\sigma}$ outside the supremum similarly to~\eqref{eq:Eoutsidesup}. As $g_n = \sigma_n \abs{g_n}$, it follows that
+% \begin{align*}
+% \expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{G_N(\psibf)}  &\leq 2\sqrt{\frac{\pi}{2}} \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ \frac{1}{\sqrt{N}}\sum_{n=1}^{N} g_n  f_{nN}} \\
+% &= \sqrt{2\pi} \; \expect \sup_{\|\psibf\|_{\infty} < \epsilon}  \abs{ Z_N(\psibf) } .
+% \end{align*}
+% \end{proof}
+
+\begin{lemma} \label{lem:chaining}
+For any $\kappa > 0$ there exists $\epsilon > 0$ such that
+\[
+\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ Z_N(\psibf) } < \kappa.
+\]
+%where $Z_N(\psibf)$ is defined by~\eqref{eq:ZpsiCondGaussProc}.
+\end{lemma}
+\begin{proof}
+Without loss of generality, assume that $\epsilon < \frac{1}{m+1}$.  Lemma~\ref{lem:chaining2} shows that
+\begin{equation}\label{eq:supZCK1} 
+\expect_\Phi \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ Z_N(\psibf) } \leq K_1 \sqrt{C_\epsilon(\{\Phi_n\})},
+\end{equation}
+where $K_1$ is a finite, positive constant, and $C_\epsilon(\{\Phi_n\})$ is the average number of times $\abs{\Phi_1}, \dots, \abs{\Phi_N}$ is greater than or equal to $\nicefrac{1}{2} - (m+1)\epsilon$.
+That is,
+\begin{equation}\label{eq:Cedefn}
+C_\epsilon(\{\Phi_n\}) = \frac{1}{N} \sum_{n=1}^{N} I_\epsilon(\abs{\Phi_n}),
+\end{equation}
+where $I_\epsilon(\abs{\Phi_n})$ is 1 when $\abs{\Phi_n} \geq \nicefrac{1}{2} - (m+1)\epsilon$ and zero otherwise.
+Recall that $f$ is the probability density function of $\Phi_n$, and (by assumption in Theorem~\ref{thm:asymp_proof}) that $f(\fracpart{x})$ is continuous at $x = -\nicefrac{1}{2}$.
+Because of this, the expected value of $C_\epsilon(\{\Phi_n\})$ is small when $\epsilon$ is small, since
+\begin{align*}
+\expect C_\epsilon(\{\Phi_n\}) &= \frac{1}{N} \sum_{n=1}^{N} \expect I_\epsilon(\abs{\Phi_n}) \\
+&= \prob\cubr{\abs{\Phi_1} \geq \nicefrac{1}{2} - (m+1)\epsilon} \\
+&= \int_{-1/2}^{-1/2 + (m+1)\epsilon} f(\phi) d\phi + \int_{1/2 - (m+1)\epsilon}^{1/2} f(\phi) d\phi \\
+&= \int_{-1/2 + (m+1)\epsilon}^{1/2 - (m+1)\epsilon} f(\fracpart{\phi}) d\phi \\
+&= 2(m+1)\epsilon \big(f(-\nicefrac{1}{2}) + o(1)\big),
+\end{align*}
+where $o(1)$ goes to zero as $\epsilon$ goes to zero.
+Since $\sqrt{\cdot}$ is a concave function on the positive real line and $C_{\epsilon}(\{\Phi_n\})$ is nonnegative, it follows from Jensen's inequality that $\expect \sqrt{C_\epsilon(\{\Phi_n\})} \leq  \sqrt{\expect  C_\epsilon(\{\Phi_n\} )} < \sqrt{K_2 \epsilon }$ for some constant $K_2$.
+Applying $\expect$ to both sides of~\eqref{eq:supZCK1} gives
+\[
+\expect \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ Z_N(\psibf) } \leq K_1 \sqrt{\expect C_\epsilon(\{\Phi_n\})} < K_1 \sqrt{K_2 \epsilon}.
+\]
+Choosing $\epsilon = \kappa^2/(K_1^2 K_2)$ completes the proof.  It remains to prove Lemma~\ref{lem:chaining2}.
+\end{proof}
+
+The proofs of Lemmas~\ref{lem:chaining2}~and~\ref{lem:chaining3} are based on a technique called \emph{chaining} (or \emph{bracketing})~\cite{Dudley_unif_central_lim_th_1999,Ossiander_clt_bracketing_1984,Pollard_asymp_empi_proc_1989,Pollard_new_ways_clts_1986,van2009empirical}.
+The proofs here follow those of Pollard~\cite{Pollard_asymp_empi_proc_1989}.
+In the remaining lemmas we consider expectation conditional on $\{\Phi_n\}$ and treat $\{\Phi_n\}$ as a fixed realisation.
+We consequently use the abbreviations $C_\epsilon = C_\epsilon(\{\Phi_n\})$ and $f_{nN}(\psibf) = f_{nN}(\psibf, \Phi_n)$.
+As in Lemma~\ref{lem:chaining} we assume, without loss of generality, that $\epsilon < \tfrac{1}{m+1}$.
+%This comes at no loss since Lemma~\ref{lem:chaining} only asks for the existence of an $\epsilon > 0$.
+
+\begin{lemma}\label{lem:chaining2}
+There exists a positive constant $K_1$ such that
+\[
+\expect_\Phi \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ Z_N(\psibf) } \leq K_1 \sqrt{C_\epsilon}.
+\]
+\end{lemma}
+\begin{proof}  Let $B_\epsilon = \{\xbf \in \reals \mid \|\xbf\|_\infty < \epsilon \}$.
+%We are now required to prove that
+%\[
+%\expect_\Phi \sup_{\psibf \in B_\epsilon} \abs{ Z_N(\psibf) } \leq K_1 \sqrt{C_\epsilon}.
+%\]
+For each non negative integer $k$, let $T_\epsilon(k)$ be a discrete subset of $\reals^{m+1}$ with the property that for every $\psibf \in B_\epsilon$ there exists some $\psibf^* \in T_\epsilon(k)$ such that the pseudometric
+\[
+d(\psibf, \psibf^*) = \sum_{n = 1}^N \big( f_{nN}(\psibf) - f_{nN}(\psibf^*) \big)^2 \leq 2^{-k} C_\epsilon N.
+\]
+We define $T_\epsilon(0)$ to contain a single point, the origin $\zerobf$.
+Defined this way $T_\epsilon(0)$ satisfies the inequality above because $d(\psibf, \zerobf) = \sum_{n = 1}^N f_{nN}(\psibf)^2  \leq C_\epsilon N$ for all $\psibf \in B_\epsilon$, as a result of Lemma~\ref{lem:epslmlemma}.
+
+The existence of $T_\epsilon(k)$ for each positive integer $k$ will be proved in Lemma~\ref{lem:metricentropy}.
+It is worth giving some intuition regarding $T_\epsilon(k)$.
+If we place a `ball' of radius $2^{-k} C_\epsilon N$  with respect to the pseudometric $d(\cdot, \cdot)$ around each point in $T_\epsilon(k)$, then, by definition, the union of these balls is a superset of $B_\epsilon$.
+The balls are said to \emph{cover} $B_\epsilon$ and $T_\epsilon(k)$ is said to form a \emph{covering} of $B_\epsilon$~\cite[Section 1.2]{Dudley_unif_central_lim_th_1999}.
+The minimum number of such balls required to cover $B_\epsilon$ is called a \emph{covering number} of $B_\epsilon$.
+In Lemma~\ref{lem:metricentropy} we show that no more than $K_3 2^{(m+1)k}$ balls of radius $2^{-k} C_\epsilon N$ are required to cover $B_\epsilon$, that is $\abs{T_\epsilon(k)} \leq K_3 2^{(m+1)k}$, where $K_3$ is a constant, independent of $N$ and $\epsilon$.
+%We make use of this upper bound now.
+%The lemma also upper bounds the number of elements by $\abs{T_\epsilon(k)} < K_3 2^{(m+1)k}$ where $K_3$ is a constant independent of $N$ and $\epsilon$.
+
+Since $f_{nN}(\psibf) = \left( \tfrac{n}{N}\right)^\ell \round{\Phi_n + p_{nN}(\psibf)}$ is a multiple of $N^{-\ell}$ for all $\psibf \in \reals^{m+1}$, it follows that $d(\psibf, \psibf^*)$ is a multiple of $N^{-2\ell}$.
+When $2^k > C_\epsilon N^{1+2\ell}$ we have $0 \leq d(\psibf, \psibf^*) < 2^{-k} C_\epsilon N < N^{-2\ell}$, and so $d(\psibf, \psibf^*) = 0$, and consequently $f_{nN}(\psibf) = f_{nN}(\psibf^*)$  for every $n = 1, \dots, N$ and $Z_N(\psibf) = Z_N(\psibf^*)$.
+Thus,
+\[
+\sup_{\|\psibf\|_{\infty} < \epsilon } \abs{ Z_N(\psibf) } = \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) }
+\]
+for all $k$ large enough that $2^{k} > C_\epsilon N^{1+2\ell}$.
+So, to analyse the supremum of $Z_N(\psibf)$ over the continuous interval $B_\epsilon$ it is enough to analyse the supremum over the discrete set $T_\epsilon(k)$ for large $k$.
+% The chaining argument~\cite[Sec.~3]{Pollard_asymp_empi_proc_1989} can  be used to show that
+% \[
+% \expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) } < \expect_{\Phi} \sup_{\psibf \in T_\epsilon(k-1)} \sabs{ Z_N(\psibf) }  + \sqrt{C_\epsilon}\frac{6 \sqrt{\log\sabs{T_\epsilon(k)}}}{2^{k/2}},
+% \]
+% which involves a recursion in $k$, where $\log(\cdot)$ is the natural logarithm.
+Since $T_\epsilon(0)$ contains only the origin, $\expect_{\Phi}\sup_{\psibf \in T_\epsilon(0)} \abs{ Z_N(\psibf) } = \expect_{\Phi}\abs{ Z_N(0) } = 0$ and, by unravelling the recursion, we obtain
+% \[
+% \expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) } < \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{6\sqrt{\log\abs{T_\epsilon(i)}}}{2^{i/2}}.
+% \]
+% Now $\log\abs{T_\epsilon(i)} \leq \log(K_3) + i(m+1)\log(2)$ by Lemma~\ref{lem:metricentropy}, and so the sum above converges as $k\rightarrow\infty$.
+The lemma holds with $K_1 =  \sum_{i=1}^{\infty}\frac{6\sqrt{\log\abs{T_\epsilon(i)}}}{2^{i/2}}$.
+% as required.
+Lemma~\ref{lem:chaining3} shows that 
+\[
+\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) } \leq \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{\sqrt{ i A_1 + A_2}}{2^{i/2}} < \infty
+\]
+for every positive integer $k$, where $A_1 = 18(m+1)\log 2$ and $A_2 = 18\log K_3$ are constants and $\log(\cdot)$ is the natural logarithm.
+The lemma holds with $K_1 =  \sum_{i=1}^{\infty}2^{-i/2}\sqrt{ i A_1 + A_2 }$.
+\end{proof}
+
+\begin{lemma}\label{lem:epslmlemma}
+For $\epsilon < \frac{1}{m+1}$ and all $\psibf \in B_\epsilon$ and $n = 1, \dots, N$,
+\[
+f_{nN}(\psibf)^2 \leq \abs{f_{nN}(\psibf)} \leq I_\epsilon(\abs{\Phi_n})
+\]
+and consequently,
+\begin{equation}\label{eq:sumf2absfCe}
+\sum_{n=1}^{N}f_{nN}(\psibf)^2 \leq \sum_{n=1}^{N}\abs{f_{nN}(\psibf)} \leq N C_\epsilon.
+\end{equation}
+\end{lemma}
+\begin{proof}
+Recall that $f_{nN}(\psibf) = \left( \tfrac{n}{N}\right)^\ell \round{\Phi_n + p_{nN}(\psibf)}$.  Because $\abs{\psi_i} < \epsilon$ for all $i = 0, \dots, m$, 
+\begin{equation}\label{eq:pnNsmall}
+\abs{p_{nN}(\psibf)} = \abs{\sum_{i=0}^{m}\big(\tfrac{n}{N}\big)^i \psi_i} \leq (m+1)\epsilon < 1.
+\end{equation}
+%for all $n = 1, \dots, N$.  
+Since $\Phi_n \in [-\nicefrac{1}{2}, \nicefrac{1}{2})$, it follows that $f_{nN}(\psibf)$ equals either $-(\tfrac{n}{N})^\ell$, $(\tfrac{n}{N})^\ell$ or $0$ and so
+\begin{equation}\label{eq:fnNleqfnNs}
+   f_{nN}(\psibf)^2 \leq \abs{f_{nN}(\psibf)} \leq 1.
+\end{equation}
+Whenever $f_{nN}(\psibf) \neq 0$ we must have 
+\[
+\abs{\Phi_n} \geq \nicefrac{1}{2} - \abs{p_{nN}(\psibf)} \geq \nicefrac{1}{2} - (m+1)\epsilon
+\]
+and $I_\epsilon(\abs{\Phi_n}) = 1$.  Thus, for all $n = 1, \dots, N$,
+\[
+f_{nN}(\psibf)^2 \leq \abs{f_{nN}(\psibf)} \leq I_\epsilon(\abs{\Phi_n}). %\qquad \text{for all $n = 1, \dots, N$.}
+\]
+Summing the terms in this inequality over $n = 1, \dots, N$ and using~\eqref{eq:Cedefn} gives~\eqref{eq:sumf2absfCe}.
+\end{proof}
+
+\begin{lemma}\label{lem:chaining3}(Chaining)
+For all positive integers $k$,
+\[
+\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) } \leq \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{\sqrt{ i A_1 + A_2 }}{2^{i/2}},
+\]
+where the constants $A_1 = 18(m+1)\log 2$ and $A_2 = 18\log K_3$.
+\end{lemma}
+\begin{proof}
+Let $b_k$ be a function that maps each $\psibf \in T_\epsilon(k)$ to $b_k(\psibf) \in T_\epsilon(k-1)$ such that  $d(\psibf,b_k(\psibf)) \leq 2^{1-k} C_\epsilon N$.
+The existence of the function $b_k$ is guaranteed by the definition of $T_\epsilon(k)$.
+By the triangle inequality,
+\[
+\abs{ Z_N(\psibf) } \leq \abs{ Z_N(b_k(\psibf)) } + \abs{ Z_N(\psibf) - Z_N(b_k(\psibf))  },
+\]
+and by taking supremums on both sides,
+\begin{align}
+\sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) } &\leq \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(b_k(\psibf)) } + \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) - Z_N(b_k(\psibf))  } \nonumber \\
+&\leq \sup_{\psibf \in T_\epsilon(k-1)} \sabs{ Z_N(\psibf) } + \sup_{\psibf \in T_\epsilon(k)} \sabs{ Z_N(\psibf) - Z_N(b_k(\psibf)) }, \label{eq:supZchain}
+\end{align}
+the last line following since $b_k(\psibf) \in T_\epsilon(k-1)$ and so 
+\[
+\sup_{\psibf \in T_\epsilon(k)} \abs{ Z_N(b_k(\psibf)) } \leq \sup_{\psibf \in T_\epsilon(k-1)} \abs{ Z_N(\psibf) }.
+\]
+Conditional on $\Phi_1, \Phi_2, \dots$ the random variable $X(\psibf) = Z_N(\psibf) - Z_N(b_k(\psibf))$ has zero mean, and is normally distributed with variance
+\begin{align*}
+\sigma_X^2 &= \expect_{\Phi} \frac{1}{N} \sum_{n=1}^N g_n^2 \big( f_{nN}(\psibf) - f_{nN}(b_k(\psibf)) \big)^2 \\
+&= \frac{1}{N} \sum_{n=1}^N \big( f_{nN}(\psibf) - f_{nN}(b_k(\psibf)) \big)^2 = d(\psibf, b_k(\psibf)) \leq 2^{1-k} C_\epsilon,
+\end{align*}
+because $\expect_{\Phi} g_n^2 = 1$.  Using Lemma~\ref{lem:maxineq}, 
+\begin{align*}
+\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k)} \sabs{X(\psibf)} \leq 3\sqrt{2^{1-k} C_\epsilon \log\sabs{T_\epsilon(k)}} \leq \sqrt{C_\epsilon}\frac{\sqrt{  k A_1 + A_2 }}{2^{k/2}}
+\end{align*}
+because $\log\abs{T_\epsilon(k)} \leq k(m+1)\log 2 +  \log K_3$.  Taking expectations on both sides of~\eqref{eq:supZchain} gives
+\begin{align*}
+\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \sabs{ Z_N(\psibf) } \leq \expect_{\Phi} \sup_{\psibf \in T_\epsilon(k-1)} \sabs{ Z_N(\psibf) }  + \sqrt{C_\epsilon}\frac{\sqrt{  k A_1 + A_2 }}{2^{k/2}},
+\end{align*}
+which involves a recursion in $k$.  By unravelling the recursion, and using the fact $T_\epsilon(0)$ contains only the origin, and therefore 
+\[
+\expect_{\Phi}\sup_{\psibf \in T_\epsilon(0)} \abs{ Z_N(\psibf) } = \expect_{\Phi}\abs{ Z_N(0) } = 0,
+\]
+we obtain $\expect_{\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ Z_N(\psibf) } < \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{\sqrt{  k A_1 + A_2 }}{2^{k/2}}$
+as required.
+\end{proof}
+
+\begin{lemma}\label{lem:maxineq}(Maximal inequality)
+Suppose $X_1, \dots, X_N$ are zero mean Gaussian random variables each with variance less than some positive constant $K$, then $\expect \sup_{n = 1, \dots, N} \abs{X_n} \leq 3  \sqrt{K \log N}$ where $\log N$ is the natural logarithm of $N$.
+\end{lemma}
+\begin{proof}
+This result is well known, see for example~\cite[Section 3]{Pollard_asymp_empi_proc_1989}  
+\end{proof}
+
+\begin{lemma}\label{lem:metricentropy} (Covering numbers)
+For $k \in \ints$ there exists a discrete set $T_\epsilon(k) \subset \reals^{m+1}$ with the property that, for every $\psibf \in B_\epsilon$, there is a $\psibf^* \in T_\epsilon(k)$ such that,
+\[
+d(\psibf,\psibf^*) = \sum_{n = 1}^N \big( f_{nN}(\psibf) - f_{nN}(\psibf^*) \big)^2 \leq \frac{C_\epsilon N}{2^k}.
+\]
+The number of elements in $T_\epsilon(k)$ is no more than $K_3 2^{(m+1)k}$ where $K_3$ is a positive constant, independent of $N$, $\epsilon$ and $k$.
+\end{lemma}
+
+Before we give the proof of this lemma we need some results from the literature on hyperplane arrangements and what are called \emph{$\epsilon$-cuttings}~\cite{Chazelle_discrepency_method_2000,Matousek_lect_disc_geom_2002}.
+Let $H$ be a set of $m$-dimensional affine hyperplanes lying in $\reals^{m+1}$.
+By \emph{affine} it is meant that  the hyperplanes need not pass through the origin.
+For each hyperplane $h \in H$ let $D(h)$ and its complement $\bar{D}(h)$ be the corresponding half spaces of $\mathbb{R}^{m+1}$.
+For a point $\xbf \in \mathbb{R}^{m+1}$, let
+\begin{equation}\label{eq:bhsi}
+b(h,\xbf) = \begin{cases} 
+1 & \xbf \in D(h) \\
+0 & \xbf \in \bar{D}(h).
+\end{cases}
+\end{equation}
+Note that $b(h,\xbf)$ is piecewise constant in $\xbf$.  Two points $\xbf$ and $\ybf$ from $\reals^{m+1}$ are in the same halfspace of $h$ if and only if $b(h, \xbf) = b(h, \ybf)$.  So the pseudometric
+\[
+\sigma(\xbf,\ybf) = \sum_{h \in H} |b(h,\xbf) - b(h,\ybf)|
+\]
+is the number of hyperplanes in $H$ that pass between the points $\xbf$ and $\ybf$.  %Observe that $\sigma(\xbf, \ybf)$ is peicewise constant in both $\xbf$ and $\ybf$.
+
+The next theorem considers the partitioning of $\reals^{m+1}$ into subsets so that not too many hyperplanes intersect with any subset.
+Proofs can be found in Theorem 5.1 on page 206 of~\cite{Chazelle_discrepency_method_2000} and also Theorem 6.5.3 on page 144 of~\cite{Matousek_lect_disc_geom_2002}.
+
+\begin{theorem}\label{thm:epscutting}
+There exists a constant $K$, independent of the set of hyperplanes $H$, such that for any positive real number $r$, we can partition $\reals^{m+1}$ into $K r^{m+1}$ generalised $(m+1)$-dimensional simplices with the property that no more than $\abs{H}/r$ hyperplanes from $H$ pass through the interior of any simplex.
+\end{theorem}
+
+By the phrase `There exists a constant $K$, independent of the set of hyperplanes $H$', it is meant that the constant $K$ is valid for every possible set of hyperplanes in $\reals^{m+1}$, regardless of the number of hyperplanes or their position and orientation.
+A \emph{generalised} $(m+1)$-dimensional simplex is the region defined by the intersection of $m+2$ half spaces in $\reals^{m+1}$.
+Note that a generalised simplex (unlike an ordinary simplex) can be unbounded.
+For our purposes Theorem~\ref{thm:epscutting} is important because of the following corollary.
+
+\begin{corollary}\label{cor:epscutting}
+There exists a constant $K$, independent of the set of hyperplanes $H$, such that for every positive real number $r$ there is a discrete subset $T \subset \reals^{m+1}$ containing no more than $K r^{m+1}$ elements with the property that for every $\xbf \in \reals^{m+1}$ there exists $\ybf \in T$ with $\sigma(\xbf, \ybf) \leq \abs{H}/r$.
+\end{corollary}
+\begin{proof}
+Let $C$ be the set of generalised simplices constructed according to Theorem~\ref{thm:epscutting}.
+Define $T$ as a set containing precisely one point from the interior of each simplex in $C$.
+Let $\xbf \in \reals^{m+1}$.
+Since $b(h, \xbf)$ is piecewise constant for each $h \in H$, and since $C$ partitions $\reals^{m+1}$, there must exist a simplex $c \in C$ with a point $\zbf$ in its interior such that $b(h,\zbf) = b(h, \xbf)$ for all $h \in H$, and correspondingly $\sigma(\zbf, \xbf) = 0$.
+Let $\ybf$ be the element from $T$ that is in the interior of $c$.
+Since at most $\abs{H}/r$ hyperplanes cross the interior of $c$ there can be at most $\abs{H}/r$ hyperplanes between $\zbf$ and $\ybf$, and so $\sigma(\zbf, \ybf) \leq \abs{H}/r$.
+Now $\sigma(\xbf, \ybf) < \sigma(\zbf, \xbf) + \sigma(\zbf, \ybf) \leq \frac{\abs{H}}{r}$ follows from the triangle inequality.
+\end{proof}
+
+The previous corollary ensures that we can \emph{cover} $\reals^{m+1}$ using $K r^{m+1}$ `balls' of radius $\abs{H}/r$ with respect to the pseudometric $\sigma(\xbf, \ybf)$.
+The balls are placed at the positions defined by points in the set $T$, and these points could be anywhere in $\reals^{m+1}$.
+The next corollary asserts that we can cover a subset of $\reals^{m+1}$, by placing the balls at points only within this subset.
+
+\begin{corollary}\label{cor:epscuttingsubset}
+Let $B$ be a subset of $\reals^{m+1}$.
+There exists a constant $K$, independent of the set of hyperplanes $H$, such that for every positive real number $r$ there is a discrete subset $T_B \subset B$ containing no more than $K r^{m+1}$ elements with the property that for every $\xbf \in B$ there exists $\ybf \in T_B$ with $\sigma(\xbf, \ybf) \leq \abs{H}/r$.
+\end{corollary}
+\begin{proof}
+Let $C$ be the set of generalised simplices constructed according to Theorem~\ref{thm:epscutting} and let $C_B$ be the subset of those indices that intersect $B$.
+Let $T_B$ contain a point from $c \cap B$ for each simplex $c \in C_B$.
+The proof now follows similarly to Corollary~\ref{cor:epscutting}.
+\end{proof}
+
+We are now ready to prove Lemma~\ref{lem:metricentropy}.
+
+\begin{proof} (Lemma~\ref{lem:metricentropy})
+Put $g_{nN}(\psibf) = (\tfrac{N}{n})^\ell f_{nN}(\psibf) = \round{ \Phi_n + p_{nN}(\psibf)}$, and let
+\[
+d_g(\psibf, \psibf^*) = \sum_{n=1}^{N} \big( g_{nN}(\psibf) - g_{nN}(\psibf^*) \big)^2.
+\]
+We have $d(\psibf, \psibf^*) \leq d_g(\psibf, \psibf^*)$, and so it suffices to prove the lemma with $d$ replaced by $d_g$.  
+%It will be more convenient to work with $d_g$ rather than $d$.
+From~\eqref{eq:pnNsmall} it follows that $\abs{p_{nN}(\psibf)} \leq (m+1)\epsilon < 1$.  Since $\Phi_n \in [-\nicefrac{1}{2}, \nicefrac{1}{2})$, when $\Phi_n \geq 0$,
+\[
+g_{nN}(\psibf) = \begin{cases}
+1 & p_{nN}(\psibf) \geq \nicefrac{1}{2} - \Phi_n \\
+0 & \text{otherwise},
+\end{cases}
+\]
+and when $\Phi_n < 0$,
+\[
+g_{nN}(\psibf) = \begin{cases}
+-1 & p_{nN}(\psibf) < -\nicefrac{1}{2} - \Phi_n \\
+0 & \text{otherwise}.
+\end{cases}
+\]
+Thus, $( g_{nN}(\psibf) - g_{nN}(\psibf^*) )^2$ is either equal to one when $g_{nN}(\psibf) \neq g_{nN}(\psibf^*)$ or zero when $g_{nN}(\psibf) = g_{nN}(\psibf^*)$.
+Now $g_{nN}(\psibf) \neq 0$ only if
+\[
+\abs{\Phi_n} \geq \nicefrac{1}{2} - \abs{p_{nN}(\psibf)} \geq \nicefrac{1}{2} - (m+1)\epsilon,
+\]
+that is, only if $I_\epsilon(\Phi_n) = 1$.
+Let $A = \{ n \in \{1, \dots, N\} \mid I_\epsilon(\Phi_n) = 1 \}$ be the subset of the indices where $I_\epsilon(\Phi_n) = 1$.
+By definition the number of elements in $A$ is $C_\epsilon N$ (see~\eqref{eq:Cedefn}).
+If both $\psibf$ and $\psibf^*$ are in $B_\epsilon$, then
+\[
+(g_{nN}(\psibf) - g_{nN}(\psibf^*) )^2 \neq 0 
+\]
+only if $n \notin A$.  Thus,
+\begin{align*}
+d_g(\psibf, \psibf^*) = \sum_{n=1}^{N} \big( g_{nN}(\psibf) - g_{nN}(\psibf^*) \big)^2 = \sum_{n \in A} \big( g_{nN}(\psibf) - g_{nN}(\psibf^*) \big)^2.
+\end{align*}
+
+We now use Corollary~\ref{cor:epscuttingsubset}.  Let $h_n$ be the $m$ dimensional hyperplane in $\reals^{m+1}$ satisfying
+\[
+p_{nN}(\psibf) =  \sum_{i=0}^{m} \big(\tfrac{n}{N}\big)^i \psi_i = \tfrac{1}{2} \sign{\Phi_n} - \Phi_n
+\]
+where $\sign{\Phi_n}$ is equal to $1$ when $\Phi_n \geq 0$ and $-1$ otherwise.
+The hyperplane $h_n$ divides $\reals^{m+1}$ into two halfspaces, $D(h_n)$ and its complement $\bar{D}(h_n)$.
+If $\psibf$ and $\psibf^*$ are in the same halfspace, then $\abs{ b(h_n,\psibf) - b(h_n,\psibf^*) } = 0$ and $g_{nN}(\psibf) = g_{nN}(\psibf^*)$, and therefore $( g_{nN}(\psibf) - g_{nN}(\psibf^*) )^2 = 0$.
+Otherwise, if $\psibf$ and $\psibf^*$ are in different halfspaces, then $\abs{ b(h_n,\psibf) - b(h_n,\psibf^*) } = 1$ and $g_{nN}(\psibf) \neq g_{nN}(\psibf^*)$, and therefore $( g_{nN}(\psibf) - g_{nN}(\psibf^*) )^2 = 1$.
+Thus,
+\[
+(g_{nN}(\psibf) - g_{nN}(\psibf^*) )^2 = \abs{ b(h_n,\psibf) - b(h_n,\psibf^*) }
+\]
+for $n = 1, \dots,  N$.
+Let $H$ be the finite set of hyperplanes $\{ h_n, n \in A\}$ and observe that the number of hyperplanes is $\abs{H} = \abs{A} = C_\epsilon N$.
+When both $\psibf$ and $\psibf^*$ are inside $B_\epsilon$, $d_g$ can be written as
+\begin{align*}
+d_g(\psibf, \psibf^*) &= \sum_{n\in A} \abs{ b(h_n,\psibf) - b(h_n,\psibf^*) } \\
+&=  \sum_{h \in H} \abs{ b(h,\psibf) - b(h,\psibf^*) } = \sigma(\psibf, \psibf^*).
+\end{align*}
+That is, when both $\psibf, \psibf^* \in B_\epsilon$, $d_g(\psibf, \psibf^*)$ is the number of hyperplanes from $H$ that pass between the points $\psibf$ and $\psibf^*$.
+
+It follows from Corollary~\ref{cor:epscuttingsubset} that for any positive $r$ there exists a finite subset $T_B$ of $B_\epsilon$ containing at most $K_3 r^{m+1}$ elements, such that for every $\psibf \in B_\epsilon$ there is a $\psibf^* \in T_B$ with
+\[
+d_g(\psibf, \psibf^*) = \sigma(\psibf, \psibf^*) \leq \frac{\abs{H}}{r} = \frac{\abs{A}}{r} = \frac{C_\epsilon N}{r}.
+\]
+Putting $r = 2^k$ and choosing $T_\epsilon(k) = T_B$ completes the proof.
+\end{proof}
+
+%\bibliographystyle{acmtrans-ims}
+\bibliographystyle{IEEEbib}
+\bibliography{bib}
+
 \end{document}

### 2012-11-02 paper.tex
removed aoap header

@@
 % natbib option: authoryear
-\documentclass[aap]{imsart}
+\documentclass[aap,preprint]{imsart}
+
+\setattribute{journal}{name}{}
 
@@
 
+Although we will not prove it here the assumption that $\Phi_1,\dots,\Phi_N$ have zero intrinsic mean is not only sufficient, but also necessary, for if $\Phi_1,\dots,\Phi_N$ have intrinsic mean $x \in [-\nicefrac{1}{2},\nicefrac{1}{2})$ with $x \neq 0$ then $\sfracpart{\widehat{\lambda}_{0,N} - x}\rightarrow 0$ almost surely as $N\rightarrow\infty$, and so $\widehat{\lambda}_{0,N}$ does not converge to zero.
+On the other hand if $\Phi_1,\dots,\Phi_N$ do not have an intrinsic mean then $\widehat{\lambda}_{0,N}$ will not converge.
+
 The proof of asymptotic normality places requirements on the probability density function $f$ of the phase noise $\Phi_1, \dots, \Phi_N$.

### 2013-02-03 paper.tex
moved stuff around.  Squashed stuff back into one papers.  Needs heay editing.

--- 2012-11-02 paper.tex
@@
 
-\title{Polynomial phase estimation by phase unwrapping 1: Identifiability and strong consistency}
-%\title{Polynomial phase estimation by phase unwrapping}
+%\title{Polynomial phase estimation by phase unwrapping 1: Identifiability and strong consistency}
+\title{Polynomial phase estimation by phase unwrapping}
 
@@
 
-%A popular estimator is based on the \emph{discrete polynomial phase transform} (DPT) first introduced by Peleg and Porat \cite{Peleg_DPT_1995}.
-The transform enables each parameter to be be estimated iteratively using the Fast Fourier transform in a way analogous to the frequency estimator of Rife and Boorstyn \cite{Rife1974, Quinn2001}.
-The DPT estimator is characterized by good statistical performance and computational efficiency.
-Variants of the DPT have been suggested by numerous authors including Golden and Friedlander \cite{Golden_mod_dpt_1998} and O'Shea \cite{Oshea_iterative_1996}.
-The recent \emph{high-order phase function} estimators are based on similar ideas \cite{Farquharson_another_poly_est_2005}.
-A substantial drawback of all of these estimators is that they only work correctly over a small set of parameters, generally much smaller than the identifiable region.
-This property appears to have been first noticed in \cite{McKilliam2009IndentifiabliltyAliasingPolyphase, McKilliam2009asilomar_polyest_lattice} and has been analysed further in \cite[Ch.
-10]{McKilliam2010thesis}.
-%
-%Another estimator, described by Kitchen \cite{Kitchen_polyphase_unwrapping_1994}, is based on the phase unwrapping approach to frequency estimation suggested by Kay \cite{Kay1989}.
-The primary advantage of Kitchen's estimator is that it requires only $O(N)$ arithmetic operations to compute.
-The primary disadvantage is that it only works correctly when the signal to noise ratio (SNR) is high.
-It was also observed in \cite{McKilliam2009asilomar_polyest_lattice} that Kitchen's estimator does not work correctly for all coefficients in the identifiable region, but, the problem is far less severe than for estimators based on the DPT \cite[Sec.
-10.3]{McKilliam2010thesis}.
-
-%An obvious estimator of the unknown coefficients is the least squares estimator (LSE).
+An obvious estimator of the unknown coefficients is the least squares estimator (LSE).
 This is also the maximum likelihood estimator (MLE) when the noise sequence $\{X_n\}$ is white and Gaussian.
@@
 
-One estimation approach attempts to perform polynomial regression on the phase of the signal~\cite{Slocumb_polynomial_1994,Tretter1985,Kitchen_polyphase_unwrapping_1994,Djuric_phase_unwrap_chirp_1990,Morelande_bayes_unwrapping_2009_tsp}.
-This is complicated by the fact that the phase is \emph{wrapped} modulo $2\pi$ and must be \emph{unwrapped} before regression can be performed.
-In this two part series of papers we consider the estimator that results from unwrapping the phase in a least squares manner.
+In this paper we consider the estimator that results from unwrapping the phase in a least squares manner.
 We call this the \emph{least squares unwrapping estimator} (LSU)~\cite{McKilliam2009asilomar_polyest_lattice, McKilliamFrequencyEstimationByPhaseUnwrapping2009}\cite[Chap.
@@
 It was shown in \cite{McKilliam2009asilomar_polyest_lattice, McKilliamFrequencyEstimationByPhaseUnwrapping2009} that the LSU estimator can be represented as a \emph{nearest lattice point problem}, and Monte-Carlo simulations were used to show the LSU estimator's favourable statistical performance.
-%Moreover, it was noticed that the LSU appears to work correctly for coefficients anywhere in the identifiable region.
+Moreover, it was noticed that the LSU appears to work correctly for coefficients anywhere in the identifiable region.
 A drawback of the LSU estimator is that computing a nearest lattice point is, in general, computationally difficult.
@@
 A major point of interest is that the lattices considered are not \emph{random}, and therefore may admit fast nearest point algorithms.
-%This has been studied in~\cite[Sec.~4.2]{McKilliam2010thesis} where polynomial time algorithms where found that compute the nearest point exactly.
+This has been studied in~\cite[Sec.~4.2]{McKilliam2010thesis} where polynomial time algorithms where found that compute the nearest point exactly.
 Unfortantely, although polynomial time, the algorithm are still computationally very slow in practice.
-% Two main questions were raised in \cite{McKilliam2009asilomar_polyest_lattice}:
-% \begin{enumerate}
-% \vspace{-0.1cm}
-% \item What are the asymptotic statistical properties of the LSU estimator?
-% \vspace{-0.18cm}
-% \item Can the structure of the lattice be exploited to obtain fast (exact or approximate) nearest point algorithms and thereby a computationally efficient estimator?
-% \vspace{-0.1cm}
-% \end{enumerate}
-% The purpose of this paper is to answer the first question. We prove that the LSU estimator is strongly consistent and asymptotically normally distributed. 
-%We also apply another known approximate nearest point algorithm, called the $K$-best method \cite{Zhan2006_K_best_sphere_decoder}, that we show provides near sphere decoder performance, but can be computed in a reasonable amount of time if $N$ is less than about 300.
-In this two part series of we derive the asymptotic statistical properties of the LSU estimator showing, under some assumptions on the noise $X_1,\dots,X_N$, that it is strongly consistent and asymptotically normally distributed.
-Strong consistency is proved in this paper, while asymptotic normality is proved in the second part~\cite{McKilliam_pps2_2012}.
+ Two main questions were raised in \cite{McKilliam2009asilomar_polyest_lattice}:
+ \begin{enumerate}
+ \vspace{-0.1cm}
+ \item What are the asymptotic statistical properties of the LSU estimator?
+ \vspace{-0.18cm}
+ \item Can the structure of the lattice be exploited to obtain fast (exact or approximate) nearest point algorithms and thereby a computationally efficient estimator?
+ \vspace{-0.1cm}
+ \end{enumerate}
+ The primary purpose of this paper is to answer the first question.
 Similar results were stated without a complete proof in~\cite{McKilliam_polyphase_est_icassp_2011}.
@@
 The results here are also more general than in~\cite{McKilliam_polyphase_est_icassp_2011}, allowing for a wider class of noise distributions.
+We prove that the LSU estimator is strongly consistent and asymptotically normally distributed.
+BLERG `small noise' assumption.
+
+We also apply another known approximate nearest point algorithm, called the $K$-best method \cite{Zhan2006_K_best_sphere_decoder}, that we show provides near sphere decoder performance, but can be computed in a reasonable amount of time if $N$ is less than about 300.
+BLERG `better than everyting else and computationally reasonably for smallish $N \leq 300$'.
 
@@
 
-It may be helpful to observe Figures~\ref{fig:circstatplot_zero} to~\ref{fig:circstatplot_cube}.
+It may be helpful to observe Figures~\ref{circstatplot_line} to~\ref{fig:circstatplot_cube}.
 In these, the phase (divided by $2\pi$) of two distinct polynomial phase signals is plotted on the left, and on the right the principal component of the phase (also divided by $2\pi$) is plotted.
@@
 
-\begin{figure}[p]
-	\centering
-		\includegraphics[width=\linewidth]{plots/circstatfigzero-1.mps}
-		\caption{The zeroth order polynomials $\tfrac{7}{10}$ (solid line) and $\tfrac{17}{10}$ (dashed line).}		
-\label{fig:circstatplot_zero}
-\end{figure}
-
-\begin{figure}[p]
+% \begin{figure}[p]
+% 	\centering
+% 		\includegraphics[width=\linewidth]{plots/circstatfigzero-1.mps}
+% 		\caption{The zeroth order polynomials $\tfrac{7}{10}$ (solid line) and $\tfrac{17}{10}$ (dashed line).}		
+% \label{fig:circstatplot_zero}
+% \end{figure}
+
+\begin{figure}[t]
 	\centering
@@
 
-\begin{figure}[p]
-	\centering
-		\includegraphics[width=\linewidth]{plots/circstatfigquad-1.mps}
-		\caption{The quadratic polynomials $\tfrac{1}{10} (15 - 15 t + 4 t^2)$ (solid line) and $\tfrac{1}{10}(25 -  t^2)$  (dashed line).}
-		\label{fig:circstatplot_quad}
-\end{figure}
-
-\begin{figure}[p]
+% \begin{figure}[p]
+% 	\centering
+% 		\includegraphics[width=\linewidth]{plots/circstatfigquad-1.mps}
+% 		\caption{The quadratic polynomials $\tfrac{1}{10} (15 - 15 t + 4 t^2)$ (solid line) and $\tfrac{1}{10}(25 -  t^2)$  (dashed line).}
+% 		\label{fig:circstatplot_quad}
+% \end{figure}
+
+\begin{figure}[t]
 	\centering
@@
 
-\section{The least squares unwrapping estimator}\label{sec:least-squar-unwr}
+\section{The Least squares unwrapping estimator}\label{sec:least-squar-unwr}
 
@@
 
-It is shown in~\cite[Sec~8.1]{McKilliam2010thesis}\cite{McKilliam2009asilomar_polyest_lattice} how this minimisation problem can be posed as that of computing a nearest lattice point in a particular lattice.
-Polynomial time algorithms that compute the nearest point are described in~\cite[Sec.~4.3]{McKilliam2010thesis}.
+It is shown in~\cite{McKilliam2009asilomar_polyest_lattice} and \cite[Sec~8.1]{McKilliam2010thesis} how this minimisation problem can be posed as a nearest lattice point problem, and we briefly discuss this here.
+Write $SS$ as
+\[
+SS(\mubf) = \sum_{n=1}^{N}{ \left( \Theta_{n} - W_n - \sum_{k = 0}^{m}{\mu_k n^k} \right)^2 },
+\]
+where $W_n = \round{\Theta_{n} - \sum_{k = 0}^{m}{\mu_k n^k} }$ for $n = 1, \dots, N$ are integers called \emph{wrapping variables}.
+If we consider $W_1,\dots,W_N$ as nuisance parameters to be estimated then $SS$ can be written as a function of both $\mu_0,\dots,\mu_m$ and $W_1,\dots,W_N$.
+The least squares unwrapping estimator is then found by jointly minimising over $\mu_0,\dots,\mu_m$ and $W_1,\dots,W_N$.
+This joint minimisation problem can be solved by computing a nearest point in a lattice.
+To see this write $SS$ as a function of both $\mu_0,\dots,\mu_m$ and $W_1,\dots,W_N$ using vectors,
+\begin{equation}\label{eq:SS_in_vector_form_with_u_nuisance}
+SS(\mubf, \wbf) = \| \thetabf - \Xbf\mubf - \wbf \|^2,
+\end{equation}
+where $\|\cdot\|^2$ denotes the squared Euclidean norm, and $\Xbf$ is the rectangular Vandermonde matrix
+\begin{equation}\label{eq:Xvandermonde}
+\Xbf = \left[ \begin{array}{ccccc} 
+1 & 1 & 1 & \cdots & 1 \\
+1 & 2 & 4 & \cdots & 2^m \\
+\vdots & \vdots  & \vdots & \ddots & \vdots \\
+1 & N & N^2 & \cdots & N^m \\
+\end{array} \right],
+\end{equation}
+and where the column vectors $\wbf = [W_1, \dots, W_N]^\prime$ and $\thetabf = [\Theta_1, \dots, \Theta_N]^\prime$.   For fixed $\wbf$ the minimiser of $SS(\mubf, \wbf)$ with respect to $\mubf$ is
+\begin{equation} \label{eq:hat_theta_cal}
+\mubf(\wbf) = (\Xbf^\prime\Xbf)^{-1}\Xbf^\prime( \thetabf - \wbf) = \Xbf^+( \thetabf - \wbf),
+\end{equation}
+where $\Xbf^+ = (\Xbf^\prime\Xbf)^{-1}\Xbf^\prime$ is the Moore-Penrose pseudoinverse of $\Xbf$.
+
+We may thus compute $\widehat{\mubf}$ as $\widehat{\mubf} = \mu(\widehat{\wbf})$, where $\widehat{\wbf}$ is the minimiser of
+\[
+SS(\wbf) = SS(\mubf(\wbf),\wbf) = \| \Qbf\thetabf - \Qbf\wbf \|^2,
+\]
+where $\Qbf = \Ibf - \Xbf\Xbf^+$ and $\Ibf$ is the $N \times N$ identity matrix\footnote{We have slightly abused notation here by reusing $SS$.
+This should not cause any confusion as $SS(\mubf)$ and $SS(\mubf,\wbf)$ and $SS(\wbf)$ are easily told apart by their inputs.}.
+
+Let $\Lambda$ be the lattice generated by $\Qbf$, that is $\Lambda = \{\Qbf\xbf \mid \xbf \in \ints^N \}$\footnote{The $N \times N$ projection matrix $\Qbf$ does not have full rank, it has rank $N-m-1$, and for this reason it is not strictly a generator matrix.
+However, a generator can be constructed by removing $m+1$ consecutive columns from $\Qbf$.
+In our implementation the last $m+1$ columns are removed.}.
+It follows that $\Qbf \wbf$ is a lattice point in $\Lambda$ and that minimising $SS(\wbf)$ is equivalent to finding the nearest lattice point in $\Lambda$ to $\Qbf\thetabf$.
+Denote this point by $\Qbf\widehat{\wbf}$.
+Now the estimate $\widehat{\mubf}$ is given by substituting $\widehat{\wbf}$ for $\wbf$ in~\eqref{eq:hat_theta_cal}.
+After this procedure it is possible that the $\widehat{\mubf}$ obtained is not in the identifiable region but is instead an aliased version of the desired estimate.
+This can be resolved by computing $\dealias(\widehat{\mubf})$.
+There exist algorithms that can compute a nearest point in $\Lambda$ that require a number of operations that is polynomial in $N$~\cite[Sec 4.3]{McKilliam2010thesis}.
 Although polynomial in complexity, these algorithms are not fast in practice.
 The existence of practically fast nearest point algorithms for these lattices is an interesting open problem.
-In this paper we focus on the asymptotic statistical properties of the LSU estimator, rather than computational aspects.
 
@@
 
-The \emph{intrinsic mean} or \emph{Fr\'{e}chet mean} of $\Phi_n$ is defined as~\cite{McKilliam_mean_dir_est_sq_arc_length2010,Bhattacharya_int_ext_means_2003,Bhattacharya_int_ext_means_2005},
+The \emph{intrinsic mean} or \emph{Fr\'{e}chet mean} of $\Phi_n$ is defined as~\cite{McKilliam_mean_dir_est_sq_arc_length2010,bwhk07a,Bhattacharya_int_ext_means_2003,Bhattacharya_int_ext_means_2005},
 \begin{equation}\label{eq:intrmeandefn}
@@
 \[
-\sigma_{\text{intr}}^2 = \expect\fracpart{\Phi_n - \mu_{\text{intr}}}^2 = \min_{\mu \in [-\nicefrac{1}{2}, \nicefrac{1}{2})} \expect \fracpart{\Phi_n - \mu}^2,
+\sigma_{\text{intr}}^2 = E\fracpart{\Theta - \mu_{\text{intr}}}^2 = \min_{\mu \in [-\nicefrac{1}{2}, \nicefrac{1}{2})} \expect \fracpart{\Phi_n - \mu}^2,
 \]
@@
 If the minimiser is not unique we say that $\Phi_n$ has no intrinsic mean.
-%We are now equipped to state the asymptotic properties of the LSU estimator.
-We are now equipped to state the main result of this paper.
+We are now equipped to state the asymptotic properties of the LSU estimator.
 
@@
 
-\begin{theorem} \label{thm:asymp_proof}  (Strong consistency) 
+\begin{theorem} \label{thm:asymp_proof} 
 Let $\widehat{\mubf}$ be defined by~\eqref{eq:hatmubfLSUdefn} and put $\widehat{\lambdabf}_N = \dealias(\tilde{\mubf} - \widehat{\mubf})$.
 Denote the elements of $\widehat{\lambdabf}_N$ by $\widehat{\lambda}_{0,N}, \dots, \widehat{\lambda}_{m,N}$.
-Suppose $\Phi_1, \dots, \Phi_N$ are independent and identically distributed with zero intrinsic mean and intrinsic variance $\sigma^2$, then $N^k \widehat{\lambda}_{k,N}$ converges almost surely to $0$ as $N\rightarrow\infty$ for all $k = 0, 1, \dots, m$.
+Suppose $\Phi_1, \dots, \Phi_N$ are independent and identically distributed with zero intrinsic mean, intrinsic variance $\sigma^2$, and pdf $f$, then:
+\begin{enumerate}
+\item (Strong consistency) $N^k \widehat{\lambda}_{k,N}$ converges almost surely to $0$ as $N\rightarrow\infty$ for all $k = 0, 1, \dots, m$.
+\item (Asymptotic normality) If $f(\fracpart{x})$ is continuous at $x = -\nicefrac{1}{2}$ and $f(-\nicefrac{1}{2}) < 1$ then the distribution of
+\[
+\left[
+\begin{array}
+[c]{cccc}%
+\sqrt{N} \widehat{\lambda}_{0,N} & N \sqrt{N}\widehat{\lambda}_{1,N}  & \dots & N^m\sqrt{N} \widehat{\lambda}_{m,N}
+\end{array}
+\right]^\prime
+\]
+converges to the normal with zero mean and covariance
+\[
+\frac{\sigma^2}{\left(1-f( -\nicefrac{1}{2}) \right)^{2}} \Cbf^{-1},
+\]
+where $\Cbf$ is the $(m+1)\times (m+1)$ Hilbert matrix with elements $C_{ik} = 1/(i + k + 1)$ for $i,k = \{0, 1, \dots, m\}$.
+\end{enumerate}
 \end{theorem}
-
-A proof of this theorem is given in the next section.
-Proofs for the case when $m=0$ was given in~\cite{McKilliam_mean_dir_est_sq_arc_length2010} and for the case when $m=1$ was given in~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.
-The proof here takes a similar approach, but requires new techniques.
-%Before giving the proof we make the following remarks.
+The proof of Theorem~\ref{thm:asymp_proof} is contained within the next two sections.
+Section~\ref{sec:strongconstproof} proves strong consistency and Section~\ref{sec:centlimitproof} proves asymptotic normality.
+The proofs for the case of $m=0$ were given in~\cite{McKilliam_mean_dir_est_sq_arc_length2010} and for the case when $m=1$ were given in~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.
+The proofs here take a similar approach, but require new techniques.
+Before giving the proof we make the following remarks.
+
 The theorem gives conditions on the \emph{dealiased} difference $\dealias(\tilde{\mubf} - \widehat{\mubf})$ between the true coefficients $\tilde{\mubf}$ and the estimated coefficients $\widehat{\mubf}$ rather than directly on the difference $\tilde{\mubf} - \widehat{\mubf}$.
@@
 In the interest of simplicity we only consider the case when $\Phi_1, \dots, \Phi_N$ are identically distributed here.
-If $X_n$ is circularly symmetric with density function nonincreasing with magnitude $\abs{X_n}$, then, the corresponding $\Phi_n$ necessarily has zero intrinsic mean~\cite[Theorem~5.2,~page~78]{McKilliam2010thesis}.
+If $X_n$ is circularly symmetric with density function nonincreasing with magnitude $\abs{X_n}$, then, the corresponding $\Phi_n$ necessarily has zero intrinsic mean~\cite[Theorem 5.2, page 78]{McKilliam2010thesis}.
 Thus, our theorem covers commonly used distributions for $X_1, \dots, X_N$, such as the normal distribution.
 
-Although we will not prove it here the assumption that $\Phi_1,\dots,\Phi_N$ have zero intrinsic mean is not only sufficient, but also necessary, for if $\Phi_1,\dots,\Phi_N$ have intrinsic mean $x \in [-\nicefrac{1}{2},\nicefrac{1}{2})$ with $x \neq 0$ then $\sfracpart{\widehat{\lambda}_{0,N} - x}\rightarrow 0$ almost surely as $N\rightarrow\infty$, and so $\widehat{\lambda}_{0,N}$ does not converge to zero.
-On the other hand if $\Phi_1,\dots,\Phi_N$ do not have an intrinsic mean then $\widehat{\lambda}_{0,N}$ will not converge.
-
-%The proof of asymptotic normality places requirements on the pdf $f$ of the phase noise.
+The proof of asymptotic normality places requirements on the pdf $f$ of the phase noise.
 The requirement that $\Phi_1, \dots, \Phi_N$ have zero intrinsic mean implies that $f(-\nicefrac{1}{2}) \leq 1$~\cite[Lemma~1]{McKilliam_mean_dir_est_sq_arc_length2010}, so the only case not handled is when equality holds, i.e., when $f(-\nicefrac{1}{2}) = 1$ or when $f(\fracpart{x})$ is discontinuous at $x = -\nicefrac{1}{2}$.
@@
 This type of result has been called a \emph{uniform law of large numbers} and follows from standard techniques~\cite{Pollard_conv_stat_proc_1984}.
-We give a full proof of~\eqref{eq:SNVNunifmlln} in Appendix~\ref{app:uniform-law-large}.
+A full proof of~\eqref{eq:SNVNunifmlln} in given in BLERG (see arxiv).
 We now concentrate attention on the minimiser of $V_N$.
@@
 
+\section{Proof of asymptotic normality}\label{sec:centlimitproof}
+
+Let $\psibf$ be the vector with $k$th component $\psi_k = N^k \lambda_k$, $k=0, \dots, m$ and let 
+\[
+T_{N}(\psibf) = S_N(\lambdabf) = \frac{1}{N} \sum_{n=1}^{N} \fracpart{ \Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \psi_k }^2.
+\]
+Let $\widehat{\psibf}_N$ be the vector with elements $\widehat{\psi}_{k,N} = N^k \widehat{\lambda}_{k,N}$ so that $\widehat{\psibf}_N$ is the minimiser of $T_N$.
+Because each of $N^k \widehat{\lambda}_{k,N}$ converges almost surely to zero as $N \rightarrow \infty$~\cite[Theorem~2]{McKilliam_pps1_2012}, it follows that $\widehat{\psibf}_{N}$ converges almost surely to $\zerobf$ as $N \rightarrow \infty$.
+We want to find the asymptotic distribution of
+\[
+\sqrt{N}\widehat{\psibf}_{N} = 
+\left[
+\begin{array}
+[c]{c}%
+\sqrt{N} \widehat{\psi}_{0,N} \\ \sqrt{N}\widehat{\psi}_{1,N}  \\ \vdots \\ \sqrt{N} \widehat{\psi}_{m,N}
+\end{array}
+\right]
+=
+\left[
+\begin{array}
+[c]{c}%
+\sqrt{N} \widehat{\lambda}_{0,N} \\ N\sqrt{N}\widehat{\lambda}_{1,N} \\ \vdots \\ N^m\sqrt{N} \widehat{\lambda}_{m,N}
+\end{array}
+\right].
+\]
+The proof is complicated by the fact that $T_N$ is not differentiable everywhere as $\fracpart{x}^2$ is not differentiable when $\fracpart{x} = \tfrac{1}{2}$.
+This precludes the use of ``standard approaches'' to proving asymptotic normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.
+However, we show in Lemma~\ref{lem:diffathatpsi} that all the partial derivatives $\frac{\partial T_N}{\partial \psi_\ell}$ for $\ell = 0, \dots, m$ exist, and are equal to zero, at the minimiser $\widehat{\psibf}_N$.
+Thus, putting
+\begin{equation}\label{eq:Wn}
+W_{n} = \round{\Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \widehat{\psi}_{k,N}},
+\end{equation}
+we have, for each $\ell = 0, \dots, m$,
+\begin{align*}
+0 & = \frac{\partial T_N}{\partial \psi_\ell}(\widehat{\psibf}_N)\\
+&= \frac{2}{N}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\left( \Phi_n - W_{n}+ \sum_{k=0}^{m}(\tfrac{n}{N})^k \widehat{\psi}_{k,N}  \right),
+\end{align*}
+so that
+\[
+D_{\ell,N} = K_{\ell,N},
+\]
+where
+\[
+D_{\ell,N} = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}(\tfrac{n}{N})^\ell \Phi_n,
+\]
+and
+\begin{equation}\label{eq:KellN}
+K_{\ell,N} = \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\left(W_{n}- \sum_{k=0}^{m}(\tfrac{n}{N})^k \widehat{\psi}_{k,N}  \right).
+\end{equation}
+Lemma~\ref{lem:Kconvfhalf} shows that, for all $\ell = 0, \dots m$,
+\begin{equation}\label{eq:KellNconv}
+K_{\ell,N} =  (h - 1) \sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big( C_{\ell k} + o_P(1) \big) + o_P(1)
+\end{equation}
+where $C_{\ell k} =  \tfrac{1}{\ell + k + 1}$, and $h = f(-\nicefrac{1}{2})$, and $o_P(1)$ denotes a random variable converging in probability to zero as $N\rightarrow\infty$.
+
+It is now convenient to write in vector form.  Let 
+\begin{equation}\label{eq:dnandkn}
+\kbf_N = 
+\left[
+\begin{array}
+[c]{c}%
+K_{0,N}  \\ \vdots \\ K_{m,N}
+\end{array}
+\right]
+\qquad \text{and} \qquad 
+\dbf_N = \left[
+\begin{array}
+[c]{c}%
+D_{0,N}  \\ \vdots \\ D_{m,N}
+\end{array}
+\right].
+\end{equation}
+From~\eqref{eq:KellNconv},
+\[
+\dbf_N = \kbf_N = \sqrt{N} (h - 1)(\Cbf + o_P(1)) \widehat{\psibf}_N + o_P(1)
+\]
+where $o_P(1)$ here means a vector or matrix of the appropriate dimension with every element converging in probability to zero as $N\rightarrow\infty$.
+Thus $\sqrt{N}\widehat{\psibf}_N$ has the same asymptotic distribution as $(h - 1)^{-1}\Cbf^{-1} \dbf_N$.
+Lemma~\ref{eq:convdn} shows that $\dbf_N$ is asymptotically normally distributed with zero mean and covariance matrix $\sigma^2\Cbf$.
+Hence $\sqrt{N}\widehat{\psibf}_N$ is asymptotically normal with zero mean and covariance matrix
+\[
+\frac{\sigma^2\Cbf^{-1}\Cbf(\Cbf^{-1})^\prime}{(1 - h)^2} = 
+\frac{\sigma^2\Cbf^{-1}}{(1 - h)^2}.
+\] 
+It remains to prove Lemmas~\ref{lem:diffathatpsi},~\ref{lem:Kconvfhalf} and~\ref{eq:convdn}.
+
+\begin{lemma}\label{lem:diffathatpsi}
+For all $\ell = 0, \dots, m$ the partial derivatives $\frac{\partial T_N}{\partial \psi_\ell}$ exist, and are equal to zero, at the minimiser $\widehat{\psibf}_N$.  That is,
+\[
+\frac{\partial T_N}{\partial \psi_\ell}(\widehat{\psibf}_N) = 0 \qquad \text{for each $\ell = 0, \dots, m$.}
+\]
+\end{lemma}
+\begin{IEEEproof}
+The function $\fracpart{x}^2$ is differentiable everywhere except if $\fracpart{x} \neq -\nicefrac{1}{2}$.  Recalling that
+\[
+T_{N}(\psibf) = \frac{1}{N} \sum_{n=1}^{N} \fracpart{ \Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \psi_k }^2
+\]
+we see that $T_N$ will be differentiable with respect to $\psibf$ at $\widehat{\psibf}_N$ if 
+\[
+\fracpart{\Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \widehat{\psi}_{k,N}} \neq -\nicefrac{1}{2} \qquad \text{for all $n = 1, \dots, N$.}
+\] 
+This is proved in Lemma~\ref{lem:boundonhatlambda}.
+So the partial derivatives $\frac{\partial T_N}{\partial \psi_\ell}$ exist for all $\ell = 0, \dots, m$ at $\widehat{\psibf}_N$.
+That each of the partial derivatives is equal to zero at $\widehat{\psibf}_N$ follows immediately from the fact that $\widehat{\psibf}_N$ is a minimiser of $T_N$.
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:boundonhatlambda} $\sabs{\sfracpart{\Phi_n + \sum_{k=0}^m (\nicefrac{n}{N})^k \widehat{\psi}_{k,N}}} \leq \frac{1}{2} - \frac{1}{2N}$ for all $n = 1, \dots, N$.
+\end{lemma}
+\begin{IEEEproof}
+To simplify our notation let 
+\[
+B_n = \Phi_n + \sum_{k=1}^m (\nicefrac{n}{N})^k \widehat{\psi}_{k,N}
+\]
+so that we now require to prove $\sabs{\sfracpart{B_n + \widehat{\psi}_{0,N}}} \leq \frac{1}{2} - \frac{1}{2N}$
+for all $n = 1, \dots, N$.  From~\eqref{eq:Wn}, $W_n = \sround{B_n + \widehat{\psi}_{0,N}}$, and 
+\begin{align*}
+T_N(\widehat{\psibf}_N) &= \frac{1}{N}\sum_{n=1}^N \sfracpart{B_n + \widehat{\psi}_{0,N}}^2
+\\ &= \frac{1}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N} - W_n)^2.
+\end{align*}
+Since $\widehat{\psi}_{0,N}$ is the minimiser of the quadratic above,
+\begin{equation}\label{eq:lamsum}
+\widehat{\psi}_{0,N} = -\frac{1}{N}\sum_{n=1}^N(B_n - W_n).
+\end{equation}
+The proof now proceeds by contradiction.  Assume that for some $k$,
+\begin{equation}\label{eq:philamfraccontra}
+\sfracpart{B_k + \widehat{\psi}_{0,N}} > \frac{1}{2} - \frac{1}{2N}.
+\end{equation}
+Let $F_n = W_n$ for all $n \neq k$ and $F_k = W_k + 1$, and let
+\[
+\phi = -\frac{1}{N}\sum_{n=1}^N(B_n - F_n) = \widehat{\psi}_{0,N} + \frac{1}{N}.
+\]
+Then,
+\begin{align}
+(B_k + &\phi - F_k)^2 = (B_k + \phi - W_k - 1)^2 \nonumber \\
+&= (B_k + \phi - W_k)^2 - 2(B_k + \phi - W_k) + 1 \nonumber \\
+&= (B_k + \phi - W_k)^2 - 2(B_k + \widehat{\psi}_{0,N} - W_k) + 1 - \frac{2}{N} \nonumber \\
+&= (B_k + \phi - W_k)^2 - 2\sfracpart{B_k + \widehat{\psi}_{0,N}} + 1 - \frac{2}{N} \nonumber \\
+&< (B_k + \phi - W_k)^2 - \frac{1}{N}, \label{eq:Bkineq}
+\end{align}
+where the inequality in the last line follows from~\eqref{eq:philamfraccontra}. Let
+\[
+\bbf = \left[
+\begin{array}
+[c]{cccc}%
+\phi & \widehat{\psi}_{1,N} & \cdots & \widehat{\psi}_{m,N}
+\end{array}
+\right]^\prime
+\]
+be the vector of length $m+1$ with components $b_0 = \phi$ and $b_\ell = \widehat{\psi}_{\ell,N}$ for $\ell = 1 , \dots m$.  Now,
+\[
+N T_N(\bbf) = \sum_{n=1}^N \fracpart{B_n + \phi }^2 \leq  \sum_{n=1}^N (B_n + \phi  - F_n)^2, 
+\]
+and using the inequality from~\eqref{eq:Bkineq},
+\begin{align*}
+N &T_N(\bbf) < - \frac{1}{N} + \sum_{n=1}^N (B_n + \phi  - W_n)^2 \\
+&= - \frac{1}{N} + \sum_{n=1}^N (B_n + \widehat{\psi}_{0,N} + \frac{1}{N}  - W_n)^2 \\
+&= \sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n)^2 +  \frac{2}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n)\\
+%&= N T_N(\widehat{\psibf}_N) + \frac{2}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n) \\
+&= N T_N(\widehat{\psibf}_N),
+\end{align*}
+because $\frac{2}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n) = 0$ as a result of~\eqref{eq:lamsum}.
+But, now $T_N(\bbf) < T_N(\widehat{\psibf}_N)$ violating the fact that $\widehat{\psibf}_N$ is a minimiser of $T_N$.
+So~\eqref{eq:philamfraccontra} is false by contradiction.
+
+If $\sfracpart{B_k + \widehat{\psi}_{0,N}} < -\frac{1}{2} + \frac{1}{2N}$ for some $k$, we set $F_k = W_k - 1$ and using the same procedure as before obtain $T_N(\bbf) < T_N(\widehat{\psibf}_N)$ again.
+The proof follows.
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:Kconvfhalf}
+With $K_{\ell,N}$ defined as in~\eqref{eq:KellN} and $h = f(-\nicefrac{1}{2})$,
+\[
+K_{\ell,N} = (h - 1) \sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big( C_{\ell k} + o_P(1) \big) + o_P(1) 
+\] 
+for all $\ell = 0, \dots, m$, where $C_{\ell k} =  \frac{1}{\ell + k + 1}$.% is an element of the $m+1$ by $m+1$ Hilbert matrix, and $o_P(1)$ denotes a random variable converging in probability to zero as $N\rightarrow\infty$.
+\end{lemma}
+\begin{IEEEproof}
+Care must be taken since $\widehat{\psibf}_N$ depends on the sequence $\{ \Phi_n \}$.  For $n = 1, \dots, N$ and positive $N$, let
+\begin{equation}
+p_{nN}(\psibf) = \sum_{k=0}^{m}\left( \tfrac{n}{N}\right)^k \psi_k,
+\end{equation}
+and put
+\[
+q_{n}(x) = \round{\Phi_n + x}, \qquad Q(x) = \expect q_{n}(x) =  \expect q_{1}(x).
+\]
+Let
+\begin{equation}\label{eq:GNdef}
+G_N(\psibf) = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}\left( \tfrac{n}{N}\right)^\ell  \big(q_n(p_{nN}(\psibf)) - Q(p_{nN}(\psibf)) \big).
+\end{equation}
+and put
+\begin{equation}\label{eq:hatpnN}
+\widehat{p}_{nN} = p_{nN}(\widehat{\psibf}_N) = \sum_{k=0}^m \left(\tfrac{n}{N}\right)^k \widehat{\psi}_{k,N}.
+\end{equation}
+Observe that $G_N$ depends on $\ell$ and we could write $G_{\ell,N}$ but have suppressed the subscript $\ell$ for notational simplicity.
+Now $W_n$ from~\eqref{eq:Wn} can be written as $W_n = \round{\Phi_n + \widehat{p}_{nN}} = q_{n}(\widehat{p}_{nN})$
+and $K_{\ell,N}$ from~\eqref{eq:KellN} can be written as
+\begin{align*}
+K_{\ell,N} &= \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\big( q_{n}(\widehat{p}_{nN}) - \widehat{p}_{nN}  \big) \\
+&= \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\big(q_{n}(\widehat{p}_{nN}) - \widehat{p}_{nN} + Q(\widehat{p}_{nN}) - Q(\widehat{p}_{nN}) \big) \\
+&= G_N(\widehat{\psibf}_N) + H_{\ell,N},
+\end{align*}
+where 
+\begin{equation}\label{eq:HellNdef}
+H_{\ell,N} = \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell \big( Q(\widehat{p}_{nN}) - \widehat{p}_{nN} \big).
+\end{equation}
+BLERG (empirical process, tighttess, ref arxiv) that for any $\delta >0$ and $\nu > 0$ there exists an $\epsilon > 0$ such that
+\[
+\prob\left\{ \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf) } > \delta   \right\} < \nu
+\]
+for all positive integers $N$, where $\|\psibf\|_\infty = \sup_{k} \abs{\psi_k}$.  Since $\widehat{\psibf}_N$ converges almost surely to zero, it follows that for any $\epsilon > 0$,
+\[
+\lim_{N\rightarrow\infty}\prob\left\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} = 0
+\] 
+and therefore $\prob\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \} < \nu$ for all sufficiently large $N$.  Now
+\begin{align*}
+  \prob&\left\{\abs{ G_N(\widehat{\psibf}_N) } > \delta \right\} \\
+&= \prob\left\{ \abs{G_N(\widehat{\psibf}_N)} > \delta \;, \; \|\widehat{\psibf}_N\|_{\infty} < \epsilon \right\} \\
+&\hspace{0.7cm} + \prob\left\{ \abs{G_N(\widehat{\psibf}_N)} > \delta  \;, \; \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} \\
+&\leq \prob\left\{  \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf) } > \delta \right\} + \prob\left\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} \\
+&\leq 2\nu
+\end{align*}
+for all sufficiently large $N$.
+Since $\nu$ and $\delta$ can be chosen arbitrarily small, it follows that $G_N(\widehat{\psibf}_N)$ converges in probability to zero as $N\rightarrow\infty$, and therefore $K_{\ell,N} = H_{\ell,N} + o_P(1)$.
+Lemma~\ref{lem:sumpNhilb} shows that
+\begin{align*}
+H_{\ell,N} =  (h-1)\sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big(C_{\ell k} + o_P(1)\big).
+\end{align*}
+The proof follows.
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:sumpNhilb}
+With $H_{\ell,N}$ defined in~\eqref{eq:HellNdef} and $\widehat{p}_{nN}$ defined in~\eqref{eq:hatpnN}, and with $h = f(-\nicefrac{1}{2})$,
+\[
+H_{\ell,N} = (h-1)\sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big(C_{\ell k} + o_P(1)\big),
+\]
+where $C_{\ell k} = \tfrac{1}{\ell + k + 1}$.
+\end{lemma}
+\begin{IEEEproof}
+If $\abs{x} < 1$, then
+\[
+q_n(x)  = \round{\Phi_n + x} = \begin{cases}
+1, & \Phi_n + x  \geq \nicefrac{1}{2} \\
+-1, & \Phi_n + x  < -\nicefrac{1}{2} \\
+0, & \text{otherwise},
+\end{cases}
+\]
+and,
+\begin{align*}
+Q(x) = Eq_1(x) &=  \begin{cases}
+\int_{\nicefrac{1}{2} -x }^{\nicefrac{1}{2}}f(t)\,dt, &   x \geq 0 \\
+-\int_{-\nicefrac{1}{2}}^{-\nicefrac{1}{2} - x }f(t)\,dt, &  x  < 0.
+\end{cases} 
+% \\
+% &= \begin{cases}
+% 1 - F(\nicefrac{1}{2} - x ), & x  \geq 0 \\
+% -F(-\nicefrac{1}{2} - x ), &  x < 0,
+% \end{cases} 
+\end{align*}
+Because $f(\fracpart{x})$ is continuous at $-\nicefrac{1}{2}$ it follows that 
+\[
+Q(x) = x \big( h + \zeta(x) \big),
+\] 
+where $\zeta(x)$ is a function that converges to zero as $x$ converges to zero.  
+%The equation above is still valid when $\abs{x} \geq 1$ by an appropriate choice of the function $\zeta(x)$.  
+Observe that $\abs{\widehat{p}_{nN}} \leq \sum_{k=0}^m\sabs{ \widehat{\psi}_{k,N}}$ and, since each of the $\widehat{\psi}_{k,N} \rightarrow 0$ almost surely as $N\rightarrow\infty$, it follows that $\widehat{p}_{nN} \rightarrow 0$ almost surely uniformly in $n = 1, \dots, N$ as $N\rightarrow\infty$.
+Thus $\zeta(\widehat{p}_{nN}) \rightarrow 0$ almost surely (and therefore also in probability) uniformly in $n = 1, \dots, N$ as $N\rightarrow\infty$.
+Now,
+\begin{align*}
+Q(\widehat{p}_{nN}) - \widehat{p}_{nN} &= \widehat{p}_{nN}\big( h - 1 + \zeta(\widehat{p}_{nN}) \big) \\
+&= \widehat{p}_{nN}\big( h - 1 + o_P(1) \big), 
+\end{align*}
+and so,
+\begin{align*}
+H_{\ell,N} &= \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell \widehat{p}_{nN}\big( h - 1 + o_P(1) \big)  \\
+&=  \frac{1}{\sqrt{N}} \sum_{n=1}^{N}(\tfrac{n}{N})^\ell \sum_{k=0}^m (\tfrac{n}{N})^k \widehat{\psi}_{k,N} \big( h - 1 + o_P(1) \big)   \\
+&= \sqrt{N}\sum_{k=0}^m \widehat{\psi}_{k,N} \frac{1}{N} \sum_{n=1}^{N} \frac{n^{\ell + k}}{N^{\ell + k + 1}}\big( h - 1 + o_P(1)  \big).
+\end{align*}
+The Riemann sum
+\[
+\frac{1}{N} \sum_{n=1}^{N} \frac{n^{\ell + k}}{N^{\ell + k + 1}} = \int_{0}^{1}x^{k+\ell+1}dx + o_P(1),
+\]
+and since the integral above evaluates to $\frac{1}{k+\ell+1}$, we have
+\[
+H_{\ell,N} = (h - 1)\sqrt{N}\sum_{k=0}^m \widehat{\psi}_{k,N}\left( \frac{1}{k+\ell+1} + o_P(1)  \right)
+\]
+as required.
+% Put
+% \[
+% C_{\ell + k,N} = \sum_{n=1}^{N}\frac{n^{\ell + k}}{N^{\ell + k+1}}.
+% \]
+% It is well known (see for example~\cite[eq. 21]{Peleg1991_CRB_PPS_1991}) that,
+% \begin{equation}\label{eq:ClkNHilbert}
+% C_{\ell + k,N} = C_{\ell k} + O(N^{-1}).
+% \end{equation}
+% Since $\zeta(\widehat{p}_{nN}) = o_P(1)$ for all $n = 1, \dots, N$ then
+% \[
+% \sum_{n=1}^{N}\frac{n^{\ell + k}}{N^{\ell + k + 1}}\zeta(\widehat{p}_{nN}) 
+% \]
+% converges in probability to zero as $N\rightarrow\infty$.  So,
+% \begin{align*}
+% \sum_{n=1}^{N}\frac{n^{\ell + k}}{N^{\ell + k + 1}}\big( h - 1 +  \zeta(\widehat{p}_{nN}) \big) = (h-1)\big( C_{\ell k} + o_P(1) \big).
+% \end{align*}
+% The proof follows by substituting this into~\eqref{eq:HellNre}.
+\end{IEEEproof}
+
+\begin{lemma}\label{eq:convdn}
+The distribution of the vector $\dbf_N$, defined in~\eqref{eq:dnandkn}, converges to the multivariate normal with zero mean and covariance matrix $\sigma^2\Cbf$.
+\end{lemma}
+\begin{IEEEproof}
+For any constant vector $\alphabf$, let
+\[
+z_N = \alphabf^\prime \dbf_N = \frac{1}{\sqrt{N}} \sum_{n=1}^{N} \Phi_n \sum_{\ell=0}^{m}\alpha_\ell\left( \frac{n}{N} \right)^\ell.
+\]
+Observe that $\Phi_n$ has zero mean and variance $\sigma^2$ as a result of Proposition~\ref{prop:zerointmeanzeromean}.
+By Lypanov's central limit theorem $\zbf_N$ is asymptotically normally distributed with zero mean and variance
+\[
+\lim_{N\rightarrow\infty} \sigma^2 \frac{1}{N} \sum_{n=1}^{N}\left( \sum_{\ell=0}^{m}\alpha_\ell\left( \frac{n}{N} \right)^\ell \right)^2 = \sigma^2 \alphabf^\prime \Cbf \alphabf.
+\]
+By the Cram\'{e}r-Wold theorem it follows that $\dbf_N$ is asymptotically normally distributed with zero mean and covariance $\sigma^2 \Cbf$. 
+% Recall that the $\ell$th element of $\dbf_N$ is
+% \[
+% D_{\ell,N} = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}(\tfrac{n}{N})^\ell \Phi_n.
+% \]
+% Since the $\Phi_1, \dots, \Phi_N$ are independent, each with variance $\sigma^2$, the distribution of $D_{\ell,N}$ converges to the normal with zero mean and variance
+% \[
+% \lim_{N\rightarrow\infty}\var D_{\ell,N} = \sigma^2 \lim_{N\rightarrow\infty} C_{2\ell,N} = \sigma^2 C_{\ell\ell}
+% \]
+% by Lypanov's central limit theorem and by using~\eqref{eq:ClkNHilbert}.  The covariance between $D_{\ell,N}$ and $D_{k,N}$ is
+% \[
+% \covar(D_{\ell,N}, D_{k,N}) = \sigma^2 C_{\ell+k,N} \rightarrow \sigma^2 C_{\ell k}
+% \]
+% as $N \rightarrow \infty$, again using~\eqref{eq:ClkNHilbert}.  The proof follows.
+\end{IEEEproof}
+
+\section{Simulations}\label{sec:simulations}
+
+% This section describes the results of Monte-Carlo simulations involving the least squares unwrapping estimator (LSU), the least squares estimator (LSE), Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, and the discrete polynomial phase transform (DPT)~\cite{Peleg_DPT_1995}.
+In all simulations the sample sizes considered are $N = 10, 50$ and $200$ and the unknown amplitude is $\rho = 1$.
+The number of replications of each experiment is 2000 and the $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ are distributed uniformly randomly over the identifiable region.
+
+% Figure~\ref{plot:lsetest} shows the performance of the LSU estimator and LSE when $m=2$ (linear chirp).
+Only results for the highest order parameter $\widehat{\mu}_2$ are displayed, the results for the lower order parameters leading to similar conclusions.
+The LSE is computed using the approach described in~\cite[Sec.~10.1]{McKilliam2010thesis} which is based on that of Abatzoglou~\cite{Abatzoglou_ml_chirp_1986}.
+The LSE is computationally intensive.
+We were able to compute it within a reasonable amount of time only with $N$ less than $50$ when $m=2$.
+Results for the exact LSU estimator computed using the sphere decoder and the approximate LSU estimator computed using Babai's algorithm are displayed.
+The sphere decoder is computationally intractable when $N = 200$.
+It is evident that a significant performance penalty is suffered by using Babai's algorithm when the noise variance is large.
+We have also displayed results using another approximate nearest point algorithm called the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder}.
+Our implementation requires $O(N^3\log{N})$ operations.
+This is more computationally expensive than Babai's algorithm but, as can be seen, the $K$-best method is more statistically accurate.
+
+% For the simulations displayed in the left hand plot of Figure~\ref{plot:lsetest} the signal model used is that from~\eqref{eq:Y_nsamplednoisey}.
+The noise terms $X_1, \dots, X_N$ are independent and identically distributed zero mean complex Gaussian random variables with independent real and imaginary parts having variance $\sigma_c^2$.
+In this case the LSE is also the maximum likelihood estimator.
+The \term{Cram\`{e}r-Rao lower bound} (CRB) for the variance of unbiased estimators of $\tilde{\mu}_j$ is given by Peleg and Porat~\cite{Peleg1991_CRB_PPS_1991} to be
+% \[ 
+% \frac{\sigma_c^2}{4\pi^2 N^{2j + 1}} C^{-1}_{jj} + o(N^{-2j-1})
+% \]
+% where $C^{-1}_{jj}$ is the $j$th diagonal element of the inverse of the $m+1$ by $m+1$ Hilbert matrix $\Cbf$.
+Observe that the performance of the exact LSU estimator (computed using the sphere decoder) is close to the maximum likelihood LSE.
+The asymptotic Cramer-Rao bound (CRB) (solid line) and the asymptotic variance predicted in Theorem~\ref{thm:asymp_proof} (dashed line) are also shown.
+Provided the noise variance is small enough (so that the `threshold' is avoided) the LSE variance is close to the CRB (as is expected) and the LSU variance is close to that predicted by Theorem~\ref{thm:asymp_proof}.
+% %Note the gap between the CRB and the variance predicted by Theorem~\ref{thm:asymp_proof}.
+In practice this gap can be overcome using a numerical technique, such as Newton's method, with the LSU estimate as a starting point.
+We have not done this here in order to show the accuracy of Theorem~\ref{thm:asymp_proof}.
+The other noticeable gap is that between the threshold behaviour of the LSU estimator and the LSE.
+When $N=50$ the threshold \emph{kicks in} a little earlier for the LSE than it does for the LSU.
+We suspect that this gap will actually \emph{close} as $N$ increases.
+This hypothesis is supported by the fact that this behaviour is observed in the case of frequency estimation (when $m=1$)~\cite[Sec.
+9.5]{McKilliam2010thesis}\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.
+Unfortunately it is not feasible to test this hypothesis because of the computational complexity of the LSE.
+% For the simulations displayed in the right hand plot of Figure~\ref{plot:lsetest} the signal model used is that from~\eqref{eq:noise_circ_poly}.
+The phase noise terms $\Phi_1, \dots, \Phi_N$ are uniform random variables with variance $\sigma^2$.
+In this case, the LSU estimator performs slightly better than the LSE.
+This result agrees with similar observations made in circular statistics~\cite{McKilliam_mean_dir_est_sq_arc_length2010}.
+The asymptotic variance predicted by Theorem~\ref{thm:asymp_proof} again accurately predicts the performance of the LSU estimator.
+
+% Figure~\ref{plot:polyest} displays the performance of the LSU estimator, the DPT and Kitchen's estimator when $m=3$.
+Both the DPT and Kitchen's estimator perform poorly because they do not work correctly for some parameters in the identifiable region $B$.
+This behaviour has been observed previously~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase,McKilliam_polyphase_est_icassp_2011}.
+Again the performance of the LSU is well predicted by Theorem~\ref{thm:asymp_proof} provided that the noise variance is small enough to avoid the threshold effect and provided that $N$ is sufficiently large.
+The asymptotic CRB is also plotted.
+
+This section describes the results of Monte-Carlo simulations with the least squares unwrapping (LSU) estimator, Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, and the discrete polynomial phase transform (DPT) estimator of Peleg and Porat~\cite{Peleg_DPT_1995}.
+The sample sizes considered are $N = 10, 50, 200$ and the unknown amplitude is $\rho = 1$.
+The $X_1, \dots, X_N$ are pseudorandomly generated independent and identically distributed circularly symmetric complex Gaussian random variables with variance $\sigma_c^2$.
+The coefficients $\tilde{\mubf} = [\tilde{\mu}_0, \dots, \tilde{\mu}_m]$ are distributed uniformly randomly in the identifiable region $B$~\cite[Sec.~3]{McKilliam_pps1_2012}.
+The number of replications of each experiment is $T = 2000$ to obtain estimates $\widehat{\mubf}_1, \dots, \widehat{\mubf}_T$ and the corresponding dealiased errors $\widehat{\lambdabf}_t = \dealias(\widehat{\mubf}_T - \tilde{\mubf})$ are computed.
+The sample mean square error (MSE) of the $k$th coefficient is computed according to $\tfrac{1}{T}\sum_{t=1}^T \widehat{\lambda}_{k,t}^2$ where $\widehat{\lambda}_{k,t}$ is the $k$th element of $\widehat{\lambdabf}_t$.
+
+Figure~\ref{plot:polyest} shows the sample MSEs obtained for a polynomial phase signal of order $m=3$.
+%Results are displayed results for the zeroth and third order coefficients $\widehat{\mu}_1$ and $\widehat{\mu}_3$.
+The results for $\widehat{\mu}_1$ and $\widehat{\mu}_2$ lead to similar conclusions.
+The LSU estimator can be computed by finding a nearest lattice point in a particular lattice~\cite{McKilliam2010thesis,McKilliam2009asilomar_polyest_lattice}.
+When $N = 10$ and $50$ the LSU estimator can be computed exactly using a general purpose algorithm for finding nearest lattice points called the sphere~decoder~\cite{Pohst_sphere_decoder_1981,Agrell2002,Viterbo_sphere_decoder_1999}.
+This is displayed by the circles in the figures.
+When $N=200$ the sphere decoder is computationally intractable and we instead use an approximate nearest point algorithm called the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder}.
+This is displayed by the dots.
+For the purpose of comparison we have also plotted the results for the $K$-best method when $N = 10$ and $50$.
+The asymptotic variance predicted in Theorem~\ref{thm:asymp_proof} is displayed by the dashed line.
+Provided the noise variance is small enough (so that the `threshold' is avoided) the sample MSE of the LSU estimator is close to that predicted by Theorem~\ref{thm:asymp_proof}.
+The Cram\'{e}r-Rao lower bound for the variance of unbiased polynomial phase estimators in Gaussian noise is also plotted using the solid line~\cite{Peleg1991_CRB_PPS_1991}.
+When the noise variance is small the asymptotic variance of the LSU estimator is close to the Cram\'{e}r-Rao lower bound.
+
+Kitchen's unwrapping estimator and the DPT estimator perform very poorly in Figure~\ref{plot:polyest}.
+The reason is that both estimators do not work correctly for all parameters in the identifiable region $B$.
+For this reason Figure~\ref{plot:polyest2} plots the sample MSEs for the estimators with the true coefficients fixed to $\tilde{\mubf} = [0.1,0.2,10^{-3},10^{-5}]$.
+This is within the range suitable for the DPT estimator~\cite{Peleg_DPT_1995}.
+Both Kitchen's unwrapping estimator and the DPT estimator perform better in this case.
+Only results for the zeroth and third order parameters $\tilde{\mu}_0$ and $\tilde{\mu}_3$ are displayed.
+The results for the other parameters are similar.
+
+\begin{figure*}[p] 
+   	\centering 
+  		\includegraphics{code/gaussianplot4-1.mps} 
+  		\caption{Mean square error (MSE) of polynomial phase estimators.
+The true coefficients $\tilde{\mubf}$ are uniformly distributed in the identifiable region $B$.
+(Top left) MSE in the phase coefficient $\mu_0$.
+(Top right) MSE in the frequency coefficient $\mu_1$.
+(Bottom left) MSE in the quadratic coefficient $\mu_2$.
+(Bottom right) MSE in the cubic coefficient $\mu_3$.}
+  		\label{plot:polyest} 
+ \end{figure*}
+
+\begin{figure}[p] 
+   	\centering 
+  		\includegraphics{code/gaussianplot4-2.mps} 
+  		\caption{Mean square error (MSE) of polynomial phase estimators.
+The true coefficients are fixed to $\tilde{\mubf} = [0.1,0.2,10^{-3},10^{-5}]$.
+(Top) MSE in the phase coefficient $\widehat{\mu}_0$.
+(Bottom right) MSE in the cubic coefficient $\widehat{\mu}_3$.}
+  		\label{plot:polyest2} 
+ \end{figure}
+
+\section{Computational considerations}
+
+When $m=0$, polynomial phase estimation is equivalent to the problem of \emph{direction estimation} from circular statistics, and the least squares estimator corresponds with the \emph{extrinsic mean} (or \emph{sample circular mean})~\cite{McKilliam_mean_dir_est_sq_arc_length2010,Fisher1993,Mardia_directional_statistics}.
+In this case the LSE can be computed efficiently in $O(N)$ by averaging $N$ complex numbers.
+When $m=1$, polynomial phase estimation is equivalent to single frequency estimation, and the least squares estimator can be computed efficiently in $O(N\log N)$ operations by maximising the \emph{periodogram} using the fast Fourier transform followed by an appropriate numerical optimisation technique~\cite{Quinn2008maximizing_the_periodogram,Rife1974,Quinn2001}.
+When $m=2$, the least squares estimator can be computed in a reasonable amount of time when $N$ is approximately less than $50$.
+The LSE is computationally intensive for any $N$ when $m > 2$.
+By comparison the exact LSU estimator implemented using the sphere decoder can be computed quickly for any $m$ with $N$ approximately less than $50$.
+In situations where $N$ is small, but high statistical accuracy is required, the sphere decoder is computationally a better choice than the LSE for polynomial phase signals of order greater than or equal to $2$.
+
+Babai's nearest plane algorithm requires $O(N^2)$ operations and can be run in a reasonable amount of time for any $m$ and reasonably large $N$.
+It should be noted that Babai's nearest plane algorithm requires a Lov\`as reduced lattice basis~\cite{Lenstra1982}.
+This requires $O(N^4)$ operations to compute, but, the Lov\`as reduced basis can be computed once offline.
+It does not need to be computed each time the estimator is run.
+The $K$-best algorithm runs in a reasonable amount of time for any $m$ and when $N$ is less than about $300$.
+If the noise variance is small, then both the sphere decoder and the $K$-best algorithm are actually fast.
+This is because both of these estimators begin with the lattice point found by Babai's nearest plane algorithm.
+If this point is close to (or \emph{is}) the nearest point then both the sphere decoder and the $K$-best algorithm terminate quickly.
+This is in contrast to the least squares estimator that is slow regardless of the noise variance.
+
+Both Kitchen's estimator and the DPT are fast to compute.
+Kitchen's estimator requires only $O(N)$ arithmetic operations and the DPT requires $O(N \log N)$ operations.
+However, these estimators are not as accurate as the LSE or the LSU.
+They also fail to operate correctly over the entire identifiable region of polynomial phase coefficients~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}~\cite[Ch.~10]{McKilliam2010thesis}.
+
 %\section{Discussion}
@@
 This paper has considered the estimation of the coefficients of a noisy polynomial phase signal by least squares phase unwrapping (LSU).
-We have derived conditions under which the polynomial phase estimation problem is identifiable.
-Under these conditions, and under some assumptions on the distribution of the noise, the LSU estimator is shown to be strongly consistent.
+It has been shown that the LSU estimator is strongly consistent and asymptotically normally distributed.
 %Simulations are used to show that the LSU has near maximum likelihood performance in the case that the noise is additive white and Gaussian.
-In the second paper in this series~\cite{McKilliam_pps2_2012} we show the LSU estimator to be asymptotically normally distributed and present the results of Monte Carlo simulations that support our asymptotic results.
+The LSU estimator can be computed by finding a nearest lattice point in a lattice~\cite{McKilliam2009asilomar_polyest_lattice}.
+%Here, we consider using the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder} to achieve this in reasonable time when the number of observations, $N$, is less than about 300.
+Polynomial time nearest point algorithms for these lattices exist~\cite[Sec 4.3]{McKilliam2010thesis}, but these algorithms are not fast in practice.
+Instead, we considered some general purpose algorithms, the \term{sphere decoder}, \term{Babai's nearest plane algorithm} and the \term{$K$-best algorithm}.
+The sphere decoder and $K$-best algorithm result in accurate estimators.
+
+The major outstanding question is whether faster nearest point algorithms exist for these specific lattices.
+Considering the excellent statistical performance (both theoretically and practically) of the LSU estimator, even fast \emph{approximate} nearest point algorithms are likely to prove useful for the estimation of polynomial phase signals.
 
@@
 \normalsize
-\appendix
-
-%\section{Tricks with fractional parts}
-%Throughout the paper, and particularly in  have made use of a number of results involving the fractional part function
-
-\subsection{A uniform law of large numbers} \label{app:uniform-law-large}
-
-During the proof of strong consistency we made use of the fact that
-\begin{equation}~\label{eq:SNVNunifmlln2}
-\sup_{\lambdabf \in B}\sabs{S_N(\lambdabf) - V_N(\lambdabf)} \rightarrow 0
-\end{equation}
-almost surely as $N\rightarrow\infty$, where $V_N(\lambdabf) = \expect S_N(\lambdabf)$.  We prove this result here.  Put
-\[
-D_N(\lambdabf) = S_N(\lambdabf) - V_N(\lambdabf).
-\]  
-Now, for any $\epsilon > 0$,
-\[
-\sum_{N=1}^\infty \prob \left\{ \sup_{\lambdabf \in B}\abs{ D_N(\lambdabf) } > \epsilon \right\} < \infty
-\]
-by Lemma~\ref{lem:vn}.
-So~\eqref{eq:SNVNunifmlln2} follows from the Borel-Cantelli lemma.
-In what follows we use order notation in the standard way, that is, for functions $h$ and $g$, we write $h(N) = O(g(N))$ to mean that there exists a constant $K > 0$ and a finite $N_0$ such that $h(N) \leq K g(N)$ for all $N > N_0$.
-
-\begin{lemma} \label{lem:vn} 
-For any $\epsilon > 0$ and $c < 2$,
-\[
-\prob \left\{ \sup_{\lambdabf \in B}\abs{ D_N(\lambdabf) } > \epsilon \right\} = O(e^{-c \epsilon^2 N}).
-\]
- \end{lemma}
-\begin{proof}
-Consider a rectangular grid of points spaced over the identifiable region $B$.  We use $\lambdabf[\rbf]$, where $\rbf \in \ints^{m+1}$, to denote the grid point 
- %\begin{align*}
-\[ 
-%\lambdabf[\rbf] = &\left[  \frac{r_0}{N^{b}} - \frac{1}{2}, \; \frac{r_1}{N^{b+1}} - \frac{1}{2}, \right.\\ 
-%&\dots, \; \frac{r_k}{N^{b+k}} - \frac{1}{2(k!)}, \; \dots \\
-%&\left. \dots, \; \frac{r_m}{N^{b+m}} - \frac{1}{2(m!) }\right]
- \lambdabf[\rbf] = \left[  \frac{r_0}{N^{b}} - \frac{1}{2}, \; \frac{r_1}{N^{b+1}} - \frac{1}{2}, \; \dots, \; \frac{r_m}{m!N^{b+m}} - \frac{1}{2(m!) }\right]
-\]
- %\end{align*}
- for some constant $b>0$.
-Adjacent grid points are separated by $\tfrac{1}{N^b}$ in the zeroth coordinate, $\tfrac{1}{N^{b+1}}$ in the first coordinate and $\tfrac{1}{k!N^{b+k}}$ in the $k$th coordinate.
-Let
- \[
- B[\rbf]=\left\{  \xbf\in\reals^{m+1}  ; \frac{r_k}{N^{b+k}}\leq x_k + \frac{1}{2(k!)} < \frac{r_k + 1}{N^{b+k}} \right\}  .
- \]
-and let $G$ be the finite set of grid points
-\[
-G = \left\{ \xbf \in \ints^{m+1} \mid x_k = 0,1,2\dots,N^{b+k}-1  \right\}.
-\]
-The total number of grid points is $|G| = N^{(m+1)(2b + m)/2}$, and the $B[\rbf]$ partition $B$, that is, $B = \cup_{\rbf \in G}B[\rbf]$.  Now 
-\begin{align}
- &\sup_{\lambdabf \in B}\sabs{ D_N(\lambdabf) }  \nonumber \\
-&= \sup_{\rbf \in G}\sup_{ \lambdabf \in B[\rbf]}\sabs{ D_N(\lambdabf[\rbf]) + D_N(\lambdabf) - D_N(\lambdabf[\rbf])  } \nonumber \\
- &\leq \sup_{\rbf \in G}\sabs{ D_N(\lambdabf[\rbf])} + \sup_{\rbf \in G}\sup_{\lambdabf\in B[\rbf]}\sabs{ D_N(\lambdabf) - D_N(\lambdabf[\rbf])  }. \label{eq:gridanddensespace}
- \end{align}
-From Lemma~\ref{lem:supVjk} it will follow that 
-\[
-\prob\left\{   \sup_{\rbf \in G }\abs{ D_N(\lambdabf[\rbf])  } > \frac{\epsilon}{2} \right\} = O(e^{-c \epsilon^2 N})
-\]
-for any $\epsilon > 0$ and $c < 2$.  In Lemma~\ref{lem:supBVn} we show that
-\[
-\sup_{\rbf \in G}\sup_{\lambdabf\in B[\rbf]}\abs{ D_N(\lambdabf) - D_N(\lambdabf[\rbf])} < 2\frac{m+1}{N^b}.
-\]
-%that is, $D_N$ does not change much within any of the regions $B[\rbf]$ neighbouring the grid points.  
-Combining these results with~\eqref{eq:gridanddensespace}, we obtain
-\[
-\prob\left(\sup_{\lambdabf \in B }\sabs{ D_N(\lambdabf)} > \frac{\epsilon}{2} +  \frac{2(m+1)}{N^b} \right) = O(e^{-c \epsilon^2 N}),
-\]
-and for sufficiently large $N$, we have $\epsilon/2 + \frac{2(m+1)}{N^b} < \epsilon$ completing the proof.  It remains to prove Lemmas~\ref{lem:supVjk} and~\ref{lem:supBVn}.
-\end{proof}
-
-% We first need the following special case of Hoeffding's inequality~\cite{Hoeffding_inequality_1963}.
-
-% \begin{lemma} \label{lem:zero_mean_indepent_sum_bound}
-%   Let $\bar{Z} = (Z_1 + \dots + Z_N)/N$ where $Z_1, \dots, Z_N$ are independent, zero-mean random variables with $|Z_n| \leq \tfrac{1}{4}$ for each $n$.  Then,
-% \[
-%   \prob\{ \sabs{\bar{Z}} > \epsilon \} \leq 2e^{ - 8 \epsilon^2 N}.
-% \]
-% \end{lemma}
-
-\begin{lemma}\label{lem:supVjk}
-For any $\epsilon > 0$ and $c < 8$,
-\[
-\prob\cubr{   \sup_{\rbf \in G}\abs{ D_N(\lambdabf[\rbf])  } > \epsilon } = O(e^{-c \epsilon^2 N}).
-\]
-\end{lemma}
- \begin{proof}
-Fix $\lambdabf$ and write 
-\[
-D_N(\lambdabf) = \bar{Z} = \frac{1}{N}  \sum_{n=1}^{N}Z_{n},
-\] 
-where
- \[
- Z_{n}=\fracpart{  \Phi_n+\sum_{k = 0}^{m}{\lambda_k n^k} }^{2} - \expect \fracpart{  \Phi_n + \sum_{k = 0}^{m}{\lambda_k n^k} }^{2}
- \]
- are independent with zero mean and $\abs{Z_n} \leq \tfrac{1}{4}$. It follows from Hoeffding's inequality~\cite{Hoeffding_inequality_1963} that,  
-\[
-\prob\{ \abs{D_N(\lambdabf)} > \epsilon \} \leq 2e^{ - 8 \epsilon^2 N},
-\] 
-and so,
- \begin{align*}
- \prob\cubr{  \sup_{\rbf\in G}\left\vert D_N(\lambdabf[\rbf]) \right\vert >\epsilon }  &\leq \sum_{\rbf \in G} \prob\scubr{  \left\vert D_N(\lambdabf[\rbf])  \right\vert >\epsilon } \nonumber \\
-&= 2 |G| e^{ - 8 \epsilon^2 N} = O(e^{-c \epsilon^2 N}), \label{eq:betaprobconv}
-\end{align*}
-where $c$ is any real number less than $8$, since $|G| = N^{(m+1)(2b + m)/2}$ is polynomial in $N$.
-\end{proof}
-
-Before proving Lemma~\ref{lem:supBVn} we need the following result.
-
-\begin{lemma}\label{lem:boundedsquarefracparts} For real numbers $x$ and $delta$,
-\[
-\fracpart{x}^2 - |\delta| \leq \fracpart{x + \delta}^2 \leq \fracpart{x}^2 + |\delta|.
-\]
-\end{lemma}
-\begin{proof}
-Since $\abs{\delta} \leq \abs{n + \delta}$ for all $\delta \in [-\nicefrac{1}{2}, \nicefrac{1}{2})$ and $n \in \ints$, the result will follow if we can show that it holds when both $x$ and $\delta$ are in $[-\nicefrac{1}{2}, \nicefrac{1}{2})$.
-Also, for reasons of symmetry, we need only show that it holds when $\delta \geq 0$.
-Now
-\[
-\fracpart{x + \delta}^2 - x^2 = \begin{cases}
-2x\delta + \delta^2, & x \in [-\nicefrac{1}{2}, \nicefrac{1}{2} - \delta) \\
-2x(\delta-1) + (\delta-1)^2, & x \in [\nicefrac{1}{2} - \delta, \nicefrac{1}{2})
-\end{cases}
-\] 
-But, when $x \in [-\nicefrac{1}{2}, \nicefrac{1}{2} - \delta)$, 
-\[
--1 \leq -1 + \delta \leq 2x + \delta < 1 - \delta \leq 1,
-\]
-and so
-\[
--\delta \leq (-1 + \delta)\delta \leq (2x + \delta)\delta < (1-\delta)\delta \leq \delta.
-\]
-Also, when $x \in [\nicefrac{1}{2} - \delta, \nicefrac{1}{2})$ we  have $-\delta \leq 2x + \delta - 1 < \delta$, and consequently
-\[
--\delta \leq -\delta(1-\delta) \leq (2x + \delta - 1)(1 - \delta) \leq \delta(1 - \delta) \leq \delta.
-\]
-%and consequenly
-%\[
-%-\delta \leq (2x + \delta - 1)(\delta - 1) \leq \delta.
-%\]
-
-% In some sense this result is `obvious' because  the function $\fracpart{x}^2$ is continuous and piecewise differentiable and the derivative has magnitude less than one whenever it exists.
-We will give a rigorous proof of the lemma when $\delta \geq 0$, the proof for $\delta \leq 0$ is similar.
-Note that $0 \leq \fracpart{x}^2 \leq \tfrac{1}{4}$ for all real numbers $x$.
-So if $\delta \geq \tfrac{1}{4}$ then the lemma is true because
-% \[
-% \fracpart{x}^2 - |\delta| \leq 0 \leq \fracpart{x + \delta}^2 \leq \tfrac{1}{4} \leq \fracpart{x}^2 + |\delta|.
-% \]
-% So we may assume that $\delta < \tfrac{1}{4}$.
-We consider two cases, firstly when $\fracpart{x} + \delta \in [-\nicefrac{1}{2}, \nicefrac{1}{2})$, and secondly when $\fracpart{x} + \delta \geq \nicefrac{1}{2}$.
-Assume that $\fracpart{x} + \delta \in [-\nicefrac{1}{2}, \nicefrac{1}{2}]$ and therefore
-% \[
-% -\tfrac{1}{2} \leq \fracpart{x} \leq \tfrac{1}{2} - \delta
-% \]
-% and also
-% \[
-% \fracpart{x + \delta}^2 = \fracpart{\fracpart{x} + \delta}^2 = (\fracpart{x} + \delta)^2 = \fracpart{x}^2 + 2\fracpart{x}\delta + \delta^2.
-% \]
-% Substituting the bounds on $\fracpart{x}$ above into the right hand side of this equation we obtain
-% \[
-% \fracpart{x}^2 - \delta(1 - \delta) \leq \fracpart{x + \delta}^2 \leq \fracpart{x}^2 + \delta(1 - \delta).
-% \]
-% The lemma is true because $\delta(1 - \delta) \leq |\delta|$ since $0 \leq \delta < \tfrac{1}{4}$. Now assume that $\fracpart{x} + \delta \geq \nicefrac{1}{2}$. We have
-% \[
-% \fracpart{x + \delta} = \fracpart{x} + \delta - 1
-% \]
-% and therefore
-% \[
-% \tfrac{1}{2} - \delta \leq \fracpart{x} \leq \tfrac{1}{2}
-% \]
-% and also 
-% \[
-% \fracpart{x + \delta}^2 = (\fracpart{x} + \delta - 1)^2 = \fracpart{x}^2 + 2\fracpart{x}(\delta  - 1) + (\delta - 1)^2.
-% \]
-% Substituting the bounds on $\fracpart{x}$ above into the right hand side of this equation we again obtain
-% \[
-% \fracpart{x}^2 - \delta(1 - \delta) \leq \fracpart{x + \delta}^2 \leq \fracpart{x}^2 + \delta(1 - \delta)
-% \]
-% and the lemma is true because $\delta(1 - \delta) \leq |\delta|$ since $0 \leq \delta < \tfrac{1}{4}$.
-\end{proof}
-
- \begin{lemma}\label{lem:supBVn} 
-For all positive integers $N$,
-\[
-\sup_{\rbf\in G}\sup_{ \lambdabf  \in B[\rbf] }\left\vert D_N\left(  \lambdabf \right) - D_N\left(  \lambdabf[\rbf] \right)\right\vert < 2\frac{m+1}{N^b}.
-\]
- \end{lemma}
- \begin{proof}
-%The proof we give is not just almost surely but \emph{surely} as the convergence will be shown to occur irrespective of the values of the $\Phi_n$. 
-Put 
-\[
-b_n = \Phi_n + \sum_{k = 0}^{m}{\lambda_k n^k} \;\;\; \text{and}  \;\;\; a_n = \Phi_n + \sum_{k = 0}^{m}{\lambda[\rbf]_k n^k},
-\]
-where $\lambda[\rbf]_k$ denotes the $k$th element of the grid point $\lambdabf[\rbf]$. For $\lambdabf \in B[\rbf]$ we have $b_n = a_n + \delta_n$,
-where 
-\[
-\sabs{\delta_n} \leq \sum_{k=0}^m\frac{n^k}{k! N^{b+k}} \leq \frac{m+1}{N^{b}}.
-\]
-From Lemma~\ref{lem:boundedsquarefracparts} it follows that 
-\[
--|\delta_n| \leq \fracpart{x + b_n}^2 -  \fracpart{x + a_n}^2  \leq |\delta_n|,
-\]
-and consequently 
-\[
-|\fracpart{x + b_n}^2 -  \fracpart{x + a_n}^2| \leq \tfrac{m+1}{N^b}
-\] 
-for all $x \in \reals$. Now
-\[
-S_N(\lambdabf) - S_N(\lambdabf[\rbf]) =  \frac{1}{N}\sum_{n=1}^{N}\big( \fracpart{\Phi_n + b_n}^2 -  \fracpart{\Phi_n + a_n}^2 \big)
-\]
-and therefore 
-\[
-\sabs{S_N(\lambdabf) - S_N(\lambdabf[\rbf])} \leq \frac{m+1}{N^b}
-\] 
-for all $\lambdabf \in B[\rbf]$.  As this bound is independent of $\Phi_1 \dots \Phi_N$, we have
-\[
-\sabs{ V_N(\lambdabf) - V_N(\lambdabf[\rbf])} \leq \expect \sabs{ S_N(\lambdabf) - S_N(\lambdabf[\rbf])} \leq \frac{m+1}{N^b}
-\]
-by Jensen's inequality.  Therefore,
- \begin{align*}
- &\sabs{ D_N(\lambdabf) - D_N(\lambdabf[\rbf]) } \\
-&= \sabs{ S_N(\lambdabf) - S_N(\lambdabf[\rbf]) + V_N(\lambdabf) - V_N(\lambdabf[\rbf]) } \\
-&\leq \sabs{ S_N(\lambdabf) - S_N(\lambdabf[\rbf]) \vert + \vert V_N(\lambdabf) - V_N(\lambdabf[\rbf]) } \\
-&\leq 2\frac{m+1}{N^{b}},
-\end{align*}
-for all $\lambdabf \in B[\rbf]$.  The lemma follows because this bound is independent of $\rbf$.
-%\[
-%\sup_{\rbf\in G}\sup_{ \lambdabf  \in B[\rbf] }\sabs{ D_N\left(  \lambdabf \right) - D_N\left(  \lambdabf[\rbf] \right) } \leq 2\frac{m+1}{N^{b}}.
-%\]
- \end{proof}
 

### 2013-02-04 paper.tex

@@
 This is complicated by the fact that the phase is \emph{wrapped} modulo $2\pi$ and must be \emph{unwrapped} before regression can be performed.
-%A recent approach suggested by the authors is to perform the unwrapping in a least squares manner.
-%It was shown how the resulting estimator could be posed as a \emph{nearest lattice point problem} in a specific lattice and could be solved using existing algorithms.
-%It was shown by Monte Carlo simulation that this produces a remarkably accurate estimator.  
-In this two part series of papers we consider an estimator that performs phase unwrapping in a least squares manner.
-In this first part we describe conditions for the identifiability of polynomial phase signals and we prove the strong consistency of our unwrapping estimator.
-%We describe the asymptotic properties of this estimator, showing that it is strongly consistent and asymptotically normally distributed.
-%We hypothesise that the estimator produces very near maximum likelihood performance.
+In this paper we consider an estimator that performs phase unwrapping in a least squares manner.
+We describe the asymptotic properties of this estimator, showing that it is strongly consistent and asymptotically normally distributed.
+No `small noise' assumption (common in the literature) is required in our analysis.
+As a result our estimator will be accurate, even for very noisy signals, provided that the number of observations is sufficiently large.
+This appears to be the first time a theoretical result of this type has be obtained in the context of polynomial phase estimation, and the proof techniques we develop may be useful for strengthening the analysis of existing polynomial phase estimators.
+Our unwrapping estimator can be computed in a reasonable amount of time for data sets of moderate size using existing general purpose algorithms from algebraic number theory.
+Monte-carlo simualtions assert the superior statistical performance of our unwrapping estimator when compared to existing state-of-the-art estimators for data sets of less than 300 observations.
 \end{abstract}
@@
 
-Polynomial phase signals arise in fields including radar, sonar, geophysics, biology, and radio communication \cite{Angeby_estimating_2000}.
+Polynomial phase signals arise in fields including radar, sonar, geophysics, biology, and radio communication~\cite{Hlawatsch_lin_quad_time_freq_spmag_1992,Ridleyspeechpolyphase1989, Suga_1975_bats_echolocation, Moss_2005echolocation,Angeby_estimating_2000}.
 In radar and sonar applications polynomial phase signals arise when acquiring radial velocity and acceleration (and higher order motion descriptors) of a target from a reflected signal, and also in continuous wave radar and low probability of intercept radar~\cite{Levanon_Radar_signals_2004}.

### 2013-02-05 paper.tex
added java code etc.

@@
 In this paper we consider an estimator that performs phase unwrapping in a least squares manner.
-We describe the asymptotic properties of this estimator, showing that it is strongly consistent and asymptotically normally distributed.
+Under mild conditions on the distribution of the noise we describe the asymptotic properties of this estimator, showing that it is strongly consistent and asymptotically normally distributed.
 No `small noise' assumption (common in the literature) is required in our analysis.
 As a result our estimator will be accurate, even for very noisy signals, provided that the number of observations is sufficiently large.
-This appears to be the first time a theoretical result of this type has be obtained in the context of polynomial phase estimation, and the proof techniques we develop may be useful for strengthening the analysis of existing polynomial phase estimators.
 Our unwrapping estimator can be computed in a reasonable amount of time for data sets of moderate size using existing general purpose algorithms from algebraic number theory.
@@
 In this paper we only consider uniform sampling, where the gap between consecutive samples is constant.
-In this case we can always consider the samples to be taken at some set of consecutive integers and our sampled polynomial phase signal looks like
+In this case we can always consider the samples to be taken at some set of consecutive integers and our sampled polynomial phase signal takes the form
 \[
@@
 This is also the maximum likelihood estimator (MLE) when the noise sequence $\{X_n\}$ is white and Gaussian.
-When $m=0$ (phase estimation) or $m=1$ (frequency estimation) the LSE is an effective approach, being both computationally efficient and statistically accurate \cite{Quinn2009_dasp_phase_only_information_loss,Hannan1973,Quinn2001,McKilliam_mean_dir_est_sq_arc_length2010}\cite[Sec.~6.4~and~9.1]{McKilliam2010thesis}.
-When $m \geq 2$ the computational complexity of the LSE is large~\cite[Sec.~10.1]{McKilliam2010thesis}\cite{Abatzoglou_ml_chirp_1986}.
+When $m=0$ (phase estimation) or $m=1$ (frequency estimation) the LSE is an effective approach, being both computationally efficient and statistically accurate \cite{Quinn2009_dasp_phase_only_information_loss,Hannan1973,Quinn2001,McKilliam_mean_dir_est_sq_arc_length2010,McKilliam2010thesis}.
+When $m \geq 2$ the computational complexity of the LSE is large~\cite{McKilliam2010thesis,Abatzoglou_ml_chirp_1986}.
 For this reason many authors have considered alternative approaches to polynomial phase estimation.
-These can loosely be grouped into two classes, estimators based on polynomial phase transforms, such as the discrete polynomial phase transform (DPT)~\cite{Peleg_DPT_1995,Peleg1991_est_class_PPS_1991,Porat_asympt_HAF_DPT_1996} and the high order phase function~\cite{Farquharson_another_poly_est_2005}, and estimators based on phase unwrapping, such as Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, the estimator of Djuric and Kay~\cite{Djuric_phase_unwrap_chirp_1990}, and Morelande's Bayesian unwrapping estimator~\cite{Morelande_bayes_unwrapping_2009_tsp}.
+These can loosely be grouped into two classes: estimators based on `multilinear transforms', such as the discrete polynomial phase transform (DPT), high order ambiguity function (HAF)~\cite{Peleg_DPT_1995,Peleg1991_est_class_PPS_1991,Porat_asympt_HAF_DPT_1996,Farquharson_another_poly_est_2005,Boashash_Oshea_1994}, or cubic phase function (CPF)~\cite{Oshea_cpf_2004,Oshea_cpf_2002_letter}; and estimators based on phase unwrapping, such as Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, the estimator of Djuric and Kay~\cite{Djuric_phase_unwrap_chirp_1990}, and Morelande's Bayesian unwrapping estimator~\cite{Morelande_bayes_unwrapping_2009_tsp}.
 
@@
 It was shown in \cite{McKilliam2009asilomar_polyest_lattice, McKilliamFrequencyEstimationByPhaseUnwrapping2009} that the LSU estimator can be represented as a \emph{nearest lattice point problem}, and Monte-Carlo simulations were used to show the LSU estimator's favourable statistical performance.
-Moreover, it was noticed that the LSU appears to work correctly for coefficients anywhere in the identifiable region.
+%Moreover, it was noticed that the LSU appears to work correctly for coefficients anywhere in the identifiable region.
 A drawback of the LSU estimator is that computing a nearest lattice point is, in general, computationally difficult.
-In \cite{McKilliam2009asilomar_polyest_lattice}, two standard techniques were considered, the \emph{sphere decoder} \cite{Pohst_sphere_decoder_1981}, and \emph{Babai's nearest plane algorithm} \cite{Babai1986}.
-The sphere decoder was observed to have excellent statistical performance but can only be computed efficiently for $N<50$ \cite{Jalden2005_sphere_decoding_complexity}.
-Computing the Babai point requires only $O(N^2)$ operations, but its statistical performance is comparatively poor at low signal to noise ratio (SNR).
-A major point of interest is that the lattices considered are not \emph{random}, and therefore may admit fast nearest point algorithms.
+In \cite{McKilliam2009asilomar_polyest_lattice}, two standard techniques were considered, the \emph{sphere decoder}~\cite{Pohst_sphere_decoder_1981} that exactly computes a nearest lattice point, and \emph{Babai's nearest plane algorithm}~\cite{Babai1986} that only produces and approximation.
+The sphere decoder was observed to have excellent statistical performance but can only be computed efficiently for small $N$ less than about $60$~\cite{Jalden2005_sphere_decoding_complexity}.
+Computing the Babai point requires only $O(N^2)$ operations, but its statistical performance was observed to be comparatively poor at low signal to noise ratio (SNR).
+A major point of interest is that the lattices considered are not \emph{random}, and may admit fast nearest point algorithms.
 This has been studied in~\cite[Sec.~4.2]{McKilliam2010thesis} where polynomial time algorithms where found that compute the nearest point exactly.
 Unfortantely, although polynomial time, the algorithm are still computationally very slow in practice.
- Two main questions were raised in \cite{McKilliam2009asilomar_polyest_lattice}:
- \begin{enumerate}
- \vspace{-0.1cm}
- \item What are the asymptotic statistical properties of the LSU estimator?
- \vspace{-0.18cm}
- \item Can the structure of the lattice be exploited to obtain fast (exact or approximate) nearest point algorithms and thereby a computationally efficient estimator?
- \vspace{-0.1cm}
- \end{enumerate}
- The primary purpose of this paper is to answer the first question.
-Similar results were stated without a complete proof in~\cite{McKilliam_polyphase_est_icassp_2011}.
-Here, we give a proof.
-The results here are also more general than in~\cite{McKilliam_polyphase_est_icassp_2011}, allowing for a wider class of noise distributions.
-We prove that the LSU estimator is strongly consistent and asymptotically normally distributed.
-BLERG `small noise' assumption.
-
-We also apply another known approximate nearest point algorithm, called the $K$-best method \cite{Zhan2006_K_best_sphere_decoder}, that we show provides near sphere decoder performance, but can be computed in a reasonable amount of time if $N$ is less than about 300.
-BLERG `better than everyting else and computationally reasonably for smallish $N \leq 300$'.
+Nevertheless fast (exact or approximate) algorithms that we have not found may still exist for these special lattices.
+
+In this paper we apply another general purpose approximate nearest point algorithm, called the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder} for approximating the LSU estimator.
+We show that this provides near sphere decoder (exact) performance, but can be computed in a reasonable amount of time if $N$ is not too large.
+BLERG benchmarking results.
+The $K$-best method is motivated by sequential decoding algorithms and, in the context of polynomial phase estimation, bares some resemblance to the Bayesian unwrapping algorithm studied by Morelande~\cite{Morelande_bayes_unwrapping_2009_tsp}.
+We show by Montecarlo simulation that the LSU estimator approximated using the $K$-best method is statistically superior to state-of-the-art estimators from the literature including the HAF and CPF.
+We also compare the computational performance of the estimators and find that BLERG.
+
+In addition to the above computational results a major purpose of this paper is to study the asymptotic properties of the LSU estimator.
+Under mild assumptions about the distribution of the noise we prove that the LSU estimator is strongly consistent, that is, the estimated coefficients converge (almost surely) to the true coefficients as the number of observations $N$ grows.
+Similar results were stated without proof in~\cite{McKilliam_polyphase_est_icassp_2011}.
+In the   BLERG `small noise' assumption.
 
@@
 Empirical process techniques~\cite{Pollard_new_ways_clts_1986,Pollard_asymp_empi_proc_1989,van2009empirical,Dudley_unif_central_lim_th_1999} and results from the literature on hyperplane arrangements~\cite{Chazelle_discrepency_method_2000,Matousek_lect_disc_geom_2002} become useful here.
-We are hopeful that the proof techniques developed here will be useful for purposes other than polynomial phase estimation, and in particular other applications involving data that is `wrapped' in some sense.
+This proof technique appears to be novel and may be useful for purposes other than polynomial phase estimation, and in particular other applications involving data that is `wrapped' in some sense.
 Potential candidates are the phase wrapped images observed in modern radar and medical imaging devices such as synthetic aperture radar and magnetic resonance imaging~\cite{Nico_phaseunwrappingSAR_2000,Friedlander_PD_phaseunwrapping_1996}.
+
+We also assert, under similarly mild conditions, that the LSU estimator is asymptotically normal.
+BLERG CLOSENESS TO CLRB in Gaussian case.
+Space restrictions make it impossible for us to include a proof of asymptotic normality in this paper.
 
@@
 The theorem is proved in Section~\ref{sec:strongconstproof}.
-In the second paper~\cite{McKilliam_pps2_2012} we prove the asymptotic normality of the LSU estimator and describe the results of Monte Carlo simulations.
-These simulations agree with our derived asymptotic statistical properties.
+Section \ref{sec:simulations} describes the results of Monte Carlo simulations that compare the performance of the LSU estimator with some existing estimators.
+These simulations agree with the derived asymptotic properties.
 
@@
 
-\section{Proof of asymptotic normality}\label{sec:centlimitproof}
-
-Let $\psibf$ be the vector with $k$th component $\psi_k = N^k \lambda_k$, $k=0, \dots, m$ and let 
-\[
-T_{N}(\psibf) = S_N(\lambdabf) = \frac{1}{N} \sum_{n=1}^{N} \fracpart{ \Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \psi_k }^2.
-\]
-Let $\widehat{\psibf}_N$ be the vector with elements $\widehat{\psi}_{k,N} = N^k \widehat{\lambda}_{k,N}$ so that $\widehat{\psibf}_N$ is the minimiser of $T_N$.
+% \section{Proof of asymptotic normality}\label{sec:centlimitproof}
+
+% Let $\psibf$ be the vector with $k$th component $\psi_k = N^k \lambda_k$, $k=0, \dots, m$ and let 
+% \[
+% T_{N}(\psibf) = S_N(\lambdabf) = \frac{1}{N} \sum_{n=1}^{N} \fracpart{ \Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \psi_k }^2.
+% \]
+% Let $\widehat{\psibf}_N$ be the vector with elements $\widehat{\psi}_{k,N} = N^k \widehat{\lambda}_{k,N}$ so that $\widehat{\psibf}_N$ is the minimiser of $T_N$.
 Because each of $N^k \widehat{\lambda}_{k,N}$ converges almost surely to zero as $N \rightarrow \infty$~\cite[Theorem~2]{McKilliam_pps1_2012}, it follows that $\widehat{\psibf}_{N}$ converges almost surely to $\zerobf$ as $N \rightarrow \infty$.
 We want to find the asymptotic distribution of
-\[
-\sqrt{N}\widehat{\psibf}_{N} = 
-\left[
-\begin{array}
-[c]{c}%
-\sqrt{N} \widehat{\psi}_{0,N} \\ \sqrt{N}\widehat{\psi}_{1,N}  \\ \vdots \\ \sqrt{N} \widehat{\psi}_{m,N}
-\end{array}
-\right]
-=
-\left[
-\begin{array}
-[c]{c}%
-\sqrt{N} \widehat{\lambda}_{0,N} \\ N\sqrt{N}\widehat{\lambda}_{1,N} \\ \vdots \\ N^m\sqrt{N} \widehat{\lambda}_{m,N}
-\end{array}
-\right].
-\]
-The proof is complicated by the fact that $T_N$ is not differentiable everywhere as $\fracpart{x}^2$ is not differentiable when $\fracpart{x} = \tfrac{1}{2}$.
+% \[
+% \sqrt{N}\widehat{\psibf}_{N} = 
+% \left[
+% \begin{array}
+% [c]{c}%
+% \sqrt{N} \widehat{\psi}_{0,N} \\ \sqrt{N}\widehat{\psi}_{1,N}  \\ \vdots \\ \sqrt{N} \widehat{\psi}_{m,N}
+% \end{array}
+% \right]
+% =
+% \left[
+% \begin{array}
+% [c]{c}%
+% \sqrt{N} \widehat{\lambda}_{0,N} \\ N\sqrt{N}\widehat{\lambda}_{1,N} \\ \vdots \\ N^m\sqrt{N} \widehat{\lambda}_{m,N}
+% \end{array}
+% \right].
+% \]
+% The proof is complicated by the fact that $T_N$ is not differentiable everywhere as $\fracpart{x}^2$ is not differentiable when $\fracpart{x} = \tfrac{1}{2}$.
 This precludes the use of ``standard approaches'' to proving asymptotic normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.
@@
 Thus, putting
-\begin{equation}\label{eq:Wn}
-W_{n} = \round{\Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \widehat{\psi}_{k,N}},
-\end{equation}
-we have, for each $\ell = 0, \dots, m$,
-\begin{align*}
-0 & = \frac{\partial T_N}{\partial \psi_\ell}(\widehat{\psibf}_N)\\
-&= \frac{2}{N}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\left( \Phi_n - W_{n}+ \sum_{k=0}^{m}(\tfrac{n}{N})^k \widehat{\psi}_{k,N}  \right),
-\end{align*}
-so that
-\[
-D_{\ell,N} = K_{\ell,N},
-\]
-where
-\[
-D_{\ell,N} = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}(\tfrac{n}{N})^\ell \Phi_n,
-\]
-and
-\begin{equation}\label{eq:KellN}
-K_{\ell,N} = \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\left(W_{n}- \sum_{k=0}^{m}(\tfrac{n}{N})^k \widehat{\psi}_{k,N}  \right).
-\end{equation}
-Lemma~\ref{lem:Kconvfhalf} shows that, for all $\ell = 0, \dots m$,
-\begin{equation}\label{eq:KellNconv}
-K_{\ell,N} =  (h - 1) \sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big( C_{\ell k} + o_P(1) \big) + o_P(1)
-\end{equation}
-where $C_{\ell k} =  \tfrac{1}{\ell + k + 1}$, and $h = f(-\nicefrac{1}{2})$, and $o_P(1)$ denotes a random variable converging in probability to zero as $N\rightarrow\infty$.
-
-It is now convenient to write in vector form.  Let 
-\begin{equation}\label{eq:dnandkn}
-\kbf_N = 
-\left[
-\begin{array}
-[c]{c}%
-K_{0,N}  \\ \vdots \\ K_{m,N}
-\end{array}
-\right]
-\qquad \text{and} \qquad 
-\dbf_N = \left[
-\begin{array}
-[c]{c}%
-D_{0,N}  \\ \vdots \\ D_{m,N}
-\end{array}
-\right].
-\end{equation}
-From~\eqref{eq:KellNconv},
-\[
-\dbf_N = \kbf_N = \sqrt{N} (h - 1)(\Cbf + o_P(1)) \widehat{\psibf}_N + o_P(1)
-\]
-where $o_P(1)$ here means a vector or matrix of the appropriate dimension with every element converging in probability to zero as $N\rightarrow\infty$.
+% \begin{equation}\label{eq:Wn}
+% W_{n} = \round{\Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \widehat{\psi}_{k,N}},
+% \end{equation}
+% we have, for each $\ell = 0, \dots, m$,
+% \begin{align*}
+% 0 & = \frac{\partial T_N}{\partial \psi_\ell}(\widehat{\psibf}_N)\\
+% &= \frac{2}{N}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\left( \Phi_n - W_{n}+ \sum_{k=0}^{m}(\tfrac{n}{N})^k \widehat{\psi}_{k,N}  \right),
+% \end{align*}
+% so that
+% \[
+% D_{\ell,N} = K_{\ell,N},
+% \]
+% where
+% \[
+% D_{\ell,N} = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}(\tfrac{n}{N})^\ell \Phi_n,
+% \]
+% and
+% \begin{equation}\label{eq:KellN}
+% K_{\ell,N} = \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\left(W_{n}- \sum_{k=0}^{m}(\tfrac{n}{N})^k \widehat{\psi}_{k,N}  \right).
+% \end{equation}
+% Lemma~\ref{lem:Kconvfhalf} shows that, for all $\ell = 0, \dots m$,
+% \begin{equation}\label{eq:KellNconv}
+% K_{\ell,N} =  (h - 1) \sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big( C_{\ell k} + o_P(1) \big) + o_P(1)
+% \end{equation}
+% where $C_{\ell k} =  \tfrac{1}{\ell + k + 1}$, and $h = f(-\nicefrac{1}{2})$, and $o_P(1)$ denotes a random variable converging in probability to zero as $N\rightarrow\infty$.
+
+% It is now convenient to write in vector form.  Let 
+% \begin{equation}\label{eq:dnandkn}
+% \kbf_N = 
+% \left[
+% \begin{array}
+% [c]{c}%
+% K_{0,N}  \\ \vdots \\ K_{m,N}
+% \end{array}
+% \right]
+% \qquad \text{and} \qquad 
+% \dbf_N = \left[
+% \begin{array}
+% [c]{c}%
+% D_{0,N}  \\ \vdots \\ D_{m,N}
+% \end{array}
+% \right].
+% \end{equation}
+% From~\eqref{eq:KellNconv},
+% \[
+% \dbf_N = \kbf_N = \sqrt{N} (h - 1)(\Cbf + o_P(1)) \widehat{\psibf}_N + o_P(1)
+% \]
+% where $o_P(1)$ here means a vector or matrix of the appropriate dimension with every element converging in probability to zero as $N\rightarrow\infty$.
 Thus $\sqrt{N}\widehat{\psibf}_N$ has the same asymptotic distribution as $(h - 1)^{-1}\Cbf^{-1} \dbf_N$.
@@
 Hence $\sqrt{N}\widehat{\psibf}_N$ is asymptotically normal with zero mean and covariance matrix
-\[
-\frac{\sigma^2\Cbf^{-1}\Cbf(\Cbf^{-1})^\prime}{(1 - h)^2} = 
-\frac{\sigma^2\Cbf^{-1}}{(1 - h)^2}.
-\] 
-It remains to prove Lemmas~\ref{lem:diffathatpsi},~\ref{lem:Kconvfhalf} and~\ref{eq:convdn}.
-
-\begin{lemma}\label{lem:diffathatpsi}
-For all $\ell = 0, \dots, m$ the partial derivatives $\frac{\partial T_N}{\partial \psi_\ell}$ exist, and are equal to zero, at the minimiser $\widehat{\psibf}_N$.  That is,
-\[
-\frac{\partial T_N}{\partial \psi_\ell}(\widehat{\psibf}_N) = 0 \qquad \text{for each $\ell = 0, \dots, m$.}
-\]
-\end{lemma}
-\begin{IEEEproof}
-The function $\fracpart{x}^2$ is differentiable everywhere except if $\fracpart{x} \neq -\nicefrac{1}{2}$.  Recalling that
-\[
-T_{N}(\psibf) = \frac{1}{N} \sum_{n=1}^{N} \fracpart{ \Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \psi_k }^2
-\]
-we see that $T_N$ will be differentiable with respect to $\psibf$ at $\widehat{\psibf}_N$ if 
-\[
-\fracpart{\Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \widehat{\psi}_{k,N}} \neq -\nicefrac{1}{2} \qquad \text{for all $n = 1, \dots, N$.}
-\] 
-This is proved in Lemma~\ref{lem:boundonhatlambda}.
+% \[
+% \frac{\sigma^2\Cbf^{-1}\Cbf(\Cbf^{-1})^\prime}{(1 - h)^2} = 
+% \frac{\sigma^2\Cbf^{-1}}{(1 - h)^2}.
+% \] 
+% It remains to prove Lemmas~\ref{lem:diffathatpsi},~\ref{lem:Kconvfhalf} and~\ref{eq:convdn}.
+
+% \begin{lemma}\label{lem:diffathatpsi}
+% For all $\ell = 0, \dots, m$ the partial derivatives $\frac{\partial T_N}{\partial \psi_\ell}$ exist, and are equal to zero, at the minimiser $\widehat{\psibf}_N$.  That is,
+% \[
+% \frac{\partial T_N}{\partial \psi_\ell}(\widehat{\psibf}_N) = 0 \qquad \text{for each $\ell = 0, \dots, m$.}
+% \]
+% \end{lemma}
+% \begin{IEEEproof}
+% The function $\fracpart{x}^2$ is differentiable everywhere except if $\fracpart{x} \neq -\nicefrac{1}{2}$.  Recalling that
+% \[
+% T_{N}(\psibf) = \frac{1}{N} \sum_{n=1}^{N} \fracpart{ \Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \psi_k }^2
+% \]
+% we see that $T_N$ will be differentiable with respect to $\psibf$ at $\widehat{\psibf}_N$ if 
+% \[
+% \fracpart{\Phi_n + \sum_{k=0}^m (\tfrac{n}{N})^k \widehat{\psi}_{k,N}} \neq -\nicefrac{1}{2} \qquad \text{for all $n = 1, \dots, N$.}
+% \] 
+% This is proved in Lemma~\ref{lem:boundonhatlambda}.
 So the partial derivatives $\frac{\partial T_N}{\partial \psi_\ell}$ exist for all $\ell = 0, \dots, m$ at $\widehat{\psibf}_N$.
 That each of the partial derivatives is equal to zero at $\widehat{\psibf}_N$ follows immediately from the fact that $\widehat{\psibf}_N$ is a minimiser of $T_N$.
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:boundonhatlambda} $\sabs{\sfracpart{\Phi_n + \sum_{k=0}^m (\nicefrac{n}{N})^k \widehat{\psi}_{k,N}}} \leq \frac{1}{2} - \frac{1}{2N}$ for all $n = 1, \dots, N$.
-\end{lemma}
-\begin{IEEEproof}
-To simplify our notation let 
-\[
-B_n = \Phi_n + \sum_{k=1}^m (\nicefrac{n}{N})^k \widehat{\psi}_{k,N}
-\]
-so that we now require to prove $\sabs{\sfracpart{B_n + \widehat{\psi}_{0,N}}} \leq \frac{1}{2} - \frac{1}{2N}$
-for all $n = 1, \dots, N$.  From~\eqref{eq:Wn}, $W_n = \sround{B_n + \widehat{\psi}_{0,N}}$, and 
-\begin{align*}
-T_N(\widehat{\psibf}_N) &= \frac{1}{N}\sum_{n=1}^N \sfracpart{B_n + \widehat{\psi}_{0,N}}^2
-\\ &= \frac{1}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N} - W_n)^2.
-\end{align*}
-Since $\widehat{\psi}_{0,N}$ is the minimiser of the quadratic above,
-\begin{equation}\label{eq:lamsum}
-\widehat{\psi}_{0,N} = -\frac{1}{N}\sum_{n=1}^N(B_n - W_n).
-\end{equation}
-The proof now proceeds by contradiction.  Assume that for some $k$,
-\begin{equation}\label{eq:philamfraccontra}
-\sfracpart{B_k + \widehat{\psi}_{0,N}} > \frac{1}{2} - \frac{1}{2N}.
-\end{equation}
-Let $F_n = W_n$ for all $n \neq k$ and $F_k = W_k + 1$, and let
-\[
-\phi = -\frac{1}{N}\sum_{n=1}^N(B_n - F_n) = \widehat{\psi}_{0,N} + \frac{1}{N}.
-\]
-Then,
-\begin{align}
-(B_k + &\phi - F_k)^2 = (B_k + \phi - W_k - 1)^2 \nonumber \\
-&= (B_k + \phi - W_k)^2 - 2(B_k + \phi - W_k) + 1 \nonumber \\
-&= (B_k + \phi - W_k)^2 - 2(B_k + \widehat{\psi}_{0,N} - W_k) + 1 - \frac{2}{N} \nonumber \\
-&= (B_k + \phi - W_k)^2 - 2\sfracpart{B_k + \widehat{\psi}_{0,N}} + 1 - \frac{2}{N} \nonumber \\
-&< (B_k + \phi - W_k)^2 - \frac{1}{N}, \label{eq:Bkineq}
-\end{align}
-where the inequality in the last line follows from~\eqref{eq:philamfraccontra}. Let
-\[
-\bbf = \left[
-\begin{array}
-[c]{cccc}%
-\phi & \widehat{\psi}_{1,N} & \cdots & \widehat{\psi}_{m,N}
-\end{array}
-\right]^\prime
-\]
-be the vector of length $m+1$ with components $b_0 = \phi$ and $b_\ell = \widehat{\psi}_{\ell,N}$ for $\ell = 1 , \dots m$.  Now,
-\[
-N T_N(\bbf) = \sum_{n=1}^N \fracpart{B_n + \phi }^2 \leq  \sum_{n=1}^N (B_n + \phi  - F_n)^2, 
-\]
-and using the inequality from~\eqref{eq:Bkineq},
-\begin{align*}
-N &T_N(\bbf) < - \frac{1}{N} + \sum_{n=1}^N (B_n + \phi  - W_n)^2 \\
-&= - \frac{1}{N} + \sum_{n=1}^N (B_n + \widehat{\psi}_{0,N} + \frac{1}{N}  - W_n)^2 \\
-&= \sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n)^2 +  \frac{2}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n)\\
-%&= N T_N(\widehat{\psibf}_N) + \frac{2}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n) \\
-&= N T_N(\widehat{\psibf}_N),
-\end{align*}
-because $\frac{2}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n) = 0$ as a result of~\eqref{eq:lamsum}.
+% \end{IEEEproof}
+
+% \begin{lemma}\label{lem:boundonhatlambda} $\sabs{\sfracpart{\Phi_n + \sum_{k=0}^m (\nicefrac{n}{N})^k \widehat{\psi}_{k,N}}} \leq \frac{1}{2} - \frac{1}{2N}$ for all $n = 1, \dots, N$.
+% \end{lemma}
+% \begin{IEEEproof}
+% To simplify our notation let 
+% \[
+% B_n = \Phi_n + \sum_{k=1}^m (\nicefrac{n}{N})^k \widehat{\psi}_{k,N}
+% \]
+% so that we now require to prove $\sabs{\sfracpart{B_n + \widehat{\psi}_{0,N}}} \leq \frac{1}{2} - \frac{1}{2N}$
+% for all $n = 1, \dots, N$.  From~\eqref{eq:Wn}, $W_n = \sround{B_n + \widehat{\psi}_{0,N}}$, and 
+% \begin{align*}
+% T_N(\widehat{\psibf}_N) &= \frac{1}{N}\sum_{n=1}^N \sfracpart{B_n + \widehat{\psi}_{0,N}}^2
+% \\ &= \frac{1}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N} - W_n)^2.
+% \end{align*}
+% Since $\widehat{\psi}_{0,N}$ is the minimiser of the quadratic above,
+% \begin{equation}\label{eq:lamsum}
+% \widehat{\psi}_{0,N} = -\frac{1}{N}\sum_{n=1}^N(B_n - W_n).
+% \end{equation}
+% The proof now proceeds by contradiction.  Assume that for some $k$,
+% \begin{equation}\label{eq:philamfraccontra}
+% \sfracpart{B_k + \widehat{\psi}_{0,N}} > \frac{1}{2} - \frac{1}{2N}.
+% \end{equation}
+% Let $F_n = W_n$ for all $n \neq k$ and $F_k = W_k + 1$, and let
+% \[
+% \phi = -\frac{1}{N}\sum_{n=1}^N(B_n - F_n) = \widehat{\psi}_{0,N} + \frac{1}{N}.
+% \]
+% Then,
+% \begin{align}
+% (B_k + &\phi - F_k)^2 = (B_k + \phi - W_k - 1)^2 \nonumber \\
+% &= (B_k + \phi - W_k)^2 - 2(B_k + \phi - W_k) + 1 \nonumber \\
+% &= (B_k + \phi - W_k)^2 - 2(B_k + \widehat{\psi}_{0,N} - W_k) + 1 - \frac{2}{N} \nonumber \\
+% &= (B_k + \phi - W_k)^2 - 2\sfracpart{B_k + \widehat{\psi}_{0,N}} + 1 - \frac{2}{N} \nonumber \\
+% &< (B_k + \phi - W_k)^2 - \frac{1}{N}, \label{eq:Bkineq}
+% \end{align}
+% where the inequality in the last line follows from~\eqref{eq:philamfraccontra}. Let
+% \[
+% \bbf = \left[
+% \begin{array}
+% [c]{cccc}%
+% \phi & \widehat{\psi}_{1,N} & \cdots & \widehat{\psi}_{m,N}
+% \end{array}
+% \right]^\prime
+% \]
+% be the vector of length $m+1$ with components $b_0 = \phi$ and $b_\ell = \widehat{\psi}_{\ell,N}$ for $\ell = 1 , \dots m$.  Now,
+% \[
+% N T_N(\bbf) = \sum_{n=1}^N \fracpart{B_n + \phi }^2 \leq  \sum_{n=1}^N (B_n + \phi  - F_n)^2, 
+% \]
+% and using the inequality from~\eqref{eq:Bkineq},
+% \begin{align*}
+% N &T_N(\bbf) < - \frac{1}{N} + \sum_{n=1}^N (B_n + \phi  - W_n)^2 \\
+% &= - \frac{1}{N} + \sum_{n=1}^N (B_n + \widehat{\psi}_{0,N} + \frac{1}{N}  - W_n)^2 \\
+% &= \sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n)^2 +  \frac{2}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n)\\
+% %&= N T_N(\widehat{\psibf}_N) + \frac{2}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n) \\
+% &= N T_N(\widehat{\psibf}_N),
+% \end{align*}
+% because $\frac{2}{N}\sum_{n=1}^N (B_n + \widehat{\psi}_{0,N}  - W_n) = 0$ as a result of~\eqref{eq:lamsum}.
 But, now $T_N(\bbf) < T_N(\widehat{\psibf}_N)$ violating the fact that $\widehat{\psibf}_N$ is a minimiser of $T_N$.
@@
 
-If $\sfracpart{B_k + \widehat{\psi}_{0,N}} < -\frac{1}{2} + \frac{1}{2N}$ for some $k$, we set $F_k = W_k - 1$ and using the same procedure as before obtain $T_N(\bbf) < T_N(\widehat{\psibf}_N)$ again.
+% If $\sfracpart{B_k + \widehat{\psi}_{0,N}} < -\frac{1}{2} + \frac{1}{2N}$ for some $k$, we set $F_k = W_k - 1$ and using the same procedure as before obtain $T_N(\bbf) < T_N(\widehat{\psibf}_N)$ again.
 The proof follows.
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:Kconvfhalf}
-With $K_{\ell,N}$ defined as in~\eqref{eq:KellN} and $h = f(-\nicefrac{1}{2})$,
-\[
-K_{\ell,N} = (h - 1) \sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big( C_{\ell k} + o_P(1) \big) + o_P(1) 
-\] 
-for all $\ell = 0, \dots, m$, where $C_{\ell k} =  \frac{1}{\ell + k + 1}$.% is an element of the $m+1$ by $m+1$ Hilbert matrix, and $o_P(1)$ denotes a random variable converging in probability to zero as $N\rightarrow\infty$.
-\end{lemma}
-\begin{IEEEproof}
-Care must be taken since $\widehat{\psibf}_N$ depends on the sequence $\{ \Phi_n \}$.  For $n = 1, \dots, N$ and positive $N$, let
-\begin{equation}
-p_{nN}(\psibf) = \sum_{k=0}^{m}\left( \tfrac{n}{N}\right)^k \psi_k,
-\end{equation}
-and put
-\[
-q_{n}(x) = \round{\Phi_n + x}, \qquad Q(x) = \expect q_{n}(x) =  \expect q_{1}(x).
-\]
-Let
-\begin{equation}\label{eq:GNdef}
-G_N(\psibf) = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}\left( \tfrac{n}{N}\right)^\ell  \big(q_n(p_{nN}(\psibf)) - Q(p_{nN}(\psibf)) \big).
-\end{equation}
-and put
-\begin{equation}\label{eq:hatpnN}
-\widehat{p}_{nN} = p_{nN}(\widehat{\psibf}_N) = \sum_{k=0}^m \left(\tfrac{n}{N}\right)^k \widehat{\psi}_{k,N}.
-\end{equation}
-Observe that $G_N$ depends on $\ell$ and we could write $G_{\ell,N}$ but have suppressed the subscript $\ell$ for notational simplicity.
+% \end{IEEEproof}
+
+% \begin{lemma}\label{lem:Kconvfhalf}
+% With $K_{\ell,N}$ defined as in~\eqref{eq:KellN} and $h = f(-\nicefrac{1}{2})$,
+% \[
+% K_{\ell,N} = (h - 1) \sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big( C_{\ell k} + o_P(1) \big) + o_P(1) 
+% \] 
+% for all $\ell = 0, \dots, m$, where $C_{\ell k} =  \frac{1}{\ell + k + 1}$.% is an element of the $m+1$ by $m+1$ Hilbert matrix, and $o_P(1)$ denotes a random variable converging in probability to zero as $N\rightarrow\infty$.
+% \end{lemma}
+% \begin{IEEEproof}
+% Care must be taken since $\widehat{\psibf}_N$ depends on the sequence $\{ \Phi_n \}$.  For $n = 1, \dots, N$ and positive $N$, let
+% \begin{equation}
+% p_{nN}(\psibf) = \sum_{k=0}^{m}\left( \tfrac{n}{N}\right)^k \psi_k,
+% \end{equation}
+% and put
+% \[
+% q_{n}(x) = \round{\Phi_n + x}, \qquad Q(x) = \expect q_{n}(x) =  \expect q_{1}(x).
+% \]
+% Let
+% \begin{equation}\label{eq:GNdef}
+% G_N(\psibf) = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}\left( \tfrac{n}{N}\right)^\ell  \big(q_n(p_{nN}(\psibf)) - Q(p_{nN}(\psibf)) \big).
+% \end{equation}
+% and put
+% \begin{equation}\label{eq:hatpnN}
+% \widehat{p}_{nN} = p_{nN}(\widehat{\psibf}_N) = \sum_{k=0}^m \left(\tfrac{n}{N}\right)^k \widehat{\psi}_{k,N}.
+% \end{equation}
+% Observe that $G_N$ depends on $\ell$ and we could write $G_{\ell,N}$ but have suppressed the subscript $\ell$ for notational simplicity.
 Now $W_n$ from~\eqref{eq:Wn} can be written as $W_n = \round{\Phi_n + \widehat{p}_{nN}} = q_{n}(\widehat{p}_{nN})$
-and $K_{\ell,N}$ from~\eqref{eq:KellN} can be written as
-\begin{align*}
-K_{\ell,N} &= \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\big( q_{n}(\widehat{p}_{nN}) - \widehat{p}_{nN}  \big) \\
-&= \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\big(q_{n}(\widehat{p}_{nN}) - \widehat{p}_{nN} + Q(\widehat{p}_{nN}) - Q(\widehat{p}_{nN}) \big) \\
-&= G_N(\widehat{\psibf}_N) + H_{\ell,N},
-\end{align*}
-where 
-\begin{equation}\label{eq:HellNdef}
-H_{\ell,N} = \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell \big( Q(\widehat{p}_{nN}) - \widehat{p}_{nN} \big).
-\end{equation}
-BLERG (empirical process, tighttess, ref arxiv) that for any $\delta >0$ and $\nu > 0$ there exists an $\epsilon > 0$ such that
-\[
-\prob\left\{ \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf) } > \delta   \right\} < \nu
-\]
-for all positive integers $N$, where $\|\psibf\|_\infty = \sup_{k} \abs{\psi_k}$.  Since $\widehat{\psibf}_N$ converges almost surely to zero, it follows that for any $\epsilon > 0$,
-\[
-\lim_{N\rightarrow\infty}\prob\left\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} = 0
-\] 
-and therefore $\prob\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \} < \nu$ for all sufficiently large $N$.  Now
-\begin{align*}
-  \prob&\left\{\abs{ G_N(\widehat{\psibf}_N) } > \delta \right\} \\
-&= \prob\left\{ \abs{G_N(\widehat{\psibf}_N)} > \delta \;, \; \|\widehat{\psibf}_N\|_{\infty} < \epsilon \right\} \\
-&\hspace{0.7cm} + \prob\left\{ \abs{G_N(\widehat{\psibf}_N)} > \delta  \;, \; \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} \\
-&\leq \prob\left\{  \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf) } > \delta \right\} + \prob\left\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} \\
-&\leq 2\nu
-\end{align*}
-for all sufficiently large $N$.
+% and $K_{\ell,N}$ from~\eqref{eq:KellN} can be written as
+% \begin{align*}
+% K_{\ell,N} &= \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\big( q_{n}(\widehat{p}_{nN}) - \widehat{p}_{nN}  \big) \\
+% &= \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell\big(q_{n}(\widehat{p}_{nN}) - \widehat{p}_{nN} + Q(\widehat{p}_{nN}) - Q(\widehat{p}_{nN}) \big) \\
+% &= G_N(\widehat{\psibf}_N) + H_{\ell,N},
+% \end{align*}
+% where 
+% \begin{equation}\label{eq:HellNdef}
+% H_{\ell,N} = \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell \big( Q(\widehat{p}_{nN}) - \widehat{p}_{nN} \big).
+% \end{equation}
+% BLERG (empirical process, tighttess, ref arxiv) that for any $\delta >0$ and $\nu > 0$ there exists an $\epsilon > 0$ such that
+% \[
+% \prob\left\{ \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf) } > \delta   \right\} < \nu
+% \]
+% for all positive integers $N$, where $\|\psibf\|_\infty = \sup_{k} \abs{\psi_k}$.  Since $\widehat{\psibf}_N$ converges almost surely to zero, it follows that for any $\epsilon > 0$,
+% \[
+% \lim_{N\rightarrow\infty}\prob\left\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} = 0
+% \] 
+% and therefore $\prob\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \} < \nu$ for all sufficiently large $N$.  Now
+% \begin{align*}
+%   \prob&\left\{\abs{ G_N(\widehat{\psibf}_N) } > \delta \right\} \\
+% &= \prob\left\{ \abs{G_N(\widehat{\psibf}_N)} > \delta \;, \; \|\widehat{\psibf}_N\|_{\infty} < \epsilon \right\} \\
+% &\hspace{0.7cm} + \prob\left\{ \abs{G_N(\widehat{\psibf}_N)} > \delta  \;, \; \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} \\
+% &\leq \prob\left\{  \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ G_N(\psibf) } > \delta \right\} + \prob\left\{ \|\widehat{\psibf}_N\|_{\infty} \geq \epsilon \right\} \\
+% &\leq 2\nu
+% \end{align*}
+% for all sufficiently large $N$.
 Since $\nu$ and $\delta$ can be chosen arbitrarily small, it follows that $G_N(\widehat{\psibf}_N)$ converges in probability to zero as $N\rightarrow\infty$, and therefore $K_{\ell,N} = H_{\ell,N} + o_P(1)$.
 Lemma~\ref{lem:sumpNhilb} shows that
-\begin{align*}
-H_{\ell,N} =  (h-1)\sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big(C_{\ell k} + o_P(1)\big).
-\end{align*}
-The proof follows.
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:sumpNhilb}
-With $H_{\ell,N}$ defined in~\eqref{eq:HellNdef} and $\widehat{p}_{nN}$ defined in~\eqref{eq:hatpnN}, and with $h = f(-\nicefrac{1}{2})$,
-\[
-H_{\ell,N} = (h-1)\sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big(C_{\ell k} + o_P(1)\big),
-\]
-where $C_{\ell k} = \tfrac{1}{\ell + k + 1}$.
-\end{lemma}
-\begin{IEEEproof}
-If $\abs{x} < 1$, then
-\[
-q_n(x)  = \round{\Phi_n + x} = \begin{cases}
-1, & \Phi_n + x  \geq \nicefrac{1}{2} \\
--1, & \Phi_n + x  < -\nicefrac{1}{2} \\
-0, & \text{otherwise},
-\end{cases}
-\]
-and,
-\begin{align*}
-Q(x) = Eq_1(x) &=  \begin{cases}
-\int_{\nicefrac{1}{2} -x }^{\nicefrac{1}{2}}f(t)\,dt, &   x \geq 0 \\
--\int_{-\nicefrac{1}{2}}^{-\nicefrac{1}{2} - x }f(t)\,dt, &  x  < 0.
-\end{cases} 
-% \\
-% &= \begin{cases}
-% 1 - F(\nicefrac{1}{2} - x ), & x  \geq 0 \\
-% -F(-\nicefrac{1}{2} - x ), &  x < 0,
+% \begin{align*}
+% H_{\ell,N} =  (h-1)\sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big(C_{\ell k} + o_P(1)\big).
+% \end{align*}
+% The proof follows.
+% \end{IEEEproof}
+
+% \begin{lemma}\label{lem:sumpNhilb}
+% With $H_{\ell,N}$ defined in~\eqref{eq:HellNdef} and $\widehat{p}_{nN}$ defined in~\eqref{eq:hatpnN}, and with $h = f(-\nicefrac{1}{2})$,
+% \[
+% H_{\ell,N} = (h-1)\sqrt{N} \sum_{k=0}^{m}  \widehat{\psi}_{k,N} \big(C_{\ell k} + o_P(1)\big),
+% \]
+% where $C_{\ell k} = \tfrac{1}{\ell + k + 1}$.
+% \end{lemma}
+% \begin{IEEEproof}
+% If $\abs{x} < 1$, then
+% \[
+% q_n(x)  = \round{\Phi_n + x} = \begin{cases}
+% 1, & \Phi_n + x  \geq \nicefrac{1}{2} \\
+% -1, & \Phi_n + x  < -\nicefrac{1}{2} \\
+% 0, & \text{otherwise},
 % \end{cases}
-\end{align*}
-Because $f(\fracpart{x})$ is continuous at $-\nicefrac{1}{2}$ it follows that 
-\[
-Q(x) = x \big( h + \zeta(x) \big),
-\] 
-where $\zeta(x)$ is a function that converges to zero as $x$ converges to zero.  
-%The equation above is still valid when $\abs{x} \geq 1$ by an appropriate choice of the function $\zeta(x)$.  
-Observe that $\abs{\widehat{p}_{nN}} \leq \sum_{k=0}^m\sabs{ \widehat{\psi}_{k,N}}$ and, since each of the $\widehat{\psi}_{k,N} \rightarrow 0$ almost surely as $N\rightarrow\infty$, it follows that $\widehat{p}_{nN} \rightarrow 0$ almost surely uniformly in $n = 1, \dots, N$ as $N\rightarrow\infty$.
+% \]
+% and,
+% \begin{align*}
+% Q(x) = Eq_1(x) &=  \begin{cases}
+% \int_{\nicefrac{1}{2} -x }^{\nicefrac{1}{2}}f(t)\,dt, &   x \geq 0 \\
+% -\int_{-\nicefrac{1}{2}}^{-\nicefrac{1}{2} - x }f(t)\,dt, &  x  < 0.
+% \end{cases} 
+% % \\
+% % &= \begin{cases}
+% % 1 - F(\nicefrac{1}{2} - x ), & x  \geq 0 \\
+% % -F(-\nicefrac{1}{2} - x ), &  x < 0,
+% % \end{cases} 
+% \end{align*}
+% Because $f(\fracpart{x})$ is continuous at $-\nicefrac{1}{2}$ it follows that 
+% \[
+% Q(x) = x \big( h + \zeta(x) \big),
+% \] 
+% where $\zeta(x)$ is a function that converges to zero as $x$ converges to zero.  
+% %The equation above is still valid when $\abs{x} \geq 1$ by an appropriate choice of the function $\zeta(x)$.  
+% Observe that $\abs{\widehat{p}_{nN}} \leq \sum_{k=0}^m\sabs{ \widehat{\psi}_{k,N}}$ and, since each of the $\widehat{\psi}_{k,N} \rightarrow 0$ almost surely as $N\rightarrow\infty$, it follows that $\widehat{p}_{nN} \rightarrow 0$ almost surely uniformly in $n = 1, \dots, N$ as $N\rightarrow\infty$.
 Thus $\zeta(\widehat{p}_{nN}) \rightarrow 0$ almost surely (and therefore also in probability) uniformly in $n = 1, \dots, N$ as $N\rightarrow\infty$.
 Now,
-\begin{align*}
-Q(\widehat{p}_{nN}) - \widehat{p}_{nN} &= \widehat{p}_{nN}\big( h - 1 + \zeta(\widehat{p}_{nN}) \big) \\
-&= \widehat{p}_{nN}\big( h - 1 + o_P(1) \big), 
-\end{align*}
-and so,
-\begin{align*}
-H_{\ell,N} &= \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell \widehat{p}_{nN}\big( h - 1 + o_P(1) \big)  \\
-&=  \frac{1}{\sqrt{N}} \sum_{n=1}^{N}(\tfrac{n}{N})^\ell \sum_{k=0}^m (\tfrac{n}{N})^k \widehat{\psi}_{k,N} \big( h - 1 + o_P(1) \big)   \\
-&= \sqrt{N}\sum_{k=0}^m \widehat{\psi}_{k,N} \frac{1}{N} \sum_{n=1}^{N} \frac{n^{\ell + k}}{N^{\ell + k + 1}}\big( h - 1 + o_P(1)  \big).
-\end{align*}
-The Riemann sum
-\[
-\frac{1}{N} \sum_{n=1}^{N} \frac{n^{\ell + k}}{N^{\ell + k + 1}} = \int_{0}^{1}x^{k+\ell+1}dx + o_P(1),
-\]
-and since the integral above evaluates to $\frac{1}{k+\ell+1}$, we have
-\[
-H_{\ell,N} = (h - 1)\sqrt{N}\sum_{k=0}^m \widehat{\psi}_{k,N}\left( \frac{1}{k+\ell+1} + o_P(1)  \right)
-\]
-as required.
-% Put
-% \[
-% C_{\ell + k,N} = \sum_{n=1}^{N}\frac{n^{\ell + k}}{N^{\ell + k+1}}.
-% \]
-% It is well known (see for example~\cite[eq. 21]{Peleg1991_CRB_PPS_1991}) that,
-% \begin{equation}\label{eq:ClkNHilbert}
-% C_{\ell + k,N} = C_{\ell k} + O(N^{-1}).
-% \end{equation}
-% Since $\zeta(\widehat{p}_{nN}) = o_P(1)$ for all $n = 1, \dots, N$ then
-% \[
-% \sum_{n=1}^{N}\frac{n^{\ell + k}}{N^{\ell + k + 1}}\zeta(\widehat{p}_{nN}) 
-% \]
-% converges in probability to zero as $N\rightarrow\infty$.  So,
 % \begin{align*}
-% \sum_{n=1}^{N}\frac{n^{\ell + k}}{N^{\ell + k + 1}}\big( h - 1 +  \zeta(\widehat{p}_{nN}) \big) = (h-1)\big( C_{\ell k} + o_P(1) \big).
+% Q(\widehat{p}_{nN}) - \widehat{p}_{nN} &= \widehat{p}_{nN}\big( h - 1 + \zeta(\widehat{p}_{nN}) \big) \\
+% &= \widehat{p}_{nN}\big( h - 1 + o_P(1) \big), 
 % \end{align*}
-% The proof follows by substituting this into~\eqref{eq:HellNre}.
-\end{IEEEproof}
-
-\begin{lemma}\label{eq:convdn}
-The distribution of the vector $\dbf_N$, defined in~\eqref{eq:dnandkn}, converges to the multivariate normal with zero mean and covariance matrix $\sigma^2\Cbf$.
-\end{lemma}
-\begin{IEEEproof}
-For any constant vector $\alphabf$, let
-\[
-z_N = \alphabf^\prime \dbf_N = \frac{1}{\sqrt{N}} \sum_{n=1}^{N} \Phi_n \sum_{\ell=0}^{m}\alpha_\ell\left( \frac{n}{N} \right)^\ell.
-\]
-Observe that $\Phi_n$ has zero mean and variance $\sigma^2$ as a result of Proposition~\ref{prop:zerointmeanzeromean}.
+% and so,
+% \begin{align*}
+% H_{\ell,N} &= \frac{1}{\sqrt{N}}\sum_{n=1}^{N}(\tfrac{n}{N})^\ell \widehat{p}_{nN}\big( h - 1 + o_P(1) \big)  \\
+% &=  \frac{1}{\sqrt{N}} \sum_{n=1}^{N}(\tfrac{n}{N})^\ell \sum_{k=0}^m (\tfrac{n}{N})^k \widehat{\psi}_{k,N} \big( h - 1 + o_P(1) \big)   \\
+% &= \sqrt{N}\sum_{k=0}^m \widehat{\psi}_{k,N} \frac{1}{N} \sum_{n=1}^{N} \frac{n^{\ell + k}}{N^{\ell + k + 1}}\big( h - 1 + o_P(1)  \big).
+% \end{align*}
+% The Riemann sum
+% \[
+% \frac{1}{N} \sum_{n=1}^{N} \frac{n^{\ell + k}}{N^{\ell + k + 1}} = \int_{0}^{1}x^{k+\ell+1}dx + o_P(1),
+% \]
+% and since the integral above evaluates to $\frac{1}{k+\ell+1}$, we have
+% \[
+% H_{\ell,N} = (h - 1)\sqrt{N}\sum_{k=0}^m \widehat{\psi}_{k,N}\left( \frac{1}{k+\ell+1} + o_P(1)  \right)
+% \]
+% as required.
+% % Put
+% % \[
+% % C_{\ell + k,N} = \sum_{n=1}^{N}\frac{n^{\ell + k}}{N^{\ell + k+1}}.
+% % \]
+% % It is well known (see for example~\cite[eq. 21]{Peleg1991_CRB_PPS_1991}) that,
+% % \begin{equation}\label{eq:ClkNHilbert}
+% % C_{\ell + k,N} = C_{\ell k} + O(N^{-1}).
+% % \end{equation}
+% % Since $\zeta(\widehat{p}_{nN}) = o_P(1)$ for all $n = 1, \dots, N$ then
+% % \[
+% % \sum_{n=1}^{N}\frac{n^{\ell + k}}{N^{\ell + k + 1}}\zeta(\widehat{p}_{nN}) 
+% % \]
+% % converges in probability to zero as $N\rightarrow\infty$.  So,
+% % \begin{align*}
+% % \sum_{n=1}^{N}\frac{n^{\ell + k}}{N^{\ell + k + 1}}\big( h - 1 +  \zeta(\widehat{p}_{nN}) \big) = (h-1)\big( C_{\ell k} + o_P(1) \big).
+% % \end{align*}
+% % The proof follows by substituting this into~\eqref{eq:HellNre}.
+% \end{IEEEproof}
+
+% \begin{lemma}\label{eq:convdn}
+% The distribution of the vector $\dbf_N$, defined in~\eqref{eq:dnandkn}, converges to the multivariate normal with zero mean and covariance matrix $\sigma^2\Cbf$.
+% \end{lemma}
+% \begin{IEEEproof}
+% For any constant vector $\alphabf$, let
+% \[
+% z_N = \alphabf^\prime \dbf_N = \frac{1}{\sqrt{N}} \sum_{n=1}^{N} \Phi_n \sum_{\ell=0}^{m}\alpha_\ell\left( \frac{n}{N} \right)^\ell.
+% \]
+% Observe that $\Phi_n$ has zero mean and variance $\sigma^2$ as a result of Proposition~\ref{prop:zerointmeanzeromean}.
 By Lypanov's central limit theorem $\zbf_N$ is asymptotically normally distributed with zero mean and variance
-\[
-\lim_{N\rightarrow\infty} \sigma^2 \frac{1}{N} \sum_{n=1}^{N}\left( \sum_{\ell=0}^{m}\alpha_\ell\left( \frac{n}{N} \right)^\ell \right)^2 = \sigma^2 \alphabf^\prime \Cbf \alphabf.
-\]
-By the Cram\'{e}r-Wold theorem it follows that $\dbf_N$ is asymptotically normally distributed with zero mean and covariance $\sigma^2 \Cbf$. 
-% Recall that the $\ell$th element of $\dbf_N$ is
-% \[
-% D_{\ell,N} = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}(\tfrac{n}{N})^\ell \Phi_n.
-% \]
-% Since the $\Phi_1, \dots, \Phi_N$ are independent, each with variance $\sigma^2$, the distribution of $D_{\ell,N}$ converges to the normal with zero mean and variance
-% \[
-% \lim_{N\rightarrow\infty}\var D_{\ell,N} = \sigma^2 \lim_{N\rightarrow\infty} C_{2\ell,N} = \sigma^2 C_{\ell\ell}
-% \]
-% by Lypanov's central limit theorem and by using~\eqref{eq:ClkNHilbert}.  The covariance between $D_{\ell,N}$ and $D_{k,N}$ is
-% \[
-% \covar(D_{\ell,N}, D_{k,N}) = \sigma^2 C_{\ell+k,N} \rightarrow \sigma^2 C_{\ell k}
-% \]
-% as $N \rightarrow \infty$, again using~\eqref{eq:ClkNHilbert}.  The proof follows.
-\end{IEEEproof}
+% \[
+% \lim_{N\rightarrow\infty} \sigma^2 \frac{1}{N} \sum_{n=1}^{N}\left( \sum_{\ell=0}^{m}\alpha_\ell\left( \frac{n}{N} \right)^\ell \right)^2 = \sigma^2 \alphabf^\prime \Cbf \alphabf.
+% \]
+% By the Cram\'{e}r-Wold theorem it follows that $\dbf_N$ is asymptotically normally distributed with zero mean and covariance $\sigma^2 \Cbf$. 
+% % Recall that the $\ell$th element of $\dbf_N$ is
+% % \[
+% % D_{\ell,N} = \frac{1}{\sqrt{N}} \sum_{n=1}^{N}(\tfrac{n}{N})^\ell \Phi_n.
+% % \]
+% % Since the $\Phi_1, \dots, \Phi_N$ are independent, each with variance $\sigma^2$, the distribution of $D_{\ell,N}$ converges to the normal with zero mean and variance
+% % \[
+% % \lim_{N\rightarrow\infty}\var D_{\ell,N} = \sigma^2 \lim_{N\rightarrow\infty} C_{2\ell,N} = \sigma^2 C_{\ell\ell}
+% % \]
+% % by Lypanov's central limit theorem and by using~\eqref{eq:ClkNHilbert}.  The covariance between $D_{\ell,N}$ and $D_{k,N}$ is
+% % \[
+% % \covar(D_{\ell,N}, D_{k,N}) = \sigma^2 C_{\ell+k,N} \rightarrow \sigma^2 C_{\ell k}
+% % \]
+% % as $N \rightarrow \infty$, again using~\eqref{eq:ClkNHilbert}.  The proof follows.
+% \end{IEEEproof}
 

### 2013-02-05 code/java/src/pubsim/poly/AmbiguityRemover.java

@@ -1,0 +1,110 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+package pubsim.poly;
+
+import Jama.Matrix;
+import java.io.Serializable;
+import pubsim.VectorFunctions;
+import pubsim.lattices.GeneralLattice;
+import pubsim.lattices.NearestPointAlgorithm;
+import pubsim.lattices.decoder.SphereDecoder;
+
+/**
+ * This uses m nearest lattice point approach to remove the
+ * ambiguities inherent in polynomial phase estimation.
+ * @author Robby McKilliam
+ */
+public class AmbiguityRemover implements Serializable{
+
+    protected int m;
+    protected Matrix M;
+    double[] p;
+    NearestPointAlgorithm np;
+
+    protected AmbiguityRemover() {
+    }
+
+    /**
+     * Public constructor.  Set the order of the polynomial phase signal.
+     * @param m = the number of parameters ie. order of polynomial + 1
+     */
+    public AmbiguityRemover(int m) {
+        this.m = m;
+        p = new double[m+1];
+        M = constructBasisMatrix();
+        GeneralLattice lattice = new GeneralLattice(M);
+        np = new SphereDecoder(lattice);
+    }
+    
+    protected final Matrix constructBasisMatrix() {
+        Matrix B = new Matrix(m+1, m+1);
+        double[] c = VectorFunctions.eVector(0, 1);
+        for (int j = 0; j < m+1; j++) {
+            for (int i = 0; i <= j; i++) {
+                B.set(i, j, c[i]);
+            }
+            c = getNextColumn(c);
+        }
+
+        //System.out.println("M = " + VectorFunctions.print(B));
+
+        return B;
+    }
+
+    /**
+     * Recursively generates the columns of the ambiguity lattice
+     * generator matrix.  Essentially generates coefficients of the
+     * integer valued polynomials.
+     * @param c previous column.
+     * @return next column
+     */
+    protected static double[] getNextColumn(double[] c) {
+        double[] r = new double[c.length + 1];
+
+        //shift c
+        for (int i = 1; i < r.length; i++) {
+            r[i] = c[i - 1];
+        }
+
+        for (int i = 0; i < c.length; i++) {
+            r[i] += c[i] * (c.length - 1);
+        }
+
+        for (int i = 0; i < r.length; i++) {
+            r[i] /= c.length;
+        }
+        return r;
+    }
+
+    /**
[…]

### 2013-02-05 code/java/src/pubsim/poly/AmbiguityRemoverRectangular.java

@@ -1,0 +1,23 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+package pubsim.poly;
+
+import pubsim.lattices.GeneralLattice;
+import pubsim.lattices.decoder.BabaiNoLLL;
+
+/**
+ *
+ * @author Robby McKilliam
+ */
+public class AmbiguityRemoverRectangular extends AmbiguityRemover {
+
+    public AmbiguityRemoverRectangular(int m) {
+        this.m = m;
+        p = new double[m+1];
+        M = constructBasisMatrix();
+        GeneralLattice lattice = new GeneralLattice(M);
+        np = new BabaiNoLLL(lattice);
+    }
+}

### 2013-02-05 code/java/src/pubsim/poly/BabaiEstimator.java

@@ -1,0 +1,96 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+package pubsim.poly;
+
+import Jama.Matrix;
+import pubsim.VectorFunctions;
+import pubsim.lattices.NearestPointAlgorithm;
+import pubsim.lattices.VnmStar;
+import pubsim.lattices.VnmStarGlued;
+import pubsim.lattices.decoder.Babai;
+
+/**
+ * Uses the Babai nearest plane algorithm
+ * @author Robby
+ */
+public class BabaiEstimator implements PolynomialPhaseEstimator {
+
+    protected double[] ya,  p;
+    protected int n,  m;
+    protected VnmStar lattice;
+    protected NearestPointAlgorithm npalgorithm;
+    protected Matrix M,  K;
+    protected AmbiguityRemover ambiguityRemover;
+    
+    protected BabaiEstimator() {}
+    
+    /** 
+     * You must set the polynomial order in the constructor
+     * @param m = polynomial order
+     */
+    public BabaiEstimator(int m, int n) {
+        lattice = new VnmStarGlued(m, n - m - 1);
+        this.m = m;
+        npalgorithm = new Babai(lattice);
+        ambiguityRemover = new AmbiguityRemover(m);
+
+        ya = new double[n];
+        p = new double[m+1];
+        this.n = n;
+
+        M = lattice.getMMatrix();
+        Matrix Mt = M.transpose();
+        K = Mt.times(M).inverse().times(Mt);
+    }
+
+    @Override
+    public double[] estimate(double[] real, double[] imag) {
+        if(n != real.length) throw new RuntimeException("Data length does not equal " + n);
+        
+        for (int i = 0; i < real.length; i++) {
+            ya[i] = Math.atan2(imag[i], real[i]) / (2 * Math.PI);
+        }
+        npalgorithm.nearestPoint(ya);
+        double[] u = npalgorithm.getIndex();
+
+        double[] ymu = new double[ya.length];
+        for (int i = 0; i < u.length; i++) {
+            ymu[i] = ya[i] - u[i];
+        }
+        System.arraycopy(ya, u.length, ymu, u.length, ya.length - u.length);
+
+        //compute the parameters
+        VectorFunctions.matrixMultVector(K, ymu, p); 
+        
+        return ambiguityRemover.disambiguate(p);
+    }
+
+
+    /**
+     * Run the estimator and return the square error wrapped modulo the the ambiguity
[…]

### 2013-02-05 code/java/src/pubsim/poly/CircularNoisePolynomialPhaseSignal.java

@@ -1,0 +1,45 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+
+package pubsim.poly;
+
+import pubsim.distributions.RealRandomVariable;
+import pubsim.distributions.circular.CircularRandomVariable;
+
+/**
+ * Like PolynommialPhaseSignal but the noise gets added to the phase.
+ * @author Robby McKilliam
+ */
+public class CircularNoisePolynomialPhaseSignal extends PolynomialPhaseSignal{
+
+    Double[] phasearray = new Double[0];
+
+    public CircularNoisePolynomialPhaseSignal(int N){
+        super(N);
+    }
+    
+    /**
+     * Returns the phase signal and places the real and imaginary parts
+     * of the complex signal into real and imag arrays that can be retrieved
+     * with getReal() and getImag()
+     * @return
+     */
+    @Override
+    public Double[] generateReceivedSignal() {
+        if(phasearray.length != n) phasearray = new Double[n];
+
+        for(int t = 0; t < n; t++){
+            double phase = 0.0;
+            for(int j = 0; j < params.length; j++)
+                phase += Math.pow(t+1,j)*params[j];
+            double pnoise = noise.getNoise();
+            phasearray[t] = phase + pnoise;
+            real[t] = Math.cos(2*Math.PI*(phase + pnoise));
+            imag[t] = Math.sin(2*Math.PI*(phase + pnoise));
+        }
+        return phasearray;
+    }
+
+}

### 2013-02-05 code/java/src/pubsim/poly/DPTEstimator.java

@@ -1,0 +1,267 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+package pubsim.poly;
+
+import flanagan.math.FourierTransform;
+import flanagan.complex.Complex;
+import pubsim.Util;
+import pubsim.VectorFunctions;
+
+/**
+ * Implementation of the estimator based on iteratively maximising
+ * the Discrete Polynomial Transform.  See Peleg and Friedlander,
+ * Trans. Sig. Proc. Vol 43 August 1995.
+ * @author Robby McKilliam
+ */
+public class DPTEstimator implements PolynomialPhaseEstimator {
+
+    protected Complex[] z;
+    protected double[] p;
+    protected int m,  n;
+    protected double tau;
+    protected int num_samples;
+    protected FourierTransform fft;
+    protected Complex[] sig;
+    
+    /**Max number of iterations for the Newton step */
+    static final int MAX_ITER = 15;
+    /**Step variable for the Newton step */
+    static final double EPSILON = 1e-10;
+
+    protected AmbiguityRemover ambiguityRemover;
+
+    protected DPTEstimator() {}
+
+    public DPTEstimator(int m, int n) {
+        this.m = m;
+        ambiguityRemover = new AmbiguityRemover(m);
+        z = new Complex[n];
+        p = new double[m+1];
+        this.n = n;
+        num_samples = 4 * n;
+
+        //set the tau parameter for the PPT
+        tau = Math.round(((double) n) / (m));
+        if (m > 4) {
+            tau = Math.round(((double) n) / (m + 2));
+         }
+
+        int oversampled = 4;
+        sig = new Complex[FourierTransform.nextPowerOfTwo(oversampled * n)];
+        fft = new FourierTransform();
+       // tau = Math.round(0.2 * n);
+    }
+
+    @Override
+    public double[] estimate(double[] real, double[] imag) {
+        if(n != real.length) throw new RuntimeException("Data length does not equal " + n);
+        
+        for (int i = 0; i < n; i++) {
+            z[i] = new Complex(real[i], imag[i]);
+        }
+
+//        System.out.println(VectorFunctions.print(real));
+//        System.out.println(VectorFunctions.print(imag));
+//        System.out.println(VectorFunctions.print(z));
+
+        for (int i = m; i >= 0; i--) {
+            p[i] = estimateM(z, i);
+            for (int j = 0; j < z.length; j++) {
[…]

### 2013-02-05 code/java/src/pubsim/poly/DPToversampled.java

@@ -1,0 +1,40 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+
+package pubsim.poly;
+
+/**
+ * The is the DPT estimator that has been oversampled so that the
+ * volume of parameters is equal to that of the identifiable region.
+ * @author Robby McKilliam
+ */
+public class DPToversampled extends DPTEstimator {
+
+    @Override
+    public double[] error(double[] real, double[] imag, double[] truth) {
+
+        //this computes the 'fair' error after oversampling
+        double dptos = Math.pow( (n*1.0)/m , (m-1.0)/(m+1.0));
+
+        double[] est = estimate(real, imag);
+        double[] err = new double[est.length];
+
+        for (int i = 0; i < err.length; i++) {
+            double oscale = Math.pow(dptos,i);
+            double oserr = oscale*(est[i] - truth[i]);
+            double osr = 1.0;
+            if( i > 1 ) osr = oscale*Math.pow(m/(n*1.0), i-1)/pubsim.Util.factorial(i);
+
+            err[i] = osr*pubsim.Util.fracpart(oserr/osr);
+        }
+
+        //err = ambiguityRemover.disambiguate(err);
+        for (int i = 0; i < err.length; i++) {
+            err[i] *= err[i];
+        }
+        return err;
+    }
+
+}

### 2013-02-05 code/java/src/pubsim/poly/HPFEstimator.java

@@ -1,0 +1,49 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+
+package pubsim.poly;
+
+import pubsim.Util;
+
+/**
+ * Implementation of the High-order phase function estimator.
+ * UNDER CONSTRUCTION!
+ * @author robertm
+ */
+public class HPFEstimator implements PolynomialPhaseEstimator {
+
+    public void setSize(int n) {
+        throw new UnsupportedOperationException("Not supported yet.");
+    }
+
+    public int getOrder() {
+        throw new UnsupportedOperationException("Not supported yet.");
+    }
+
+    public double[] estimate(double[] real, double[] imag) {
+        throw new UnsupportedOperationException("Not supported yet.");
+    }
+
+    public double[] error(double[] real, double[] imag, double[] truth) {
+        throw new UnsupportedOperationException("Not supported yet.");
+    }
+
+    /**
+     * Returns the volume of the functional regions of
+     * the HPF estimator.  This is just for comparison
+     * with the identifiable region.
+     * @return volume of functional region
+     */
+    public static double volumeOfFunctionalRegion(int n, int m){
+        double prod = 1.0;
+        for(int k = 2; k <= m; k++){
+            prod *= 1.0/( Util.factorial(k) * Math.pow(n/2.0, k-1) );
+        }
+        return prod;
+    }
+
+
+
+}

### 2013-02-05 code/java/src/pubsim/poly/KitchenEstimator.java

@@ -1,0 +1,117 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+
+package pubsim.poly;
+
+import pubsim.Util;
+import pubsim.VectorFunctions;
+import pubsim.bearing.SampleCircularMean;
+
+/**
+ * Implementation of Kitchen's polynomial phase estimate.
+ *
+ * J. Kitchen, "A method for estimating the coefficients of a polynomial
+ * phase signal", Signal Processing, vol 37, 1994.
+ * 
+ * This is essentially a generalisation of Kay's frequency estimator.
+ * @author Robby McKilliam
+ */
+public class KitchenEstimator implements PolynomialPhaseEstimator{
+
+    protected Double[] y;
+    protected double[] p;
+    final protected int m, N;
+    
+    protected AmbiguityRemover ambiguityRemover;
+
+    public KitchenEstimator(int m, int N) {
+        this.m = m;
+        ambiguityRemover = new AmbiguityRemover(m);
+        this.N = N;
+        y = new Double[N];
+        p = new double[m+1];
+    }
+
+    public int getOrder() {
+        return m;
+    }
+
+    public double[] estimate(double[] real, double[] imag) {
+        if(N != real.length) throw new RuntimeException("Data length does not equal " + N);
+
+        for(int i = 0; i < N; i++)
+            y[i] = Math.atan2(imag[i], real[i])/(2*Math.PI);
+
+        for (int i = m; i >= 0; i--) {
+            p[i] = estimateM(y, i);
+            for (int j = 0; j < y.length; j++) {
+                y[j] -= p[i]*Math.pow(j+1, i);
+            }
+        }
+        
+        return ambiguityRemover.disambiguate(p);
+    }
+
+    /** Estimate the Mth order parameter */
+    protected double estimateM(Double[] y, int M){
+
+        //if it's the last parameter just run sample circular mean
+        //otherwise the estimator needs the phase parameter bounded,
+        //which is really silly.
+        if(M == 0){
+            SampleCircularMean vm = new SampleCircularMean();
+            return vm.estimateBearing(y);
+        }
+
+        double[] d = VectorFunctions.mthDifference(y, M);
+
+        //System.out.println("y = " + VectorFunctions.print(y));
+        //System.out.println("d = " + VectorFunctions.print(d));
+
+        //compute the denominator
+        double dprod = 1.0;
[…]

### 2013-02-05 code/java/src/pubsim/poly/MaximumLikelihood.java

@@ -1,0 +1,305 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+
+package pubsim.poly;
+
+import Jama.Matrix;
+import pubsim.lattices.util.PointInParallelepiped;
+import pubsim.optimisation.AutoDerivativeFunction;
+import pubsim.optimisation.FunctionAndDerivatives;
+import pubsim.optimisation.NewtonRaphson;
+import pubsim.Complex;
+import pubsim.VectorFunctions;
+import pubsim.fes.PeriodogramFFTEstimator;
+
+/**
+ * Implements m (approximate) maximum likelihood estimator for
+ * polynomial phase signals.  This samples the identifiable region
+ * performing Newton's method m whole lot of times.
+ * @author Robby McKilliam
+ */
+public class MaximumLikelihood implements PolynomialPhaseEstimator{
+
+    protected int m;
+    protected int N;
+
+    protected AmbiguityRemover ambiguityRemover;
+    protected PeriodogramFFTEstimator freqest;
+
+    int samples[];
+    double[] realp, imagp;
+    Complex[] z;
+
+    protected MaximumLikelihood() {}
+    
+    /**
+     * @param m : polynomial order
+     * @param samples : number of samples used per parameter in ML search.
+     * Default samples = 100
+     */
+    public MaximumLikelihood(int m, int n){
+        this.m = m;
+        this.samples = new int[m-1];
+        ambiguityRemover = new AmbiguityRemover(m);
+        N = n;
+        realp = new double[N];
+        imagp = new double[N];
+        z = new Complex[N];
+        freqest = new PeriodogramFFTEstimator(N);
+        for(int i = 0; i < samples.length; i++)
+            samples[i] = (int)Math.round(4*Math.pow( N, i+2 ));
+    }
+    
+    /**
+     * @param samples : number of samples used per parameter in ML search.
+     */
+    public void setSamples(int[] samples){
+        this.samples = samples;
+    }
+
+    @Override
+    public int getOrder() {
+        return m;
+    }
+
+    @Override
+    public double[] estimate(double[] real, double[] imag) {
+        if(N != real.length) throw new RuntimeException("Data length does not equal " + N);
+
+        PolynomialPhaseLikelihood func
+                = new PolynomialPhaseLikelihood(real, imag);
+        NewtonRaphson newtonRaphson
+                = new NewtonRaphson(func);
+
+        PointInParallelepiped points
[…]

### 2013-02-05 code/java/src/pubsim/poly/MbestEstimator.java

@@ -1,0 +1,36 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+
+package pubsim.poly;
+
+import Jama.Matrix;
+import pubsim.lattices.VnmStarGlued;
+import pubsim.lattices.decoder.Mbest;
+
+/**
+ *
+ * @author Robby McKilliam
+ */
+public class MbestEstimator extends BabaiEstimator {
+
+    /**
+     * You must set the polynomial order in the constructor
+     * @param m = polynomial order
+     */
+    public MbestEstimator(int m, int n, int M) {
+        lattice = new VnmStarGlued(m, n - m - 1);
+        this.m = m;
+        npalgorithm = new Mbest(lattice,M);
+        ambiguityRemover = new AmbiguityRemover(m);
+
+        ya = new double[n];
+        p = new double[m+1];
+        this.n = n;
+
+        this.M = lattice.getMMatrix();
+        Matrix Mt = this.M.transpose();
+        K = Mt.times(this.M).inverse().times(Mt);
+    }
+}

### 2013-02-05 code/java/src/pubsim/poly/PolynomialPhaseEstimator.java

@@ -1,0 +1,35 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+
+package pubsim.poly;
+
+import java.io.Serializable;
+
+/**
+ * 
+ * @author Robby McKilliam
+ */
+public interface PolynomialPhaseEstimator extends Serializable{
+
+    /** 
+     * Return the order of this estimator.
+     * This is the number of parameters.
+     */
+    public int getOrder();
+    
+    /** 
+     * Run the estimator on received real and imaginary signal.
+     * Returns an array of polynomial parameters.
+     */
+    public double[] estimate(double[] real, double[] imag);
+
+    /**
+     * Run the estimator and return the square error
+     * wrapped modulo the the ambiguity region between
+     * the estimate and the truth.
+     */
+    public double[] error(double[] real, double[] imag, double[] truth);
+    
+}

### 2013-02-05 code/java/src/pubsim/poly/PolynomialPhaseSignal.java

@@ -1,0 +1,153 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+
+package pubsim.poly;
+
+import java.util.Random;
+import java.io.Serializable;
+import pubsim.SignalGenerator;
+import pubsim.VectorFunctions;
+import pubsim.distributions.NoiseGenerator;
+
+/**
+ * Generates m polynomial phase signal with parameters specified
+ * by the setParameters method.
+ * @author Robby McKilliam
+ */
+public class PolynomialPhaseSignal implements SignalGenerator<Double>{
+    
+    protected double[] params;
+    protected double[] real;
+    protected double[] imag;
+    
+    protected int n;
+    
+    protected NoiseGenerator<Double> noise;
+    
+    public PolynomialPhaseSignal(int N){
+        this.n = N;
+        real = new double[n];
+        imag = new double[n];
+    }
+
+    @Override
+    public Double[] generateReceivedSignal() {
+        for(int t = 0; t < n; t++){
+            double phase = 0.0;
+            for(int j = 0; j < params.length; j++)
+                phase += Math.pow(t+1,j)*params[j];
+            real[t] = Math.cos(2*Math.PI*(phase)) + noise.getNoise();
+            imag[t] = Math.sin(2*Math.PI*(phase)) + noise.getNoise();
+        }
+        return null;
+    }
+
+    @Override
+    public void setNoiseGenerator(NoiseGenerator<Double> noise) {
+        this.noise = noise;
+    }
+
+    @Override
+    public NoiseGenerator<Double> getNoiseGenerator() {
+        return noise;
+    }
+
+    @Override
+    public int getLength() {
+        return n;
+    }
+    
+    /** Return the noisy real component of the signal */
+    public double[] getReal() { return real; }
+    
+    /** Return the noisy imaginary component of the signal */
+    public double[] getImag() { return imag; }
+    
+    /** 
+     * Set the parameters for the polynomial phase signal with phase
+     * p[0] + p[1]t + p[2]t^2 + ...
+     */
+    public void setParameters(double[] p){
+        this.params = p;
+    }
+
+    RandomParameterGenerator pgen = new RandomParameterGenerator(0);
+    /**
+     * Generates parameters of m polynomial phase signal of order m
+     * that are uniformly distributed in the identifiable range.  See:
+     * R. G. McKilliam, I. V. L. Clarkson, "Identifiability and aliasing
+     * of polynomial phase signals", working paper.
+     */
[…]

### 2013-02-05 code/java/src/pubsim/poly/SphereDecoderEstimator.java

@@ -1,0 +1,39 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+
+package pubsim.poly;
+
+import Jama.Matrix;
+import pubsim.lattices.VnmStarGlued;
+import pubsim.lattices.decoder.SphereDecoderSchnorrEuchner;
+
+/**
+ * Runs the VnmStarSampledEfficient nearest lattice point algorithm to
+ * estimate m polynomial phase signal.
+ * @author Robby McKilliam
+ */
+public class SphereDecoderEstimator extends BabaiEstimator {
+    
+    /** 
+     * You must set the polynomial order in the constructor
+     * @param m = polynomial order
+     */
+    public SphereDecoderEstimator(int m, int n) {
+        lattice = new VnmStarGlued(m, n - m - 1);
+        this.m = m;
+        npalgorithm = new SphereDecoderSchnorrEuchner(lattice);
+        ambiguityRemover = new AmbiguityRemover(m);
+
+        ya = new double[n];
+        p = new double[m+1];
+        this.n = n;
+
+        M = lattice.getMMatrix();
+        Matrix Mt = M.transpose();
+        K = Mt.times(M).inverse().times(Mt);
+    }
+   
+
+}

### 2013-02-05 code/java/src/pubsim/poly/bounds/AngularLeastSquaresVariance.java

@@ -1,0 +1,112 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+
+package pubsim.poly.bounds;
+
+import pubsim.distributions.circular.CircularRandomVariable;
+import pubsim.distributions.circular.ProjectedNormalDistribution;
+import pubsim.distributions.circular.VonMises;
+import pubsim.distributions.circular.WrappedUniform;
+import java.io.BufferedWriter;
+import java.io.File;
+import java.io.FileWriter;
+import java.util.Vector;
+
+/**
+ *
+ * @author Robby McKilliam
+ */
+public class AngularLeastSquaresVariance extends BoundCalculator{
+
+    CircularRandomVariable dist;
+
+    public AngularLeastSquaresVariance(int N, int m){
+        super(N,m);
+    }
+
+    public double getBound(int m, CircularRandomVariable dist) {
+        double sigma2 = dist.unwrappedVariance();
+        double d = 1 - dist.pdf(-0.5);
+        return sigma2/(Math.pow(N, 2*m+1)*d*d) * C.get(m,m);
+    }
+
+
+    //** Write bound data to a file */
+    public static void main(String[] args) throws Exception  {
+
+        int N = 200;
+        int m = 2;
+
+        CircularRandomVariable dist = new WrappedUniform(0,0);
+        //CircularRandomVariable dist = new VonMises.Mod1();
+        //CircularRandomVariable dist = new ProjectedNormalDistribution();
+
+        AngularLeastSquaresVariance bound =
+                new AngularLeastSquaresVariance(N, m);
+
+
+        //this commented bit is for the projected normal
+        double from_log_snr = 18.0;
+        double to_log_snr = -6.0;
+        double step_log_snr = -0.2;
+
+        Vector<Double> snr_array = new Vector<Double>();
+        Vector<Double> snr_db_array = new Vector<Double>();
+        for(double snrdb = from_log_snr; snrdb >= to_log_snr; snrdb += step_log_snr){
+            snr_db_array.add(new Double(snrdb));
+            snr_array.add(new Double(Math.pow(10.0, ((snrdb)/10.0))));
+        }
+
+
+
+        for(int j = 0; j <= m; j++){
+            String fname = bound.getClass().getName() + dist.getClass().getCanonicalName();
+            fname = fname.concat(Integer.toString(N) + "_p" + j);
+            File file = new File(fname);
[…]

### 2013-02-05 code/java/src/pubsim/poly/bounds/GaussianCRB.java

@@ -1,0 +1,65 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+
+package pubsim.poly.bounds;
+
+import pubsim.distributions.RealRandomVariable;
+import java.io.BufferedWriter;
+import java.io.File;
+import java.io.FileWriter;
+import java.util.Vector;
+
+/**
+ *
+ * @author Robby McKilliam
+ */
+public class GaussianCRB extends BoundCalculator{
+
+    public GaussianCRB(int N, int m){
+        super(N,m);
+    }
+
+    public double getBound(int m, double variance){
+        return variance/(4*Math.PI * Math.PI * Math.pow(N,2*m+1)) * C.get(m, m);
+    }
+
+
+    //** Write CRB data to a file */
+    public static void main(String[] args) throws Exception  {
+
+        int N = 200;
+        int m = 2;
+
+        GaussianCRB bound = new GaussianCRB(N, m);
+        
+        double from_log_snr = 18.0;
+        double to_log_snr = -6.0;
+        double step_log_snr = -0.2;
+
+        Vector<Double> snr_array = new Vector<Double>();
+        Vector<Double> snr_db_array = new Vector<Double>();
+        for(double snrdb = from_log_snr; snrdb >= to_log_snr; snrdb += step_log_snr){
+            snr_db_array.add(new Double(snrdb));
+            snr_array.add(new Double(Math.pow(10.0, ((snrdb)/10.0))));
+        }
+
+        for(int j = 0; j <= m; j++){
+            String fname = bound.getClass().getName();
+            fname = fname.concat(Integer.toString(N) + "_p" + j);
+            File file = new File(fname);
+            BufferedWriter writer =  new BufferedWriter(new FileWriter(file));
+            for(int i = 0; i < snr_array.size(); i++){
+                double var = 0.5/snr_array.get(i);
+                writer.write(
+                        (new Double(var)).toString().replace('E', 'e')
+                        + "\t" + (new Double(bound.getBound(j,var))).toString().replace('E', 'e'));
+                writer.newLine();
+            }
+            writer.close();
+        }
+
+    }
+
+}

### 2013-02-05 code/java/src/pubsim/poly/dptspecialm2/AngularLeastSquares.java

@@ -1,0 +1,117 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+
+package pubsim.poly.dptspecialm2;
+
+import Jama.Matrix;
+import pubsim.lattices.Anstar.Anstar;
+import pubsim.lattices.VnmStar;
+import pubsim.Complex;
+import pubsim.VectorFunctions;
+import pubsim.lattices.Anstar.AnstarLinear;
+import pubsim.poly.AmbiguityRemover;
+import pubsim.poly.PolynomialPhaseEstimator;
+
+/**
+ * Returns the angular least squares estimator but only searches in the interval
+ * on which the DPT estimator works.
+ * @author Robby McKilliam
+ */
+public class AngularLeastSquares implements PolynomialPhaseEstimator{
+
+    protected final int m = 2;
+    protected int N;
+    protected Matrix M,  K;
+
+    private Anstar anstar;
+    private double[] phi;
+    private double[] z;
+    private double[] pmw;
+    private final AmbiguityRemover ambiguityRemover;
+
+    public AngularLeastSquares(){
+        ambiguityRemover = new AmbiguityRemover(m);
+    }
+
+    public void setSize(int n) {
+        N = n;
+        anstar = new AnstarLinear(N-1);
+        M = VnmStar.getMMatrix(N-m-1, m);
+        Matrix Mt = M.transpose();
+        K = Mt.times(M).inverse().times(Mt);
+        z = new double[N];
+        pmw = new double[N];
+        phi = new double[N];
+    }
+
+    public int getOrder() {
+        return m;
+    }
+
+    public double[] estimate(double[] real, double[] imag) {
+        if(real.length != N) setSize(real.length);
+
+        final int numsamples = 4*N;
+
+        for(int i = 0; i < N; i++){
+            phi[i] = (new Complex(real[i], imag[i])).phase()/(2*Math.PI);
+        }
+
+        double p2start = -(2.0/N)/4.0;
+        double p2end = (2.0/N)/4.0;
+        double step = (p2end - p2start)/numsamples;
+
+        double D = Double.POSITIVE_INFINITY;
+        double[] hatp = new double[3];
+
+        for(double p2 = p2start; p2 < p2end; p2 += step){
+            for(double f = -0.5; f < 0.5; f+= 1.0/numsamples){
+
+                for (int i = 0; i < N; i++) {
[…]

### 2013-02-05 code/java/src/pubsim/poly/dptspecialm2/MaximumLikelihood.java

@@ -1,0 +1,116 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+
+package pubsim.poly.dptspecialm2;
+
+import Jama.Matrix;
+import pubsim.optimisation.NewtonRaphson;
+import pubsim.Complex;
+import pubsim.VectorFunctions;
+import pubsim.fes.PeriodogramFFTEstimator;
+import pubsim.poly.AmbiguityRemover;
+
+/**
+ * Polynomial phase estimator that for m=2 that only searches the DPT region.
+ * O(N^2logN) complexity.
+ * @author Robby McKilliam
+ */
+public class MaximumLikelihood extends pubsim.poly.MaximumLikelihood {
+
+    Complex[] z;
+    double[] realp, imagp;
+
+    public MaximumLikelihood(int n) {
+        this.m = 2;
+        ambiguityRemover = new AmbiguityRemover(2);
+        N = n;
+        realp = new double[N];
+        imagp = new double[N];
+        z = new Complex[N];
+        freqest = new PeriodogramFFTEstimator(N);
+    }
+
+    
+    @Override
+    public double[] estimate(double[] real, double[] imag) {
+        if(N != real.length) throw new RuntimeException("Data length does not equal " + N);
+        
+        PolynomialPhaseLikelihood func
+                = new PolynomialPhaseLikelihood(real, imag);
+        NewtonRaphson newtonRaphson
+                = new NewtonRaphson(func);
+
+        int numsamples = 4*N;
+
+        double p2start = -(2.0/N)/4.0;
+        double p2end = (2.0/N)/4.0;
+        double step = (p2end - p2start)/numsamples;
+
+        Matrix p = null;
+        double D = Double.NEGATIVE_INFINITY;
+
+        for(double p2 = p2start; p2 < p2end; p2 += step){
+            for (int j = 0; j < real.length; j++)
+                z[j] = new Complex(real[j], imag[j]);
+
+            for (int i = 0; i < N; i++) {
+                double cs = Math.cos(-2.0 * Math.PI * p2 * (i+1)*(i+1) );
+                double ss = Math.sin(-2.0 * Math.PI * p2 * (i+1)*(i+1) );
+                z[i] = z[i].times(new Complex(cs, ss));
+            }
+
+            for (int j = 0; j < real.length; j++){
+                realp[j] = z[j].re();
+                imagp[j] = z[j].im();
+            }
+
[…]

### 2013-02-05 code/java/test/pubsim/poly/AmbiguityRemoverTest.java

@@ -1,0 +1,105 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+
+package pubsim.poly;
+
+import pubsim.poly.AmbiguityRemover;
+import Jama.Matrix;
+import org.junit.After;
+import org.junit.AfterClass;
+import org.junit.Before;
+import org.junit.BeforeClass;
+import org.junit.Test;
+import pubsim.VectorFunctions;
+import static org.junit.Assert.*;
+
+/**
+ *
+ * @author Robby McKilliam
+ */
+public class AmbiguityRemoverTest {
+
+    public AmbiguityRemoverTest() {
+    }
+
+    @BeforeClass
+    public static void setUpClass() throws Exception {
+    }
+
+    @AfterClass
+    public static void tearDownClass() throws Exception {
+    }
+
+    @Before
+    public void setUp() {
+    }
+
+    @After
+    public void tearDown() {
+    }
+
+
+    @Test
+    public void throwsExceptionWhenIncorrectSize() {
+        System.out.println("throwsExceptionWhenIncorrectSize");
+        int m = 4;
+        double[] p = {3, 2, 4, 5};
+        AmbiguityRemover instance = new AmbiguityRemover(m);
+
+        boolean caught = false;
+        try{
+            instance.disambiguate(p);
+        } catch(RuntimeException e){
+            caught = true;
+        }
+        assertTrue(caught);
+    }
+
+    @Test
+    public void testDisambiguate() {
+        System.out.println("testDisambiguate");
+        double[] p1 = {1.3, -2.1};
+        int m = p1.length - 1;
+        AmbiguityRemover instance = new AmbiguityRemover(m);
+        p1 = instance.disambiguate(p1);
+        double[] y1 = {0.3, -0.1};
+        assertTrue(VectorFunctions.distance_between2(p1, y1) < 0.000001);
+
+        double[] p2 = {1.3, -2.1, 2.6};
+        m = p2.length - 1;
+        instance = new AmbiguityRemover(m);
+        p2 = instance.disambiguate(p2);
+        double[] y2 = {0.3, 0.4, 0.1};
+        assertTrue(VectorFunctions.distance_between2(p2, y2) < 0.000001);
+        
+    }
+
+
+        /**
+     * Test of setSize method, of class AmbiguityRemover.
+     */
+    @Test
+    public void testNextColumn() {
[…]

### 2013-02-05 code/java/test/pubsim/poly/BabaiEstimatorTest.java

@@ -1,0 +1,77 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+
+package pubsim.poly;
+
+import pubsim.poly.PolynomialPhaseEstimator;
+import pubsim.poly.BabaiEstimator;
+import pubsim.poly.PolynomialPhaseSignal;
+import Jama.Matrix;
+import pubsim.distributions.GaussianNoise;
+import pubsim.lattices.VnmStar;
+import org.junit.After;
+import org.junit.AfterClass;
+import org.junit.Before;
+import org.junit.BeforeClass;
+import org.junit.Test;
+import pubsim.VectorFunctions;
+import static org.junit.Assert.*;
+
+/**
+ *
+ * @author harprobey
+ */
+public class BabaiEstimatorTest {
+
+    public BabaiEstimatorTest() {
+    }
+
+    @BeforeClass
+    public static void setUpClass() throws Exception {
+    }
+
+    @AfterClass
+    public static void tearDownClass() throws Exception {
+    }
+
+    @Before
+    public void setUp() {
+    }
+
+    @After
+    public void tearDown() {
+    }
+
+            /**
+     * Test of estimate method, of class DPTEstimator.
+     */
+    @Test
+    public void testEstimate() {
+        System.out.println("testEstimate");
+
+        int n = 24;
+        double[] params = {0.11, 0.05002, 0.0205, 0.0001};
+        int m = params.length-1;
+
+        Matrix M = VnmStar.getGeneratorMatrix(m, n-m-1);
+        System.out.println(VectorFunctions.print(M));
+
+        PolynomialPhaseSignal siggen = new PolynomialPhaseSignal(n);
+        siggen.setParameters(params);
+        siggen.setNoiseGenerator(new GaussianNoise(0, 0.00001));
+
+        siggen.generateReceivedSignal();
+
+        PolynomialPhaseEstimator inst = new BabaiEstimator(m,n);
+
+        double[] p = inst.estimate(siggen.getReal(), siggen.getImag());
+
+        System.out.println(VectorFunctions.print(p));
+
+        assertTrue(VectorFunctions.distance_between(p, params) < 0.001);
+
+    }
+
+}

### 2013-02-05 code/java/test/pubsim/poly/DPTEstimatorTest.java

@@ -1,0 +1,142 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+
+package pubsim.poly;
+
+import pubsim.poly.DPTEstimator;
+import pubsim.poly.PolynomialPhaseSignal;
+import pubsim.distributions.GaussianNoise;
+import org.junit.After;
+import org.junit.AfterClass;
+import org.junit.Before;
+import org.junit.BeforeClass;
+import org.junit.Test;
+import static org.junit.Assert.*;
+import pubsim.Complex;
+import pubsim.VectorFunctions;
+
+/**
+ *
+ * @author Robby McKilliam
+ */
+public class DPTEstimatorTest {
+
+    public DPTEstimatorTest() {
+    }
+
+    @BeforeClass
+    public static void setUpClass() throws Exception {
+    }
+
+    @AfterClass
+    public static void tearDownClass() throws Exception {
+    }
+
+    @Before
+    public void setUp() {
+    }
+
+    @After
+    public void tearDown() {
+    }
+
+
+    /**
+     * Test of PPT method, of class DPTEstimator.
+     */
+//    @Test
+//    public void PPTHasFirstMElementsZero() {
+//        int m = 4;
+//        int n = 10;
+//        Complex[] y = VectorFunctions.randomComplex(n);
+//        DPTEstimator instance = new DPTEstimator(m);
+//        instance.setSize(n);
+//        Complex[] result = instance.PPT(m, y);
+//        //System.out.print(VectorFunctions.print(result));
+//        for(int i = 0; i <= 2; i++){
+//            assertTrue(result[i].re() == 0.0);
+//            assertTrue(result[i].im() == 0.0);
+//        }
+//    }
+
+    /**
+     * Test of PPT2 method, of class DPTEstimator.
+     */
+    @Test
+    public void PPT2HasLastElementCorrect() {
+        System.out.println("PPT2");
+        int m = 4;
+        int n = 10;
+        Complex[] y = VectorFunctions.randomComplex(n);
+        DPTEstimator instance = new DPTEstimator(m,n);
+        flanagan.complex.Complex[] result = instance.PPT2(VectorFunctions.simComplexArrayToFlanComplexArray(y));
+        //System.out.print(VectorFunctions.print(result));
+
+        //test first element is zero
+        //assertTrue(result[0].re() == 0.0);
+        //assertTrue(result[0].im() == 0.0);
+
+        //test last element
[…]

### 2013-02-05 code/java/test/pubsim/poly/KitchenEstimatorTest.java

@@ -1,0 +1,97 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+
+package pubsim.poly;
+
+import pubsim.poly.KitchenEstimator;
+import pubsim.poly.PolynomialPhaseSignal;
+import pubsim.distributions.GaussianNoise;
+import org.junit.After;
+import org.junit.AfterClass;
+import org.junit.Before;
+import org.junit.BeforeClass;
+import org.junit.Test;
+import pubsim.VectorFunctions;
+import static org.junit.Assert.*;
+
+/**
+ *
+ * @author Robby
+ */
+public class KitchenEstimatorTest {
+
+    public KitchenEstimatorTest() {
+    }
+
+    @BeforeClass
+    public static void setUpClass() throws Exception {
+    }
+
+    @AfterClass
+    public static void tearDownClass() throws Exception {
+    }
+
+    @Before
+    public void setUp() {
+    }
+
+    @After
+    public void tearDown() {
+    }
+
+    /**
+     * Test of estimate method, of class KitchenEstimator.
+     */
+    @Test
+    public void testEstimate() {
+        System.out.println("testEstimate");
+
+        int n = 20;
+        double[] params = {0.1, 0.2, 0.05};
+        int m = params.length-1;
+
+        PolynomialPhaseSignal siggen = new PolynomialPhaseSignal(n);
+        siggen.setParameters(params);
+        siggen.setNoiseGenerator(new GaussianNoise(0, 0.000001));
+
+        siggen.generateReceivedSignal();
+
+        KitchenEstimator inst = new KitchenEstimator(m,n);
+
+        double[] p = inst.estimate(siggen.getReal(), siggen.getImag());
+
+        System.out.println(VectorFunctions.print(p));
+
+        assertTrue(VectorFunctions.distance_between(p, params) < 0.001);
+    }
+
+    /**
+     * Test of estimateM method, of class KitchenEstimator.
+     */
+    @Test
+    public void testHighestOrderParameter() {
+        System.out.println("testHighestOrderParameter");
+
+        int n = 20;
+        double[] params = {0.1, 0.2, 0.05};
+        int a = params.length;
+
+        PolynomialPhaseSignal siggen = new PolynomialPhaseSignal(n);
+        siggen.setParameters(params);
+        siggen.setNoiseGenerator(new GaussianNoise(0, 0.000001));
+
[…]

### 2013-02-05 code/java/test/pubsim/poly/MaximumLikelihoodTest.java

@@ -1,0 +1,122 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+
+package pubsim.poly;
+
+import Jama.Matrix;
+import static org.junit.Assert.assertEquals;
+import static org.junit.Assert.assertTrue;
+import org.junit.*;
+import pubsim.VectorFunctions;
+import pubsim.distributions.GaussianNoise;
+
+/**
+ *
+ * @author Robby McKilliam
+ */
+public class MaximumLikelihoodTest {
+
+    public MaximumLikelihoodTest() {
+    }
+
+    @BeforeClass
+    public static void setUpClass() throws Exception {
+    }
+
+    @AfterClass
+    public static void tearDownClass() throws Exception {
+    }
+
+    @Before
+    public void setUp() {
+    }
+
+    @After
+    public void tearDown() {
+    }
+
+    @Test
+    public void testPolyFunctionValue() {
+        System.out.println("testPolyFunctionValue");
+        double[] yr = {0.1, 0.1, 0.1};
+        double[] yi = {0.1, 0.1, 0.1};
+        double[] p = {0.1, 0.1, 0.1};
+        Matrix P = VectorFunctions.columnMatrix(p);
+        MaximumLikelihood.PolynomialPhaseLikelihood func
+                = new MaximumLikelihood.PolynomialPhaseLikelihood(yr, yi);
+        //calculated in Matlab
+        double expr = -3.055198893365938;
+        double res = func.value(P);
+        assertEquals(res, expr, 0.000001);
+        
+    }
+
+    @Test
+    public void testPolyFunctionGradient() {
+        System.out.println("testPolyFunctionGradient");
+        double[] yr = {0.1, 0.1, 0.1};
+        double[] yi = {0.1, 0.1, 0.1};
+        double[] p = {0.1, 0.1, 0.1};
+        Matrix P = VectorFunctions.columnMatrix(p);
+        MaximumLikelihood.PolynomialPhaseLikelihood func
+                = new MaximumLikelihood.PolynomialPhaseLikelihood(yr, yi);
+        MaximumLikelihood.PolynomialPhaseLikelihoodAutoDerivative funcA
+                = new MaximumLikelihood.PolynomialPhaseLikelihoodAutoDerivative(yr, yi);
+        //calculated in Matlab
+        Matrix gf = func.gradient(P);
+        Matrix gfA = funcA.gradient(P);
[…]

### 2013-02-05 code/java/test/pubsim/poly/PolynomialPhaseSignalTest.java

@@ -1,0 +1,65 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+
+package pubsim.poly;
+
+import pubsim.poly.PolynomialPhaseSignal;
+import pubsim.distributions.GaussianNoise;
+import junit.framework.TestCase;
+import pubsim.VectorFunctions;
+
+/**
+ *
+ * @author Robby
+ */
+public class PolynomialPhaseSignalTest extends TestCase {
+    
+    public PolynomialPhaseSignalTest(String testName) {
+        super(testName);
+    }            
+
+    /**
+     * Test of generateReceivedSignal method, of class PolynomialPhaseSignal.
+     */
+    public void testGenerateReceivedSignal() {
+        System.out.println("generateReceivedSignal");
+        
+        int n = 4;
+        double[] params = {1, 0.1, 0.2};
+        
+        PolynomialPhaseSignal instance = new PolynomialPhaseSignal(n);
+        instance.setParameters(params);
+        instance.setNoiseGenerator(new GaussianNoise(0, 0.0));
+        
+        //these are taken from matlab
+        double[] expreal = {-0.3090, 1.0000, 0.8090, -0.8090};
+        double[] expimag = {0.9511, -0.0000, 0.5878, -0.5878};
+        
+        instance.generateReceivedSignal();
+        
+        System.out.println("real = " + VectorFunctions.print(instance.getReal()));
+        System.out.println("imag = " + VectorFunctions.print(instance.getImag()));
+        
+        assertEquals(VectorFunctions.distance_between(expreal, instance.getReal())<0.001, true);
+        assertEquals(VectorFunctions.distance_between(expimag, instance.getImag())<0.001, true);
+        
+    }
+
+        /**
+     * Test of generateReceivedSignal method, of class PolynomialPhaseSignal.
+     */
+    public void testGenerateRandomParameters() {
+        System.out.println("testGenerateRandomParameters");
+        int m = 2;
+        PolynomialPhaseSignal.RandomParameterGenerator pgen
+                = new PolynomialPhaseSignal.RandomParameterGenerator(m);
+
+        double[] p = pgen.generateParameters();
+        assertEquals(3, p.length);
+        System.out.println(VectorFunctions.print(p));
+
+    }
+
+}

### 2013-02-05 code/java/test/pubsim/poly/TestWeakConvergenceOfApproximatingPolynomial.java

@@ -1,0 +1,59 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+package pubsim.poly;
+
+import org.junit.*;
+import static pubsim.Util.fracpart;
+import pubsim.distributions.circular.CircularRandomVariable;
+import pubsim.distributions.circular.CircularUniform;
+
+/**
+ *
+ * @author Robby McKilliam
+ */
+public class TestWeakConvergenceOfApproximatingPolynomial {
+    
+    @BeforeClass
+    public static void setUpClass() throws Exception {
+    }
+
+    @AfterClass
+    public static void tearDownClass() throws Exception {
+    }
+
+    @Before
+    public void setUp() {
+    }
+
+    @After
+    public void tearDown() {
+    }
+
+    protected double polyn(double x, double N){
+        if(Math.abs(x) < 1.0/2.0 - 1.0/2.0/N ) return 2.0;
+        double y = fracpart(x - 0.5);
+        return (4 - 15*N)/2.0 + 4*3*10*N*N*N*y*y - 6*5*12*N*N*N*N*N*y*y*y*y;
+    }
+    
+    @Test
+    public void test() {
+        System.out.println("testPolyFunctionValue");
+        
+        long fromN = 100000;
+        int numruns = 30;
+
+        
+        CircularRandomVariable rv = new CircularUniform();
+        
+        for(long N = fromN; N < fromN + numruns; N++){
+            double sum = 0.0;
+            for(int n = 0; n < N; n++) sum += polyn(rv.getNoise(),Math.round(Math.pow(N,1.0/2 + 0.01)));
+            System.out.println("N = " + N + ", mean = " + (sum/N));
+        }
+                
+        
+    }
+    
+}

### 2013-02-06 paper.tex
moved data

@@
 Our unwrapping estimator can be computed in a reasonable amount of time for data sets of moderate size using existing general purpose algorithms from algebraic number theory.
-Monte-carlo simualtions assert the superior statistical performance of our unwrapping estimator when compared to existing state-of-the-art estimators for data sets of less than 300 observations.
+Monte-carlo simualtions assert the statistical performance of our unwrapping estimator when compared to existing state-of-the-art estimators for data sets of less than 250 observations.
+A feature of our estimator is that it functions correctly over a far wider range of parameters than most existing estimators.
 \end{abstract}
@@
    	\centering 
-  		\includegraphics{code/gaussianplot4-1.mps} 
-  		\caption{Mean square error (MSE) of polynomial phase estimators.
-The true coefficients $\tilde{\mubf}$ are uniformly distributed in the identifiable region $B$.
-(Top left) MSE in the phase coefficient $\mu_0$.
-(Top right) MSE in the frequency coefficient $\mu_1$.
-(Bottom left) MSE in the quadratic coefficient $\mu_2$.
-(Bottom right) MSE in the cubic coefficient $\mu_3$.}
+  		\includegraphics{code/data/plot-1.mps} 
+  		\caption{BLERG} 
   		\label{plot:polyest} 
  \end{figure*}
-
-\begin{figure}[p] 
-   	\centering 
-  		\includegraphics{code/gaussianplot4-2.mps} 
-  		\caption{Mean square error (MSE) of polynomial phase estimators.
-The true coefficients are fixed to $\tilde{\mubf} = [0.1,0.2,10^{-3},10^{-5}]$.
-(Top) MSE in the phase coefficient $\widehat{\mu}_0$.
-(Bottom right) MSE in the cubic coefficient $\widehat{\mu}_3$.}
-  		\label{plot:polyest2} 
- \end{figure}
 

### 2013-02-06 build.sh

@@ -10,10 +10,9 @@ mpost -interaction=nonstopmode
 distplots.class.distributions.circular.ProjectedNormalDistribution.var3.0.mp 
 cd .. 
 
-cd code 
-mpost -interaction=nonstopmode gaussianplot4.mp 
-mpost -interaction=nonstopmode lseplot.mp 
-cd .. 
+cd ./code/data/ 
+mpost -interaction=nonstopmode plot.mp 
+cd ../../ 
 
 pdflatex paper.tex 
 bibtex paper.aux 

### 2013-02-06 code/java/src/pubsim/poly/BabaiEstimator.java

@@ -6,7 +6,6 @@ import Jama.Matrix;
 import pubsim.VectorFunctions;
 import pubsim.lattices.NearestPointAlgorithmInterface;
 import pubsim.lattices.VnmStar;
-import pubsim.lattices.VnmStarGlued;
 import pubsim.lattices.decoder.Babai;
 
 /**
@@ -27,7 +26,7 @@ public class BabaiEstimator extends AbstractPolynomialPhaseEstimator {
      */
     public BabaiEstimator(int m, int n) {
         super(m);
-        lattice = new VnmStarGlued(m, n - m - 1);
+        lattice = new VnmStar(m, n - m - 1);
         npalgorithm = new Babai(lattice);
         ya = new double[n];
         p = new double[m+1];

### 2013-02-06 code/java/src/pubsim/poly/CubicPhaseFunction.java

@@ -50,7 +50,7 @@ public class CubicPhaseFunction extends AbstractPolynomialPhaseEstimator {
     /** Cubic phase function estimator of dimension N.  Will use sample the CP function
      * 4N times to approximate maximum. */
     public CubicPhaseFunction(int N){
-        this(N,4*N);
+        this(N,10*N);
     } 
 
     @Override
@@ -128,7 +128,7 @@ public class CubicPhaseFunction extends AbstractPolynomialPhaseEstimator {
                 return -CP(n,x).abs();
             }
         };
-        Brent optimiser = new Brent(f, what - step, what, what + step);
+        Brent optimiser = new Brent(f, what - step, what, what + step, 1e-10);
         //System.out.println();
         return what;     
     }

### 2013-02-06 code/java/test/pubsim/poly/BabaiEstimatorTest.java

@@ -63,9 +63,7 @@ public class BabaiEstimatorTest {
         siggen.setNoiseGenerator(new GaussianNoise(0, 0.00001));
 
         siggen.generateReceivedSignal();
-
         PolynomialPhaseEstimatorInterface inst = new BabaiEstimator(m,n);
-
         double[] p = inst.estimate(siggen.getReal(), siggen.getImag());
 
         System.out.println(VectorFunctions.print(p));

### 2013-02-06 code/montecarlo.scala

@@ -1,0 +1,126 @@
+/**
+ * Run simulations for the sparse noisy period estimation problem.
+ * Various estimators are available.
+ */
+import pubsim.poly.PolynomialPhaseSignal
+import pubsim.poly.KitchenEstimator
+import pubsim.poly.DPTEstimator
+import pubsim.poly.MbestEstimator
+import pubsim.poly.BabaiEstimator
+import pubsim.poly.MbestEstimator
+import pubsim.poly.MaximumLikelihood
+import pubsim.poly.SphereDecoderEstimator
+import pubsim.poly.PolynomialPhaseEstimatorInterface
+import pubsim.poly.CubicPhaseFunction
+import pubsim.poly.bounds.AngularLeastSquaresVariance
+import pubsim.poly.bounds.GaussianCRB
+import pubsim.distributions.GaussianNoise
+import pubsim.distributions.circular.ProjectedNormalDistribution
+import pubsim.lattices.reduction.None
+
+val iters = 500 //number of Monte-Carlo trials.
+val Ns = List(59) //values of N we will generate curves for
+val ms = List(3) //order of our polynomial phase signals
+
+//returns an array of noise distributions with a logarithmic scale
+val SNRdBs = -5 to 15
+val SNRs = SNRdBs.map(db => scala.math.pow(10.0, db/10.0))
+def noises =  SNRs.map( snr => new GaussianNoise(0,1.0/snr/2.0) ) //variance for real and imaginary parts (divide by 2)
+
+//Returns a list of functions that return estimators we will run (factory patern to enable parallelism)
+def estfactory(m : Int, N : Int) : List[() => PolynomialPhaseEstimatorInterface] = {
+  var ret = List( 
+    () => new KitchenEstimator(m,N),
+    () => new DPTEstimator(m,N),
+    //() => new BabaiEstimator(m,N),
+    () => new MbestEstimator(m,N,4*N) 
+  )
+  //add the sphere decoder and Least squares estimators if N and m are small
+  if( N < 60 ) ret = ret :+ ( () => new SphereDecoderEstimator(m,N) )
+  //if( N < 60 && m <= 2 ) ret = ret :+ ( () => new MaximumLikelihood(m,N) )
+if( m==3 ) ret = ret :+ ( () => new CubicPhaseFunction(N) )
+  return ret
+}
+
+val starttime = (new java.util.Date).getTime
+
+//for all the the values of N and m.
+for( N <-  Ns; m <- ms ) {
+
[…]

### 2013-02-06 code/runsim.sh

@@ -1,3 +1,10 @@
 #export JAVA_OPTS="-Xprof -server -Xms1g -Xmx1g"
 export JAVA_OPTS="-d64 -server -Xms1g -Xmx1g"
-scala -nocompdaemon -cp java/lib/PubSim.jar:java/lib/Jama-1.0.3.jar:java/lib/flanagan.jar:java/lib/colt.jar:java/lib/RngPack.jar:java/lib/Poly.jar simgaussian.scala
+
+CP=""
+for f in java/lib/*.jar
+do
+CP=$CP:${f}
+done
+
+scala -nocompdaemon -cp $CP montecarlo.scala

### 2013-02-06 code/simgaussian.scala

@@ -1,118 +1,0 @@
-/**
- * Run simulations for the sparse noisy period estimation problem.
- * Various estimators are available.
- */
-import pubsim.poly.PolynomialPhaseSignal
-import pubsim.poly.KitchenEstimator
-import pubsim.poly.DPTEstimator
-import pubsim.poly.MbestEstimator
-import pubsim.poly.BabaiEstimator
-import pubsim.poly.MbestEstimator
-import pubsim.poly.MaximumLikelihood
-import pubsim.poly.SphereDecoderEstimator
-import pubsim.poly.PolynomialPhaseEstimator
-import pubsim.poly.CubicPhaseFunction
-import pubsim.poly.bounds.AngularLeastSquaresVariance
-import pubsim.poly.bounds.GaussianCRB
-import pubsim.distributions.GaussianNoise
-import pubsim.distributions.circular.ProjectedNormalDistribution
-
-val iters = 100 //number of Monte-Carlo trials.
-val Ns = List(39)//,249) //values of N we will generate curves for
-val ms = List(3) //order of our polynomial phase signals
-
-def gp( a : Double, r : Double, n : Int ) : Double = a * scala.math.pow(r,n)
-
-//returns an array of noise distributions with a logarithmic scale
-val SNRdBs = 0 to 21
-val SNRs = SNRdBs.map(db => scala.math.pow(10.0, db/10.0))
-def noises =  SNRs.map( snr => new GaussianNoise(0,1.0/snr/2.0) ) //variance for real and imaginary parts (divide by 2)
-
-//Returns a list of functions that return estimators we will run (factory patern to enable parallelism)
-def estfactory(m : Int, N : Int) : List[() => PolynomialPhaseEstimator] = {
-  var ret = List( 
-    () => new KitchenEstimator(m,N),
-    () => new DPTEstimator(m,N),
-    //() => new BabaiEstimator(m,N),
-    () => new MbestEstimator(m,N,4*N) 
-  )
-  //add the sphere decoder and Least squares estimators if N and m are small
-  if( N < 60 ) ret = ret :+ ( () => new SphereDecoderEstimator(m,N) )
-  //if( N < 60 && m <= 2 ) ret = ret :+ ( () => new MaximumLikelihood(m,N) )
-if( m==3 ) ret = ret :+ ( () => new CubicPhaseFunction(N) )
-  return ret
-}
-
-val starttime = (new java.util.Date).getTime
-
-//for all the the values of N and m.
-for( N <-  Ns; m <- ms ) {
-
[…]

### 2013-02-06 code/simwrappedu.scala

@@ -1,98 +1,0 @@
-/**
- * Run simulations for the sparse noisy period estimation problem.
- * Various estimators are available.
- */
-import pubsim.poly.CircularNoisePolynomialPhaseSignal
-import pubsim.poly.KitchenEstimator
-import pubsim.poly.DPTEstimator
-import pubsim.poly.MbestEstimator
-import pubsim.poly.BabaiEstimator
-import pubsim.poly.MbestEstimator
-import pubsim.poly.MaximumLikelihood
-import pubsim.poly.SphereDecoderEstimator
-import pubsim.distributions.ContinuousRandomVariable
-import pubsim.poly.PolynomialPhaseEstimator
-import pubsim.poly.bounds.AngularLeastSquaresVariance
-import pubsim.distributions.circular.WrappedUniform
-
-val iters = 2000 //number of Monte-Carlo trials.
-val Ns = List(10,50,200) //values of N we will generate curves for
-val ms = List(2,3) //order of our polynomial phase signals
-
-def gp( a : Double, r : Double, n : Int ) : Double = a * scala.math.pow(r,n)
-
-//returns an array of noise distributions with a logarithmic scale
-def noises = (0 to 25).map( n => gp(0.002,1.15,n) ).map( v => new WrappedUniform(0,v) )
-
-//Returns a list of functions that return estimators we will run (factory patern to enable parallelism)
-def estfactory(m : Int, N : Int) : List[() => PolynomialPhaseEstimator] = {
-  var ret = List( 
-    () => new KitchenEstimator(m,N),
-    () => new DPTEstimator(m,N),
-    () => new BabaiEstimator(m,N),
-    () => new MbestEstimator(m,N,4*N) 
-  )
-  //add the sphere decoder and Least squares estimators if N and m are small
-  if( N < 60 ) ret = ret :+ ( () => new SphereDecoderEstimator(m,N) )
-  if( N < 60 && m <= 2 ) ret = ret :+ ( () => new MaximumLikelihood(m,N) )
-  return ret
-}
-
-val starttime = (new java.util.Date).getTime
-
-//for all the the values of N and m.
-for( N <-  Ns; m <- ms ) {
-
-  for(estf <- estfactory(m,N) ){
-    
-    val estname = estf().getClass.getSimpleName + "N" + N.toString + "m" + m + "WU"
-    print("Running " + estname)
-    val eststarttime = (new java.util.Date).getTime
-
-    //for all the noise distributions (in parallel threads)
[…]

### 2013-02-06 code/bench.scala
added code for benchmarking

@@ -1,0 +1,43 @@
+/**
+ * Run simulations for the sparse noisy period estimation problem.
+ * Various estimators are available.
+ */
+
+import pubsim.poly.MbestEstimator
+import pubsim.poly.PolynomialPhaseSignal
+import pubsim.distributions.GaussianNoise
+
+val m = 8;
+val Ns = List(30, 60, 100, 200, 350, 500, 1000);
+
+val MIN_BENCH_DURATION : Long = 40000000000L; // (20 secs)  
+
+val snrdB = 7
+val snr = scala.math.pow(10.0, snrdB/10.0)
+val noise =  new GaussianNoise(0,1.0/snr/2.0) //variance for real and imaginary parts (divide by 2)
+
+for(N <- Ns){
+
+val siggen =  new PolynomialPhaseSignal(N) //construct a signal generator
+siggen.setNoiseGenerator(noise)
+siggen.generateRandomParameters(m)
+siggen.generateReceivedSignal
+
+val est = new MbestEstimator(m,N,4*N) //contruct an estimator to benchmark
+
+print(" ... Warming up ... ")
+var numiters = 0
+val warmupstarttime = System.nanoTime
+while(System.nanoTime - warmupstarttime < MIN_BENCH_DURATION/2){
+   est.estimate(siggen.getReal, siggen.getImag)
+   numiters = numiters+2
+}
+    
+print("Benchmarking ... ")
+val benchstarttime = System.nanoTime
+for( i <- 1 to numiters) est.estimate(siggen.getReal, siggen.getImag)
+val tNano = (System.nanoTime - benchstarttime +(numiters)/2) / numiters //copied from Alan Eliasen java BigInteger benchmarks
+val tSec = tNano/1000000000.0 //time in milliseconds per iteration
+println(tSec + " seconds per iteration")
+
+}

### 2013-02-06 code/benchmark.sh

@@ -1,0 +1,10 @@
+#export JAVA_OPTS="-Xprof -server -Xms1g -Xmx1g"
+export JAVA_OPTS="-d64 -server -Xms1g -Xmx1g"
+
+CP=""
+for f in java/lib/*.jar
+do
+CP=$CP:${f}
+done
+
+scala -nocompdaemon -cp $CP bench.scala

### 2013-02-06 paper.tex
more sims

@@
 Our unwrapping estimator can be computed in a reasonable amount of time for data sets of moderate size using existing general purpose algorithms from algebraic number theory.
-Monte-carlo simualtions assert the statistical performance of our unwrapping estimator when compared to existing state-of-the-art estimators for data sets of less than 250 observations.
+Monte-carlo simualtions assert the excellent statistical performance of our unwrapping estimator when compared to existing state-of-the-art estimators for data sets of less than 200 observations.
 A feature of our estimator is that it functions correctly over a far wider range of parameters than most existing estimators.
@@
 \]
-It will follow that
+Using standard techniques~\cite{Pollard_conv_stat_proc_1984,van2009empirical} one can show that
  \begin{equation}\label{eq:SNVNunifmlln}
@@
 almost surely as $N\rightarrow\infty$.
-This type of result has been called a \emph{uniform law of large numbers} and follows from standard techniques~\cite{Pollard_conv_stat_proc_1984}.
+This type of resul is called a~\emph{uniform law of large numbers}~\cite{Pollard_conv_stat_proc_1984,van2009empirical}.
 A full proof of~\eqref{eq:SNVNunifmlln} in given in BLERG (see arxiv).
@@
 
-\begin{figure*}[p] 
+\begin{figure}[t] 
    	\centering 
@@
   		\label{plot:polyest} 
- \end{figure*}
+ \end{figure}
+
+\begin{figure}[t] 
+   	\centering 
+  		\includegraphics{code/data/plot-2.mps} 
+  		\caption{BLERG} 
+  		\label{plot:polyest} 
+ \end{figure}
+
+%\begin{figure*}[t] 
+%\begin{gnuplot}[terminal=epslatex]
+%plot sin(x), cos(x)
+%\end{gnuplot}
+%\end{figure*}
 

### 2013-02-06 build.sh

@@ -14,10 +14,10 @@ cd ./code/data/
 mpost -interaction=nonstopmode plot.mp 
 cd ../../ 
 
-pdflatex paper.tex 
+pdflatex -shell-escape paper.tex 
 bibtex paper.aux 
-pdflatex paper.tex 
-pdflatex paper.tex 
+pdflatex -shell-escape paper.tex 
+pdflatex -shell-escape paper.tex 
 
 #pdflatex paper2.tex 
 #bibtex paper2.aux 

### 2013-02-06 code/java/src/pubsim/poly/BabaiEstimator.java

@@ -7,10 +7,12 @@ import pubsim.VectorFunctions;
 import pubsim.lattices.NearestPointAlgorithmInterface;
 import pubsim.lattices.VnmStar;
 import pubsim.lattices.decoder.Babai;
+import pubsim.lattices.reduction.LatticeReduction;
+import pubsim.lattices.reduction.LLL;
 
 /**
  * Uses the Babai nearest plane algorithm
- * @author Robby
+ * @author Robby McKilliam
  */
 public class BabaiEstimator extends AbstractPolynomialPhaseEstimator {
 
@@ -20,19 +22,24 @@ public class BabaiEstimator extends AbstractPolynomialPhaseEstimator {
     protected NearestPointAlgorithmInterface npalgorithm;
     final protected Matrix M,  K;
     
+    public BabaiEstimator(int m, int n){
+        this(m,n,new LLL());
+    } 
+    
     /** 
      * You must set the polynomial order in the constructor
      * @param m = polynomial order
      */
-    public BabaiEstimator(int m, int n) {
+    public BabaiEstimator(int m, int n, LatticeReduction lr) {
         super(m);
         lattice = new VnmStar(m, n - m - 1);
-        npalgorithm = new Babai(lattice);
+        npalgorithm = new Babai(lattice, lr);
         ya = new double[n];
         p = new double[m+1];
         this.n = n;
         M = lattice.getMMatrix();
         Matrix Mt = M.transpose();
+        //System.out.println(Mt.times(M).inverse().cond());
         K = Mt.times(M).inverse().times(Mt);
     }
 

### 2013-02-06 code/java/src/pubsim/poly/CubicPhaseFunction.java

@@ -4,9 +4,8 @@ package pubsim.poly;
 
 import Jama.Matrix;
 import pubsim.Complex;
-import pubsim.optimisation.Brent;
 import pubsim.optimisation.FunctionAndDerivatives;
-import pubsim.optimisation.SingleVariateFunction;
+import pubsim.optimisation.NewtonRaphson;
 
 /**
  * Implements O'Shea's cubic phase function estimator
@@ -124,13 +123,26 @@ public class CubicPhaseFunction extends AbstractPolynomialPhaseEstimator {
             //System.out.print(w + ", " + CPthis + "; ");
         }
         //refine what
-        SingleVariateFunction f = new SingleVariateFunction() {
-            public double value(double x) {
-                return -CP(n,x).abs();
+        FunctionAndDerivatives f = new FunctionAndDerivatives() {
+            public double value(Matrix x) {
+                return CP(n,x.get(0,0)).abs2();
+            }
+            public Matrix hessian(Matrix x) {
+                Complex f = CP(n,x.get(0,0));
+                Complex fd = CPdw(n,x.get(0,0));
+                Complex fdd = CPdw2(n,x.get(0,0));
+                double hes = 2*fd.abs() + 2*(fdd*f.conjugate()).re();
+                return Matrix.identity(1,1).times(hes);
+            }
+            public Matrix gradient(Matrix x) {
+                Complex fd = CPdw(n,x.get(0,0));
+                Complex f = CP(n,x.get(0,0));
+                double grad = 2 * (fd * f.conjugate()).re();
+                return Matrix.identity(1,1).times(grad);
             }
         };
-        Brent optimiser = new Brent(f, what - step, what, what + step, 1e-10);
-        return optimiser.xmin();     
+        NewtonRaphson optimiser = new NewtonRaphson(f);
+        return optimiser.maximise(Matrix.identity(1,1).times(what)).get(0,0);
     }
     
     /** 
@@ -167,20 +179,20 @@ public class CubicPhaseFunction extends AbstractPolynomialPhaseEstimator {
 //            y[t] = new Complex(wreal[t],wimag[t]) * Complex.polar(1, -2*Math.PI*freq*(t+1));
 //    }
     
-//    /** First derivative of the cubic phase function with respect to w*/
-//    final protected Complex CPdw(int n, double w){
-//        Complex sum = Complex.zero;
[…]

### 2013-02-06 code/java/src/pubsim/poly/DPTEstimator.java

@@ -25,9 +25,9 @@ public class DPTEstimator extends AbstractPolynomialPhaseEstimator {
     protected Complex[] sig;
     
     /**Max number of iterations for the Newton step */
-    static final int MAX_ITER = 15;
+    static final int MAX_ITER = 20;
     /**Step variable for the Newton step */
-    static final double EPSILON = 1e-10;
+    static final double EPSILON = 1e-15;
 
     public DPTEstimator(int m, int n) {
         super(m);

### 2013-02-06 code/java/test/pubsim/poly/CubicPhaseFunctionTest.java

@@ -73,6 +73,7 @@ public class CubicPhaseFunctionTest {
         double[] oparams = {1.0, Math.PI/8, 0.005, 0.00001};
         CubicPhaseFunction inst = new CubicPhaseFunction(N,10*N);
         double[] params = inst.transformToStandardBasis(oparams);
+        System.out.println(print(params));
         
         PolynomialPhaseSignal siggen = new PolynomialPhaseSignal(N);
         siggen.setParameters(params);

### 2013-02-06 code/montecarlo.scala

@@ -17,13 +17,14 @@ import pubsim.poly.bounds.GaussianCRB
 import pubsim.distributions.GaussianNoise
 import pubsim.distributions.circular.ProjectedNormalDistribution
 import pubsim.lattices.reduction.None
+import pubsim.lattices.reduction.HKZ
 
-val iters = 500 //number of Monte-Carlo trials.
-val Ns = List(59) //values of N we will generate curves for
-val ms = List(3) //order of our polynomial phase signals
+val iters = 50 //number of Monte-Carlo trials.
+val Ns = List(199) //values of N we will generate curves for
+val ms = List(4) //order of our polynomial phase signals
 
 //returns an array of noise distributions with a logarithmic scale
-val SNRdBs = -5 to 15
+val SNRdBs = 0 to 20
 val SNRs = SNRdBs.map(db => scala.math.pow(10.0, db/10.0))
 def noises =  SNRs.map( snr => new GaussianNoise(0,1.0/snr/2.0) ) //variance for real and imaginary parts (divide by 2)
 
@@ -32,7 +33,7 @@ def estfactory(m : Int, N : Int) : List[() => PolynomialPhaseEstimatorInterface]
   var ret = List( 
     () => new KitchenEstimator(m,N),
     () => new DPTEstimator(m,N),
-    //() => new BabaiEstimator(m,N),
+    () => new BabaiEstimator(m,N,new HKZ()),
     () => new MbestEstimator(m,N,4*N) 
   )
   //add the sphere decoder and Least squares estimators if N and m are small
@@ -47,7 +48,10 @@ val starttime = (new java.util.Date).getTime
 //for all the the values of N and m.
 for( N <-  Ns; m <- ms ) {
 
-  val params = Array(0.1,0.1,1e-3,1e-5) //parameters suitable for DPT and CPF
+  def params : Array[Double] = {
+    if(m==3) return Array(2.0,-5e-2,1e-4,1e-6) //parameters suitable for DPT and CPF
+    else return Array(2.0,-5e-2,1e-4,1e-6,1e-8) //parameters suitable for DPT and CPF
+  }
 
   for(estf <- estfactory(m,N) ){
     

### 2013-02-12 paper.tex
modified various stuff.  Made simulations easier to run

@@
 Monte-carlo simualtions assert the excellent statistical performance of our unwrapping estimator when compared to existing state-of-the-art estimators for data sets of less than 200 observations.
-A feature of our estimator is that it functions correctly over a far wider range of parameters than most existing estimators.
+A key feature of our estimator is that it functions correctly over a far wider range of parameters than most existing estimators.
 \end{abstract}
@@
 We are now in a position to complete the proof of strong consistency.
-As is customary, let $(\Omega, \mathcal F, \prob)$ be the probability space over which the random variables $\{X_i\}$ and $\{\Phi_i\}$ are defined.
+As is customary, let $\Omega$ be the sample space on which the random variables $\{X_i\}$ and $\{\Phi_i\}$ are defined.
 Let $A$ be the subset of the sample space $\Omega$ on which  $V_N(\widehat{\lambdabf}_N) - \sigma^2 \rightarrow 0$ as $N\rightarrow\infty$.
@@
 
-This section describes the results of Monte-Carlo simulations with the least squares unwrapping (LSU) estimator, Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, and the discrete polynomial phase transform (DPT) estimator of Peleg and Porat~\cite{Peleg_DPT_1995}.
-The sample sizes considered are $N = 10, 50, 200$ and the unknown amplitude is $\rho = 1$.
-The $X_1, \dots, X_N$ are pseudorandomly generated independent and identically distributed circularly symmetric complex Gaussian random variables with variance $\sigma_c^2$.
-The coefficients $\tilde{\mubf} = [\tilde{\mu}_0, \dots, \tilde{\mu}_m]$ are distributed uniformly randomly in the identifiable region $B$~\cite[Sec.~3]{McKilliam_pps1_2012}.
+This section describes the results of Monte-Carlo simulations with the least squares unwrapping (LSU) estimator, the high order ambiguity function (HAF) estimator~\cite{Peleg_DPT_1995}, the cubic phase function (CPF) estimator~\cite{Oshea_cpf_2004} and estimator that result from combining the HAF and CPF~[BLERG].
+In all simulations the unknown amplitude is $\rho = 1$ and $N = 199$ (an odd number is chosen to satisfy the requirements of the CPF estimator).
+The $X_1, \dots, X_N$ are pseudorandomly generated independent and identically distributed circularly symmetric complex Gaussian random variables with variance $\expect\abs{X_1}^2 = \sigma_c^2$.
+The corresponding signal to noise ratio (SNR) is $\rho^2 \sigma_c^{-2} = \sigma^{-2}$.
 The number of replications of each experiment is $T = 2000$ to obtain estimates $\widehat{\mubf}_1, \dots, \widehat{\mubf}_T$ and the corresponding dealiased errors $\widehat{\lambdabf}_t = \dealias(\widehat{\mubf}_T - \tilde{\mubf})$ are computed.
@@
 
-Figure~\ref{plot:polyest} shows the sample MSEs obtained for a polynomial phase signal of order $m=3$.
-%Results are displayed results for the zeroth and third order coefficients $\widehat{\mu}_1$ and $\widehat{\mu}_3$.
-The results for $\widehat{\mu}_1$ and $\widehat{\mu}_2$ lead to similar conclusions.
-The LSU estimator can be computed by finding a nearest lattice point in a particular lattice~\cite{McKilliam2010thesis,McKilliam2009asilomar_polyest_lattice}.
-When $N = 10$ and $50$ the LSU estimator can be computed exactly using a general purpose algorithm for finding nearest lattice points called the sphere~decoder~\cite{Pohst_sphere_decoder_1981,Agrell2002,Viterbo_sphere_decoder_1999}.
-This is displayed by the circles in the figures.
-When $N=200$ the sphere decoder is computationally intractable and we instead use an approximate nearest point algorithm called the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder}.
-This is displayed by the dots.
-For the purpose of comparison we have also plotted the results for the $K$-best method when $N = 10$ and $50$.
 The asymptotic variance predicted in Theorem~\ref{thm:asymp_proof} is displayed by the dashed line.
@@
 
-Kitchen's unwrapping estimator and the DPT estimator perform very poorly in Figure~\ref{plot:polyest}.
-The reason is that both estimators do not work correctly for all parameters in the identifiable region $B$.
-For this reason Figure~\ref{plot:polyest2} plots the sample MSEs for the estimators with the true coefficients fixed to $\tilde{\mubf} = [0.1,0.2,10^{-3},10^{-5}]$.
-This is within the range suitable for the DPT estimator~\cite{Peleg_DPT_1995}.
-Both Kitchen's unwrapping estimator and the DPT estimator perform better in this case.
-Only results for the zeroth and third order parameters $\tilde{\mu}_0$ and $\tilde{\mu}_3$ are displayed.
-The results for the other parameters are similar.
-
-\begin{figure}[t] 
+\begin{figure}[p] 
    	\centering 
-  		\includegraphics{code/data/plot-1.mps} 
+  		\includegraphics{code/data/plot3-1.mps} 
   		\caption{BLERG} 
@@
 
-\begin{figure}[t] 
+\begin{figure}[p] 
    	\centering 
-  		\includegraphics{code/data/plot-2.mps} 
+  		\includegraphics{code/data/plot3-2.mps} 
   		\caption{BLERG} 
@@
  \end{figure}
+
+\begin{figure}[p] 
+   	\centering 
+  		\includegraphics{code/data/plot3-3.mps} 
+  		\caption{BLERG} 
+  		\label{plot:polyest} 
+ \end{figure}
+
+\begin{figure}[t] 
+   	\centering 
+  		\includegraphics{code/data/plot5-1.mps} 
+  		\caption{BLERG} 
+  		\label{plot:polyest} 
+ \end{figure}
+
+%\begin{figure}[t] 
+%   	\centering 
+%  		\includegraphics{code/data/plot5-2.mps} 
+%  		\caption{BLERG} 
+%  		\label{plot:polyest} 
+% \end{figure}
 
@@
 
-%\begin{figure}[p] 
-%   	\centering 
-%  		\includegraphics[width=\linewidth]{polyfig/polyphase4plot.4} 
-%  		\caption{MSE in the third order coefficient $\mu_3$ versus $\var{X_n} = \sigma^2$. The true coefficients are uniformly spread in the identifiable region.} 
-%  		\label{plot:polyest_projnorm_mu3} 
-% \end{figure}
-
-% \begin{figure}[p] 
-%  	\centering 
-% 		\includegraphics[width=\linewidth]{polyfig/polyphase4plot_unfair.4} 
-% 		\caption{MSE in the third order coefficient $\mu_3$ versus $\var{X_n} = \sigma_c^2$. The coefficient have been restricted for Kitchen's estimator and the DPT.} 
-% 		\label{plot:polyest_projnorm_mu3_unfair} 
-%\end{figure} 
-% 
-% \begin{figure}[p] 
-%  	\centering 
-% 		\includegraphics[width=\linewidth]{polyfig/polyphase4plot_fair.4} 
-% 		\caption{MSE in the third order coefficient $\mu_3$ versus $\var{X_n} = \sigma_c^2$. The DPT estimator runs at the higher sampling rate $\delta$ so that the volumes $V_{DPT}(\delta) = V_m$.} 
-% 		\label{plot:polyest_projnorm_mu3_fair} 
-%\end{figure}
-
-%\addtolength{\parskip}{-0.2cm} 
-%\bibliographystyle{../../bib/IEEEbib} 
 \small 

### 2013-02-12 build.sh

@@ -8,7 +8,8 @@ mpost -interaction=nonstopmode tesselationfigures.mp
 cd .. 
 
 cd code/data 
-mpost -interaction=nonstopmode plot.mp 
+mpost -interaction=nonstopmode plot3.mp 
+mpost -interaction=nonstopmode plot5.mp 
 cd ../../ 
 
 pdflatex -shell-escape paper.tex 

### 2013-02-12 code/montecarlo.scala

@@ -18,111 +18,122 @@ import pubsim.lattices.reduction.None
 import pubsim.lattices.reduction.HKZ
 import pubsim.lattices.reduction.LLL
 
-val iters = 500 //number of Monte-Carlo trials.
-val Ns = List(199) //values of N we will generate curves for
-val ms = List(3) //order of our polynomial phase signals
-
-//Returns a list of functions that return estimators we will run (factory patern to enable parallelism)
-def estfactory(m : Int, N : Int) : List[() => PolynomialPhaseEstimatorInterface] = {
-  var ret = List( 
-    //() => new Kitchen(m,N),
-    () => new HAF(m,N),
-    //() => new Babai(m,N, if(m<=3) new HKZ() else new LLL()),
-    //() => new Mbest(m,N,4*N, if(m<=3) new HKZ() else new LLL()),
-    () => new ZW(m,N,N/m,N/m+1)
-  )
-  //add the sphere decoder and Least squares estimators if N and m are small
-  if( N < 60 ) ret = ret :+ ( () => new SphereDecoder(m,N) )
-  //if( N < 60 && m <= 2 ) ret = ret :+ ( () => new MaximumLikelihood(m,N) )
-if( m==3 ) ret = ret :+ ( () => new CPF(N) )
-  return ret
-}
+val iters = 500
+val N = 199
+val sp3 = (0 to 3).map( k => 0.5/pubsim.Util.factorial(k)/scala.math.pow(N,k-1) ).toArray //3rd order paramater that work for the HAF and CPF
+val bp3 = (0 to 3).map( k => 0.5/pubsim.Util.factorial(k)/2.0 ).toArray //3rd order paramater that do not work for the HAF and CPF
+val sp5 = (0 to 5).map( k => 0.5/pubsim.Util.factorial(k)/scala.math.pow(N,k-1) ).toArray //5rd order paramater that work for the HAF
+val bp5 = (0 to 5).map( k => 0.5/pubsim.Util.factorial(k)/2.0 ).toArray //5rd order paramater that do not work for the HAF
 
 val starttime = (new java.util.Date).getTime
 
-//for all the the values of N and m.
-for( N <-  Ns; m <- ms ) {
+runsim(-5 to 15, sp3, N, iters, () => new HAF(3,N), "HAFm3small")
+runsim(-5 to 15, bp3, N, iters, () => new HAF(3,N), "HAFm3big")
+runsim(-5 to 15, sp3, N, iters, () => new CPF(N), "CPFm3small")
[…]

### 2013-02-13 code/java/src/pubsim/poly/CPFHAF.java
Added the hybrid CPF HAF, and some tests.  Appears to work, but with numerical stability problems?

@@ -1,0 +1,85 @@
+/*
+ */
+package pubsim.poly;
+
+import Jama.Matrix;
+import flanagan.complex.Complex;
+import flanagan.math.FourierTransform;
+import static pubsim.Util.factorial;
+
+/**
+ * Combines the HAF and CPF, i.e. does phase differencing to get to a cubic phase signal
+ * and then uses the CPF.  See:
+ * I. Djurovic and M. Simeunovic and S. Djukanovic and P. Wang 
+ * "A Hybrid CPF-HAF Estimation of Polynomial-Phase Signals: Detailed Statistical Analysis"
+ * IEEE Transactions on Signal Processing, 60:10, Oct. 2012.
+ * TO DO: Estimate m-2 lower order parameters after dechirping
+ * @author Robby McKilliam
+ */
+public class CPFHAF extends HAF {
+    
+    protected final CPF cpf;
+    protected final double[] realcpf;
+    protected final double[] imagcpf;
+    final protected Matrix T, Tinv; //transformation between polynomial bases
+     
+    /** tau is set to default round(n/m) */
+    public CPFHAF(int m, int n) {
+        this(m,n,(n+1)/2/m); //tau default to half usual because of the symmetric definition of PPT here.
+    }
+
+    public CPFHAF(int m, int n, int tau) {
+        super(m,n,tau);
+        if(n%2 == 0) throw new RuntimeException("n must be odd be the cubic phase function");
+        int cpfn = n - 2*tau*(m-3); //length adjusted for size of CPF after apply PPT m-3 times.
+        cpf = new CPF(cpfn); 
+        realcpf = new double[cpfn];
+        imagcpf = new double[cpfn];
+        T = CPF.constructOsheaBasisTransformaton(m,n);
+        Tinv = T.inverse();
+    }
+    
+    @Override
+    public double[] estimate(double[] real, double[] imag) {
+        for (int i = 0; i < n; i++) z[i] = new Complex(real[i], imag[i]); //complex version of recieved signal
+        Complex[] d = PPT(m-2, z, tau); //compute the PPT to get cubic phase signal
+        for(int i = 0; i < d.length; i++){
+            realcpf[i] = d[i].getReal();
+            imagcpf[i] = d[i].getImag();
+        }
+        double [] pcpf = cpf.estimateInOsheaBasis(realcpf, imagcpf);
+        double C2 = Math.pow(2,m-4) * Math.pow(tau, m-3) * factorial(m-1);
+        double C3 = C2 * m / 3.0;
[…]

### 2013-02-13 code/java/src/pubsim/poly/HAF.java

@@ -135,15 +135,10 @@ public class HAF extends AbstractPolynomialPhaseEstimator {
         return PPT2(y,tau);
     }
     protected Complex[] PPT2(Complex[] y, int tau) {
-        Complex[] trans = new Complex[y.length - tau];
-        for (int i = 0; i < y.length; i++) {
-            if (i - tau >= 0) {
+        int N = y.length;
+        Complex[] trans = new Complex[N - tau];
+        for (int i = tau; i < N; i++) 
                 trans[i - tau] = y[i].times(y[i - tau].conjugate());
-            }
-        //else {
-        //    trans[i] = new Complex(0, 0);
-        //}
-        }
         return trans;
     }
 

### 2013-02-13 paper.tex
added new data

@@
 In this paper we consider an estimator that performs phase unwrapping in a least squares manner.
+We call this the \emph{least squares unwrapping} (LSU) estimator.
 Under mild conditions on the distribution of the noise we describe the asymptotic properties of this estimator, showing that it is strongly consistent and asymptotically normally distributed.
 No `small noise' assumption (common in the literature) is required in our analysis.
-As a result our estimator will be accurate, even for very noisy signals, provided that the number of observations is sufficiently large.
-Our unwrapping estimator can be computed in a reasonable amount of time for data sets of moderate size using existing general purpose algorithms from algebraic number theory.
-Monte-carlo simualtions assert the excellent statistical performance of our unwrapping estimator when compared to existing state-of-the-art estimators for data sets of less than 200 observations.
-A key feature of our estimator is that it functions correctly over a far wider range of parameters than most existing estimators.
+As a result the LSU estimator will be accurate, even for very noisy signals, provided that the number of observations is sufficiently large.
+The LSU estimator can be computed in a reasonable amount of time for data sets of moderate size using existing general purpose algorithms from algebraic number theory.
+Monte-carlo simualtions assert the excellent statistical performance of the LSU estimator when compared to existing state-of-the-art estimators.
+A key feature is that the LSU estimator functions accurately over a far wider range of parameters than most existing estimators.
 \end{abstract}
@@
 For this reason many authors have considered alternative approaches to polynomial phase estimation.
-These can loosely be grouped into two classes: estimators based on `multilinear transforms', such as the discrete polynomial phase transform (DPT), high order ambiguity function (HAF)~\cite{Peleg_DPT_1995,Peleg1991_est_class_PPS_1991,Porat_asympt_HAF_DPT_1996,Farquharson_another_poly_est_2005,Boashash_Oshea_1994}, or cubic phase function (CPF)~\cite{Oshea_cpf_2004,Oshea_cpf_2002_letter}; and estimators based on phase unwrapping, such as Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, the estimator of Djuric and Kay~\cite{Djuric_phase_unwrap_chirp_1990}, and Morelande's Bayesian unwrapping estimator~\cite{Morelande_bayes_unwrapping_2009_tsp}.
+These can loosely be grouped into two classes: estimators based on `multilinear transforms', such as the high order ambiguity function (HAF)~\cite{Peleg_DPT_1995,Peleg1991_est_class_PPS_1991,Porat_asympt_HAF_DPT_1996,Farquharson_another_poly_est_2005,Boashash_Oshea_1994}, and cubic phase function (CPF)~\cite{Oshea_cpf_2004,Oshea_cpf_2002_letter,Djurovic_haf_cpf_2012}; and estimators based on phase unwrapping, such as Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, the estimator of Djuric and Kay~\cite{Djuric_phase_unwrap_chirp_1990}, and Morelande's Bayesian unwrapping estimator~\cite{Morelande_bayes_unwrapping_2009_tsp}.
 
 In this paper we consider the estimator that results from unwrapping the phase in a least squares manner.
-We call this the \emph{least squares unwrapping estimator} (LSU)~\cite{McKilliam2009asilomar_polyest_lattice, McKilliamFrequencyEstimationByPhaseUnwrapping2009}\cite[Chap.
-8]{McKilliam2010thesis}.
-It was shown in \cite{McKilliam2009asilomar_polyest_lattice, McKilliamFrequencyEstimationByPhaseUnwrapping2009} that the LSU estimator can be represented as a \emph{nearest lattice point problem}, and Monte-Carlo simulations were used to show the LSU estimator's favourable statistical performance.
+We call this the \emph{least squares unwrapping} (LSU) estimator~\cite{McKilliam2009asilomar_polyest_lattice, McKilliamFrequencyEstimationByPhaseUnwrapping2009,McKilliam2010thesis}.
+It was shown in \cite{McKilliam2009asilomar_polyest_lattice, McKilliamFrequencyEstimationByPhaseUnwrapping2009} that the LSU estimator can be represented as a \emph{nearest lattice point problem}~\cite{Agrell2002}, and Monte-Carlo simulations were used to show the LSU estimator's favourable statistical performance.
 %Moreover, it was noticed that the LSU appears to work correctly for coefficients anywhere in the identifiable region.
 A drawback of the LSU estimator is that computing a nearest lattice point is, in general, computationally difficult.
-In \cite{McKilliam2009asilomar_polyest_lattice}, two standard techniques were considered, the \emph{sphere decoder}~\cite{Pohst_sphere_decoder_1981} that exactly computes a nearest lattice point, and \emph{Babai's nearest plane algorithm}~\cite{Babai1986} that only produces and approximation.
-The sphere decoder was observed to have excellent statistical performance but can only be computed efficiently for small $N$ less than about $60$~\cite{Jalden2005_sphere_decoding_complexity}.
+In \cite{McKilliam2009asilomar_polyest_lattice}, two standard techniques were considered, the \emph{sphere decoder}~\cite{Pohst_sphere_decoder_1981,Viterbo_sphere_decoder_1999} that exactly computes a nearest lattice point, and \emph{Babai's nearest plane algorithm}~\cite{Babai1986} that only produces an approximation.
+The sphere decoder was observed to have excellent statistical performance but can only be computed efficiently for small $N$ less than about $50$.
 Computing the Babai point requires only $O(N^2)$ operations, but its statistical performance was observed to be comparatively poor at low signal to noise ratio (SNR).
 A major point of interest is that the lattices considered are not \emph{random}, and may admit fast nearest point algorithms.
-This has been studied in~\cite[Sec.~4.2]{McKilliam2010thesis} where polynomial time algorithms where found that compute the nearest point exactly.
-Unfortantely, although polynomial time, the algorithm are still computationally very slow in practice.
-Nevertheless fast (exact or approximate) algorithms that we have not found may still exist for these special lattices.
-
+This has been studied in~\cite{McKilliam2010thesis} where polynomial time algorithms where found that compute the nearest point exactly.
+Although polynomial time, the algorithm are still computationally demanding in practice.
+Nevertheless fast (exact or approximate) algorithms may still exist for these special lattices, that are yet to be discovered.
 In this paper we apply another general purpose approximate nearest point algorithm, called the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder} for approximating the LSU estimator.
-We show that this provides near sphere decoder (exact) performance, but can be computed in a reasonable amount of time if $N$ is not too large.
-BLERG benchmarking results.
-The $K$-best method is motivated by sequential decoding algorithms and, in the context of polynomial phase estimation, bares some resemblance to the Bayesian unwrapping algorithm studied by Morelande~\cite{Morelande_bayes_unwrapping_2009_tsp}.
-We show by Montecarlo simulation that the LSU estimator approximated using the $K$-best method is statistically superior to state-of-the-art estimators from the literature including the HAF and CPF.
+We show that this provides near sphere decoder (exact) performance, but can be computed in a reasonable amount of time if $N$ is not too large (approximately $N$ less than $1000$).
+%The $K$-best method is motivated by sequential decoding algorithms and, in the context of polynomial phase estimation, bares some resemblance to the Bayesian unwrapping algorithm studied by Morelande~\cite{Morelande_bayes_unwrapping_2009_tsp}.
+%We show by Montecarlo simulation that the LSU estimator approximated using the $K$-best method is statistically superior to state-of-the-art estimators from the literature including the HAF and CPF.
 We also compare the computational performance of the estimators and find that BLERG.
@@
 Under mild assumptions about the distribution of the noise we prove that the LSU estimator is strongly consistent, that is, the estimated coefficients converge (almost surely) to the true coefficients as the number of observations $N$ grows.
-Similar results were stated without proof in~\cite{McKilliam_polyphase_est_icassp_2011}.
+%Similar results were stated without proof in~\cite{McKilliam_polyphase_est_icassp_2011}.
 In the   BLERG `small noise' assumption.
@@
 However, asserting that convergence at this rate occurs in the polynomial phase setting is not trivial.
-For this purpose we make use of an elementary result about the number of arithmetic progressions contained inside subsets of $\{1,2,\dots,N\}$~\cite{Erdos_on_some_sequence_of_integers1936,Szemeredi_setint_no_k_arth1975,Gowers_new_proof2001}.
+For this purpose we make use of an elementary result about the number of arithmetic progressions contained inside subsets of $\{1,2,\dots,N\}$~\cite{Gowers_new_proof2001}.
 %This proof technique appears to be novel, and may be useful in other problems.
@@
 Potential candidates are the phase wrapped images observed in modern radar and medical imaging devices such as synthetic aperture radar and magnetic resonance imaging~\cite{Nico_phaseunwrappingSAR_2000,Friedlander_PD_phaseunwrapping_1996}.
-
 We also assert, under similarly mild conditions, that the LSU estimator is asymptotically normal.
-BLERG CLOSENESS TO CLRB in Gaussian case.
+%In the case where the noise is normally distributed the asymptotic variance of the LSU estimator is close to the Cramer-Rao bound at high signal to noise ratio (SNR) and is close to $\tfrac{2}{3}\pi$ times that Cramer-Rao bound at low (SNR).
 Space restrictions make it impossible for us to include a proof of asymptotic normality in this paper.
@@
 
-When $m=0$, polynomial phase estimation is equivalent to the problem of \emph{direction estimation} from circular statistics, and the least squares estimator corresponds with the \emph{extrinsic mean} (or \emph{sample circular mean})~\cite{McKilliam_mean_dir_est_sq_arc_length2010,Fisher1993,Mardia_directional_statistics}.
-In this case the LSE can be computed efficiently in $O(N)$ by averaging $N$ complex numbers.
-When $m=1$, polynomial phase estimation is equivalent to single frequency estimation, and the least squares estimator can be computed efficiently in $O(N\log N)$ operations by maximising the \emph{periodogram} using the fast Fourier transform followed by an appropriate numerical optimisation technique~\cite{Quinn2008maximizing_the_periodogram,Rife1974,Quinn2001}.
-When $m=2$, the least squares estimator can be computed in a reasonable amount of time when $N$ is approximately less than $50$.
-The LSE is computationally intensive for any $N$ when $m > 2$.
-By comparison the exact LSU estimator implemented using the sphere decoder can be computed quickly for any $m$ with $N$ approximately less than $50$.
+The exact LSU estimator implemented using the sphere decoder can be computed quickly for any $m$ with $N$ approximately less than $50$.
 In situations where $N$ is small, but high statistical accuracy is required, the sphere decoder is computationally a better choice than the LSE for polynomial phase signals of order greater than or equal to $2$.

### 2013-02-13 code/montecarlo.scala

@@ -9,6 +9,7 @@ import pubsim.poly.MaximumLikelihood
 import pubsim.poly.SphereDecoder
 import pubsim.poly.PolynomialPhaseEstimatorInterface
 import pubsim.poly.CPF
+import pubsim.poly.CPFHAF
 import pubsim.poly.ZW
 import pubsim.poly.bounds.AngularLeastSquaresVariance
 import pubsim.poly.bounds.GaussianCRB
@@ -18,15 +19,16 @@ import pubsim.lattices.reduction.None
 import pubsim.lattices.reduction.HKZ
 import pubsim.lattices.reduction.LLL
 
-val iters = 2000
+val iters = 100
 val N = 199
 val sp3 = (0 to 3).map( k => 0.5/pubsim.Util.factorial(k)/scala.math.pow(N,k-1) ).toArray //3rd order paramater that work for the HAF and CPF
 val bp3 = (0 to 3).map( k => 0.5/pubsim.Util.factorial(k)/2.0 ).toArray //3rd order paramater that do not work for the HAF and CPF
 val sp5 = (0 to 5).map( k => 0.5/pubsim.Util.factorial(k)/scala.math.pow(N,k-1) ).toArray //5rd order paramater that work for the HAF
 val bp5 = (0 to 5).map( k => 0.5/pubsim.Util.factorial(k)/2.0 ).toArray //5rd order paramater that do not work for the HAF
+val sp6 = (0 to 6).map( k => 0.5/pubsim.Util.factorial(k)/scala.math.pow(N,k-1) ).toArray //5rd order paramater that work for the HAF
 
 val starttime = (new java.util.Date).getTime
-
+/*
 runsim(-5 to 15, sp3, N, iters, () => new HAF(3,N), "HAFm3small")
 runsim(-5 to 15, bp3, N, iters, () => new HAF(3,N), "HAFm3big")
 runsim(-5 to 15, sp3, N, iters, () => new CPF(N), "CPFm3small")
@@ -45,10 +47,19 @@ runcrb(-5 to 35, 3, N, "crbm3range")
 runlsuclt(-5 to 35, 3, N, "lsucltm3range")
 
 runsim(5 to 30, sp5, N, iters, () => new HAF(5,N), "HAFm5small")
+runsim(5 to 30, sp5, N, iters, () => new CPFHAF(5,N), "CPFHAFm5small")
[…]

### 2013-02-13 code/montecarlo.scala
added new data

@@ -19,7 +19,7 @@ import pubsim.lattices.reduction.None
 import pubsim.lattices.reduction.HKZ
 import pubsim.lattices.reduction.LLL
 
-val iters = 100
+val iters = 2000
 val N = 199
 val sp3 = (0 to 3).map( k => 0.5/pubsim.Util.factorial(k)/scala.math.pow(N,k-1) ).toArray //3rd order paramater that work for the HAF and CPF
 val bp3 = (0 to 3).map( k => 0.5/pubsim.Util.factorial(k)/2.0 ).toArray //3rd order paramater that do not work for the HAF and CPF

### 2013-02-13 paper.tex
edited plot and simulations section

@@
 
-This section describes the results of Monte-Carlo simulations with the least squares unwrapping (LSU) estimator, the high order ambiguity function (HAF) estimator~\cite{Peleg_DPT_1995}, the cubic phase function (CPF) estimator~\cite{Oshea_cpf_2004} and estimator that result from combining the HAF and CPF~[BLERG].
-In all simulations the unknown amplitude is $\rho = 1$ and $N = 199$ (an odd number is chosen to satisfy the requirements of the CPF estimator).
+This section describes the results of Monte-Carlo simulations with the least squares unwrapping (LSU) estimator, the high order ambiguity function (HAF) estimator~\cite{Peleg_DPT_1995}, the cubic phase function (CPF) estimator~\cite{Oshea_cpf_2004} and the estimator that results from combining the HAF and CPF~\cite{Djurovic_haf_cpf_2012}.
+The CPF estimator only applies to polynomial phase signals of order $m=3$.
+In all simulations the unknown amplitude is $\rho = 1$ and the number of observations is $N = 199$ (an odd number is chosen to satisfy the requirements of the CPF estimator).
 The $X_1, \dots, X_N$ are pseudorandomly generated independent and identically distributed circularly symmetric complex Gaussian random variables with variance $\expect\abs{X_1}^2 = \sigma_c^2$.
@@
 The sample mean square error (MSE) of the $k$th coefficient is computed according to $\tfrac{1}{T}\sum_{t=1}^T \widehat{\lambda}_{k,t}^2$ where $\widehat{\lambda}_{k,t}$ is the $k$th element of $\widehat{\lambdabf}_t$.
-
+In all simulation the LSU estimator is approximately computed using the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder}.
+
+Figure~\ref{plot:m3small} show the sample MSE of the highest order coefficient $\mu_3$ for a polynomial phase signal of order $m=3$.
+All the estimators perform poorly at low SNR until a `threshold' is reached, after which the variance is close the Cramer-Rao bound~\cite{Peleg1991_CRB_PPS_1991} (indicated by the solid line).
 The asymptotic variance predicted in Theorem~\ref{thm:asymp_proof} is displayed by the dashed line.
 Provided the noise variance is small enough (so that the `threshold' is avoided) the sample MSE of the LSU estimator is close to that predicted by Theorem~\ref{thm:asymp_proof}.
-The Cram\'{e}r-Rao lower bound for the variance of unbiased polynomial phase estimators in Gaussian noise is also plotted using the solid line~\cite{Peleg1991_CRB_PPS_1991}.
-When the noise variance is small the asymptotic variance of the LSU estimator is close to the Cram\'{e}r-Rao lower bound.
+Figure~\ref{plot:m3small} the CPF estimator appears to have the best performance at low SNR, reaching a threshold at \unit[1]{dB}.
+The LSU estimator reaches a threshold at \unit[3]{dB} and the HAF estimator at \unit[5]{dB}.
+In Figure~\ref{plot:m3small} the true coefficients are $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}  \right]$.
+These are specifically chosen to satisfy the requirements of the HAF and CPF estimator.
+The HAF estimator requires that the coefficients satisfy
+\[
+2|\mu_k| \leq \frac{1}{k! \tau^{k-1}} \leq \frac{1}{k!}\left( \frac{m}{N} \right)^{k-1}
+\]
+where $\tau$ is the \emph{lag}.
+Here, we choose $\tau = \floor{N/m}$ as suggested in~\cite{Peleg_DPT_1995}.
+The CPF estimator requires similar, but slightly more complicated conditions on the coeffcients~\cite{Oshea_cpf_2004}.
+It is informative to compare the volume of the set of coefficients suitable for the HAF and CPF estimators with the volume of the identifiable region $B$ from~\eqref{eq:rectangular_identifiable_region}.
+The identifiable region has volume
+\[
+\vol(B) = \prod_{k=0}^m {\frac{1}{k!}}.
+\]
+The volume of coefficients suitable for the HAF estimator is less than
+\[
+\prod_{k=0}^m \frac{1}{k!}\left( \frac{m}{N} \right)^{k-1} = \vol(B) \prod_{k=0}^m \left( \frac{m}{N} \right)^{k-1}.
+\]
+The volume of the coefficients suitable for the CPF estimator is $\tfrac{3}{2N^3} = 18 \vol(B) N^{-3}$.
+It is worth reconsidering the results in Figure~\ref{plot:m3small}.
+Substituting $m=3$ and $N=199$ in the formula above, we see that the HAF and CPF estimators search a space of parameters that is almost $5 \times 10^5$ times smaller than that searched by the LSU estimator.
+In this context we might expect the CPF and HAF estimators to have thresholds at lower SNR than LSU estimator.
+It is encouraging that the LSU estimator has comparible performance.
+To put these results further into context, Figure~\ref{plot:m3big} repeats the simulations in Figure~\ref{plot:m3small}, but where the true coefficients are $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$.
+The CPF and HAF estimators fail spectacularly.
+The performance of the LSU estimator is unchanged, as expected in view of Theorem~\ref{thm:asymp_proof}.
+
+The limited set of allowable parameters for the HAF estimator has been studied previously~\cite{Angeby_PPS_aliasing_2000,Djurovic_aliasing_2012,McKilliam2009IndentifiabliltyAliasingPolyphase,Zhou_lag_diverity_PPT_1997,Xia_dynamic_range_pps_2001} and can be a significant limitation for some applications, for example, the problem of interference supression in the pressence of a jammer~\cite{Djukanovic:2011:PMN:1950994.1951252,Djurovic_aliasing_2012}.
+Zhou and Wang~\cite{Zhou_lag_diverity_PPT_1997} describe a method based on the extended Euclidean algorithm for increasing the range of allowable coefficients for the HAF estimator to be the entire identifiable region $B$.
+Unfortunately this comes at the price of statistical accuracy.
+Figure~\ref{plot:m3bigzw} shows the performance of their estimator when the coefficients are $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$ as in Figure~\ref{plot:m3big}.
+The threshold for this estimator is approximately~\unit[27]{dB}.
+There does not appear to exist (atleast in the literature) a similar method for increasing the range of coefficients for the CPF estimator.
 
@@
   		\caption{BLERG} 
-  		\label{plot:polyest} 
+  		\label{plot:m3small} 
  \end{figure}
@@
   		\caption{BLERG} 
-  		\label{plot:polyest} 
+  		\label{plot:m3big} 
  \end{figure}
@@
   		\caption{BLERG} 
-  		\label{plot:polyest} 
+  		\label{plot:m3bigzw} 
  \end{figure}
@@
   		\includegraphics{code/data/plot5-1.mps} 
+  		\caption{BLERG} 
+  		\label{plot:polyest} 
+ \end{figure}
+
+\begin{figure}[t] 
+   	\centering 
+  		\includegraphics{code/data/plot6-1.mps} 
   		\caption{BLERG} 

### 2013-02-13 build.sh

@@ -10,6 +10,7 @@ cd ..
 cd code/data 
 mpost -interaction=nonstopmode plot3.mp 
 mpost -interaction=nonstopmode plot5.mp 
+mpost -interaction=nonstopmode plot6.mp 
 cd ../../ 
 
 pdflatex -shell-escape paper.tex 

### 2013-02-13 code/montecarlo.scala

@@ -21,11 +21,12 @@ import pubsim.lattices.reduction.LLL
 
 val iters = 100
 val N = 199
-val sp3 = (0 to 3).map( k => 0.5/pubsim.Util.factorial(k)/scala.math.pow(N,k-1) ).toArray //3rd order paramater that work for the HAF and CPF
-val bp3 = (0 to 3).map( k => 0.5/pubsim.Util.factorial(k)/2.0 ).toArray //3rd order paramater that do not work for the HAF and CPF
-val sp5 = (0 to 5).map( k => 0.5/pubsim.Util.factorial(k)/scala.math.pow(N,k-1) ).toArray //5rd order paramater that work for the HAF
-val bp5 = (0 to 5).map( k => 0.5/pubsim.Util.factorial(k)/2.0 ).toArray //5rd order paramater that do not work for the HAF
-val sp6 = (0 to 6).map( k => 0.5/pubsim.Util.factorial(k)/scala.math.pow(N,k-1) ).toArray //5rd order paramater that work for the HAF
+def npow(x : Int, t : Int) = if(k<=0) 1.0 else scala.math.pow(x,t)
+val sp3 = (0 to 3).map( k => 0.25/pubsim.Util.factorial(k)/npow(N,k-1) ).toArray //3rd order paramater that work for the HAF and CPF
+val bp3 = (0 to 3).map( k => 0.25/pubsim.Util.factorial(k) ).toArray //3rd order paramater that do not work for the HAF and CPF
+val sp5 = (0 to 5).map( k => 0.25/pubsim.Util.factorial(k)/npow(N,k-1) ).toArray //5rd order paramater that work for the HAF
+val bp5 = (0 to 5).map( k => 0.25/pubsim.Util.factorial(k) ).toArray //5rd order paramater that do not work for the HAF
+val sp6 = (0 to 6).map( k => 0.25/pubsim.Util.factorial(k)/npow(N,k-1) ).toArray //5rd order paramater that work for the HAF
 
 val starttime = (new java.util.Date).getTime
 /*

### 2013-02-13 code/montecarlo.scala
added m=7 to montecarlo

@@ -53,7 +53,6 @@ runsim(5 to 30, sp5, N, iters, () => new Babai(5,N, new LLL()), "Babaim5small")
 runsim(5 to 30, sp5, N, iters, () => new Mbest(5,N, 20*N, new LLL()), "Mbestm5small")
 runcrb(5 to 30, 5, N, "crbm5")
 runlsuclt(5 to 30, 5, N, "lsucltm5")
-*/
 
 runsim(5 to 30, sp6, N, iters, () => new HAF(6,N), "HAFm6small")
 runsim(5 to 30, sp6, N, iters, () => new CPFHAF(6,N), "CPFHAFm6small")
@@ -61,6 +60,14 @@ runsim(5 to 30, sp6, N, iters, () => new Babai(6,N, new LLL()), "Babaim6small")
 runsim(5 to 30, sp6, N, iters, () => new Mbest(6,N, 20*N, new LLL()), "Mbestm6small")
 runcrb(5 to 30, 6, N, "crbm6")
 runlsuclt(5 to 30, 6, N, "lsucltm6")
+*/
+
+runsim(5 to 30, sp7, N, iters, () => new HAF(7,N), "HAFm7small")
+runsim(5 to 30, sp7, N, iters, () => new CPFHAF(7,N), "CPFHAFm7small")
+runsim(5 to 30, sp7, N, iters, () => new Babai(7,N, new LLL()), "Babaim7small")
+runsim(5 to 30, sp7, N, iters, () => new Mbest(7,N, 20*N, new LLL()), "Mbestm7small")
+runcrb(5 to 30, 7, N, "crbm7")
+runlsuclt(5 to 30, 7, N, "lsucltm7")
 
 val runtime = (new java.util.Date).getTime - starttime
 println("Simulation finshed in " + (runtime/1000.0) + " seconds.\n")

### 2013-02-13 code/montecarlo.scala
minor

@@ -27,6 +27,7 @@ val bp3 = (0 to 3).map( k => 0.25/pubsim.Util.factorial(k) ).toArray //3rd order
 val sp5 = (0 to 5).map( k => 0.25/pubsim.Util.factorial(k)/npow(N,k-1) ).toArray //5rd order paramater that work for the HAF
 val bp5 = (0 to 5).map( k => 0.25/pubsim.Util.factorial(k) ).toArray //5rd order paramater that do not work for the HAF
 val sp6 = (0 to 6).map( k => 0.25/pubsim.Util.factorial(k)/npow(N,k-1) ).toArray //5rd order paramater that work for the HAF
+val sp7 = (0 to 7).map( k => 0.25/pubsim.Util.factorial(k)/npow(N,k-1) ).toArray //5rd order paramater that work for the HAF
 
 val starttime = (new java.util.Date).getTime
 /*

### 2013-02-13 code/montecarlo.scala
added new data

@@ -21,7 +21,7 @@ import pubsim.lattices.reduction.LLL
 
 val iters = 2000
 val N = 199
-def npow(x : Int, t : Int) = if(k<=0) 1.0 else scala.math.pow(x,t)
+def npow(x : Int, t : Int) = if(t<=0) 1.0 else scala.math.pow(x,t)
 val sp3 = (0 to 3).map( k => 0.25/pubsim.Util.factorial(k)/npow(N,k-1) ).toArray //3rd order paramater that work for the HAF and CPF
 val bp3 = (0 to 3).map( k => 0.25/pubsim.Util.factorial(k) ).toArray //3rd order paramater that do not work for the HAF and CPF
 val sp5 = (0 to 5).map( k => 0.25/pubsim.Util.factorial(k)/npow(N,k-1) ).toArray //5rd order paramater that work for the HAF
@@ -63,12 +63,12 @@ runcrb(5 to 30, 6, N, "crbm6")
 runlsuclt(5 to 30, 6, N, "lsucltm6")
 */
 
-runsim(5 to 30, sp7, N, iters, () => new HAF(7,N), "HAFm7small")
-runsim(5 to 30, sp7, N, iters, () => new CPFHAF(7,N), "CPFHAFm7small")
-runsim(5 to 30, sp7, N, iters, () => new Babai(7,N, new LLL()), "Babaim7small")
-runsim(5 to 30, sp7, N, iters, () => new Mbest(7,N, 20*N, new LLL()), "Mbestm7small")
-runcrb(5 to 30, 7, N, "crbm7")
-runlsuclt(5 to 30, 7, N, "lsucltm7")
+runsim(10 to 35, sp7, N, iters, () => new HAF(7,N), "HAFm7small")
+runsim(10 to 35, sp7, N, iters, () => new CPFHAF(7,N), "CPFHAFm7small")
+runsim(10 to 35, sp7, N, iters, () => new Babai(7,N, new LLL()), "Babaim7small")
+runsim(10 to 35, sp7, N, iters, () => new Mbest(7,N, 20*N, new LLL()), "Mbestm7small")
+runcrb(10 to 35, 7, N, "crbm7")
+runlsuclt(10 to 35, 7, N, "lsucltm7")
 
 val runtime = (new java.util.Date).getTime - starttime
 println("Simulation finshed in " + (runtime/1000.0) + " seconds.\n")

### 2013-02-13 paper.tex
moved stuff

@@
 The $X_1, \dots, X_N$ are pseudorandomly generated independent and identically distributed circularly symmetric complex Gaussian random variables with variance $\expect\abs{X_1}^2 = \sigma_c^2$.
-The corresponding signal to noise ratio (SNR) is $\rho^2 \sigma_c^{-2} = \sigma^{-2}$.
+The corresponding signal to noise ratio (SNR) is $\rho^2 \sigma_c^{-2} = \sigma_c^{-2}$.
 The number of replications of each experiment is $T = 2000$ to obtain estimates $\widehat{\mubf}_1, \dots, \widehat{\mubf}_T$ and the corresponding dealiased errors $\widehat{\lambdabf}_t = \dealias(\widehat{\mubf}_T - \tilde{\mubf})$ are computed.
@@
 Figure~\ref{plot:m3small} show the sample MSE of the highest order coefficient $\mu_3$ for a polynomial phase signal of order $m=3$.
-All the estimators perform poorly at low SNR until a `threshold' is reached, after which the variance is close the Cramer-Rao bound~\cite{Peleg1991_CRB_PPS_1991} (indicated by the solid line).
-The asymptotic variance predicted in Theorem~\ref{thm:asymp_proof} is displayed by the dashed line.
+All the estimators perform poorly at low SNR until a `threshold' is reached, after which the variance is close the Cramer-Rao bound (CRB)~\cite{Peleg1991_CRB_PPS_1991} indicated by the solid line.
+The asymptotic variance of the LSU estimator predicted by Theorem~\ref{thm:asymp_proof} is displayed by the dashed line.
 Provided the noise variance is small enough (so that the `threshold' is avoided) the sample MSE of the LSU estimator is close to that predicted by Theorem~\ref{thm:asymp_proof}.
-Figure~\ref{plot:m3small} the CPF estimator appears to have the best performance at low SNR, reaching a threshold at \unit[1]{dB}.
+In Figure~\ref{plot:m3small} the CPF estimator appears to have the best performance at low SNR, reaching a threshold at \unit[1]{dB}.
 The LSU estimator reaches a threshold at \unit[3]{dB} and the HAF estimator at \unit[5]{dB}.
 In Figure~\ref{plot:m3small} the true coefficients are $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}  \right]$.
-These are specifically chosen to satisfy the requirements of the HAF and CPF estimator.
-The HAF estimator requires that the coefficients satisfy
-\[
-2|\mu_k| \leq \frac{1}{k! \tau^{k-1}} \leq \frac{1}{k!}\left( \frac{m}{N} \right)^{k-1}
-\]
-where $\tau$ is the \emph{lag}.
-Here, we choose $\tau = \floor{N/m}$ as suggested in~\cite{Peleg_DPT_1995}.
+These are specifically chosen to satisfy the requirements of the HAF and CPF estimators.
+The HAF estimator requires that the coefficients satisfy $\tilde{\mu}_0, \tilde{\mu}_1 \in [-0.5, 0.5)$ and
+\[
+\tilde{\mu}_k \in \left[ -\frac{0.5}{k! \tau^{k-1}}, \frac{0.5}{k! \tau^{k-1}}\right) \qquad k = 2,3,\dots,m,
+\]
+where $\tau$ is the \emph{lag} and we choose $\tau = \floor{N/m}$ as suggested in~\cite{Peleg_DPT_1995}.
 The CPF estimator requires similar, but slightly more complicated conditions on the coeffcients~\cite{Oshea_cpf_2004}.
@@
 \]
-The volume of coefficients suitable for the HAF estimator is less than
-\[
-\prod_{k=0}^m \frac{1}{k!}\left( \frac{m}{N} \right)^{k-1} = \vol(B) \prod_{k=0}^m \left( \frac{m}{N} \right)^{k-1}.
-\]
-The volume of the coefficients suitable for the CPF estimator is $\tfrac{3}{2N^3} = 18 \vol(B) N^{-3}$.
-It is worth reconsidering the results in Figure~\ref{plot:m3small}.
+The volume of coefficients suitable for the HAF estimator is
+\[
+\prod_{k=2}^m \frac{1}{k! \tau^{k-1}} =  \vol(B) \prod_{k=2}^m \frac{1}{\tau^{k-1}} \leq \vol(B) \prod_{k=2}^m \left( \frac{m}{N} \right)^{k-1}.
+\]
+The volume of the coefficients suitable for the CPF estimator can be shown to be $\tfrac{3}{2N^3} = 18 \vol(B) N^{-3}$.
+It is now worth reconsidering the results in Figure~\ref{plot:m3small}.
 Substituting $m=3$ and $N=199$ in the formula above, we see that the HAF and CPF estimators search a space of parameters that is almost $5 \times 10^5$ times smaller than that searched by the LSU estimator.
@@
 
-The limited set of allowable parameters for the HAF estimator has been studied previously~\cite{Angeby_PPS_aliasing_2000,Djurovic_aliasing_2012,McKilliam2009IndentifiabliltyAliasingPolyphase,Zhou_lag_diverity_PPT_1997,Xia_dynamic_range_pps_2001} and can be a significant limitation for some applications, for example, the problem of interference supression in the pressence of a jammer~\cite{Djukanovic:2011:PMN:1950994.1951252,Djurovic_aliasing_2012}.
-Zhou and Wang~\cite{Zhou_lag_diverity_PPT_1997} describe a method based on the extended Euclidean algorithm for increasing the range of allowable coefficients for the HAF estimator to be the entire identifiable region $B$.
+The comparitively small set of coefficients suitable for the HAF and CPF estimators has been studied previously~\cite{Angeby_PPS_aliasing_2000,Djurovic_aliasing_2012,McKilliam2009IndentifiabliltyAliasingPolyphase,Zhou_lag_diverity_PPT_1997,Xia_dynamic_range_pps_2001} and can be a significant limitation for some applications, for example, interference supression in the pressence of a jammer~\cite{Djukanovic:2011:PMN:1950994.1951252,Djurovic_aliasing_2012}.
+Zhou and Wang~\cite{Zhou_lag_diverity_PPT_1997} describe a method based on the extended Euclidean algorithm that enables the HAF estimator to operate on the entire identifiable region $B$.
 Unfortunately this comes at the price of statistical accuracy.
-Figure~\ref{plot:m3bigzw} shows the performance of their estimator when the coefficients are $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$ as in Figure~\ref{plot:m3big}.
-The threshold for this estimator is approximately~\unit[27]{dB}.
+Figure~\ref{plot:m3bigzw} shows the performance of the Zhou and Wang estimator when the coefficients are $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$, the same as in Figure~\ref{plot:m3big}.
+The threshold for the estimator is approximately~\unit[29]{dB}.
 There does not appear to exist (atleast in the literature) a similar method for increasing the range of coefficients for the CPF estimator.
@@
    	\centering 
-  		\includegraphics{code/data/plot5-1.mps} 
+  		\includegraphics{code/data/plot567-1.mps} 
   		\caption{BLERG} 
@@
    	\centering 
-  		\includegraphics{code/data/plot6-1.mps} 
+  		\includegraphics{code/data/plot567-2.mps} 
+  		\caption{BLERG} 
+  		\label{plot:polyest} 
+ \end{figure}
+
+\begin{figure}[t] 
+   	\centering 
+  		\includegraphics{code/data/plot567-3.mps} 
   		\caption{BLERG} 

### 2013-02-13 code/montecarlo.scala

@@ -21,7 +21,7 @@ import pubsim.lattices.reduction.LLL
 
 val iters = 2000
 val N = 199
-def npow(x : Int, t : Int) = if(k<=0) 1.0 else scala.math.pow(x,t)
+def npow(x : Int, t : Int) = if(t<=0) 1.0 else scala.math.pow(x,t)
 val sp3 = (0 to 3).map( k => 0.25/pubsim.Util.factorial(k)/npow(N,k-1) ).toArray //3rd order paramater that work for the HAF and CPF
 val bp3 = (0 to 3).map( k => 0.25/pubsim.Util.factorial(k) ).toArray //3rd order paramater that do not work for the HAF and CPF
 val sp5 = (0 to 5).map( k => 0.25/pubsim.Util.factorial(k)/npow(N,k-1) ).toArray //5rd order paramater that work for the HAF

### 2013-02-13 build.sh
minor

@@ -9,8 +9,7 @@ cd ..
 
 cd code/data 
 mpost -interaction=nonstopmode plot3.mp 
-mpost -interaction=nonstopmode plot5.mp 
-mpost -interaction=nonstopmode plot6.mp 
+mpost -interaction=nonstopmode plot567.mp 
 cd ../../ 
 
 pdflatex -shell-escape paper.tex 

### 2013-02-13 code/montecarlo.scala

@@ -30,7 +30,7 @@ val sp6 = (0 to 6).map( k => 0.25/pubsim.Util.factorial(k)/npow(N,k-1) ).toArray
 val sp7 = (0 to 7).map( k => 0.25/pubsim.Util.factorial(k)/npow(N,k-1) ).toArray //5rd order paramater that work for the HAF
 
 val starttime = (new java.util.Date).getTime
-/*
+
 runsim(-5 to 15, sp3, N, iters, () => new HAF(3,N), "HAFm3small")
 runsim(-5 to 15, bp3, N, iters, () => new HAF(3,N), "HAFm3big")
 runsim(-5 to 15, sp3, N, iters, () => new CPF(N), "CPFm3small")
@@ -61,7 +61,6 @@ runsim(5 to 30, sp6, N, iters, () => new Babai(6,N, new LLL()), "Babaim6small")
 runsim(5 to 30, sp6, N, iters, () => new Mbest(6,N, 20*N, new LLL()), "Mbestm6small")
 runcrb(5 to 30, 6, N, "crbm6")
 runlsuclt(5 to 30, 6, N, "lsucltm6")
-*/
 
 runsim(5 to 30, sp7, N, iters, () => new HAF(7,N), "HAFm7small")
 runsim(5 to 30, sp7, N, iters, () => new CPFHAF(7,N), "CPFHAFm7small")

### 2013-02-13 code/montecarlo.scala
minor

@@ -35,36 +35,36 @@ runsim(-5 to 15, sp3, N, iters, () => new HAF(3,N), "HAFm3small")
 runsim(-5 to 15, bp3, N, iters, () => new HAF(3,N), "HAFm3big")
 runsim(-5 to 15, sp3, N, iters, () => new CPF(N), "CPFm3small")
 runsim(-5 to 15, bp3, N, iters, () => new CPF(N), "CPFm3big")
-runsim(-5 to 15, sp3, N, iters, () => new Babai(3,N, new HKZ()), "Babaim3small")
-runsim(-5 to 15, bp3, N, iters, () => new Babai(3,N, new HKZ()), "Babaim3big")
+//runsim(-5 to 15, sp3, N, iters, () => new Babai(3,N, new HKZ()), "Babaim3small")
+//runsim(-5 to 15, bp3, N, iters, () => new Babai(3,N, new HKZ()), "Babaim3big")
 runsim(-5 to 15, sp3, N, iters, () => new Mbest(3,N, 20*N, new HKZ()), "Mbestm3small")
 runsim(-5 to 15, bp3, N, iters, () => new Mbest(3,N, 20*N, new HKZ()), "Mbestm3big")
 runcrb(-5 to 15, 3, N, "crbm3")
 runlsuclt(-5 to 15, 3, N, "lsucltm3")
 
 runsim(-5 to 35 by 2, bp3, N, iters, () => new ZW(3,N,N/3,N/3+1), "ZWm3big")
-runsim(-5 to 35 by 2, bp3, N, iters, () => new Babai(3,N, new HKZ()), "Babaim3bigrange")
+//runsim(-5 to 35 by 2, bp3, N, iters, () => new Babai(3,N, new HKZ()), "Babaim3bigrange")
 runsim(-5 to 35 by 2, bp3, N, iters, () => new Mbest(3,N, 20*N, new HKZ()), "Mbestm3bigrange")
 runcrb(-5 to 35, 3, N, "crbm3range")
 runlsuclt(-5 to 35, 3, N, "lsucltm3range")
 
 runsim(5 to 30, sp5, N, iters, () => new HAF(5,N), "HAFm5small")
 runsim(5 to 30, sp5, N, iters, () => new CPFHAF(5,N), "CPFHAFm5small")
-runsim(5 to 30, sp5, N, iters, () => new Babai(5,N, new LLL()), "Babaim5small")
[…]

### 2013-02-18 paper.tex
minor changes to paper.

@@
 The sphere decoder was observed to have excellent statistical performance but can only be computed efficiently for small $N$ less than about $50$.
-Computing the Babai point requires only $O(N^2)$ operations, but its statistical performance was observed to be comparatively poor at low signal to noise ratio (SNR).
+Babai's algorithm is computationally simple, but its statistical performance was observed to be comparatively poor at low signal to noise ratio (SNR).
 A major point of interest is that the lattices considered are not \emph{random}, and may admit fast nearest point algorithms.
 This has been studied in~\cite{McKilliam2010thesis} where polynomial time algorithms where found that compute the nearest point exactly.
-Although polynomial time, the algorithm are still computationally demanding in practice.
+Although polynomial time, the algorithms are still computationally demanding in practice.
 Nevertheless fast (exact or approximate) algorithms may still exist for these special lattices, that are yet to be discovered.
@@
 %Similar results were stated without proof in~\cite{McKilliam_polyphase_est_icassp_2011}.
-In the   BLERG `small noise' assumption.
-
-An interesting property is that the estimator of the $k$th polynomial phase coefficient converges to $\tilde{\mu}_k$ at rate $o(N^{-k})$.
-This is perhaps not surprising, since it is the same rate observed in polynomial regression.
-However, asserting that convergence at this rate occurs in the polynomial phase setting is not trivial.
-For this purpose we make use of an elementary result about the number of arithmetic progressions contained inside subsets of $\{1,2,\dots,N\}$~\cite{Gowers_new_proof2001}.
+The proof makes use of an elementary result about the number of arithmetic progressions contained inside subsets of $\{1,2,\dots,N\}$~\cite{Gowers_new_proof2001}.
 %This proof technique appears to be novel, and may be useful in other problems.
@@
 Potential candidates are the phase wrapped images observed in modern radar and medical imaging devices such as synthetic aperture radar and magnetic resonance imaging~\cite{Nico_phaseunwrappingSAR_2000,Friedlander_PD_phaseunwrapping_1996}.
-We also assert, under similarly mild conditions, that the LSU estimator is asymptotically normal.
+
+In the analysis of most existing polynomial phase estimators it is common to make what is known as a `small noise' assumption.  This name
+
+Along with strong consistency, we also assert, under similarly mild conditions, that the LSU estimator is asymptotically normal.
 %In the case where the noise is normally distributed the asymptotic variance of the LSU estimator is close to the Cramer-Rao bound at high signal to noise ratio (SNR) and is close to $\tfrac{2}{3}\pi$ times that Cramer-Rao bound at low (SNR).
@@
 
-\begin{figure}[t] 
+\begin{figure}[p] 
    	\centering 

### 2013-02-18 code/java/src/pubsim/poly/Mbest.java

@@ -3,13 +3,10 @@
 
 package pubsim.poly;
 
-import Jama.Matrix;
-import pubsim.lattices.Vnmstar.VnmStarGlued;
 import pubsim.lattices.reduction.LLL;
 import pubsim.lattices.reduction.LatticeReduction;
 
 /**
- *
  * @author Robby McKilliam
  */
 public class Mbest extends Babai {

### 2013-02-18 code/montecarlo.scala

@@ -19,7 +19,7 @@ import pubsim.lattices.reduction.None
 import pubsim.lattices.reduction.HKZ
 import pubsim.lattices.reduction.LLL
 
-val iters = 2000
+val iters = 100
 val N = 199
 def npow(x : Int, t : Int) = if(t<=0) 1.0 else scala.math.pow(x,t)
 val sp3 = (0 to 3).map( k => 0.25/pubsim.Util.factorial(k)/npow(N,k-1) ).toArray //3rd order paramater that work for the HAF and CPF
@@ -31,7 +31,7 @@ val sp7 = (0 to 7).map( k => 0.25/pubsim.Util.factorial(k)/npow(N,k-1) ).toArray
 
 val starttime = (new java.util.Date).getTime
 
-runsim(-5 to 15, sp3, N, iters, () => new HAF(3,N), "HAFm3small")
+/*runsim(-5 to 15, sp3, N, iters, () => new HAF(3,N), "HAFm3small")
 runsim(-5 to 15, bp3, N, iters, () => new HAF(3,N), "HAFm3big")
 runsim(-5 to 15, sp3, N, iters, () => new CPF(N), "CPFm3small")
 runsim(-5 to 15, bp3, N, iters, () => new CPF(N), "CPFm3big")
@@ -53,7 +53,7 @@ runsim(5 to 30, sp5, N, iters, () => new CPFHAF(5,N), "CPFHAFm5small")
 //runsim(5 to 30, sp5, N, iters, () => new Babai(5,N, new LLL()), "Babaim5small")
 runsim(5 to 30, sp5, N, iters, () => new Mbest(5,N, 20*N, new LLL()), "Mbestm5small")
 runcrb(5 to 30, 5, N, "crbm5")
-runlsuclt(5 to 30, 5, N, "lsucltm5")
+runlsuclt(5 to 30, 5, N, "lsucltm5")*/
 
 runsim(5 to 30, sp6, N, iters, () => new HAF(6,N), "HAFm6small")
 runsim(5 to 30, sp6, N, iters, () => new CPFHAF(6,N), "CPFHAFm6small")
@@ -62,12 +62,12 @@ runsim(5 to 30, sp6, N, iters, () => new Mbest(6,N, 20*N, new LLL()), "Mbestm6sm
 runcrb(5 to 30, 6, N, "crbm6")
 runlsuclt(5 to 30, 6, N, "lsucltm6")
 
[…]

### 2013-02-18 code/java/src/pubsim/poly/AbstractPolynomialPhaseEstimator.java
fixed bug in round

@@ -8,7 +8,7 @@ package pubsim.poly;
  */
 abstract public class AbstractPolynomialPhaseEstimator implements PolynomialPhaseEstimatorInterface {
 
-    final protected AmbiguityRemover ambiguityRemover;
+    final protected AmbiguityRemoverRectangular ambiguityRemover;
     final int m;
 
     private AbstractPolynomialPhaseEstimator() {

### 2013-02-18 paper.tex
Worked out that it's the error method that has numerical difficulties, probably need to implement that with BigRational

@@
 
-In the analysis of most existing polynomial phase estimators it is common to make what is known as a `small noise' assumption.  This name
+The analysis of most existing polynomial phase estimators make what is known as a `small errors' or `small noise' assumption~\cite{Peleg_linear_FM,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.
+This amounts to assuming that that estimator $\hat{\mu}_k$ of the $k$th coefficient is such that $N^{k}(\tilde{\mu}_k -  \hat{\mu}_k) \rightarrow 0$ in probability as $N \rightarrow \infty$.
+This assumption is used to justify the use of first and second order expansions~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats}, typically called a `perturbation analysis' in this literature~\cite{Peleg_linear_FM,Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004}, to obtain expression for the asymptotic mean square error of the estimator in question.
+The `small errors' assumption is usually justified by numerical Monte Carlo simulations, showing the asymptotic mean square error obtained by the perturbation analysis to be accurate provided that the signal to noise ratio (SNR) is larger than a `threshold' value (see the simulations in Section~\ref{sec:simulations} for examples).
+Typically one observes the threshold to move to smaller values of SNR as the $N$ increases.
+The problem with the `small errors' assumption is that it is not known if the threshold continues to move to smaller and smaller values of SNR as $N$ increase or if it infact converges to some finite (but perhaps small) value of SNR, the consequence being that the estimator will be innaccurate if the SNR is smaller than this finte value, regardless of the number of observations $N$ obtained.
+
+An analysis of the HAF estimator under the assumption that the noise $\{X_n\}$ is what and Gaussian and without the `small errors' assumption has been made in~\cite{Porat_asympt_HAF_DPT_1996}.
+There does not appear to have been an analysis of any phase unwrapping estimator without the small noise assumption.
+Infact, there appears to be a misconception that phase unwrapping estimators are only accurate with the noise is `small' in some sense.
+The misconceptions would appear to be motivated by early papers such as~\cite{Tretter1985,Kay1989} where it was observed that the noise observed in the phase of the signal only looks `Gaussian' when the complex noise producing it has small variance.
+The results
 

### 2013-02-18 code/montecarlo.scala

@@ -19,15 +19,18 @@ import pubsim.lattices.reduction.None
 import pubsim.lattices.reduction.HKZ
 import pubsim.lattices.reduction.LLL
 
-val iters = 100
+val iters = 20
 val N = 199
-def npow(x : Int, t : Int) = if(t<=0) 1.0 else scala.math.pow(x,t)
+def npow(x : Int, t : Int) : Double = if(t<=0) 1.0 else scala.math.pow(x,t)
+//def npow(x : Int, t : Int) = scala.math.pow(x,t)
 val sp3 = (0 to 3).map( k => 0.25/pubsim.Util.factorial(k)/npow(N,k-1) ).toArray //3rd order paramater that work for the HAF and CPF
 val bp3 = (0 to 3).map( k => 0.25/pubsim.Util.factorial(k) ).toArray //3rd order paramater that do not work for the HAF and CPF
-val sp5 = (0 to 5).map( k => 0.25/pubsim.Util.factorial(k)/npow(N,k-1) ).toArray //5rd order paramater that work for the HAF
-val bp5 = (0 to 5).map( k => 0.25/pubsim.Util.factorial(k) ).toArray //5rd order paramater that do not work for the HAF
-val sp6 = (0 to 6).map( k => 0.25/pubsim.Util.factorial(k)/npow(N,k-1) ).toArray //5rd order paramater that work for the HAF
-val sp7 = (0 to 7).map( k => 0.25/pubsim.Util.factorial(k)/npow(N,k-1) ).toArray //5rd order paramater that work for the HAF
+val sp5 = (0 to 5).map( k => 0.25/pubsim.Util.factorial(k)/npow(N,k-1) ).toArray //5th order paramater that work for the HAF
+val bp5 = (0 to 5).map( k => 0.25/pubsim.Util.factorial(k) ).toArray //5th order paramater that do not work for the HAF
+val sp6 = (0 to 6).map( k => 0.25/pubsim.Util.factorial(k)/npow(N,k-1) ).toArray //6th order paramater that work for the HAF
+val sp7 = (0 to 7).map( k => 0.25/pubsim.Util.factorial(k)/npow(N,k-1) ).toArray //7th order paramater that work for the HAF
+
+sp6.foreach(println)
 
 val starttime = (new java.util.Date).getTime
 

### 2013-02-18 paper.tex
added to intro

--- 2013-02-18 paper.tex
@@
 Polynomial phase signals arise in fields including radar, sonar, geophysics, biology, and radio communication~\cite{Hlawatsch_lin_quad_time_freq_spmag_1992,Ridleyspeechpolyphase1989, Suga_1975_bats_echolocation, Moss_2005echolocation,Angeby_estimating_2000}.
-In radar and sonar applications polynomial phase signals arise when acquiring radial velocity and acceleration (and higher order motion descriptors) of a target from a reflected signal, and also in continuous wave radar and low probability of intercept radar~\cite{Levanon_Radar_signals_2004}.
+%In radar and sonar applications polynomial phase signals arise when acquiring radial velocity and acceleration (and higher order motion descriptors) of a target from a reflected signal, and also in continuous wave radar and low probability of intercept radar~\cite{Levanon_Radar_signals_2004}.
 In biology, polynomial phase signals are used to describe the sounds emitted by bats and dolphins for echo location \citep{Suga_1975_bats_echolocation, Moss_2005echolocation}.
-
 A polynomial phase signal of order $m$ is a function of the form
@@
 \]
-where $j = \sqrt{-1}$, and $t$ is a real number, often representing time, and 
+where $j = \sqrt{-1}$, and $t$ is a real number and 
 \[
@@
 It was shown in \cite{McKilliam2009IndentifiabliltyAliasingPolyphase} that an identifiable region tessellates a particular $m+1$ dimensional lattice.
-We discuss this in Section \ref{sec:ident_aliasing}.
+We discuss this further in Section \ref{sec:ident_aliasing}.
 
@@
 A major point of interest is that the lattices considered are not \emph{random}, and may admit fast nearest point algorithms.
-This has been studied in~\cite{McKilliam2010thesis} where polynomial time algorithms where found that compute the nearest point exactly.
+This has been studied in~\cite{McKilliam2010thesis} where polynomial time algorithms were found that compute the nearest point exactly.
 Although polynomial time, the algorithms are still computationally demanding in practice.
-Nevertheless fast (exact or approximate) algorithms may still exist for these special lattices, that are yet to be discovered.
+Nevertheless fast (exact or approximate) algorithms may exist for these special lattices, that are yet to be discovered.
 In this paper we apply another general purpose approximate nearest point algorithm, called the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder} for approximating the LSU estimator.
@@
 This proof technique appears to be novel and may be useful for purposes other than polynomial phase estimation, and in particular other applications involving data that is `wrapped' in some sense.
-Potential candidates are the phase wrapped images observed in modern radar and medical imaging devices such as synthetic aperture radar and magnetic resonance imaging~\cite{Nico_phaseunwrappingSAR_2000,Friedlander_PD_phaseunwrapping_1996}.
-
-In the analysis of most existing polynomial phase estimators it is common to make what is known as a `small noise' assumption.  This name
-
-Along with strong consistency, we also assert, under similarly mild conditions, that the LSU estimator is asymptotically normal.
+Potential candidates are the phase wrapped images observed in modern radar and medical imaging devices such as synthetic aperture radar and magnetic resonance imaging~\cite{Katkovnik_phasela_2009, Nico_phaseunwrappingSAR_2000,Friedlander_PD_phaseunwrapping_1996}.
+
+The analysis of most existing polynomial phase estimators make what is known as a `small noise' assumption~\cite{Peleg_linear_FM,Oshea_cpf_2004,Djurovic_haf_cpf_2012,Djuric_phase_unwrap_chirp_1990}.
+This amounts to assuming that that estimator $\hat{\mu}_k$ of the $k$th coefficient is such that $N^{k}(\tilde{\mu}_k -  \hat{\mu}_k) \rightarrow 0$ in probability as $N \rightarrow \infty$.
+This assumption is used to justify the use of first and second order expansions~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats}, typically called a `perturbation analysis' in this literature~\cite{Peleg_linear_FM,Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004}, to obtain expressions for the asymptotic mean square error of the estimator.
+The `small noise' assumption is usually justified by Monte-Carlo simulations showing the asymptotic mean square error obtained by the perturbation analysis to be accurate provided that the signal to noise ratio (SNR) is larger than a `threshold' value (see the simulations in Section~\ref{sec:simulations} for examples).
+Typically one observes the threshold to occur at smaller values of SNR for larger $N$.
+A problem with the `small noise' assumption is that it is not know that the threshold continues to occur at increasingly smaller values of SNR as $N$ increases or if it infact converges to some finite (but perhaps small) value of SNR.
+The consequence of the later being that the estimator will be innaccurate if the SNR is smaller than this finite value, regardless of the number of observations $N$ obtained.
+
+An analysis of the HAF estimator under the assumption that the noise $\{X_n\}$ is white and Gaussian and without the `small noise' assumption has been made in~\cite{Porat_asympt_HAF_DPT_1996} (see inparticular the proof of Lemma~1 and the remark that follows).
+To date, the literature does not contain a corresponding analysis of any estimator based on phase unwrapping.
+This has apparently lead to the misconception that phase unwrapping estimators are only accurate with the noise is `small' in some sense.
+This misconception would appear to be motivated by early papers such as~\cite{Tretter1985,Kay1989} where it was observed that the noise occuring in the phase of the signal only looks `Gaussian' when the complex noise producing it has small variance.
+The results in this paper make it clear that phase unwrapping estimators can be accurate even at very low signal to noise ratios.
+The information contained in the amplitude is not essential for accurate estimation of polynomial phase signals, and can be discarded with only small loss.
+For example, in the case that $\{X_n\}$ is white and Gaussian, our results show that the asymptotic variance of the LSU estimator converges to the Cramer-Rao lower bound (CRB) as the SNR becomes large, and converges to approximately 3 times the CRB as the SNR approaches zero.
+Related results have been presented previously~\cite{Quinn2009_dasp_phase_only_information_loss}.
+It is worth noting that this factor of 3 times the CRB at low SNR is better than results obtained with estimators based on multi-linear transforms, such as the HAF and CPF, for which the ratio between the asymptotic variance and the CRB diverges as the SNR approaches zero~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.
+
+%Along with strong consistency, we also assert, under similarly mild conditions, that the LSU estimator is asymptotically normal.
 %In the case where the noise is normally distributed the asymptotic variance of the LSU estimator is close to the Cramer-Rao bound at high signal to noise ratio (SNR) and is close to $\tfrac{2}{3}\pi$ times that Cramer-Rao bound at low (SNR).
-Space restrictions make it impossible for us to include a proof of asymptotic normality in this paper.
+%Space restrictions make it impossible for us to include a proof of asymptotic normality in this paper.
+
+BLERG: Appears to be the first estimator for which theoretical results have been derived and which works over the entire identifiable region of parameters.
 
@@
 These identifiability results are required in order to properly understand the statistical properties of polynomial phase estimators.
-Section~\ref{sec:least-squar-unwr} describes the LSU estimator and states a theorem asserting the estimator to be strongly consistent under some assumptions on the noise $X_1,\dots,X_N$.
-The theorem is proved in Section~\ref{sec:strongconstproof}.
+Section~\ref{sec:least-squar-unwr} describes the LSU estimator and states a theorem asserting the estimator to be strongly consistent and asymptotically normally distributed under some mild assumptions on the noise $\{X_n\}$.
+The proof of strong consistency is given in Section~\ref{sec:strongconstproof}.
 Section \ref{sec:simulations} describes the results of Monte Carlo simulations that compare the performance of the LSU estimator with some existing estimators.

### 2013-02-19 paper.tex
added now new data.  Added to paper

@@
 We call this the \emph{least squares unwrapping} (LSU) estimator.
+The LSU estimator can be computed in a reasonable amount of time for data sets of moderate size using existing general purpose algorithms from algebraic number theory.
 Under mild conditions on the distribution of the noise we describe the asymptotic properties of this estimator, showing that it is strongly consistent and asymptotically normally distributed.
@@
 As a result the LSU estimator will be accurate, even for very noisy signals, provided that the number of observations is sufficiently large.
-The LSU estimator can be computed in a reasonable amount of time for data sets of moderate size using existing general purpose algorithms from algebraic number theory.
-Monte-carlo simualtions assert the excellent statistical performance of the LSU estimator when compared to existing state-of-the-art estimators.
-A key feature is that the LSU estimator functions accurately over a far wider range of parameters than most existing estimators.
+A key feature is that the LSU estimator functions accurately over a far wider range of parameters than many popular existing estimators.
+Monte-carlo simualtions assert the excellent statistical performance of the LSU estimator when compared to existing state-of-the-art estimators and support our theoretical results.
 \end{abstract}
@@
 \begin{equation}\label{eq:Y_nsamplednoisey}
-Y_n = \rho s_n + X_n,
+Y_n = \rho s_n + X_n \qquad n = 1, \dots, N,
 \end{equation}
@@
 When $m=0$ (phase estimation) or $m=1$ (frequency estimation) the LSE is an effective approach, being both computationally efficient and statistically accurate \cite{Quinn2009_dasp_phase_only_information_loss,Hannan1973,Quinn2001,McKilliam_mean_dir_est_sq_arc_length2010,McKilliam2010thesis}.
-When $m \geq 2$ the computational complexity of the LSE is large~\cite{McKilliam2010thesis,Abatzoglou_ml_chirp_1986}.
+When $m \geq 2$ the computational complexity of the LSE is large~\cite{Oshea_cpf_2004,McKilliam2010thesis,Abatzoglou_ml_chirp_1986}.
 For this reason many authors have considered alternative approaches to polynomial phase estimation.
@@
 In \cite{McKilliam2009asilomar_polyest_lattice}, two standard techniques were considered, the \emph{sphere decoder}~\cite{Pohst_sphere_decoder_1981,Viterbo_sphere_decoder_1999} that exactly computes a nearest lattice point, and \emph{Babai's nearest plane algorithm}~\cite{Babai1986} that only produces an approximation.
-The sphere decoder was observed to have excellent statistical performance but can only be computed efficiently for small $N$ less than about $50$.
+The sphere decoder was observed to have excellent statistical performance but can only be computed efficiently for small $N$ less than about $60$.
 Babai's algorithm is computationally simple, but its statistical performance was observed to be comparatively poor at low signal to noise ratio (SNR).
@@
 The analysis of most existing polynomial phase estimators make what is known as a `small noise' assumption~\cite{Peleg_linear_FM,Oshea_cpf_2004,Djurovic_haf_cpf_2012,Djuric_phase_unwrap_chirp_1990}.
-This amounts to assuming that that estimator $\hat{\mu}_k$ of the $k$th coefficient is such that $N^{k}(\tilde{\mu}_k -  \hat{\mu}_k) \rightarrow 0$ in probability as $N \rightarrow \infty$.
-This assumption is used to justify the use of first and second order expansions~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats}, typically called a `perturbation analysis' in this literature~\cite{Peleg_linear_FM,Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004}, to obtain expressions for the asymptotic mean square error of the estimator.
+This amounts to assuming that the estimator $\hat{\mu}_k$ of the $k$th coefficient is such that $N^{k}(\tilde{\mu}_k -  \hat{\mu}_k) \rightarrow 0$ (in some sense) as $N \rightarrow \infty$.
+This assumption is used to justify the use of first and second order expansions~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats} (typically called a `perturbation analysis' in this literature~\cite{Peleg_linear_FM,Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004}) to obtain expressions for the asymptotic mean square error of the estimator in question.
 The `small noise' assumption is usually justified by Monte-Carlo simulations showing the asymptotic mean square error obtained by the perturbation analysis to be accurate provided that the signal to noise ratio (SNR) is larger than a `threshold' value (see the simulations in Section~\ref{sec:simulations} for examples).
-Typically one observes the threshold to occur at smaller values of SNR for larger $N$.
-A problem with the `small noise' assumption is that it is not know that the threshold continues to occur at increasingly smaller values of SNR as $N$ increases or if it infact converges to some finite (but perhaps small) value of SNR.
-The consequence of the later being that the estimator will be innaccurate if the SNR is smaller than this finite value, regardless of the number of observations $N$ obtained.
+Typically one observes the threshold to occur at smaller values of SNR as $N$ increases.
+A problem with the `small noise' assumption is that it is not known that the threshold continues to occur at increasingly smaller values of SNR as $N$ increases, or if it instead converges to some finite (but perhaps small) value of SNR.
+The consequence of the later occurance being that the estimator will be innaccurate if the SNR is smaller than this finite value, regardless of the number of observations $N$ obtained.
+Our analysis of the LSU estimator requires no `small noise' assumpion and guarantees that no such finite value of SNR exists for the LSU estimator.
+As a result the LSU estimator will be accurate for signals with arbitrarily small SNR provided that the number of observations is sufficiently large.
 
@@
 To date, the literature does not contain a corresponding analysis of any estimator based on phase unwrapping.
-This has apparently lead to the misconception that phase unwrapping estimators are only accurate with the noise is `small' in some sense.
+This has apparently lead to the misconception that phase unwrapping estimators are only accurate when the noise is `small' in some sense.
 This misconception would appear to be motivated by early papers such as~\cite{Tretter1985,Kay1989} where it was observed that the noise occuring in the phase of the signal only looks `Gaussian' when the complex noise producing it has small variance.
-The results in this paper make it clear that phase unwrapping estimators can be accurate even at very low signal to noise ratios.
+The results in this paper make it clear that phase unwrapping estimators can be accurate at arbitrarily low SNR.
 The information contained in the amplitude is not essential for accurate estimation of polynomial phase signals, and can be discarded with only small loss.
@@
 Related results have been presented previously~\cite{Quinn2009_dasp_phase_only_information_loss}.
-It is worth noting that this factor of 3 times the CRB at low SNR is better than results obtained with estimators based on multi-linear transforms, such as the HAF and CPF, for which the ratio between the asymptotic variance and the CRB diverges as the SNR approaches zero~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.
-
-%Along with strong consistency, we also assert, under similarly mild conditions, that the LSU estimator is asymptotically normal.
-%In the case where the noise is normally distributed the asymptotic variance of the LSU estimator is close to the Cramer-Rao bound at high signal to noise ratio (SNR) and is close to $\tfrac{2}{3}\pi$ times that Cramer-Rao bound at low (SNR).
-%Space restrictions make it impossible for us to include a proof of asymptotic normality in this paper.
-
-BLERG: Appears to be the first estimator for which theoretical results have been derived and which works over the entire identifiable region of parameters.
+It is worth noting that this factor of 3 times the CRB at low SNR is better than popular estimators such as the HAF and CPF, for which the ratio between the asymptotic variance and the CRB diverges as the SNR approaches zero~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.
+
+A key property of the LSU estimator is that it functions correctly for polynomial phase parameters contained anywhere inside the identifiable region described in Section~\ref{sec:ident_aliasing} and~\cite{Xia_dynamic_range_pps_2001,McKilliam2009IndentifiabliltyAliasingPolyphase}.
+This is the maximum possible range.
+By constrast popular estimators such as the HAF and CPF operate correctly only for parameters in a range much smaller than the identifiable region.
+This property has been studied previously~\cite{Djurovic_aliasing_2012,McKilliam2009IndentifiabliltyAliasingPolyphase,Zhou_lag_diverity_PPT_1997,Xia_dynamic_range_pps_2001} and can be a critical limitation for some applications, for example, interference supression in the pressence of a jammer~\cite{Djukanovic:2011:PMN:1950994.1951252,Djurovic_aliasing_2012}.
+Approaches exist that extend the range of parameters suitable for the HAF~\cite{Zhou_lag_diverity_PPT_1997,Djurovic_aliasing_2012}, but the resulting estimators typically have poor statistical performance.
+For this reason, the LSU estimator might be the only presently existing estimator that is viable for certain applications, such as interference supression.
+The theoretical results given in Section~\ref{sec:least-squar-unwr} prove that the LSU estimator functions correctly over the entire identifiable region.
+To our knowledge this is the first time a result of this type has been obtained for any polynomial phase estimator.
 
@@
 In Section~\ref{sec:ident_aliasing} we describe the identifiable region that was also derived in~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}.
-These identifiability results are required in order to properly understand the statistical properties of polynomial phase estimators.
+These identifiability results are required in order to properly understand the statistical properties of the LSU estimator.
 Section~\ref{sec:least-squar-unwr} describes the LSU estimator and states a theorem asserting the estimator to be strongly consistent and asymptotically normally distributed under some mild assumptions on the noise $\{X_n\}$.
 The proof of strong consistency is given in Section~\ref{sec:strongconstproof}.
-Section \ref{sec:simulations} describes the results of Monte Carlo simulations that compare the performance of the LSU estimator with some existing estimators.
+Space restrictions make it impossible to give a proof of asymptotic normality in this paper.
+A complete proof of normality is available in~\cite{McKilliam_arxiv_pps_unwrapping_2012}.
+Section~\ref{sec:simulations} describes the results of Monte Carlo simulations that compare the performance of the LSU estimator with existing estimators including the HAF and CPF.
 These simulations agree with the derived asymptotic properties.
+BLERG: benchmarks.
 
@@
 \end{figure}
+
+A fundamental problem in lattice theory is the \emph{nearest lattice point problem}~\cite{Agrell2002}.
+The nearest lattice point problem is: Given $\ybf\in\reals^n$ and some lattice $\Lambda \subset \reals^n$, find the lattice point $\xbf \in \Lambda$ such that the Euclidean distance between $\ybf$ and $\xbf$ is minimised, that is, solve
+\[
+\arg\min_{\xbf \in \Lambda} \|\ybf - \xbf\|^2.
+\]
+The nearest lattice point problem is known to be NP-hard under certain conditions when the lattice itself, or rather a basis thereof, is considered as an additional input parameter \cite{micciancio_hardness_2001, AjtaiShortestVecProbNPHard1998}.
+Nevertheless, algorithms exist that can compute the nearest lattice point in reasonable time if the dimension is small \cite{Agrell2002, Viterbo_sphere_decoder_1999, Pohst_sphere_decoder_1981}.
+One such algorithm introduced by Pohst \cite{Pohst_sphere_decoder_1981} in 1981 was popularised in signal processing and communications fields by Viterbo and Boutros \cite{Viterbo_sphere_decoder_1999} and has since been called the \emph{sphere decoder}.
+With modern computers, the sphere decoder is computationally viable if the dimension of the lattice is less than approximately $60$.
+Approximate algorithms for computing the nearest point have also been studied.
+One example is Babai's nearest plane algorithm \cite{Babai1986} which has worst case polynomial complexity.
+Yet another approximate algorithm is the $K$-best algorithm studied in~\cite{Zhan2006_K_best_sphere_decoder} that operates similarly to the sequential $M$-algorithm~\citep{Anderson1984_seq_coding_alg} used in coding theory.
+The $K$-best algorithm is more expensive that Babai's algorithm, but typically produces a better approximation of the nearest lattice point.
+In this paper we make use of the $K$-best algorithm for approximating the LSU estimator.
+We will find that is produces excellent estimates of polynomial phase coefficients and can be computed in a reasonable amount.
+We will not detail the workings of this algorithm as an excellent description has been given in~\cite{Zhan2006_K_best_sphere_decoder}.
 
@@
 
-It may be helpful to observe Figures~\ref{circstatplot_line} to~\ref{fig:circstatplot_cube}.
+It may be helpful to observe Figures~\ref{fig:circstatplot_line} and~\ref{fig:circstatplot_cube}.
 In these, the phase (divided by $2\pi$) of two distinct polynomial phase signals is plotted on the left, and on the right the principal component of the phase (also divided by $2\pi$) is plotted.
@@
 
-The comparitively small set of coefficients suitable for the HAF and CPF estimators has been studied previously~\cite{Angeby_PPS_aliasing_2000,Djurovic_aliasing_2012,McKilliam2009IndentifiabliltyAliasingPolyphase,Zhou_lag_diverity_PPT_1997,Xia_dynamic_range_pps_2001} and can be a significant limitation for some applications, for example, interference supression in the pressence of a jammer~\cite{Djukanovic:2011:PMN:1950994.1951252,Djurovic_aliasing_2012}.
+The comparitively small set of coefficients suitable for the HAF and CPF estimators has been studied previously~\cite{Angeby_PPS_aliasing_2000,Djurovic_aliasing_2012,McKilliam2009IndentifiabliltyAliasingPolyphase,Zhou_lag_diverity_PPT_1997,Xia_dynamic_range_pps_2001} and can be a critical limitation for some applications, for example, interference supression in the pressence of a jammer~\cite{Djukanovic:2011:PMN:1950994.1951252,Djurovic_aliasing_2012}.
 Zhou and Wang~\cite{Zhou_lag_diverity_PPT_1997} describe a method based on the extended Euclidean algorithm that enables the HAF estimator to operate on the entire identifiable region $B$.
@@
 
-\begin{figure}[p] 
+\begin{figure}[t] 
    	\centering 
@@
 
-\begin{figure}[p] 
-   	\centering 
-  		\includegraphics{code/data/plot567-3.mps} 
-  		\caption{BLERG} 
-  		\label{plot:polyest} 
- \end{figure}
+% \begin{figure}[p] 
+%    	\centering 
+%   		\includegraphics{code/data/plot567-3.mps} 
+%   		\caption{BLERG} 
+%   		\label{plot:polyest} 
+%  \end{figure}
 

### 2013-02-19 code/montecarlo.scala

@@ -19,22 +19,19 @@ import pubsim.lattices.reduction.None
 import pubsim.lattices.reduction.HKZ
 import pubsim.lattices.reduction.LLL
 
-val iters = 20
+val iters = 2000
 val N = 199
-def npow(x : Int, t : Int) : Double = if(t<=0) 1.0 else scala.math.pow(x,t)
-//def npow(x : Int, t : Int) = scala.math.pow(x,t)
+//def npow(x : Int, t : Int) : Double = if(t<=0) 1.0 else scala.math.pow(x,t)
+def npow(x : Int, t : Int) = scala.math.pow(x,t)
 val sp3 = (0 to 3).map( k => 0.25/pubsim.Util.factorial(k)/npow(N,k-1) ).toArray //3rd order paramater that work for the HAF and CPF
 val bp3 = (0 to 3).map( k => 0.25/pubsim.Util.factorial(k) ).toArray //3rd order paramater that do not work for the HAF and CPF
 val sp5 = (0 to 5).map( k => 0.25/pubsim.Util.factorial(k)/npow(N,k-1) ).toArray //5th order paramater that work for the HAF
 val bp5 = (0 to 5).map( k => 0.25/pubsim.Util.factorial(k) ).toArray //5th order paramater that do not work for the HAF
 val sp6 = (0 to 6).map( k => 0.25/pubsim.Util.factorial(k)/npow(N,k-1) ).toArray //6th order paramater that work for the HAF
-val sp7 = (0 to 7).map( k => 0.25/pubsim.Util.factorial(k)/npow(N,k-1) ).toArray //7th order paramater that work for the HAF
-
-sp6.foreach(println)
 
 val starttime = (new java.util.Date).getTime
 
-/*runsim(-5 to 15, sp3, N, iters, () => new HAF(3,N), "HAFm3small")
+runsim(-5 to 15, sp3, N, iters, () => new HAF(3,N), "HAFm3small")
 runsim(-5 to 15, bp3, N, iters, () => new HAF(3,N), "HAFm3big")
 runsim(-5 to 15, sp3, N, iters, () => new CPF(N), "CPFm3small")
 runsim(-5 to 15, bp3, N, iters, () => new CPF(N), "CPFm3big")
[…]

### 2013-02-21 paper.tex
added minor

@@
 This result is well known~\cite[Chapter IX, Theorem IV]{Cassels_geom_numbers_1997}~\cite[Proposition 2.1]{McKilliam2010thesis}.
-This result is for lattices with full rank.
+%This result is for lattices with full rank.
 A result in the general case can be obtained similarly, but is not required here.
@@
 A fundamental problem in lattice theory is the \emph{nearest lattice point problem}~\cite{Agrell2002}.
-The nearest lattice point problem is: Given $\ybf\in\reals^n$ and some lattice $\Lambda \subset \reals^n$, find the lattice point $\xbf \in \Lambda$ such that the Euclidean distance between $\ybf$ and $\xbf$ is minimised, that is, solve
-\[
-\arg\min_{\xbf \in \Lambda} \|\ybf - \xbf\|^2.
-\]
-The nearest lattice point problem is known to be NP-hard under certain conditions when the lattice itself, or rather a basis thereof, is considered as an additional input parameter \cite{micciancio_hardness_2001, AjtaiShortestVecProbNPHard1998}.
-Nevertheless, algorithms exist that can compute the nearest lattice point in reasonable time if the dimension is small \cite{Agrell2002, Viterbo_sphere_decoder_1999, Pohst_sphere_decoder_1981}.
-One such algorithm introduced by Pohst \cite{Pohst_sphere_decoder_1981} in 1981 was popularised in signal processing and communications fields by Viterbo and Boutros \cite{Viterbo_sphere_decoder_1999} and has since been called the \emph{sphere decoder}.
+The nearest lattice point problem is: Given $\ybf\in\reals^n$ and some lattice $\Lambda \subset \reals^n$, find a lattice point $\xbf \in \Lambda$ such that the Euclidean distance between $\ybf$ and $\xbf$ is minimised, that is, solve $\arg\min_{\xbf \in \Lambda} \|\ybf - \xbf\|^2.$  The nearest lattice point problem is known to be NP-hard under certain conditions when the lattice itself is considered as an additional input parameter \cite{micciancio_hardness_2001, AjtaiShortestVecProbNPHard1998}.
+On the other hand,  fast algorithms that compute a nearest point often exist for specific lattices with known structure~\cite{Conway1982FastQuantDec,McKilliam2009CoxeterLattices,McKilliam2008}.
+Even in the general case, where the lattice has no known structure, algorithms exist that can compute a nearest lattice point in reasonable time if the dimension is small \cite{Agrell2002, Viterbo_sphere_decoder_1999, Pohst_sphere_decoder_1981}.
+One such algorithm is the \emph{sphere decoder}~\cite{Pohst_sphere_decoder_1981,Viterbo_sphere_decoder_1999}.
 With modern computers, the sphere decoder is computationally viable if the dimension of the lattice is less than approximately $60$.
-Approximate algorithms for computing the nearest point have also been studied.
-One example is Babai's nearest plane algorithm \cite{Babai1986} which has worst case polynomial complexity.
+Algorithms that approximate a nearest point have also been studied.
+One example is Babai's nearest plane algorithm \cite{Babai1986}.
 Yet another approximate algorithm is the $K$-best algorithm studied in~\cite{Zhan2006_K_best_sphere_decoder} that operates similarly to the sequential $M$-algorithm~\citep{Anderson1984_seq_coding_alg} used in coding theory.
-The $K$-best algorithm is more expensive that Babai's algorithm, but typically produces a better approximation of the nearest lattice point.
+The $K$-best algorithm is more expensive that Babai's algorithm, but typically produces a better approximation.
 In this paper we make use of the $K$-best algorithm for approximating the LSU estimator.
-We will find that is produces excellent estimates of polynomial phase coefficients and can be computed in a reasonable amount.
-We will not detail the workings of this algorithm as an excellent description has been given in~\cite{Zhan2006_K_best_sphere_decoder}.
+We will find that it produces excellent estimates of polynomial phase coefficients and can be computed in a reasonable amount of time for moderate numbers of observations ($N$ less that about 1000).
+We will not detail the workings of the $K$-best algorithm in this paper.
+A thorough description is given in~\cite{Zhan2006_K_best_sphere_decoder}.
 
@@
 \]
-where superscript $^\prime$ indicates the transpose, denote the column vector of length $m+1$ containing the coefficients of $g$.
+denote the column vector of length $m+1$ containing the coefficients of $g$.
+Superscript $^\prime$ indicates the transpose.
 If $y$ and $z$ differ by a polynomial from $\mathcal{Z}$ then we can write $y = z + p$ where $p \in \mathcal{Z}$ and then also $\coef(y) = \coef(z) + \coef(p)$.
@@
 This corresponds with the \term{Nyquist criterion}.
-The lattice $L_2$ is equal to $\ints^2$ so the box $[-\nicefrac{1}{2}, \nicefrac{1}{2})^2$ corresponds with the Voronoi cell of $L_2$.
+The lattice $L_2$ is equal to $\ints^2$ and the box $[-\nicefrac{1}{2}, \nicefrac{1}{2})^2$ corresponds with the Voronoi cell of $L_2$.
 When $m > 1$ the identifiable region becomes more complicated and $L_{m+1} \neq \ints^{m+1}$.
@@
 In general there are infinitely many choices for the identifiable region.
-A natural choice is the Voronoi cell of $L_{m+1}$ used in~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}.
+A natural choice is the Voronoi cell of $L_{m+1}$ as used in~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}.
 Another potential choice is a fundamental parallelepiped of $L_{m+1}$.
@@
 %The effect of $\dealias(\xbf)$ is to \emph{dealias} the polynomial phase coefficients.
-When $m = 0$ or $1$ then $\dealias(\xbf) = \fracpart{\xbf}$ where $\fracpart{\xbf} = \xbf - \round{\xbf}$ denotes the (centered) fractional part and $\round{\xbf}$ denotes the nearest integer to $\xbf$ with half integers rounded upwards and both $\fracpart{\cdot}$ and $\round{\cdot}$ operate on vectors elementwise.
+When $m = 0$ or $1$ then $\dealias(\xbf) = \fracpart{\xbf}$ where $\fracpart{\xbf} = \xbf - \round{\xbf}$ denotes the (centered) fractional part and $\round{\xbf}$ denotes the nearest integer to $\xbf$ and both $\fracpart{\cdot}$ and $\round{\cdot}$ operate on vectors elementwise.
+The direction of rounding for half-integers is not important so long as it is consistent.
+We have chosen to round up half-integers here.
 For $m \geq 2$ the function $\dealias(\xbf)$ can be computed by a simple sequential algorithm \cite[Sec.
@@
 Recall that we desire to estimate the coefficients $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ from the noisy samples $Y_1, \dots, Y_N$ given in~\eqref{eq:Y_nsamplednoisey}.
-We take the complex argument of the $Y_n$ and divide by $2\pi$ to obtain
+Take the complex argument of the $Y_n$ and divide by $2\pi$ to obtain
 \begin{equation}\label{eq:noise_circ_poly}
@@
 \end{equation}
-where $\angle{z}$ denotes the complex argument of the complex number $z$, and 
+where $\angle{z}$ denotes the complex argument of $z$, and 
 \[
@@
 
-It is shown in~\cite{McKilliam2009asilomar_polyest_lattice} and \cite[Sec~8.1]{McKilliam2010thesis} how this minimisation problem can be posed as a nearest lattice point problem, and we briefly discuss this here.
-Write $SS$ as
+This minimisation problem can be posed as a nearest lattice point problem~\cite{McKilliam2009asilomar_polyest_lattice, McKilliam2010thesis}.  Write $SS$ as
 \[
@@
 \end{equation}
-and where the column vectors $\wbf = [W_1, \dots, W_N]^\prime$ and $\thetabf = [\Theta_1, \dots, \Theta_N]^\prime$.   For fixed $\wbf$ the minimiser of $SS(\mubf, \wbf)$ with respect to $\mubf$ is
+and the column vectors $\wbf = [W_1, \dots, W_N]^\prime$ and $\thetabf = [\Theta_1, \dots, \Theta_N]^\prime$.   For fixed $\wbf$ the minimiser of $SS(\mubf, \wbf)$ with respect to $\mubf$ is
 \begin{equation} \label{eq:hat_theta_cal}
@@
 where $\Xbf^+ = (\Xbf^\prime\Xbf)^{-1}\Xbf^\prime$ is the Moore-Penrose pseudoinverse of $\Xbf$.
-
 We may thus compute $\widehat{\mubf}$ as $\widehat{\mubf} = \mu(\widehat{\wbf})$, where $\widehat{\wbf}$ is the minimiser of
@@
 
-The \emph{intrinsic mean} or \emph{Fr\'{e}chet mean} of $\Phi_n$ is defined as~\cite{McKilliam_mean_dir_est_sq_arc_length2010,bwhk07a,Bhattacharya_int_ext_means_2003,Bhattacharya_int_ext_means_2005},
+The \emph{intrinsic mean} (or \emph{Fr\'{e}chet mean}) of $\Phi_n$ is defined as~\cite{McKilliam_mean_dir_est_sq_arc_length2010,bwhk07a,Bhattacharya_int_ext_means_2003,Bhattacharya_int_ext_means_2005},
 \begin{equation}\label{eq:intrmeandefn}
@@
 \]
-where $\Cbf$ is the $(m+1)\times (m+1)$ Hilbert matrix with elements $C_{ik} = 1/(i + k + 1)$ for $i,k = \{0, 1, \dots, m\}$.
+where $\Cbf$ is the $(m+1)\times (m+1)$ Hilbert matrix with elements $C_{ik} = 1/(i + k + 1)$ for $i,k \in \{0, 1, \dots, m\}$.
 \end{enumerate}
 \end{theorem}
-The proof of Theorem~\ref{thm:asymp_proof} is contained within the next two sections.
-Section~\ref{sec:strongconstproof} proves strong consistency and Section~\ref{sec:centlimitproof} proves asymptotic normality.
-The proofs for the case of $m=0$ were given in~\cite{McKilliam_mean_dir_est_sq_arc_length2010} and for the case when $m=1$ were given in~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.
-The proofs here take a similar approach, but require new techniques.
-Before giving the proof we make the following remarks.
+
+A proof of the first part of this theorem (strong consistency) is given in Section~\ref{sec:strongconstproof}.
+Space restrictions make it impossible to give a proof of asymptotic normality in this paper.
+A complete proof of normality is available in~\cite{McKilliam_arxiv_pps_unwrapping_2012}.
+Proofs for the case of $m=0$ were given in~\cite{McKilliam_mean_dir_est_sq_arc_length2010} and for the case when $m=1$ were given in~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.
+The proof here takes a similar approach, but requires new techniques.
+Before giving the proof of strong consistency we make the following remarks.
 

### 2013-02-22 paper.tex
added some data.  Changes plots.

@@
 %Similar results were stated without proof in~\cite{McKilliam_polyphase_est_icassp_2011}.
-The proof makes use of an elementary result about the number of arithmetic progressions contained inside subsets of $\{1,2,\dots,N\}$~\cite{Gowers_new_proof2001}.
+The proof makes use of an elementary result about the number of arithmetic progressions contained inside subsets of $\{1,2,\dots,N\}$~\cite{Erdos_on_some_sequence_of_integers1936,Szemeredi_setint_no_k_arth1975,Gowers_new_proof2001}.
 %This proof technique appears to be novel, and may be useful in other problems.
@@
 In the interest of simplicity we only consider the case when $\Phi_1, \dots, \Phi_N$ are identically distributed here.
-If $X_n$ is circularly symmetric with density function nonincreasing with magnitude $\abs{X_n}$, then, the corresponding $\Phi_n$ necessarily has zero intrinsic mean~\cite[Theorem 5.2, page 78]{McKilliam2010thesis}.
+If $X_n$ is circularly symmetric with density function nonincreasing with magnitude $\abs{X_n}$, then, the corresponding $\Phi_n$ necessarily has zero intrinsic mean~\cite[Theorem 5.2]{McKilliam2010thesis}.
 Thus, our theorem covers commonly used distributions for $X_1, \dots, X_N$, such as the normal distribution.
+No assumptions about the magnitude of $X_1, \dots, X_N$ are required, that is, we do not require the noise to be `small' in any sense.
+For example, if $X_1, \dots, X_N$ are normally distributed with variance $\sigma_c^2$, then the theorem holds regardless of how large $\sigma_c^2$ is, and as a result the LSU estimator will be accurate provided $N$ is sufficiently large.
 
@@
 This type of resul is called a~\emph{uniform law of large numbers}~\cite{Pollard_conv_stat_proc_1984,van2009empirical}.
-A full proof of~\eqref{eq:SNVNunifmlln} in given in BLERG (see arxiv).
+A full proof of~\eqref{eq:SNVNunifmlln} in given in~\cite{McKilliam_mean_dir_est_sq_arc_length2010}.
 We now concentrate attention on the minimiser of $V_N$.
@@
 This section describes the results of Monte-Carlo simulations with the least squares unwrapping (LSU) estimator, the high order ambiguity function (HAF) estimator~\cite{Peleg_DPT_1995}, the cubic phase function (CPF) estimator~\cite{Oshea_cpf_2004} and the estimator that results from combining the HAF and CPF~\cite{Djurovic_haf_cpf_2012}.
-The CPF estimator only applies to polynomial phase signals of order $m=3$.
-In all simulations the unknown amplitude is $\rho = 1$ and the number of observations is $N = 199$ (an odd number is chosen to satisfy the requirements of the CPF estimator).
-The $X_1, \dots, X_N$ are pseudorandomly generated independent and identically distributed circularly symmetric complex Gaussian random variables with variance $\expect\abs{X_1}^2 = \sigma_c^2$.
+The CPF estimator only applies to polynomial phase signals of order $m=3$ and when the number of observations $N$ is odd.
+In all our simulations the unknown amplitude is $\rho = 1$ and $N = 199$.
+The $X_1, \dots, X_N$ are pseudorandomly generated independent and identically distributed circularly symmetric complex Gaussian with variance $\expect\abs{X_1}^2 = \sigma_c^2$.
 The corresponding signal to noise ratio (SNR) is $\rho^2 \sigma_c^{-2} = \sigma_c^{-2}$.
-The number of replications of each experiment is $T = 2000$ to obtain estimates $\widehat{\mubf}_1, \dots, \widehat{\mubf}_T$ and the corresponding dealiased errors $\widehat{\lambdabf}_t = \dealias(\widehat{\mubf}_T - \tilde{\mubf})$ are computed.
+The number of replications of each Monte-Carlo experiment is $T = 2000$ to obtain estimates $\widehat{\mubf}_1, \dots, \widehat{\mubf}_T$ and the corresponding dealiased errors $\widehat{\lambdabf}_t = \dealias(\widehat{\mubf}_T - \tilde{\mubf})$ are computed.
 The sample mean square error (MSE) of the $k$th coefficient is computed according to $\tfrac{1}{T}\sum_{t=1}^T \widehat{\lambda}_{k,t}^2$ where $\widehat{\lambda}_{k,t}$ is the $k$th element of $\widehat{\lambdabf}_t$.
-In all simulation the LSU estimator is approximately computed using the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder}.
-
-Figure~\ref{plot:m3small} show the sample MSE of the highest order coefficient $\mu_3$ for a polynomial phase signal of order $m=3$.
-All the estimators perform poorly at low SNR until a `threshold' is reached, after which the variance is close the Cramer-Rao bound (CRB)~\cite{Peleg1991_CRB_PPS_1991} indicated by the solid line.
+In all simulations the LSU estimator is approximated using the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder}.
+
+Figure~\ref{plot:m3small} shows the sample MSE of the highest order coefficient $\mu_3$ for a polynomial phase signal of order $m=3$.
+All the estimators perform poorly at low SNR until a `threshold' is reached, after which the variance is close to the Cramer-Rao bound (CRB)~\cite{Peleg1991_CRB_PPS_1991} indicated by the solid line.
 The asymptotic variance of the LSU estimator predicted by Theorem~\ref{thm:asymp_proof} is displayed by the dashed line.
 Provided the noise variance is small enough (so that the `threshold' is avoided) the sample MSE of the LSU estimator is close to that predicted by Theorem~\ref{thm:asymp_proof}.
-In Figure~\ref{plot:m3small} the CPF estimator appears to have the best performance at low SNR, reaching a threshold at \unit[1]{dB}.
-The LSU estimator reaches a threshold at \unit[3]{dB} and the HAF estimator at \unit[5]{dB}.
+Using Theorem~\ref{thm:asymp_proof} we predict the ratio between the CRB and the asymptotic variance of the LSU estimator to approach 1 as SNR approaches $\infty$ and approximately $3$ as SNR approaches zero~\cite{Quinn2009_dasp_phase_only_information_loss}.
+Contrastingly, the ratio between the CRB and asymptotic variance of the HAF and CPF estimators is known to diverge as SNR goes to zero~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.
+%For this reason we can expect the LSU estimator to be more accurate at low SNR for sufficiently large $N$.
+
+In Figure~\ref{plot:m3small} the CPF estimator reaches a threshold at \unit[1]{dB}, the LSU estimator at \unit[3]{dB} and the HAF estimator at \unit[5]{dB}.
 In Figure~\ref{plot:m3small} the true coefficients are $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}  \right]$.
@@
 
+Figures~\ref{plot:m5} displays the sample MSE of the highest order coefficient $\mu_5$  for a polynomial phase signal of order $m=5$.
+In this case the LSU estimator reaches a threshold at \unit[12]{dB}, the HAF estimator at \unit[18]{dB}, and the estimator that combines the CPF and HAF at \unit[10]{dB}.
+In this figure the true coefficients are $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}, \tfrac{1}{96N^3}, \tfrac{1}{480N^4}  \right]$ and these are chosen specifically to satisfy the requirement of the HAF and CPF-HAF estimators.
+As in the case where $m=3$, the region of coefficients suitable for the HAF and CPF-HAF estimators is much smaller that the identifiable region $B$.
+The effect becomes more servere as $m$ increases.
+When $m=5$ and $N=199$ the space of coefficients searched by the HAF and CPF-HAF estimators is approximately $10^{16}$ times smaller than the identifiable region.
+This constitutes is a dramatic reduction in the range of detectable polynomial phase signals.
+Simulations have been run with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}, \tfrac{1}{96}, \tfrac{1}{480}  \right]$, and in this case we obtain results resembling those in Figure~\ref{plot:m3big}.
+The performance of the LSU estimator remains unchanged, but the HAF and CPF-HAF estimators fail completely.
+
 \begin{figure}[p] 
@@
   		\includegraphics{code/data/plot3-1.mps} 
-  		\caption{BLERG} 
+  		\caption{Polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}  \right]$.} 
   		\label{plot:m3small} 
@@
   		\includegraphics{code/data/plot3-2.mps} 
-  		\caption{BLERG} 
+  		\caption{Polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$.} 
   		\label{plot:m3big} 
@@
   		\includegraphics{code/data/plot3-3.mps} 
-  		\caption{BLERG} 
+  		\caption{Polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$.} 
   		\label{plot:m3bigzw} 
@@
    	\centering 
-  		\includegraphics{code/data/plot567-1.mps} 
-  		\caption{BLERG} 
-  		\label{plot:polyest} 
+  		\includegraphics{code/data/plot5-1.mps} 
+  		\caption{Polynomial phase signal of order 5 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}, \tfrac{1}{96N^3}, \tfrac{1}{480N^4}   \right]$.} 
+  		\label{plot:m5} 
 \end{figure}
@@
    	\centering 
-  		\includegraphics{code/data/plot567-2.mps} 
-  		\caption{BLERG} 
-  		\label{plot:polyest} 
+  		\includegraphics{code/data/plot5-2.mps} 
+  		\caption{Polynomial phase signal of order 5 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}, \tfrac{1}{96}, \tfrac{1}{480}   \right]$.} 
+  		\label{plot:m5} 
 \end{figure}
+
+% \begin{figure}[t] 
+%    	\centering 
+%   		\includegraphics{code/data/plot567-2.mps} 
+%   		\caption{BLERG} 
+%   		\label{plot:m6} 
+%  \end{figure}
 
@@
 
-\section{Computational considerations}
-
-The exact LSU estimator implemented using the sphere decoder can be computed quickly for any $m$ with $N$ approximately less than $50$.
+% \section{Computational considerations}
+
+% The exact LSU estimator implemented using the sphere decoder can be computed quickly for any $m$ with $N$ approximately less than $50$.
 In situations where $N$ is small, but high statistical accuracy is required, the sphere decoder is computationally a better choice than the LSE for polynomial phase signals of order greater than or equal to $2$.
 
-Babai's nearest plane algorithm requires $O(N^2)$ operations and can be run in a reasonable amount of time for any $m$ and reasonably large $N$.
+% Babai's nearest plane algorithm requires $O(N^2)$ operations and can be run in a reasonable amount of time for any $m$ and reasonably large $N$.
 It should be noted that Babai's nearest plane algorithm requires a Lov\`as reduced lattice basis~\cite{Lenstra1982}.

### 2013-02-22 build.sh

@@ -9,7 +9,7 @@ cd ..
 
 cd code/data 
 mpost -interaction=nonstopmode plot3.mp 
-mpost -interaction=nonstopmode plot567.mp 
+mpost -interaction=nonstopmode plot5.mp 
 cd ../../ 
 
 pdflatex -shell-escape paper.tex 

### 2013-02-22 code/bench.scala

@@ -2,37 +2,48 @@
  * Run simulations for the sparse noisy period estimation problem.
  * Various estimators are available.
  */
-
-import pubsim.poly.MbestEstimator
-import pubsim.poly.BabaiEstimator
 import pubsim.poly.PolynomialPhaseSignal
+import pubsim.poly.HAF
+import pubsim.poly.Babai
+import pubsim.poly.Mbest
+import pubsim.poly.MaximumLikelihood
+import pubsim.poly.SphereDecoder
+import pubsim.poly.PolynomialPhaseEstimatorInterface
+import pubsim.poly.CPF
+import pubsim.poly.CPFHAF
+import pubsim.poly.ZW
+import pubsim.poly.bounds.GaussianCRB
 import pubsim.distributions.GaussianNoise
 import pubsim.lattices.reduction.None
 import pubsim.lattices.reduction.HKZ
 import pubsim.lattices.reduction.LLL
 
-val m = 5;
-//val Ns = List(30, 60, 100, 200, 350, 500, 1000);
-val Ns = List(199);
-
 val MIN_BENCH_DURATION : Long = 40000000000L; // (20 secs)  
+val Ns = List(30, 60, 100, 200, 350, 500);
 
-val snrdB = 7
+def npow(x : Int, t : Int) : Double = if(t<=0) 1.0 else scala.math.pow(x,t)
+val p3 = (0 to 3).map( k => 0.25/pubsim.Util.factorial(k) ).toArray //3rd order paramaters
+val p5 = (0 to 5).map( k => 0.25/pubsim.Util.factorial(k) ).toArray //5th order paramaters
+
+runbench(Ns, sp3, N, iters, () => new HAF(3,N), "benchHAF3")
+runbench(Ns, sp3, N, iters, () => new Mbest(3,N), "benchMbest3")
+runbench(Ns, sp5, N, iters, () => new HAF(5,N), "benchHAF5")
+runbench(Ns, sp5, N, iters, () => new Mbest(5,N), "benchMbest5")
+
+def runbench(N : Seq[Int], params : Array[Double], snrdB : Double, benchtime : Double, estf : () => PolynomialPhaseEstimatorInterface, name : String) {
+
 val snr = scala.math.pow(10.0, snrdB/10.0)
 val noise =  new GaussianNoise(0,1.0/snr/2.0) //variance for real and imaginary parts (divide by 2)
 
-for(N <- Ns){
+val timelist = Ns.map{ N =>
 
[…]

### 2013-02-22 code/montecarlo.scala

@@ -21,15 +21,15 @@ import pubsim.lattices.reduction.LLL
 
 val iters = 2000
 val N = 199
-//def npow(x : Int, t : Int) : Double = if(t<=0) 1.0 else scala.math.pow(x,t)
-def npow(x : Int, t : Int) = scala.math.pow(x,t)
+def npow(x : Int, t : Int) : Double = if(t<=0) 1.0 else scala.math.pow(x,t)
+//def npow(x : Int, t : Int) = scala.math.pow(x,t)
 val sp3 = (0 to 3).map( k => 0.25/pubsim.Util.factorial(k)/npow(N,k-1) ).toArray //3rd order paramater that work for the HAF and CPF
 val bp3 = (0 to 3).map( k => 0.25/pubsim.Util.factorial(k) ).toArray //3rd order paramater that do not work for the HAF and CPF
 val sp5 = (0 to 5).map( k => 0.25/pubsim.Util.factorial(k)/npow(N,k-1) ).toArray //5th order paramater that work for the HAF
 val bp5 = (0 to 5).map( k => 0.25/pubsim.Util.factorial(k) ).toArray //5th order paramater that do not work for the HAF
-val sp6 = (0 to 6).map( k => 0.25/pubsim.Util.factorial(k)/npow(N,k-1) ).toArray //6th order paramater that work for the HAF
 
 val starttime = (new java.util.Date).getTime
+
 
 runsim(-5 to 15, sp3, N, iters, () => new HAF(3,N), "HAFm3small")
 runsim(-5 to 15, bp3, N, iters, () => new HAF(3,N), "HAFm3big")
@@ -48,19 +48,18 @@ runsim(-5 to 35 by 2, bp3, N, iters, () => new Mbest(3,N, 20*N, new HKZ()), "Mbe
 runcrb(-5 to 35, 3, N, "crbm3range")
 runlsuclt(-5 to 35, 3, N, "lsucltm3range")
 
+
 runsim(5 to 30, sp5, N, iters, () => new HAF(5,N), "HAFm5small")
 runsim(5 to 30, sp5, N, iters, () => new CPFHAF(5,N), "CPFHAFm5small")
 //runsim(5 to 30, sp5, N, iters, () => new Babai(5,N, new LLL()), "Babaim5small")
[…]

### 2013-02-22 code/bench.scala
added benching scripts.  This tells quite a nice story!

@@ -18,20 +18,21 @@ import pubsim.lattices.reduction.None
 import pubsim.lattices.reduction.HKZ
 import pubsim.lattices.reduction.LLL
 
-val MIN_BENCH_DURATION : Long = 40000000000L; // (20 secs)  
+val MIN_BENCH_DURATION : Long = 5000000000L; // (20 secs)  
 val Ns = List(30, 60, 100, 200, 350, 500);
 
 def npow(x : Int, t : Int) : Double = if(t<=0) 1.0 else scala.math.pow(x,t)
 val p3 = (0 to 3).map( k => 0.25/pubsim.Util.factorial(k) ).toArray //3rd order paramaters
 val p5 = (0 to 5).map( k => 0.25/pubsim.Util.factorial(k) ).toArray //5th order paramaters
 
-runbench(Ns, sp3, N, iters, () => new HAF(3,N), "benchHAF3")
-runbench(Ns, sp3, N, iters, () => new Mbest(3,N), "benchMbest3")
-runbench(Ns, sp5, N, iters, () => new HAF(5,N), "benchHAF5")
-runbench(Ns, sp5, N, iters, () => new Mbest(5,N), "benchMbest5")
+runbench(Ns, p3, 15, MIN_BENCH_DURATION, (N : Int) => new HAF(3,N), "benchHAF3")
+runbench(Ns, p3, 15, MIN_BENCH_DURATION, (N : Int) => new Mbest(3,N, 20*N), "benchMbest3")
+runbench(Ns, p5, 15, MIN_BENCH_DURATION, (N : Int) => new HAF(5,N), "benchHAF5")
+runbench(Ns, p5, 15, MIN_BENCH_DURATION, (N : Int) => new Mbest(5,N, 20*N), "benchMbest5")
 
-def runbench(N : Seq[Int], params : Array[Double], snrdB : Double, benchtime : Double, estf : () => PolynomialPhaseEstimatorInterface, name : String) {
+def runbench(N : Seq[Int], params : Array[Double], snrdB : Double, benchtime : Double, estf : Int => PolynomialPhaseEstimatorInterface, name : String) {
 
+val m = params.length-1
 val snr = scala.math.pow(10.0, snrdB/10.0)
 val noise =  new GaussianNoise(0,1.0/snr/2.0) //variance for real and imaginary parts (divide by 2)
 
@@ -39,12 +40,12 @@ val timelist = Ns.map{ N =>
 
[…]

### 2013-02-22 code/benchmark.sh

@@ -1,5 +1,5 @@
-export JAVA_OPTS="-Xprof -server -Xms1g -Xmx1g"
-#export JAVA_OPTS="-d64 -server -Xms1g -Xmx1g"
+#export JAVA_OPTS="-Xprof -server -Xms1g -Xmx1g"
+export JAVA_OPTS="-d64 -server -Xms1g -Xmx1g"
 
 CP=""
 for f in java/lib/*.jar

### 2013-02-22 paper.tex
minor

@@
 A drawback of the LSU estimator is that computing a nearest lattice point is, in general, computationally difficult.
-In \cite{McKilliam2009asilomar_polyest_lattice}, two standard techniques were considered, the \emph{sphere decoder}~\cite{Pohst_sphere_decoder_1981,Viterbo_sphere_decoder_1999} that exactly computes a nearest lattice point, and \emph{Babai's nearest plane algorithm}~\cite{Babai1986} that only produces an approximation.
+In \cite{McKilliam2009asilomar_polyest_lattice}, two standard techniques were considered, the \emph{sphere decoder}~\cite{Pohst_sphere_decoder_1981,Viterbo_sphere_decoder_1999,Agrell2002} that exactly computes a nearest lattice point, and \emph{Babai's nearest plane algorithm}~\cite{Babai1986} that only produces an approximation.
 The sphere decoder was observed to have excellent statistical performance but can only be computed efficiently for small $N$ less than about $60$.
 Babai's algorithm is computationally simple, but its statistical performance was observed to be comparatively poor at low signal to noise ratio (SNR).
-A major point of interest is that the lattices considered are not \emph{random}, and may admit fast nearest point algorithms.
+A major point of interest is that the lattices considered has significant structure, and may admit fast nearest point algorithms.
 This has been studied in~\cite{McKilliam2010thesis} where polynomial time algorithms were found that compute the nearest point exactly.
@@
 In addition to the above computational results a major purpose of this paper is to study the asymptotic properties of the LSU estimator.
-Under mild assumptions about the distribution of the noise we prove that the LSU estimator is strongly consistent, that is, the estimated coefficients converge (almost surely) to the true coefficients as the number of observations $N$ grows.
+Under mild assumptions about the distribution of the noise we prove that the LSU estimator is strongly consistent, that is, the estimated coefficients converge almost surely to the true coefficients as the number of observations $N$ grows.
 %Similar results were stated without proof in~\cite{McKilliam_polyphase_est_icassp_2011}.
@@
 A problem with the `small noise' assumption is that it is not known that the threshold continues to occur at increasingly smaller values of SNR as $N$ increases, or if it instead converges to some finite (but perhaps small) value of SNR.
-The consequence of the later occurance being that the estimator will be innaccurate if the SNR is smaller than this finite value, regardless of the number of observations $N$ obtained.
+The consequence of the later being that the estimator will be innaccurate if the SNR is smaller than this finite value, regardless of the number of observations $N$ obtained.
 Our analysis of the LSU estimator requires no `small noise' assumpion and guarantees that no such finite value of SNR exists for the LSU estimator.
-As a result the LSU estimator will be accurate for signals with arbitrarily small SNR provided that the number of observations is sufficiently large.
+As a result the LSU estimator can be accurate for signals with arbitrarily small SNR provided that the number of observations is sufficiently large.
 
@@
 A key property of the LSU estimator is that it functions correctly for polynomial phase parameters contained anywhere inside the identifiable region described in Section~\ref{sec:ident_aliasing} and~\cite{Xia_dynamic_range_pps_2001,McKilliam2009IndentifiabliltyAliasingPolyphase}.
-This is the maximum possible range.
+This is the maximum possible range of parameters.
 By constrast popular estimators such as the HAF and CPF operate correctly only for parameters in a range much smaller than the identifiable region.
@@
 Approaches exist that extend the range of parameters suitable for the HAF~\cite{Zhou_lag_diverity_PPT_1997,Djurovic_aliasing_2012}, but the resulting estimators typically have poor statistical performance.
-For this reason, the LSU estimator might be the only presently existing estimator that is viable for certain applications, such as interference supression.
+For this reason, the LSU estimator may be the only presently existing estimator that is viable for certain applications, such as interference supression.
 The theoretical results given in Section~\ref{sec:least-squar-unwr} prove that the LSU estimator functions correctly over the entire identifiable region.
@@
 A complete proof of normality is available in~\cite{McKilliam_arxiv_pps_unwrapping_2012}.
-Section~\ref{sec:simulations} describes the results of Monte Carlo simulations that compare the performance of the LSU estimator with existing estimators including the HAF and CPF.
+Section~\ref{sec:simulations} describes the results of Monte Carlo simulations that compare the LSU estimator with existing estimators including the HAF and CPF.
 These simulations agree with the derived asymptotic properties.

### 2013-02-22 code/bench.scala

@@ -18,8 +18,8 @@ import pubsim.lattices.reduction.None
 import pubsim.lattices.reduction.HKZ
 import pubsim.lattices.reduction.LLL
 
-val MIN_BENCH_DURATION : Long = 5000000000L; // (20 secs)  
-val Ns = List(30, 60, 100, 200, 350, 500);
+val MIN_BENCH_DURATION : Long = 20000000000L; // (20 secs)  
+val Ns = List(10, 20, 30, 60, 100, 200, 350, 500,700,1000);
 
 def npow(x : Int, t : Int) : Double = if(t<=0) 1.0 else scala.math.pow(x,t)
 val p3 = (0 to 3).map( k => 0.25/pubsim.Util.factorial(k) ).toArray //3rd order paramaters

### 2013-02-23 paper.tex
more paper edits

@@
 Nevertheless fast (exact or approximate) algorithms may exist for these special lattices, that are yet to be discovered.
-In this paper we apply another general purpose approximate nearest point algorithm, called the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder} for approximating the LSU estimator.
+In this paper we apply another general purpose approximate nearest point algorithm, called the $K$-best algorithm~\cite{Baro_list_sphere_dec_2003,Zhan2006_K_best_sphere_decoder,Mondal_Kbest_again_2009} for approximating the LSU estimator.
 We show that this provides near sphere decoder (exact) performance, but can be computed in a reasonable amount of time if $N$ is not too large (approximately $N$ less than $1000$).
@@
 The information contained in the amplitude is not essential for accurate estimation of polynomial phase signals, and can be discarded with only small loss.
-For example, in the case that $\{X_n\}$ is white and Gaussian, our results show that the asymptotic variance of the LSU estimator converges to the Cramer-Rao lower bound (CRB) as the SNR becomes large, and converges to approximately 3 times the CRB as the SNR approaches zero.
+For example, in the case that $\{X_n\}$ is white and Gaussian, our results show that the asymptotic variance of the LSU estimator converges to the Cramer-Rao lower bound (CRB) as the SNR becomes large, and converges to $\tfrac{2}{3}\pi$ times the CRB as the SNR approaches zero.
 Related results have been presented previously~\cite{Quinn2009_dasp_phase_only_information_loss}.
-It is worth noting that this factor of 3 times the CRB at low SNR is better than popular estimators such as the HAF and CPF, for which the ratio between the asymptotic variance and the CRB diverges as the SNR approaches zero~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.
+It is worth noting that this factor of $\tfrac{2}{3}\pi$ times the CRB at low SNR is better than popular estimators such as the HAF and CPF, for which the ratio between the asymptotic variance and the CRB diverges as the SNR approaches zero~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.
 
@@
 The nearest lattice point problem is: Given $\ybf\in\reals^n$ and some lattice $\Lambda \subset \reals^n$, find a lattice point $\xbf \in \Lambda$ such that the Euclidean distance between $\ybf$ and $\xbf$ is minimised, that is, solve $\arg\min_{\xbf \in \Lambda} \|\ybf - \xbf\|^2.$  The nearest lattice point problem is known to be NP-hard under certain conditions when the lattice itself is considered as an additional input parameter \cite{micciancio_hardness_2001, AjtaiShortestVecProbNPHard1998}.
-On the other hand,  fast algorithms that compute a nearest point often exist for specific lattices with known structure~\cite{Conway1982FastQuantDec,McKilliam2009CoxeterLattices,McKilliam2008}.
-Even in the general case, where the lattice has no known structure, algorithms exist that can compute a nearest lattice point in reasonable time if the dimension is small \cite{Agrell2002, Viterbo_sphere_decoder_1999, Pohst_sphere_decoder_1981}.
-One such algorithm is the \emph{sphere decoder}~\cite{Pohst_sphere_decoder_1981,Viterbo_sphere_decoder_1999}.
+On the other hand,  fast algorithms that compute a nearest point often exist for specific lattices with known structure~\cite{SPLAG,Conway1982FastQuantDec,McKilliam2009CoxeterLattices,McKilliam2008}.
+Even in the general case, where the lattice has no known structure, algorithms exist that can compute a nearest lattice point in reasonable time if the dimension is small~\cite{Agrell2002, Viterbo_sphere_decoder_1999, Pohst_sphere_decoder_1981}.
+These algorithms typically operate using a tree search to enumerate all lattice points inside a ball of radius sufficiently large to contain a nearest point.
+The algorithms vary in the order that branches in the tree are searched.
+A popular approach due to Schnorr and Eucher~\cite{schnorr_euchner_sd_1994,Agrell2002} searches branches in a greedy order, and this is the approach we implement here.
+We refer to this algorithm as the~\emph{sphere decoder} in this paper, although other variants of the algorithm also go by this name.
+
 With modern computers, the sphere decoder is computationally viable if the dimension of the lattice is less than approximately $60$.
@@
 One example is Babai's nearest plane algorithm \cite{Babai1986}.
-Yet another approximate algorithm is the $K$-best algorithm studied in~\cite{Zhan2006_K_best_sphere_decoder} that operates similarly to the sequential $M$-algorithm~\citep{Anderson1984_seq_coding_alg} used in coding theory.
-The $K$-best algorithm is more expensive that Babai's algorithm, but typically produces a better approximation.
-In this paper we make use of the $K$-best algorithm for approximating the LSU estimator.
-We will find that it produces excellent estimates of polynomial phase coefficients and can be computed in a reasonable amount of time for moderate numbers of observations ($N$ less that about 1000).
-We will not detail the workings of the $K$-best algorithm in this paper.
-A thorough description is given in~\cite{Zhan2006_K_best_sphere_decoder}.
+Another pragmatic approximate approach is to employ the sphere decoder, but prune branches in the search tree to ensure that there are atmost a finite number, say $K$, remaining branches at any time.
+This approach has been studied previously~\cite{Baro_list_sphere_dec_2003,Zhan2006_K_best_sphere_decoder,Mondal_Kbest_again_2009} and resembles what is called the sequential $M$-algorithm~\citep{Anderson1984_seq_coding_alg} from coding theory.
+The approaches vary in how branches are pruned.
+In our implementation we prune branches using the same greedy metric as that used by the Schnorr and Euchner sphere deocder~\cite{schnorr_euchner_sd_1994,Agrell2002} for deciding the order branches are searched.
+Following~\cite{Zhan2006_K_best_sphere_decoder} we refer to this as \emph{$K$-best algorithm}.
+We will find that this algorithm produces excellent estimates of polynomial phase coefficients and can be computed in a reasonable amount of time for moderate numbers of observations ($N$ less that about 1000).
 
@@
 The sample mean square error (MSE) of the $k$th coefficient is computed according to $\tfrac{1}{T}\sum_{t=1}^T \widehat{\lambda}_{k,t}^2$ where $\widehat{\lambda}_{k,t}$ is the $k$th element of $\widehat{\lambdabf}_t$.
-In all simulations the LSU estimator is approximated using the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder}.
+In all simulations the LSU estimator is approximated using the $K$-best method described in Section~\ref{sec:lattice-theory}.
+The maximum number of branches allowed is set to $K=20*N$.
 
@@
 The asymptotic variance of the LSU estimator predicted by Theorem~\ref{thm:asymp_proof} is displayed by the dashed line.
-Provided the noise variance is small enough (so that the `threshold' is avoided) the sample MSE of the LSU estimator is close to that predicted by Theorem~\ref{thm:asymp_proof}.
-Using Theorem~\ref{thm:asymp_proof} we predict the ratio between the CRB and the asymptotic variance of the LSU estimator to approach 1 as SNR approaches $\infty$ and approximately $3$ as SNR approaches zero~\cite{Quinn2009_dasp_phase_only_information_loss}.
+Provided the SNR is large enough (so that the `threshold' is avoided) the sample MSE of the LSU estimator is close to that predicted by Theorem~\ref{thm:asymp_proof}.
+Using Theorem~\ref{thm:asymp_proof} we can predict the ratio between the CRB and the asymptotic variance of the LSU estimator to approach 1 as SNR approaches $\infty$ and $\tfrac{2}{3}\pi$ as SNR approaches zero~(see Lemma~3 and the subsequent discussion in~\cite{Quinn2009_dasp_phase_only_information_loss}).
 Contrastingly, the ratio between the CRB and asymptotic variance of the HAF and CPF estimators is known to diverge as SNR goes to zero~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.
@@
 When $m=5$ and $N=199$ the space of coefficients searched by the HAF and CPF-HAF estimators is approximately $10^{16}$ times smaller than the identifiable region.
-This constitutes is a dramatic reduction in the range of detectable polynomial phase signals.
-Simulations have been run with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}, \tfrac{1}{96}, \tfrac{1}{480}  \right]$, and in this case we obtain results resembling those in Figure~\ref{plot:m3big}.
+This constitutes a dramatic reduction in the range of detectable polynomial phase signals.
+Figure~\ref{plot:m3big} displays results under the same conditions but with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}, \tfrac{1}{96}, \tfrac{1}{480}  \right]$.
+The conclusion is similar to that from Figure~\ref{plot:m3big}.
 The performance of the LSU estimator remains unchanged, but the HAF and CPF-HAF estimators fail completely.
@@
   		\includegraphics{code/data/plot3-1.mps} 
-  		\caption{Polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}  \right]$.} 
+  		\caption{Sample MSE for highest order coefficientl $\tilde{\mu}_3$ for a polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}  \right]$.}
   		\label{plot:m3small} 
@@
   		\includegraphics{code/data/plot3-2.mps} 
-  		\caption{Polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$.} 
+  		\caption{Sample MSE for the highest order coefficientl $\tilde{\mu}_3$ for a polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$.}
   		\label{plot:m3big} 
@@
   		\includegraphics{code/data/plot3-3.mps} 
-  		\caption{Polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$.} 
+  		\caption{Sample MSE for the highest order coefficientl $\tilde{\mu}_3$ for a polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$.}
   		\label{plot:m3bigzw} 
@@
   		\includegraphics{code/data/plot5-1.mps} 
-  		\caption{Polynomial phase signal of order 5 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}, \tfrac{1}{96N^3}, \tfrac{1}{480N^4}   \right]$.} 
+  		\caption{Sample MSE for the highest order coefficientl $\tilde{\mu}_5$ for a polynomial phase signal of order 5 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}, \tfrac{1}{96N^3}, \tfrac{1}{480N^4}   \right]$.}
   		\label{plot:m5} 
@@
   		\includegraphics{code/data/plot5-2.mps} 
-  		\caption{Polynomial phase signal of order 5 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}, \tfrac{1}{96}, \tfrac{1}{480}   \right]$.} 
+  		\caption{Sample MSE for the highest order coefficientl $\tilde{\mu}_5$ for a polynomial phase signal of order 5 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}, \tfrac{1}{96}, \tfrac{1}{480}   \right]$.}
   		\label{plot:m5} 
@@
 This paper has considered the estimation of the coefficients of a noisy polynomial phase signal by least squares phase unwrapping (LSU).
-It has been shown that the LSU estimator is strongly consistent and asymptotically normally distributed.
-%Simulations are used to show that the LSU has near maximum likelihood performance in the case that the noise is additive white and Gaussian.
-The LSU estimator can be computed by finding a nearest lattice point in a lattice~\cite{McKilliam2009asilomar_polyest_lattice}.
-%Here, we consider using the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder} to achieve this in reasonable time when the number of observations, $N$, is less than about 300.
+The LSU estimator was shown to be strongly consistent and asymptotically normally distributed under mild conditions on the distribution of the noise.
+The estimator can be computed by finding a nearest lattice point in a lattice.
 Polynomial time nearest point algorithms for these lattices exist~\cite[Sec 4.3]{McKilliam2010thesis}, but these algorithms are not fast in practice.
-Instead, we considered some general purpose algorithms, the \term{sphere decoder}, \term{Babai's nearest plane algorithm} and the \term{$K$-best algorithm}.
-The sphere decoder and $K$-best algorithm result in accurate estimators.
-
-The major outstanding question is whether faster nearest point algorithms exist for these specific lattices.
+We instead make used on the appromxiate $K$-best algorithm~\cite{Baro_list_sphere_dec_2003,Zhan2006_K_best_sphere_decoder,Mondal_Kbest_again_2009}.
+This algorithm results in an accurate estimator and can be computed in a reasoable amount of time for moderately large numbers of obervations $N$.
+
+A key property of the LSU estimator is that it functions correctly for polynomial phase parameters contained anywhere inside the identifiable region described in Section~\ref{sec:ident_aliasing}.
+This is the maximum possible range of parameters.
+By constrast popular estimators such as the HAF and CPF operate correctly only for parameters in a range much smaller than the identifiable region.
+For this reason, the LSU estimator may be the only presently existing estimator that is viable for certain applications, such as interference supression~\cite{Djukanovic:2011:PMN:1950994.1951252,Djurovic_aliasing_2012}.
+
+The major outstanding question is whether faster nearest point algorithms exist for these specific lattices, particularly for the case when $N$ is large.
 Considering the excellent statistical performance (both theoretically and practically) of the LSU estimator, even fast \emph{approximate} nearest point algorithms are likely to prove useful for the estimation of polynomial phase signals.

### 2013-02-23 code/montecarlo.scala

@@ -30,7 +30,7 @@ val bp5 = (0 to 5).map( k => 0.25/pubsim.Util.factorial(k) ).toArray //5th order
 
 val starttime = (new java.util.Date).getTime
 
-
+/*
 runsim(-5 to 15, sp3, N, iters, () => new HAF(3,N), "HAFm3small")
 runsim(-5 to 15, bp3, N, iters, () => new HAF(3,N), "HAFm3big")
 runsim(-5 to 15, sp3, N, iters, () => new CPF(N), "CPFm3small")
@@ -47,8 +47,9 @@ runsim(-5 to 35 by 2, bp3, N, iters, () => new ZW(3,N,N/3,N/3+1), "ZWm3big")
 runsim(-5 to 35 by 2, bp3, N, iters, () => new Mbest(3,N, 20*N, new HKZ()), "Mbestm3bigrange")
 runcrb(-5 to 35, 3, N, "crbm3range")
 runlsuclt(-5 to 35, 3, N, "lsucltm3range")
+*/
 
-
+/*
 runsim(5 to 30, sp5, N, iters, () => new HAF(5,N), "HAFm5small")
 runsim(5 to 30, sp5, N, iters, () => new CPFHAF(5,N), "CPFHAFm5small")
 //runsim(5 to 30, sp5, N, iters, () => new Babai(5,N, new LLL()), "Babaim5small")
@@ -56,10 +57,12 @@ runsim(5 to 30, sp5, N, iters, () => new Mbest(5,N, 20*N, new LLL()), "Mbestm5sm
 runsim(5 to 30, bp5, N, iters, () => new HAF(5,N), "HAFm5big")
 runsim(5 to 30, bp5, N, iters, () => new CPFHAF(5,N), "CPFHAFm5big")
 //runsim(5 to 30, bp5, N, iters, () => new Babai(5,N, new LLL()), "Babaim5big")
+*/
 runsim(5 to 30, bp5, N, iters, () => new Mbest(5,N, 20*N, new LLL()), "Mbestm5big")
+/*
 runcrb(5 to 30, 5, N, "crbm5")
 runlsuclt(5 to 30, 5, N, "lsucltm5")
-
+*/
 
 val runtime = (new java.util.Date).getTime - starttime
 println("Simulation finshed in " + (runtime/1000.0) + " seconds.\n")

### 2013-02-24 code/montecarlo.scala
minor

@@ -66,7 +66,6 @@ println("Simulation finshed in " + (runtime/1000.0) + " seconds.\n")
 
 
 
-
 /** Runs a simulation with given parameters and stores output in a file */
 def runsim(snrdbs : Seq[Int], params : Array[Double], N : Int, iters : Int, estf : () => PolynomialPhaseEstimatorInterface, name : String) {
 

### 2013-02-25 code/bench.scala
added some new data

@@ -18,15 +18,15 @@ import pubsim.lattices.reduction.None
 import pubsim.lattices.reduction.HKZ
 import pubsim.lattices.reduction.LLL
 
-val MIN_BENCH_DURATION : Long = 10000000000L; // (20 secs)  
-val Ns = List(10, 15, 20, 30, 45, 65, 100, 200, 350, 500,700,1000);
+val MIN_BENCH_DURATION : Long = 50000000000L; // (20 secs)  
+val Ns = List(10,15,20,30,45,65,100,150,250,400,550,750,1000);
 
 def npow(x : Int, t : Int) : Double = if(t<=0) 1.0 else scala.math.pow(x,t)
 val p3 = (0 to 3).map( k => 0.25/pubsim.Util.factorial(k) ).toArray //3rd order paramaters
 val p5 = (0 to 5).map( k => 0.25/pubsim.Util.factorial(k) ).toArray //5th order paramaters
 
 runbench(Ns, p3, 10, MIN_BENCH_DURATION, (N : Int) => new HAF(3,N), "benchHAF3")
-runbench(Ns.filter(_<=500), p3, 0, MIN_BENCH_DURATION, (N : Int) => new Mbest(3,N, 20*N), "benchMbest3snr0")
+runbench(Ns.filter(_<=550), p3, 0, MIN_BENCH_DURATION, (N : Int) => new Mbest(3,N, 20*N), "benchMbest3snr0")
 runbench(Ns, p3, 10, MIN_BENCH_DURATION, (N : Int) => new Mbest(3,N, 20*N), "benchMbest3snr10")
 runbench(Ns, p3, 20, MIN_BENCH_DURATION, (N : Int) => new Mbest(3,N, 20*N), "benchMbest3snr20")
 

### 2013-02-25 paper.tex

@@
 
-The \emph{intrinsic mean} (or \emph{Fr\'{e}chet mean}) of $\Phi_n$ is defined as~\cite{McKilliam_mean_dir_est_sq_arc_length2010,bwhk07a,Bhattacharya_int_ext_means_2003,Bhattacharya_int_ext_means_2005},
+The \emph{intrinsic mean} (or \emph{Fr\'{e}chet mean}) of $\Phi_n$ is defined as~\cite{McKilliam_mean_dir_est_sq_arc_length2010,Bhattacharya_int_ext_means_2003},
 \begin{equation}\label{eq:intrmeandefn}
@@
 
+Figure~\ref{plot:bench} shows the running time of the LSU estimator computed using the $K$-best method versus the HAF estimator for polynomial phase signals of order 3.
+The HAF estimator is computed using the fast Fourier transform and has complexity $O(N \log N)$.
+The $K$-best method tends to be faster when the SNR is large.
+This is due to the algorithm rapidly finding a nearest point, and subsequently terminating branches in its search tree early.
+On the other hand, when the SNR is small, all branches in the tree will be searched.
+Interestingly the LSU estimator computed using the $K$-best algorithm is \emph{faster} than the HAF when the SNR is large and when $N$ is approximately less than 100.
+The LSU estimator can be computed in a reasonable amount of time (a few minutes) when $N=1000$.
+The algorithm becomes prohibitively expensive for large $N$.
+Benchmarks have been performed with polynomial phase signal of order larger than $3$.
+The results are ommited due to space constraints.
+The running time of the LSU estimator does not appear to change significantly as $m$ increases.
+This is expected since a nearest point in a lattice of approximately the same dimension $N-m-1$ must be computed.
+The complexity of the HAF estimator grows linearly with $m$.
+
 \begin{figure}[p] 
@@
 
-% \begin{figure}[t] 
-%    	\centering 
-%   		\includegraphics{code/data/plot567-2.mps} 
-%   		\caption{BLERG} 
-%   		\label{plot:m6} 
-%  \end{figure}
-
-% \begin{figure}[p] 
-%    	\centering 
-%   		\includegraphics{code/data/plot567-3.mps} 
-%   		\caption{BLERG} 
-%   		\label{plot:polyest} 
-%  \end{figure}
-
-%\begin{figure}[t] 
-%   	\centering 
-%  		\includegraphics{code/data/plot5-2.mps} 
-%  		\caption{BLERG} 
-%  		\label{plot:polyest} 
-% \end{figure}
+\begin{figure}[p] 
+   	\centering 
+  		%\includegraphics{code/data/bench-1.mps} 
+  		\caption{Running time in seconds with the LSU and HAF estimators.} 
+  		\label{plot:bench} 
+\end{figure}
 

### 2013-02-25 code/bench.scala
added to bench plot

@@ -26,11 +26,12 @@ val p3 = (0 to 3).map( k => 0.25/pubsim.Util.factorial(k) ).toArray //3rd order 
 val p5 = (0 to 5).map( k => 0.25/pubsim.Util.factorial(k) ).toArray //5th order paramaters
 
 runbench(Ns, p3, 10, MIN_BENCH_DURATION, (N : Int) => new HAF(3,N), "benchHAF3")
-runbench(Ns.filter(_<=550), p3, 0, MIN_BENCH_DURATION, (N : Int) => new Mbest(3,N, 20*N), "benchMbest3snr0")
+runbench(Ns, p3, 0, MIN_BENCH_DURATION, (N : Int) => new Mbest(3,N, 20*N), "benchMbest3snr0")
+runbench(Ns, p3, 5, MIN_BENCH_DURATION, (N : Int) => new Mbest(3,N, 20*N), "benchMbest3snr5")
 runbench(Ns, p3, 10, MIN_BENCH_DURATION, (N : Int) => new Mbest(3,N, 20*N), "benchMbest3snr10")
 runbench(Ns, p3, 20, MIN_BENCH_DURATION, (N : Int) => new Mbest(3,N, 20*N), "benchMbest3snr20")
 
-def runbench(N : Seq[Int], params : Array[Double], snrdB : Double, benchtime : Double, estf : Int => PolynomialPhaseEstimatorInterface, name : String) {
+def runbench(Ns : Seq[Int], params : Array[Double], snrdB : Double, benchtime : Double, estf : Int => PolynomialPhaseEstimatorInterface, name : String) {
 
 val m = params.length-1
 val snr = scala.math.pow(10.0, snrdB/10.0)
@@ -47,7 +48,7 @@ val est = estf(N)
 print(name + " N = " + N + " warming up ... ")
 var numiters = 0
 val warmupstarttime = System.nanoTime
-while(System.nanoTime - warmupstarttime < MIN_BENCH_DURATION){
+while(System.nanoTime - warmupstarttime < benchtime){
    siggen.generateReceivedSignal
    est.estimate(siggen.getReal, siggen.getImag)
    numiters = numiters+1

### 2013-02-26 paper.tex
added to bench plot
added to paper and bib and bench.  Need to implement the PHAF, which is annoying

@@
 For this reason many authors have considered alternative approaches to polynomial phase estimation.
-These can loosely be grouped into two classes: estimators based on `multilinear transforms', such as the high order ambiguity function (HAF)~\cite{Peleg_DPT_1995,Peleg1991_est_class_PPS_1991,Porat_asympt_HAF_DPT_1996,Farquharson_another_poly_est_2005,Boashash_Oshea_1994}, and cubic phase function (CPF)~\cite{Oshea_cpf_2004,Oshea_cpf_2002_letter,Djurovic_haf_cpf_2012}; and estimators based on phase unwrapping, such as Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, the estimator of Djuric and Kay~\cite{Djuric_phase_unwrap_chirp_1990}, and Morelande's Bayesian unwrapping estimator~\cite{Morelande_bayes_unwrapping_2009_tsp}.
+These can loosely be grouped into two classes: estimators based on `multilinear transforms', such as the high order ambiguity function (HAF)~\cite{Peleg_DPT_1995,Peleg1991_est_class_PPS_1991,Porat_asympt_HAF_DPT_1996,Farquharson_another_poly_est_2005,Boashash_Oshea_1994,Barbarossa_analysis_of_PPS_1997}, and cubic phase function (CPF)~\cite{Oshea_cpf_2004,Oshea_cpf_2002_letter,Djurovic_haf_cpf_2012}; and estimators based on phase unwrapping, such as Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, the estimator of Djuric and Kay~\cite{Djuric_phase_unwrap_chirp_1990}, and Morelande's Bayesian unwrapping estimator~\cite{Morelande_bayes_unwrapping_2009_tsp}.
 
@@
 
-The analysis of most existing polynomial phase estimators make what is known as a `small noise' assumption~\cite{Peleg_linear_FM,Oshea_cpf_2004,Djurovic_haf_cpf_2012,Djuric_phase_unwrap_chirp_1990}.
+The analysis of most existing polynomial phase estimators make what is known as a `small noise' assumption~\cite{Peleg_linear_FM,Oshea_cpf_2004,Djurovic_haf_cpf_2012,Djuric_phase_unwrap_chirp_1990,Barbarossa_analysis_of_PPS_1997,Barbarossa_PHAF_1998}.
 This amounts to assuming that the estimator $\hat{\mu}_k$ of the $k$th coefficient is such that $N^{k}(\tilde{\mu}_k -  \hat{\mu}_k) \rightarrow 0$ (in some sense) as $N \rightarrow \infty$.
-This assumption is used to justify the use of first and second order expansions~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats} (typically called a `perturbation analysis' in this literature~\cite{Peleg_linear_FM,Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004}) to obtain expressions for the asymptotic mean square error of the estimator in question.
+This assumption is used to justify the use what is called a `perturbation analysis'~\cite{Porat_asympt_HAF_DPT_1996,Peleg_linear_FM,Oshea_cpf_2004,Djurovic_haf_cpf_2012,Djuric_phase_unwrap_chirp_1990,Barbarossa_analysis_of_PPS_1997,Barbarossa_PHAF_1998}) to obtain expressions for the asymptotic mean square error of the estimator in question.
 The `small noise' assumption is usually justified by Monte-Carlo simulations showing the asymptotic mean square error obtained by the perturbation analysis to be accurate provided that the signal to noise ratio (SNR) is larger than a `threshold' value (see the simulations in Section~\ref{sec:simulations} for examples).
-Typically one observes the threshold to occur at smaller values of SNR as $N$ increases.
+Often the threshold is observed to occur at smaller values of SNR as $N$ increases.
 A problem with the `small noise' assumption is that it is not known that the threshold continues to occur at increasingly smaller values of SNR as $N$ increases, or if it instead converges to some finite (but perhaps small) value of SNR.
@@
 These simulations agree with the derived asymptotic properties.
-BLERG: benchmarks.
 
@@
 The algorithm becomes prohibitively expensive for large $N$.
-Benchmarks have been performed with polynomial phase signal of order larger than $3$.
+Benchmarks have been performed with polynomial phase signals of order larger than $3$.
 The results are ommited due to space constraints.
@@
    	\centering 
-  		%\includegraphics{code/data/bench-1.mps} 
-  		\caption{Running time in seconds with the LSU and HAF estimators.} 
+  		\includegraphics{code/data/bench-1.mps} 
+  		\caption{Running time in seconds with the LSU and HAF estimators.  The LSU estimator is benchmarked at SNR equal to \unit[0]{dB}, \unit[5]{dB}, \unit[10]{dB} and \unit[20]{dB}.} 
   		\label{plot:bench} 
@@
 
+\nocite{OShea_refiningpps_2010}
+
 \small 

### 2013-02-26 code/java/src/pubsim/poly/CPF.java

@@ -11,7 +11,7 @@ import pubsim.optimisation.NewtonRaphson;
  * Implements O'Shea's cubic phase function estimator
  * Peter O'Shea "A Fast Algorithm for Estimating the Parameters of a Quadratic FM Signal"
  * IEEE Trans. Signal Proc. Vol 53, Feb 2004.
- * @todo Dechirp and estimate phase and frequency
+ * @todo Dechirp and estimate phase and frequency, only quadratic and cubic term currently estimated.
  * @author Robby McKilliam
  */
 public class CPF extends AbstractPolynomialPhaseEstimator {

### 2013-02-26 paper.tex
added to paper.

@@
 The analysis of most existing polynomial phase estimators make what is known as a `small noise' assumption~\cite{Peleg_linear_FM,Oshea_cpf_2004,Djurovic_haf_cpf_2012,Djuric_phase_unwrap_chirp_1990,Barbarossa_analysis_of_PPS_1997,Barbarossa_PHAF_1998}.
-This amounts to assuming that the estimator $\hat{\mu}_k$ of the $k$th coefficient is such that $N^{k}(\tilde{\mu}_k -  \hat{\mu}_k) \rightarrow 0$ (in some sense) as $N \rightarrow \infty$.
-This assumption is used to justify the use what is called a `perturbation analysis'~\cite{Porat_asympt_HAF_DPT_1996,Peleg_linear_FM,Oshea_cpf_2004,Djurovic_haf_cpf_2012,Djuric_phase_unwrap_chirp_1990,Barbarossa_analysis_of_PPS_1997,Barbarossa_PHAF_1998}) to obtain expressions for the asymptotic mean square error of the estimator in question.
+%This amounts to assuming that the estimator $\hat{\mu}_k$ of the $k$th coefficient is such that $N^{k}(\tilde{\mu}_k -  \hat{\mu}_k) \rightarrow 0$ (in some sense) as $N \rightarrow \infty$.
+This assumption is used to justify the use what is called a `perturbation analysis'~\cite{Porat_asympt_HAF_DPT_1996,Peleg_linear_FM,Oshea_cpf_2004,Djurovic_haf_cpf_2012,Djuric_phase_unwrap_chirp_1990,Barbarossa_analysis_of_PPS_1997,Barbarossa_PHAF_1998} to obtain expressions for the asymptotic mean square error of the estimator in question.
 The `small noise' assumption is usually justified by Monte-Carlo simulations showing the asymptotic mean square error obtained by the perturbation analysis to be accurate provided that the signal to noise ratio (SNR) is larger than a `threshold' value (see the simulations in Section~\ref{sec:simulations} for examples).
@@
 A problem with the `small noise' assumption is that it is not known that the threshold continues to occur at increasingly smaller values of SNR as $N$ increases, or if it instead converges to some finite (but perhaps small) value of SNR.
-The consequence of the later being that the estimator will be innaccurate if the SNR is smaller than this finite value, regardless of the number of observations $N$ obtained.
+The consequence of the later being that the estimator will be innaccurate if the SNR is smaller than this finite value, regardless of the number of observations $N$.
 Our analysis of the LSU estimator requires no `small noise' assumpion and guarantees that no such finite value of SNR exists for the LSU estimator.
@@
 This is due to the algorithm rapidly finding a nearest point, and subsequently terminating branches in its search tree early.
-On the other hand, when the SNR is small, all branches in the tree will be searched.
+On the other hand, when the SNR is small, all branches in the tree tend to be searched.
 Interestingly the LSU estimator computed using the $K$-best algorithm is \emph{faster} than the HAF when the SNR is large and when $N$ is approximately less than 100.

### 2013-02-26 paper.tex
spellchecked

@@
 A key feature is that the LSU estimator functions accurately over a far wider range of parameters than many popular existing estimators.
-Monte-carlo simualtions assert the excellent statistical performance of the LSU estimator when compared to existing state-of-the-art estimators and support our theoretical results.
+Monte-Carlo simulations assert the excellent statistical performance of the LSU estimator when compared to existing state-of-the-art estimators and support our theoretical results.
 \end{abstract}
@@
 A problem with the `small noise' assumption is that it is not known that the threshold continues to occur at increasingly smaller values of SNR as $N$ increases, or if it instead converges to some finite (but perhaps small) value of SNR.
-The consequence of the later being that the estimator will be innaccurate if the SNR is smaller than this finite value, regardless of the number of observations $N$.
-Our analysis of the LSU estimator requires no `small noise' assumpion and guarantees that no such finite value of SNR exists for the LSU estimator.
+The consequence of the later being that the estimator will be inaccurate if the SNR is smaller than this finite value, regardless of the number of observations $N$.
+Our analysis of the LSU estimator requires no `small noise' assumption and guarantees that no such finite value of SNR exists for the LSU estimator.
 As a result the LSU estimator can be accurate for signals with arbitrarily small SNR provided that the number of observations is sufficiently large.
 
-An analysis of the HAF estimator under the assumption that the noise $\{X_n\}$ is white and Gaussian and without the `small noise' assumption has been made in~\cite{Porat_asympt_HAF_DPT_1996} (see inparticular the proof of Lemma~1 and the remark that follows).
+An analysis of the HAF estimator under the assumption that the noise $\{X_n\}$ is white and Gaussian and without the `small noise' assumption has been made in~\cite{Porat_asympt_HAF_DPT_1996} (see in particular the proof of Lemma~1 and the remark that follows).
 To date, the literature does not contain a corresponding analysis of any estimator based on phase unwrapping.
 This has apparently lead to the misconception that phase unwrapping estimators are only accurate when the noise is `small' in some sense.
-This misconception would appear to be motivated by early papers such as~\cite{Tretter1985,Kay1989} where it was observed that the noise occuring in the phase of the signal only looks `Gaussian' when the complex noise producing it has small variance.
+This misconception would appear to be motivated by early papers such as~\cite{Tretter1985,Kay1989} where it was observed that the noise occurring in the phase of the signal only looks `Gaussian' when the complex noise producing it has small variance.
 The results in this paper make it clear that phase unwrapping estimators can be accurate at arbitrarily low SNR.
@@
 This is the maximum possible range of parameters.
-By constrast popular estimators such as the HAF and CPF operate correctly only for parameters in a range much smaller than the identifiable region.
-This property has been studied previously~\cite{Djurovic_aliasing_2012,McKilliam2009IndentifiabliltyAliasingPolyphase,Zhou_lag_diverity_PPT_1997,Xia_dynamic_range_pps_2001} and can be a critical limitation for some applications, for example, interference supression in the pressence of a jammer~\cite{Djukanovic:2011:PMN:1950994.1951252,Djurovic_aliasing_2012}.
+By contrast popular estimators such as the HAF and CPF operate correctly only for parameters in a range much smaller than the identifiable region.
+This property has been studied previously~\cite{Djurovic_aliasing_2012,McKilliam2009IndentifiabliltyAliasingPolyphase,Zhou_lag_diverity_PPT_1997,Xia_dynamic_range_pps_2001} and can be a critical limitation for some applications, for example, interference suppression in the presence of a jammer~\cite{Djukanovic:2011:PMN:1950994.1951252,Djurovic_aliasing_2012}.
 Approaches exist that extend the range of parameters suitable for the HAF~\cite{Zhou_lag_diverity_PPT_1997,Djurovic_aliasing_2012}, but the resulting estimators typically have poor statistical performance.
-For this reason, the LSU estimator may be the only presently existing estimator that is viable for certain applications, such as interference supression.
+For this reason, the LSU estimator may be the only presently existing estimator that is viable for certain applications, such as interference suppression.
 The theoretical results given in Section~\ref{sec:least-squar-unwr} prove that the LSU estimator functions correctly over the entire identifiable region.
@@
 The algorithms vary in the order that branches in the tree are searched.
-A popular approach due to Schnorr and Eucher~\cite{schnorr_euchner_sd_1994,Agrell2002} searches branches in a greedy order, and this is the approach we implement here.
+A popular approach due to Schnorr and Euchner~\cite{schnorr_euchner_sd_1994,Agrell2002} searches branches in a greedy order, and this is the approach we implement here.
 We refer to this algorithm as the~\emph{sphere decoder} in this paper, although other variants of the algorithm also go by this name.
@@
 One example is Babai's nearest plane algorithm \cite{Babai1986}.
-Another pragmatic approximate approach is to employ the sphere decoder, but prune branches in the search tree to ensure that there are atmost a finite number, say $K$, remaining branches at any time.
+Another pragmatic approximate approach is to employ the sphere decoder, but prune branches in the search tree to ensure that there are at most a finite number, say $K$, remaining branches at any time.
 This approach has been studied previously~\cite{Baro_list_sphere_dec_2003,Zhan2006_K_best_sphere_decoder,Mondal_Kbest_again_2009} and resembles what is called the sequential $M$-algorithm~\citep{Anderson1984_seq_coding_alg} from coding theory.
 The approaches vary in how branches are pruned.
-In our implementation we prune branches using the same greedy metric as that used by the Schnorr and Euchner sphere deocder~\cite{schnorr_euchner_sd_1994,Agrell2002} for deciding the order branches are searched.
+In our implementation we prune branches using the same greedy metric as that used by the Schnorr and Euchner sphere decoder~\cite{schnorr_euchner_sd_1994,Agrell2002} for deciding the order branches are searched.
 Following~\cite{Zhan2006_K_best_sphere_decoder} we refer to this as \emph{$K$-best algorithm}.
@@
 almost surely as $N\rightarrow\infty$.
-This type of resul is called a~\emph{uniform law of large numbers}~\cite{Pollard_conv_stat_proc_1984,van2009empirical}.
+This type of result is called a~\emph{uniform law of large numbers}~\cite{Pollard_conv_stat_proc_1984,van2009empirical}.
 A full proof of~\eqref{eq:SNVNunifmlln} in given in~\cite{McKilliam_mean_dir_est_sq_arc_length2010}.
@@
 where $\tau$ is the \emph{lag} and we choose $\tau = \floor{N/m}$ as suggested in~\cite{Peleg_DPT_1995}.
-The CPF estimator requires similar, but slightly more complicated conditions on the coeffcients~\cite{Oshea_cpf_2004}.
+The CPF estimator requires similar, but slightly more complicated conditions on the coefficients~\cite{Oshea_cpf_2004}.
 It is informative to compare the volume of the set of coefficients suitable for the HAF and CPF estimators with the volume of the identifiable region $B$ from~\eqref{eq:rectangular_identifiable_region}.
@@
 In this context we might expect the CPF and HAF estimators to have thresholds at lower SNR than LSU estimator.
-It is encouraging that the LSU estimator has comparible performance.
+It is encouraging that the LSU estimator has comparable performance.
 To put these results further into context, Figure~\ref{plot:m3big} repeats the simulations in Figure~\ref{plot:m3small}, but where the true coefficients are $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$.
@@
 
-The comparitively small set of coefficients suitable for the HAF and CPF estimators has been studied previously~\cite{Angeby_PPS_aliasing_2000,Djurovic_aliasing_2012,McKilliam2009IndentifiabliltyAliasingPolyphase,Zhou_lag_diverity_PPT_1997,Xia_dynamic_range_pps_2001} and can be a critical limitation for some applications, for example, interference supression in the pressence of a jammer~\cite{Djukanovic:2011:PMN:1950994.1951252,Djurovic_aliasing_2012}.
+The comparatively small set of coefficients suitable for the HAF and CPF estimators has been studied previously~\cite{Angeby_PPS_aliasing_2000,Djurovic_aliasing_2012,McKilliam2009IndentifiabliltyAliasingPolyphase,Zhou_lag_diverity_PPT_1997,Xia_dynamic_range_pps_2001} and can be a critical limitation for some applications, for example, interference suppression in the presence of a jammer~\cite{Djukanovic:2011:PMN:1950994.1951252,Djurovic_aliasing_2012}.
 Zhou and Wang~\cite{Zhou_lag_diverity_PPT_1997} describe a method based on the extended Euclidean algorithm that enables the HAF estimator to operate on the entire identifiable region $B$.
@@
 The threshold for the estimator is approximately~\unit[29]{dB}.
-There does not appear to exist (atleast in the literature) a similar method for increasing the range of coefficients for the CPF estimator.
+There does not appear to exist (at least in the literature) a similar method for increasing the range of coefficients for the CPF estimator.
 
@@
 As in the case where $m=3$, the region of coefficients suitable for the HAF and CPF-HAF estimators is much smaller that the identifiable region $B$.
-The effect becomes more servere as $m$ increases.
+The effect becomes more severe as $m$ increases.
 When $m=5$ and $N=199$ the space of coefficients searched by the HAF and CPF-HAF estimators is approximately $10^{16}$ times smaller than the identifiable region.
@@
 Benchmarks have been performed with polynomial phase signals of order larger than $3$.
-The results are ommited due to space constraints.
+The results are omitted due to space constraints.
 The running time of the LSU estimator does not appear to change significantly as $m$ increases.
@@
   		\includegraphics{code/data/plot3-1.mps} 
-  		\caption{Sample MSE for highest order coefficientl $\tilde{\mu}_3$ for a polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}  \right]$.}
+  		\caption{Sample MSE for highest order coefficient $\tilde{\mu}_3$ for a polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}  \right]$.}
   		\label{plot:m3small} 
@@
   		\includegraphics{code/data/plot3-2.mps} 
-  		\caption{Sample MSE for the highest order coefficientl $\tilde{\mu}_3$ for a polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$.}
+  		\caption{Sample MSE for the highest order coefficient $\tilde{\mu}_3$ for a polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$.}
   		\label{plot:m3big} 
@@
   		\includegraphics{code/data/plot3-3.mps} 
-  		\caption{Sample MSE for the highest order coefficientl $\tilde{\mu}_3$ for a polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$.}
+  		\caption{Sample MSE for the highest order coefficient $\tilde{\mu}_3$ for a polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$.}
   		\label{plot:m3bigzw} 
@@
   		\includegraphics{code/data/plot5-1.mps} 
-  		\caption{Sample MSE for the highest order coefficientl $\tilde{\mu}_5$ for a polynomial phase signal of order 5 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}, \tfrac{1}{96N^3}, \tfrac{1}{480N^4} \right]$.}
+  		\caption{Sample MSE for the highest order coefficient $\tilde{\mu}_5$ for a polynomial phase signal of order 5 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}, \tfrac{1}{96N^3}, \tfrac{1}{480N^4} \right]$.}
   		\label{plot:m5} 
@@
   		\includegraphics{code/data/plot5-2.mps} 
-  		\caption{Sample MSE for the highest order coefficientl $\tilde{\mu}_5$ for a polynomial phase signal of order 5 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}, \tfrac{1}{96}, \tfrac{1}{480}   \right]$.}
+  		\caption{Sample MSE for the highest order coefficient $\tilde{\mu}_5$ for a polynomial phase signal of order 5 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}, \tfrac{1}{96}, \tfrac{1}{480}   \right]$.}
   		\label{plot:m5} 
@@
 Polynomial time nearest point algorithms for these lattices exist~\cite[Sec 4.3]{McKilliam2010thesis}, but these algorithms are not fast in practice.
-We instead make used on the appromxiate $K$-best algorithm~\cite{Baro_list_sphere_dec_2003,Zhan2006_K_best_sphere_decoder,Mondal_Kbest_again_2009}.
-This algorithm results in an accurate estimator and can be computed in a reasoable amount of time for moderately large numbers of obervations $N$.
+We instead make used on the approximate $K$-best algorithm~\cite{Baro_list_sphere_dec_2003,Zhan2006_K_best_sphere_decoder,Mondal_Kbest_again_2009}.
+This algorithm results in an accurate estimator and can be computed in a reasonable amount of time for moderately large numbers of observations $N$.
 
@@
 This is the maximum possible range of parameters.
-By constrast popular estimators such as the HAF and CPF operate correctly only for parameters in a range much smaller than the identifiable region.
-For this reason, the LSU estimator may be the only presently existing estimator that is viable for certain applications, such as interference supression~\cite{Djukanovic:2011:PMN:1950994.1951252,Djurovic_aliasing_2012}.
+By contrast popular estimators such as the HAF and CPF operate correctly only for parameters in a range much smaller than the identifiable region.
+For this reason, the LSU estimator may be the only presently existing estimator that is viable for certain applications, such as interference suppression~\cite{Djukanovic:2011:PMN:1950994.1951252,Djurovic_aliasing_2012}.
 

### 2013-02-27 paper.tex
added data and to paper

@@
 For this reason many authors have considered alternative approaches to polynomial phase estimation.
-These can loosely be grouped into two classes: estimators based on `multilinear transforms', such as the high order ambiguity function (HAF)~\cite{Peleg_DPT_1995,Peleg1991_est_class_PPS_1991,Porat_asympt_HAF_DPT_1996,Farquharson_another_poly_est_2005,Boashash_Oshea_1994,Barbarossa_analysis_of_PPS_1997}, and cubic phase function (CPF)~\cite{Oshea_cpf_2004,Oshea_cpf_2002_letter,Djurovic_haf_cpf_2012}; and estimators based on phase unwrapping, such as Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, the estimator of Djuric and Kay~\cite{Djuric_phase_unwrap_chirp_1990}, and Morelande's Bayesian unwrapping estimator~\cite{Morelande_bayes_unwrapping_2009_tsp}.
+These can loosely be grouped into two classes: estimators based on `multilinear transforms', such as the high order ambiguity function (HAF)~\cite{Peleg_DPT_1995,Peleg1991_est_class_PPS_1991,Porat_asympt_HAF_DPT_1996,Farquharson_another_poly_est_2005,Boashash_Oshea_1994,Barbarossa_analysis_of_PPS_1997}, the product HAF (PHAF)~\cite{Barbarossa_PHAF_1998}, and cubic phase function (CPF)~\cite{Oshea_cpf_2004,Oshea_cpf_2002_letter,Djurovic_haf_cpf_2012,OShea_new_class_multilinear_2009}; and estimators based on phase unwrapping, such as Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, the estimator of Djuric and Kay~\cite{Djuric_phase_unwrap_chirp_1990}, and Morelande's Bayesian unwrapping estimator~\cite{Morelande_bayes_unwrapping_2009_tsp}.
 
@@
 Babai's algorithm is computationally simple, but its statistical performance was observed to be comparatively poor at low signal to noise ratio (SNR).
-A major point of interest is that the lattices considered has significant structure, and may admit fast nearest point algorithms.
+A major point of interest is that the lattices considered have significant structure, and may admit fast nearest point algorithms.
 This has been studied in~\cite{McKilliam2010thesis} where polynomial time algorithms were found that compute the nearest point exactly.
@@
 Nevertheless fast (exact or approximate) algorithms may exist for these special lattices, that are yet to be discovered.
-In this paper we apply another general purpose approximate nearest point algorithm, called the $K$-best algorithm~\cite{Baro_list_sphere_dec_2003,Zhan2006_K_best_sphere_decoder,Mondal_Kbest_again_2009} for approximating the LSU estimator.
+In this paper we apply another general purpose approximate nearest point algorithm called the $K$-best algorithm~\cite{Baro_list_sphere_dec_2003,Zhan2006_K_best_sphere_decoder,Mondal_Kbest_again_2009} for approximating the LSU estimator.
 We show that this provides near sphere decoder (exact) performance, but can be computed in a reasonable amount of time if $N$ is not too large (approximately $N$ less than $1000$).
@@
 %This amounts to assuming that the estimator $\hat{\mu}_k$ of the $k$th coefficient is such that $N^{k}(\tilde{\mu}_k -  \hat{\mu}_k) \rightarrow 0$ (in some sense) as $N \rightarrow \infty$.
-This assumption is used to justify the use what is called a `perturbation analysis'~\cite{Porat_asympt_HAF_DPT_1996,Peleg_linear_FM,Oshea_cpf_2004,Djurovic_haf_cpf_2012,Djuric_phase_unwrap_chirp_1990,Barbarossa_analysis_of_PPS_1997,Barbarossa_PHAF_1998} to obtain expressions for the asymptotic mean square error of the estimator in question.
-The `small noise' assumption is usually justified by Monte-Carlo simulations showing the asymptotic mean square error obtained by the perturbation analysis to be accurate provided that the signal to noise ratio (SNR) is larger than a `threshold' value (see the simulations in Section~\ref{sec:simulations} for examples).
+This assumption is used to justify what is called a `perturbation analysis'~%\cite{Porat_asympt_HAF_DPT_1996,Peleg_linear_FM,Oshea_cpf_2004,Djurovic_haf_cpf_2012,Djuric_phase_unwrap_chirp_1990,Barbarossa_analysis_of_PPS_1997,Barbarossa_PHAF_1998}
+to obtain expressions for the asymptotic mean square error of the estimator in question.
+The `small noise' assumption is usually supported by Monte-Carlo simulations showing the asymptotic mean square error obtained by the perturbation analysis to be accurate provided that the signal to noise ratio (SNR) is larger than a `threshold' value (see the simulations in Section~\ref{sec:simulations} for examples).
 Often the threshold is observed to occur at smaller values of SNR as $N$ increases.
@@
 
-An analysis of the HAF estimator under the assumption that the noise $\{X_n\}$ is white and Gaussian and without the `small noise' assumption has been made in~\cite{Porat_asympt_HAF_DPT_1996} (see in particular the proof of Lemma~1 and the remark that follows).
+An analysis of the HAF estimator under the assumption that the noise $\{X_n\}$ is white and Gaussian and without the `small noise' assumption has been conducted in~\cite{Porat_asympt_HAF_DPT_1996} (see in particular the proof of Lemma~1 and the remark that follows).
 To date, the literature does not contain a corresponding analysis of any estimator based on phase unwrapping.
@@
 For example, in the case that $\{X_n\}$ is white and Gaussian, our results show that the asymptotic variance of the LSU estimator converges to the Cramer-Rao lower bound (CRB) as the SNR becomes large, and converges to $\tfrac{2}{3}\pi$ times the CRB as the SNR approaches zero.
-Related results have been presented previously~\cite{Quinn2009_dasp_phase_only_information_loss}.
+Related results have been presented in~\cite{Quinn2009_dasp_phase_only_information_loss}.
 It is worth noting that this factor of $\tfrac{2}{3}\pi$ times the CRB at low SNR is better than popular estimators such as the HAF and CPF, for which the ratio between the asymptotic variance and the CRB diverges as the SNR approaches zero~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.
@@
 A complete proof of normality is available in~\cite{McKilliam_arxiv_pps_unwrapping_2012}.
-Section~\ref{sec:simulations} describes the results of Monte Carlo simulations that compare the LSU estimator with existing estimators including the HAF and CPF.
+Section~\ref{sec:simulations} describes the results of Monte Carlo simulations that compare the LSU estimator with existing estimators including the HAF, PHAF and CPF.
 These simulations agree with the derived asymptotic properties.
@@
 
-This section describes the results of Monte-Carlo simulations with the least squares unwrapping (LSU) estimator, the high order ambiguity function (HAF) estimator~\cite{Peleg_DPT_1995}, the cubic phase function (CPF) estimator~\cite{Oshea_cpf_2004} and the estimator that results from combining the HAF and CPF~\cite{Djurovic_haf_cpf_2012}.
+This section describes the results of Monte-Carlo simulations with the least squares unwrapping (LSU) estimator, the high order ambiguity function (HAF) estimator~\cite{Peleg_DPT_1995}, the product high order ambiguity function (PHAF) estimator~\cite{Barbarossa_PHAF_1998}, the cubic phase function (CPF) estimator~\cite{Oshea_cpf_2004} and the estimator that results from combining the HAF and CPF~\cite{Djurovic_haf_cpf_2012}.
 The CPF estimator only applies to polynomial phase signals of order $m=3$ and when the number of observations $N$ is odd.
-In all our simulations the unknown amplitude is $\rho = 1$ and $N = 199$.
+In all simulations the unknown amplitude is $\rho = 1$ and $N = 199$.
 The $X_1, \dots, X_N$ are pseudorandomly generated independent and identically distributed circularly symmetric complex Gaussian with variance $\expect\abs{X_1}^2 = \sigma_c^2$.
@@
 The sample mean square error (MSE) of the $k$th coefficient is computed according to $\tfrac{1}{T}\sum_{t=1}^T \widehat{\lambda}_{k,t}^2$ where $\widehat{\lambda}_{k,t}$ is the $k$th element of $\widehat{\lambdabf}_t$.
-In all simulations the LSU estimator is approximated using the $K$-best method described in Section~\ref{sec:lattice-theory}.
-The maximum number of branches allowed is set to $K=20*N$.
+In all simulations four \emph{lag sets} are used for the PHAF estimator~\cite{Barbarossa_PHAF_1998} and the LSU estimator is approximated using the $K$-best method described in Section~\ref{sec:lattice-theory}.
+The maximum number of branches allowed is $K=20N$.
 
@@
 Using Theorem~\ref{thm:asymp_proof} we can predict the ratio between the CRB and the asymptotic variance of the LSU estimator to approach 1 as SNR approaches $\infty$ and $\tfrac{2}{3}\pi$ as SNR approaches zero~(see Lemma~3 and the subsequent discussion in~\cite{Quinn2009_dasp_phase_only_information_loss}).
-Contrastingly, the ratio between the CRB and asymptotic variance of the HAF and CPF estimators is known to diverge as SNR goes to zero~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.
+Contrastingly, the ratio between the CRB and asymptotic variance of the HAF, PHAF and CPF estimators is known to diverge as SNR goes to zero~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.
 %For this reason we can expect the LSU estimator to be more accurate at low SNR for sufficiently large $N$.
 
-In Figure~\ref{plot:m3small} the CPF estimator reaches a threshold at \unit[1]{dB}, the LSU estimator at \unit[3]{dB} and the HAF estimator at \unit[5]{dB}.
+In Figure~\ref{plot:m3small} the CPF estimator reaches a threshold at \unit[1]{dB}, the LSU and PHAF estimators at \unit[3]{dB} and the HAF estimator at \unit[5]{dB}.
 In Figure~\ref{plot:m3small} the true coefficients are $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}  \right]$.
-These are specifically chosen to satisfy the requirements of the HAF and CPF estimators.
-The HAF estimator requires that the coefficients satisfy $\tilde{\mu}_0, \tilde{\mu}_1 \in [-0.5, 0.5)$ and
+These are specifically chosen to satisfy the requirements of the HAF, PHAF and CPF estimators.
+The HAF and PHAF estimators require that the coefficients satisfy $\tilde{\mu}_0, \tilde{\mu}_1 \in [-0.5, 0.5)$ and
 \[
@@
 \]
-where $\tau$ is the \emph{lag} and we choose $\tau = \floor{N/m}$ as suggested in~\cite{Peleg_DPT_1995}.
+where $\tau$ is the \emph{lag} and we choose $\tau = \floor{N/m}$ as suggested in~\cite{Peleg_DPT_1995,Barbarossa_PHAF_1998}.
 The CPF estimator requires similar, but slightly more complicated conditions on the coefficients~\cite{Oshea_cpf_2004}.
-It is informative to compare the volume of the set of coefficients suitable for the HAF and CPF estimators with the volume of the identifiable region $B$ from~\eqref{eq:rectangular_identifiable_region}.
+It is informative to compare the volume of the set of coefficients suitable for the HAF, PHAF and CPF estimators with the volume of the identifiable region $B$ from~\eqref{eq:rectangular_identifiable_region}.
 The identifiable region has volume
@@
 \]
-The volume of coefficients suitable for the HAF estimator is
+The volume of coefficients suitable for the HAF and PHAF estimators is
 \[
@@
 Substituting $m=3$ and $N=199$ in the formula above, we see that the HAF and CPF estimators search a space of parameters that is almost $5 \times 10^5$ times smaller than that searched by the LSU estimator.
-In this context we might expect the CPF and HAF estimators to have thresholds at lower SNR than LSU estimator.
+In this context we might expect the HAF, PHAF and CPF estimators to have thresholds at lower SNR than LSU estimator.
 It is encouraging that the LSU estimator has comparable performance.
@@
 
-The comparatively small set of coefficients suitable for the HAF and CPF estimators has been studied previously~\cite{Angeby_PPS_aliasing_2000,Djurovic_aliasing_2012,McKilliam2009IndentifiabliltyAliasingPolyphase,Zhou_lag_diverity_PPT_1997,Xia_dynamic_range_pps_2001} and can be a critical limitation for some applications, for example, interference suppression in the presence of a jammer~\cite{Djukanovic:2011:PMN:1950994.1951252,Djurovic_aliasing_2012}.
+The comparatively small set of coefficients suitable for the HAF, PHAF and CPF estimators has been studied previously~\cite{Angeby_PPS_aliasing_2000,Djurovic_aliasing_2012,McKilliam2009IndentifiabliltyAliasingPolyphase,Zhou_lag_diverity_PPT_1997,Xia_dynamic_range_pps_2001,Morelande_bayes_unwrapping_2009_tsp} and can be a critical limitation for some applications, for example, interference suppression in the presence of a jammer~\cite{Djukanovic:2011:PMN:1950994.1951252,Djurovic_aliasing_2012}.
 Zhou and Wang~\cite{Zhou_lag_diverity_PPT_1997} describe a method based on the extended Euclidean algorithm that enables the HAF estimator to operate on the entire identifiable region $B$.
@@
 The threshold for the estimator is approximately~\unit[29]{dB}.
-There does not appear to exist (at least in the literature) a similar method for increasing the range of coefficients for the CPF estimator.
-
-Figures~\ref{plot:m5} displays the sample MSE of the highest order coefficient $\mu_5$  for a polynomial phase signal of order $m=5$.
-In this case the LSU estimator reaches a threshold at \unit[12]{dB}, the HAF estimator at \unit[18]{dB}, and the estimator that combines the CPF and HAF at \unit[10]{dB}.
-In this figure the true coefficients are $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}, \tfrac{1}{96N^3}, \tfrac{1}{480N^4}  \right]$ and these are chosen specifically to satisfy the requirement of the HAF and CPF-HAF estimators.
-As in the case where $m=3$, the region of coefficients suitable for the HAF and CPF-HAF estimators is much smaller that the identifiable region $B$.
+There does not appear to exist (at least in the literature) a similar method for increasing the range of coefficients for the PHAF and CPF estimators.
+
+Figure~\ref{plot:m5} displays the sample MSE of the highest order coefficient $\mu_5$  for a polynomial phase signal of order $m=5$.
+In this case the LSU and PHAF estimators reach a threshold at \unit[12]{dB}, the HAF estimator at \unit[18]{dB}, and the estimator that combines the CPF and HAF at \unit[10]{dB}.
+In this figure the true coefficients are $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}, \tfrac{1}{96N^3}, \tfrac{1}{480N^4}  \right]$ and these are chosen specifically to satisfy the requirement of the HAF, PHAF and CPF-HAF estimators.
+As in the case where $m=3$, the region of coefficients suitable for the HAF, PHAF and CPF-HAF estimators is much smaller that the identifiable region.
 The effect becomes more severe as $m$ increases.
-When $m=5$ and $N=199$ the space of coefficients searched by the HAF and CPF-HAF estimators is approximately $10^{16}$ times smaller than the identifiable region.
-This constitutes a dramatic reduction in the range of detectable polynomial phase signals.
-Figure~\ref{plot:m3big} displays results under the same conditions but with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}, \tfrac{1}{96}, \tfrac{1}{480}  \right]$.
+When $m=5$ and $N=199$ the space of coefficients searched these estimators is approximately $10^{16}$ times smaller than the identifiable region.
+This is a dramatic reduction in the range of detectable polynomial phase signals.
+Figure~\ref{plot:m5big} displays results under the same conditions as Figure~\ref{plot:m5} but with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}, \tfrac{1}{96}, \tfrac{1}{480}  \right]$.
 The conclusion is similar to that from Figure~\ref{plot:m3big}.
-The performance of the LSU estimator remains unchanged, but the HAF and CPF-HAF estimators fail completely.
+The performance of the LSU estimator remains unchanged, but the HAF, PHAF and CPF-HAF estimators fail completely.
+The Zhou and Wang estimator can again be used to increase the range of parameters for HAF, but simulations indicate that the threshold for the estimator is approximately~\unit[103]{dB}.
 
@@
   		\caption{Sample MSE for the highest order coefficient $\tilde{\mu}_5$ for a polynomial phase signal of order 5 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}, \tfrac{1}{96}, \tfrac{1}{480}   \right]$.}
-  		\label{plot:m5} 
+  		\label{plot:m5big} 
 \end{figure}

### 2013-02-27 code/java/src/pubsim/poly/HAF.java

@@ -106,7 +106,7 @@ public class HAF extends AbstractPolynomialPhaseEstimator {
 
         //note that the FFT is generally defined with exp(-jw) but
         //periodogram has exp(jw) so freq are -ve here.
-        double maxp = 0;
+        double maxp = 0.0;
         double fhat = 0.0;
         double f = 0.0;
         double fstep = 1.0 / ft.length;

### 2013-02-27 code/montecarlo.scala

@@ -31,21 +31,21 @@ val bp5 = (0 to 5).map( k => 0.25/pubsim.Util.factorial(k) ).toArray //5th order
 
 //lags for the PHAF estimator
 val tau3 = Array(
-Array((N+1)/3,(N+1)/3),
-Array((N+1)/3-4,(N+1)/3+4),
-Array((N+1)/3-14,(N+1)/3+14)
-//Array((N+1)/3-7,(N+1)/3+5)
+Array(66,66),
+Array(44,99),
+Array(36,121),
+Array(33,132)
 )
-
 val tau5 = Array(
-Array((N+1)/5,(N+1)/5,(N+1)/5,(N+1)/5),
-Array((N+1)/5-4,(N+1)/5+4,(N+1)/5-4,(N+1)/5+4),
-Array((N+1)/5-14,(N+1)/5+14,(N+1)/5-14,(N+1)/5+14)
+Array(40,40,40,40),
+Array(40,40,32,50),
+Array(32,50,25,64),
+Array(25,64,20,80)
 )
 
 val starttime = (new java.util.Date).getTime
 
-/*
+
 runsim(-5 to 15, sp3, N, iters, () => new HAF(3,N), "HAFm3small")
 runsim(-5 to 15, bp3, N, iters, () => new HAF(3,N), "HAFm3big")
 runsim(-5 to 15, sp3, N, iters, () => new CPF(N), "CPFm3small")
@@ -63,22 +63,24 @@ runsim(-5 to 35 by 2, bp3, N, iters, () => new Mbest(3,N, 20*N, new HKZ()), "Mbe
 runcrb(-5 to 35, 3, N, "crbm3range")
 runlsuclt(-5 to 35, 3, N, "lsucltm3range")
 
-runsim(5 to 30, sp5, N, iters, () => new HAF(5,N), "HAFm5small")
-runsim(5 to 30, sp5, N, iters, () => new CPFHAF(5,N), "CPFHAFm5small")
-//runsim(5 to 30, sp5, N, iters, () => new Babai(5,N, new LLL()), "Babaim5small")
-runsim(5 to 30, sp5, N, iters, () => new Mbest(5,N, 20*N, new LLL()), "Mbestm5small")
-runsim(5 to 30, bp5, N, iters, () => new HAF(5,N), "HAFm5big")
-runsim(5 to 30, bp5, N, iters, () => new CPFHAF(5,N), "CPFHAFm5big")
[…]

### 2013-02-27 paper.tex
minor

@@
 It is now worth reconsidering the results in Figure~\ref{plot:m3small}.
-Substituting $m=3$ and $N=199$ in the formula above, we see that the HAF and CPF estimators search a space of parameters that is almost $5 \times 10^5$ times smaller than that searched by the LSU estimator.
+Substituting $m=3$ and $N=199$ in the formula above, we see that the HAF, PHAF and CPF estimators search a space of parameters that is more than $10^5$ times smaller than that searched by the LSU estimator.
 In this context we might expect the HAF, PHAF and CPF estimators to have thresholds at lower SNR than LSU estimator.
@@
 To put these results further into context, Figure~\ref{plot:m3big} repeats the simulations in Figure~\ref{plot:m3small}, but where the true coefficients are $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$.
-The CPF and HAF estimators fail spectacularly.
+The HAF, PHAF and CPF estimators fail spectacularly.
 The performance of the LSU estimator is unchanged, as expected in view of Theorem~\ref{thm:asymp_proof}.
@@
 The performance of the LSU estimator remains unchanged, but the HAF, PHAF and CPF-HAF estimators fail completely.
-The Zhou and Wang estimator can again be used to increase the range of parameters for HAF, but simulations indicate that the threshold for the estimator is approximately~\unit[103]{dB}.
+The Zhou and Wang estimator can again be used to increase the range of parameters for the HAF, but simulations indicate that the threshold for the estimator is approximately~\unit[103]{dB}.
 

### 2013-02-28 paper.tex

@@
 This assumption is used to justify what is called a `perturbation analysis'~%\cite{Porat_asympt_HAF_DPT_1996,Peleg_linear_FM,Oshea_cpf_2004,Djurovic_haf_cpf_2012,Djuric_phase_unwrap_chirp_1990,Barbarossa_analysis_of_PPS_1997,Barbarossa_PHAF_1998}
-to obtain expressions for the asymptotic mean square error of the estimator in question.
-The `small noise' assumption is usually supported by Monte-Carlo simulations showing the asymptotic mean square error obtained by the perturbation analysis to be accurate provided that the signal to noise ratio (SNR) is larger than a `threshold' value (see the simulations in Section~\ref{sec:simulations} for examples).
+to obtain expressions for the mean square error of the estimator in question.
+The `small noise' assumption is usually supported by Monte-Carlo simulations showing the mean square error predicted by the perturbation analysis to be accurate provided that the signal to noise ratio (SNR) is larger than a `threshold' value (see the simulations in Section~\ref{sec:simulations} for examples).
 Often the threshold is observed to occur at smaller values of SNR as $N$ increases.
@@
 This can be resolved by computing $\dealias(\widehat{\mubf})$.
+%The optimisation problem corresponding with~\eqref{eq:sumofsquaresfunction} for the LSU estimator involves minimising over $m+1$ real parameters.
+Computing a nearest lattice point in the $N-m-1$ dimensional lattice $\Lambda$ involves a minimisation over $N-m-1$ integers.
+It may at first seem that we have made the problem~\emph{harder} by converting it into a nearest lattice point problem.
+However, there is a substantial body of work dedicated to the development of algorithms for computing a nearest lattice point~\cite{Agrell2002,schnorr_euchner_sd_1994,Viterbo_sphere_decoder_1999}, and we exploit this for the purpose of computing the LSU estimator (see Section~\ref{sec:lattice-theory}).
 There exist algorithms that can compute a nearest point in $\Lambda$ that require a number of operations that is polynomial in $N$~\cite[Sec 4.3]{McKilliam2010thesis}.
@@
 On the other hand, when the SNR is small, all branches in the tree tend to be searched.
-Interestingly the LSU estimator computed using the $K$-best algorithm is \emph{faster} than the HAF when the SNR is large and when $N$ is approximately less than 100.
+Interestingly the LSU estimator computed using the $K$-best algorithm can be \emph{faster} than the HAF when the SNR is large and when $N$ is approximately less than 100.
 The LSU estimator can be computed in a reasonable amount of time (a few minutes) when $N=1000$.

### 2013-03-14 paper.tex
added Vaughan's comments

@@
 B.
-Moran is with the Department of Electrical Engineering and Computer
-Science, Melbourne Systems Lab, Dept of Elec \& Electronic Eng, Uni of Melbourne, Vic. 3010, Australia.}}
+Moran is with the Defense Science Institute, University of Melbourne, Vic.
+3010, Australia.}}
 % The paper headers 
@@
 A key feature is that the LSU estimator functions accurately over a far wider range of parameters than many popular existing estimators.
-Monte-Carlo simulations assert the excellent statistical performance of the LSU estimator when compared to existing state-of-the-art estimators and support our theoretical results.
+Monte-Carlo simulations support our theoretical results and demonstrate the excellent statistical performance of the LSU estimator when compared to existing state-of-the-art estimators.
 \end{abstract}
@@
 \]
-where $j = \sqrt{-1}$, and $t$ is a real number and 
+where $j = \sqrt{-1}$, $t$ is a real number usually representing time and 
 \[
@@
 is a polynomial of order $m$.
-In practice the signal is typically sampled at discrete points in `time', $t$.
+In practice the signal is typically sampled at discrete points in time $t$.
 In this paper we only consider uniform sampling, where the gap between consecutive samples is constant.
@@
 where $\rho$ is a positive real number representing the (usually unknown) signal amplitude and $\{X_n, n \in \ints\}$ is a sequence of complex noise variables.
-In order to ensure identifiability it is necessary to restrict the $m+1$ coefficients to a region of $m+1$ dimensional Euclidean space $\reals^{m+1}$ called an \emph{identifiable region}.
-It was shown in \cite{McKilliam2009IndentifiabliltyAliasingPolyphase} that an identifiable region tessellates a particular $m+1$ dimensional lattice.
+In order to ensure identifiability it is necessary to restrict the $m+1$ coefficients to a region of $m+1$ dimensional Euclidean space $\reals^{m+1}$ called an \emph{identifiable region}~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}.
+%It was shown in \cite{McKilliam2009IndentifiabliltyAliasingPolyphase} that an identifiable region tessellates a particular $m+1$ dimensional lattice.
 We discuss this further in Section \ref{sec:ident_aliasing}.
@@
 Babai's algorithm is computationally simple, but its statistical performance was observed to be comparatively poor at low signal to noise ratio (SNR).
-A major point of interest is that the lattices considered have significant structure, and may admit fast nearest point algorithms.
+Of interest is that the lattices considered have significant structure, and may admit fast nearest point algorithms.
 This has been studied in~\cite{McKilliam2010thesis} where polynomial time algorithms were found that compute the nearest point exactly.
@@
 In addition to the above computational results a major purpose of this paper is to study the asymptotic properties of the LSU estimator.
-Under mild assumptions about the distribution of the noise we prove that the LSU estimator is strongly consistent, that is, the estimated coefficients converge almost surely to the true coefficients as the number of observations $N$ grows.
+Under mild assumptions about the distribution of the noise we prove that the LSU estimator is strongly consistent, that is, the estimated coefficients converge almost surely to the true coefficients as the number of observations, $N$, grows.
 %Similar results were stated without proof in~\cite{McKilliam_polyphase_est_icassp_2011}.
@@
 to obtain expressions for the mean square error of the estimator in question.
-The `small noise' assumption is usually supported by Monte-Carlo simulations showing the mean square error predicted by the perturbation analysis to be accurate provided that the signal to noise ratio (SNR) is larger than a `threshold' value (see the simulations in Section~\ref{sec:simulations} for examples).
+The small noise assumption is usually supported by Monte-Carlo simulations showing the mean square error predicted by the perturbation analysis to be accurate provided that the signal to noise ratio (SNR) is larger than a `threshold' value (see the simulations in Section~\ref{sec:simulations} for examples).
 Often the threshold is observed to occur at smaller values of SNR as $N$ increases.
-A problem with the `small noise' assumption is that it is not known that the threshold continues to occur at increasingly smaller values of SNR as $N$ increases, or if it instead converges to some finite (but perhaps small) value of SNR.
-The consequence of the later being that the estimator will be inaccurate if the SNR is smaller than this finite value, regardless of the number of observations $N$.
-Our analysis of the LSU estimator requires no `small noise' assumption and guarantees that no such finite value of SNR exists for the LSU estimator.
+A problem with the small noise assumption is that it is not known if the threshold can be made to occur at arbitrarily small values of SNR by increasing $N$, or if it instead converges to some finite (but perhaps small) value of SNR as $N$ increases.
+The consequence of the later is that the estimator will be inaccurate if the SNR is smaller than this finite value, regardless of the number of observations $N$.
+Our analysis of the LSU estimator requires no small noise assumption and guarantees that no such finite value of SNR exists for the LSU estimator.
 As a result the LSU estimator can be accurate for signals with arbitrarily small SNR provided that the number of observations is sufficiently large.
 
-An analysis of the HAF estimator under the assumption that the noise $\{X_n\}$ is white and Gaussian and without the `small noise' assumption has been conducted in~\cite{Porat_asympt_HAF_DPT_1996} (see in particular the proof of Lemma~1 and the remark that follows).
+An analysis of the HAF estimator under the assumption that the noise $\{X_n\}$ is white and Gaussian and without the small noise assumption has been conducted in~\cite{Porat_asympt_HAF_DPT_1996} (see Lemma~1 and the remark that follows in~\cite{Porat_asympt_HAF_DPT_1996}).
 To date, the literature does not contain a corresponding analysis of any estimator based on phase unwrapping.
-This has apparently lead to the misconception that phase unwrapping estimators are only accurate when the noise is `small' in some sense.
-This misconception would appear to be motivated by early papers such as~\cite{Tretter1985,Kay1989} where it was observed that the noise occurring in the phase of the signal only looks `Gaussian' when the complex noise producing it has small variance.
+This has apparently led to the misconception that phase unwrapping estimators are only accurate when the noise is `small' in some sense.
+This misconception may have been motivated by early papers such as~\cite{Tretter1985,Kay1989} where it was observed that the noise occurring in the phase of the signal only looks `Gaussian' when the complex noise producing it has small variance.
 The results in this paper make it clear that phase unwrapping estimators can be accurate at arbitrarily low SNR.
-The information contained in the amplitude is not essential for accurate estimation of polynomial phase signals, and can be discarded with only small loss.
+The information contained in the amplitude is not essential for accurate estimation of polynomial phase signals, and can be discarded with only small loss in accuracy.
 For example, in the case that $\{X_n\}$ is white and Gaussian, our results show that the asymptotic variance of the LSU estimator converges to the Cramer-Rao lower bound (CRB) as the SNR becomes large, and converges to $\tfrac{2}{3}\pi$ times the CRB as the SNR approaches zero.
 Related results have been presented in~\cite{Quinn2009_dasp_phase_only_information_loss}.
-It is worth noting that this factor of $\tfrac{2}{3}\pi$ times the CRB at low SNR is better than popular estimators such as the HAF and CPF, for which the ratio between the asymptotic variance and the CRB diverges as the SNR approaches zero~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.
+This factor of $\tfrac{2}{3}\pi$ times the CRB at low SNR is better than popular estimators such as the HAF and CPF, for which the ratio between the asymptotic variance and the CRB diverges as the SNR approaches zero~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.
 
@@
 This property has been studied previously~\cite{Djurovic_aliasing_2012,McKilliam2009IndentifiabliltyAliasingPolyphase,Zhou_lag_diverity_PPT_1997,Xia_dynamic_range_pps_2001} and can be a critical limitation for some applications, for example, interference suppression in the presence of a jammer~\cite{Djukanovic:2011:PMN:1950994.1951252,Djurovic_aliasing_2012}.
-Approaches exist that extend the range of parameters suitable for the HAF~\cite{Zhou_lag_diverity_PPT_1997,Djurovic_aliasing_2012}, but the resulting estimators typically have poor statistical performance.
-For this reason, the LSU estimator may be the only presently existing estimator that is viable for certain applications, such as interference suppression.
+Approaches are known that extend the range of parameters suitable for the HAF~\cite{Zhou_lag_diverity_PPT_1997,Djurovic_aliasing_2012}, but the resulting estimators typically have poor statistical performance.
+For this reason, the LSU estimator may be the only known estimator that is viable for certain applications, such as interference suppression.
 The theoretical results given in Section~\ref{sec:least-squar-unwr} prove that the LSU estimator functions correctly over the entire identifiable region.
@@
 Section~\ref{sec:lattice-theory} describes some required concepts from lattice theory.
-In Section~\ref{sec:ident_aliasing} we describe the identifiable region that was also derived in~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}.
+In Section~\ref{sec:ident_aliasing} we describe the identifiable region that was derived in~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}.
 These identifiability results are required in order to properly understand the statistical properties of the LSU estimator.
@@
 The nearest lattice point problem is: Given $\ybf\in\reals^n$ and some lattice $\Lambda \subset \reals^n$, find a lattice point $\xbf \in \Lambda$ such that the Euclidean distance between $\ybf$ and $\xbf$ is minimised, that is, solve $\arg\min_{\xbf \in \Lambda} \|\ybf - \xbf\|^2.$  The nearest lattice point problem is known to be NP-hard under certain conditions when the lattice itself is considered as an additional input parameter \cite{micciancio_hardness_2001, AjtaiShortestVecProbNPHard1998}.
-On the other hand,  fast algorithms that compute a nearest point often exist for specific lattices with known structure~\cite{SPLAG,Conway1982FastQuantDec,McKilliam2009CoxeterLattices,McKilliam2008}.
-Even in the general case, where the lattice has no known structure, algorithms exist that can compute a nearest lattice point in reasonable time if the dimension is small~\cite{Agrell2002, Viterbo_sphere_decoder_1999, Pohst_sphere_decoder_1981}.
+On the other hand, fast algorithms exist for specific lattices~\cite{SPLAG,Conway1982FastQuantDec,McKilliam2009CoxeterLattices,McKilliam2008}.
+Even in the general case, where the lattice has no known structure, there are algorithms that compute a nearest lattice point in reasonable time if the dimension is small~\cite{Agrell2002, Viterbo_sphere_decoder_1999, Pohst_sphere_decoder_1981}.
 These algorithms typically operate using a tree search to enumerate all lattice points inside a ball of radius sufficiently large to contain a nearest point.
@@
 We refer to this algorithm as the~\emph{sphere decoder} in this paper, although other variants of the algorithm also go by this name.
-
 With modern computers, the sphere decoder is computationally viable if the dimension of the lattice is less than approximately $60$.
+
 Algorithms that approximate a nearest point have also been studied.
@@
 Following~\cite{Zhan2006_K_best_sphere_decoder} we refer to this as \emph{$K$-best algorithm}.
-We will find that this algorithm produces excellent estimates of polynomial phase coefficients and can be computed in a reasonable amount of time for moderate numbers of observations ($N$ less that about 1000).
+We will find that this algorithm produces excellent estimates of polynomial phase coefficients and can be computed in a reasonable amount of time for moderate numbers of observations ($N$ less than about 1000).
 
@@
 
-As discussed in the introduction, a polynomial phase signal of order $m$ is a complex valued function of the form $s(t) = e^{2\pi j y(t)}$ where $t$ is a real number and $y(t)$ is a polynomial of order $m$.
+A polynomial phase signal of order $m$ is a complex valued function of the form $s(t) = e^{2\pi j y(t)}$ where $t$ is a real number and $y(t)$ is a polynomial of order $m$.
 We will often drop the $(t)$ and just write the polynomial as $y$ and the polynomial phase signal as $s$ whenever there is no chance of ambiguity.
@@
 The proof here takes a similar approach, but requires new techniques.
-Before giving the proof of strong consistency we make the following remarks.
-
-The theorem gives conditions on the \emph{dealiased} difference $\dealias(\tilde{\mubf} - \widehat{\mubf})$ between the true coefficients $\tilde{\mubf}$ and the estimated coefficients $\widehat{\mubf}$ rather than directly on the difference $\tilde{\mubf} - \widehat{\mubf}$.
+%Before giving the proof of strong consistency we make the following remarks.
+
+The theorem makes statements about the \emph{dealiased} difference $\dealias(\tilde{\mubf} - \widehat{\mubf})$ between the true coefficients $\tilde{\mubf}$ and the estimated coefficients $\widehat{\mubf}$ rather than directly on the difference $\tilde{\mubf} - \widehat{\mubf}$.
 To see why this makes sense, consider the case when $m=0$, $\tilde{\mu}_0 = -0.5$ and $\widehat{\mu}_0 = 0.49$, so that $\tilde{\mu}_0 - \widehat{\mu}_0 = -0.99$.
@@
 To put these results further into context, Figure~\ref{plot:m3big} repeats the simulations in Figure~\ref{plot:m3small}, but where the true coefficients are $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$.
-The HAF, PHAF and CPF estimators fail spectacularly.
+The HAF, PHAF and CPF estimators fail consistently.
 The performance of the LSU estimator is unchanged, as expected in view of Theorem~\ref{thm:asymp_proof}.
@@
   		\includegraphics{code/data/plot3-1.mps} 
-  		\caption{Sample MSE for highest order coefficient $\tilde{\mu}_3$ for a polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}  \right]$.}
+  		\caption{Sample MSE for highest order coefficient $\tilde{\mu}_3$ for a polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}  \right]$ and $N=199$.}
   		\label{plot:m3small} 
@@
   		\includegraphics{code/data/plot3-2.mps} 
-  		\caption{Sample MSE for the highest order coefficient $\tilde{\mu}_3$ for a polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$.}
+  		\caption{Sample MSE for the highest order coefficient $\tilde{\mu}_3$ for a polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$ and $N=199$.}
   		\label{plot:m3big} 
@@
   		\includegraphics{code/data/plot3-3.mps} 
-  		\caption{Sample MSE for the highest order coefficient $\tilde{\mu}_3$ for a polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$.}
+  		\caption{Sample MSE for the highest order coefficient $\tilde{\mu}_3$ for a polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$ and $N=199$.}
   		\label{plot:m3bigzw} 
@@
   		\includegraphics{code/data/plot5-1.mps} 
-  		\caption{Sample MSE for the highest order coefficient $\tilde{\mu}_5$ for a polynomial phase signal of order 5 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}, \tfrac{1}{96N^3}, \tfrac{1}{480N^4} \right]$.}
+  		\caption{Sample MSE for the highest order coefficient $\tilde{\mu}_5$ for a polynomial phase signal of order 5 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}, \tfrac{1}{96N^3}, \tfrac{1}{480N^4} \right]$ and $N=199$.}
   		\label{plot:m5} 
@@
   		\includegraphics{code/data/plot5-2.mps} 
-  		\caption{Sample MSE for the highest order coefficient $\tilde{\mu}_5$ for a polynomial phase signal of order 5 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}, \tfrac{1}{96}, \tfrac{1}{480}   \right]$.}
+  		\caption{Sample MSE for the highest order coefficient $\tilde{\mu}_5$ for a polynomial phase signal of order 5 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}, \tfrac{1}{96}, \tfrac{1}{480}   \right]$ and $N=199$.}
   		\label{plot:m5big} 
@@
 Polynomial time nearest point algorithms for these lattices exist~\cite[Sec 4.3]{McKilliam2010thesis}, but these algorithms are not fast in practice.
-We instead make used on the approximate $K$-best algorithm~\cite{Baro_list_sphere_dec_2003,Zhan2006_K_best_sphere_decoder,Mondal_Kbest_again_2009}.
+We instead make use of the approximate $K$-best algorithm~\cite{Baro_list_sphere_dec_2003,Zhan2006_K_best_sphere_decoder,Mondal_Kbest_again_2009}.
 This algorithm results in an accurate estimator and can be computed in a reasonable amount of time for moderately large numbers of observations $N$.

### 2013-04-04 paper.tex
added to paper, barry's edits

@@
 Under mild conditions on the distribution of the noise we describe the asymptotic properties of this estimator, showing that it is strongly consistent and asymptotically normally distributed.
-No `small noise' assumption (common in the literature) is required in our analysis.
-As a result the LSU estimator will be accurate, even for very noisy signals, provided that the number of observations is sufficiently large.
-A key feature is that the LSU estimator functions accurately over a far wider range of parameters than many popular existing estimators.
-Monte-Carlo simulations support our theoretical results and demonstrate the excellent statistical performance of the LSU estimator when compared to existing state-of-the-art estimators.
+No `high SNR' assumption (common in the literature) is required in our analysis.
+As a result the LSU estimator is accurate, even for very noisy signals, if the number of observations is sufficiently large.
+A key feature is that the LSU estimator is accurate over a far wider range of parameters than many popular existing estimators.
+Monte-Carlo simulations support our theoretical results and demonstrate the excellent statistical performance of the LSU estimator when compared with existing state-of-the-art estimators.
 \end{abstract}
@@
 
-Polynomial phase signals arise in fields including radar, sonar, geophysics, biology, and radio communication~\cite{Hlawatsch_lin_quad_time_freq_spmag_1992,Ridleyspeechpolyphase1989, Suga_1975_bats_echolocation, Moss_2005echolocation,Angeby_estimating_2000}.
+Polynomial phase signals arise in fields including radar, sonar, geophysics, biology, and radio communication~\cite{Hlawatsch_lin_quad_time_freq_spmag_1992,Ridleyspeechpolyphase1989, Suga_1975_bats_echolocation,Angeby_estimating_2000}.
 %In radar and sonar applications polynomial phase signals arise when acquiring radial velocity and acceleration (and higher order motion descriptors) of a target from a reflected signal, and also in continuous wave radar and low probability of intercept radar~\cite{Levanon_Radar_signals_2004}.
@@
 \]
-where $j = \sqrt{-1}$, $t$ is a real number usually representing time and 
+where $t$ is a real number usually representing time and 
 \[
@@
 \]
-is a polynomial of order $m$.
-In practice the signal is typically sampled at discrete points in time $t$.
-In this paper we only consider uniform sampling, where the gap between consecutive samples is constant.
-In this case we can always consider the samples to be taken at some set of consecutive integers and our sampled polynomial phase signal takes the form
-\[
-s_n = s(n) = e^{2\pi j y(n)},
-\] 
-where $n$ is an integer.  Of practical importance is the estimation of the coefficients $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ from a number, say $N$, of observations of the noisy sampled signal
+is a polynomial of order $m$.  In this paper we assume that the signal is sampled at the integers and the sampled polynomial phase signal takes the form
+\[
+s_n = s(n) = e^{2\pi j y(n)} \qquad n\in\ints.
+\] 
+Of practical importance is the estimation of the coefficients $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ from a number, say $N$, of observations of the noisy sampled signal
 \begin{equation}\label{eq:Y_nsamplednoisey}
@@
 \end{equation}
-where $\rho$ is a positive real number representing the (usually unknown) signal amplitude and $\{X_n, n \in \ints\}$ is a sequence of complex noise variables.
+where $\rho$ is a positive number representing the (usually unknown) signal amplitude and $\{X_n, n \in \ints\}$ is a sequence of complex random variables representing noise.
 In order to ensure identifiability it is necessary to restrict the $m+1$ coefficients to a region of $m+1$ dimensional Euclidean space $\reals^{m+1}$ called an \emph{identifiable region}~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}.
@@
 
-An obvious estimator of the unknown coefficients is the least squares estimator (LSE).
-This is also the maximum likelihood estimator (MLE) when the noise sequence $\{X_n\}$ is white and Gaussian.
-When $m=0$ (phase estimation) or $m=1$ (frequency estimation) the LSE is an effective approach, being both computationally efficient and statistically accurate \cite{Quinn2009_dasp_phase_only_information_loss,Hannan1973,Quinn2001,McKilliam_mean_dir_est_sq_arc_length2010,McKilliam2010thesis}.
-When $m \geq 2$ the computational complexity of the LSE is large~\cite{Oshea_cpf_2004,McKilliam2010thesis,Abatzoglou_ml_chirp_1986}.
+An obvious estimator of the unknown coefficients is the least squares estimator.
+This is also the maximum likelihood estimator when the noise sequence $\{X_n\}$ is white and Gaussian.
+When $m=0$ (phase estimation) or $m=1$ (frequency estimation) the least squares estimator is an effective approach, being both computationally efficient and statistically accurate \cite{Quinn2009_dasp_phase_only_information_loss,Hannan1973,Quinn2001,McKilliam_mean_dir_est_sq_arc_length2010,McKilliam2010thesis}.
+When $m \geq 2$ the computational complexity of the least squares estimator is large~\cite{Oshea_cpf_2004,McKilliam2010thesis,Abatzoglou_ml_chirp_1986}.
 For this reason many authors have considered alternative approaches to polynomial phase estimation.
@@
 A drawback of the LSU estimator is that computing a nearest lattice point is, in general, computationally difficult.
-In \cite{McKilliam2009asilomar_polyest_lattice}, two standard techniques were considered, the \emph{sphere decoder}~\cite{Pohst_sphere_decoder_1981,Viterbo_sphere_decoder_1999,Agrell2002} that exactly computes a nearest lattice point, and \emph{Babai's nearest plane algorithm}~\cite{Babai1986} that only produces an approximation.
-The sphere decoder was observed to have excellent statistical performance but can only be computed efficiently for small $N$ less than about $60$.
-Babai's algorithm is computationally simple, but its statistical performance was observed to be comparatively poor at low signal to noise ratio (SNR).
+In~\cite{McKilliam2009asilomar_polyest_lattice}, two standard techniques were considered, the \emph{sphere decoder}~\cite{Pohst_sphere_decoder_1981,Viterbo_sphere_decoder_1999,Agrell2002} that exactly computes a nearest lattice point, and \emph{Babai's nearest plane algorithm}~\cite{Babai1986} that only produces an approximation.
+The sphere decoder was observed to have excellent statistical performance but can only be computed efficiently for $N$ less than about $60$.
+Babai's algorithm is computationally simple, but its statistical performance has been observed to be comparatively poor at low signal to noise ratio (SNR)~\cite{McKilliam2009asilomar_polyest_lattice}.
 Of interest is that the lattices considered have significant structure, and may admit fast nearest point algorithms.
@@
 
-The analysis of most existing polynomial phase estimators make what is known as a `small noise' assumption~\cite{Peleg_linear_FM,Oshea_cpf_2004,Djurovic_haf_cpf_2012,Djuric_phase_unwrap_chirp_1990,Barbarossa_analysis_of_PPS_1997,Barbarossa_PHAF_1998}.
+The analyses of most existing polynomial phase estimators make what is known as a `high SNR' assumption~\cite{Peleg_linear_FM,Oshea_cpf_2004,Djurovic_haf_cpf_2012,Djuric_phase_unwrap_chirp_1990,Barbarossa_analysis_of_PPS_1997,Barbarossa_PHAF_1998}.
 %This amounts to assuming that the estimator $\hat{\mu}_k$ of the $k$th coefficient is such that $N^{k}(\tilde{\mu}_k -  \hat{\mu}_k) \rightarrow 0$ (in some sense) as $N \rightarrow \infty$.
@@
 to obtain expressions for the mean square error of the estimator in question.
-The small noise assumption is usually supported by Monte-Carlo simulations showing the mean square error predicted by the perturbation analysis to be accurate provided that the signal to noise ratio (SNR) is larger than a `threshold' value (see the simulations in Section~\ref{sec:simulations} for examples).
+The high SNR assumption is usually supported by Monte-Carlo simulations showing the mean square error predicted by the perturbation analysis to be accurate provided that the signal to noise ratio (SNR) is larger than a `threshold' value (see the simulations in Section~\ref{sec:simulations} for examples).
 Often the threshold is observed to occur at smaller values of SNR as $N$ increases.
-A problem with the small noise assumption is that it is not known if the threshold can be made to occur at arbitrarily small values of SNR by increasing $N$, or if it instead converges to some finite (but perhaps small) value of SNR as $N$ increases.
-The consequence of the later is that the estimator will be inaccurate if the SNR is smaller than this finite value, regardless of the number of observations $N$.
-Our analysis of the LSU estimator requires no small noise assumption and guarantees that no such finite value of SNR exists for the LSU estimator.
-As a result the LSU estimator can be accurate for signals with arbitrarily small SNR provided that the number of observations is sufficiently large.
-
-An analysis of the HAF estimator under the assumption that the noise $\{X_n\}$ is white and Gaussian and without the small noise assumption has been conducted in~\cite{Porat_asympt_HAF_DPT_1996} (see Lemma~1 and the remark that follows in~\cite{Porat_asympt_HAF_DPT_1996}).
+A problem with the high SNR assumption is that it is not known if the threshold occurs at smaller SNR with increasing $N$, or if it instead converges to some finite (but perhaps small) value of SNR as $N$ increases.
+The consequence of the latter is that the estimator will be inaccurate if the SNR is smaller than this finite value, regardless of the number of observations $N$.
+Our analysis of the LSU estimator requires no high SNR assumption and guarantees that no such finite value of SNR exists for the LSU estimator.
+As a result the LSU estimator is accurate for signals with arbitrarily small SNR provided that the number of observations is sufficiently large.
+
+An analysis of the HAF estimator under the assumption that the noise process $\{X_n\}$ is white and Gaussian and without the high SNR assumption has been conducted in~\cite{Porat_asympt_HAF_DPT_1996} (see Lemma~1 and the remark that follows in~\cite{Porat_asympt_HAF_DPT_1996}).
 To date, the literature does not contain a corresponding analysis of any estimator based on phase unwrapping.
@@
 The information contained in the amplitude is not essential for accurate estimation of polynomial phase signals, and can be discarded with only small loss in accuracy.
-For example, in the case that $\{X_n\}$ is white and Gaussian, our results show that the asymptotic variance of the LSU estimator converges to the Cramer-Rao lower bound (CRB) as the SNR becomes large, and converges to $\tfrac{2}{3}\pi$ times the CRB as the SNR approaches zero.
+For example, in the case that $\{X_n\}$ is white and Gaussian, our results show that the asymptotic variance of the LSU estimator converges to the asymptotic Cramer-Rao lower bound (CRB) as the SNR becomes large, and converges to $\tfrac{2}{3}\pi$ times the asymptotic CRB as the SNR approaches zero.
 Related results have been presented in~\cite{Quinn2009_dasp_phase_only_information_loss}.
-This factor of $\tfrac{2}{3}\pi$ times the CRB at low SNR is better than popular estimators such as the HAF and CPF, for which the ratio between the asymptotic variance and the CRB diverges as the SNR approaches zero~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.
-
-A key property of the LSU estimator is that it functions correctly for polynomial phase parameters contained anywhere inside the identifiable region described in Section~\ref{sec:ident_aliasing} and~\cite{Xia_dynamic_range_pps_2001,McKilliam2009IndentifiabliltyAliasingPolyphase}.
+This factor of $\tfrac{2}{3}\pi$ times the CRB at low SNR is better than popular estimators such as the HAF and CPF, for which the ratio of the asymptotic variance and the asymptotic CRB diverges as the SNR approaches zero~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.
+
+A key property of the LSU estimator is that it works for polynomial phase parameters contained anywhere inside the identifiable region described in Section~\ref{sec:ident_aliasing} and~\cite{Xia_dynamic_range_pps_2001,McKilliam2009IndentifiabliltyAliasingPolyphase}.
 This is the maximum possible range of parameters.
-By contrast popular estimators such as the HAF and CPF operate correctly only for parameters in a range much smaller than the identifiable region.
+By contrast popular estimators such as the HAF and CPF work only for parameters in a range much smaller than the identifiable region.
 This property has been studied previously~\cite{Djurovic_aliasing_2012,McKilliam2009IndentifiabliltyAliasingPolyphase,Zhou_lag_diverity_PPT_1997,Xia_dynamic_range_pps_2001} and can be a critical limitation for some applications, for example, interference suppression in the presence of a jammer~\cite{Djukanovic:2011:PMN:1950994.1951252,Djurovic_aliasing_2012}.
@@
 For this reason, the LSU estimator may be the only known estimator that is viable for certain applications, such as interference suppression.
-The theoretical results given in Section~\ref{sec:least-squar-unwr} prove that the LSU estimator functions correctly over the entire identifiable region.
+The theoretical results given in Section~\ref{sec:least-squar-unwr} prove that the LSU estimator work over the entire identifiable region.
 To our knowledge this is the first time a result of this type has been obtained for any polynomial phase estimator.
@@
 In Section~\ref{sec:ident_aliasing} we describe the identifiable region that was derived in~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}.
-These identifiability results are required in order to properly understand the statistical properties of the LSU estimator.
+These identifiability results are required in order to understand the statistical properties of the LSU estimator.
 Section~\ref{sec:least-squar-unwr} describes the LSU estimator and states a theorem asserting the estimator to be strongly consistent and asymptotically normally distributed under some mild assumptions on the noise $\{X_n\}$.
@@
 Space restrictions make it impossible to give a proof of asymptotic normality in this paper.
-A complete proof of normality is available in~\cite{McKilliam_arxiv_pps_unwrapping_2012}.
+A complete proof of this is available in~\cite{McKilliam_arxiv_pps_unwrapping_2012}.
 Section~\ref{sec:simulations} describes the results of Monte Carlo simulations that compare the LSU estimator with existing estimators including the HAF, PHAF and CPF.
@@
 
-A \term{lattice},  $\Lambda$, is a discrete subset of points in $\reals^n$ such that
+A \term{lattice},  $\Lambda$, is a discrete subset of points in $\reals^n$ for which
 \[
@@
 \]
-where $\Bbf \in \reals^{n \times d}$ is an $n \times d$ matrix of rank $d$, called the generator matrix.
-If $n = d$ the lattice is said to be full rank.
+where $\Bbf \in \reals^{n \times d}$ has rank $d$ and is called the generator matrix.
+If $n = d$ the lattice is said to be of full rank.
 Lattices are discrete Abelian groups under vector addition.
@@
 Lattices naturally give rise to tessellations of $\reals^n$ by the specification of a set of coset representatives for the quotient $\reals^n / \Lambda$.
-One choice for a set of coset representatives is a fundamental parallelepiped; the parallelepiped generated by the columns of a generator matrix.
-Another choice is based on the Voronoi cell; those points from $\reals^n$ nearest (with respect to the Euclidean norm in this paper) to the lattice point at the origin.
+One choice for a set of coset representatives is a fundamental parallelepiped, the parallelepiped generated by the columns of a generator matrix.
+Another choice is based on the Voronoi cell, those points from $\reals^n$ nearest (with respect to the Euclidean norm in this paper) to the lattice point at the origin.
 It is always possible to construct a rectangular set of representatives, as the next proposition will show.
-We will use these rectangular regions for describing the aliasing properties of polynomial phase signals in Section~\ref{sec:ident_aliasing}.
+We will use these rectangular regions to describe the aliasing properties of polynomial phase signals in Section~\ref{sec:ident_aliasing}.
 These rectangular regions will be important for the derivation of the asymptotic properties of the LSU estimator.
@@
 
-A fundamental problem in lattice theory is the \emph{nearest lattice point problem}~\cite{Agrell2002}.
-The nearest lattice point problem is: Given $\ybf\in\reals^n$ and some lattice $\Lambda \subset \reals^n$, find a lattice point $\xbf \in \Lambda$ such that the Euclidean distance between $\ybf$ and $\xbf$ is minimised, that is, solve $\arg\min_{\xbf \in \Lambda} \|\ybf - \xbf\|^2.$  The nearest lattice point problem is known to be NP-hard under certain conditions when the lattice itself is considered as an additional input parameter \cite{micciancio_hardness_2001, AjtaiShortestVecProbNPHard1998}.
+A fundamental problem in lattice theory is the \emph{nearest lattice point problem}~\cite{Agrell2002}: Given $\ybf\in\reals^n$ and some lattice $\Lambda \subset \reals^n$, find a lattice point $\xbf \in \Lambda$ such that the Euclidean distance between $\ybf$ and $\xbf$ is minimised, that is, find $\arg\min_{\xbf \in \Lambda} \|\ybf - \xbf\|^2.$  The nearest lattice point problem is known to be NP-hard under certain conditions when the lattice itself is considered as an additional input parameter \cite{micciancio_hardness_2001}.
 On the other hand, fast algorithms exist for specific lattices~\cite{SPLAG,Conway1982FastQuantDec,McKilliam2009CoxeterLattices,McKilliam2008}.
@@
 The approaches vary in how branches are pruned.
-In our implementation we prune branches using the same greedy metric as that used by the Schnorr and Euchner sphere decoder~\cite{schnorr_euchner_sd_1994,Agrell2002} for deciding the order branches are searched.
-Following~\cite{Zhan2006_K_best_sphere_decoder} we refer to this as \emph{$K$-best algorithm}.
-We will find that this algorithm produces excellent estimates of polynomial phase coefficients and can be computed in a reasonable amount of time for moderate numbers of observations ($N$ less than about 1000).
+In our implementation we prune branches using the same greedy metric as that used by the Schnorr and Euchner sphere decoder~\cite{schnorr_euchner_sd_1994,Agrell2002} for deciding the order in which branches are searched.
+Following~\cite{Zhan2006_K_best_sphere_decoder} we refer to this as the \emph{$K$-best algorithm}.
+We have found that this algorithm produces excellent estimates of polynomial phase coefficients and can be computed in a reasonable amount of time for $N$ less than about 1000.
 
@@
 
-A polynomial phase signal of order $m$ is a complex valued function of the form $s(t) = e^{2\pi j y(t)}$ where $t$ is a real number and $y(t)$ is a polynomial of order $m$.
+%A polynomial phase signal of order $m$ is a complex valued function of the form $s(t) = e^{2\pi j y(t)}$ where $t$ is a real number and $y(t)$ is a polynomial of order $m$.
 We will often drop the $(t)$ and just write the polynomial as $y$ and the polynomial phase signal as $s$ whenever there is no chance of ambiguity.
@@
 %\]
-Aliasing can occur when polynomial-phase signals are sampled.
-That is, two or more distinct polynomial-phase signals can take exactly the same values at the sample points.
-These aliasing results are also given in~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}, but the presentation here is different, and is better suited to studying the asymptotic properties of the LSU estimator.
+Aliasing can occur when polynomial phase signals are sampled~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}.
+That is, two or more distinct polynomial phase signals can take exactly the same values at the sample points.
+%These aliasing results are also given in, but the presentation here is different, and is better suited to studying the asymptotic properties of the LSU estimator.
 %A description of the aliasing properties of polynomial phase signals of order $2$ has been independently discovered in~\cite{Abatzoglou_ml_chirp_1986}~and~\cite{Angeby_PPS_aliasing_2000}.
 The results here are for polynomial phase signals of any order.
-
 Let $\mathcal{Z}$ be the set of polynomials of order at most $m$ that take integer values when evaluated at integers.
@@
 The polynomial phase signals $s$ and $r$ are equal at the integers, and although they are distinct, they are indistinguishable from their samples.
-We call such polynomial phase signals \term{aliases}\index{alias} and immediately obtain the following theorem.
-
-\begin{theorem}\label{thm:circpolysampledthm}
+We call such polynomial phase signals \term{aliases}\index{alias} and immediately obtain the following lemma.
+
+\begin{lemma}\label{lem:circpolysampledthm}
 Two polynomial phase signals $s(t) = e^{2\pi j y(t)}$  and $r(t) = e^{2\pi j z(t)}$  are aliases if and only if the polynomials that define their phase, $y$ and $z$, differ by a polynomial from the set $\mathcal{Z}$, that is, $y - z \in \mathcal{Z}$.
-\end{theorem}
+\end{lemma}
 
@@
 
-It may be helpful to observe Figures~\ref{fig:circstatplot_line} and~\ref{fig:circstatplot_cube}.
-In these, the phase (divided by $2\pi$) of two distinct polynomial phase signals is plotted on the left, and on the right the principal component of the phase (also divided by $2\pi$) is plotted.
+The above is illustrated by Figures~\ref{fig:circstatplot_line} and~\ref{fig:circstatplot_cube}, where the phase (divided by $2\pi$) of two distinct polynomial phase signals is plotted on the left, and the principal component of the phase (also divided by $2\pi$) is plotted on the right.
 The circles display the samples at the integers.
@@
 \begin{lemma}\label{lem:intvalpol}
-  The integer valued polynomials $p_0,\dots,p_m$ are an integer basis for $\mathcal{Z}$.  That is, every polynomial in $\mathcal{Z}$ can be uniquely written as
+  The integer valued polynomials $p_0,\dots,p_m$ form an integer basis for $\mathcal{Z}$.  That is, every polynomial in $\mathcal{Z}$ can be uniquely written as
 \begin{equation} \label{eq:lem_polynomial}
@@
 \]
-denote the column vector of length $m+1$ containing the coefficients of $g$.
-Superscript $^\prime$ indicates the transpose.
-If $y$ and $z$ differ by a polynomial from $\mathcal{Z}$ then we can write $y = z + p$ where $p \in \mathcal{Z}$ and then also $\coef(y) = \coef(z) + \coef(p)$.
+denote the column vector of length $m+1$ containing the coefficients of $g$, where $^\prime$ indicates transpose.
+If $y$ and $z$ differ by a polynomial in $\mathcal{Z}$ then we can write $y = z + p$ where $p \in \mathcal{Z}$ and then also $\coef(y) = \coef(z) + \coef(p)$.
 %\footnote{In group theory terminology $\coef(\cdot)$ coupled with vector addition is called a \term{group homomorphism}.\index{group homomorphism}}. 
@@
 \]
-containing the coefficient vectors corresponding to the polynomials in $\mathcal{Z}$.  Since the integer valued polynomials are a basis for $\mathcal{Z}$,
+containing the coefficient vectors corresponding to the polynomials in $\mathcal{Z}$.  Since the integer valued polynomials form a basis for $\mathcal{Z}$,
 \begin{align*}
@@
 That is, the set of coefficients of the polynomials from $\mathcal{Z}$ forms a lattice with generator matrix $\Pbf$.
-We can restate Theorem~\ref{thm:circpolysampledthm} as:
+We can restate Lemma~\ref{lem:circpolysampledthm} as:
 \begin{corollary}\label{cor:circpolysampledcoef}
@@
 For the purpose of estimating the coefficients of a polynomial phase signal we must (in order to ensure identifiability) restrict the set of allowable coefficients so that no two polynomial phase signals are aliases of each other.
-In consideration of Corollary~\ref{cor:circpolysampledcoef} we require that the coefficients of $y(t)$, written in vector form $\mubf$, are contained in a set of coset representatives for the quotient $\reals^{m+1}/L_{m+1}$.
+In consideration of Corollary~\ref{cor:circpolysampledcoef} we require that the coefficients of $y(t)$, written in vector form $\tilde{\mubf}$, are contained in a set of coset representatives for the quotient $\reals^{m+1}/L_{m+1}$.
 We call the chosen set of representatives the \term{identifiable region}\index{identifiable region}.
@@
 A natural choice is the interval $[-\nicefrac{1}{2}, \nicefrac{1}{2})$.
-The lattice $L_1$ is the 1-dimensional integer lattice $\ints$ and the interval $[-\nicefrac{1}{2}, \nicefrac{1}{2})$ corresponds to the Voronoi cell of $L_1$.
-When $m=1$ it turns out that a natural choice of identifiable region is the square box $[-\nicefrac{1}{2}, \nicefrac{1}{2})^2$.
-This corresponds with the \term{Nyquist criterion}.
-The lattice $L_2$ is equal to $\ints^2$ and the box $[-\nicefrac{1}{2}, \nicefrac{1}{2})^2$ corresponds with the Voronoi cell of $L_2$.
+The lattice $L_1$ is the 1-dimensional integer lattice $\ints$ and the interval $[-\nicefrac{1}{2}, \nicefrac{1}{2})$ is the Voronoi cell of $L_1$.
+When $m=1$ it turns out that a natural choice of identifiable region is the square box $[-\nicefrac{1}{2}, \nicefrac{1}{2})^2$ in accordance with the \term{Nyqvist criterion}.
+The lattice $L_2$ is $\ints^2$ and $[-\nicefrac{1}{2}, \nicefrac{1}{2})^2$ is the Voronoi cell of $L_2$.
 When $m > 1$ the identifiable region becomes more complicated and $L_{m+1} \neq \ints^{m+1}$.
@@
 In this paper we will use the rectangular set constructed using Proposition~\ref{prop:lattice-theory-constructing_a_rectangular_tesselating_region}.
-Observe that $\Pbf$ is upper triangular with $k$th diagonal element equal to $\tfrac{1}{k!}$.
+Observe that the matrix $\Pbf$ is upper triangular with $k$th diagonal element equal to $\tfrac{1}{k!}$.
 So this rectangular region is
@@
 
-Given vectors $\xbf$ and $\ybf$ in $\reals^{m+1}$ we say that $\xbf \equiv \ybf \bmod L_{m+1}$ if $\xbf$ and $\ybf$ differ by a lattice point in $L_{m+1}$.
+Given $\xbf$ and $\ybf$ in $\reals^{m+1}$ we say that $\xbf \equiv \ybf \bmod L_{m+1}$ if $\xbf$ and $\ybf$ differ by a lattice point in $L_{m+1}$.
 We define the function $\dealias(\xbf)$ to take $\xbf$ to its coset representative inside $B$.
@@
 We now describe the least squares unwrapping (LSU) estimator of the polynomial coefficients.
-Recall that we desire to estimate the coefficients $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ from the noisy samples $Y_1, \dots, Y_N$ given in~\eqref{eq:Y_nsamplednoisey}.
-Take the complex argument of the $Y_n$ and divide by $2\pi$ to obtain
+Recall that we desire to estimate the coefficients $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ from the observations $Y_1, \dots, Y_N$ defined in~\eqref{eq:Y_nsamplednoisey}.
+Let
 \begin{equation}\label{eq:noise_circ_poly}
@@
 \end{equation}
-where $\angle{z}$ denotes the complex argument of $z$, and 
+where $\angle$ denotes the complex argument, and 
 \[
@@
 \end{equation}
-The least squares unwrapping estimator is defined as those coefficients $\widehat{\mu}_0, \dots, \widehat{\mu}_m$ that minimise $SS$ over the identifiable region $B$. That is, the LSU estimator is,
+The LSU estimator $\widehat{\mubf}$ is the minimiser of $SS$ over the identifiable region $B$, that is,
 \begin{equation}\label{eq:hatmubfLSUdefn}
@@
 where $W_n = \round{\Theta_{n} - \sum_{k = 0}^{m}{\mu_k n^k} }$ for $n = 1, \dots, N$ are integers called \emph{wrapping variables}.
-If we consider $W_1,\dots,W_N$ as nuisance parameters to be estimated then $SS$ can be written as a function of both $\mu_0,\dots,\mu_m$ and $W_1,\dots,W_N$.
-The least squares unwrapping estimator is then found by jointly minimising over $\mu_0,\dots,\mu_m$ and $W_1,\dots,W_N$.
+If we consider $W_1,\dots,W_N$ as nuisance parameters to be estimated then $SS$ can be considered as a function of both $\mu_0,\dots,\mu_m$ and $W_1,\dots,W_N$, the minimiser with respect to $W_1,\dots,W_N$ being $SS(\mubf)$.
+The LSU estimator can be found by jointly minimising over $\mu_0,\dots,\mu_m$ and $W_1,\dots,W_N$.
 This joint minimisation problem can be solved by computing a nearest point in a lattice.
-To see this write $SS$ as a function of both $\mu_0,\dots,\mu_m$ and $W_1,\dots,W_N$ using vectors,
+To see this write $SS$ as a function of both $\mu_0,\dots,\mu_m$ and $W_1,\dots,W_N$,
 \begin{equation}\label{eq:SS_in_vector_form_with_u_nuisance}
@@
 \end{equation}
-where $\|\cdot\|^2$ denotes the squared Euclidean norm, and $\Xbf$ is the rectangular Vandermonde matrix
+where $\|\cdot\|^2$ denotes the squared Euclidean norm, $\Xbf$ is the rectangular Vandermonde matrix
 \begin{equation}\label{eq:Xvandermonde}
@@
 However, there is a substantial body of work dedicated to the development of algorithms for computing a nearest lattice point~\cite{Agrell2002,schnorr_euchner_sd_1994,Viterbo_sphere_decoder_1999}, and we exploit this for the purpose of computing the LSU estimator (see Section~\ref{sec:lattice-theory}).
-There exist algorithms that can compute a nearest point in $\Lambda$ that require a number of operations that is polynomial in $N$~\cite[Sec 4.3]{McKilliam2010thesis}.
+There exist algorithms that can compute a nearest point in $\Lambda$ using a number of operations that is polynomial in $N$~\cite[Sec 4.3]{McKilliam2010thesis}.
 Although polynomial in complexity, these algorithms are not fast in practice.
@@
 \]
-where $\expect$ denotes the expected value.
+where $\expect$ denotes expected value.
 Depending on the distribution of $\Phi_n$ the argument that minimises~\eqref{eq:intrmeandefn} may not be unique.
@@
 \]
-converges to the normal with zero mean and covariance
+converges to the normal with zero mean and covariance matrix
 \[
@@
 A proof of the first part of this theorem (strong consistency) is given in Section~\ref{sec:strongconstproof}.
-Space restrictions make it impossible to give a proof of asymptotic normality in this paper.
+Space restrictions make it impossible prove asymptotic normality in this paper.
 A complete proof of normality is available in~\cite{McKilliam_arxiv_pps_unwrapping_2012}.
-Proofs for the case of $m=0$ were given in~\cite{McKilliam_mean_dir_est_sq_arc_length2010} and for the case when $m=1$ were given in~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.
+Proofs for the case of $m=0$ were given in~\cite{McKilliam_mean_dir_est_sq_arc_length2010} and $m=1$ in~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.
 The proof here takes a similar approach, but requires new techniques.
@@
 To see why this makes sense, consider the case when $m=0$, $\tilde{\mu}_0 = -0.5$ and $\widehat{\mu}_0 = 0.49$, so that $\tilde{\mu}_0 - \widehat{\mu}_0 = -0.99$.
-However, the two phases are obviously close, since the phase $\pm 0.5$ are actually the same.
+However, the two phases are obviously close, since the phases $\pm 0.5$ are actually the same.
 In this case
@@
 In the interest of simplicity we only consider the case when $\Phi_1, \dots, \Phi_N$ are identically distributed here.
-If $X_n$ is circularly symmetric with density function nonincreasing with magnitude $\abs{X_n}$, then, the corresponding $\Phi_n$ necessarily has zero intrinsic mean~\cite[Theorem 5.2]{McKilliam2010thesis}.
+If $X_n$ is circularly symmetric with pdf nonincreasing with magnitude $\abs{X_n}$, then, the corresponding $\Phi_n$ necessarily has zero intrinsic mean~\cite[Theorem 5.2]{McKilliam2010thesis}.
 Thus, our theorem covers commonly used distributions for $X_1, \dots, X_N$, such as the normal distribution.
@@
 Let $\lambdabf = \dealias(\tilde{\mubf} - \mubf) = \tilde{\mubf} - \mubf - \pbf$ where $\pbf$ is a lattice point from $L_{m+1}$.
-From the definition of $L_{m+1}$ we have $p_0 + p_1 n + \dots + p_{m} n^m$ an integer whenever $n$ is an integer, so
+From the definition of $L_{m+1}$ we have $p_0 + p_1 n + \dots + p_{m} n^m$ an integer whenever $n$ is an integer, and so
 \begin{align*}
@@
 \]
-Using standard techniques~\cite{Pollard_conv_stat_proc_1984,van2009empirical} one can show that
+Using techniques described in~\cite{Pollard_conv_stat_proc_1984,van2009empirical} one can show that
  \begin{equation}\label{eq:SNVNunifmlln}
@@
 \begin{lemma}\label{lem:ES_Nminimisedzero}
-For $\lambdabf \in B$ the function $V_N(\lambdabf)$ is minimised uniquely at $\zerobf$, the vector of all zeros.  At this minimum $V_N(\zerobf) = \sigma^2$.
+The minimum value of $V_N(\lambdabf)$ over $\lambdabf \in B$ is $\sigma^2$, uniquely attained at the origin $\zerobf$.
 \end{lemma}
@@
 \end{align*}
-We know that $\expect\fracpart{ \Phi_n + \fracpart{z(n)} }^2$ is minimised uniquely when $\fracpart{z(n)} = 0$ at which point it takes the value $\sigma^2$.
-Now $\fracpart{z(n)}$ is equal to zero for all integers $n$ if and only if $z \in \mathcal{Z}$, or equivalently if $\coef(z)$ is a lattice point in $L_{m+1}$.
-By definition $B$ contains precisely one lattice point from $L_{m+1}$, this being the origin $\zerobf$.
+From above $\expect\fracpart{ \Phi_n + \fracpart{z(n)} }^2$ has minimum value $\sigma^2$, uniquely when $\fracpart{z(n)} = 0$.
+But, $\fracpart{z(n)}$ is zero for all integers $n$ if and only if $z \in \mathcal{Z}$, or equivalently if $\coef(z)$ is a lattice point in $L_{m+1}$.
+By definition $B$ contains precisely one lattice point from $L_{m+1}$, namely $\zerobf$.
 Therefore $V_N$ is minimised uniquely at $\zerobf$, at which point it takes the value $\sigma^2$.
@@
 \begin{proof}
-By definition $\widehat{\lambdabf}_N =\arg\min_{\lambdabf\in B} S_N(\lambdabf)$ so 
+Since $\widehat{\lambdabf}_N =\arg\min_{\lambdabf\in B} S_N(\lambdabf)$, 
 \[
@@
 We have now shown that $V_N$ is uniquely minimised at $\zerobf$, that $V_N(\zerobf) = \sigma^2$, and that $V_N(\widehat{\lambdabf}_N)$ converges almost surely to $\sigma^2$.
-These results are enough to show that $\widehat{\lambdabf}_N$ converges almost surely to zero.
-However, this tells us nothing about the rate at which the components of $\widehat{\lambdabf}_N$ approach zero as required by Theorem~\ref{thm:asymp_proof}.
+We could now proceed to prove that $\widehat{\lambdabf}_N$ converges almost surely to $\zerobf$, but this would tell us nothing about the superior rates of convergence stated in Theorem~\ref{thm:asymp_proof}.
 To prove these stronger properties we need some preliminary results about arithmetic progressions, and from the calculus of finite differences.
@@
 of length $m+1$ is contained in the subset $K$.
-If $K$ is a small subset of $W$ then $A(h,K)$ might be empty.
+If $K$ is small subset of $W$ then $A(h,K)$ might be empty.
 However, the next two lemmas and the following corollary will show that if $K$ is sufficiently large then it always contains at least one arithmetic progression (for all sufficiently small $h$) and therefore $A(h,K)$ is not empty.
@@
 The study of arithmetic progressions within subsets of $W$ has a considerable history~\cite{Erdos_on_some_sequence_of_integers1936,Szemeredi_setint_no_k_arth1975,Gowers_new_proof2001}.
-In particular, Gower's~\cite[Theorem 1.3]{Gowers_new_proof2001} gives a result far stronger than we require here.
+In particular, Gowers~\cite[Theorem 1.3]{Gowers_new_proof2001} gives a result far stronger than we require here.
 Denote by $K \backslash \{r\}$ the set $K$ with the element $r$ removed.
@@
  Let $K \subseteq W$ such that $|K| > \frac{2m+1}{2m+2}N$.
-For all $h$ such that $1\leq h \leq\frac{N}{2m}$ the set $K$ contains at least one arithmetic progression $n, n+h, \dots, n+mh$ of length $m+1$.
+For all integers $h$ such that $1\leq h \leq\frac{N}{2m}$ the set $K$ contains at least one arithmetic progression $n, n+h, \dots, n+mh$ of length $m+1$.
 That is, $|A(h,K)| > 0$.
@@
 
-The next result we require comes from the calculus of finite differences. For any function $d(n)$ mapping $\reals$ to $\reals$, let 
+The next result we require comes from the calculus of finite differences. For any function $d(n)$ from $\reals$ to $\reals$, let 
 \[
@@
  \end{equation}
-So, the $r$th difference of the polynomial is a constant depending on $h$, $r$ and the $r$th coefficient $a_r$~\cite[page 51]{Jordan_Calculus_of_finite_difference_1965}.
+Thus, the $r$th difference of $d(n)$ is a constant depending on $h$, $r$ and the $r$th coefficient $a_r$~\cite[page 51]{Jordan_Calculus_of_finite_difference_1965}.
 We can now continue the proof of strong consistency.
@@
 \begin{lemma}\label{lem:moran2}
-Suppose $\lambdabf_1, \lambdabf_2,\dots$ is a sequence of vectors from $B$ with $V_N(\lambdabf_N) - \sigma^2\rightarrow 0$ as $N\rightarrow\infty$.
+Suppose $\{\lambdabf_n\}$ is a sequence from $B$ with $V_N(\lambdabf_N) - \sigma^2\rightarrow 0$ as $N\rightarrow\infty$.
 Then the elements $\lambda_{0,N}, \dots \lambda_{m,N}$ of $\lambdabf_N$ satisfy $N^k\lambda_{k, N}\rightarrow0$ as $N\rightarrow\infty$.
@@
 \end{equation}
-which is continuous in $z$. Because of~\eqref{eq:Efracpartmined} and~\eqref{eq:Efracpartphi}, $g(z) \geq 0$ with equality only at $z = 0$ for $z \in [-\nicefrac{1}{2}, \nicefrac{1}{2})$. Now
+which is continuous in $z$. Because of~\eqref{eq:Efracpartmined} and~\eqref{eq:Efracpartphi}, $g(z) \geq 0$ with equality when $\fracpart{z} = 0$.  Now
 \[
@@
 \end{align*}
-violating that $V_N(\lambdabf_N) - \sigma^2$ converges to zero as $N \rightarrow \infty$.  We will assume $N > N_0$ in what follows.
+contradicting the fact that $V_N(\lambdabf_N) - \sigma^2$ converges to zero as $N \rightarrow \infty$.  We will assume that $N > N_0$ in what follows.
 
@@
 \]
-So, when $h = 2$, 
+Thus, when $h = 2$, 
 \[
@@
  \begin{lemma} \label{lem:fracpartsumanddelta}
- Let $a_1, a_2, \dots, a_r$ be $r$ real numbers such  that $\left|\fracpart{a_n}\right| < \delta$ for all $n = 1,2,\dots,r$.  Then $\left|\fracpart{\sum_{n=1}^r{a_n}}\right| < r\delta.$
+ Let $a_1, a_2, \dots, a_r$ be $r$ real numbers for which $\left|\fracpart{a_n}\right| < \delta$ for all $n = 1,2,\dots,r$.  Then $\left|\fracpart{\sum_{n=1}^r{a_n}}\right| < r\delta.$
  \end{lemma}
@@
 We are now in a position to complete the proof of strong consistency.
-As is customary, let $\Omega$ be the sample space on which the random variables $\{X_i\}$ and $\{\Phi_i\}$ are defined.
-Let $A$ be the subset of the sample space $\Omega$ on which  $V_N(\widehat{\lambdabf}_N) - \sigma^2 \rightarrow 0$ as $N\rightarrow\infty$.
+Let $\Omega$ be the sample space on which the random variables $\{X_i\}$ and $\{\Phi_i\}$ are defined.
+Let $A$ be the subset of $\Omega$ on which  $V_N(\widehat{\lambdabf}_N) - \sigma^2 \rightarrow 0$ as $N\rightarrow\infty$.
 Now $\prob\{A\} =1$ as a result of Lemma~\ref{lem:ESNconv}.
@@
 The corresponding signal to noise ratio (SNR) is $\rho^2 \sigma_c^{-2} = \sigma_c^{-2}$.
-The number of replications of each Monte-Carlo experiment is $T = 2000$ to obtain estimates $\widehat{\mubf}_1, \dots, \widehat{\mubf}_T$ and the corresponding dealiased errors $\widehat{\lambdabf}_t = \dealias(\widehat{\mubf}_T - \tilde{\mubf})$ are computed.
+The number of replications of each Monte-Carlo experiment is $T = 2000$ to obtain estimates $\widehat{\mubf}_1, \dots, \widehat{\mubf}_T$ and the corresponding dealiased errors $\widehat{\lambdabf}_t = \dealias(\tilde{\mubf} - \widehat{\mubf}_t)$ are computed.
 The sample mean square error (MSE) of the $k$th coefficient is computed according to $\tfrac{1}{T}\sum_{t=1}^T \widehat{\lambda}_{k,t}^2$ where $\widehat{\lambda}_{k,t}$ is the $k$th element of $\widehat{\lambdabf}_t$.
@@
 It is informative to compare the volume of the set of coefficients suitable for the HAF, PHAF and CPF estimators with the volume of the identifiable region $B$ from~\eqref{eq:rectangular_identifiable_region}.
-The identifiable region has volume
-\[
-\vol(B) = \prod_{k=0}^m {\frac{1}{k!}}.
-\]
+The identifiable region has volume $\vol(B) = \prod_{k=0}^m {\frac{1}{k!}}$.
 The volume of coefficients suitable for the HAF and PHAF estimators is
@@
 It is now worth reconsidering the results in Figure~\ref{plot:m3small}.
-Substituting $m=3$ and $N=199$ in the formula above, we see that the HAF, PHAF and CPF estimators search a space of parameters that is more than $10^5$ times smaller than that searched by the LSU estimator.
-In this context we might expect the HAF, PHAF and CPF estimators to have thresholds at lower SNR than LSU estimator.
+Substituting $m=3$ and $N=199$ in the formula above, we see that the HAF, PHAF and CPF estimators search a space of parameters that is approximately $10^{5}$ times smaller than that searched by the LSU estimator.
+In this context we might expect the HAF, PHAF and CPF estimators to have thresholds at lower SNR than the LSU estimator.
 It is encouraging that the LSU estimator has comparable performance.
 To put these results further into context, Figure~\ref{plot:m3big} repeats the simulations in Figure~\ref{plot:m3small}, but where the true coefficients are $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$.
-The HAF, PHAF and CPF estimators fail consistently.
+The HAF, PHAF and CPF estimators fail consistently, as expected.
 The performance of the LSU estimator is unchanged, as expected in view of Theorem~\ref{thm:asymp_proof}.
@@
   		\includegraphics{code/data/plot3-1.mps} 
-  		\caption{Sample MSE for highest order coefficient $\tilde{\mu}_3$ for a polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}  \right]$ and $N=199$.}
+  		\caption{Sample MSE for coefficient $\tilde{\mu}_3$ for a polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}  \right]$ and $N=199$.}
   		\label{plot:m3small} 
@@
   		\includegraphics{code/data/plot3-2.mps} 
-  		\caption{Sample MSE for the highest order coefficient $\tilde{\mu}_3$ for a polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$ and $N=199$.}
+  		\caption{Sample MSE for the coefficient $\tilde{\mu}_3$ for a polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$ and $N=199$.}
   		\label{plot:m3big} 
@@
   		\includegraphics{code/data/plot3-3.mps} 
-  		\caption{Sample MSE for the highest order coefficient $\tilde{\mu}_3$ for a polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$ and $N=199$.}
+  		\caption{Sample MSE for the coefficient $\tilde{\mu}_3$ for a polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$ and $N=199$.}
   		\label{plot:m3bigzw} 
@@
   		\includegraphics{code/data/plot5-1.mps} 
-  		\caption{Sample MSE for the highest order coefficient $\tilde{\mu}_5$ for a polynomial phase signal of order 5 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}, \tfrac{1}{96N^3}, \tfrac{1}{480N^4} \right]$ and $N=199$.}
+  		\caption{Sample MSE for the coefficient $\tilde{\mu}_5$ for a polynomial phase signal of order 5 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}, \tfrac{1}{96N^3}, \tfrac{1}{480N^4} \right]$ and $N=199$.}
   		\label{plot:m5} 
@@
   		\includegraphics{code/data/plot5-2.mps} 
-  		\caption{Sample MSE for the highest order coefficient $\tilde{\mu}_5$ for a polynomial phase signal of order 5 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}, \tfrac{1}{96}, \tfrac{1}{480}   \right]$ and $N=199$.}
+  		\caption{Sample MSE for the coefficient $\tilde{\mu}_5$ for a polynomial phase signal of order 5 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}, \tfrac{1}{96}, \tfrac{1}{480}   \right]$ and $N=199$.}
   		\label{plot:m5big} 
@@
 
-A key property of the LSU estimator is that it functions correctly for polynomial phase parameters contained anywhere inside the identifiable region described in Section~\ref{sec:ident_aliasing}.
+A key property of the LSU estimator is that it works for polynomial phase parameters contained anywhere inside the identifiable region described in Section~\ref{sec:ident_aliasing}.
 This is the maximum possible range of parameters.
-By contrast popular estimators such as the HAF and CPF operate correctly only for parameters in a range much smaller than the identifiable region.
+By contrast popular estimators such as the HAF and CPF work only for parameters in a range much smaller than the identifiable region.
 For this reason, the LSU estimator may be the only presently existing estimator that is viable for certain applications, such as interference suppression~\cite{Djukanovic:2011:PMN:1950994.1951252,Djurovic_aliasing_2012}.

### 2013-04-04 paper.tex
added more barry edits

@@
 The information contained in the amplitude is not essential for accurate estimation of polynomial phase signals, and can be discarded with only small loss in accuracy.
-For example, in the case that $\{X_n\}$ is white and Gaussian, our results show that the asymptotic variance of the LSU estimator converges to the asymptotic Cramer-Rao lower bound (CRB) as the SNR becomes large, and converges to $\tfrac{2}{3}\pi$ times the asymptotic CRB as the SNR approaches zero.
+For example, in the case that $\{X_n\}$ is white and Gaussian, our results show that, asymptotically in $N$, the ratio of the variance of the LSU estimator to the Cramer-Rao lower bound (CRB) converges to $1$ for high SNR, and to $\tfrac{2}{3}\pi$ for low SNR.
 Related results have been presented in~\cite{Quinn2009_dasp_phase_only_information_loss}.
-This factor of $\tfrac{2}{3}\pi$ times the CRB at low SNR is better than popular estimators such as the HAF and CPF, for which the ratio of the asymptotic variance and the asymptotic CRB diverges as the SNR approaches zero~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.
+This factor of $\tfrac{2}{3}\pi$ for low SNR is better than popular estimators such as the HAF and CPF, for which the ratio of the variance and the CRB diverges as the SNR become small~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.
 
@@
 For this reason, the LSU estimator may be the only known estimator that is viable for certain applications, such as interference suppression.
-The theoretical results given in Section~\ref{sec:least-squar-unwr} prove that the LSU estimator work over the entire identifiable region.
+The theoretical results given in Section~\ref{sec:least-squar-unwr} prove that the LSU estimator works over the entire identifiable region.
 To our knowledge this is the first time a result of this type has been obtained for any polynomial phase estimator.
@@
 These identifiability results are required in order to understand the statistical properties of the LSU estimator.
-Section~\ref{sec:least-squar-unwr} describes the LSU estimator and states a theorem asserting the estimator to be strongly consistent and asymptotically normally distributed under some mild assumptions on the noise $\{X_n\}$.
+Section~\ref{sec:least-squar-unwr} describes the LSU estimator and states a theorem asserting the estimator to be strongly consistent and asymptotically normally distributed under some mild assumptions on the noise process $\{X_n\}$.
 The proof of strong consistency is given in Section~\ref{sec:strongconstproof}.
@@
 Section~\ref{sec:simulations} describes the results of Monte Carlo simulations that compare the LSU estimator with existing estimators including the HAF, PHAF and CPF.
-These simulations agree with the derived asymptotic properties.
+These simulations agree with the asymptotic properties.
 
@@
 
-The above is illustrated by Figures~\ref{fig:circstatplot_line} and~\ref{fig:circstatplot_cube}, where the phase (divided by $2\pi$) of two distinct polynomial phase signals is plotted on the left, and the principal component of the phase (also divided by $2\pi$) is plotted on the right.
+The above is illustrated by Figures~\ref{fig:circstatplot_line} and~\ref{fig:circstatplot_cube}, where the phase (divided by $2\pi$) of two distinct polynomial phase signals is plotted on the left, and the principal component of the phase (also divided by $2\pi$) on the right.
 The circles display the samples at the integers.
@@
 We now describe the least squares unwrapping (LSU) estimator of the polynomial coefficients.
-Recall that we desire to estimate the coefficients $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ from the observations $Y_1, \dots, Y_N$ defined in~\eqref{eq:Y_nsamplednoisey}.
+Recall that we desire to estimate $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ from the observations $Y_1, \dots, Y_N$ defined in~\eqref{eq:Y_nsamplednoisey}.
 Let
@@
 \end{equation}
-where $\angle$ denotes the complex argument, and 
+where $\angle z$ denotes the argument (or angle) of the complex number $z$, and 
 \[
@@
 \begin{lemma}\label{lem:ES_Nminimisedzero}
-The minimum value of $V_N(\lambdabf)$ over $\lambdabf \in B$ is $\sigma^2$, uniquely attained at the origin $\zerobf$.
+The minimum value of $V_N(\lambdabf)$ over $\lambdabf \in B$ is $\sigma^2$, uniquely attained at the $\lambdabf = \zerobf$ where $\zerobf$ is the origin.
 \end{lemma}
@@
 of length $m+1$ is contained in the subset $K$.
-If $K$ is small subset of $W$ then $A(h,K)$ might be empty.
+As the size of the subset $K$ decreases it is possible that $A(h,K)$ might be empty.
 However, the next two lemmas and the following corollary will show that if $K$ is sufficiently large then it always contains at least one arithmetic progression (for all sufficiently small $h$) and therefore $A(h,K)$ is not empty.
@@
 
-The next result we require comes from the calculus of finite differences. For any function $d(n)$ from $\reals$ to $\reals$, let 
+The next result we require comes from the calculus of finite differences. For any function $d(n)\from\reals\to\reals$, let 
 \[
@@
 \end{equation}
-which is continuous in $z$. Because of~\eqref{eq:Efracpartmined} and~\eqref{eq:Efracpartphi}, $g(z) \geq 0$ with equality when $\fracpart{z} = 0$.  Now
+which is continuous in $z$. Because of~\eqref{eq:Efracpartmined} and~\eqref{eq:Efracpartphi}, $g(z) \geq 0$ with equality only when $\fracpart{z} = 0$.  Now
 \[
@@
 Provided the SNR is large enough (so that the `threshold' is avoided) the sample MSE of the LSU estimator is close to that predicted by Theorem~\ref{thm:asymp_proof}.
-Using Theorem~\ref{thm:asymp_proof} we can predict the ratio between the CRB and the asymptotic variance of the LSU estimator to approach 1 as SNR approaches $\infty$ and $\tfrac{2}{3}\pi$ as SNR approaches zero~(see Lemma~3 and the subsequent discussion in~\cite{Quinn2009_dasp_phase_only_information_loss}).
-Contrastingly, the ratio between the CRB and asymptotic variance of the HAF, PHAF and CPF estimators is known to diverge as SNR goes to zero~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.
+Using Theorem~\ref{thm:asymp_proof} we can predict that, asymptotically in $N$, the ratio between the CRB and the variance of the LSU estimator will approach $1$ for high SNR and $\tfrac{2}{3}\pi$ for low SNR~(see Lemma~3 and the subsequent discussion in~\cite{Quinn2009_dasp_phase_only_information_loss}).
+Contrastingly, the ratio between the CRB and variance of the HAF, PHAF and CPF estimators is known to diverge as the SNR becomes small~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.
 %For this reason we can expect the LSU estimator to be more accurate at low SNR for sufficiently large $N$.
@@
 It is now worth reconsidering the results in Figure~\ref{plot:m3small}.
-Substituting $m=3$ and $N=199$ in the formula above, we see that the HAF, PHAF and CPF estimators search a space of parameters that is approximately $10^{5}$ times smaller than that searched by the LSU estimator.
+Substituting $m=3$ and $N=199$ in the formula above, we see that the HAF, PHAF and CPF estimators search a space of parameters that is less than $10^{-5}$ the size of the identifiable region searched by the LSU estimator.
 In this context we might expect the HAF, PHAF and CPF estimators to have thresholds at lower SNR than the LSU estimator.
@@
 The effect becomes more severe as $m$ increases.
-When $m=5$ and $N=199$ the space of coefficients searched these estimators is approximately $10^{16}$ times smaller than the identifiable region.
+When $m=5$ and $N=199$ the space of coefficients searched by these estimators is approximately $10^{-16}$ the size of the identifiable region.
 This is a dramatic reduction in the range of detectable polynomial phase signals.

### 2013-04-15 paper.tex
minor paper mod

@@
-%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
-\documentclass[journal]{IEEEtran}
+\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
+%\documentclass[journal]{IEEEtran}
 
@@
 Related results have been presented in~\cite{Quinn2009_dasp_phase_only_information_loss}.
-This factor of $\tfrac{2}{3}\pi$ for low SNR is better than popular estimators such as the HAF and CPF, for which the ratio of the variance and the CRB diverges as the SNR become small~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.
+This factor of $\tfrac{2}{3}\pi$ for low SNR is better than popular estimators such as the HAF and CPF, for which the ratio of the variance to the CRB diverges as the SNR becomes small~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.
 
@@
 Provided the SNR is large enough (so that the `threshold' is avoided) the sample MSE of the LSU estimator is close to that predicted by Theorem~\ref{thm:asymp_proof}.
-Using Theorem~\ref{thm:asymp_proof} we can predict that, asymptotically in $N$, the ratio between the CRB and the variance of the LSU estimator will approach $1$ for high SNR and $\tfrac{2}{3}\pi$ for low SNR~(see Lemma~3 and the subsequent discussion in~\cite{Quinn2009_dasp_phase_only_information_loss}).
-Contrastingly, the ratio between the CRB and variance of the HAF, PHAF and CPF estimators is known to diverge as the SNR becomes small~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.
+Using Theorem~\ref{thm:asymp_proof} we can predict that, asymptotically in $N$, the ratio of the variance of the LSU estimator to the CRB will approach $1$ for high SNR and $\tfrac{2}{3}\pi$ for low SNR~(see Lemma~3 and the subsequent discussion in~\cite{Quinn2009_dasp_phase_only_information_loss}).
+Contrastingly, the ratio of the variance of the HAF, PHAF and CPF estimators to the CRB is known to diverge as the SNR becomes small~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.
 %For this reason we can expect the LSU estimator to be more accurate at low SNR for sufficiently large $N$.

### 2013-06-26 paper.tex
added to paper, needs another read, and to go over responses again

@@
-\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
-%\documentclass[journal]{IEEEtran}
+%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
+\documentclass[journal]{IEEEtran}
 
@@
 %\title{Polynomial phase estimation by phase unwrapping 1: Identifiability and strong consistency}
-\title{Polynomial phase estimation by phase unwrapping}
+\title{Polynomial phase estimation by least squares phase unwrapping}
 
@@
 Following~\cite{Zhan2006_K_best_sphere_decoder} we refer to this as the \emph{$K$-best algorithm}.
-We have found that this algorithm produces excellent estimates of polynomial phase coefficients and can be computed in a reasonable amount of time for $N$ less than about 1000.
+The $K$-best algorithm requires $O(NK^2 \log K)$ arithmetic operations.
+The maximum number of branches $K$ is free to be chosen.
+For polynomial phase estimation we have found that setting $K=10N$ yields excellent results, and in this case the algorithm requires $O(N^3 \log N)$ operations.
+Simulations in Section~\ref{sec:simulations} show that this algorithm produces excellent estimates of polynomial phase coefficients and can be computed in a reasonable amount of time for $N$ less than about 1000.
 
@@
 
-A proof of the first part of this theorem (strong consistency) is given in Section~\ref{sec:strongconstproof}.
+A proof of the first part of this theorem (strong consistency) is given in Appendix~\ref{sec:strongconstproof}.
 Space restrictions make it impossible prove asymptotic normality in this paper.
@@
 The proof here takes a similar approach, but requires new techniques.
-%Before giving the proof of strong consistency we make the following remarks.
+%Before giving the proof of strong consistency we make the following remarks
+A statement similar to Theorem~\ref{thm:asymp_proof} without a complete proof was given in~\cite{McKilliam_polyphase_est_icassp_2011}.
+Here, we give a proof.
+The results here are also more general than in~\cite{McKilliam_polyphase_est_icassp_2011}, allowing for a wider class of noise distributions.
 
@@
 
-\section{Proof of strong consistency}\label{sec:strongconstproof}
+\section{Simulations}\label{sec:simulations}
+
+% This section describes the results of Monte-Carlo simulations involving the least squares unwrapping estimator (LSU), the least squares estimator (LSE), Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, and the discrete polynomial phase transform (DPT)~\cite{Peleg_DPT_1995}.
+In all simulations the sample sizes considered are $N = 10, 50$ and $200$ and the unknown amplitude is $\rho = 1$.
+The number of replications of each experiment is 2000 and the $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ are distributed uniformly randomly over the identifiable region.
+
+% Figure~\ref{plot:lsetest} shows the performance of the LSU estimator and LSE when $m=2$ (linear chirp).
+Only results for the highest order parameter $\widehat{\mu}_2$ are displayed, the results for the lower order parameters leading to similar conclusions.
+The LSE is computed using the approach described in~\cite[Sec.~10.1]{McKilliam2010thesis} which is based on that of Abatzoglou~\cite{Abatzoglou_ml_chirp_1986}.
+The LSE is computationally intensive.
+We were able to compute it within a reasonable amount of time only with $N$ less than $50$ when $m=2$.
+Results for the exact LSU estimator computed using the sphere decoder and the approximate LSU estimator computed using Babai's algorithm are displayed.
+The sphere decoder is computationally intractable when $N = 200$.
+It is evident that a significant performance penalty is suffered by using Babai's algorithm when the noise variance is large.
+We have also displayed results using another approximate nearest point algorithm called the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder}.
+Our implementation requires $O(N^3\log{N})$ operations.
+This is more computationally expensive than Babai's algorithm but, as can be seen, the $K$-best method is more statistically accurate.
+
+% For the simulations displayed in the left hand plot of Figure~\ref{plot:lsetest} the signal model used is that from~\eqref{eq:Y_nsamplednoisey}.
+The noise terms $X_1, \dots, X_N$ are independent and identically distributed zero mean complex Gaussian random variables with independent real and imaginary parts having variance $\sigma_c^2$.
+In this case the LSE is also the maximum likelihood estimator.
+The \term{Cram\`{e}r-Rao lower bound} (CRB) for the variance of unbiased estimators of $\tilde{\mu}_j$ is given by Peleg and Porat~\cite{Peleg1991_CRB_PPS_1991} to be
+% \[ 
+% \frac{\sigma_c^2}{4\pi^2 N^{2j + 1}} C^{-1}_{jj} + o(N^{-2j-1})
+% \]
+% where $C^{-1}_{jj}$ is the $j$th diagonal element of the inverse of the $m+1$ by $m+1$ Hilbert matrix $\Cbf$.
+Observe that the performance of the exact LSU estimator (computed using the sphere decoder) is close to the maximum likelihood LSE.
+The asymptotic Cramer-Rao bound (CRB) (solid line) and the asymptotic variance predicted in Theorem~\ref{thm:asymp_proof} (dashed line) are also shown.
+Provided the noise variance is small enough (so that the `threshold' is avoided) the LSE variance is close to the CRB (as is expected) and the LSU variance is close to that predicted by Theorem~\ref{thm:asymp_proof}.
+% %Note the gap between the CRB and the variance predicted by Theorem~\ref{thm:asymp_proof}.
+In practice this gap can be overcome using a numerical technique, such as Newton's method, with the LSU estimate as a starting point.
+We have not done this here in order to show the accuracy of Theorem~\ref{thm:asymp_proof}.
+The other noticeable gap is that between the threshold behaviour of the LSU estimator and the LSE.
+When $N=50$ the threshold \emph{kicks in} a little earlier for the LSE than it does for the LSU.
+We suspect that this gap will actually \emph{close} as $N$ increases.
+This hypothesis is supported by the fact that this behaviour is observed in the case of frequency estimation (when $m=1$)~\cite[Sec.
+9.5]{McKilliam2010thesis}\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.
+Unfortunately it is not feasible to test this hypothesis because of the computational complexity of the LSE.
+% For the simulations displayed in the right hand plot of Figure~\ref{plot:lsetest} the signal model used is that from~\eqref{eq:noise_circ_poly}.
+The phase noise terms $\Phi_1, \dots, \Phi_N$ are uniform random variables with variance $\sigma^2$.
+In this case, the LSU estimator performs slightly better than the LSE.
+This result agrees with similar observations made in circular statistics~\cite{McKilliam_mean_dir_est_sq_arc_length2010}.
+The asymptotic variance predicted by Theorem~\ref{thm:asymp_proof} again accurately predicts the performance of the LSU estimator.
+
+% Figure~\ref{plot:polyest} displays the performance of the LSU estimator, the DPT and Kitchen's estimator when $m=3$.
+Both the DPT and Kitchen's estimator perform poorly because they do not work correctly for some parameters in the identifiable region $B$.
+This behaviour has been observed previously~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase,McKilliam_polyphase_est_icassp_2011}.
+Again the performance of the LSU is well predicted by Theorem~\ref{thm:asymp_proof} provided that the noise variance is small enough to avoid the threshold effect and provided that $N$ is sufficiently large.
+The asymptotic CRB is also plotted.
+
+This section describes the results of Monte-Carlo simulations with the least squares unwrapping (LSU) estimator, the high order ambiguity function (HAF) estimator~\cite{Peleg_DPT_1995}, the product high order ambiguity function (PHAF) estimator~\cite{Barbarossa_PHAF_1998}, the cubic phase function (CPF) estimator~\cite{Oshea_cpf_2004} and the estimator that results from combining the HAF and CPF~\cite{Djurovic_haf_cpf_2012}.
+The CPF estimator only applies to polynomial phase signals of order $m=3$ and when the number of observations $N$ is odd.
+In all simulations the unknown amplitude is $\rho = 1$ and $N = 199$.
+The $X_1, \dots, X_N$ are pseudorandomly generated independent and identically distributed circularly symmetric complex Gaussian with variance $\expect\abs{X_1}^2 = \sigma_c^2$.
+The corresponding signal to noise ratio (SNR) is $\rho^2 \sigma_c^{-2} = \sigma_c^{-2}$.
+The number of replications of each Monte-Carlo experiment is $T = 2000$ to obtain estimates $\widehat{\mubf}_1, \dots, \widehat{\mubf}_T$ and the corresponding dealiased errors $\widehat{\lambdabf}_t = \dealias(\tilde{\mubf} - \widehat{\mubf}_t)$ are computed.
+The sample mean square error (MSE) of the $k$th coefficient is computed according to $\tfrac{1}{T}\sum_{t=1}^T \widehat{\lambda}_{k,t}^2$ where $\widehat{\lambda}_{k,t}$ is the $k$th element of $\widehat{\lambdabf}_t$.
+In all simulations four \emph{lag sets} are used for the PHAF estimator~\cite{Barbarossa_PHAF_1998} and the LSU estimator is approximated using the $K$-best method described in Section~\ref{sec:lattice-theory}.
+The maximum number of branches allowed is $K=20N$.
+
+Figure~\ref{plot:m3small} shows the sample MSE of the highest order coefficient $\mu_3$ for a polynomial phase signal of order $m=3$.
+All the estimators perform poorly at low SNR until a `threshold' is reached, after which the variance is close to the Cramer-Rao bound (CRB)~\cite{Peleg1991_CRB_PPS_1991} indicated by the solid line.
+The asymptotic variance of the LSU estimator predicted by Theorem~\ref{thm:asymp_proof} is displayed by the dashed line.
+Provided the SNR is large enough (so that the `threshold' is avoided) the sample MSE of the LSU estimator is close to that predicted by Theorem~\ref{thm:asymp_proof}.
+Using Theorem~\ref{thm:asymp_proof} we can predict that, asymptotically in $N$, the ratio of the variance of the LSU estimator to the CRB will approach $1$ for high SNR and $\tfrac{2}{3}\pi$ for low SNR~(see Lemma~3 and the subsequent discussion in~\cite{Quinn2009_dasp_phase_only_information_loss}).
+Contrastingly, the ratio of the variance of the HAF, PHAF and CPF estimators to the CRB is known to diverge as the SNR becomes small~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.
+%For this reason we can expect the LSU estimator to be more accurate at low SNR for sufficiently large $N$.
+
+In Figure~\ref{plot:m3small} the CPF estimator reaches a threshold at \unit[1]{dB}, the LSU and PHAF estimators at \unit[3]{dB} and the HAF estimator at \unit[5]{dB}.
+In Figure~\ref{plot:m3small} the true coefficients are $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}  \right]$.
+These are specifically chosen to satisfy the requirements of the HAF, PHAF and CPF estimators.
+The HAF and PHAF estimators require that the coefficients satisfy $\tilde{\mu}_0, \tilde{\mu}_1 \in [-0.5, 0.5)$ and
+\[
+\tilde{\mu}_k \in \left[ -\frac{0.5}{k! \tau^{k-1}}, \frac{0.5}{k! \tau^{k-1}}\right) \qquad k = 2,3,\dots,m,
+\]
+where $\tau$ is the \emph{lag} and we choose $\tau = \floor{N/m}$ as suggested in~\cite{Peleg_DPT_1995,Barbarossa_PHAF_1998}.
+The CPF estimator requires similar, but slightly more complicated conditions on the coefficients~\cite{Oshea_cpf_2004}.
+It is informative to compare the volume of the set of coefficients suitable for the HAF, PHAF and CPF estimators with the volume of the identifiable region $B$ from~\eqref{eq:rectangular_identifiable_region}.
+The identifiable region has volume $\vol(B) = \prod_{k=0}^m {\frac{1}{k!}}$.
+The volume of coefficients suitable for the HAF and PHAF estimators is
+\[
+\prod_{k=2}^m \frac{1}{k! \tau^{k-1}} =  \vol(B) \prod_{k=2}^m \frac{1}{\tau^{k-1}} \leq \vol(B) \prod_{k=2}^m \left( \frac{m}{N} \right)^{k-1}.
+\]
+The volume of the coefficients suitable for the CPF estimator can be shown to be $\tfrac{3}{2N^3} = 18 \vol(B) N^{-3}$.
+It is now worth reconsidering the results in Figure~\ref{plot:m3small}.
+Substituting $m=3$ and $N=199$ in the formula above, we see that the HAF, PHAF and CPF estimators search a space of parameters that is less than $10^{-5}$ the size of the identifiable region searched by the LSU estimator.
+In this context we might expect the HAF, PHAF and CPF estimators to have thresholds at lower SNR than the LSU estimator.
+It is encouraging that the LSU estimator has comparable performance.
+To put these results further into context, Figure~\ref{plot:m3big} repeats the simulations in Figure~\ref{plot:m3small}, but where the true coefficients are $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$.
+The HAF, PHAF and CPF estimators fail consistently, as expected.
+The performance of the LSU estimator is unchanged, as expected in view of Theorem~\ref{thm:asymp_proof}.
+
+The comparatively small set of coefficients suitable for the HAF, PHAF and CPF estimators has been studied previously~\cite{Angeby_PPS_aliasing_2000,Djurovic_aliasing_2012,McKilliam2009IndentifiabliltyAliasingPolyphase,Zhou_lag_diverity_PPT_1997,Xia_dynamic_range_pps_2001,Morelande_bayes_unwrapping_2009_tsp} and can be a critical limitation for some applications, for example, interference suppression in the presence of a jammer~\cite{Djukanovic:2011:PMN:1950994.1951252,Djurovic_aliasing_2012}.
+Zhou and Wang~\cite{Zhou_lag_diverity_PPT_1997} describe a method based on the extended Euclidean algorithm that enables the HAF estimator to operate on the entire identifiable region $B$.
+Unfortunately this comes at the price of statistical accuracy.
+Figure~\ref{plot:m3bigzw} shows the performance of the Zhou and Wang estimator when the coefficients are $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$, the same as in Figure~\ref{plot:m3big}.
+The threshold for the estimator is approximately~\unit[29]{dB}.
+There does not appear to exist (at least in the literature) a similar method for increasing the range of coefficients for the PHAF and CPF estimators.
+
+Figure~\ref{plot:m5} displays the sample MSE of the highest order coefficient $\mu_5$  for a polynomial phase signal of order $m=5$.
+In this case the LSU and PHAF estimators reach a threshold at \unit[12]{dB}, the HAF estimator at \unit[18]{dB}, and the estimator that combines the CPF and HAF at \unit[10]{dB}.
+In this figure the true coefficients are $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}, \tfrac{1}{96N^3}, \tfrac{1}{480N^4}  \right]$ and these are chosen specifically to satisfy the requirement of the HAF, PHAF and CPF-HAF estimators.
+As in the case where $m=3$, the region of coefficients suitable for the HAF, PHAF and CPF-HAF estimators is much smaller that the identifiable region.
+The effect becomes more severe as $m$ increases.
+When $m=5$ and $N=199$ the space of coefficients searched by these estimators is approximately $10^{-16}$ the size of the identifiable region.
+This is a dramatic reduction in the range of detectable polynomial phase signals.
+Figure~\ref{plot:m5big} displays results under the same conditions as Figure~\ref{plot:m5} but with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}, \tfrac{1}{96}, \tfrac{1}{480}  \right]$.
+The conclusion is similar to that from Figure~\ref{plot:m3big}.
+The performance of the LSU estimator remains unchanged, but the HAF, PHAF and CPF-HAF estimators fail completely.
+The Zhou and Wang estimator can again be used to increase the range of parameters for the HAF, but simulations indicate that the threshold for the estimator is approximately~\unit[103]{dB}.
+
+Figure~\ref{plot:bench} shows the running time of the LSU estimator computed using the $K$-best method versus the HAF estimator for polynomial phase signals of order 3.
+The HAF estimator is computed using the fast Fourier transform and has complexity $O(N \log N)$.
+As described in Section~\ref{sec:lattice-theory} the LSU estimator computed using the $K$-best method requires in the worst case $O(N^3 \log N)$ operations.
+The $K$-best method tends to be faster when the SNR is large.
+This is due to the algorithm rapidly finding a nearest point, and subsequently terminating branches in its search tree early.
+On the other hand, when the SNR is small, all branches in the tree tend to be searched.
+Interestingly the LSU estimator computed using the $K$-best algorithm is comparable with the HAF when the SNR is large and when $N$ is approximately less than 100.
+The LSU estimator can be computed in a reasonable amount of time (a few minutes) when $N=1000$.
+The algorithm becomes prohibitively expensive for large $N$.
+Benchmarks have been performed with polynomial phase signals of order larger than $3$.
+The results are omitted due to space constraints.
+The running time of the LSU estimator does not appear to change significantly as $m$ increases.
+This is expected since a nearest point in a lattice of approximately the same dimension $N-m-1$ must be computed.
+The complexity of the HAF estimator grows linearly with $m$.
+
+\begin{figure}[p] 
+   	\centering 
+  		\includegraphics{code/data/plot3-1.mps} 
+  		\caption{Sample MSE for coefficient $\tilde{\mu}_3$ for a polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}  \right]$ and $N=199$.}
+  		\label{plot:m3small} 
+ \end{figure}
+
+\begin{figure}[p] 
+   	\centering 
+  		\includegraphics{code/data/plot3-2.mps} 
+  		\caption{Sample MSE for the coefficient $\tilde{\mu}_3$ for a polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$ and $N=199$.}
+  		\label{plot:m3big} 
+ \end{figure}
+
+\begin{figure}[p] 
+   	\centering 
+  		\includegraphics{code/data/plot3-3.mps} 
+  		\caption{Sample MSE for the coefficient $\tilde{\mu}_3$ for a polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$ and $N=199$.}
+  		\label{plot:m3bigzw} 
+ \end{figure}
+
+\begin{figure}[p] 
+   	\centering 
+  		\includegraphics{code/data/plot5-1.mps} 
+  		\caption{Sample MSE for the coefficient $\tilde{\mu}_5$ for a polynomial phase signal of order 5 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}, \tfrac{1}{96N^3}, \tfrac{1}{480N^4} \right]$ and $N=199$.}
+  		\label{plot:m5} 
+\end{figure}
+
+\begin{figure}[p] 
+   	\centering 
+  		\includegraphics{code/data/plot5-2.mps} 
+  		\caption{Sample MSE for the coefficient $\tilde{\mu}_5$ for a polynomial phase signal of order 5 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}, \tfrac{1}{96}, \tfrac{1}{480}   \right]$ and $N=199$.}
+  		\label{plot:m5big} 
+\end{figure}
+
+\begin{figure}[p] 
+   	\centering 
+  		\includegraphics{code/data/bench-1.mps} 
+  		\caption{Running time in seconds with the LSU and HAF estimators.  The LSU estimator is benchmarked at SNR equal to \unit[0]{dB}, \unit[5]{dB}, \unit[10]{dB} and \unit[20]{dB}.} 
+  		\label{plot:bench} 
+\end{figure}
+
+%\begin{figure*}[t] 
+%\begin{gnuplot}[terminal=epslatex]
+%plot sin(x), cos(x)
+%\end{gnuplot}
+%\end{figure*}
+
+% \section{Computational considerations}
+
+% The exact LSU estimator implemented using the sphere decoder can be computed quickly for any $m$ with $N$ approximately less than $50$.
+In situations where $N$ is small, but high statistical accuracy is required, the sphere decoder is computationally a better choice than the LSE for polynomial phase signals of order greater than or equal to $2$.
+
+% Babai's nearest plane algorithm requires $O(N^2)$ operations and can be run in a reasonable amount of time for any $m$ and reasonably large $N$.
+It should be noted that Babai's nearest plane algorithm requires a Lov\`as reduced lattice basis~\cite{Lenstra1982}.
+This requires $O(N^4)$ operations to compute, but, the Lov\`as reduced basis can be computed once offline.
+It does not need to be computed each time the estimator is run.
+The $K$-best algorithm runs in a reasonable amount of time for any $m$ and when $N$ is less than about $300$.
+If the noise variance is small, then both the sphere decoder and the $K$-best algorithm are actually fast.
+This is because both of these estimators begin with the lattice point found by Babai's nearest plane algorithm.
+If this point is close to (or \emph{is}) the nearest point then both the sphere decoder and the $K$-best algorithm terminate quickly.
+This is in contrast to the least squares estimator that is slow regardless of the noise variance.
+
+% Both Kitchen's estimator and the DPT are fast to compute.
+Kitchen's estimator requires only $O(N)$ arithmetic operations and the DPT requires $O(N \log N)$ operations.
+However, these estimators are not as accurate as the LSE or the LSU.
+They also fail to operate correctly over the entire identifiable region of polynomial phase coefficients~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}~\cite[Ch.~10]{McKilliam2010thesis}.
+
+%\section{Discussion}
+
+\section{Conclusion} \label{sec:conclusion}
+
+This paper has considered the estimation of the coefficients of a noisy polynomial phase signal by least squares phase unwrapping (LSU).
+The LSU estimator was shown to be strongly consistent and asymptotically normally distributed under mild conditions on the distribution of the noise.
+The estimator can be computed by finding a nearest lattice point in a lattice.
+Polynomial time nearest point algorithms for these lattices exist~\cite[Sec 4.3]{McKilliam2010thesis}, but these algorithms are not fast in practice.
+We instead make use of the approximate $K$-best algorithm~\cite{Baro_list_sphere_dec_2003,Zhan2006_K_best_sphere_decoder,Mondal_Kbest_again_2009}.
+This algorithm results in an accurate estimator and can be computed in a reasonable amount of time for moderately large numbers of observations $N$.
+
+A key property of the LSU estimator is that it works for polynomial phase parameters contained anywhere inside the identifiable region described in Section~\ref{sec:ident_aliasing}.
+This is the maximum possible range of parameters.
+By contrast popular estimators such as the HAF and CPF work only for parameters in a range much smaller than the identifiable region.
+For this reason, the LSU estimator may be the only presently existing estimator that is viable for certain applications, such as interference suppression~\cite{Djukanovic:2011:PMN:1950994.1951252,Djurovic_aliasing_2012}.
+
+The major outstanding question is whether faster nearest point algorithms exist for these specific lattices, particularly for the case when $N$ is large.
+Considering the excellent statistical performance (both theoretically and practically) of the LSU estimator, even fast \emph{approximate} nearest point algorithms are likely to prove useful for the estimation of polynomial phase signals.
+
+\nocite{OShea_refiningpps_2010}
+
+\appendix
+
+\subsection{Proof of strong consistency}\label{sec:strongconstproof}
  Substituting \eqref{eq:noise_circ_poly} into $SS$ we obtain
@@
 
-\section{Simulations}\label{sec:simulations}
-
-% This section describes the results of Monte-Carlo simulations involving the least squares unwrapping estimator (LSU), the least squares estimator (LSE), Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, and the discrete polynomial phase transform (DPT)~\cite{Peleg_DPT_1995}.
-In all simulations the sample sizes considered are $N = 10, 50$ and $200$ and the unknown amplitude is $\rho = 1$.
-The number of replications of each experiment is 2000 and the $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ are distributed uniformly randomly over the identifiable region.
-
-% Figure~\ref{plot:lsetest} shows the performance of the LSU estimator and LSE when $m=2$ (linear chirp).
-Only results for the highest order parameter $\widehat{\mu}_2$ are displayed, the results for the lower order parameters leading to similar conclusions.
-The LSE is computed using the approach described in~\cite[Sec.~10.1]{McKilliam2010thesis} which is based on that of Abatzoglou~\cite{Abatzoglou_ml_chirp_1986}.
-The LSE is computationally intensive.
-We were able to compute it within a reasonable amount of time only with $N$ less than $50$ when $m=2$.
-Results for the exact LSU estimator computed using the sphere decoder and the approximate LSU estimator computed using Babai's algorithm are displayed.
-The sphere decoder is computationally intractable when $N = 200$.
-It is evident that a significant performance penalty is suffered by using Babai's algorithm when the noise variance is large.
-We have also displayed results using another approximate nearest point algorithm called the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder}.
-Our implementation requires $O(N^3\log{N})$ operations.
-This is more computationally expensive than Babai's algorithm but, as can be seen, the $K$-best method is more statistically accurate.
-
-% For the simulations displayed in the left hand plot of Figure~\ref{plot:lsetest} the signal model used is that from~\eqref{eq:Y_nsamplednoisey}.
-The noise terms $X_1, \dots, X_N$ are independent and identically distributed zero mean complex Gaussian random variables with independent real and imaginary parts having variance $\sigma_c^2$.
-In this case the LSE is also the maximum likelihood estimator.
-The \term{Cram\`{e}r-Rao lower bound} (CRB) for the variance of unbiased estimators of $\tilde{\mu}_j$ is given by Peleg and Porat~\cite{Peleg1991_CRB_PPS_1991} to be
-% \[ 
-% \frac{\sigma_c^2}{4\pi^2 N^{2j + 1}} C^{-1}_{jj} + o(N^{-2j-1})
-% \]
-% where $C^{-1}_{jj}$ is the $j$th diagonal element of the inverse of the $m+1$ by $m+1$ Hilbert matrix $\Cbf$.
-Observe that the performance of the exact LSU estimator (computed using the sphere decoder) is close to the maximum likelihood LSE.
-The asymptotic Cramer-Rao bound (CRB) (solid line) and the asymptotic variance predicted in Theorem~\ref{thm:asymp_proof} (dashed line) are also shown.
-Provided the noise variance is small enough (so that the `threshold' is avoided) the LSE variance is close to the CRB (as is expected) and the LSU variance is close to that predicted by Theorem~\ref{thm:asymp_proof}.
-% %Note the gap between the CRB and the variance predicted by Theorem~\ref{thm:asymp_proof}.
-In practice this gap can be overcome using a numerical technique, such as Newton's method, with the LSU estimate as a starting point.
-We have not done this here in order to show the accuracy of Theorem~\ref{thm:asymp_proof}.
-The other noticeable gap is that between the threshold behaviour of the LSU estimator and the LSE.
-When $N=50$ the threshold \emph{kicks in} a little earlier for the LSE than it does for the LSU.
-We suspect that this gap will actually \emph{close} as $N$ increases.
-This hypothesis is supported by the fact that this behaviour is observed in the case of frequency estimation (when $m=1$)~\cite[Sec.
-9.5]{McKilliam2010thesis}\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.
-Unfortunately it is not feasible to test this hypothesis because of the computational complexity of the LSE.
-% For the simulations displayed in the right hand plot of Figure~\ref{plot:lsetest} the signal model used is that from~\eqref{eq:noise_circ_poly}.
-The phase noise terms $\Phi_1, \dots, \Phi_N$ are uniform random variables with variance $\sigma^2$.
-In this case, the LSU estimator performs slightly better than the LSE.
-This result agrees with similar observations made in circular statistics~\cite{McKilliam_mean_dir_est_sq_arc_length2010}.
-The asymptotic variance predicted by Theorem~\ref{thm:asymp_proof} again accurately predicts the performance of the LSU estimator.
-
-% Figure~\ref{plot:polyest} displays the performance of the LSU estimator, the DPT and Kitchen's estimator when $m=3$.
-Both the DPT and Kitchen's estimator perform poorly because they do not work correctly for some parameters in the identifiable region $B$.
-This behaviour has been observed previously~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase,McKilliam_polyphase_est_icassp_2011}.
-Again the performance of the LSU is well predicted by Theorem~\ref{thm:asymp_proof} provided that the noise variance is small enough to avoid the threshold effect and provided that $N$ is sufficiently large.
-The asymptotic CRB is also plotted.
-
-This section describes the results of Monte-Carlo simulations with the least squares unwrapping (LSU) estimator, the high order ambiguity function (HAF) estimator~\cite{Peleg_DPT_1995}, the product high order ambiguity function (PHAF) estimator~\cite{Barbarossa_PHAF_1998}, the cubic phase function (CPF) estimator~\cite{Oshea_cpf_2004} and the estimator that results from combining the HAF and CPF~\cite{Djurovic_haf_cpf_2012}.
-The CPF estimator only applies to polynomial phase signals of order $m=3$ and when the number of observations $N$ is odd.
-In all simulations the unknown amplitude is $\rho = 1$ and $N = 199$.
-The $X_1, \dots, X_N$ are pseudorandomly generated independent and identically distributed circularly symmetric complex Gaussian with variance $\expect\abs{X_1}^2 = \sigma_c^2$.
-The corresponding signal to noise ratio (SNR) is $\rho^2 \sigma_c^{-2} = \sigma_c^{-2}$.
-The number of replications of each Monte-Carlo experiment is $T = 2000$ to obtain estimates $\widehat{\mubf}_1, \dots, \widehat{\mubf}_T$ and the corresponding dealiased errors $\widehat{\lambdabf}_t = \dealias(\tilde{\mubf} - \widehat{\mubf}_t)$ are computed.
-The sample mean square error (MSE) of the $k$th coefficient is computed according to $\tfrac{1}{T}\sum_{t=1}^T \widehat{\lambda}_{k,t}^2$ where $\widehat{\lambda}_{k,t}$ is the $k$th element of $\widehat{\lambdabf}_t$.
-In all simulations four \emph{lag sets} are used for the PHAF estimator~\cite{Barbarossa_PHAF_1998} and the LSU estimator is approximated using the $K$-best method described in Section~\ref{sec:lattice-theory}.
-The maximum number of branches allowed is $K=20N$.
-
-Figure~\ref{plot:m3small} shows the sample MSE of the highest order coefficient $\mu_3$ for a polynomial phase signal of order $m=3$.
-All the estimators perform poorly at low SNR until a `threshold' is reached, after which the variance is close to the Cramer-Rao bound (CRB)~\cite{Peleg1991_CRB_PPS_1991} indicated by the solid line.
-The asymptotic variance of the LSU estimator predicted by Theorem~\ref{thm:asymp_proof} is displayed by the dashed line.
-Provided the SNR is large enough (so that the `threshold' is avoided) the sample MSE of the LSU estimator is close to that predicted by Theorem~\ref{thm:asymp_proof}.
-Using Theorem~\ref{thm:asymp_proof} we can predict that, asymptotically in $N$, the ratio of the variance of the LSU estimator to the CRB will approach $1$ for high SNR and $\tfrac{2}{3}\pi$ for low SNR~(see Lemma~3 and the subsequent discussion in~\cite{Quinn2009_dasp_phase_only_information_loss}).
-Contrastingly, the ratio of the variance of the HAF, PHAF and CPF estimators to the CRB is known to diverge as the SNR becomes small~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.
-%For this reason we can expect the LSU estimator to be more accurate at low SNR for sufficiently large $N$.
-
-In Figure~\ref{plot:m3small} the CPF estimator reaches a threshold at \unit[1]{dB}, the LSU and PHAF estimators at \unit[3]{dB} and the HAF estimator at \unit[5]{dB}.
-In Figure~\ref{plot:m3small} the true coefficients are $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}  \right]$.
-These are specifically chosen to satisfy the requirements of the HAF, PHAF and CPF estimators.
-The HAF and PHAF estimators require that the coefficients satisfy $\tilde{\mu}_0, \tilde{\mu}_1 \in [-0.5, 0.5)$ and
-\[
-\tilde{\mu}_k \in \left[ -\frac{0.5}{k! \tau^{k-1}}, \frac{0.5}{k! \tau^{k-1}}\right) \qquad k = 2,3,\dots,m,
-\]
-where $\tau$ is the \emph{lag} and we choose $\tau = \floor{N/m}$ as suggested in~\cite{Peleg_DPT_1995,Barbarossa_PHAF_1998}.
-The CPF estimator requires similar, but slightly more complicated conditions on the coefficients~\cite{Oshea_cpf_2004}.
-It is informative to compare the volume of the set of coefficients suitable for the HAF, PHAF and CPF estimators with the volume of the identifiable region $B$ from~\eqref{eq:rectangular_identifiable_region}.
-The identifiable region has volume $\vol(B) = \prod_{k=0}^m {\frac{1}{k!}}$.
-The volume of coefficients suitable for the HAF and PHAF estimators is
-\[
-\prod_{k=2}^m \frac{1}{k! \tau^{k-1}} =  \vol(B) \prod_{k=2}^m \frac{1}{\tau^{k-1}} \leq \vol(B) \prod_{k=2}^m \left( \frac{m}{N} \right)^{k-1}.
-\]
-The volume of the coefficients suitable for the CPF estimator can be shown to be $\tfrac{3}{2N^3} = 18 \vol(B) N^{-3}$.
-It is now worth reconsidering the results in Figure~\ref{plot:m3small}.
-Substituting $m=3$ and $N=199$ in the formula above, we see that the HAF, PHAF and CPF estimators search a space of parameters that is less than $10^{-5}$ the size of the identifiable region searched by the LSU estimator.
-In this context we might expect the HAF, PHAF and CPF estimators to have thresholds at lower SNR than the LSU estimator.
-It is encouraging that the LSU estimator has comparable performance.
-To put these results further into context, Figure~\ref{plot:m3big} repeats the simulations in Figure~\ref{plot:m3small}, but where the true coefficients are $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$.
-The HAF, PHAF and CPF estimators fail consistently, as expected.
-The performance of the LSU estimator is unchanged, as expected in view of Theorem~\ref{thm:asymp_proof}.
-
-The comparatively small set of coefficients suitable for the HAF, PHAF and CPF estimators has been studied previously~\cite{Angeby_PPS_aliasing_2000,Djurovic_aliasing_2012,McKilliam2009IndentifiabliltyAliasingPolyphase,Zhou_lag_diverity_PPT_1997,Xia_dynamic_range_pps_2001,Morelande_bayes_unwrapping_2009_tsp} and can be a critical limitation for some applications, for example, interference suppression in the presence of a jammer~\cite{Djukanovic:2011:PMN:1950994.1951252,Djurovic_aliasing_2012}.
-Zhou and Wang~\cite{Zhou_lag_diverity_PPT_1997} describe a method based on the extended Euclidean algorithm that enables the HAF estimator to operate on the entire identifiable region $B$.
-Unfortunately this comes at the price of statistical accuracy.
-Figure~\ref{plot:m3bigzw} shows the performance of the Zhou and Wang estimator when the coefficients are $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$, the same as in Figure~\ref{plot:m3big}.
-The threshold for the estimator is approximately~\unit[29]{dB}.
-There does not appear to exist (at least in the literature) a similar method for increasing the range of coefficients for the PHAF and CPF estimators.
-
-Figure~\ref{plot:m5} displays the sample MSE of the highest order coefficient $\mu_5$  for a polynomial phase signal of order $m=5$.
-In this case the LSU and PHAF estimators reach a threshold at \unit[12]{dB}, the HAF estimator at \unit[18]{dB}, and the estimator that combines the CPF and HAF at \unit[10]{dB}.
-In this figure the true coefficients are $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}, \tfrac{1}{96N^3}, \tfrac{1}{480N^4}  \right]$ and these are chosen specifically to satisfy the requirement of the HAF, PHAF and CPF-HAF estimators.
-As in the case where $m=3$, the region of coefficients suitable for the HAF, PHAF and CPF-HAF estimators is much smaller that the identifiable region.
-The effect becomes more severe as $m$ increases.
-When $m=5$ and $N=199$ the space of coefficients searched by these estimators is approximately $10^{-16}$ the size of the identifiable region.
-This is a dramatic reduction in the range of detectable polynomial phase signals.
-Figure~\ref{plot:m5big} displays results under the same conditions as Figure~\ref{plot:m5} but with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}, \tfrac{1}{96}, \tfrac{1}{480}  \right]$.
-The conclusion is similar to that from Figure~\ref{plot:m3big}.
-The performance of the LSU estimator remains unchanged, but the HAF, PHAF and CPF-HAF estimators fail completely.
-The Zhou and Wang estimator can again be used to increase the range of parameters for the HAF, but simulations indicate that the threshold for the estimator is approximately~\unit[103]{dB}.
-
-Figure~\ref{plot:bench} shows the running time of the LSU estimator computed using the $K$-best method versus the HAF estimator for polynomial phase signals of order 3.
-The HAF estimator is computed using the fast Fourier transform and has complexity $O(N \log N)$.
-The $K$-best method tends to be faster when the SNR is large.
-This is due to the algorithm rapidly finding a nearest point, and subsequently terminating branches in its search tree early.
-On the other hand, when the SNR is small, all branches in the tree tend to be searched.
-Interestingly the LSU estimator computed using the $K$-best algorithm can be \emph{faster} than the HAF when the SNR is large and when $N$ is approximately less than 100.
-The LSU estimator can be computed in a reasonable amount of time (a few minutes) when $N=1000$.
-The algorithm becomes prohibitively expensive for large $N$.
-Benchmarks have been performed with polynomial phase signals of order larger than $3$.
-The results are omitted due to space constraints.
-The running time of the LSU estimator does not appear to change significantly as $m$ increases.
-This is expected since a nearest point in a lattice of approximately the same dimension $N-m-1$ must be computed.
-The complexity of the HAF estimator grows linearly with $m$.
-
-\begin{figure}[p] 
-   	\centering 
-  		\includegraphics{code/data/plot3-1.mps} 
-  		\caption{Sample MSE for coefficient $\tilde{\mu}_3$ for a polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}  \right]$ and $N=199$.}
-  		\label{plot:m3small} 
- \end{figure}
-
-\begin{figure}[p] 
-   	\centering 
-  		\includegraphics{code/data/plot3-2.mps} 
-  		\caption{Sample MSE for the coefficient $\tilde{\mu}_3$ for a polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$ and $N=199$.}
-  		\label{plot:m3big} 
- \end{figure}
-
-\begin{figure}[p] 
-   	\centering 
-  		\includegraphics{code/data/plot3-3.mps} 
-  		\caption{Sample MSE for the coefficient $\tilde{\mu}_3$ for a polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$ and $N=199$.}
-  		\label{plot:m3bigzw} 
- \end{figure}
-
-\begin{figure}[p] 
-   	\centering 
-  		\includegraphics{code/data/plot5-1.mps} 
-  		\caption{Sample MSE for the coefficient $\tilde{\mu}_5$ for a polynomial phase signal of order 5 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}, \tfrac{1}{96N^3}, \tfrac{1}{480N^4} \right]$ and $N=199$.}
-  		\label{plot:m5} 
-\end{figure}
-
-\begin{figure}[p] 
-   	\centering 
-  		\includegraphics{code/data/plot5-2.mps} 
-  		\caption{Sample MSE for the coefficient $\tilde{\mu}_5$ for a polynomial phase signal of order 5 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}, \tfrac{1}{96}, \tfrac{1}{480}   \right]$ and $N=199$.}
-  		\label{plot:m5big} 
-\end{figure}
-
-\begin{figure}[p] 
-   	\centering 
-  		\includegraphics{code/data/bench-1.mps} 
-  		\caption{Running time in seconds with the LSU and HAF estimators.  The LSU estimator is benchmarked at SNR equal to \unit[0]{dB}, \unit[5]{dB}, \unit[10]{dB} and \unit[20]{dB}.} 
-  		\label{plot:bench} 
-\end{figure}
-
-%\begin{figure*}[t] 
-%\begin{gnuplot}[terminal=epslatex]
-%plot sin(x), cos(x)
-%\end{gnuplot}
-%\end{figure*}
-
-% \section{Computational considerations}
-
-% The exact LSU estimator implemented using the sphere decoder can be computed quickly for any $m$ with $N$ approximately less than $50$.
-In situations where $N$ is small, but high statistical accuracy is required, the sphere decoder is computationally a better choice than the LSE for polynomial phase signals of order greater than or equal to $2$.
-
-% Babai's nearest plane algorithm requires $O(N^2)$ operations and can be run in a reasonable amount of time for any $m$ and reasonably large $N$.
-It should be noted that Babai's nearest plane algorithm requires a Lov\`as reduced lattice basis~\cite{Lenstra1982}.
-This requires $O(N^4)$ operations to compute, but, the Lov\`as reduced basis can be computed once offline.
-It does not need to be computed each time the estimator is run.
-The $K$-best algorithm runs in a reasonable amount of time for any $m$ and when $N$ is less than about $300$.
-If the noise variance is small, then both the sphere decoder and the $K$-best algorithm are actually fast.
-This is because both of these estimators begin with the lattice point found by Babai's nearest plane algorithm.
-If this point is close to (or \emph{is}) the nearest point then both the sphere decoder and the $K$-best algorithm terminate quickly.
-This is in contrast to the least squares estimator that is slow regardless of the noise variance.
-
-% Both Kitchen's estimator and the DPT are fast to compute.
-Kitchen's estimator requires only $O(N)$ arithmetic operations and the DPT requires $O(N \log N)$ operations.
-However, these estimators are not as accurate as the LSE or the LSU.
-They also fail to operate correctly over the entire identifiable region of polynomial phase coefficients~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}~\cite[Ch.~10]{McKilliam2010thesis}.
-
-%\section{Discussion}
-
-\section{Conclusion} \label{sec:conclusion}
-
-This paper has considered the estimation of the coefficients of a noisy polynomial phase signal by least squares phase unwrapping (LSU).
-The LSU estimator was shown to be strongly consistent and asymptotically normally distributed under mild conditions on the distribution of the noise.
-The estimator can be computed by finding a nearest lattice point in a lattice.
-Polynomial time nearest point algorithms for these lattices exist~\cite[Sec 4.3]{McKilliam2010thesis}, but these algorithms are not fast in practice.
-We instead make use of the approximate $K$-best algorithm~\cite{Baro_list_sphere_dec_2003,Zhan2006_K_best_sphere_decoder,Mondal_Kbest_again_2009}.
-This algorithm results in an accurate estimator and can be computed in a reasonable amount of time for moderately large numbers of observations $N$.
-
-A key property of the LSU estimator is that it works for polynomial phase parameters contained anywhere inside the identifiable region described in Section~\ref{sec:ident_aliasing}.
-This is the maximum possible range of parameters.
-By contrast popular estimators such as the HAF and CPF work only for parameters in a range much smaller than the identifiable region.
-For this reason, the LSU estimator may be the only presently existing estimator that is viable for certain applications, such as interference suppression~\cite{Djukanovic:2011:PMN:1950994.1951252,Djurovic_aliasing_2012}.
-
-The major outstanding question is whether faster nearest point algorithms exist for these specific lattices, particularly for the case when $N$ is large.
-Considering the excellent statistical performance (both theoretically and practically) of the LSU estimator, even fast \emph{approximate} nearest point algorithms are likely to prove useful for the estimation of polynomial phase signals.
-
-\nocite{OShea_refiningpps_2010}
-
 \small 

### 2013-06-28 code/java/src/pubsim/poly/PHAF.java
minor

@@ -166,7 +166,7 @@ public class PHAF extends AbstractPolynomialPhaseEstimator {
         }
 
         /**
-         * Compute the product of hafs using the FFT
+         * Compute the product of HAFs using the FFT
          */
         public Complex[] computeFFTproduct(Complex[] x, int M) {
             Complex[] fftprod = hafs[0].FFTHAF(x, M);

### 2013-06-29 paper.tex
minor
added to pps, dumb tables

@@
 Under mild conditions on the distribution of the noise we describe the asymptotic properties of this estimator, showing that it is strongly consistent and asymptotically normally distributed.
-No `high SNR' assumption (common in the literature) is required in our analysis.
+%No `high SNR' assumption (common in the literature) is required in our analysis.
 As a result the LSU estimator is accurate, even for very noisy signals, if the number of observations is sufficiently large.
@@
 
-%\bibliographystyle{unsrt}
+% \bibliographystyle{unsrt}
 
@@
 In this paper we consider the estimator that results from unwrapping the phase in a least squares manner.
-We call this the \emph{least squares unwrapping} (LSU) estimator~\cite{McKilliam2009asilomar_polyest_lattice, McKilliamFrequencyEstimationByPhaseUnwrapping2009,McKilliam2010thesis}.
+We call this the \emph{least squares unwrapping} (LSU) estimator~\cite{McKilliam2009asilomar_polyest_lattice, McKilliamFrequencyEstimationByPhaseUnwrapping2009,McKilliam2010thesis,McKilliam_polyphase_est_icassp_2011}.
 It was shown in \cite{McKilliam2009asilomar_polyest_lattice, McKilliamFrequencyEstimationByPhaseUnwrapping2009} that the LSU estimator can be represented as a \emph{nearest lattice point problem}~\cite{Agrell2002}, and Monte-Carlo simulations were used to show the LSU estimator's favourable statistical performance.
@@
 Although polynomial time, the algorithms are still computationally demanding in practice.
-Nevertheless fast (exact or approximate) algorithms may exist for these special lattices, that are yet to be discovered.
+Fast (exact or approximate) algorithms may exist for these special lattices, that are yet to be discovered.
 In this paper we apply another general purpose approximate nearest point algorithm called the $K$-best algorithm~\cite{Baro_list_sphere_dec_2003,Zhan2006_K_best_sphere_decoder,Mondal_Kbest_again_2009} for approximating the LSU estimator.
@@
 Section~\ref{sec:least-squar-unwr} describes the LSU estimator and states a theorem asserting the estimator to be strongly consistent and asymptotically normally distributed under some mild assumptions on the noise process $\{X_n\}$.
-The proof of strong consistency is given in Section~\ref{sec:strongconstproof}.
+The proof of strong consistency is given in the appendix.
 Space restrictions make it impossible to give a proof of asymptotic normality in this paper.
@@
 The algorithms vary in the order that branches in the tree are searched.
-A popular approach due to Schnorr and Euchner~\cite{schnorr_euchner_sd_1994,Agrell2002} searches branches in a greedy order, and this is the approach we implement here.
+A popular approach due to Schnorr and Euchner~\cite{schnorr_euchner_sd_1994,Agrell2002} searches branches in a greedy order.
 We refer to this algorithm as the~\emph{sphere decoder} in this paper, although other variants of the algorithm also go by this name.
@@
 The maximum number of branches $K$ is free to be chosen.
-For polynomial phase estimation we have found that setting $K=10N$ yields excellent results, and in this case the algorithm requires $O(N^3 \log N)$ operations.
+For polynomial phase estimation we have found that setting $K=20N$ yields excellent results, and in this case the algorithm requires $O(N^3 \log N)$ operations.
 Simulations in Section~\ref{sec:simulations} show that this algorithm produces excellent estimates of polynomial phase coefficients and can be computed in a reasonable amount of time for $N$ less than about 1000.
@@
 
+Tables
+
+\begin{table}[htbp]
+\centering
+\begin{tabular}{lrrrrr}
+ & LSU & CPF & PHAF & HAF & ZW \\ \toprule
+$N=99$ &  &  &  &  & \\ 
+$N=199$ & 3 & 1 & 3 & 5 & 23 \\ 
+$N=299$ &  &  &  & & \\ 
+$N=399$ &  &  &  & & \\ 
+$N=499$ &  &  &  & & \\ 
+$N\to\infty$ & $-\infty$ & ? & ? & 3.77  & ? \\ \bottomrule
+\end{tabular}
+\caption{Threshold SNR in decibels for estimators with $m=3$}
+\label{tablm3}
+\end{table}
+
+\begin{table}[htbp]
+\centering
+\begin{tabular}{lrrrrr}
+ & LSU & CPF-HAF & PHAF & HAF & ZW \\ \toprule
+$N=99$ &  &  &  &  & \\ 
+$N=199$ & 12 & 10 & 12 & 18 & 103 \\ 
+$N=299$ &  &  &  & & \\ 
+$N=399$ &  &  &  & & \\ 
+$N=499$ &  &  &  & & \\ 
+$N\to\infty$ & $-\infty$ & $\approx 6.5$ & ?  & 16.6 & ? \\ \bottomrule
+\end{tabular}
+\caption{Threshold SNR in decibels for estimators with $m=5$}
+\label{tablm3}
+\end{table}
+
 Figure~\ref{plot:bench} shows the running time of the LSU estimator computed using the $K$-best method versus the HAF estimator for polynomial phase signals of order 3.
 The HAF estimator is computed using the fast Fourier transform and has complexity $O(N \log N)$.
-As described in Section~\ref{sec:lattice-theory} the LSU estimator computed using the $K$-best method requires in the worst case $O(N^3 \log N)$ operations.
+Wtih $K=20N$ the LSU estimator computed using the $K$-best method requires $O(N^3 \log N)$ operations in the worst case.
 The $K$-best method tends to be faster when the SNR is large.
@@
 
-\nocite{OShea_refiningpps_2010}
+\nocite{OShea_refiningpps_2010,Amar_low_complex_polyphase_2010}
 

### 2013-06-30 paper.tex
added to tables

@@
 
-Tables
+Tables~\ref{tablm3} and~\ref{tablm5} \dots BLERG
 
@@
  & LSU & CPF & PHAF & HAF & ZW \\ \toprule
-$N=99$ &  &  &  &  & \\ 
+$N=49$ &  &  &  &  &  \\ 
+$N=99$ & 6 & 3 & 5 & 7 & 27 \\ 
 $N=199$ & 3 & 1 & 3 & 5 & 23 \\ 
-$N=299$ &  &  &  & & \\ 
-$N=399$ &  &  &  & & \\ 
-$N=499$ &  &  &  & & \\ 
+$N=299$ & 2 & -1 &  & 4 & \\ 
 $N\to\infty$ & $-\infty$ & ? & ? & 3.77  & ? \\ \bottomrule
 \end{tabular}
-\caption{Threshold SNR in decibels for estimators with $m=3$}
+\caption{Estimated SNR threshold in decibels with $m=3$}
 \label{tablm3}
@@
  & LSU & CPF-HAF & PHAF & HAF & ZW \\ \toprule
-$N=99$ &  &  &  &  & \\ 
+$N=49$ &  &  &  & &  \\ 
+$N=99$ &  15 &  12 & 13 &  19 & $>100$ \\ 
 $N=199$ & 12 & 10 & 12 & 18 & 103 \\ 
 $N=299$ &  &  &  & &  \\ 
-$N=399$ &  &  &  & & \\ 
-$N=499$ &  &  &  & & \\ 
 $N\to\infty$ & $-\infty$ & $\approx 6.5$ & ?  & 16.6 & ? \\ \bottomrule
 \end{tabular}
-\caption{Threshold SNR in decibels for estimators with $m=5$}
-\label{tablm3}
+\caption{Estimated SNR threshold in decibels with $m=5$}
+\label{tablm5}
 \end{table}

### 2013-07-01 paper.tex
added to plots and paper

@@
 
-Tables~\ref{tablm3} and~\ref{tablm5} \dots BLERG
-
-\begin{table}[htbp]
+Tables~\ref{tablm1},~\ref{tablm3} and~\ref{tablm5} show the values of SNR at which the estimators reach a threshold for polynomial phase signals of order $m=1,3$ and $5$.
+The values in the tables are approximated using Monte-Carlo simulation with $X_1, \dots, X_N$ pseudorandomly generated independent and identically distributed circularly symmetric complex Gaussian.
+When $m=3$ the true coefficients are as in Figure~\ref{plot:m3small} and when $m=5$ the true coefficients are as in Figure~\ref{plot:m5} so that the coefficients are inside the range acceptable for the HAF, CPF and CPF-HAF estimators.
+When $m=1$ (single frequency estimation) the LSU estimator can be computed exactly using the algorithm from~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.
+In this case the value at which the LSU estimator reaches a threshold is similar to the maximum likelihood periodogram estimator~\cite{Quinn2001,Hannan1973,Rife1974}.
+This property was observed in~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.
+Both the LSU and periodogram estimators are strongly consistent, and so, the threshold values can be made arbitrarily small by choosing $N$ large.
+This is indicated in the bottom row of Table~\ref{tablm1} by the $-\infty$ symbols.
+When $m>1$ the LSU estimator can be exactly computed using the sphere decoder~\cite{Agrell2002,Viterbo_sphere_decoder_1999} in a reasonable amount of time when $N < 60$.
+When $N > 60$ we use the approximate $K$-best algorithm.
+We observe that the $K$-best algorithm produces estimates with the same threshold as the exactly computed LSU estimator when $N = 19$ and $49$.
+As $N\rightarrow\infty$ the asymptotic threshold of the HAF estimator has been studied in~\cite{Porat_asympt_HAF_DPT_1996} and these values are indicated in the bottom rows of tables~\ref{tablm3} and~\ref{tablm5}.
+By Theorem~\ref{thm:asymp_proof} the LSU estimator is strongly consistent and the threshold can be made as small as desired by choosing $N$ large.
+Whether the LSU estimator computed approximately using the $K$-best algorithm is strongly consistent is an open theoretical question.
+That the $K$-best algorithm provides a close approximation of the exact LSU estimator for small $N$ is an encouraging observation.
+
+\begin{table}[t]
+\centering
+\begin{tabular}{lrr}
+ & LSU & ML  \\ \toprule
+$N=16$ & 5 & 2  \\ 
+$N=64$ & -2 & -4 \\ 
+$N=256$ & -6 & -8 \\ 
+$N=1024$ & -14 & -14 \\ 
+$N\to\infty$ & $-\infty$ & $-\infty$  \\ \bottomrule
+\end{tabular}
+\caption{Estimated SNR threshold in decibels with $m=1$}
+\label{tablm1}
+\end{table}
+
+\begin{table}[t]
 \centering
 \begin{tabular}{lrrrrr}
- & LSU & CPF & PHAF & HAF & ZW \\ \toprule
-$N=49$ &  &  &  &  &  \\ 
-$N=99$ & 6 & 3 & 5 & 7 & 27 \\ 
-$N=199$ & 3 & 1 & 3 & 5 & 23 \\ 
-$N=299$ & 2 & -1 &  & 4 & \\ 
-$N\to\infty$ & $-\infty$ & ? & ? & 3.77  & ? \\ \bottomrule
+ & LSU & LSU ($K$-best) & CPF & HAF & ZW \\ \toprule
+$N=19$ & 10 & 10 & 8 & 11 & 31 \\ 
+$N=49$ & 7 & 7 & 5 & 9 & 29 \\ 
+$N=99$ & & 6 & 3 & 5 & 27 \\ 
+$N=199$ & & 3 & 1 & 3 & 23 \\ 
+$N=299$ & & 2 & -1 &  & \\ 
+$N\to\infty$ & $-\infty$ & & & 3.77 & \\ \bottomrule
 \end{tabular}
@@
 
-\begin{table}[htbp]
+\begin{table}[t]
 \centering
 \begin{tabular}{lrrrrr}
- & LSU & CPF-HAF & PHAF & HAF & ZW \\ \toprule
-$N=49$ &  &  &  & &  \\ 
-$N=99$ &  15 &  12 & 13 &  19 & $>100$ \\ 
-$N=199$ & 12 & 10 & 12 & 18 & 103 \\ 
+ & LSU & LSU ($K$-best) & CPF-HAF & HAF & ZW \\ \toprule
+$N=19$ & 22  & 22 & 17 & 23 & $>100$ \\ 
+$N=49$ & 17 & 17 & 14 & 21 & $>100$  \\ 
+$N=99$ &  & 15 &  12 &  19 & $>100$ \\ 
+$N=199$ & & 12 & 10 & 18 & 103 \\ 
 $N=299$ &  &  &  & &  \\ 
-$N\to\infty$ & $-\infty$ & $\approx 6.5$ & ?  & 16.6 & ? \\ \bottomrule
+$N\to\infty$ & $-\infty$ &  &  & 16.6 &  \\ \bottomrule
 \end{tabular}
@@
 The HAF estimator is computed using the fast Fourier transform and has complexity $O(N \log N)$.
-Wtih $K=20N$ the LSU estimator computed using the $K$-best method requires $O(N^3 \log N)$ operations in the worst case.
+With $K=20N$ the LSU estimator computed using the $K$-best method requires $O(N^3 \log N)$ operations in the worst case.
 The $K$-best method tends to be faster when the SNR is large.

### 2013-07-01 code/java/src/pubsim/poly/SphereDecoder.java

@@ -1,8 +1,3 @@
-/*
- * To change this template, choose Tools | Templates
- * and open the template in the editor.
- */
-
 package pubsim.poly;
 
 import pubsim.lattices.decoder.SphereDecoderSchnorrEuchner;

### 2013-07-06 paper.tex
minor

@@
 The paper is organised in the following way.
-Section~\ref{sec:lattice-theory} describes some required concepts from lattice theory.
-In Section~\ref{sec:ident_aliasing} we describe the identifiable region that was derived in~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}.
+Section~\ref{sec:lattice-theory} describes some required concepts from lattice theory and Section~\ref{sec:ident_aliasing} describes the identifiable region that was derived in~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}.
 These identifiability results are required in order to understand the statistical properties of the LSU estimator.
@@
 It is always possible to construct a rectangular set of representatives, as the next proposition will show.
-We will use these rectangular regions to describe the aliasing properties of polynomial phase signals in Section~\ref{sec:ident_aliasing}.
+These rectangular regions are used to describe the aliasing properties of polynomial phase signals in Section~\ref{sec:ident_aliasing}.
 These rectangular regions will be important for the derivation of the asymptotic properties of the LSU estimator.
@@
 \]
-where $\Cbf$ is the $(m+1)\times (m+1)$ Hilbert matrix with elements $C_{ik} = 1/(i + k + 1)$ for $i,k \in \{0, 1, \dots, m\}$.
+where $\Cbf$ is the $m+1$ by $m+1$ Hilbert matrix with elements $C_{ik} = 1/(i + k + 1)$ for $i,k \in \{0, 1, \dots, m\}$.
 \end{enumerate}
@@
 A proof of the first part of this theorem (strong consistency) is given in Appendix~\ref{sec:strongconstproof}.
-Space restrictions make it impossible prove asymptotic normality in this paper.
+Space restrictions make it impossible to prove asymptotic normality in this paper.
 A complete proof of normality is available in~\cite{McKilliam_arxiv_pps_unwrapping_2012}.
@@
 The maximum number of branches allowed is $K=20N$.
+The arithmetic complexity of the LSU estimator is $O(N^3\log N)$ in this case.
 
@@
 Figure~\ref{plot:m3bigzw} shows the performance of the Zhou and Wang estimator when the coefficients are $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$, the same as in Figure~\ref{plot:m3big}.
-The threshold for the estimator is approximately~\unit[29]{dB}.
+The threshold for the estimator is approximately~\unit[28]{dB}.
 There does not appear to exist (at least in the literature) a similar method for increasing the range of coefficients for the PHAF and CPF estimators.
@@
 We observe that the $K$-best algorithm produces estimates with the same threshold as the exactly computed LSU estimator when $N = 19$ and $49$.
-As $N\rightarrow\infty$ the asymptotic threshold of the HAF estimator has been studied in~\cite{Porat_asympt_HAF_DPT_1996} and these values are indicated in the bottom rows of tables~\ref{tablm3} and~\ref{tablm5}.
+As $N\rightarrow\infty$ the asymptotic threshold of the HAF estimator has been studied in~\cite{Porat_asympt_HAF_DPT_1996} and these values are indicated in the bottom rows of Tables~\ref{tablm3} and~\ref{tablm5}.
 By Theorem~\ref{thm:asymp_proof} the LSU estimator is strongly consistent and the threshold can be made as small as desired by choosing $N$ large.
@@
 $N=49$ & 7 & 7 & 5 & 9 & 29 \\ 
-$N=99$ & & 6 & 3 & 5 & 27 \\ 
-$N=199$ & & 3 & 1 & 3 & 23 \\ 
-$N=299$ & & 2 & -1 &  & \\ 
+$N=99$ & & 6 & 3 & 7 & 29 \\ 
+$N=199$ & & 3 & 1 & 5 & 28 \\ 
+$N=299$ & & 1 & -1 & 5 & 27 \\ 
 $N\to\infty$ & $-\infty$ & & & 3.77 & \\ \bottomrule
@@
 $N=99$ &  & 15 &  12 &  19 & $>100$ \\ 
-$N=199$ & & 12 & 10 & 18 & 103 \\ 
-$N=299$ &  &  &  & &  \\ 
+$N=199$ & & 12 & 10 & 18 & $>100$ \\ 
+$N=299$ &  & 11 & 9 & 17 & $>100$ \\ 
 $N\to\infty$ & $-\infty$ &  &  & 16.6 &  \\ \bottomrule
@@
 The LSU estimator was shown to be strongly consistent and asymptotically normally distributed under mild conditions on the distribution of the noise.
-The estimator can be computed by finding a nearest lattice point in a lattice.
+The estimator can be computed by finding a nearest point in a lattice.
 Polynomial time nearest point algorithms for these lattices exist~\cite[Sec 4.3]{McKilliam2010thesis}, but these algorithms are not fast in practice.

### 2013-07-08 paper.tex
minor?

--- 2013-07-01 paper.tex
@@
 $N=99$ & & 6 & 3 & 5 & 27 \\ 
-$N=199$ & & 3 & 1 & 3 & 23 \\ 
-$N=299$ & & 2 & -1 &  & \\ 
+$N=199$ & & 3 & 1 & 5 & 23 \\ 
+$N=299$ & & 2 & -2 & 5 & \\ 
 $N\to\infty$ & $-\infty$ & & & 3.77 & \\ \bottomrule

### 2013-12-16 paper.tex
minor
minor?
added product CPF to simulation software and ran some simulation with it. It does have a lower threshold than the other estimators

--- 2013-07-06 paper.tex
@@
 % The paper headers 
-\markboth{Polynomial phase estimation by phase unwrapping}{DRAFT \today}
+\markboth{Polynomial phase estimation by least squares phase unwrapping}{DRAFT \today}
 
@@
 
+One drawback of the phase unwrapping approach is that it does not appear to generalise well to the problem of estimating the parameters in a signal that contains the sum of multiple polynomial phase signals.
+This appears to be a negative feature of all polynomial phase estimators based on phase unwrapping.
+However, the LSU estimator does generalise in other ways.
+For example, the LSU estimator can be extended to deal with polynomial phase signals in more dimensions than one.
+Our algorithms and analysis would carry over to this case with only minor modification.
+Even more generally, the ideas behind the LSU estimator might apply to other types of phase unwrapping problems such as observed in synthetic aperture radar and magnetic resonance imaging~\cite{Katkovnik_phasela_2009, Nico_phaseunwrappingSAR_2000,Friedlander_PD_phaseunwrapping_1996}.
+This is beyond the scope of our present paper, but may be the subject of future investigation.
+
 The paper is organised in the following way.
@@
 \[
-   \Lambda = \{\xbf = \Bbf\ubf \mid \ubf \in \ints^d \}
+   \Lambda = \{\Bbf\ubf \mid \ubf \in \ints^d \}
 \]
@@
 \[
-L_{m+1} = \{ \xbf = \Pbf\ubf \mid \ubf \in \ints^{m+1} \}
+L_{m+1} = \{ \Pbf\ubf \mid \ubf \in \ints^{m+1} \}
 \]
@@
 This joint minimisation problem can be solved by computing a nearest point in a lattice.
-To see this write $SS$ as a function of both $\mu_0,\dots,\mu_m$ and $W_1,\dots,W_N$,
+To see this write $SS$ as a function of both $\mu_0,\dots,\mu_m$ and $W_1,\dots,W_N$, that is,
 \begin{equation}\label{eq:SS_in_vector_form_with_u_nuisance}
@@
 where $\Xbf^+ = (\Xbf^\prime\Xbf)^{-1}\Xbf^\prime$ is the Moore-Penrose pseudoinverse of $\Xbf$.
-We may thus compute $\widehat{\mubf}$ as $\widehat{\mubf} = \mu(\widehat{\wbf})$, where $\widehat{\wbf}$ is the minimiser of
+We may thus compute $\widehat{\mubf}$ as $\widehat{\mubf} = \mubf(\widehat{\wbf})$, where $\widehat{\wbf}$ is the minimiser of
 \[
@@
 
-The theorem makes statements about the \emph{dealiased} difference $\dealias(\tilde{\mubf} - \widehat{\mubf})$ between the true coefficients $\tilde{\mubf}$ and the estimated coefficients $\widehat{\mubf}$ rather than directly on the difference $\tilde{\mubf} - \widehat{\mubf}$.
+The theorem makes statements about the \emph{dealiased} difference $\dealias(\tilde{\mubf} - \widehat{\mubf})$ between the true coefficients $\tilde{\mubf}$ and the estimated coefficients $\widehat{\mubf}$ rather than directly about the difference $\tilde{\mubf} - \widehat{\mubf}$.
 To see why this makes sense, consider the case when $m=0$, $\tilde{\mu}_0 = -0.5$ and $\widehat{\mu}_0 = 0.49$, so that $\tilde{\mu}_0 - \widehat{\mu}_0 = -0.99$.
@@
 When $m>1$ the LSU estimator can be exactly computed using the sphere decoder~\cite{Agrell2002,Viterbo_sphere_decoder_1999} in a reasonable amount of time when $N < 60$.
-When $N > 60$ we use the approximate $K$-best algorithm.
+When $N \geq 60$ we use the approximate $K$-best algorithm.
 We observe that the $K$-best algorithm produces estimates with the same threshold as the exactly computed LSU estimator when $N = 19$ and $49$.
@@
 \begin{tabular}{lrr}
- & LSU & ML  \\ \toprule
-$N=16$ & 5 & 2  \\ 
-$N=64$ & -2 & -4 \\ 
-$N=256$ & -6 & -8 \\ 
-$N=1024$ & -14 & -14 \\ 
-$N\to\infty$ & $-\infty$ & $-\infty$  \\ \bottomrule
+$N$ & LSU & ML  \\ \toprule
+$16$ & 5 & 2  \\ 
+$64$ & -2 & -4 \\ 
+$256$ & -6 & -8 \\ 
+$1024$ & -14 & -14 \\ 
+$\infty$ & $-\infty$ & $-\infty$  \\ \bottomrule
 \end{tabular}
@@
 \begin{tabular}{lrrrrr}
- & LSU & LSU ($K$-best) & CPF & HAF & ZW \\ \toprule
-$N=19$ & 10 & 10 & 8 & 11 & 31 \\ 
-$N=49$ & 7 & 7 & 5 & 9 & 29 \\ 
-$N=99$ & & 6 & 3 & 7 & 29 \\ 
-$N=199$ & & 3 & 1 & 5 & 28 \\ 
-$N=299$ & & 1 & -1 & 5 & 27 \\ 
-$N\to\infty$ & $-\infty$ & & & 3.77 & \\ \bottomrule
+$N$ & LSU & LSU ($K$-best) & CPF & HAF & ZW \\ \toprule
+$19$ & 10 & 10 & 8 & 11 & 31 \\ 
+$49$ & 7 & 7 & 5 & 9 & 29 \\ 
+$99$ & & 6 & 3 & 7 & 29 \\ 
+$199$ & & 3 & 1 & 5 & 28 \\ 
+$299$ & & 1 & -1 & 5 & 27 \\ 
+$\infty$ & $-\infty$ & & & 3.77 & \\ \bottomrule
 \end{tabular}
@@
 \begin{tabular}{lrrrrr}
- & LSU & LSU ($K$-best) & CPF-HAF & HAF & ZW \\ \toprule
-$N=19$ & 22  & 22 & 17 & 23 & $>100$ \\ 
-$N=49$ & 17 & 17 & 14 & 21 & $>100$  \\ 
-$N=99$ &  & 15 &  12 &  19 & $>100$ \\ 
-$N=199$ & & 12 & 10 & 18 & $>100$ \\ 
-$N=299$ &  & 11 & 9 & 17 & $>100$ \\ 
-$N\to\infty$ & $-\infty$ &  &  & 16.6 &  \\ \bottomrule
+$N$ & LSU & LSU ($K$-best) & CPF-HAF & HAF & ZW \\ \toprule
+$19$ & 22  & 22 & 17 & 23 & $>100$ \\ 
+$49$ & 17 & 17 & 14 & 21 & $>100$  \\ 
+$99$ &  & 15 &  12 &  19 & $>100$ \\ 
+$199$ & & 12 & 10 & 18 & $>100$ \\ 
+$299$ &  & 11 & 9 & 17 & $>100$ \\ 
+$\infty$ & $-\infty$ &  &  & 16.6 &  \\ \bottomrule
 \end{tabular}
@@
 The remaining coefficients will be shown to converge by induction.
-Assume that $N^k \lambda_{k,N} \rightarrow 0$ for all $k=r+1, r+2, \dots, m$, that is, assume that the $m-r$ highest order coefficients all converge as required.
+Assume that $N^k \lambda_{k,N} \rightarrow 0$ for all $k=r+1, r+2, \dots, m$.
+That is, assume that the $m-r$ highest order coefficients all converge as required by the lemma.
 Let

### 2013-12-16 code/java/test/pubsim/poly/CPFHAFTest.java

@@ -2,7 +2,6 @@
  */
 package pubsim.poly;
 
-import flanagan.complex.Complex;
 import org.junit.After;
 import org.junit.AfterClass;
 import static org.junit.Assert.*;
@@ -36,34 +35,34 @@ public class CPFHAFTest {
     public void tearDown() {
     }
 
-    /**
-     * Test of estimate method, of class HAF.
-     */
-    @Test
-    public void testSameAsCPFforOrder3() {
-        System.out.println("testEstimate");
-        double tol = 1e-8;
-        
-        int n = 25;
-        double[] params = {0.11, 0.05002, 0.0205, 0.0001};
-        int m = params.length-1;
-
-        PolynomialPhaseSignal siggen = new PolynomialPhaseSignal(n);
-        siggen.setParameters(params);
-        siggen.setNoiseGenerator(new GaussianNoise(0, 0.00001));
-
-        siggen.generateReceivedSignal();
-
-        CPFHAF inst = new CPFHAF(m,n);
-        CPF test = new CPF(n);
-
-        double[] pinst = inst.estimate(siggen.getReal(), siggen.getImag());
-        double[] ptest = inst.estimate(siggen.getReal(), siggen.getImag());
-
-        assertEquals(pinst[3], ptest[3],tol);
-        assertEquals(pinst[2], ptest[2],tol);
-
-    }
+//    /**
+//     * Test of estimate method, of class HAF.
+//     */
+//    @Test
+//    public void testSameAsCPFforOrder3() {
+//        System.out.println("testEstimate");
+//        double tol = 1e-8;
+//        
+//        int n = 25;
+//        double[] params = {0.11, 0.05002, 0.0205, 0.0001};
+//        int m = params.length-1;
+//
+//        PolynomialPhaseSignal siggen = new PolynomialPhaseSignal(n);
+//        siggen.setParameters(params);
+//        siggen.setNoiseGenerator(new GaussianNoise(0, 0.00001));
+//
+//        siggen.generateReceivedSignal();
+//
+//        CPFHAF inst = new CPFHAF(m,n);
+//        CPF test = new CPF(n);
+//
+//        double[] pinst = inst.estimate(siggen.getReal(), siggen.getImag());
+//        double[] ptest = inst.estimate(siggen.getReal(), siggen.getImag());
+//
+//        assertEquals(pinst[3], ptest[3],tol);
+//        assertEquals(pinst[2], ptest[2],tol);
+//
+//    }
     
     /**
[…]

### 2013-12-16 code/montecarlo.scala

@@ -10,6 +10,7 @@ import pubsim.poly.SphereDecoder
 import pubsim.poly.PolynomialPhaseEstimatorInterface
 import pubsim.poly.CPF
 import pubsim.poly.CPFHAF
+import pubsim.poly.PCPFHAF
 import pubsim.poly.PHAF
 import pubsim.poly.ZW
 import pubsim.poly.bounds.AngularLeastSquaresVariance
@@ -42,6 +43,13 @@ Array(40,40,32,50),
 Array(32,50,25,64),
 Array(25,64,20,80)
 )
+//lags for the PCPFHAF estimator
+val tau5cpf = Array(
+Array(20,20),
+Array(15,25),
+Array(17,23),
+Array(13,29)
+)
 
 val starttime = (new java.util.Date).getTime
 
@@ -72,13 +80,15 @@ runsim(5 to 25, bp5, N, iters, () => new CPFHAF(5,N), "CPFHAFm5big")
 //runsim(5 to 25, bp5, N, iters, () => new Babai(5,N, new LLL()), "Babaim5big")
 runsim(5 to 25, bp5, N, iters, () => new Mbest(5,N, 20*N, new LLL()), "Mbestm5big")
 runcrb(5 to 25, 5, N, "crbm5")
-runlsuclt(5 to 25, 5, N, "lsucltm5")*/
+runlsuclt(5 to 25, 5, N, "lsucltm5")
 
 runsim(-5 to 15, sp3, N, iters, () => new PHAF.PHAFfft(3,N,tau3), "PHAFm3small")
 runsim(-5 to 15, bp3, N, iters, () => new PHAF.PHAFfft(3,N,tau3), "PHAFm3big")
 runsim(5 to 25, sp5, N, iters, () => new PHAF.PHAFfft(5,N,tau5), "PHAFm5small")
-runsim(5 to 25, bp5, N, iters, () => new PHAF.PHAFfft(5,N,tau5), "PHAFm5big")
+runsim(5 to 25, bp5, N, iters, () => new PHAF.PHAFfft(5,N,tau5), "PHAFm5big")*/
 
+runsim(5 to 25, sp5, N, iters, () => new PCPFHAF(5,N,tau5cpf), "PCPFHAFm5small")
+runsim(5 to 25, bp5, N, iters, () => new PCPFHAF(5,N,tau5cpf), "PCPFHAFm5big")
 
 //runsim(90 to 110 by 1, bp5, N, iters, () => new ZW(5,N,N/5,N/5+1), "ZWm5big")
 

### 2013-12-16 paper.tex
added plots and modified simulation section to include product CPF-HAF

@@
 
-This section describes the results of Monte-Carlo simulations with the least squares unwrapping (LSU) estimator, the high order ambiguity function (HAF) estimator~\cite{Peleg_DPT_1995}, the product high order ambiguity function (PHAF) estimator~\cite{Barbarossa_PHAF_1998}, the cubic phase function (CPF) estimator~\cite{Oshea_cpf_2004} and the estimator that results from combining the HAF and CPF~\cite{Djurovic_haf_cpf_2012}.
+This section describes the results of Monte-Carlo simulations with the least squares unwrapping (LSU) estimator, the high order ambiguity function (HAF) estimator~\cite{Peleg_DPT_1995}, the product high order ambiguity function (PHAF) estimator~\cite{Barbarossa_PHAF_1998}, the cubic phase function (CPF) estimator~\cite{Oshea_cpf_2004} and the estimators that result from combining the HAF and CPF~\cite{Djurovic_haf_cpf_2012}.
 The CPF estimator only applies to polynomial phase signals of order $m=3$ and when the number of observations $N$ is odd.
@@
 Figure~\ref{plot:m5} displays the sample MSE of the highest order coefficient $\mu_5$  for a polynomial phase signal of order $m=5$.
-In this case the LSU and PHAF estimators reach a threshold at \unit[12]{dB}, the HAF estimator at \unit[18]{dB}, and the estimator that combines the CPF and HAF at \unit[10]{dB}.
-In this figure the true coefficients are $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}, \tfrac{1}{96N^3}, \tfrac{1}{480N^4}  \right]$ and these are chosen specifically to satisfy the requirement of the HAF, PHAF and CPF-HAF estimators.
-As in the case where $m=3$, the region of coefficients suitable for the HAF, PHAF and CPF-HAF estimators is much smaller that the identifiable region.
+In this case the LSU and PHAF estimators reach a threshold at \unit[12]{dB}, the HAF estimator at \unit[18]{dB}, the estimator that combines the CPF and HAF at \unit[10]{dB}, and the product form of this estimator (PCH) at \unit{6}[dB] (see~\cite[Section~5.A]{Djurovic_haf_cpf_2012}).
+In this figure the true coefficients are $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}, \tfrac{1}{96N^3}, \tfrac{1}{480N^4}  \right]$ and these are chosen specifically to satisfy the requirement of the HAF, PHAF, CPF-HAF and PCH estimators.
+As in the case where $m=3$, the region of coefficients suitable for the HAF, PHAF, CPF-HAF and PCH estimators is much smaller than the identifiable region.
 The effect becomes more severe as $m$ increases.
@@
 This is a dramatic reduction in the range of detectable polynomial phase signals.
-Figure~\ref{plot:m5big} displays results under the same conditions as Figure~\ref{plot:m5} but with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}, \tfrac{1}{96}, \tfrac{1}{480}  \right]$.
+Figure~\ref{plot:m5big} displays results under the same conditions as Figure~\ref{plot:m5} but with $\tilde{\mubf} = \left[\tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}, \tfrac{1}{96}, \tfrac{1}{480}  \right]$.
 The conclusion is similar to that from Figure~\ref{plot:m3big}.
-The performance of the LSU estimator remains unchanged, but the HAF, PHAF and CPF-HAF estimators fail completely.
+The performance of the LSU estimator remains unchanged, but the HAF, PHAF, CPF-HAF and PCH estimators fail completely.
 The Zhou and Wang estimator can again be used to increase the range of parameters for the HAF, but simulations indicate that the threshold for the estimator is approximately~\unit[103]{dB}.
@@
 The values in the tables are approximated using Monte-Carlo simulation with $X_1, \dots, X_N$ pseudorandomly generated independent and identically distributed circularly symmetric complex Gaussian.
-When $m=3$ the true coefficients are as in Figure~\ref{plot:m3small} and when $m=5$ the true coefficients are as in Figure~\ref{plot:m5} so that the coefficients are inside the range acceptable for the HAF, CPF and CPF-HAF estimators.
+When $m=3$ the true coefficients are as in Figure~\ref{plot:m3small} and when $m=5$ the true coefficients are as in Figure~\ref{plot:m5} so that the coefficients are inside the range acceptable for the HAF, CPF, CPF-HAF and PCH estimators.
 When $m=1$ (single frequency estimation) the LSU estimator can be computed exactly using the algorithm from~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.
@@
 \begin{tabular}{lrrrrr}
-$N$ & LSU & LSU ($K$-best) & CPF & HAF & ZW \\ \toprule
+$N$ & LSU & $K$-best & CPF & HAF & ZW \\ \toprule
 $19$ & 10 & 10 & 8 & 11 & 31 \\ 
@@
 \centering
-\begin{tabular}{lrrrrr}
-$N$ & LSU & LSU ($K$-best) & CPF-HAF & HAF & ZW \\ \toprule
-$19$ & 22  & 22 & 17 & 23 & $>100$ \\ 
-$49$ & 17 & 17 & 14 & 21 & $>100$  \\ 
-$99$ &  & 15 &  12 &  19 & $>100$ \\ 
-$199$ & & 12 & 10 & 18 & $>100$ \\ 
-$299$ &  & 11 & 9 & 17 & $>100$ \\ 
-$\infty$ & $-\infty$ &  &  & 16.6 &  \\ \bottomrule
+\begin{tabular}{lrrrrrr}
+$N$ & LSU & $K$-best & CPF-HAF & PCH  & HAF & ZW \\ \toprule
+$19$ & 22  & 22 & 17 & 13 & 23 & $>100$ \\ 
+$49$ & 17 & 17 & 14 & 10 & 21 & $>100$  \\ 
+$99$ &  & 15 &  12 &  8 & 19 & $>100$ \\ 
+$199$ & & 12 & 10 & 6 & 18 & $>100$ \\ 
+$299$ &  & 11 & 9 & 5 & 17 & $>100$ \\ 
+$\infty$ & $-\infty$ &  &  & & 16.6 &  \\ \bottomrule
 \end{tabular}

### 2013-12-16 paper.tex
added minor

@@
 
-This section describes the results of Monte-Carlo simulations with the least squares unwrapping (LSU) estimator, the high order ambiguity function (HAF) estimator~\cite{Peleg_DPT_1995}, the product high order ambiguity function (PHAF) estimator~\cite{Barbarossa_PHAF_1998}, the cubic phase function (CPF) estimator~\cite{Oshea_cpf_2004} and the estimators that result from combining the HAF and CPF~\cite{Djurovic_haf_cpf_2012}.
+This section describes the results of Monte-Carlo simulations with the least squares unwrapping (LSU) estimator, the high order ambiguity function (HAF) estimator~\cite{Peleg_DPT_1995}, the product high order ambiguity function (PHAF) estimator~\cite{Barbarossa_PHAF_1998}, and the cubic phase function (CPF) estimator~\cite{Oshea_cpf_2004}.
+Simulation are also performed with two recently proposed estimators that combine the HAF and CPF~\cite{Djurovic_haf_cpf_2012}.
+The standard version of this estimator we refer to using the abreviation CPF-HAF~\cite[Sectoion~3]{Djurovic_haf_cpf_2012} and the product form of the estimator we abreviate by PCH~\cite[Section~5.A]{Djurovic_haf_cpf_2012}.
 The CPF estimator only applies to polynomial phase signals of order $m=3$ and when the number of observations $N$ is odd.
@@
 The sample mean square error (MSE) of the $k$th coefficient is computed according to $\tfrac{1}{T}\sum_{t=1}^T \widehat{\lambda}_{k,t}^2$ where $\widehat{\lambda}_{k,t}$ is the $k$th element of $\widehat{\lambdabf}_t$.
-In all simulations four \emph{lag sets} are used for the PHAF estimator~\cite{Barbarossa_PHAF_1998} and the LSU estimator is approximated using the $K$-best method described in Section~\ref{sec:lattice-theory}.
+In all simulations four \emph{lag sets} are used for the PHAF and PCH estimators~\cite{Barbarossa_PHAF_1998,Djurovic_haf_cpf_2012} and the LSU estimator is approximated using the $K$-best method described in Section~\ref{sec:lattice-theory}.
 The maximum number of branches allowed is $K=20N$.
@@
 Figure~\ref{plot:m5} displays the sample MSE of the highest order coefficient $\mu_5$  for a polynomial phase signal of order $m=5$.
-In this case the LSU and PHAF estimators reach a threshold at \unit[12]{dB}, the HAF estimator at \unit[18]{dB}, the estimator that combines the CPF and HAF at \unit[10]{dB}, and the product form of this estimator (PCH) at \unit{6}[dB] (see~\cite[Section~5.A]{Djurovic_haf_cpf_2012}).
+In this case the LSU and PHAF estimators reach a threshold at \unit[12]{dB}, the HAF estimator at \unit[18]{dB}, the estimator that combines the CPF and HAF at \unit[10]{dB}, and the product form of this estimator (PCH) at \unit[6]{dB} (see~\cite[Section~5.A]{Djurovic_haf_cpf_2012}).
 In this figure the true coefficients are $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}, \tfrac{1}{96N^3}, \tfrac{1}{480N^4}  \right]$ and these are chosen specifically to satisfy the requirement of the HAF, PHAF, CPF-HAF and PCH estimators.

### 2013-12-16 paper.tex
minor

@@
 Simulation are also performed with two recently proposed estimators that combine the HAF and CPF~\cite{Djurovic_haf_cpf_2012}.
-The standard version of this estimator we refer to using the abreviation CPF-HAF~\cite[Sectoion~3]{Djurovic_haf_cpf_2012} and the product form of the estimator we abreviate by PCH~\cite[Section~5.A]{Djurovic_haf_cpf_2012}.
+The standard version of this estimator we refer to using the abbreviation CPF-HAF~\cite[Section~3]{Djurovic_haf_cpf_2012} and the product form of the estimator we abbreviate by PCH~\cite[Section~5.A]{Djurovic_haf_cpf_2012}.
 The CPF estimator only applies to polynomial phase signals of order $m=3$ and when the number of observations $N$ is odd.

### 2013-12-19 paper.tex
added Badri's edits

@@
 The sphere decoder was observed to have excellent statistical performance but can only be computed efficiently for $N$ less than about $60$.
-Babai's algorithm is computationally simple, but its statistical performance has been observed to be comparatively poor at low signal to noise ratio (SNR)~\cite{McKilliam2009asilomar_polyest_lattice}.
+Babai's algorithm is computationally simple, but its statistical performance has been observed to be comparatively poor at low signal-to-noise ratio (SNR)~\cite{McKilliam2009asilomar_polyest_lattice}.
 Of interest is that the lattices considered have significant structure, and may admit fast nearest point algorithms.
@@
 Although polynomial time, the algorithms are still computationally demanding in practice.
-Fast (exact or approximate) algorithms may exist for these special lattices, that are yet to be discovered.
+Fast (exact or approximate) algorithms, yet to be discovered, may exist for these special lattices.
 In this paper we apply another general purpose approximate nearest point algorithm called the $K$-best algorithm~\cite{Baro_list_sphere_dec_2003,Zhan2006_K_best_sphere_decoder,Mondal_Kbest_again_2009} for approximating the LSU estimator.
@@
 to obtain expressions for the mean square error of the estimator in question.
-The high SNR assumption is usually supported by Monte-Carlo simulations showing the mean square error predicted by the perturbation analysis to be accurate provided that the signal to noise ratio (SNR) is larger than a `threshold' value (see the simulations in Section~\ref{sec:simulations} for examples).
+The high SNR assumption is usually supported by Monte-Carlo simulations showing the mean square error predicted by the perturbation analysis to be accurate provided that the signal-to-noise ratio (SNR) is larger than a `threshold' value (see the simulations in Section~\ref{sec:simulations} for examples).
 Often the threshold is observed to occur at smaller values of SNR as $N$ increases.
@@
 \[
-p_k(x) = \binom{x}{k} = \frac{x(x-1)(x-2)\dots(x-k+1)}{k!},
+p_k(x) = \binom{x}{k} = \frac{x(x-1)(x-2)\cdots(x-k+1)}{k!},
 \]
@@
 \[
-\coef(g) = \left[ \begin{array}{ccccc} a_0 & a_1 & a_2 & \dots & a_m \end{array} \right]^\prime,
-\]
-denote the column vector of length $m+1$ containing the coefficients of $g$, where $^\prime$ indicates transpose.
-If $y$ and $z$ differ by a polynomial in $\mathcal{Z}$ then we can write $y = z + p$ where $p \in \mathcal{Z}$ and then also $\coef(y) = \coef(z) + \coef(p)$.
+\coef(g) = \left[ \begin{array}{ccccc} a_0 & a_1 & a_2 & \cdots & a_m \end{array} \right]^\prime,
+\]
+denote the column vector of length $m+1$ containing the coefficients of $g$, where $^\prime$ indicates the transpose operation.
+If $y$ and $z$ differ by a polynomial in $\mathcal{Z}$ then we can write $y = z + p$ where $p \in \mathcal{Z}$ and $\coef(y) = \coef(z) + \coef(p)$.
 %\footnote{In group theory terminology $\coef(\cdot)$ coupled with vector addition is called a \term{group homomorphism}.\index{group homomorphism}}. 
@@
 \[
-\Pbf = \left[ \begin{array}{cccc} \coef(p_0)& \coef(p_1)& \dots& \coef(p_m)  \end{array} \right]
+\Pbf = \left[ \begin{array}{cccc} \coef(p_0)& \coef(p_1)& \cdots& \coef(p_m)  \end{array} \right]
 \]
@@
 
-Let $\mubf$ be the vector $[\mu_0, \mu_1, \dots, \mu_m]$ and put,
+Let $\mubf$ be the vector $[\mu_0, \mu_1, \dots, \mu_m]$ and let
 \begin{equation} \label{eq:sumofsquaresfunction}
@@
 The $X_1, \dots, X_N$ are pseudorandomly generated independent and identically distributed circularly symmetric complex Gaussian with variance $\expect\abs{X_1}^2 = \sigma_c^2$.
-The corresponding signal to noise ratio (SNR) is $\rho^2 \sigma_c^{-2} = \sigma_c^{-2}$.
+The corresponding signal-to-noise ratio (SNR) is $\rho^2 \sigma_c^{-2} = \sigma_c^{-2}$.
 The number of replications of each Monte-Carlo experiment is $T = 2000$ to obtain estimates $\widehat{\mubf}_1, \dots, \widehat{\mubf}_T$ and the corresponding dealiased errors $\widehat{\lambdabf}_t = \dealias(\tilde{\mubf} - \widehat{\mubf}_t)$ are computed.
@@
 
-The next result we require comes from the calculus of finite differences. For any function $d(n)\from\reals\to\reals$, let 
+The next result we require comes from the calculus of finite differences. For any function $d\from\reals\to\reals$, let 
 \[
@@
 \end{equation}
-denote the $r$th difference with interval $h$. Since $\sum_{k=0}^{r}\binom{r}{k} = 2^r$ it follows that $\Delta_h^r d(n)$ can be represented by adding and subtracting the 
+denote the $r$th difference with interval $h$. Since $\sum_{k=0}^{r}\binom{r}{k} = 2^r$ it follows that $\Delta_h^r d(n)$ can be represented by adding and subtracting 
 \[

### 2014-04-28 plots/circstatfig.rb
fixed some metapost warnings.

@@ -1,5 +1,6 @@
 require 'rubypost'
 
+require 'redefine_float_to_s_for_metapost.rb'
 require 'matrix'
 require 'mathn'
 
@@ -7,14 +8,11 @@ include RubyPost
 
 file = RubyPost::File.new('circstatfig')
 
-#def dround(d) return (10000.0*d).round/10000.0 end
-def dround(d) return d end
-
 def drawpoly(e, s, func)
 	polypath = Path.new
 	t = 0
 	while(t < e) do
-		polypath.add_pair(Pair.new(dround(t).cm, dround(func.call(t)).cm))
+		polypath.add_pair(Pair.new((t).cm, (func.call(t)).cm))
 		t+=s
 	end
 	return polypath
@@ -23,7 +21,7 @@ end
 def drawsamplepoly(e, s, func, picture)
 	t = 0
 	while(t < e) do
-		picture.add_drawable(Draw.new(Circle.new()).scale(0.15.cm).translate(dround(t).cm, dround(func.call(t)).cm))
+		picture.add_drawable(Draw.new(Circle.new()).scale(0.15.cm).translate((t).cm, (func.call(t)).cm))
 		t+=s
 	end
 	return picture
@@ -34,8 +32,8 @@ def drawcircpoly(e, s, func)
 	t = 0
 	while(t < e) do
 		a = 2*Math::PI*func.call(t)
-		x = dround(t*Math.cos(a)/2.0)
-		y= dround(t*Math.sin(a)/2.0)		
+		x = (t*Math.cos(a)/2.0)
+		y= (t*Math.sin(a)/2.0)		
 		polypath.add_pair(Pair.new(x.cm, y.cm))
 		t+=s
 	end
@@ -46,8 +44,8 @@ def drawsamplecircpoly(e, s, func, picture)
 	t = 0
 	while(t < e) do
     a = 2*Math::PI*func.call(t)
-		x = dround(t*Math.cos(a)/2.0)
-		y= dround(t*Math.sin(a)/2.0)		
+		x = (t*Math.cos(a)/2.0)
+		y= (t*Math.sin(a)/2.0)		
     picture.add_drawable(Draw.new(Circle.new()).scale(0.12.cm).translate(x.cm, y.cm))
 		t+=s
 	end

### 2014-04-28 plots/circstatfigcube.rb

@@ -1,5 +1,6 @@
 require 'rubypost'
 
+require 'redefine_float_to_s_for_metapost.rb'
 require 'matrix'
 require 'mathn'
 
@@ -7,14 +8,11 @@ include RubyPost
 
 file = RubyPost::File.new('circstatfigcube')
 
-#def dround(d) return (1000.0*d).round/1000.0 end
-def dround(d) return d end
-
 def drawpoly(e, s, func)
 	polypath = Path.new
 	t = 0
 	while(t < e) do
-		polypath.add_pair(Pair.new(dround(t).cm, dround(func.call(t)).cm))
+		polypath.add_pair(Pair.new((t).cm, (func.call(t)).cm))
 		t+=s
 	end
 	return polypath
@@ -23,7 +21,7 @@ end
 def drawsamplepoly(e, s, func, picture)
 	t = 0
 	while(t < e) do
-		picture.add_drawable(Draw.new(Circle.new()).scale(0.15.cm).translate(dround(t).cm, dround(func.call(t)).cm))
+		picture.add_drawable(Draw.new(Circle.new()).scale(0.15.cm).translate((t).cm, (func.call(t)).cm))
 		t+=s
 	end
 	return picture
@@ -34,8 +32,8 @@ def drawcircpoly(e, s, func)
 	t = 0
 	while(t < e) do
 		a = 2*Math::PI*func.call(t)
-		x = dround(t*Math.cos(a)/2.0)
-		y= dround(t*Math.sin(a)/2.0)		
+		x = (t*Math.cos(a)/2.0)
+		y= (t*Math.sin(a)/2.0)		
 		polypath.add_pair(Pair.new(x.cm, y.cm))
 		t+=s
 	end
@@ -46,8 +44,8 @@ def drawsamplecircpoly(e, s, func, picture)
 	t = 0
 	while(t < e) do
     a = 2*Math::PI*func.call(t)
-		x = dround(t*Math.cos(a)/2.0)
-		y= dround(t*Math.sin(a)/2.0)		
+		x = (t*Math.cos(a)/2.0)
+		y= (t*Math.sin(a)/2.0)		
     picture.add_drawable(Draw.new(Circle.new()).scale(0.12.cm).translate(x.cm, y.cm))
 		t+=s
 	end

### 2014-04-28 plots/circstatfigquad.rb

@@ -1,5 +1,6 @@
 require 'rubypost'
 
+require 'redefine_float_to_s_for_metapost.rb'
 require 'matrix'
 require 'mathn'
 
@@ -7,14 +8,11 @@ include RubyPost
 
 file = RubyPost::File.new('circstatfigquad')
 
-#def dround(d) return (10000.0*d).round/10000.0 end
-def dround(d) return d end
-
 def drawpoly(e, s, func)
 	polypath = Path.new
 	t = 0
 	while(t < e) do
-		polypath.add_pair(Pair.new(dround(t).cm, dround(func.call(t)).cm))
+		polypath.add_pair(Pair.new((t).cm, (func.call(t)).cm))
 		t+=s
 	end
 	return polypath
@@ -23,7 +21,7 @@ end
 def drawsamplepoly(e, s, func, picture)
 	t = 0
 	while(t < e) do
-		picture.add_drawable(Draw.new(Circle.new()).scale(0.15.cm).translate(dround(t).cm, dround(func.call(t)).cm))
+		picture.add_drawable(Draw.new(Circle.new()).scale(0.15.cm).translate((t).cm, (func.call(t)).cm))
 		t+=s
 	end
 	return picture
@@ -34,8 +32,8 @@ def drawcircpoly(e, s, func)
 	t = 0
 	while(t < e) do
 		a = 2*Math::PI*func.call(t)
-		x = dround(t*Math.cos(a)/2.0)
-		y= dround(t*Math.sin(a)/2.0)		
+		x = (t*Math.cos(a)/2.0)
+		y= (t*Math.sin(a)/2.0)		
 		polypath.add_pair(Pair.new(x.cm, y.cm))
 		t+=s
 	end
@@ -46,8 +44,8 @@ def drawsamplecircpoly(e, s, func, picture)
 	t = 0
 	while(t < e) do
     a = 2*Math::PI*func.call(t)
-		x = dround(t*Math.cos(a)/2.0)
-		y= dround(t*Math.sin(a)/2.0)		
+		x = (t*Math.cos(a)/2.0)
+		y= (t*Math.sin(a)/2.0)		
     picture.add_drawable(Draw.new(Circle.new()).scale(0.12.cm).translate(x.cm, y.cm))
 		t+=s
 	end

### 2014-04-28 plots/circstatfigzero.rb

@@ -1,5 +1,6 @@
 require 'rubypost'
 
+require 'redefine_float_to_s_for_metapost.rb'
 require 'matrix'
 require 'mathn'
 
@@ -7,14 +8,11 @@ include RubyPost
 
 file = RubyPost::File.new('circstatfigzero')
 
-#def dround(d) return (10000.0*d).round/10000.0 end
-def dround(d) return d end
-
 def drawpoly(e, s, func)
 	polypath = Path.new
 	t = 0
 	while(t < e) do
-		polypath.add_pair(Pair.new(dround(t).cm, dround(func.call(t)).cm))
+		polypath.add_pair(Pair.new((t).cm, (func.call(t)).cm))
 		t+=s
 	end
 	return polypath
@@ -23,7 +21,7 @@ end
 def drawsamplepoly(e, s, func, picture)
 	t = 0
 	while(t < e) do
-		picture.add_drawable(Draw.new(Circle.new()).scale(0.15.cm).translate(dround(t).cm, dround(func.call(t)).cm))
+		picture.add_drawable(Draw.new(Circle.new()).scale(0.15.cm).translate((t).cm, (func.call(t)).cm))
 		t+=s
 	end
 	return picture
@@ -34,8 +32,8 @@ def drawcircpoly(e, s, func)
 	t = 0
 	while(t < e) do
 		a = 2*Math::PI*func.call(t)
-		x = dround(t*Math.cos(a)/2.0)
-		y= dround(t*Math.sin(a)/2.0)		
+		x = (t*Math.cos(a)/2.0)
+		y= (t*Math.sin(a)/2.0)		
 		polypath.add_pair(Pair.new(x.cm, y.cm))
 		t+=s
 	end
@@ -46,8 +44,8 @@ def drawsamplecircpoly(e, s, func, picture)
 	t = 0
 	while(t < e) do
     a = 2*Math::PI*func.call(t)
-		x = dround(t*Math.cos(a)/2.0)
-		y= dround(t*Math.sin(a)/2.0)		
+		x = (t*Math.cos(a)/2.0)
+		y= (t*Math.sin(a)/2.0)		
     picture.add_drawable(Draw.new(Circle.new()).scale(0.12.cm).translate(x.cm, y.cm))
 		t+=s
 	end

### 2014-04-28 plots/tesselationfigures.rb

@@ -1,5 +1,6 @@
 require 'rubypost'
 
+require 'redefine_float_to_s_for_metapost.rb'
 require 'matrix'
 require 'mathn'
 
